### 2017-01-19 hicks-summary/melt/summary-doc.Rmd
hicks rmd

---
title: "Summary of consumption in Medicare Part D"
author: "Scott Olesen"
date: "1/19/2017"
bibliography: "refs.bib"
output: html_document
---

This summary is modeled after @hicks-us-2015. That study, which used the IMS Health Xponent database to infer the number of prescriptions per capita, reported higher prescription rates than what we observe here. In that paper, the number of prescriptions per 1,000 people year (PKPY) is approximately 1,000 for people over 65. In this data, we find that number to be between 600 (in 2011) and 750 (in 2012). Our value is more similar to a report that used the National Ambulatory Care Survey (NAMCS) and National Hospital Ambulatory Care Survey (NHAMCS) rather than the Xponent data. That report estimated about 500 PKPY [@flemingdutra-prevalence-2016]. A third report, which used the Xponent database, reported that prescriptions fell from about 1,000 PKPY in 1996 to about 600 PKPY in 2003 [@hicks-outpatient-2011].

```{r load_data, echo=FALSE}
library(knitr)
dat = read_tsv('dat.tsv', progress=FALSE, col_types='cccd') %>%
  rename(rx=rx_per_1k_ppl) %>%
  mutate(rx=round(rx))
```

# Antibiotic consumption by class

Our data for antibiotic class roughly correspond to those reported in @hicks-us-2015, when allowing for the fact that their results have about 40% higher consumption overall. The striking difference is a in consumption of quinolones: in our 2011 data, quinolones are the most-prescribed class and account for 127 PKPY. In their data, quinolones are the fourth-most-prescribed class and account for 89 PKPY (adjusted to approximately $90 \times 0.6 = 54$).

```{r by_class, echo=FALSE}
top_abx_names = dat %>%
  filter(year=='2011', type=='rx_by_class') %>%
  arrange(desc(rx)) %>%
  head(9) %$%
  field

dat %>%
  filter(type=='rx_by_class') %>%
  mutate(field=if_else(field %in% top_abx_names, field, 'other')) %>% 
  group_by(year, field) %>%
  summarize(rx=sum(rx)) %>%
  ungroup %>%
  spread(year, rx) %>%
  {
    other_row = filter(., field=='other')
    most_rows = filter(., field!='other') %>% arrange(desc(`2011`))
    bind_rows(most_rows, other_row)
  } %>%
  rename(antibiotic=field) %>%
  kable
```

# Most-prescribed antibiotics

Our data for the most prescribed individual antibiotics also generally accords with @hicks-us-2015 (again, allowing for the 40% discrepancy). The major difference is that ciprofloxacin consumption is greater in our data and amoxicillin (and augmentin) consumption are greater in their data. The reduced consumption of amoxicillin in our data could be explained by the fact that our population is older.

```{r, echo=FALSE}
dat %>%
  filter(type=='top_abx_rx') %>%
  select(-type) %>% 
  spread(year, rx) %>%
  arrange(desc(`2011`)) %>%
  rename(antibiotic=field) %>%
  kable
```

# Consumption by sex

Our result, that the consumption of antibiotics by women and men is in a 3:2 ratio, agrees with the result in @hicks-us-2015.

```{r, echo=FALSE}
table_f = function(x) {
  dat %>%
    filter(type==x) %>%
    select(-type) %>%
    spread(year, rx) %>%
    arrange(desc(`2011`))
}
table_f('rx_by_sex') %>% rename(sex=field) %>% filter(sex != 'unknown') %>% kable
```

# Consumption by region

```{r, echo=FALSE}
table_f('rx_by_region') %>% rename(region=field) %>% kable
```

# Consumption by age

When stratifying by age, @hicks-us-2015 places all people 65 or older into one stratum. Here I stratify within this population. Excepting the beneficiaries under 65, older beneficiaries tend to consume more antibiotics. The beneficiaries under 65 qualify for Medicare because they are disabled, have end stage renal disease, or ALS, which might contribute to their increased consumption of antibiotics.

```{r, echo=FALSE}
dat %>%
  filter(type=='rx_by_age') %>%
  select(-type) %>%
  spread(year, rx) %>%
  arrange(field) %>%
  rename(age=field) %>%
  kable
```

### 2017-01-20 hicks-summary/melt/summary-doc.Rmd
added temporal commentary

@@
 date: "1/19/2017"
-bibliography: "refs.bib"
 output: html_document
+bibliography: refs.bib
 ---
@@
 In their data, quinolones are the fourth-most-prescribed class and account for 89 PKPY (adjusted to approximately $90 \times 0.6 = 54$).
+
+Across years, the consumption of drugs seems to follow similar patterns (excepting that consumption in 2012 was about 20% higher than in 2011 and consumption in 2013-2014 was about 10% higher).
+
+All consumption rates are reported as PKPY.
 
@@
 
-```{r, echo=FALSE}
+Allowing for the overall differences in consumption, consumption of azithromycin appears to have declined during 2011 to 2014, while the consumption of levofloxacin increased.
+
+```{r top_abx, echo=FALSE}
 dat %>%
@@
 
-Our result, that the consumption of antibiotics by women and men is in a 3:2 ratio, agrees with the result in @hicks-us-2015.
+Our result, that the consumption of antibiotics by women and men is in a 3:2 ratio, agrees with the result in @hicks-us-2015. The ratio also remained steady across years in our dataset.
 
@@
 
+In @hicks-us-2015, stratifying consumption in 2011 by census region led to one major divide: the South, West, and Northeast consumed more antibiotics than the West (931, 897, and 848 vs.
+647 PKPY).
+In our results, the major divide is between the South and the Midwest, West, and Northeast (641 vs.
+583, 563, and 533 PKPY).
+
+In our data, 2012 appears exceptional: consumption in the Midwest and West was relatively high, and consumption in the Northeast was relatively low.
+Results from 2013 and 2014 accord more closely with the results from @hicks-us-2015, in which the West is the region with the lowest consumption.
+
 ```{r, echo=FALSE}
@@
 The beneficiaries under 65 qualify for Medicare because they are disabled, have end stage renal disease, or ALS, which might contribute to their increased consumption of antibiotics.
+However, this trend across age groups does not hold in 2012.
 

### 2017-02-01 states/resistance/report/state-report.Rmd
state report

---
title: 'Consumption and resistance across states'
---

```{r echo=FALSE, message=FALSE}
res = read_tsv('../model-results-p005.tsv')
plot_dat = read_tsv('../plotting-dat.tsv')

plot_f = function(this_bug, this_resistance_drug, this_explanatory, this_response) {
  this_res = res %>%
    filter(bug==this_bug, resistance_drug==this_resistance_drug, explanatory==this_explanatory, response==this_response)

  if (dim(this_res)[1] != 1) stop('wrong number of rows')

  z = this_res$correlation %>% head
  p.value = this_res$p.value %>% head

  plot_dat %>%
    filter(bug==this_bug, resistance_drug==this_resistance_drug) %>%
    ggplot(aes_string(this_explanatory, this_response)) +
    geom_point() +
    geom_smooth(method='lm', se=FALSE) +
    theme_minimal() +
    xlab(this_explanatory) +
    ylab(this_response) +
    ggtitle(sprintf('%s susceptibility to %s', this_bug, this_resistance_drug))
}
```

# Methodology

1. I determined which bug/drug combinations were coherent at the state level (i.e., separating antibiograms into their states significantly decreased variance below what would be expected by chance)
2. For each *consumption* drug, I asserted a matched *resistance* drug
3. I correlated metrics of consumption (of the consumption drug) against metrics of resistance (against the resistance drug) across states.

## Metrics

The consumption metrics were:

- *DID*: Defined daily doses per 1,000 inhabitants per day. In fact, I don't have proper DDDs, so these are the "days supply" supplied in the Medicare claims data. This is what Goossens *et al.* uses.
- *Claims per 1,000 people*: Rather than days supply, this is just the raw number of claims. It also has a different denominator. This is what the Hicks papers use.

In general, PKPY will be larger than DID. If all people get $d$ days of drug per prescription, the relationship is:

$$
\mathrm{PKPY} = \frac{365}{d} \times \mathrm{DID}
$$

The resistance metrics were:

- *Mean* susceptibility
- *Median* susceptibility
- *Maximum* susceptibility
- *Minimum* susceptibility

## Correlations

I used Spearman correlations and examined only those bug/drug combinations that produced $p < 0.05$. I manually examined the results and show some representative ones here.

## Consumption and resistance drugs

Trimethoprim-sulfamethoxazole, ciprofloxacin, levofloxacin, ceftriaxone, nitrofurantoin, and penicillin were present in both the consumption and resistance data; they are each their own pair. Consumption of amoxicillin was matched with resistance to ampicillin; doxycycline with tetracycline/doxycycline, and azithromycin with erythromycin.

## Organisms

I examined these organisms:

Gram negatives:
- Escherichia coli
- Klebsiella pneumoniae
- Proteus mirabilis
- Enterobacter cloacae
- Klebisella oxytoca

Gram positives:
- Coagulase-negative staphylococci
- Methicillin-resistant S. aureus (MRSA)
- Methicillin-susceptible S. aureus (MSSA)
- Streptococcus pneumoniae (non-meningitis)
- Staphylococcus aureus

# Results

## Strong results

I call these strong results because:

- Their p-values were small.
- The correlations went in the right direction (i.e., increased consumption was associated with decreased susceptibility).
- The same bug/drug/drug combinations had significant correlations for multiple metrics (e.g., *E. coli* susceptibility to cipro anticorrelated with cipro consumption regardless of whether consumption was measured with claims or days and whether the mean or median susceptibility was used).

In general, the coefficients for these correlations were in the range 0.4--0.6.

N.B.: The plots have a linear best fit line added for clarity, but the correlation was measured with Spearman (i.e., rank ordered).

```{r echo=FALSE}
plot_f('Escherichia coli', 'ciprofloxacin', 'claims_per_1k_ppl', 'mean_sus')
plot_f('Escherichia coli', 'levofloxacin', 'claims_per_1k_ppl', 'mean_sus')
plot_f('Methicillin-susceptible S. aureus (MSSA)', 'ciprofloxacin', 'claims_per_1k_ppl', 'mean_sus')
plot_f('Methicillin-susceptible S. aureus (MSSA)', 'erythromycin', 'claims_per_1k_ppl', 'mean_sus')
plot_f('Streptococcus pneumoniae (non-meningitis)', 'erythromycin', 'claims_per_1k_ppl', 'mean_sus')
```

## Dubious Results

I call these results "dubious" because, in general:

- The p-values were not as strong.
- The correlations were sometimes in the wrong direction.
- The susceptibility metrics were weird (e.g., consumption correlations with the minimum susceptibility reported in a state).
- The correlation was significant for one consumption metric but not the other.

```{r echo=FALSE}
plot_f('Coagulase-negative staphylococci', 'levofloxacin', 'claims_per_1k_ppl', 'min_sus') # why only min sus?
plot_f('Enterobacter cloacae', 'ceftriaxone', 'did', 'mean_sus') # why only DID?
plot_f('Escherichia coli', 'ampicillin', 'did', 'median_sus') # why only DID?
plot_f('Escherichia coli', 'trimethoprim-sulfamethoxazole', 'claims_per_1k_ppl', 'mean_sus') # why not claims per 1k ppl?
plot_f('Klebsiella pneumoniae', 'ceftriaxone', 'did', 'mean_sus') # why only DID?
plot_f('Methicillin-resistant S. aureus (MRSA)', 'levofloxacin', 'claims_per_1k_ppl', 'max_sus') # why only max sus?
plot_f('Methicillin-resistant S. aureus (MRSA)', 'trimethoprim-sulfamethoxazole', 'did', 'mean_sus') # why only DID?
plot_f('Proteus mirabilis', 'levofloxacin', 'claims_per_1k_ppl', 'min_sus') # why only min sus?
plot_f('Streptococcus pneumoniae (non-meningitis)', 'trimethoprim-sulfamethoxazole', 'claims_per_1k_ppl', 'mean_sus') # why only some?
```

# Discussion

## Comparison to Goossens

Goossens *et al.* found significant correlations for:

- Consumption of macrolides with *S. pneumoniae* susceptibility to erythromycin
- Consumption of penicillins and cephalosporins with *S. pneumoniae* susceptibility to penicillin
- Consumption of macrolides and lincosamids with *S. pyogenes* susceptibility to erythromycin
- Consumption of quinolones with *E. coli* susceptibility to ciprofloxacin
- Consumption of co-trimoxazole with *E. coli* susceptibility to co-trimoxazole

The Spearman correlation coefficients range from 0.68 to 0.83, which are higher than our results.

The Goossens *et al.* graph shows a best-fit line of penicillin consumption against *S. pneumoniae* non-susceptibility to penicillin: 5 DID corresponds to 0% non-susceptible, while 16 DID corresponds to 45% non-susceptible. The translates to a 4 percentage point change in susceptibility per 1 DID.

I can't immediately compare this to the US data (because *S. pneumoniae* resistance to penicillin was not coherent at the state level).

However, *E. coli* resistance to ciprofloxacin is interesting: 85% susceptibility at 50 PKPY and 75% at 200, that is, 1 percentage point in susceptibility per 15 PKPY. That's a very weak response!

### 2017-02-02 states/resistance/penicillin-pneumococcus/penicillin-pneumo-report.Rmd
pneumococcus penicillin

---
title: "Pen-pneumo"
author: "Scott Olesen"
output: pdf_document
fontsize: 10pt
---

# Background

To compare against Goossens *et al.*, I wanted to compare the consumption of penicillin against the susceptibility of pneumococcus to penicillin. Because *S. pneumoniae* susceptibility to penicillin was not coherent at the state level, I wrote this result up separately. I do not filter the data used for coherence: I just run the correlations.

# Methods

```{r echo=FALSE}

library(broom)

bug = 'Streptococcus pneumoniae (non-meningitis)'
consumption_drug=c('amoxicillin', 'penicillin')
resistance_drug=c('ampicillin', 'penicillin')

bug_drug = expand.grid(bug=bug,
                       consumption_drug=consumption_drug,
                       resistance_drug=resistance_drug,
                       stringsAsFactors=FALSE)
```

I'm going to look at this one organism but a few consumption drugs and resistance drugs:

```{r echo=FALSE}
bug_drug
```

```{r echo=FALSE, message=FALSE}
abg = read_tsv('../../../../../antibiogram/data/abg.tsv') %>%
  mutate(drug=tolower(drug)) %>%
  rename(resistance_drug=drug)

known_states = abg %$% state %>% unique %>% sort

state_bene = read_tsv('../../bene-consumption/state-bene-2011.tsv') %>%
  filter(state %in% known_states)
state_cons = read_tsv('../../bene-consumption/state-consumption-2011.tsv') %>%
  filter(state %in% known_states) %>%
  rename(consumption_drug=drug)

sus = bug_drug %>%
  left_join(abg, by=c('bug', 'resistance_drug')) %>%
  group_by(bug, consumption_drug, resistance_drug, state) %>%
  summarize(n_abg=n(),
            n_isolates=sum(n_isolates, na.rm=TRUE),
            n_years=length(unique(year)),
            median_sus=median(percent_susceptible, na.rm=TRUE),
            mean_sus=mean(percent_susceptible, na.rm=TRUE),
            sd_sus=sd(percent_susceptible, na.rm=TRUE),
            min_sus=min(percent_susceptible, na.rm=TRUE),
            max_sus=max(percent_susceptible, na.rm=TRUE))
```

The susceptibility data:
```{r echo=FALSE}
sus
```

```{r echo=FALSE}
models = function(df) {
  out = data_frame()
  for (resp in c('median_sus', 'mean_sus', 'min_sus', 'max_sus')) {
    for (expl in c('claims_per_1k_ppl', 'did')) {
      new_rows = cor.test(df[[resp]], df[[expl]], method='spearman') %>%
        tidy %>%
        mutate(response=resp, explanatory=expl)
      out = bind_rows(out, new_rows)
    }
  }
  out
}

dat = sus %>%
  left_join(bug_drug, by=c('bug', 'consumption_drug', 'resistance_drug')) %>%
  left_join(state_cons, by=c('state', 'consumption_drug')) %>%
  group_by(bug, consumption_drug, resistance_drug)
```

The data used for correlations:
```{r echo=FALSE}
dat
```

The model results:
```{r echo=FALSE, warning=FALSE}
res = dat %>%
  do(models(.))
res
```

# Results

Of which, the statistically significant correlations are:

```{r echo=FALSE}
res %>%
  filter(p.value < 0.05) %>%
  select(bug, consumption_drug, resistance_drug, explanatory, response, correlation=estimate, p.value) %>%
  mutate(correlation=signif(correlation, digits=3), p.value=signif(p.value, digits=3)) %>%
  ungroup() %>%
  select(-bug) %>%
  t
```

Weirdly, these are *positive* correlations: more consumption means more susceptibility.

```{r echo=FALSE}
dat %>%
  filter(consumption_drug=='penicillin',
         resistance_drug=='penicillin') %>%
  ggplot(aes(x=claims_per_1k_ppl, y=mean_sus)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE) +
  xlab('claims per 1k people') +
  ylab('mean susceptibility (%)') +
  ggtitle('Pneumococcus sus. to pen. vs. pen. cons. by state') +
  theme_minimal()
```

### 2017-02-02 states/resistance/penicillin-pneumococcus/first.R

@@ -1,0 +1,76 @@
+#!/usr/bin/env Rscript
+
+library(broom)
+
+# Do the analysis like before, but just for this one bug/drug combo, and not
+# asking if the antibiograms are coherent or anything
+
+bug = 'Streptococcus pneumoniae (non-meningitis)'
+consumption_drug=c('amoxicillin', 'penicillin')
+resistance_drug=c('ampicillin', 'penicillin')
+
+bug_drug = expand.grid(bug=bug,
+                       consumption_drug=consumption_drug,
+                       resistance_drug=resistance_drug,
+                       stringsAsFactors=FALSE)
+
+# do the correlations
+
+abg = read_tsv('../../../../../antibiogram/data/abg.tsv') %>%
+  mutate(drug=tolower(drug)) %>%
+  rename(resistance_drug=drug)
+
+known_states = abg %$% state %>% unique %>% sort
+
+state_bene = read_tsv('../../bene-consumption/state-bene-2011.tsv') %>%
+  filter(state %in% known_states)
+state_cons = read_tsv('../../bene-consumption/state-consumption-2011.tsv') %>%
+  filter(state %in% known_states) %>%
+  rename(consumption_drug=drug)
+
+sus = bug_drug %>%
+  left_join(abg, by=c('bug', 'resistance_drug')) %>%
+  group_by(bug, consumption_drug, resistance_drug, state) %>%
+  summarize(n_abg=n(),
+            n_isolates=sum(n_isolates, na.rm=TRUE),
+            n_years=length(unique(year)),
+            median_sus=median(percent_susceptible, na.rm=TRUE),
+            mean_sus=mean(percent_susceptible, na.rm=TRUE),
+            sd_sus=sd(percent_susceptible, na.rm=TRUE),
+            min_sus=min(percent_susceptible, na.rm=TRUE),
+            max_sus=max(percent_susceptible, na.rm=TRUE))
+
+models = function(df) {
+  out = data_frame()
+  for (resp in c('median_sus', 'mean_sus', 'min_sus', 'max_sus')) {
+    for (expl in c('claims_per_1k_ppl', 'did')) {
+      sprintf("%s %s %s %s\n", resp, expl, head(df$resistance_drug, 1), head(df$consumption_drug, 1), head(df$bug, 1)) %>% cat
+      new_rows = cor.test(df[[resp]], df[[expl]], method='spearman') %>%
[…]

### 2017-02-02 states/resistance/penicillin-pneumococcus/make-plots.R

@@ -1,0 +1,32 @@
+#!/usr/bin/env Rscript
+
+res = read_tsv('model-results-p005.tsv')
+plot_dat = read_tsv('plotting-dat.tsv')
+
+plot_f = function(this_consumption_drug, this_resistance_drug, this_explanatory, this_response) {
+  this_res = res %>%
+    filter(consumption_drug==this_consumption_drug, resistance_drug==this_resistance_drug, explanatory==this_explanatory, response==this_response)
+
+  if (dim(this_res)[1] != 1) stop('wrong number of rows')
+
+  z = this_res$correlation %>% head
+  p.value = this_res$p.value %>% head
+
+  plot_dat %>%
+    filter(consumption_drug==this_consumption_drug, resistance_drug==this_resistance_drug) %>%
+    ggplot(aes_string(this_explanatory, this_response)) +
+    geom_point() +
+    geom_smooth(method='lm', se=FALSE) +
+    theme_minimal() +
+    xlab(this_explanatory) +
+    ylab(this_response) +
+    ggtitle(sprintf('Pneumo. sus. to %s, cons. of %s', this_resistance_drug, this_consumption_drug))
+}
+
+save_plot = function(cons, res, expl, resp, file_nickname) {
+  fn = sprintf('plots/%s.pdf', file_nickname)
+  p = plot_f(cons, res, expl, resp)
+  ggsave(fn, plot=p, useDingbats=FALSE)
+}
+
+save_plot('penicillin', 'penicillin', 'claims_per_1k_ppl', 'mean_sus', 'pen')

### 2017-02-03 states/resistance/penicillin-pneumococcus/penicillin-pneumo-report.Rmd
kable

@@
 ```{r echo=FALSE}
-
+library(knitr)
 library(broom)
@@
 
-```{r echo=FALSE}
-bug_drug
+```{r echo=FALSE, result='asis'}
+bug_drug %>% kable
 ```
 
-```{r echo=FALSE, message=FALSE}
+```{r echo=FALSE, message=FALSE, cache=TRUE}
 abg = read_tsv('../../../../../antibiogram/data/abg.tsv') %>%

### 2017-02-03 states/resistance/report2/state-report.Rmd
second report

---
title: "Consumption and resistance across states"
output:
  html_document:
    fig_height: 4
  pdf_document:
    fig_height: 3
---

```{r setup}
library(broom)
library(knitr)
library(pander)

coherence = read_tsv('../../../../../antibiogram/analysis/geographic-coherence/coherences.tsv') %>%
  rename(resistance_drug=drug) %>%
  mutate(resistance_drug=tolower(resistance_drug)) %>%
  select(bug, resistance_drug, level, n_groups)

gram_negs = c('Escherichia coli', 'Klebsiella pneumoniae', 'Proteus mirabilis',
              'Enterobacter cloacae', 'Klebisella oxytoca')
gram_poss = c('Coagulase-negative staphylococci', 'Methicillin-resistant S. aureus (MRSA)',
              'Methicillin-susceptible S. aureus (MSSA)', 'Streptococcus pneumoniae (non-meningitis)',
              'Staphylococcus aureus')
organisms = c(gram_negs, gram_poss)

consumption_drugs=c('trimethoprim-sulfamethoxazole', 'ciprofloxacin', 'levofloxacin',
                    'ceftriaxone', 'nitrofurantoin', 'amoxicillin', 'penicillin',
                    'doxycycline', 'azithromycin')
resistance_drugs=c('trimethoprim-sulfamethoxazole', 'ciprofloxacin', 'levofloxacin',
                   'ceftriaxone', 'nitrofurantoin', 'ampicillin', 'penicillin',
                   'tetracycline/doxycycline', 'erythromycin')

consumption_resistance = data_frame(consumption_drug=consumption_drugs,
                                    resistance_drug=resistance_drugs)

# get all bug/drug combos (i.e., all bug/cons drug combos with appropriate res drugs)
bug_drug = expand.grid(bug=organisms,
                       consumption_drug=consumption_drugs,
                       stringsAsFactors=FALSE) %>%
  left_join(consumption_resistance, by='consumption_drug') %>%
  left_join(coherence, by=c('resistance_drug', 'bug')) %>%
  mutate(state_coherent=level %in% c('state', 'hrr', 'zipcode'))
```

```{r load_data}
abg = read_tsv('../../../../../antibiogram/data/abg.tsv') %>%
  mutate(drug=tolower(drug)) %>%
  rename(resistance_drug=drug)

known_states = abg %$% state %>% unique %>% sort

state_bene = read_tsv('../../bene-consumption/state-bene-2011.tsv') %>%
  filter(state %in% known_states)
state_cons = read_tsv('../../bene-consumption/state-consumption-2011.tsv') %>%
  filter(state %in% known_states) %>%
  rename(consumption_drug=drug)

sus = bug_drug %>%
  left_join(abg, by=c('bug', 'resistance_drug')) %>%
  group_by(bug, resistance_drug, state) %>%
  summarize(n_abg=n(),
            n_isolates=sum(n_isolates, na.rm=TRUE),
            mean_sus=mean(percent_susceptible, na.rm=TRUE))
```

```{r run_models}
safe_spearman = function(df, expl) {
  if (dim(df)[1] < 3L) {
    data_frame(estimate=NA, statistic=NA, p.value=NA, method=NA, alternative=NA)
  } else {
    cor.test(df[['mean_sus']], df[[expl]], method='spearman') %>% tidy
  }
}

models = function(df) {
  out = data_frame()
  for (expl in c('claims_per_1k_ppl', 'did')) {
    new_rows = safe_spearman(df, expl) %>% 
      mutate(explanatory=expl)

    out = bind_rows(out, new_rows)
  }
  out
}

dat = sus %>%
  left_join(bug_drug, by=c('bug', 'resistance_drug')) %>%
  left_join(state_cons, by=c('state', 'consumption_drug')) %>%
  group_by(bug, resistance_drug)

res = dat %>%
  do(models(.))
```

```{r echo=FALSE, message=FALSE}

plot_f = function(this_bug, this_resistance_drug, this_explanatory, this_response) {
  this_res = res %>%
    filter(bug==this_bug, resistance_drug==this_resistance_drug, explanatory==this_explanatory, response==this_response)

  if (dim(this_res)[1] != 1) stop('wrong number of rows')

  z = this_res$correlation %>% head
  p.value = this_res$p.value %>% head

  plot_dat %>%
    filter(bug==this_bug, resistance_drug==this_resistance_drug) %>%
    ggplot(aes_string(this_explanatory, this_response)) +
    geom_point() +
    geom_smooth(method='lm', se=FALSE) +
    theme_minimal() +
    xlab(this_explanatory) +
    ylab(this_response) +
    ggtitle(sprintf('%s susceptibility to %s', this_bug, this_resistance_drug))
}
```

```{r}
res %>%
  head %>%
  select(bug, resistance_drug, explanatory, correlation=estimate, p.value) %>%
  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
  kable
```

### 2017-02-07 states/resistance/report2/state-report.Rmd
mostly fixed up report2

@@
 
-```{r setup}
+```{r setup, echo=FALSE}
 library(broom)
@@
 library(pander)
-
-coherence = read_tsv('../../../../../antibiogram/analysis/geographic-coherence/coherences.tsv') %>%
-  rename(resistance_drug=drug) %>%
-  mutate(resistance_drug=tolower(resistance_drug)) %>%
-  select(bug, resistance_drug, level, n_groups)
 
@@
   left_join(consumption_resistance, by='consumption_drug') %>%
-  left_join(coherence, by=c('resistance_drug', 'bug')) %>%
-  mutate(state_coherent=level %in% c('state', 'hrr', 'zipcode'))
+  select(bug, consumption_drug, resistance_drug)
 ```
 
-```{r load_data}
+```{r load_data, echo=FALSE, message=FALSE}
+coherence = read_tsv('../../../../../antibiogram/analysis/geographic-coherence/coherences.tsv') %>%
+  rename(resistance_drug=drug) %>%
+  mutate(resistance_drug=tolower(resistance_drug),
+         state_coherent=level %in% c('state', 'hrr', 'zipcode')) %>%
+  select(bug, resistance_drug, state_coherent)
+
+# Antibiogram data, with individual records
 abg = read_tsv('../../../../../antibiogram/data/abg.tsv') %>%
   mutate(drug=tolower(drug)) %>%
-  rename(resistance_drug=drug)
+  rename(resistance_drug=drug) %>%
+  select(bug, resistance_drug, state, percent_susceptible)
 
-known_states = abg %$% state %>% unique %>% sort
+# swo> State beneficiary information could be used as control variables
+# in regressions
+#state_bene = read_tsv('../../bene-consumption/state-bene-2011.tsv') %>%
+  #filter(state %in% known_states)
 
-state_bene = read_tsv('../../bene-consumption/state-bene-2011.tsv') %>%
-  filter(state %in% known_states)
-state_cons = read_tsv('../../bene-consumption/state-consumption-2011.tsv') %>%
-  filter(state %in% known_states) %>%
-  rename(consumption_drug=drug)
+# State-level consumption data
+state_consumption = read_tsv('../../bene-consumption/state-consumption-2011.tsv') %>%
+  rename(consumption_drug=drug) %>%
+  select(state, consumption_drug, claims_per_1k_ppl, did)
 
-sus = bug_drug %>%
+dat = bug_drug %>%
   left_join(abg, by=c('bug', 'resistance_drug')) %>%
-  group_by(bug, resistance_drug, state) %>%
-  summarize(n_abg=n(),
-            n_isolates=sum(n_isolates, na.rm=TRUE),
-            mean_sus=mean(percent_susceptible, na.rm=TRUE))
+  left_join(state_consumption, by=c('state', 'consumption_drug'))
 ```
 
-```{r run_models}
+```{r run_models, echo=FALSE, warning=FALSE}
 safe_spearman = function(df, expl) {
@@
 
-dat = sus %>%
+res = dat %>%
+  group_by(bug, resistance_drug, state) %>%
+  summarize(mean_sus=mean(percent_susceptible, na.rm=TRUE)) %>%
   left_join(bug_drug, by=c('bug', 'resistance_drug')) %>%
-  left_join(state_cons, by=c('state', 'consumption_drug')) %>%
-  group_by(bug, resistance_drug)
-
-res = dat %>%
-  do(models(.))
+  left_join(state_consumption, by=c('state', 'consumption_drug')) %>%
+  group_by(bug, resistance_drug, consumption_drug) %>%
+  do(models(.)) %>%
+  ungroup %>%
+  filter(!is.na(estimate)) %>%
+  left_join(coherence, by=c('bug', 'resistance_drug')) %>%
+  mutate(record_number=1:n(),
+         plot_fn=sprintf('plots/%03i.pdf', record_number),
+         plot_link=pandoc.link.return(url=plot_fn, text='plot')) %>%
+  select(bug, consumption_drug, resistance_drug, state_coherent,
+         explanatory, correlation=estimate, p.value,
+         plot_fn, plot_link)
 ```
 
-```{r echo=FALSE, message=FALSE}
+```{r make_plots, echo=FALSE, message=FALSE}
+unlink('plots/*.pdf')
 
-plot_f = function(this_bug, this_resistance_drug, this_explanatory, this_response) {
-  this_res = res %>%
-    filter(bug==this_bug, resistance_drug==this_resistance_drug, explanatory==this_explanatory, response==this_response)
+for (i in 1:nrow(res)) {
+  row = res[i,]
 
-  if (dim(this_res)[1] != 1) stop('wrong number of rows')
+  plot_dat = dat %>%
+    filter(bug==row$bug, consumption_drug==row$consumption_drug)
 
-  z = this_res$correlation %>% head
-  p.value = this_res$p.value %>% head
+  p = plot_dat %>%
+    filter(bug==row$bug, consumption_drug==row$consumption_drug) %>%
+    ggplot(aes_string(row$explanatory, 'percent_susceptible')) +
+    geom_point() +
+    xlab(row$explanatory) +
+    ylab('percent susceptible') +
+    theme_minimal()
 
-  plot_dat %>%
-    filter(bug==this_bug, resistance_drug==this_resistance_drug) %>%
-    ggplot(aes_string(this_explanatory, this_response)) +
-    geom_point() +
-    geom_smooth(method='lm', se=FALSE) +
-    theme_minimal() +
-    xlab(this_explanatory) +
-    ylab(this_response) +
-    ggtitle(sprintf('%s susceptibility to %s', this_bug, this_resistance_drug))
+  ggsave(row$plot_fn, plot=p)
 }
@@
 
-```{r}
+```{r display_results}
 res %>%
-  head %>%
-  select(bug, resistance_drug, explanatory, correlation=estimate, p.value) %>%
   mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
+  select(bug, consumption_drug, resistance_drug,
+         state_coherent, explanatory, correlation, p.value,
+         plot_link) %>%
   kable

### 2017-02-07 states/resistance/report2/state-report.Rmd
state report 2 done

@@
     ggplot(aes_string(row$explanatory, 'percent_susceptible')) +
-    geom_point() +
+    geom_point(shape=1) +
     xlab(row$explanatory) +
@@
 
-```{r display_results}
+The column `state_coherent` means that the variance in susceptibilities of the bug to the resistance drug reported in antibiograms was significantly improved by dividing the antibiogram records into states.
+The column `explanatory` is the explanatory variable in the Spearman correlation: either claims per 1,000 people or DID (days of drug per 1,000 people per day).
+
+```{r display_results, echo=FALSE}
 res %>%

### 2017-02-15 temporal/seasonal-traces/seasonal-trace.Rmd
seasonal

---
title: "Seasonal patterns for azithromycin across states"
author: "Scott Olesen"
date: "2/15/2017"
output: html_document
---

# Motivation

Azithromycin has a seasonal pattern in prescriptions. Does that pattern vary by geography? I used the Medicare data.

```{r load_data, echo=FALSE, message=FALSE}
dat2011 = read_tsv('../temporal-abx-2011.tsv')

full_dat = lapply(2011:2014, function(x) { read_tsv(sprintf('../temporal-abx-%i.tsv', x)) %>% mutate(year=x) }) %>%
  bind_rows()
```

# Data

By eye, it looks like the peak in consumption is similar across states (it occurs just after New Year's), and the minimum occurs at around the same time (weeks 27-36, which are the months of July and August).

```{r plot_azithro_two_states, echo=FALSE}
tick_labels = seq(1, 52) %>% { if_else(. %% 2 == 1, as.character(.), '') }
dat2011 %>%
  filter(abx == 'azithromycin', state %in% c('Florida', 'New York', 'Texas', 'California', 'Michigan'), week < 53.0) %>%
  ggplot(aes(x=factor(week), y=n_rx, group=state, color=state)) +
  geom_point() + geom_line() +
  xlab('week of year 2011') + ylab('# of claims') +
  scale_x_discrete(labels=tick_labels) +
  theme_minimal() +
  ggtitle('Azithromycin claims across states in 2011')
```

The pattern within a state looks fairly consistent across years

```{r plot_azithro_years, echo=FALSE}
full_dat %>%
  filter(abx == 'azithromycin', state == 'Florida', week < 53.0) %>%
  mutate(year=factor(year)) %>%
  ggplot(aes(x=factor(week), y=n_rx, group=year, color=year)) +
  geom_point() + geom_line() +
  xlab('week of year') + ylab('# of claims') +
  scale_x_discrete(labels=tick_labels) +
  theme_minimal() +
  ggtitle('Azithromycin claims in Florida across years')
```

I wrote a function to find the time until the number of prescriptions falls by half. Specifically, I:

- fit a smoothed spline (loess) to timepoints between weeks a start week and end week
- find the mean value between the predicted (i.e., the spline's values) for between the start and end weeks
- find the time at which the spline hits that value

```{r midpoint_function}
start_week = 5 # after the fast dropoff from New Years
end_week = 27 # just before the July/August lull

midpoint_time = function(n_rx, week) {
  f = loess(n_rx ~ week)
  midval = mean(c(predict(f, start_week), predict(f, end_week)))
  uniroot(function(x) predict(f, x) - midval, c(start_week, end_week))$root
}
```

It looks like there is some relationship between geography and midpoint week: in the northernmost states, consumption falls to half about 5 weeks after the southernmost states. The values are super-beautiful, but the metric I used to measure them is pretty janky.

```{r show_midpoints, echo=FALSE}
library(knitr)

full_dat %>%
  filter(abx == 'azithromycin', state %in% c('Florida', 'New York', 'Texas', 'California', 'Michigan', 'North Dakota', 'Mississippi'), week <= end_week) %>%
  group_by(year, state) %>%
  summarize(midpoint_time=midpoint_time(n_rx, week)) %>%
  mutate(midpoint_time=round(midpoint_time, 1)) %>%
  spread(year, midpoint_time) %>%
  arrange(desc(`2011`)) %>%
  kable
```

### 2017-02-15 temporal/seasonal-traces/seasonal-trace.Rmd
cirpo, doxy, etc.

@@
 ---
-title: "Seasonal patterns for azithromycin across states"
+title: "Seasonal patterns across states"
 author: "Scott Olesen"
@@
 
-Azithromycin has a seasonal pattern in prescriptions. Does that pattern vary by geography? I used the Medicare data.
+Azithromycin has a seasonal pattern in prescriptions. Does that pattern vary by geography? What about for other drugs, like ciprofloxacin or tetracycline?
+
+# Data
+
+I used the Medicare data.
 
 ```{r load_data, echo=FALSE, message=FALSE}
-dat2011 = read_tsv('../temporal-abx-2011.tsv')
-
-full_dat = lapply(2011:2014, function(x) { read_tsv(sprintf('../temporal-abx-%i.tsv', x)) %>% mutate(year=x) }) %>%
+dat = lapply(2011:2014, function(x) { read_tsv(sprintf('../temporal-abx-%i.tsv', x)) %>% mutate(year=x) }) %>%
   bind_rows()
@@
 
-# Data
+# Results
+
+## Azithromycin
 
@@
 tick_labels = seq(1, 52) %>% { if_else(. %% 2 == 1, as.character(.), '') }
-dat2011 %>%
+dat %>%
+  filter(year == 2011) %>%
   filter(abx == 'azithromycin', state %in% c('Florida', 'New York', 'Texas', 'California', 'Michigan'), week < 53.0) %>%
@@
 ```{r plot_azithro_years, echo=FALSE}
-full_dat %>%
+dat %>%
   filter(abx == 'azithromycin', state == 'Florida', week < 53.0) %>%
@@
 
-full_dat %>%
+dat %>%
   filter(abx == 'azithromycin', state %in% c('Florida', 'New York', 'Texas', 'California', 'Michigan', 'North Dakota', 'Mississippi'), week <= end_week) %>%
@@
 ```
+
+## Tetracycline
+
+There are multiple drugs with the substring "cycline". I grouped the consumption of these together for this analysis.
+
+```{r tetracycline_names, echo=FALSE}
+library(stringr)
+tetracyclines = dat$abx %>% unique %>% { .[str_detect(., 'cycline')] }
+tetracyclines
+```
+
+Tetracycline does appear to have some seasonal patterns, but they apparently differ by states: in California, Florida, and Texas (southern states), they are prescribed more in winter than in summer (although the seasonal variation seems to be around half as large as for azithromycin).
+In New York and Michigan (northern states), tetracycline claims appear to peak in the fall.
+
+```{r tetracycline_seasonality, echo=FALSE}
+dat %>%
+  filter(year == 2011) %>%
+  # group all the tetracyclines together
+  filter(abx %in% tetracyclines) %>%
+  group_by(week, state) %>%
+  summarize(n_rx=sum(n_rx)) %>%
+  filter(state %in% c('Florida', 'New York', 'Texas', 'California', 'Michigan'), week < 53.0) %>%
+  ggplot(aes(x=factor(week), y=n_rx, group=state, color=state)) +
+  geom_point() + geom_line() +
+  xlab('week of year 2011') + ylab('# of claims') +
+  scale_x_discrete(labels=tick_labels) +
+  theme_minimal() +
+  ggtitle('Tetracycline (etc.) claims across states in 2011')
+```
+
+## Ciprofloxacin
+
+For comparison, I also reproduced the plots for ciprofloxacin claims. These appear flat.
+
+```{r cipro_seasonality, echo=FALSE}
+dat %>%
+  filter(year == 2011) %>%
+  filter(abx=='ciprofloxacin', state %in% c('Florida', 'New York', 'Texas', 'California', 'Michigan'), week < 53.0) %>%
+  ggplot(aes(x=factor(week), y=n_rx, group=state, color=state)) +
+  geom_point() + geom_line() +
+  xlab('week of year 2011') + ylab('# of claims') +
+  scale_x_discrete(labels=tick_labels) +
+  theme_minimal() +
+  ggtitle('Cipro claims across states in 2011')
+```

### 2017-02-15 hicks-summary/new-hicks/hicks-summary.Rmd
new hicks

---
title: "Summary of antibiotic consumption"
author: "Scott Olesen"
date: "2/13/2017"
output: html_document
---

- Change `consumption_by_class` to `rx_by_class`

```{r setup, echo=FALSE}
# Group values in x into ranges determined by upper_limits.
# Values below the lowest limit are named as the lowest limit, and so forth.
# You should probably put Inf as the last limit.
bounds_group = function(x, upper_limits) {
  range_strings = lapply(seq(from=1, to=length(upper_limits)), function(i) {
    if (i == 1) {
      sprintf('(%i) < %.1f', i, upper_limits[i])
    } else {
      sprintf('(%i) %.1f -- %.1f', i, upper_limits[i - 1], upper_limits[i])
    }
  })
  formula_list = mapply(function(str, lim) as.formula(sprintf('x < %.1f ~ "%s"', lim, str)),
                        range_strings, upper_limits)
  do.call(case_when, formula_list)
}

summarize_by = function(x, bene, by) {
  group_by_(x, by) %>%
    summarize(n_claims=sum(n_claims)) %>%
    left_join(bene %>% group_by_(by) %>% summarize(n_ppl=n()), by=by) %>%
    mutate(cpkp=n_claims*1000/n_ppl) %>%
    select_(by, 'cpkp')
}
```

```{r load_db, echo=FALSE}
library(knitr)

db = src_sqlite('../../abx.sqlite')

regions = tbl(db, 'census_regions') %>%
  select(state, region)

abx_class = tbl(db, 'abx_class') %>%
  select(abx, abx_class)
```

```{r analyze_data, cache=TRUE, echo=FALSE}
analyze = function(this_year) {
  bene = tbl(db, sprintf('bene_%i', this_year)) %>%
    filter(!is.na(state)) %>%
    left_join(regions, by='state') %>%
    mutate(age=as.integer(age)) %>%
    compute

  pde = tbl(db, sprintf('abx_pde_%i', this_year)) %>%
    left_join(abx_class, by='abx') %>%
    left_join(bene, by='bene') %>%
    mutate(n_claims=as.integer(n_claims)) %>%
    compute

  total_n_bene = bene %>% count() %>% head(1) %>% collect %$% n

  # claims per 1,000 people by abx class
  consumption_by_class = pde %>%
    group_by(abx_class) %>%
    summarize(n_claims=sum(n_claims)) %>%
    mutate(cpkp=n_claims*1000/total_n_bene) %>%
    mutate(year=this_year) %>%
    select(abx_class, cpkp, year) %>% 
    collect

  # the top few inidividual antibiotics
  top_abx = pde %>%
    group_by(abx) %>%
    summarize(n_claims=sum(n_claims)) %>% 
    mutate(cpkp=n_claims*1000/total_n_bene) %>%
    mutate(year=this_year) %>%
    select(abx, cpkp, year) %>%
    collect

  # from here on, the denominators are taken from the beneficiary data grouped
  # in the same way as the consumption data, so we can use the summarize_by function
  claims_by_sex = summarize_by(pde, bene, 'sex') %>% mutate(year=this_year) %>% collect

  # then by age group
  # (the input to summarize_by is modified, since we want to group by age *group*,
  # not just by raw age)
  age_group_f = function(x) bounds_group(x, c(65, 75, 85, Inf))
  bene_age = bene %>%
    filter(!is.na(age)) %>%
    collect(n=Inf) %>% 
    mutate(age_group=age_group_f(age)) %>%
    count(age_group) %>%
    rename(n_bene=n)

  claims_by_age = pde %>%
    collect(n=Inf) %>%
    mutate(age_group=age_group_f(age)) %>%
    group_by(age_group) %>%
    summarize(n_claims=sum(n_claims)) %>%
    left_join(bene_age, by='age_group') %>%
    mutate(cpkp=n_claims*1000/n_bene, year=this_year) %>%
    select(age_group, cpkp, year)

  # then by census region
  claims_by_region = summarize_by(pde, bene, 'region') %>%
    mutate(year=this_year) %>% collect

  # claims by state (broken into sextiles)
  claims_by_state = summarize_by(pde, bene, 'state') %>%
    collect(n=Inf) %>%
    mutate(sextile=ntile(cpkp, 6), year=this_year)

  list(consumption_by_class=consumption_by_class,
       top_abx=top_abx,
       claims_by_sex=claims_by_sex,
       claims_by_age=claims_by_age,
       claims_by_region=claims_by_region,
       claims_by_state=claims_by_state)
}

data_2011 = analyze(2011)
data_2012 = analyze(2012)
data_2013 = analyze(2013)
data_2014 = analyze(2014)
```

```{r arrange_data}
library(knitr)
data = list("2011"=data_2011,
            "2012"=data_2012,
            "2013"=data_2013,
            "2014"=data_2014)

data_by_field = function(field) lapply(data, function(x) x[[field]]) %>% bind_rows()

claims_by_class = data_by_field('consumption_by_class')

top_classes = claims_by_class %>%
  filter(year==2011) %>%
  arrange(desc(cpkp)) %>%
  head(9) %$%
  abx_class

claims_by_class %<>%
  mutate(abx_class=if_else(abx_class %in% top_classes, abx_class, 'other')) %>%
  group_by(year, abx_class) %>%
  summarize(cpkp=sum(cpkp)) %>%
  spread(year, cpkp) %>%
  # add this is_other variable to make sure that 'other' appears in the bottom row
  # swo> could also do this with an ordered factor
  mutate(is_other=abx_class=='other') %>%
  arrange(is_other, desc(`2011`)) %>%
  select(-is_other) %>%
  kable

top_abx = data_by_field('top_abx') %>%
  spread(year, cpkp) %>%
  arrange(desc(`2011`)) %>%
  head(11) %>%
  kable

claims_by_sex = data_by_field('claims_by_sex') %>%
  filter(sex %in% c('male', 'female')) %>%
  spread(year, cpkp) %>%
  kable

claims_by_age = data_by_field('claims_by_age') %>%
  filter(!is.na(age_group)) %>%
  spread(year, cpkp) %>%
  kable

claims_by_region = data_by_field('claims_by_region') %>%
  filter(!is.na(region)) %>%
  spread(year, cpkp) %>%
  kable

state_sextiles_2011 = data_by_field('claims_by_state') %>%
  filter(!is.na(state), state != 'NA', year==2011) %>%
  select(sextile, state) %>%
  arrange(desc(sextile), state)
```

This summary is modeled after @hicks-us-2015. That study, which used the IMS Health Xponent database to infer the number of prescriptions per capita, reported higher prescription rates than what we observe here. In that paper, the number of prescriptions per 1,000 people year (PKPY) is approximately 1,000 for people over 65. In this data, we find that number to be between 600 (in 2011) and 750 (in 2012). Our value is more similar to a report that used the National Ambulatory Care Survey (NAMCS) and National Hospital Ambulatory Care Survey (NHAMCS) rather than the Xponent data. That report estimated about 500 PKPY [@flemingdutra-prevalence-2016]. A third report, which used the Xponent database, reported that prescriptions fell from about 1,000 PKPY in 1996 to about 600 PKPY in 2003 [@hicks-outpatient-2011].

# Antibiotic consumption by class

Our data for antibiotic class roughly correspond to those reported in @hicks-us-2015, when allowing for the fact that their results have about 40% higher consumption overall. The striking difference is a in consumption of quinolones: in our 2011 data, quinolones are the most-prescribed class and account for 127 PKPY. In their data, quinolones are the fourth-most-prescribed class and account for 89 PKPY (adjusted to approximately $90 \times 0.6 = 54$).

Across years, the consumption of drugs seems to follow similar patterns (excepting that consumption in 2012 was about 20% higher than in 2011 and consumption in 2013-2014 was about 10% higher).

All consumption rates are reported as PKPY.

`r claims_by_class`

# Most-prescribed antibiotics

Our data for the most prescribed individual antibiotics also generally accords with @hicks-us-2015 (again, allowing for the 40% discrepancy). The major difference is that ciprofloxacin consumption is greater in our data and amoxicillin (and augmentin) consumption are greater in their data. The reduced consumption of amoxicillin in our data could be explained by the fact that our population is older.

Allowing for the overall differences in consumption, consumption of azithromycin appears to have declined during 2011 to 2014, while the consumption of levofloxacin increased.

`r top_abx`

# Consumption by sex

Our result, that the consumption of antibiotics by women and men is in a 3:2 ratio, agrees with the result in @hicks-us-2015. The ratio also remained steady across years in our dataset.

`r claims_by_sex`

# Consumption by region

In @hicks-us-2015, stratifying consumption in 2011 by census region led to one major divide: the South, West, and Northeast consumed more antibiotics than the West (931, 897, and 848 vs. 647 PKPY). In our results, the major divide is between the South and the Midwest, West, and Northeast (641 vs. 583, 563, and 533 PKPY).

In our data, 2012 appears exceptional: consumption in the Midwest and West was relatively high, and consumption in the Northeast was relatively low. Results from 2013 and 2014 accord more closely with the results from @hicks-us-2015, in which the West is the region with the lowest consumption.

`r claims_by_region`

# Consumption by age

When stratifying by age, @hicks-us-2015 places all people 65 or older into one stratum. Here I stratify within this population. Excepting the beneficiaries under 65, older beneficiaries tend to consume more antibiotics. The beneficiaries under 65 qualify for Medicare because they are disabled, have end stage renal disease, or ALS, which might contribute to their increased consumption of antibiotics. However, this trend across age groups does not hold in 2012.

`r claims_by_age`

```{r}
# In Hicks, they also do:
# logistic regression on...
# dependent: county in top quartile for prescription rate
# independent:
#  - prescribers per capita above median,
#  - proportion of people obese above median,
#  - proportion of people <2 years old above median,
#  - proportion of people black above median,
#  - per capita income in top 2/3s,
#  - proportion with college education in top 2/3s,
#  - proportion female in top 3/4s
#
# Univariate regressions of each independent against the dependent
# Multivariate regression somehow using adjusted ratios with confounders
```

### 2017-02-16 verification/all-pde-days/all-pde-days.Rmd
pde by day

---
title: "PDE special days"
author: "Scott Olesen"
date: "2/16/2017"
output: html_document
---

The PDE data has funny bumps and wiggles. I think these are due to things like the start of the month or holidays.

# Data

Using SAS on the cluster, I collected the counts for each unique date among all the PDEs.

# Results

```{r load_data, message=FALSE, echo=FALSE}
library(lubridate)
library(knitr)
dow_levels = c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday', 'Saturday', 'Sunday')
d11 = read_tsv('pde_date_counts_2011.tsv') %>%
  mutate(date=dmy(date)) %>%
  mutate(week_n=week(date)) %>%
  mutate(day_of_week=date %>% weekdays %>% factor(levels=dow_levels, ordered=TRUE)) %>%
  group_by(day_of_week) %>%
  mutate(day_of_week_z_score=as.vector(scale(count))) %>%
  ungroup()
```

There is a clear pattern by day:
```{r plot_days}
d11 %>%
  ggplot(aes(x=date, y=count)) +
  geom_line() + geom_point() +
  xlab('day of the year 2011') + ylab('PDEs on that day') +
  theme_minimal()
```

It looks like there are a few things going on in that day:

- a fast, weekly rhythm
- a slower, monthly rhythm
- a few weirdly low points

# Patterns within a week

It looks like there is a pattern within each week:
```{r plot_weekday, echo=FALSE}
d11 %>%
  ggplot(aes(x=day_of_week, y=count)) +
  geom_boxplot() +
  xlab('') + ylab('PDEs on that day') +
  theme_minimal()
```

Some of these days fall outside the normal range. For example, there are three or four Mondays and one Thursday that fall far below the normal range.
```{r outliers_plot, echo=FALSE}
d11 %>%
  ggplot(aes(x=day_of_week, y=day_of_week_z_score)) +
  geom_boxplot() +
  xlab('day of week') + ylab('z score (within the week)') +
  theme_minimal()
```

Which days are the ones that have large z-scores?
```{r outliers}
d11 %>%
  select(day_of_week, date, day_of_week_z_score) %>%
  filter(abs(day_of_week_z_score) >= 2) %>%
  arrange(day_of_week) %>%
  kable
```

The dates with abnormally low numbers were:

- April 24 (Easter Sunday)
- May 30 (Memorial Day)
- July 4
- September 5 (Labor Day)
- November 24 (Thanksgiving)
- December 24 (Christmas Eve)
- December 25 (Christmas)

The dates with abnormally high numbers of PDEs were all the first of the month or the first work day after a holiday:

- January 2 (the Sunday after New Year's)
- January 3 (the first Monday after New Year's)
- April 1 (Friday)
- May 1 (Sunday)
- May 31 (the Tuesday after Memorial Day Monday)
- June 1 (Wednesday)
- July 1 (Friday)
- July 5 (the Tuesday after July 4 Monday)
- September 1 (Thursday)
- September 6 (the Tuesday after Labor Day Monday)
- October 1 (Saturday)
- December 1 (Thursday)

The first-of-months that are missing from this list are all Mondays and Tuesdays (i.e., otherwise busy days):

- February 1 (Tuesday)
- March 1 (Tuesday)
- August 1 (Monday)
- November 1 (Tuesday)

# Patterns across weeks

There is variation week-to-week:
```{r plot_week}
d11 %>%
  group_by(week_n) %>%
  summarize(count=sum(count)) %>%
  ggplot(aes(x=week_n, y=count)) +
  geom_point() + geom_line() +
  xlab('week of year 2011') + ylab('count') +
  theme_minimal()
```

The most obvious features are the weird last week and an apparently monthly pattern.

The last week is weird because it has only one day in it:
```{r last_week}
d11 %>% filter(week_n==53)
```

# Patterns across months

What about the monthly pattern? Which are the weeks that 
```{r exceptional_weeks}
summarize_week = function(x) {
  data_frame(start_day=min(x$date),
             end_day=max(x$date),
             count=sum(x$count))
}

d11 %>%
  group_by(week_n) %>%
  do(summarize_week(.)) %>%
  ungroup() %>%
  mutate(week_z_score=as.vector(scale(count))) %>%
  filter(abs(week_z_score) > 0.25) %>%
  arrange(week_z_score) %>%
  kable
```

The three lowest weeks are the weird last day, the week of Thanksgiving, and the week of Christmas. The highest weeks are:

- first week of December
- first week of January
- first week of March
- first week of October

and so forth.

### 2017-02-20 verification/all-pde-days/all-pde-days.Rmd
verification

@@
 There is a clear pattern by day:
-```{r plot_days}
+```{r plot_days, echo=FALSE}
 d11 %>%
@@
 Which days are the ones that have large z-scores?
-```{r outliers}
+```{r outliers, echo=FALSE}
 d11 %>%
@@
 and so forth.
+
+# 2012
+
+Was there something weird happening in 2012?
+
+```{r load_2012}
+d12 = read_tsv('pde_date_counts_2012.tsv') %>%
+  mutate(date=dmy(date)) %>%
+  mutate(week_n=week(date)) %>%
+  mutate(day_of_week=date %>% weekdays %>% factor(levels=dow_levels, ordered=TRUE)) %>%
+  group_by(day_of_week) %>%
+  mutate(day_of_week_z_score=as.vector(scale(count))) %>%
+  ungroup()
+```
+
+```{r plot_2012_days}
+d12 %>%
+  ggplot(aes(x=date, y=count)) +
+  geom_point() + geom_line() +
+  theme_minimal()
+```
+
+```{r}
+
+bind_rows(d11 %>% mutate(data_year=2011),
+          d12 %>% mutate(data_year=2012)) %>%
+  mutate(data_year=factor(data_year)) %>%
+  ggplot(aes(x=date, y=count, color=data_year, group=data_year)) +
+  geom_point() + geom_line()
+```

### 2017-02-20 verification/all-pde-days/pde-days-between-years.Rmd
verification

---
title: "PDEs between years"
author: "Scott Olesen"
date: "2/19/2017"
output: html_document
---

```{r load_data, message=FALSE, echo=FALSE}
library(lubridate)
library(knitr)

read_data = function(fn) {
  read_tsv(fn) %>%
    mutate(date=dmy(date)) %>%
    mutate(week_n=week(date))
}

read_data_year = function(y) {
  sprintf('pde_date_counts_%i.tsv', y) %>%
    read_data() %>%
    mutate(data_year=y)
}

sum_by_week = function(x) {
  data_frame(count=sum(x$count),
             date=head(x$date, 1))
}

dat = 2011:2014 %>%
  lapply(read_data_year) %>%
  bind_rows() %>%
  mutate(data_year=factor(data_year)) %>%
  group_by(data_year, year(date), week_n) %>%
  do(sum_by_week(.)) %>%
  ungroup
```

The Medicare data on the server contains Medicare Part D data for 2011 through 2014. I've been focusing on analyzing this data, especially in the files ``pdesaf2011.sas7bdat`` through ``pdesaf2014.sas7bdat``. These files contain the Part D events (PDEs), which are individual prescription drug claim records.

It seems like there is something wrong with the 2012 data. I'm hoping you can help me understand what is going on.

First, I noticed that the number of PDEs is greater in the 2012 data than in the other years:

<style type="text/css">
.table {
    width: 40%;
}
</style>

```{r counts_table, echo=FALSE}
dat %>%
  group_by(data_year) %>%
  summarize(n=sum(count)) %>%
  mutate(n=round(n/1e6, 1)) %>%
  rename(`data year`=data_year, `number of PDEs (millions)`=n) %>%
  kable(align='l')
```

This makes it seem like there were many more claims in 2012 than in the other years. However, the 2012 data file includes claims from 2011.

```{r plot_days, echo=FALSE}
dat %>%
  filter(week_n < 53) %>%
  ggplot(aes(x=date, y=count, group=data_year, color=data_year)) +
  geom_line() + geom_point() +
  ylab('PDEs in that week') +
  theme_minimal()
```

This plot shows the number of PDEs that occurred during each week, colored by the year of the data file. The 2012 data file (``pdesaf2012.sas7bdat``) contains claims information for both 2011 and 2011, while the other three files only contain claims from their years.

Furthermore, the number of claims in the 2012 data doesn't match up with the number of claims from the other years. The 2013 and 2014 data match up nicely, but it looks like there are missing claims for 2012.

```{r compute_mean_missing, echo=FALSE}
mean_missing_percent = dat %>%
  filter(year(date) == 2011) %>%
  select(date, count, data_year) %>%
  spread(data_year, count) %>%
  mutate(f=`2011`/`2012`) %$%
  mean(f) %>%
  { (. - 1) * 100 } %>% round
```

To see how many claims are "missing", I looked at PDEs from 2011. In the 2011 data file, there are on average `r mean_missing_percent`% more claims in the 2011 data file than in the 2012 data file. I'm not sure what this means. Do we have a 20% sample for the 2011, 2013, and 2014 data but only a 15% sample for the 2012 data (such that the 2011 data has $20/15 = 4/3$ fold, or 33% more, claims as the 2012 data)?

### 2017-02-20 verification/bump-in-2012/bump-analysis.Rmd
verification

---
title: "Differences in Medicare years"
author: "Scott Olesen"
date: "2/13/2017"
output:
  pdf_document:
    fig_caption: yes
---

# Motivation

There is a clear trend in the Medicare Part D data: more claims are reported in 2012 compared with 2011, 2013, and 2014.

I expected that this trend could be explained by three things:

1. There are more people enrolled in 2012 compared to the other years.
2. Enrollees in 2012 were more likely to use Part D claims.
3. Enrollees using Part D claims were more likely to have a higher number of claims in 2012.

It is easy to rule out the first two possibilities. First, the number of beneficiaries in the data only increases with each year (10.3 million in 2011, 10.7 in 2012, 11.1 in 2013, and 11.2 in 2014). Thus, there are also more claims per capita in the 2012 data.

Second, the number of enrollees who had any antibiotic prescription claim was similar across years and only slightly higher in 2012 compared to other years. In 2011, 31% of beneficiaries in the data had at least one antibiotic claim associated with them. In 2012, 38%; in 2013, 34%; in 2014, 36%.

This leaves the possibility that beneficiaries had, on average, more claims associated with them in 2012 than in other years.

# Method

I examined beneficiaries that were enrolled in the program during 2011, 2012, and 2013. To take a smaller slice, I considered beneficiaries from a single state, Florida (although the results were comparable when examining other states). I also considered only a single drug, azithromycin (although the results were comparable when considering the consumption of other drugs or of all drugs together).

For each set of beneficiaries, each drug, and each pair of consecutive years (2011-2012 and 2012-2013), I plotted the distribution of differences in number of days of drugs claimed. For example, if you took the number of days of azithromycin prescriptions for all beneficiaries in 2011 and subtracted the number of days of prescriptions for all beneficiaries in 2012 (keeping only those beneficiaries present in the data in both years), how are those numbers distributed?

If the distribution is not centered around zero, then individual users were more likely to have more days of claims in one year than the others. To ensure that peaks at zero difference have a clear meaning, I excluded beneficiaries from the plots who had zero claims in the two years. Thus, a zero value means an equal, nonzero days of claims in the two years.

# Results

```{r load_data, cache=TRUE, echo=FALSE, message=FALSE}
summarize_year = function(year) {
  fn = sprintf('../bene-abx-%s.tsv', year)
  read_tsv(fn) %>%
    mutate(state=factor(state)) %>%
    select(bene, days, azithromycin, state) %>%
    rename(abx_days=days, azi_days=azithromycin)
}

dat11 = summarize_year('2011') %>% select(bene, days11=azi_days, state)
dat12 = summarize_year('2012') %>% select(bene, days12=azi_days)
dat13 = summarize_year('2013') %>% select(bene, days13=azi_days)

dat = dat11 %>%
  filter(state=='Florida') %>%
  left_join(dat12, by='bene') %>%
  left_join(dat13, by='bene') %>%
  mutate(diff11.12=days11-days12, diff12.13=days12-days13)
```

```{r figure1, fig.width=4, fig.height=3, echo=FALSE, warning=FALSE, message=FALSE, fig.cap='Differences in azithromycin consumption in 2011 and 2012'}
dat %>%
  filter(days11 > 0 | days12 > 0) %>%
  ggplot(aes(x=diff11.12)) +
  geom_histogram() +
  xlab('days in 2011 minus days in 2012') + ylab('number of beneficiaries') +
  scale_y_log10() +
  theme_minimal()
```

```{r figure2, fig.width=4, fig.height=3, echo=FALSE, warning=FALSE, message=FALSE, fig.cap='Differences in azithromycin consumption in 2012 and 2013'}
dat %>%
  filter(days12 > 0 | days13 > 0) %>%
  ggplot(aes(x=diff12.13)) +
  geom_histogram() +
  xlab('days in 2012 minus days in 2013') + ylab('number of beneficiaries') +
  scale_y_log10() +
  theme_minimal()
```

# Discussion

Comparing 2012 and 2013, the mode of the distribution is zero: the most common scenario was for a beneficiary to have claims for the same number of days of azithromycin in the two years. However, the distribution is skewed to the left: it is more likely for a beneficiary to have had claims for more days azithromycin in 2013 than in 2012.

The comparison between 2011 and 2012 is more confusing. It appears that not one beneficiary had claims for more days of azithromycin in 2011 than in 2012.

### 2017-02-20 verification/pde-years/pde-years.Rmd
verification

--- 2017-02-20 verification/bump-in-2012/bump-analysis.Rmd
@@
 ---
-title: "Differences in Medicare years"
+title: "What entries are in which files?"
 author: "Scott Olesen"
-date: "2/13/2017"
-output:
-  pdf_document:
-    fig_caption: yes
+output: html_document
 ---
@@
 
-There is a clear trend in the Medicare Part D data: more claims are reported in 2012 compared with 2011, 2013, and 2014.
+The 2012 PDE data seems to be exceptional. Why is this the case?
 

### 2017-02-20 verification/bump-in-2012/bump-analysis.R

@@ -1,0 +1,16 @@
+summarize_year = function(year) {
+  fn = sprintf('../bene-abx-%s.tsv', year)
+  read_tsv(fn) %>%
+    mutate(state=factor(state)) %>%
+    select(bene, days, ciprofloxacin, state) %>%
+    rename(abx_days=days, azi_days=ciprofloxacin)
+}
+
+dat11 = summarize_year('2011') %>% select(bene, days11=abx_days, state)
+dat12 = summarize_year('2012') %>% select(bene, days12=abx_days)
+dat13 = summarize_year('2013') %>% select(bene, days13=abx_days)
+
+x = dat11 %>%
+  filter(state=='Florida') %>%
+  left_join(dat12, by='bene', suffix=c('.11', '.12')) %>%
+  left_join(dat13, by='bene')

### 2017-02-20 verification/pde-years/pde-years.py

@@ -1,0 +1,37 @@
+#!/usr/bin/env python3
+
+import re, argparse, sys
+
+def parse_line(line):
+    date = line.rstrip().split('\t')[1]
+    assert re.match('\d{2}[A-Z]{3}\d{2}', date)
+    day = date[0:2]
+    month = date[2:5]
+    year = date[5:9]
+    return (year, month, day)
+
+def summarize(lines):
+    data = {}
+
+    for line_i, line in enumerate(lines):
+        if line_i == 0:
+            # this is the header
+            continue
+        
+        date = parse_line(line)
+        if date not in data:
+            data[date] = 0
+
+        data[date] += 1
+
+    return data
+
+if __name__ == '__main__':
+    p = argparse.ArgumentParser(description='parse PDE files and summarize dates')
+    p.add_argument('input', type=argparse.FileType('r'), nargs='?', default=sys.stdin)
+    args = p.parse_args()
+
+    result = summarize(args.input)
+    print('year', 'month', 'day', 'count', sep='\t')
+    for k in sorted(result.keys()):
+        print(k[0], k[1], k[2], result[k], sep='\t')

### 2017-02-21 dosage-form/dosage-form.Rmd
dosage form

---
title: "Dosage forms in antibiotic PDEs"
author: "Scott Olesen"
date: "2/20/2017"
output: html_document
---

```{r load_library, echo=FALSE}
library(knitr)
```

```{r load_data, message=FALSE}
dat = 2011:2014 %>%
  lapply(function(y) read_tsv(sprintf('../abx_pde_%g.tsv', y)) %>%
           select(dosage_form_code=GCDF, dosage_form_desc=GCDF_DESC) %>%
           mutate(year=y)) %>%
  bind_rows()
```

Most codes have a unique descriptions. There are a few exceptions:
```{r code_desc}
filter_repeated_block = function(x) {
  if (nrow(x) == 1) {
    data_frame()
  } else {
    x
  }
}

dat %>%
  select(dosage_form_code, dosage_form_desc) %>%
  distinct %>%
  group_by(dosage_form_code) %>%
  do(filter_repeated_block(.)) %>%
  arrange(dosage_form_code, dosage_form_desc) %>%
  kable
```

I like the first of each of these sets, so that's the ones I'll keep.

```{r codes_by_year}
canonical_codes = dat %>%
  select(dosage_form_code, dosage_form_desc) %>%
  distinct %>%
  arrange(dosage_form_code, dosage_form_desc) %>%
  group_by(dosage_form_code) %>%
  do(head(., 1))

codes_by_year = dat %>%
  select(dosage_form_code, year) %>%
  group_by(year) %>%
  count(dosage_form_code) %>%
  spread(year, n) %>%
  left_join(canonical_codes, by='dosage_form_code') %>%
  arrange(desc(`2011`)) %>%
  kable
```

Which ones of the codes should I keep? Let's consider, for each code, the top drugs that match that code. Of these, I should only keep some.

```{r drugs_by_code, message=FALSE}
pde11 = read_tsv('../abx_pde_2011.tsv')

drugs_by_code = pde11 %>%
  select(generic_name=GNN, dosage_form_code=GCDF) %>%
  group_by(dosage_form_code) %>%
  count(generic_name) %>%
  group_by(dosage_form_code) %>%
  do(head(., 3))

codes_to_keep = c('AF', 'CA', 'CE', 'CJ', 'HH',
                  'HK', 'HM', 'HP', 'HQ', 'HS',
                  'HV', 'IF', 'IG', 'PD', 'PI',
                  'PP', 'QI', 'RG', 'RH', 'SC',
                  'SJ', 'SN', 'SO', 'ST', 'TA',
                  'TC', 'TE', 'TM', 'TS', 'UP',
                  'ZD')

drugs_by_code %>%
  mutate(keep_code=dosage_form_code %in% codes_to_keep) %>%
  rename(`generic names of top drugs with that code`=generic_name) %>%
  kable
```

### 2017-02-21 dosage-form/dosage-form.Rmd
keeping a codes-to-keep list
fixed scan
removed drops

@@
 
-```{r load_library, echo=FALSE}
-library(knitr)
-```
-
 ```{r load_data, message=FALSE}
 dat = 2011:2014 %>%
-  lapply(function(y) read_tsv(sprintf('../abx_pde_%g.tsv', y)) %>%
-           select(dosage_form_code=GCDF, dosage_form_desc=GCDF_DESC) %>%
+  lapply(function(y) read_tsv(sprintf('../../data/pde_%i.tsv', y)) %>%
+           select(generic_name=GNN, dosage_form_code=GCDF, dosage_form_desc=GCDF_DESC) %>%
            mutate(year=y)) %>%
@@
 
-Most codes have a unique descriptions. There are a few exceptions:
+Most codes have a unique descriptions.
 ```{r code_desc}
@@
 
-dat %>%
+unique_combos = dat %>%
   select(dosage_form_code, dosage_form_desc) %>%
-  distinct %>%
+  distinct
+
+unique_combos %>% kable
+```
+
+There are a few exceptions:
+
+```{r repeated_codes}
+unique_combos %>%
   group_by(dosage_form_code) %>%
@@
 
-I like the first of each of these sets, so that's the ones I'll keep.
+None of these repetitions are meaningfully different from one another, so I'll
+conclude that the code uniquely identifies the dosage type.
 
 ```{r codes_by_year}
-canonical_codes = dat %>%
-  select(dosage_form_code, dosage_form_desc) %>%
-  distinct %>%
-  arrange(dosage_form_code, dosage_form_desc) %>%
+canonical_codes = unique_combos %>%
   group_by(dosage_form_code) %>%
@@
 
-codes_by_year = dat %>%
+dat %>%
   select(dosage_form_code, year) %>%
@@
 ```{r drugs_by_code, message=FALSE}
-pde11 = read_tsv('../abx_pde_2011.tsv')
 
-drugs_by_code = pde11 %>%
-  select(generic_name=GNN, dosage_form_code=GCDF) %>%
+drugs_by_code = dat %>%
+  filter(year == 2011) %>%
+  select(generic_name, dosage_form_code) %>%
   group_by(dosage_form_code) %>%
@@
 
-codes_to_keep = c('AF', 'CA', 'CE', 'CJ', 'HH',
-                  'HK', 'HM', 'HP', 'HQ', 'HS',
-                  'HV', 'IF', 'IG', 'PD', 'PI',
-                  'PP', 'QI', 'RG', 'RH', 'SC',
-                  'SJ', 'SN', 'SO', 'ST', 'TA',
-                  'TC', 'TE', 'TM', 'TS', 'UP',
-                  'ZD')
+codes_to_keep = scan('oral_injected_codes.txt', what=character())
 

### 2017-02-21 dosage-form/dosage-form.Rmd
removed another code

@@
   lapply(function(y) read_tsv(sprintf('../../data/pde_%i.tsv', y)) %>%
-           select(generic_name=GNN, dosage_form_code=GCDF, dosage_form_desc=GCDF_DESC) %>%
+           select(brand_name=BN, generic_name=GNN, dosage_form_code=GCDF, dosage_form_desc=GCDF_DESC) %>%
            mutate(year=y)) %>%
@@
   select(dosage_form_code, dosage_form_desc) %>%
-  distinct
+  distinct %>%
+  arrange(dosage_form_code)
 
@@
   filter(year == 2011) %>%
-  select(generic_name, dosage_form_code) %>%
+  select(generic_name, brand_name, dosage_form_code) %>%
   group_by(dosage_form_code) %>%
@@
   mutate(keep_code=dosage_form_code %in% codes_to_keep) %>%
-  rename(`generic names of top drugs with that code`=generic_name) %>%
   kable

### 2017-02-22 hicks-summary/new-hicks/hicks-summary2.Rmd
new hicks

--- 2017-02-15 hicks-summary/new-hicks/hicks-summary.Rmd
@@
 
-- Change `consumption_by_class` to `rx_by_class`
+```{r setup, echo=FALSE}
+library(assertive)
 
-```{r setup, echo=FALSE}
 # Group values in x into ranges determined by upper_limits.
@@
 bounds_group = function(x, upper_limits) {
-  range_strings = lapply(seq(from=1, to=length(upper_limits)), function(i) {
+  levels = lapply(seq(from=1, to=length(upper_limits)), function(i) {
     if (i == 1) {
-      sprintf('(%i) < %.1f', i, upper_limits[i])
+      sprintf('< %.1f', upper_limits[i])
     } else {
-      sprintf('(%i) %.1f -- %.1f', i, upper_limits[i - 1], upper_limits[i])
+      sprintf('>= %.1f and < %.1f', upper_limits[i - 1], upper_limits[i])
     }
@@
   formula_list = mapply(function(str, lim) as.formula(sprintf('x < %.1f ~ "%s"', lim, str)),
-                        range_strings, upper_limits)
-  do.call(case_when, formula_list)
+                        levels, upper_limits)
+  do.call(case_when, formula_list) %>%
+    factor(levels=levels, ordered=TRUE)
 }
 
-summarize_by = function(x, bene, by) {
-  group_by_(x, by) %>%
-    summarize(n_claims=sum(n_claims)) %>%
-    left_join(bene %>% group_by_(by) %>% summarize(n_ppl=n()), by=by) %>%
+summarize_by = function(usage, bene, by) {
+  denom = group_by_(bene, by) %>%
+    summarize(n_ppl=n())
+
+  group_by_(usage, by) %>%
+    summarize(n_claims=n()) %>%
+    left_join(denom, by=by) %>%
     mutate(cpkp=n_claims*1000/n_ppl) %>%
@@
 
-```{r load_db, echo=FALSE}
-library(knitr)
+```{r load_db, echo=FALSE, message=FALSE}
 
-db = src_sqlite('../../abx.sqlite')
-
-regions = tbl(db, 'census_regions') %>%
+regions = read_tsv('../../../db/census-regions/census-regions.tsv') %>%
   select(state, region)
 
-abx_class = tbl(db, 'abx_class') %>%
-  select(abx, abx_class)
+antibiotic_class = read_tsv('../../abx_class.tsv') %>%
+  select(antibiotic, antibiotic_class) %>%
+  distinct
+
+assert_has_no_duplicates(antibiotic_class$antibiotic)
 ```
 
-```{r analyze_data, cache=TRUE, echo=FALSE}
+```{r analyze_data, cache=TRUE, echo=FALSE, message=FALSE}
 analyze = function(this_year) {
-  bene = tbl(db, sprintf('bene_%i', this_year)) %>%
-    filter(!is.na(state)) %>%
-    left_join(regions, by='state') %>%
-    mutate(age=as.integer(age)) %>%
-    compute
+  bene = read_tsv(sprintf('../../bene_%i.tsv', this_year)) %>%
+    mutate_at(vars(age, plan_coverage_months), as.integer) %>%
+    filter(!is.na(state), plan_coverage_months == 12, age >= 65) %>%
+    left_join(regions, by='state')
 
-  pde = tbl(db, sprintf('abx_pde_%i', this_year)) %>%
-    left_join(abx_class, by='abx') %>%
-    left_join(bene, by='bene') %>%
-    mutate(n_claims=as.integer(n_claims)) %>%
-    compute
+  assert_has_no_duplicates(bene$bene_id)
+  total_n_bene = nrow(bene)
 
-  total_n_bene = bene %>% count() %>% head(1) %>% collect %$% n
+  pde = read_tsv(sprintf('../../pde_%i.tsv', this_year)) %>%
+    mutate_at(vars(days_supply), as.integer) %>%
+    left_join(antibiotic_class, by='antibiotic')
+
+  usage = pde %>%
+    inner_join(bene, by='bene_id') # inner join to keep only matching rows
+
+  # usage only has PDEs from beneficiaries in our list,
+  # so there should be fewer
+  is_greater_than_or_equal_to(nrow(pde), nrow(usage))
 
   # claims per 1,000 people by abx class
-  consumption_by_class = pde %>%
-    group_by(abx_class) %>%
-    summarize(n_claims=sum(n_claims)) %>%
+  claims_by_class = usage %>%
+    group_by(antibiotic_class) %>%
+    summarize(n_claims=n()) %>%
     mutate(cpkp=n_claims*1000/total_n_bene) %>%
-    mutate(year=this_year) %>%
-    select(abx_class, cpkp, year) %>% 
-    collect
+    select(antibiotic_class, cpkp) %>%
+    mutate(year=this_year)
 
   # the top few inidividual antibiotics
-  top_abx = pde %>%
-    group_by(abx) %>%
-    summarize(n_claims=sum(n_claims)) %>% 
+  top_abx = usage %>%
+    group_by(antibiotic) %>%
+    summarize(n_claims=n()) %>%
     mutate(cpkp=n_claims*1000/total_n_bene) %>%
-    mutate(year=this_year) %>%
-    select(abx, cpkp, year) %>%
-    collect
+    select(antibiotic, cpkp) %>% 
+    mutate(year=this_year)
 
@@
   # in the same way as the consumption data, so we can use the summarize_by function
-  claims_by_sex = summarize_by(pde, bene, 'sex') %>% mutate(year=this_year) %>% collect
+  claims_by_sex = summarize_by(usage, bene, 'sex') %>%
+    mutate(year=this_year)
 
@@
   age_group_f = function(x) bounds_group(x, c(65, 75, 85, Inf))
-  bene_age = bene %>%
-    filter(!is.na(age)) %>%
-    collect(n=Inf) %>% 
-    mutate(age_group=age_group_f(age)) %>%
-    count(age_group) %>%
-    rename(n_bene=n)
-
-  claims_by_age = pde %>%
-    collect(n=Inf) %>%
-    mutate(age_group=age_group_f(age)) %>%
-    group_by(age_group) %>%
-    summarize(n_claims=sum(n_claims)) %>%
-    left_join(bene_age, by='age_group') %>%
-    mutate(cpkp=n_claims*1000/n_bene, year=this_year) %>%
-    select(age_group, cpkp, year)
+  claims_by_age = summarize_by(
+    usage %>% mutate(age_group=age_group_f(age)),
+    bene %>% mutate(age_group=age_group_f(age)),
+    'age_group') %>%
+    mutate(year=this_year)
 
   # then by census region
-  claims_by_region = summarize_by(pde, bene, 'region') %>%
-    mutate(year=this_year) %>% collect
+  claims_by_region = summarize_by(usage, bene, 'region') %>%
+    mutate(year=this_year)
 
   # claims by state (broken into sextiles)
-  claims_by_state = summarize_by(pde, bene, 'state') %>%
-    collect(n=Inf) %>%
+  claims_by_state = summarize_by(usage, bene, 'state') %>%
     mutate(sextile=ntile(cpkp, 6), year=this_year)
 
-  list(consumption_by_class=consumption_by_class,
+  list(claims_by_class=claims_by_class,
        top_abx=top_abx,
@@
 
-```{r arrange_data}
+```{r arrange_data, echo=FALSE}
 library(knitr)
@@
 
-claims_by_class = data_by_field('consumption_by_class')
+claims_by_class = data_by_field('claims_by_class')
 
@@
   head(9) %$%
-  abx_class
+  antibiotic_class %>%
+  c('other')
 
 claims_by_class %<>%
-  mutate(abx_class=if_else(abx_class %in% top_classes, abx_class, 'other')) %>%
-  group_by(year, abx_class) %>%
+  mutate(antibiotic_class=if_else(antibiotic_class %in% top_classes,
+                                  antibiotic_class,
+                                  'other')) %>%
+  group_by(year, antibiotic_class) %>%
   summarize(cpkp=sum(cpkp)) %>%
   spread(year, cpkp) %>%
-  # add this is_other variable to make sure that 'other' appears in the bottom row
-  # swo> could also do this with an ordered factor
-  mutate(is_other=abx_class=='other') %>%
-  arrange(is_other, desc(`2011`)) %>%
-  select(-is_other) %>%
+  mutate(antibiotic_class=factor(antibiotic_class, levels=top_classes, ordered=TRUE)) %>%
+  arrange(antibiotic_class) %>%
   kable
@@
   spread(year, cpkp) %>%
+  arrange(desc(`2011`)) %>%
   kable

### 2017-02-22 hicks-summary/hicks-summary.Rmd
refactor hicks

--- 2017-02-22 hicks-summary/new-hicks/hicks-summary2.Rmd
@@
 
-```{r setup, echo=FALSE}
+```{r global_options, include=FALSE}
+knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='fig/',
+                      echo=FALSE, warning=FALSE, message=FALSE)
+```
+
+```{r setup}
 library(assertive)
@@
 
-```{r load_db, echo=FALSE, message=FALSE}
+```{r load_db}
 
-regions = read_tsv('../../../db/census-regions/census-regions.tsv') %>%
+regions = read_tsv('../../db/census-regions/census-regions.tsv') %>%
   select(state, region)
 
-antibiotic_class = read_tsv('../../abx_class.tsv') %>%
+antibiotic_class = read_tsv('../abx_class.tsv') %>%
   select(antibiotic, antibiotic_class) %>%
@@
 
-```{r analyze_data, cache=TRUE, echo=FALSE, message=FALSE}
+```{r analyze_data, cache=TRUE}
 analyze = function(this_year) {
-  bene = read_tsv(sprintf('../../bene_%i.tsv', this_year)) %>%
+  bene = read_tsv(sprintf('../bene_%i.tsv', this_year)) %>%
     mutate_at(vars(age, plan_coverage_months), as.integer) %>%
@@
 
-  pde = read_tsv(sprintf('../../pde_%i.tsv', this_year)) %>%
+  pde = read_tsv(sprintf('../pde_%i.tsv', this_year)) %>%
     mutate_at(vars(days_supply), as.integer) %>%
@@
 
-```{r arrange_data, echo=FALSE}
+```{r arrange_data}
 library(knitr)

### 2017-02-22 hicks-summary/hicks-summary.R

@@ -1,121 +1,0 @@
-#!/usr/bin/env Rscript
-
-bounds_group = function(x, bounds_pairs) {
-  out = data_frame(x=x, y='NA')
-  for (i in seq_along(bounds_pairs)) {
-    bounds = bounds_pairs[[i]]
-    if (length(bounds) != 2) {
-      print(bounds)
-      sprintf("bad bounds: the %i-th bounds has length %i", i, length(bounds)) %>% stop
-    }
-
-    lower = bounds[1]
-    upper = bounds[2]
-    if (lower == -Inf) {
-      name = sprintf("(%i) < %i", i, upper)
-    } else if(upper == Inf) {
-      name = sprintf("(%i) > %i", i, lower)
-    } else {
-      name = sprintf("(%i) %i -- %i", i, lower, upper)
-    }
-    out %<>% mutate(y=if_else(between(x, lower, upper), name, y))
-  }
-  factor(out$y)
-}
-
-summarize_by = function(dat, bene, by) {
-  dat %>%
-    group_by_(by) %>%
-    summarize(n_rx=n()) %>%
-    left_join(bene %>% group_by_(by) %>% summarize(n_ppl=n()), by=by) %>%
-    mutate(rx_per_1k_ppl=n_rx*1000/n_ppl)
-}
-
-regions = read_tsv('../../db/census-regions/census-regions.tsv') %>%
-  select(state, region)
-
-analyze = function(year) {
-  bene_fn = sprintf('../bene-%s.tsv', year)
-  pde_fn = sprintf('../abx-pde-%s.tsv', year)
-
-  bene = read_tsv(bene_fn) %>%
-    # keep only the beneficiaries that don't have bogus states
-    filter(!is.na(state)) %>%
-    # merge in the census region information
-    left_join(regions, by='state')
-
-  # read in the abx class information
-  # merge the different cephalosporins into the same class
-  abx_class = read_tsv('../abx-classes.tsv') %>%
-    mutate(abx_class=if_else(startsWith(abx_class, 'cephalosporin'), 'cephalosporin', abx_class))
-
-  # read in PDE data
-  dat = read_tsv(pde_fn) %>%
-    # merge the abx class information
-    left_join(abx_class, by='abx') %>%
-    # merge the beneficiary information, keeping only claims mapping to beneficiaries
-    # for which we have data (i.e., inner join)
-    inner_join(bene, by='bene')
-
-  # total prescriptions and prescriptions per 1000 people
[…]

### 2017-02-22 states/resistance/report2/state-report.Rmd
new summaries

@@
 
-gram_negs = c('Escherichia coli', 'Klebsiella pneumoniae', 'Proteus mirabilis',
-              'Enterobacter cloacae', 'Klebisella oxytoca')
-gram_poss = c('Coagulase-negative staphylococci', 'Methicillin-resistant S. aureus (MRSA)',
-              'Methicillin-susceptible S. aureus (MSSA)', 'Streptococcus pneumoniae (non-meningitis)',
+gram_negs = c('Escherichia coli', 'Klebsiella pneumoniae',
+              'Proteus mirabilis', 'Enterobacter cloacae',
+              'Klebisella oxytoca', 'Enterococcus faecium')
+gram_poss = c('Coagulase-negative staphylococci',
+              'Methicillin-resistant S. aureus (MRSA)',
+              'Methicillin-susceptible S. aureus (MSSA)',
+              'Streptococcus pneumoniae (non-meningitis)',
               'Staphylococcus aureus')
@@
 
-consumption_drugs=c('trimethoprim-sulfamethoxazole', 'ciprofloxacin', 'levofloxacin',
-                    'ceftriaxone', 'nitrofurantoin', 'amoxicillin', 'penicillin',
-                    'doxycycline', 'azithromycin')
-resistance_drugs=c('trimethoprim-sulfamethoxazole', 'ciprofloxacin', 'levofloxacin',
-                   'ceftriaxone', 'nitrofurantoin', 'ampicillin', 'penicillin',
+consumption_drugs=c('trimethoprim-sulfamethoxazole', 'ciprofloxacin',
+                    'levofloxacin', 'ceftriaxone', 'nitrofurantoin',
+                    'amoxicillin', 'penicillin', 'doxycycline',
+                    'azithromycin')
+resistance_drugs=c('trimethoprim-sulfamethoxazole', 'ciprofloxacin',
+                   'levofloxacin', 'ceftriaxone', 'nitrofurantoin',
+                   'ampicillin', 'penicillin',
                    'tetracycline/doxycycline', 'erythromycin')
@@
 # State-level consumption data
-state_consumption = read_tsv('../../bene-consumption/state-consumption-2011.tsv') %>%
-  rename(consumption_drug=drug) %>%
+state_consumption = read_tsv('../../bene-consumption/state_usage_2011.tsv') %>%
+  rename(consumption_drug=antibiotic) %>%
   select(state, consumption_drug, claims_per_1k_ppl, did)

### 2017-02-22 states/bene-consumption/summarize-states.R

@@ -1,63 +1,37 @@
 #!/usr/bin/env Rscript
 
-f_state_bene_summary = function(bene) {
-  # chronic conditions are missing, they should come here
-  bene %>%
-    mutate(is_female=if_else(.$sex=='female', 1, 0),
-           is_white=if_else(.$race=='white', 1, 0)) %>%
-    group_by(state) %>%
-    summarize(n_bene=n(),
-              fraction_white=sum(is_white)/n_bene,
-              fraction_female=sum(is_female)/n_bene,
-              fraction_geq_65=sum(age >= 65)/n_bene,
-              mean_age=mean(age))
+# get the common chronic condition data
+all_chronic = read_tsv("../../../data/cc_all.tsv", col_types='cdddd') %>%
+  #swo> hack for column name
+  rename(bene_id=bene)
+
+summarize_year = function(year, n_max=Inf) {
+  bene_fn = sprintf('../../bene_%i.tsv', year)
+  pde_fn = sprintf('../../pde_%i.tsv', year)
+  usage_fn = sprintf('state_usage_%i.tsv', year)
+
+  year_column_name = sprintf('y%i', year)
+  chronic = all_chronic %>%
+    select_('bene_id', year_column_name) %>%
+    rename_(.dots=setNames(year_column_name, 'chronic'))
+
+  bene = read_tsv(bene_fn, n_max=n_max) %>%
+    filter(!is.na(state), plan_coverage_months==12, age >= 65) %>%
+    left_join(chronic, by='bene_id')
+
+  bene_by_state = bene %>% count(state) %>% rename(n_ppl=n)
+
+  pde = read_tsv(pde_fn, n_max=n_max) %>%
+    inner_join(bene %>% select(bene_id), by='bene_id') # keep only PDEs for beneficiaries we have
+
+  usage = pde %>%
+    left_join(bene, by='bene_id') %>%
+    group_by(state, antibiotic) %>%
+    summarize(n_claims=n(), n_days=sum(days_supply)) %>%
+    left_join(bene_by_state, by='state') %>%
+    mutate(cpkp=n_claims*1000/n_ppl, did=n_days*1000/(365*n_ppl))
+
+  write_tsv(usage, usage_fn)
 }
 
-f_state_consumption = function(abx, bene, state_bene_summary) {
-  bene %>%
-    select(bene, state) %>%
-    left_join(abx, by='bene') %>%
[…]

### 2017-02-22 summarize-bene-usage.R

@@ -3,9 +3,12 @@
 # create summary beneficiary x abx file
 
 # get the common chronic condition data
-all_chronic = read_tsv("../data/cc_all.tsv", col_types='cdddd') %>% 
+all_chronic = read_tsv("../data/cc_all.tsv", col_types='cdddd') %>%
   #swo> hack for column name
   rename(bene_id=bene)
+
+regions = read_tsv('../db/census-regions/census-regions.tsv') %>%
+  select(state, region)
 
 summarize_year = function(year, n_max=Inf) {
   bene_fn = sprintf('bene_%i.tsv', year)
@@ -18,12 +21,12 @@ summarize_year = function(year, n_max=Inf) {
     rename_(.dots=setNames(year_column_name, 'chronic'))
 
   bene = read_tsv(bene_fn, n_max=n_max) %>%
-    filter(!is.na(state), plan_coverage_months==12) %>%
-    left_join(chronic, by='bene_id')
-  #swo> filter for only those beneficiaries that appear in the chronic data?
+    filter(!is.na(state), plan_coverage_months==12, age >= 65) %>%
+    left_join(chronic, by='bene_id') %>%
+    #swo> filter for only those beneficiaries that appear in the chronic data?
+    left_join(regions, by='state')
 
   pde = read_tsv(pde_fn, n_max=n_max) %>%
-    filter(year(service_date) == year) %>%
     inner_join(bene %>% select(bene_id), by='bene_id') # keep only PDEs for beneficiaries we have
 
   usage = pde %>%

### 2017-02-22 states/resistance/report2/state-report.Rmd
new state report

@@
 
-```{r setup, echo=FALSE}
+```{r global_options, include=FALSE}
+knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/',
+                      echo=FALSE, warning=FALSE, message=FALSE)
+```
+
+```{r setup}
 library(broom)
-library(knitr)
 library(pander)
@@
 
-consumption_drugs=c('trimethoprim-sulfamethoxazole', 'ciprofloxacin',
+consumption_drugs=c('trimethoprim/sulfamethoxazole', 'ciprofloxacin',
                     'levofloxacin', 'ceftriaxone', 'nitrofurantoin',
@@
 
-```{r load_data, echo=FALSE, message=FALSE}
+```{r load_data}
 coherence = read_tsv('../../../../../antibiogram/analysis/geographic-coherence/coherences.tsv') %>%
@@
   rename(consumption_drug=antibiotic) %>%
-  select(state, consumption_drug, claims_per_1k_ppl, did)
+  select(state, consumption_drug, cpkp, did)
 
@@
 
-```{r run_models, echo=FALSE, warning=FALSE}
+```{r run_models}
 safe_spearman = function(df, expl) {
@@
   out = data_frame()
-  for (expl in c('claims_per_1k_ppl', 'did')) {
+  for (expl in c('cpkp', 'did')) {
     new_rows = safe_spearman(df, expl) %>% 
@@
 
+pvalue_stars = function(x) {
+  case_when(x < 0.001 ~ '***',
+            x < 0.01 ~ '**',
+            x < 0.05 ~ '*',
+            x < 0.1 ~ 'o',
+            TRUE ~ '')
+}
+
 res = dat %>%
@@
   left_join(state_consumption, by=c('state', 'consumption_drug')) %>%
+  mutate_at(vars(cpkp, did), function(x) if_else(is.na(x), 0, x)) %>%
   group_by(bug, resistance_drug, consumption_drug) %>%
@@
   filter(!is.na(estimate)) %>%
+  mutate(adj.p.value=p.adjust(p.value, method='BH'),
+         sig=pvalue_stars(adj.p.value)) %>%
   left_join(coherence, by=c('bug', 'resistance_drug')) %>%
@@
          explanatory, correlation=estimate, p.value,
+         adj.p.value, sig,
          plot_fn, plot_link)
@@
 
-```{r make_plots, echo=FALSE, message=FALSE}
+```{r make_plots}
 unlink('plots/*.pdf')
@@
 
-```{r display_results, echo=FALSE}
+```{r display_results}
 res %>%
   mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
-  select(bug, consumption_drug, resistance_drug,
-         state_coherent, explanatory, correlation, p.value,
+  select(bug, consumption_drug,
+         state_coherent, explanatory, correlation,
+         p.value, adj.p.value, sig,
          plot_link) %>%

### 2017-02-22 states/resistance/report/state-report.Rmd
adjust fig sizes
cleaned up report

@@
 ---
-title: 'Consumption and resistance across states'
+title: "Consumption and resistance across states"
+bibliography: refs.bib
+output:
+  html_document:
+    fig_height: 4
+  pdf_document:
+    fig_height: 3
 ---
 
-```{r echo=FALSE, message=FALSE}
-res = read_tsv('../model-results-p005.tsv')
-plot_dat = read_tsv('../plotting-dat.tsv')
+# Background
 
-plot_f = function(this_bug, this_resistance_drug, this_explanatory, this_response) {
-  this_res = res %>%
-    filter(bug==this_bug, resistance_drug==this_resistance_drug, explanatory==this_explanatory, response==this_response)
+Here I report results of comparisons antibiotic consumption and resistance across states.
+States they were a convenient middle ground between census regions (in which there is certainly a consumption/resistance association, but where having only 6 data points is disappointing) and HRRs or zipcodes (in which there might be relationships, but would require throwing out a lot of consumption data because the antibiogram data is sparse).
 
-  if (dim(this_res)[1] != 1) stop('wrong number of rows')
+In a separate analysis, I determined that the antibiogram data in certain locales, separating the data into those locales decreased variance more than could be expected by chance.
+Thus, for some bug/drug combinations, it is sensible to break the antibiogram data from census regions down into states, while for other bug/drug combinations, it does not.
 
-  z = this_res$correlation %>% head
-  p.value = this_res$p.value %>% head
+# Methods
 
-  plot_dat %>%
-    filter(bug==this_bug, resistance_drug==this_resistance_drug) %>%
-    ggplot(aes_string(this_explanatory, this_response)) +
-    geom_point() +
-    geom_smooth(method='lm', se=FALSE) +
-    theme_minimal() +
-    xlab(this_explanatory) +
-    ylab(this_response) +
-    ggtitle(sprintf('%s susceptibility to %s', this_bug, this_resistance_drug))
+I used the Medicare Part D data from 2011. Only beneficiaries greater than 65 years old were used. I also used the antibiogram data. I pooled the antiobiograms from all years.
+
+I selected a set of organisms in the antibiogram to examine.
+They mostly represent the most abundant organisms in the data.
+I also matched particular drugs in the consumption data to match to drugs in the resistance data.
+
+For each bug, consumption/resistance drug, and state, I ran a Spearman correlation test and adjusted the $p$-values using the Benjamini-Hochberg method.
+I also separately adjusted the $p$-values considering only the tests performed on coherent bug/drug/state combinations.
+
+The explanatory variable in the correlation was either claims per 1,000 people per year (CPKP) or days' supply (my proxy for daily defined dose) per 1,000 people per day (DID).
+The dependent variable was always mean percent susceptibility of that bug to that resistance drug across antibiograms from that state.
+
+# Results
+
+```{r global_options, include=FALSE}
+knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/',
+                      echo=FALSE, warning=FALSE, message=FALSE)
+```
+
+```{r setup}
+library(broom)
+library(pander)
+
+gram_negs = c('Escherichia coli', 'Klebsiella pneumoniae',
+              'Proteus mirabilis', 'Enterobacter cloacae',
+              'Klebisella oxytoca', 'Enterococcus faecium')
+gram_poss = c('Coagulase-negative staphylococci',
+              'Methicillin-resistant S. aureus (MRSA)',
+              'Methicillin-susceptible S. aureus (MSSA)',
+              'Streptococcus pneumoniae (non-meningitis)',
+              'Staphylococcus aureus')
+organisms = c(gram_negs, gram_poss)
+
+consumption_drugs=c('trimethoprim/sulfamethoxazole', 'ciprofloxacin',
+                    'levofloxacin', 'ceftriaxone', 'nitrofurantoin',
+                    'amoxicillin', 'penicillin', 'doxycycline',
+                    'azithromycin')
+resistance_drugs=c('trimethoprim-sulfamethoxazole', 'ciprofloxacin',
+                   'levofloxacin', 'ceftriaxone', 'nitrofurantoin',
+                   'ampicillin', 'penicillin',
+                   'tetracycline/doxycycline', 'erythromycin')
+
+consumption_resistance = data_frame(consumption_drug=consumption_drugs,
+                                    resistance_drug=resistance_drugs)
+
+# get all bug/drug combos (i.e., all bug/cons drug combos with appropriate res drugs)
+bug_drug = expand.grid(bug=organisms,
+                       consumption_drug=consumption_drugs,
+                       stringsAsFactors=FALSE) %>%
+  left_join(consumption_resistance, by='consumption_drug') %>%
+  select(bug, consumption_drug, resistance_drug)
+```
+
+## Matching consumption and resistance drugs
+
+I performed this matching by hand.
+
+`r consumption_resistance %>% kable`
+
+```{r load_data}
+coherence = read_tsv('../../../../../antibiogram/analysis/geographic-coherence/coherences.tsv') %>%
+  rename(resistance_drug=drug) %>%
+  mutate(resistance_drug=tolower(resistance_drug),
+         state_coherent=level %in% c('state', 'hrr', 'zipcode')) %>%
+  select(bug, resistance_drug, state_coherent)
+
+# Antibiogram data, with individual records
+abg = read_tsv('../../../../../antibiogram/data/abg.tsv') %>%
+  mutate(drug=tolower(drug)) %>%
+  rename(resistance_drug=drug) %>%
+  select(bug, resistance_drug, state, percent_susceptible)
+
+# swo> State beneficiary information could be used as control variables
+# in regressions
+#state_bene = read_tsv('../../bene-consumption/state-bene-2011.tsv') %>%
+  #filter(state %in% known_states)
+
+# State-level consumption data
+state_consumption = read_tsv('../../bene-consumption/state_usage_2011.tsv') %>%
+  rename(consumption_drug=antibiotic) %>%
+  select(state, consumption_drug, cpkp, did)
+
+dat = bug_drug %>%
+  left_join(abg, by=c('bug', 'resistance_drug')) %>%
+  left_join(state_consumption, by=c('state', 'consumption_drug'))
+```
+
+```{r run_models}
+safe_spearman = function(df, expl) {
+  if (dim(df)[1] < 3L) {
+    data_frame(estimate=NA, statistic=NA, p.value=NA, method=NA, alternative=NA)
+  } else {
+    cor.test(df[['mean_sus']], df[[expl]], method='spearman') %>% tidy
+  }
+}
+
+models = function(df) {
+  out = data_frame()
+  for (expl in c('cpkp', 'did')) {
+    new_rows = safe_spearman(df, expl) %>%
+      mutate(explanatory=expl)
+
+    out = bind_rows(out, new_rows)
+  }
+  out
+}
+
+pvalue_stars = function(x) {
+  case_when(x < 0.001 ~ '***',
+            x < 0.01 ~ '**',
+            x < 0.05 ~ '*',
+            x < 0.1 ~ 'o',
+            TRUE ~ '')
+}
+
+res = dat %>%
+  group_by(bug, resistance_drug, state) %>%
+  summarize(mean_sus=mean(percent_susceptible, na.rm=TRUE)) %>%
+  left_join(bug_drug, by=c('bug', 'resistance_drug')) %>%
+  left_join(state_consumption, by=c('state', 'consumption_drug')) %>%
+  mutate_at(vars(cpkp, did), function(x) if_else(is.na(x), 0, x)) %>%
+  group_by(bug, resistance_drug, consumption_drug) %>%
+  do(models(.)) %>%
+  ungroup %>%
+  filter(!is.na(estimate)) %>%
+  mutate(adj.p.value=p.adjust(p.value, method='BH'),
+         sig=pvalue_stars(adj.p.value)) %>%
+  left_join(coherence, by=c('bug', 'resistance_drug')) %>%
+  mutate(record_number=1:n(),
+         plot_fn=sprintf('plots/%03i.pdf', record_number),
+         plot_link=pandoc.link.return(url=plot_fn, text='plot')) %>%
+  select(bug, consumption_drug, resistance_drug, state_coherent,
+         explanatory, correlation=estimate, p.value,
+         adj.p.value, sig,
+         plot_fn, plot_link)
+
+coherent_res = res %>%
+  filter(state_coherent) %>%
+  mutate(adj.p.value=p.adjust(p.value, method='BH'),
+         sig=pvalue_stars(adj.p.value))
+```
+
+```{r make_plots, cache=TRUE}
+unlink('plots/*.pdf')
+
+for (i in 1:nrow(res)) {
+  row = res[i,]
+
+  plot_dat = dat %>%
+    filter(bug==row$bug, consumption_drug==row$consumption_drug)
+
+  p = plot_dat %>%
+    filter(bug==row$bug, consumption_drug==row$consumption_drug) %>%
+    ggplot(aes_string(row$explanatory, 'percent_susceptible')) +
+    geom_point(shape=1) +
+    xlab(row$explanatory) +
+    ylab('percent susceptible') +
+    theme_minimal()
+
+  ggsave(row$plot_fn, plot=p)
 }
@@
 
-# Methodology
+## Consumption and resistance
 
-1. I determined which bug/drug combinations were coherent at the state level (i.e., separating antibiograms into their states significantly decreased variance below what would be expected by chance)
-2. For each *consumption* drug, I asserted a matched *resistance* drug
-3. I correlated metrics of consumption (of the consumption drug) against metrics of resistance (against the resistance drug) across states.
+First, I show the results from all correlations:
 
-## Metrics
-
-The consumption metrics were:
-
-- *DID*: Defined daily doses per 1,000 inhabitants per day. In fact, I don't have proper DDDs, so these are the "days supply" supplied in the Medicare claims data. This is what Goossens *et al.* uses.
-- *Claims per 1,000 people*: Rather than days supply, this is just the raw number of claims. It also has a different denominator. This is what the Hicks papers use.
-
-In general, PKPY will be larger than DID. If all people get $d$ days of drug per prescription, the relationship is:
-
-$$
-\mathrm{PKPY} = \frac{365}{d} \times \mathrm{DID}
-$$
-
-The resistance metrics were:
-
-- *Mean* susceptibility
-- *Median* susceptibility
-- *Maximum* susceptibility
-- *Minimum* susceptibility
-
-## Correlations
-
-I used Spearman correlations and examined only those bug/drug combinations that produced $p < 0.05$. I manually examined the results and show some representative ones here.
-
-## Consumption and resistance drugs
-
-Trimethoprim-sulfamethoxazole, ciprofloxacin, levofloxacin, ceftriaxone, nitrofurantoin, and penicillin were present in both the consumption and resistance data; they are each their own pair.
-Consumption of amoxicillin was matched with resistance to ampicillin; doxycycline with tetracycline/doxycycline, and azithromycin with erythromycin.
-
-## Organisms
-
-I examined these organisms:
-
-Gram negatives:
-- Escherichia coli
-- Klebsiella pneumoniae
-- Proteus mirabilis
-- Enterobacter cloacae
-- Klebisella oxytoca
-
-Gram positives:
-- Coagulase-negative staphylococci
-- Methicillin-resistant S. aureus (MRSA)
-- Methicillin-susceptible S. aureus (MSSA)
-- Streptococcus pneumoniae (non-meningitis)
-- Staphylococcus aureus
-
-# Results
-
-## Strong results
-
-I call these strong results because:
-
-- Their p-values were small.
-- The correlations went in the right direction (i.e., increased consumption was associated with decreased susceptibility).
-- The same bug/drug/drug combinations had significant correlations for multiple metrics (e.g., *E.
-coli* susceptibility to cipro anticorrelated with cipro consumption regardless of whether consumption was measured with claims or days and whether the mean or median susceptibility was used).
-
-In general, the coefficients for these correlations were in the range 0.4--0.6.
-
-N.B.: The plots have a linear best fit line added for clarity, but the correlation was measured with Spearman (i.e., rank ordered).
-
-```{r echo=FALSE}
-plot_f('Escherichia coli', 'ciprofloxacin', 'claims_per_1k_ppl', 'mean_sus')
-plot_f('Escherichia coli', 'levofloxacin', 'claims_per_1k_ppl', 'mean_sus')
-plot_f('Methicillin-susceptible S. aureus (MSSA)', 'ciprofloxacin', 'claims_per_1k_ppl', 'mean_sus')
-plot_f('Methicillin-susceptible S. aureus (MSSA)', 'erythromycin', 'claims_per_1k_ppl', 'mean_sus')
-plot_f('Streptococcus pneumoniae (non-meningitis)', 'erythromycin', 'claims_per_1k_ppl', 'mean_sus')
+```{r display_results}
+res %>%
+  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
+  select(bug, consumption_drug,
+         state_coherent, explanatory, correlation,
+         adj.p.value, sig,
+         plot_link) %>%
+  kable
 ```
 
-## Dubious Results
+Second, I show the results from the coherent tests only. The $p$-values here are different because I adjusted those values using only this subset of tests.
 
-I call these results "dubious" because, in general:
-
-- The p-values were not as strong.
-- The correlations were sometimes in the wrong direction.
-- The susceptibility metrics were weird (e.g., consumption correlations with the minimum susceptibility reported in a state).
-- The correlation was significant for one consumption metric but not the other.
-
-```{r echo=FALSE}
-plot_f('Coagulase-negative staphylococci', 'levofloxacin', 'claims_per_1k_ppl', 'min_sus') # why only min sus?
-plot_f('Enterobacter cloacae', 'ceftriaxone', 'did', 'mean_sus') # why only DID?
-plot_f('Escherichia coli', 'ampicillin', 'did', 'median_sus') # why only DID?
-plot_f('Escherichia coli', 'trimethoprim-sulfamethoxazole', 'claims_per_1k_ppl', 'mean_sus') # why not claims per 1k ppl?
-plot_f('Klebsiella pneumoniae', 'ceftriaxone', 'did', 'mean_sus') # why only DID?
-plot_f('Methicillin-resistant S. aureus (MRSA)', 'levofloxacin', 'claims_per_1k_ppl', 'max_sus') # why only max sus?
-plot_f('Methicillin-resistant S. aureus (MRSA)', 'trimethoprim-sulfamethoxazole', 'did', 'mean_sus') # why only DID?
-plot_f('Proteus mirabilis', 'levofloxacin', 'claims_per_1k_ppl', 'min_sus') # why only min sus?
-plot_f('Streptococcus pneumoniae (non-meningitis)', 'trimethoprim-sulfamethoxazole', 'claims_per_1k_ppl', 'mean_sus') # why only some?
+```{r display_coherent_result}
+coherent_res %>%
+  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
+  select(bug, consumption_drug,
+     explanatory, correlation,
+     adj.p.value, sig,
+     plot_link) %>%
+  kable
 ```
@@
 
-## Comparison to Goossens
+## Future work
 
-Goossens *et al.* found significant correlations for:
+In the future, I intend to refine this analysis by:
 
-- Consumption of macrolides with *S. pneumoniae* susceptibility to erythromycin
-- Consumption of penicillins and cephalosporins with *S. pneumoniae* susceptibility to penicillin
-- Consumption of macrolides and lincosamids with *S. pyogenes* susceptibility to erythromycin
-- Consumption of quinolones with *E. coli* susceptibility to ciprofloxacin
-- Consumption of co-trimoxazole with *E. coli* susceptibility to co-trimoxazole
+- pooling various consumption drugs (e.g., cipro and levo consumption probably have comparable effects on cipro resistance)
+- more intelligently pooling data from the different *S. aureus* designations (in the antibiogram data, *S. aureus* is reported as one of *S. aureus*, MRSA, or MSSA)
+- comparing data from various years (i.e., is there any evidence that a consumption/resistance relationship changed drastically?)
+- pooling data from various years (i.e., if the relationship doesn't change over years, we might get more power by including
 
-The Spearman correlation coefficients range from 0.68 to 0.83, which are higher than our results.
+## Correlations are overall weak
 
-The Goossens *et al.* graph shows a best-fit line of penicillin consumption against *S.
-pneumoniae* non-susceptibility to penicillin: 5 DID corresponds to 0% non-susceptible, while 16 DID corresponds to 45% non-susceptible.
-The translates to a 4 percentage point change in susceptibility per 1 DID.
+There were only a few significant results from these tests:
 
-I can't immediately compare this to the US data (because *S. pneumoniae* resistance to penicillin was not coherent at the state level).
+- *S. aureus* and azithromycin
+- Potentially, a backward relationship for *S. aureus* and TMP-SMX
+- *S. pneumoniae* and azithromycin
+- *E. coli* and quinolones
 
-However, *E. coli* resistance to ciprofloxacin is interesting: 85% susceptibility at 50 PKPY and 75% at 200, that is, 1 percentage point in susceptibility per 15 PKPY. That's a very weak response!
+Interestingly, this accords with some of the conclusions from @bell-systematic-2014, namely:
+
+- the relationship between antibiotic consumption and resistance inside the US is weak (compared to across European countries, especially southern European countries)
+- reliable consumption/resistance relationships were found for enteric bacteria and streptococcus (though, in contrast, not with *Staphylococcus*, which is one of our strongest correlations)
+- among enteric bacteria, the only reliable consumption/resistance relationship was for quinolones

### 2017-02-22 states/resistance/state-report.Rmd
fixed filepaths

--- 2017-02-22 states/resistance/report/state-report.Rmd
@@
 ```{r load_data}
-coherence = read_tsv('../../../../../antibiogram/analysis/geographic-coherence/coherences.tsv') %>%
+coherence = read_tsv('../../../../antibiogram/analysis/geographic-coherence/coherences.tsv') %>%
   rename(resistance_drug=drug) %>%
@@
 # Antibiogram data, with individual records
-abg = read_tsv('../../../../../antibiogram/data/abg.tsv') %>%
+abg = read_tsv('../../../../antibiogram/data/abg.tsv') %>%
   mutate(drug=tolower(drug)) %>%
@@
 # State-level consumption data
-state_consumption = read_tsv('../../bene-consumption/state_usage_2011.tsv') %>%
+state_consumption = read_tsv('../bene-consumption/state_usage_2011.tsv') %>%
   rename(consumption_drug=antibiotic) %>%

### 2017-02-23 inequality/consumption-inequality.Rmd
inequality report

---
title: "Inequality in consumption"
author: "Scott Olesen"
date: "2/22/2017"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

# Background

What is the inequality in consumption for antibiotics? How does it depend on region or the antibiotic used?

# Methods

I define the Juran inequality index $J$: for a random variable $X$ with a domain $\{0, 1\}$ and a cumulative distribution function $F_X$, $J(X)$ is that value such that $F(J(X)) = 1 - J(X)$. For example, if the top 20% of beneficiaries make 80% of the claims, then $J = 0.8$.

For each group under consideration, I considered 5 different metrics:

- the fraction of beneficiaries with at least one claim
- the Juran index among all beneficiaries
- the Juran index among beneficiaries with at least one claim
- the Gini index among all beneficiaries
- the Gini index among beneficiaries with at least one claim

```{r setup}
library(ineq)
is_nonzero = function(x) x > 0
nonzero = function(x) x[is_nonzero(x)]
fraction_nonzero = function(x) sum(is_nonzero(x)) / length(x)
gini = function(x) ineq(x, type='Gini')

normalize = function(x) x / max(x)
juran = function(x) {
  fraction_consumed = cumsum(sort(x, decreasing=TRUE)) %>% normalize
  fraction_users = 1:length(x) %>% normalize
  fraction_nonzero(fraction_consumed + fraction_users - 1)
}

add_zeros = function(x, n) {
  if (n < length(x)) {
    stop(sprintf('cannot extend x of length %i to %i', length(x), n))
  }

  c(x, rep(0, n - length(x)))
}

summarize_inequality = function(x, ...) {
  do(x, inequalities(., ...))
}

inequalities = function(x, what, denom=NULL) {
  z = x[[what]]
  if (!is.null(denom)) z = add_zeros(z, denom)

  data_frame(fraction_nonzero=fraction_nonzero(z),
             juran=juran(z),
             juran_nonzero=juran(nonzero(z)),
             gini=gini(z),
             gini_nonzero=gini(nonzero(z)))
}

top_antibiotics = c('azithromycin', 'ciprofloxacin', 'levofloxacin', 'amoxicillin', 'cephalexin', 'trimethoprim/sulfamethoxazole', 'nitrofurantoin', 'doxycycline')
```

# Results

```{r load_data}
usage = read_tsv('../bene_usage_2011.tsv')
```

Overall, about half of beneficiaries took at least one antibiotic. Overall, 20% of beneficiaries claimed 80% of the antibiotic days' supply.

```{r inequality_overall, cache=TRUE}
# put in both metrics of consumption
# run it through summarize_inequality
# then group by region or by drug also
usage %>%
  group_by(bene_id) %>%
  summarize(n_claims=sum(n_claims),
            n_days=sum(n_days)) %>%
  gather('consumption_metric', 'amount', n_claims, n_days) %>%
  group_by(consumption_metric) %>%
  summarize_inequality('amount') %>%
  kable
```

Breaking down these results by region shows that the South, which consumed the most antibiotic per capita, has a slightly more even usage (Juran index of 79% versus 80% in other regions). I'm not sure that's a meaningful difference.

```{r inequality_by_region, cache=TRUE}
usage %>% 
  group_by(region, bene_id) %>%
  summarize(n_claims=sum(n_claims),
            n_days=sum(n_days)) %>%
  gather('consumption_metric', 'amount', n_claims, n_days) %>%
  group_by(region, consumption_metric) %>%
  summarize_inequality('amount') %>%
  kable
```

Among the most popular drugs, unevenness varies only slightly. Among users, azithromycin consumption has an approximately 60% Juran index, and ciprofloxacin has 65%, while TMP-SMX and vancomycin have 70%.

```{r inequality_by_drug, cache=TRUE}
n_bene = usage$bene_id %>% unique %>% length
usage %>%
  filter(!is.na(antibiotic)) %>%
  group_by(antibiotic, bene_id) %>%
  summarize(n_claims=sum(n_claims),
            n_days=sum(n_days)) %>%
  gather('consumption_metric', 'amount', n_claims, n_days) %>%
  group_by(antibiotic, consumption_metric) %>%
  summarize_inequality('amount', n_bene) %>%
  filter(antibiotic %in% top_antibiotics) %>%
  kable
```

# Discussion

I used only the 2011 data in this analysis. I intend to extend the analysis to other years. Data from other years could help confirm any of the (relatively small) differences in inequality in consumption.

### 2017-02-27 chronic-conditions/chronic-condition-drivers.Rmd
chronic

---
title: "What chronic conditions are associated with increased antibiotic consumption?"
author: "Scott Olesen"
date: "2/27/2017"
output: html_document
---

# Background

Just summing up the chronic conditions doesn't seem like a very elegant way to
summarize people's health. Choosing which chronic conditions to consider could
be a matter of expert opinion. It's also worth asking which ones are
interesting, from the data's viewpoint.

# Methods

I'll use the Medicare 2011 data.

For each chronic condition, I did a Poisson regression of the number of claims for a beneficiary against the beneficiary's age and their status for that condition. I hoped that including age in the regression would prevent a confounding between age, conditions, and consumption.

To report the results of the regression, I show the "percent increase" due to a variable in the regression. A Poisson regression estimates that
$E(Y|x) = \exp(\alpha + \beta x)$. Let $\beta_i$ be the term for the boolean variable $x_i$ of interest. Then the expected increase in consumption
due to that $x_i$ being true is $\exp(\beta_i)$.

# Results

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, 
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
```

```{r load_data}
usage = read_tsv('../bene_usage_2011.tsv')
chronic_conditions = read_tsv('cc_2011.tsv') %>%
  rename(bene_id=BENE_ID) %>%
  select(-STRCTFLG)
condition_names = read_tsv('names.txt', col_names=c('condition', 'condition_name'))
```

```{r summarize_users}
usage_summary = usage %>%
  filter(age >= 65) %>%
  group_by(bene_id) %>%
  summarize(n_claims=sum(n_claims), age=max(age)) %>%
  left_join(chronic_conditions, by='bene_id')
```

```{r compute_models}
conditions = c("AMI", "ALZH", "ALZHDMTA", "ATRIALFB", "CATARACT", "CHRNKIDN", "COPD", "CHF", "DIABETES", "GLAUCOMA", "HIPFRAC", "ISCHMCHT", "DEPRESSN", "OSTEOPRS", "RA_OA", "STRKETIA", "CNCRBRST", "CNCRCLRC", "CNCRPRST", "CNCRLUNG", "CNCRENDM", "ANEMIA", "ASTHMA", "HYPERL", "HYPERP", "HYPERT", "HYPOTH")

library(broom)
summarize_condition = function(x, condition) {
  has_condition = x[[condition]] == 3

  x %>%
    glm(n_claims ~ has_condition + age, family='poisson', data=.) %>%
    tidy %>%
    mutate(condition=condition)
}

res = lapply(conditions, function(cond) summarize_condition(usage_summary, cond)) %>%
  bind_rows() %>%
  left_join(condition_names, by='condition') %>%
  select(-condition) %>%
  rename(condition=condition_name)
```

Almost all the conditions are significantly associated with greater consumption. Here I show the results of the Poisson regression: the column `estimate` is the estimate for the parameters associated with each condition, and `p.value` measures whether the data are consistent with that parameter being zero. The column `pct.increase` shows that percent increase in number of claims derived from that estimate.

```{r show_poisson}
res %>%
  filter(term=='has_conditionTRUE') %>%
  mutate(pct.increase=round((exp(estimate)-1.0)*100)) %>%
  select(condition, pct.increase, estimate, p.value) %>%
  arrange(desc(estimate)) %>%
  kable
```

### 2017-02-27 inequality/consumption-inequality.Rmd
updates ineq, states

@@
 juran = function(x) {
+  # better to do with percent_rank and cume_dist
   fraction_consumed = cumsum(sort(x, decreasing=TRUE)) %>% normalize
@@
 
+```{r inequality_hist, cache=TRUE}
+usage %>%
+  group_by(bene_id) %>%
+  summarize(n_claims=sum(n_claims)) %>%
+  ggplot(aes(x=n_claims)) +
+  geom_histogram() +
+  scale_y_log10() +
+  xlab('number of claims') + ylab('number of beneficiaries') +
+  theme_minimal()
+```
+
 Breaking down these results by region shows that the South, which consumed the most antibiotic per capita, has a slightly more even usage (Juran index of 79% versus 80% in other regions).
@@
 
+## Characteristics by consumption
+
+I broke down the beneficiaries based on their consumption: those who had no antibiotic claims were in one "tile", and I separated the rest into $n$-tiles.
+
+```{r chars_by_consumption, cache=TRUE}
+ntile_zero = function(x, n) {
+  out = x
+  out[x > 0] = ntile(out[x > 0], n)
+  out
+}
+
+bounds_group = function(x, upper_limits) {
+  upper_limits = as.integer(upper_limits)
+  levels = lapply(seq(from=1, to=length(upper_limits)), function(i) {
+    if (i == 1) {
+      sprintf('<= %i', upper_limits[i])
+    } else {
+      sprintf('> %i and <= %i', upper_limits[i - 1], upper_limits[i])
+    }
+  })
+  top_level = sprintf('> %i', tail(upper_limits, 1))
+  formula_list = mapply(function(str, lim) as.formula(sprintf('x <= %i ~ "%s"', lim, str)),
+                        levels, upper_limits) %>%
+    c(as.formula(sprintf('TRUE ~ "%s"', top_level)))
+
+  levels = c(levels, top_level)
+
+  do.call(case_when, formula_list) %>%
+    factor(levels=levels, ordered=TRUE)
+}
+
+usage %>%
+  group_by(age, sex, race, chronic, bene_id) %>%
+  summarize(n_claims=sum(n_claims),
+            n_days=sum(n_days)) %>%
+  ungroup() %>% 
+  mutate(claims_group=bounds_group(n_claims, c(0, 1, 2, 5, 10, 20, 50, 100)),
+         is_female=if_else(sex=='female', 1, 0),
+         is_white=if_else(race=='white', 1, 0)) %>%
+  group_by(claims_group) %>%
+  summarize(n_beneficiaries=n(),
+            mean_age=mean(age),
+            percent_female=mean(is_female)*100,
+            percent_white=mean(is_white)*100,
+            mean_n_chronic_cond=mean(chronic),
+            mean_n_claims=mean(n_claims)) %>%
+  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>% 
+  kable
+```
+
 # Discussion
@@
 Data from other years could help confirm any of the (relatively small) differences in inequality in consumption.
+
+Surprisingly, the characteristics of the beneficiaries are fairly constant by consumption $n$-tile among those who took antibiotics. Those who took no antibiotics where more male and less white.

### 2017-02-27 states/resistance/state-report.Rmd
updates ineq, states

@@
 ```{r run_models}
-safe_spearman = function(df, expl) {
+safe_spearman = function(df, expl, indep) {
   if (dim(df)[1] < 3L) {
@@
   } else {
-    cor.test(df[['mean_sus']], df[[expl]], method='spearman') %>% tidy
+    cor.test(df[[indep]], df[[expl]], method='spearman') %>% tidy
   }
@@
 
-models = function(df) {
+models = function(df, indep) {
   out = data_frame()
   for (expl in c('cpkp', 'did')) {
-    new_rows = safe_spearman(df, expl) %>%
+    new_rows = safe_spearman(df, expl, indep) %>%
       mutate(explanatory=expl)
@@
   group_by(bug, resistance_drug, consumption_drug) %>%
-  do(models(.)) %>%
+  do(models(., 'mean_sus')) %>%
   ungroup %>%
@@
 
+And I try the same thing again, but this time keeping all the data points:
+
+```{r correlations_with_all_points}
+dat %>%
+  mutate_at(vars(cpkp, did), function(x) if_else(is.na(x), 0, x)) %>%
+  group_by(bug, resistance_drug, consumption_drug) %>%
+  do(models(., 'percent_susceptible')) %>%
+  ungroup %>%
+  left_join(coherence, by=c('bug', 'resistance_drug')) %>%
+  filter(!is.na(estimate), state_coherent) %>%
+  mutate(adj.p.value=p.adjust(p.value, method='BH'),
+         sig=pvalue_stars(adj.p.value)) %>%
+  select(bug, consumption_drug, resistance_drug,
+         explanatory, correlation=estimate, p.value,
+         adj.p.value, sig) %>%
+  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
+  kable
+```
+
 # Discussion

### 2017-02-27 inequality/consumption-inequality.Rmd
fixed error

@@
 knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/',
-                      echo=FALSE, warning=FALSE, message=FALSE)
+                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
 ```
@@
 
+summarize_claims = function(x) {
+  summarize(x, n_claims=sum(n_claims), n_days=sum(days_supply)) %>%
+    mutate(n_days=if_else(is.na(n_days), 0L, n_days))
+}
+
 summarize_inequality = function(x, ...) {
@@
 
-```{r inequality_overall, cache=TRUE}
+```{r inequality_overall}
 # put in both metrics of consumption
@@
   group_by(bene_id) %>%
-  summarize(n_claims=sum(n_claims),
-            n_days=sum(n_days)) %>%
+  summarize_claims() %>%
   gather('consumption_metric', 'amount', n_claims, n_days) %>%
@@
 
-```{r inequality_hist, cache=TRUE}
+```{r inequality_hist}
 usage %>%
@@
 
-```{r inequality_by_region, cache=TRUE}
+```{r inequality_by_region}
 usage %>%
   group_by(region, bene_id) %>%
-  summarize(n_claims=sum(n_claims),
-            n_days=sum(n_days)) %>%
+  summarize_claims() %>%
   gather('consumption_metric', 'amount', n_claims, n_days) %>%
@@
 
-```{r inequality_by_drug, cache=TRUE}
+```{r inequality_by_drug}
 n_bene = usage$bene_id %>% unique %>% length
 usage %>%
-  filter(!is.na(antibiotic)) %>%
   group_by(antibiotic, bene_id) %>%
-  summarize(n_claims=sum(n_claims),
-            n_days=sum(n_days)) %>%
+  summarize_claims() %>%
   gather('consumption_metric', 'amount', n_claims, n_days) %>%
@@
 
+## Inequality by state
+```{r compute_inequality_by_state}
+inequality_by_state = usage %>%
+  group_by(state, bene_id) %>%
+  summarize_claims() %>%
+  gather('consumption_metric', 'amount', n_claims, n_days) %>%
+  group_by(state, consumption_metric) %>%
+  summarize(n_bene=n(),
+             fraction_nonzero=fraction_nonzero(amount),
+             juran=juran(amount),
+             juran_nonzero=juran(nonzero(amount)),
+             gini=gini(amount),
+             gini_nonzero=gini(nonzero(amount))) %>%
+  ungroup()
+```
+
+The distribution of inequalities seems small: all states have (non-zero) Juran
+indices that fall between about 0.7 and 0.75. That being said, there does seem
+to be a trend by region: the fraction nonzero is highest in the South?
+
+```{r show_inequality_by_state}
+inequality_by_state %>%
+  filter(consumption_metric=='n_days') %>%
+  ggplot(aes(x=juran_nonzero)) + geom_histogram() +
+  theme_minimal()
+
+inequality_by_state %>%
+  filter(fraction_nonzero > 0) %>%
+  filter(consumption_metric=='n_days') %>%
+  select(-consumption_metric) %>%
+  arrange(state) %>%
+  kable
+```
+
 ## Characteristics by consumption
@@
 
-```{r chars_by_consumption, cache=TRUE}
+```{r chars_by_consumption}
 ntile_zero = function(x, n) {

### 2017-02-28 states/bene-consumption/summarize-hrrs.R
hrrs

@@ -1,0 +1,27 @@
+#!/usr/bin/env Rscript
+
+summarize_year = function(year, n_max=Inf) {
+  hrr_fn = sprintf('../../../db/hrr/hrr_%i.tsv', year)
+  usage_fn = sprintf('../../bene_usage_%i.tsv', year)
+  output_fn = sprintf('hrr_usage_%i.tsv', year)
+
+  hrr = read_tsv(hrr_fn)
+
+  usage = read_tsv(usage_fn, n_max=n_max) %>%
+    filter(age >= 65) %>%
+    left_join(hrr, by='zipcode')
+
+  denom = usage %>%
+    count(hrr) %>%
+    rename(n_bene=n)
+
+  usage %>%
+    filter(!is.na(antibiotic)) %>%
+    group_by(hrr, antibiotic) %>%
+    summarize(n_claims=sum(n_claims), n_days=sum(days_supply)) %>%
+    left_join(denom, by='hrr') %>%
+    mutate(cpkp=n_claims*1000/n_bene, did=n_days*1000/(365*n_bene)) %>%
+    write_tsv(output_fn)
+}
+
+for (y in 2011:2014) summarize_year(y)

### 2017-02-28 inequality/consumption-inequality.Rmd
more plots for state-ineq

@@
 
+These data break down, to some degree, by region: the fraction nonzero is highest for states in the South, but the inequality among users is highest for users in the Midwest.
+
+```{r show_inequality_by_region}
+regions = read_tsv('../../db/census-regions/census-regions.tsv') %>%
+  select(state, region)
+
+inequality_by_region = inequality_by_state %>%
+  filter(consumption_metric=='n_claims') %>%
+  left_join(regions, by='state')
+
+inequality_by_region %>%
+  ggplot(aes(x=n_bene, y=fraction_nonzero, color=region)) +
+  geom_point() +
+  xlab('number of beneficiaries in state') + ylab('fraction of beneficiaries with at least one claim') +
+  theme_minimal()
+
+inequality_by_region %>%
+  ggplot(aes(x=n_bene, y=juran_nonzero, color=region)) +
+  geom_point() +
+  xlab('number of beneficiaries in state') + ylab('Juran index for days of antibiotics among users') +
+  theme_minimal()
+```
+
 ## Characteristics by consumption

### 2017-02-28 states/resistance/select_plots.Rmd
select plots for presentation

---
title: "Select consumption vs resistance plots"
bibliography: refs.bib
output:
  html_document:
    fig_height: 4
  pdf_document:
    fig_height: 3
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r load_data}
# Antibiogram data, with individual records
abg = read_tsv('../../../../antibiogram/data/abg.tsv') %>%
  mutate(drug=tolower(drug)) %>%
  rename(resistance_drug=drug) %>%
  select(bug, resistance_drug, state, percent_susceptible)

# State-level consumption data
state_consumption = read_tsv('../bene-consumption/state_usage_2011.tsv') %>%
  rename(consumption_drug=antibiotic) %>%
  select(state, consumption_drug, cpkp, did)
```

E. coli against cipro resistance and consumption. First plot is against state means; second includes all data points.

```{r ecoli_cipro}
cons = state_consumption %>%
  filter(consumption_drug=='ciprofloxacin')

res = abg %>%
  filter(resistance_drug=='ciprofloxacin', bug=='Escherichia coli') %>%
  mutate(pct_nonsus=100-percent_susceptible)

res_sum = res %>%
  group_by(state) %>%
  summarize(mean_pct_nonsus=mean(pct_nonsus))

full_join(cons, res_sum, by='state') %>%
  filter(!is.na(mean_pct_nonsus), !is.na(cpkp)) %>%
  ggplot(aes(x=cpkp, y=mean_pct_nonsus)) +
  geom_point() +
  geom_smooth(method='glm', se=FALSE) +
  ylim(c(0, 50)) +
  xlab('ciprofloxacin claims per 1,000 people per year') + ylab('mean percent cirpo. non-susceptible E. coli') +
  theme_minimal()
```

```{r ecoli_cipro_all}
dat_all = bind_rows(
  res %>% mutate(summary=FALSE),
  res_sum %>% rename(pct_nonsus=mean_pct_nonsus) %>% mutate(summary=TRUE)
) %>% 
  full_join(cons, by='state') %>% 
  filter(!is.na(pct_nonsus), !is.na(cpkp))
  ggplot(dat_all %>% filter(summary), aes(x=cpkp, y=pct_nonsus)) +
  geom_point() +
  geom_point(dat=dat_all %>% filter(!summary), shape=1, size=0.5) +
  geom_smooth(dat=dat_all %>% filter(!summary), method='glm', se=FALSE, col='gray') +
  geom_smooth(method='glm', se=FALSE) +
  ylim(c(0, 75)) +
  xlab('ciprofloxacin claims per 1,000 people per year') + ylab('percent cipro. non-susceptible E. coli') +
  theme_minimal()
```

S. pneumoniae against azithromycin consumption and erythromycin resistance.

```{r strep_azi}
cons = state_consumption %>%
  filter(consumption_drug=='azithromycin')

res = abg %>%
  filter(resistance_drug=='erythromycin', bug=='Streptococcus pneumoniae (non-meningitis)') %>%
  mutate(pct_nonsus=100-percent_susceptible)

res_sum = res %>%
  group_by(state) %>%
  summarize(mean_pct_nonsus=mean(pct_nonsus))

full_join(cons, res_sum, by='state') %>%
  filter(!is.na(mean_pct_nonsus), !is.na(cpkp)) %>%
  ggplot(aes(x=cpkp, y=mean_pct_nonsus)) +
  geom_point() +
  geom_smooth(method='glm', se=FALSE) +
  ylim(c(0, 75)) +
  xlab('azithromycin claims per 1,000 people per year') + ylab('mean percent azith. non-susceptible S. pneumoniae') +
  theme_minimal()
```

```{r strep_azi_all}
dat_all = bind_rows(
  res %>% mutate(summary=FALSE),
  res_sum %>% rename(pct_nonsus=mean_pct_nonsus) %>% mutate(summary=TRUE)
) %>% 
  full_join(cons, by='state') %>% 
  filter(!is.na(pct_nonsus), !is.na(cpkp))
  ggplot(dat_all %>% filter(summary), aes(x=cpkp, y=pct_nonsus)) +
  geom_point() +
  geom_point(dat=dat_all %>% filter(!summary), shape=1, size=0.5) +
  geom_smooth(dat=dat_all %>% filter(!summary), method='glm', se=FALSE, col='gray') +
  geom_smooth(method='glm', se=FALSE) +
  ylim(c(0, 75)) +
  xlab('azithromycin claims per 1,000 people per year') + ylab('percent eryth. non-susceptible S. pneumoniae') +
  theme_minimal()
```

### 2017-02-28 states/resistance/state-report.Rmd
added parallel HRR

@@
 ---
-title: "Consumption and resistance across states"
+title: "Consumption and resistance across states and HRRs"
 bibliography: refs.bib
@@
 
-Here I report results of comparisons antibiotic consumption and resistance across states.
-States they were a convenient middle ground between census regions (in which there is certainly a consumption/resistance association, but where having only 6 data points is disappointing) and HRRs or zipcodes (in which there might be relationships, but would require throwing out a lot of consumption data because the antibiogram data is sparse).
+Here I report results of comparisons antibiotic consumption and resistance across states and HRRs.
 
@@
 Thus, for some bug/drug combinations, it is sensible to break the antibiogram data from census regions down into states, while for other bug/drug combinations, it does not.
+Similarly, HRRs can be coherent or not.
 
@@
 I also separately adjusted the $p$-values considering only the tests performed on coherent bug/drug/state combinations.
+I performed a similar analysis using HRRs.
 
 The explanatory variable in the correlation was either claims per 1,000 people per year (CPKP) or days' supply (my proxy for daily defined dose) per 1,000 people per day (DID).
-The dependent variable was always mean percent susceptibility of that bug to that resistance drug across antibiograms from that state.
+The dependent variable was always mean percent susceptibility of that bug to that resistance drug across antibiograms from that state/HRR.
 
@@
 ```{r global_options, include=FALSE}
-knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/',
+knitr::opts_chunk$set(fig.width=7, fig.height=4, cache=TRUE,
                       echo=FALSE, warning=FALSE, message=FALSE)
@@
   mutate(resistance_drug=tolower(resistance_drug),
-         state_coherent=level %in% c('state', 'hrr', 'zipcode')) %>%
-  select(bug, resistance_drug, state_coherent)
+         state_coherent=level %in% c('state', 'hrr', 'zipcode'),
+         hrr_coherent=level %in% c('hrr', 'zipcode')) %>%
+  select(bug, resistance_drug, state_coherent, hrr_coherent)
+
+hrr = read_tsv('../../../db/hrr/hrr_2011.tsv') %>%
+  select(zipcode, hrr)
 
 # Antibiogram data, with individual records
-abg = read_tsv('../../../../antibiogram/data/abg.tsv') %>%
+abg = read_tsv('../../../../antibiogram/data/abg.tsv', col_types=cols(zipcode='c')) %>%
   mutate(drug=tolower(drug)) %>%
   rename(resistance_drug=drug) %>%
-  select(bug, resistance_drug, state, percent_susceptible)
+  left_join(hrr, by='zipcode') %>%
+  select(bug, resistance_drug, state, hrr, percent_susceptible)
 
@@
 
-dat = bug_drug %>%
+hrr_consumption = read_tsv('../bene-consumption/hrr_usage_2011.tsv') %>%
+  rename(consumption_drug=antibiotic) %>%
+  select(hrr, consumption_drug, cpkp, did)
+
+state_dat = bug_drug %>%
   left_join(abg, by=c('bug', 'resistance_drug')) %>%
   left_join(state_consumption, by=c('state', 'consumption_drug'))
+
+hrr_dat = bug_drug %>%
+  left_join(abg, by=c('bug', 'resistance_drug')) %>%
+  left_join(hrr_consumption, by=c('hrr', 'consumption_drug'))
 ```
@@
 
-res = dat %>%
+state_res = state_dat %>%
   group_by(bug, resistance_drug, state) %>%
@@
 
-coherent_res = res %>%
+hrr_res = hrr_dat %>%
+  group_by(bug, resistance_drug, hrr) %>%
+  summarize(mean_sus=mean(percent_susceptible, na.rm=TRUE)) %>%
+  left_join(bug_drug, by=c('bug', 'resistance_drug')) %>%
+  left_join(hrr_consumption, by=c('hrr', 'consumption_drug')) %>%
+  mutate_at(vars(cpkp, did), function(x) if_else(is.na(x), 0, x)) %>%
+  group_by(bug, resistance_drug, consumption_drug) %>%
+  do(models(., 'mean_sus')) %>%
+  ungroup %>%
+  filter(!is.na(estimate)) %>%
+  mutate(adj.p.value=p.adjust(p.value, method='BH'),
+         sig=pvalue_stars(adj.p.value)) %>%
+  left_join(coherence, by=c('bug', 'resistance_drug')) %>%
+  mutate(record_number=1:n(),
+         plot_fn=sprintf('plots/hrr-%03i.pdf', record_number),
+         plot_link=pandoc.link.return(url=plot_fn, text='plot')) %>%
+  select(bug, consumption_drug, resistance_drug, hrr_coherent,
+         explanatory, correlation=estimate, p.value,
+         adj.p.value, sig,
+         plot_fn, plot_link)
+
+coherent_state_res = state_res %>%
   filter(state_coherent) %>%
+  mutate(adj.p.value=p.adjust(p.value, method='BH'),
+         sig=pvalue_stars(adj.p.value))
+
+coherent_hrr_res = hrr_res %>%
+  filter(hrr_coherent) %>%
   mutate(adj.p.value=p.adjust(p.value, method='BH'),
@@
 
-for (i in 1:nrow(res)) {
-  row = res[i,]
+for (i in 1:nrow(state_res)) {
+  row = state_res[i,]
 
-  plot_dat = dat %>%
+  plot_dat = state_dat %>%
+    filter(bug==row$bug, consumption_drug==row$consumption_drug)
+
+  p = plot_dat %>%
+    filter(bug==row$bug, consumption_drug==row$consumption_drug) %>%
+    ggplot(aes_string(row$explanatory, 'percent_susceptible')) +
+    geom_point(shape=1) +
+    xlab(row$explanatory) +
+    ylab('percent susceptible') +
+    theme_minimal()
+
+  ggsave(row$plot_fn, plot=p)
+}
+
+for (i in 1:nrow(hrr_res)) {
+  row = hrr_res[i,]
+
+  plot_dat = hrr_dat %>%
     filter(bug==row$bug, consumption_drug==row$consumption_drug)
@@
 
+### State
+
 ```{r display_results}
-res %>%
+state_res %>%
   mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
@@
 
+### HRR
+
+```{r}
+hrr_res %>%
+  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
+  select(bug, consumption_drug,
+         hrr_coherent, explanatory, correlation,
+         adj.p.value, sig,
+         plot_link) %>%
+  kable
+```
+
+### Coherent state
+
 Second, I show the results from the coherent tests only. The $p$-values here are different because I adjusted those values using only this subset of tests.
@@
 ```{r display_coherent_result}
-coherent_res %>%
+coherent_state_res %>%
   mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
@@
 
+### Coherent results from the HRR level
+
+```{r}
+coherent_hrr_res %>%
+  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
+  select(bug, consumption_drug,
+     explanatory, correlation,
+     adj.p.value, sig,
+     plot_link) %>%
+  kable
+```
+
+### State, all points
+
 And I try the same thing again, but this time keeping all the data points:
@@
 ```{r correlations_with_all_points}
-dat %>%
+state_dat %>%
+  mutate_at(vars(cpkp, did), function(x) if_else(is.na(x), 0, x)) %>%
+  group_by(bug, resistance_drug, consumption_drug) %>%
+  do(models(., 'percent_susceptible')) %>%
+  ungroup %>%
+  left_join(coherence, by=c('bug', 'resistance_drug')) %>%
+  filter(!is.na(estimate), hrr_coherent) %>%
+  mutate(adj.p.value=p.adjust(p.value, method='BH'),
+         sig=pvalue_stars(adj.p.value)) %>%
+  select(bug, consumption_drug, resistance_drug,
+         explanatory, correlation=estimate, p.value,
+         adj.p.value, sig) %>%
+  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
+  kable
+```
+
+### HRR, all points
+
+```{r}
+hrr_dat %>%
   mutate_at(vars(cpkp, did), function(x) if_else(is.na(x), 0, x)) %>%

### 2017-02-28 states/resistance/state-report.Rmd
refactoring to integrate

@@
   mutate(resistance_drug=tolower(resistance_drug),
-         state_coherent=level %in% c('state', 'hrr', 'zipcode'),
-         hrr_coherent=level %in% c('hrr', 'zipcode')) %>%
-  select(bug, resistance_drug, state_coherent, hrr_coherent)
+         state=level %in% c('state', 'hrr', 'zipcode'),
+         hrr=level %in% c('hrr', 'zipcode')) %>%
+  select(bug, resistance_drug, state, hrr) %>%
+  gather('place', 'coherent', state, hrr)
 
@@
 abg = read_tsv('../../../../antibiogram/data/abg.tsv', col_types=cols(zipcode='c')) %>%
-  mutate(drug=tolower(drug)) %>%
+  mutate(drug=tolower(drug), percent_nonsusceptible=100-percent_susceptible) %>%
   rename(resistance_drug=drug) %>%
   left_join(hrr, by='zipcode') %>%
-  select(bug, resistance_drug, state, hrr, percent_susceptible)
+  select(bug, resistance_drug, state, hrr, percent_nonsusceptible)
 
@@
 
-models = function(df, indep) {
-  out = data_frame()
-  for (expl in c('cpkp', 'did')) {
-    new_rows = safe_spearman(df, expl, indep) %>%
+models = function(df, indep, keep=NULL) {
+  lapply(c('cpkp', 'did'), function(expl) {
+    safe_spearman(df, expl, indep) %>%
       mutate(explanatory=expl)
-
-    out = bind_rows(out, new_rows)
-  }
-  out
+  }) %>%
+    bind_rows
 }
@@
 
-state_res = state_dat %>%
-  group_by(bug, resistance_drug, state) %>%
-  summarize(mean_sus=mean(percent_susceptible, na.rm=TRUE)) %>%
+models_result = function(x, this_place, consumption) {
+  result = x %>%
+    group_by_('bug', 'resistance_drug', this_place) %>%
+    summarize(mean_nonsus=mean(percent_nonsusceptible, na.rm=TRUE)) %>%
     left_join(bug_drug, by=c('bug', 'resistance_drug')) %>%
-  left_join(state_consumption, by=c('state', 'consumption_drug')) %>%
+    left_join(consumption, by=c(this_place, 'consumption_drug')) %>%
     mutate_at(vars(cpkp, did), function(x) if_else(is.na(x), 0, x)) %>%
     group_by(bug, resistance_drug, consumption_drug) %>%
-  do(models(., 'mean_sus')) %>%
+    do(models(., 'mean_nonsus')) %>%
     ungroup %>%
@@
            sig=pvalue_stars(adj.p.value)) %>%
-  left_join(coherence, by=c('bug', 'resistance_drug')) %>%
+    left_join(filter(coherence, place==this_place), by=c('bug', 'resistance_drug')) %>%
     mutate(record_number=1:n(),
-         plot_fn=sprintf('plots/%03i.pdf', record_number),
+           plot_fn=sprintf('plots/%s_%03i.pdf', this_place, record_number),
            plot_link=pandoc.link.return(url=plot_fn, text='plot')) %>%
-  select(bug, consumption_drug, resistance_drug, state_coherent,
+    select(bug, consumption_drug, resistance_drug, coherent,
            explanatory, correlation=estimate, p.value,
@@
 
-hrr_res = hrr_dat %>%
-  group_by(bug, resistance_drug, hrr) %>%
-  summarize(mean_sus=mean(percent_susceptible, na.rm=TRUE)) %>%
-  left_join(bug_drug, by=c('bug', 'resistance_drug')) %>%
-  left_join(hrr_consumption, by=c('hrr', 'consumption_drug')) %>%
-  mutate_at(vars(cpkp, did), function(x) if_else(is.na(x), 0, x)) %>%
-  group_by(bug, resistance_drug, consumption_drug) %>%
-  do(models(., 'mean_sus')) %>%
-  ungroup %>%
-  filter(!is.na(estimate)) %>%
-  mutate(adj.p.value=p.adjust(p.value, method='BH'),
-         sig=pvalue_stars(adj.p.value)) %>%
-  left_join(coherence, by=c('bug', 'resistance_drug')) %>%
-  mutate(record_number=1:n(),
-         plot_fn=sprintf('plots/hrr-%03i.pdf', record_number),
-         plot_link=pandoc.link.return(url=plot_fn, text='plot')) %>%
-  select(bug, consumption_drug, resistance_drug, hrr_coherent,
-         explanatory, correlation=estimate, p.value,
-         adj.p.value, sig,
-         plot_fn, plot_link)
+  # add adjusted p-values for just the coherent stuff
+  # swo> there's probably a more elegant way to do this
+  coherent.adj.p.value = result %>%
+    filter(coherent) %$%
+    p.adjust(p.value, method='BH')
+  coherent.sig = pvalue_stars(coherent.adj.p.value)
+  result$coherent.adj.p.value = NA
+  result$coherent.sig = NA
+  result$coherent.adj.p.value[result$coherent] = coherent.adj.p.value
+  result$coherent.sig[result$coherent] = coherent.sig
 
-coherent_state_res = state_res %>%
-  filter(state_coherent) %>%
-  mutate(adj.p.value=p.adjust(p.value, method='BH'),
-         sig=pvalue_stars(adj.p.value))
+  result
+}
 
-coherent_hrr_res = hrr_res %>%
-  filter(hrr_coherent) %>%
-  mutate(adj.p.value=p.adjust(p.value, method='BH'),
-         sig=pvalue_stars(adj.p.value))
+state_res = models_result(state_dat, 'state', state_consumption)
+hrr_res = models_result(hrr_dat, 'hrr', hrr_consumption)
+
+result = bind_rows(
+  state_res %>% mutate(place='state'),
+  hrr_res %>% mutate(place='hrr')
+)
 ```
 
-```{r make_plots, cache=TRUE}
+```{r make_plots}
 unlink('plots/*.pdf')
 
-for (i in 1:nrow(state_res)) {
-  row = state_res[i,]
+for (i in 1:nrow(result)) {
+  row = result[i,]
 
-  plot_dat = state_dat %>%
-    filter(bug==row$bug, consumption_drug==row$consumption_drug)
+  if (row$place == 'state') {
+    plot_dat = state_dat
+  } else if (row$place == 'hrr') {
+    plot_dat = hrr_dat
+  } else {
+    plot_dat = NULL
+  }
 
@@
     filter(bug==row$bug, consumption_drug==row$consumption_drug) %>%
-    ggplot(aes_string(row$explanatory, 'percent_susceptible')) +
+    ggplot(aes_string(row$explanatory, 'percent_nonsusceptible')) +
     geom_point(shape=1) +
     xlab(row$explanatory) +
-    ylab('percent susceptible') +
-    theme_minimal()
-
-  ggsave(row$plot_fn, plot=p)
-}
-
-for (i in 1:nrow(hrr_res)) {
-  row = hrr_res[i,]
-
-  plot_dat = hrr_dat %>%
-    filter(bug==row$bug, consumption_drug==row$consumption_drug)
-
-  p = plot_dat %>%
-    filter(bug==row$bug, consumption_drug==row$consumption_drug) %>%
-    ggplot(aes_string(row$explanatory, 'percent_susceptible')) +
-    geom_point(shape=1) +
-    xlab(row$explanatory) +
-    ylab('percent susceptible') +
+    ylab('percent non-susceptible') +
     theme_minimal()
@@
 
-First, I show the results from all correlations:
+Here a relationship where increased consumption leads to increased nonsusceptibility is encoded in a positive correlation.
 
-### State
+### Geography-averaged results
 
-```{r display_results}
-state_res %>%
+In these results, the percent nonsusceptible in each geography (state or HRR) was averaged, and the averages of the nonsusceptibilities were compared against the consumption data.
+
+#### State-level geography-averaged correlations
+
+```{r state_average_results_table}
+results %>%
+  filter(place=='state') %>%
   mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
   select(bug, consumption_drug,
-         state_coherent, explanatory, correlation,
+         coherent, explanatory, correlation,
          adj.p.value, sig,
@@
   kable
+```
+
+There correlations are mostly positive, but only a small number are significant (red line marks $p = 0.05$):
+
+```{r state_average_histogram}
+result %>%
+  filter(place=='state') %>%
+  ggplot(aes(x=correlation, y=adj.p.value)) +
+  geom_point() +
+  geom_hline(yintercept=0.05, col='red') +
+  xlab('correlation coefficient') + ylab('p value (reversed)') +
+  scale_y_reverse() +
+  theme_minimal()
 ```

### 2017-02-28 states/resistance/state-report.Rmd
refactored

@@
                        stringsAsFactors=FALSE) %>%
+  as_data_frame() %>%
   left_join(consumption_resistance, by='consumption_drug') %>%
@@
 
-### Geography-averaged results
+### All resistance-averaged results
 
@@
 
-#### State-level geography-averaged correlations
+#### All state-level averaged-resistance correlations
 
 ```{r state_average_results_table}
-results %>%
+result %>%
   filter(place=='state') %>%
@@
 
-There correlations are mostly positive, but only a small number are significant (red line marks $p = 0.05$):
+There correlations are mostly positive, but only a small number are significant (red line marks $p = 0.05$). The significant negative correlation is for MRSA and TMP-SMX.
 
-```{r state_average_histogram}
+```{r state_average_volcano}
 result %>%
@@
 
-### HRR
+#### All HRR-level averaged-resistance
 
-```{r}
-hrr_res %>%
+```{r hrr_average_results_table}
+result %>%
+  filter(place=='hrr') %>%
   mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
   select(bug, consumption_drug,
-         hrr_coherent, explanatory, correlation,
+         coherent, explanatory, correlation,
          adj.p.value, sig,
@@
 
-### Coherent state
+Like with the state-level data, only a small number are significant (red line marks $p = 0.05$), but the correlations seem more split positive and negative.
+
+```{r hrr_average_histogram}
+result %>%
+  filter(place=='hrr') %>%
+  ggplot(aes(x=correlation, y=adj.p.value)) +
+  geom_point() +
+  geom_hline(yintercept=0.05, col='red') +
+  xlab('correlation coefficient') + ylab('p value (reversed)') +
+  scale_y_reverse() +
+  theme_minimal()
+```
+
+The results from the HRR level differ: both identified a relationship between *E.
+coli* and cipro and MSSA and azithromycin.
+However, the state-level data  identified a negative relationship between MRSA and Bactrim while the HRR data identified a negative relationship between *K.
+pneumoniae* and Bactrim.
+The state-level data also identified *S.
+pneumoniae* and azithromycin.
+
+### Coherent resistance-averaged
 
@@
 
-```{r display_coherent_result}
-coherent_state_res %>%
+The plots (in the links) here are the same as before. They're actually showing all the data. I should probably have plots that show just the averaged points.
+
+#### Coherent state-level averaged-resistance correlations
+
+```{r coherent_state_average_results_table}
+result %>%
+  filter(place=='state', coherent) %>%
   mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
@@
          explanatory, correlation,
-     adj.p.value, sig,
+         coherent.adj.p.value, coherent.sig,
          plot_link) %>%
@@
 
-### Coherent results from the HRR level
+There correlations are mostly positive, and one or two more are significant (red line marks $p = 0.05$):
 
-```{r}
-coherent_hrr_res %>%
+```{r coherent_state_average_histogram}
+result %>%
+  filter(place=='state', coherent) %>%
+  ggplot(aes(x=correlation, y=coherent.adj.p.value)) +
+  geom_point() +
+  geom_hline(yintercept=0.05, col='red') +
+  xlab('correlation coefficient') + ylab('p value (reversed)') +
+  scale_y_reverse() +
+  theme_minimal()
+```
+
+#### All HRR-level averaged-resistance
+
+```{r coherent_hrr_average_results_table}
+result %>%
+  filter(place=='hrr', coherent) %>%
   mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
@@
 
-### State, all points
+```{r coherent_hrr_average_histogram}
+result %>%
+  filter(place=='hrr', coherent) %>%
+  ggplot(aes(x=correlation, y=coherent.adj.p.value)) +
+  geom_point() +
+  geom_hline(yintercept=0.05, col='red') +
+  xlab('correlation coefficient') + ylab('p value (reversed)') +
+  scale_y_reverse() +
+  theme_minimal()
+```
 
-And I try the same thing again, but this time keeping all the data points:
+### All bug/drug, all-antibiogram results
 
-```{r correlations_with_all_points}
-state_dat %>%
+In these data, I don't average the resistance before running the correlation, but I do associate each antibiogram with the same geography-wide consumption.
+
+```{r}
+all_models_result = function(x, this_place, consumption) {
+  result = x %>%
     mutate_at(vars(cpkp, did), function(x) if_else(is.na(x), 0, x)) %>%
     group_by(bug, resistance_drug, consumption_drug) %>%
-  do(models(., 'percent_susceptible')) %>%
+    do(models(., 'percent_nonsusceptible')) %>%
     ungroup %>%
-  left_join(coherence, by=c('bug', 'resistance_drug')) %>%
-  filter(!is.na(estimate), hrr_coherent) %>%
+    filter(!is.na(estimate)) %>%
     mutate(adj.p.value=p.adjust(p.value, method='BH'),
            sig=pvalue_stars(adj.p.value)) %>%
-  select(bug, consumption_drug, resistance_drug,
+    left_join(filter(coherence, place==this_place), by=c('bug', 'resistance_drug')) %>%
+    select(bug, consumption_drug, resistance_drug, coherent,
            explanatory, correlation=estimate, p.value,
+           adj.p.value, sig)
+
+  # add adjusted p-values for just the coherent stuff
+  # swo> there's probably a more elegant way to do this
+  coherent.adj.p.value = result %>%
+    filter(coherent) %$%
+    p.adjust(p.value, method='BH')
+  coherent.sig = pvalue_stars(coherent.adj.p.value)
+  result$coherent.adj.p.value = NA
+  result$coherent.sig = NA
+  result$coherent.adj.p.value[result$coherent] = coherent.adj.p.value
+  result$coherent.sig[result$coherent] = coherent.sig
+
+  result
+}
+
+all_state_res = all_models_result(state_dat, 'state', state_consumption)
+all_hrr_res = all_models_result(hrr_dat, 'hrr', hrr_consumption)
+
+all_result = bind_rows(
+  all_state_res %>% mutate(place='state'),
+  all_hrr_res %>% mutate(place='hrr')
+)
+```
+
+#### All bug/drug, state-wide consumption, all-antibiogram
+
+```{r}
+all_result %>%
+  filter(place=='state') %>% 
+  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
+  select(bug, consumption_drug, 
+         explanatory, correlation, p.value,
          adj.p.value, sig) %>%
-  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
   kable
@@
 
-### HRR, all points
+#### All bug/drug, HRR-wide consumption, all-antibiogram
 
 ```{r}
-hrr_dat %>%
-  mutate_at(vars(cpkp, did), function(x) if_else(is.na(x), 0, x)) %>%
-  group_by(bug, resistance_drug, consumption_drug) %>%
-  do(models(., 'percent_susceptible')) %>%
-  ungroup %>%
-  left_join(coherence, by=c('bug', 'resistance_drug')) %>%
-  filter(!is.na(estimate), state_coherent) %>%
-  mutate(adj.p.value=p.adjust(p.value, method='BH'),
-         sig=pvalue_stars(adj.p.value)) %>%
-  select(bug, consumption_drug, resistance_drug,
-         explanatory, correlation=estimate, p.value,
+all_result %>%
+  filter(place=='hrr') %>% 
+  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
+  select(bug, consumption_drug, 
+         explanatory, correlation, p.value,
          adj.p.value, sig) %>%
+  kable
+```
+
+#### Coherent, state-wide consumption, all-antibiogram
+
+```{r}
+all_result %>%
+  filter(place=='state', coherent) %>%
   mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
+  select(bug, consumption_drug,
+         explanatory, correlation,
+         coherent.adj.p.value, coherent.sig) %>%
+  kable
+```
+
+```{r}
+all_result %>%
+  filter(place=='state', coherent, explanatory=='cpkp') %>%
+  ggplot(aes(x=correlation, y=coherent.adj.p.value)) +
+  geom_point() +
+  geom_abline(yintercept=0.05, col='red') +
+  scale_y_reverse() +
+  theme_minimal()
+```
+
+#### Coherent, HRR-wide consumption, all-antibiogram
+
+```{r}
+all_result %>%
+  filter(place=='hrr', coherent) %>%
+  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
+  select(bug, consumption_drug,
+         explanatory, correlation,
+         coherent.adj.p.value, coherent.sig) %>%
   kable

### 2017-03-01 states/resistance/state-report.Rmd
simplifying

@@
 
-# Results
-
 ```{r global_options, include=FALSE}
@@
 organisms = c(gram_negs, gram_poss)
+```
 
-consumption_drugs=c('trimethoprim/sulfamethoxazole', 'ciprofloxacin',
-                    'levofloxacin', 'ceftriaxone', 'nitrofurantoin',
-                    'amoxicillin', 'penicillin', 'doxycycline',
-                    'azithromycin')
-resistance_drugs=c('trimethoprim-sulfamethoxazole', 'ciprofloxacin',
-                   'levofloxacin', 'ceftriaxone', 'nitrofurantoin',
-                   'ampicillin', 'penicillin',
-                   'tetracycline/doxycycline', 'erythromycin')
+## Organisms
 
-consumption_resistance = data_frame(consumption_drug=consumption_drugs,
-                                    resistance_drug=resistance_drugs)
+I selected a list of organisms from the ones that were most abundant in the data:
 
-# get all bug/drug combos (i.e., all bug/cons drug combos with appropriate res drugs)
-bug_drug = expand.grid(bug=organisms,
-                       consumption_drug=consumption_drugs,
-                       stringsAsFactors=FALSE) %>%
-  as_data_frame() %>%
-  left_join(consumption_resistance, by='consumption_drug') %>%
-  select(bug, consumption_drug, resistance_drug)
-```
+`r data_frame(organsisms=organisms) %>% kable`
 
@@
 
-I performed this matching by hand.
+I performed this matching by hand. I placed the consumption drugs into groups:
 
-`r consumption_resistance %>% kable`
+`r consumption_groups %>% kable`
+
+And did the same with the resistance drugs:
+
+`r resistance_groups %>% kable`
+
+I chose not to merge the penicillin and amoxicillin groups because antibiograms often reported very different susceptibilities for the two drugs.
+
+## Consumption and resistance data
 
 ```{r load_data}
+consumption_groups = read_tsv('consumption_groups.tsv')
+resistance_groups = read_tsv('resistance_groups.tsv')
+
 coherence = read_tsv('../../../../antibiogram/analysis/geographic-coherence/coherences.tsv') %>%
@@
   mutate(drug=tolower(drug), percent_nonsusceptible=100-percent_susceptible) %>%
-  rename(resistance_drug=drug) %>%
   left_join(hrr, by='zipcode') %>%
-  select(bug, resistance_drug, state, hrr, percent_nonsusceptible)
+  right_join(resistance_groups, by='drug') %>%
+  filter(big %in% organisms) %>%
+  select(bug, drug_group, state, hrr, percent_nonsusceptible)
 
@@
 state_consumption = read_tsv('../bene-consumption/state_usage_2011.tsv') %>%
-  rename(consumption_drug=antibiotic) %>%
-  select(state, consumption_drug, cpkp, did)
+  filter(!is.na(state)) %>%
+  right_join(consumption_groups, by=c('antibiotic'='drug')) %>%
+  group_by(state, drug_group) %>%
+  summarize_at(vars(cpkp, did), sum)
 
 hrr_consumption = read_tsv('../bene-consumption/hrr_usage_2011.tsv') %>%
-  rename(consumption_drug=antibiotic) %>%
-  select(hrr, consumption_drug, cpkp, did)
+  filter(!is.na(hrr)) %>%
+  right_join(consumption_groups, by=c('antibiotic'='drug')) %>%
+  group_by(hrr, drug_group) %>%
+  summarize_at(vars(cpkp, did), sum)
 
-state_dat = bug_drug %>%
-  left_join(abg, by=c('bug', 'resistance_drug')) %>%
-  left_join(state_consumption, by=c('state', 'consumption_drug'))
+state_dat = left_join(abg, state_consumption, by=c('drug_group', 'state'))
+hrr_dat = left_join(abg, hrr_consumption, by=c('drug_group', 'hrr'))
+```
 
-hrr_dat = bug_drug %>%
-  left_join(abg, by=c('bug', 'resistance_drug')) %>%
-  left_join(hrr_consumption, by=c('hrr', 'consumption_drug'))
-```
+## Coherence of bug/drug combinations in resistance data
+
+I want to be able to filter the bug/drug combinations by whether they are coherent in the resistance data. Does the geography (state or HRR) actually have any effect on the
 
@@
 ```
+
+# Results
 
@@
 - *E. coli* and quinolones
-
 Interestingly, this accords with some of the conclusions from @bell-systematic-2014, namely:

### 2017-03-01 states/resistance/state-report.Rmd
cosmetic

@@
 organisms = c(gram_negs, gram_poss)
+
+consumption_groups = read_tsv('consumption_groups.tsv')
+resistance_groups = read_tsv('resistance_groups.tsv')
 ```
@@
 ```{r load_data}
-consumption_groups = read_tsv('consumption_groups.tsv')
-resistance_groups = read_tsv('resistance_groups.tsv')
 
@@
   right_join(resistance_groups, by='drug') %>%
-  filter(big %in% organisms) %>%
+  filter(bug %in% organisms) %>%
   select(bug, drug_group, state, hrr, percent_nonsusceptible)
+
+# Keep only those bug/drug combinations that are in at least 20 states
+abg %<>%
+  select(bug, drug_group, state) %>%
+  distinct %>%
+  count(bug, drug_group) %>%
+  filter(n >= 20) %>%
+  left_join(abg, by=c('bug', 'drug_group'))
 
@@
 
-I want to be able to filter the bug/drug combinations by whether they are coherent in the resistance data. Does the geography (state or HRR) actually have any effect on the
+To ensure coherence, I only use the bug/drug combinations that appear in antibiograms in at least 20 states. This is equivalent to filtering those bug/drug combinations that have at least 125 records.
 
@@
   geom_point() +
-  geom_abline(yintercept=0.05, col='red') +
+  geom_hline(yintercept=0.05, col='red') +
   scale_y_reverse() +

### 2017-03-01 states/resistance/state-report.Rmd
moved to filter

@@
 
-`r data_frame(organsisms=organisms) %>% kable`
+`r organisms`
 
@@
 
+I used the 2011 Medicare data.
+
 ```{r load_data}
-
-coherence = read_tsv('../../../../antibiogram/analysis/geographic-coherence/coherences.tsv') %>%
-  rename(resistance_drug=drug) %>%
-  mutate(resistance_drug=tolower(resistance_drug),
-         state=level %in% c('state', 'hrr', 'zipcode'),
-         hrr=level %in% c('hrr', 'zipcode')) %>%
-  select(bug, resistance_drug, state, hrr) %>%
-  gather('place', 'coherent', state, hrr)
-
 hrr = read_tsv('../../../db/hrr/hrr_2011.tsv') %>%
@@
 safe_spearman = function(df, expl, indep) {
-  if (dim(df)[1] < 3L) {
+  try(return(cor.test(df[[indep]], df[[expl]], method='spearman') %>% tidy), silent=TRUE)
   data_frame(estimate=NA, statistic=NA, p.value=NA, method=NA, alternative=NA)
-  } else {
-    cor.test(df[[indep]], df[[expl]], method='spearman') %>% tidy
-  }
 }
@@
       mutate(explanatory=expl)
-  }) %>%
-    bind_rows
+  }) %>% bind_rows
 }
@@
 
-models_result = function(x, this_place, consumption) {
-  result = x %>%
-    group_by_('bug', 'resistance_drug', this_place) %>%
-    summarize(mean_nonsus=mean(percent_nonsusceptible, na.rm=TRUE)) %>%
-    left_join(bug_drug, by=c('bug', 'resistance_drug')) %>%
-    left_join(consumption, by=c(this_place, 'consumption_drug')) %>%
-    mutate_at(vars(cpkp, did), function(x) if_else(is.na(x), 0, x)) %>%
-    group_by(bug, resistance_drug, consumption_drug) %>%
+models_result = function(x, this_place) {
+  x %>%
+    group_by_('bug', 'drug_group', this_place) %>%
+    summarize(mean_nonsus=mean(percent_nonsusceptible, na.rm=TRUE),
+              cpkp=head(cpkp, 1),
+              did=head(did, 1)) %>%
+    group_by(bug, drug_group) %>%
     do(models(., 'mean_nonsus')) %>%
@@
            sig=pvalue_stars(adj.p.value)) %>%
-    left_join(filter(coherence, place==this_place), by=c('bug', 'resistance_drug')) %>%
     mutate(record_number=1:n(),
@@
            plot_link=pandoc.link.return(url=plot_fn, text='plot')) %>%
-    select(bug, consumption_drug, resistance_drug, coherent,
+    select(bug, drug_group,
            explanatory, correlation=estimate, p.value,
@@
            plot_fn, plot_link)
-
-  # add adjusted p-values for just the coherent stuff
-  # swo> there's probably a more elegant way to do this
-  coherent.adj.p.value = result %>%
-    filter(coherent) %$%
-    p.adjust(p.value, method='BH')
-  coherent.sig = pvalue_stars(coherent.adj.p.value)
-  result$coherent.adj.p.value = NA
-  result$coherent.sig = NA
-  result$coherent.adj.p.value[result$coherent] = coherent.adj.p.value
-  result$coherent.sig[result$coherent] = coherent.sig
-
-  result
 }
 
-state_res = models_result(state_dat, 'state', state_consumption)
-hrr_res = models_result(hrr_dat, 'hrr', hrr_consumption)
+state_res = models_result(state_dat, 'state')
+hrr_res = models_result(hrr_dat, 'hrr')
 
@@
   p = plot_dat %>%
-    filter(bug==row$bug, consumption_drug==row$consumption_drug) %>%
+    filter(bug==row$bug, drug_group==row$drug_group) %>%
     ggplot(aes_string(row$explanatory, 'percent_nonsusceptible')) +
@@
 
-### All resistance-averaged results
+### Resistance-averaged results
 
@@
 
-#### All state-level averaged-resistance correlations
+#### State-level resistance-averaged correlations
 
@@
   mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
-  select(bug, consumption_drug,
-         coherent, explanatory, correlation,
+  select(bug, drug_group,
+         explanatory, correlation,
          adj.p.value, sig,
@@
 
-There correlations are mostly positive, but only a small number are significant (red line marks $p = 0.05$). The significant negative correlation is for MRSA and TMP-SMX.
+(red line marks $p = 0.05$).
 
@@
 
-#### All HRR-level averaged-resistance
+#### HRR-level resistance-averaged
 
@@
 
-Like with the state-level data, only a small number are significant (red line marks $p = 0.05$), but the correlations seem more split positive and negative.
+(red line marks $p = 0.05$)
 
@@
 
-The results from the HRR level differ: both identified a relationship between *E.
+**swo adjust** The results from the HRR level differ: both identified a relationship between *E.
 coli* and cipro and MSSA and azithromycin.
@@
 
-### Coherent resistance-averaged
-
-Second, I show the results from the coherent tests only. The $p$-values here are different because I adjusted those values using only this subset of tests.
-
-The plots (in the links) here are the same as before. They're actually showing all the data. I should probably have plots that show just the averaged points.
-
-#### Coherent state-level averaged-resistance correlations
-
-```{r coherent_state_average_results_table}
-result %>%
-  filter(place=='state', coherent) %>%
-  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
-  select(bug, consumption_drug,
-         explanatory, correlation,
-         coherent.adj.p.value, coherent.sig,
-         plot_link) %>%
-  kable
-```
-
-There correlations are mostly positive, and one or two more are significant (red line marks $p = 0.05$):
-
-```{r coherent_state_average_histogram}
-result %>%
-  filter(place=='state', coherent) %>%
-  ggplot(aes(x=correlation, y=coherent.adj.p.value)) +
-  geom_point() +
-  geom_hline(yintercept=0.05, col='red') +
-  xlab('correlation coefficient') + ylab('p value (reversed)') +
-  scale_y_reverse() +
-  theme_minimal()
-```
-
-#### All HRR-level averaged-resistance
-
-```{r coherent_hrr_average_results_table}
-result %>%
-  filter(place=='hrr', coherent) %>%
-  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
-  select(bug, consumption_drug,
-         explanatory, correlation,
-         coherent.adj.p.value, coherent.sig,
-         plot_link) %>%
-  kable
-```
-
-```{r coherent_hrr_average_histogram}
-result %>%
-  filter(place=='hrr', coherent) %>%
-  ggplot(aes(x=correlation, y=coherent.adj.p.value)) +
-  geom_point() +
-  geom_hline(yintercept=0.05, col='red') +
-  xlab('correlation coefficient') + ylab('p value (reversed)') +
-  scale_y_reverse() +
-  theme_minimal()
-```
-
-### All bug/drug, all-antibiogram results
+### All-antibiogram results
 
@@
 all_models_result = function(x, this_place, consumption) {
-  result = x %>%
+  x %>%
     mutate_at(vars(cpkp, did), function(x) if_else(is.na(x), 0, x)) %>%
-    group_by(bug, resistance_drug, consumption_drug) %>%
+    group_by(bug, drug_group) %>%
     do(models(., 'percent_nonsusceptible')) %>%
@@
            sig=pvalue_stars(adj.p.value)) %>%
-    left_join(filter(coherence, place==this_place), by=c('bug', 'resistance_drug')) %>%
-    select(bug, consumption_drug, resistance_drug, coherent,
+    select(bug, drug_group,
            explanatory, correlation=estimate, p.value,
            adj.p.value, sig)
-
-  # add adjusted p-values for just the coherent stuff
-  # swo> there's probably a more elegant way to do this
-  coherent.adj.p.value = result %>%
-    filter(coherent) %$%
-    p.adjust(p.value, method='BH')
-  coherent.sig = pvalue_stars(coherent.adj.p.value)
-  result$coherent.adj.p.value = NA
-  result$coherent.sig = NA
-  result$coherent.adj.p.value[result$coherent] = coherent.adj.p.value
-  result$coherent.sig[result$coherent] = coherent.sig
-
-  result
 }
@@
 
-#### All bug/drug, state-wide consumption, all-antibiogram
+#### State-wide consumption, all-antibiogram
 
@@
 
-#### All bug/drug, HRR-wide consumption, all-antibiogram
+#### HRR-wide consumption, all-antibiogram
 
@@
          adj.p.value, sig) %>%
-  kable
-```
-
-#### Coherent, state-wide consumption, all-antibiogram
-
-```{r}
-all_result %>%
-  filter(place=='state', coherent) %>%
-  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
-  select(bug, consumption_drug,
-         explanatory, correlation,
-         coherent.adj.p.value, coherent.sig) %>%
-  kable
-```
-
-```{r}
-all_result %>%
-  filter(place=='state', coherent, explanatory=='cpkp') %>%
-  ggplot(aes(x=correlation, y=coherent.adj.p.value)) +
-  geom_point() +
-  geom_hline(yintercept=0.05, col='red') +
-  scale_y_reverse() +
-  theme_minimal()
-```
-
-#### Coherent, HRR-wide consumption, all-antibiogram
-
-```{r}
-all_result %>%
-  filter(place=='hrr', coherent) %>%
-  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
-  select(bug, consumption_drug,
-         explanatory, correlation,
-         coherent.adj.p.value, coherent.sig) %>%
   kable

### 2017-03-01 states/resistance/state-report.Rmd
improving state report

@@
 library(pander)
+library(assertr)
 
@@
   filter(bug %in% organisms) %>%
-  select(bug, drug_group, state, hrr, percent_nonsusceptible)
+  select(bug, drug_group, state, hrr, percent_nonsusceptible, n_isolates)
 
@@
   distinct %>%
-  count(bug, drug_group) %>%
-  filter(n >= 20) %>%
+  count(bug, drug_group) %>% rename(n_states=n) %>%
+  filter(n_states >= 20) %>%
   left_join(abg, by=c('bug', 'drug_group'))
@@
 ```
+
+Linear models:
+
+```{r}
+safe_lm = function(data, indep, expl) {
+  try(return(lm(sprintf('%s ~ %s', indep, expl), data) %>% tidy), silent=TRUE)
+  data_frame(term=expl, estimate=NA, std.error=NA, statistic=NA, p.value=1.0)
+}
+
+linear_models = function(df, indep) {
+  lapply(c('cpkp', 'did'), function(expl) {
+    safe_lm(df, indep, expl) %>%
+      filter(term==expl) %>%
+      rename(explanatory=term) %>%
+      verify(nrow(.) == 1)
+  }) %>% bind_rows
+}
+
+linear_models_result = function(x, this_place) {
+  x %>%
+    group_by_('bug', 'drug_group', this_place) %>%
+    summarize(mean_nonsus=mean(percent_nonsusceptible, na.rm=TRUE),
+              cpkp=head(cpkp, 1),
+              did=head(did, 1)) %>%
+    group_by(bug, drug_group) %>%
+    do(linear_models(., 'mean_nonsus')) %>%
+    ungroup %>%
+    filter(!is.na(estimate)) %>%
+    mutate(adj.p.value=p.adjust(p.value, method='BH'),
+           sig=pvalue_stars(adj.p.value)) %>%
+    select(bug, drug_group,
+           explanatory, slope=estimate, p.value,
+           adj.p.value, sig)
+}
+
+linear_state_res = linear_models_result(state_dat, 'state')
+
+linear_state_res %>% kable
+```
+
+end linear models
 
@@
   mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
-  select(bug, consumption_drug, 
+  select(bug, drug_group, 
          explanatory, correlation, p.value,

### 2017-03-01 states/resistance/state-report.Rmd
removed wrong bugs

@@
 
-gram_negs = c('Escherichia coli', 'Klebsiella pneumoniae',
+# swo> I removed Klebsiella because it's mostly in immune-compromised people
+# and K. oxytoca because it's from neonatal septicemia?
+gram_negs = c('Escherichia coli',
               'Proteus mirabilis', 'Enterobacter cloacae',
-              'Klebisella oxytoca', 'Enterococcus faecium')
+              'Enterococcus faecium')
 gram_poss = c('Coagulase-negative staphylococci',
@@
 
+#### Linear
+
+```{r}
+linear_all_models_result = function(x, this_place) {
+  x %>%
+    group_by(bug, drug_group) %>%
+    do(linear_models(., 'percent_nonsusceptible')) %>%
+    ungroup %>%
+    filter(!is.na(estimate)) %>%
+    mutate(adj.p.value=p.adjust(p.value, method='BH'),
+           sig=pvalue_stars(adj.p.value)) %>%
+    select(bug, drug_group,
+           explanatory, slope=estimate, p.value,
+           adj.p.value, sig)
+}
+
+linear_all_state_res = linear_all_models_result(state_dat, 'state')
+linear_all_hrr_res = linear_all_models_result(hrr_dat, 'hrr')
+
+linear_all_state_res %>% kable
+```
+
 # Discussion

### 2017-03-02 chronic-conditions/chronic-condition-drivers.Rmd
cleaning up Rmds

@@
 ---
-title: "What chronic conditions are associated with increased antibiotic consumption?"
+title: "Chronic conditions and increased antibiotic consumption"
 author: "Scott Olesen"
@@
 
-I'll use the Medicare 2011 data.
+I'll use the Medicare 2011 data with normal filters (age 65 or above, only beneficiaries on Medicare for 12 months of the year).
 
@@
 ```
+
+As a verification, I'll look at the distribution of number of claims among patients who have COPD or not.
+The mean for patients with COPD is more than double than that for patients without, which makes the age-adjusted 2.23 fold increase seem reasonable.
+
+```{r copd}
+usage_summary %>%
+  mutate(has_copd=COPD==3) %>%
+  group_by(has_copd) %>%
+  do(summary(.$n_claims) %>% tidy) %>%
+  kable
+```

### 2017-03-02 inequality/consumption-inequality.Rmd
cleaning up Rmds

@@
 
+```{r inequality_cumrelfreq}
+normalize = function(x) x / max(x)
+usage %>%
+  group_by(bene_id) %>%
+  summarize(n_claims=sum(n_claims)) %>%
+  arrange(n_claims) %>%
+  mutate(cdf_x=cume_dist(n_claims), cdf_y=normalize(cumsum(n_claims))) %>%
+  group_by(cdf_x) %>%
+  summarize(cdf_y=max(cdf_y)) %>%
+  ggplot(aes(x=cdf_x, y=cdf_y)) +
+  geom_point() + geom_line() +
+  xlab('cumulative fraction of beneficiaries') +
+  ylab('cumulative fraction of claims') +
+  xlim(0, 1) + ylim(0, 1) +
+  theme_minimal()
+```
+
 Breaking down these results by region shows that the South, which consumed the most antibiotic per capita, has a slightly more even usage (Juran index of 79% versus 80% in other regions).
@@
 to be a trend by region: the fraction nonzero is highest in the South?
+
+These results all show inequalities measured using number of days.
 
@@
 
+As expected, greater consumption is correlated with greater age and comorbidity.
+There are strange trends among gender and race (higher consumption is associated with more female and more white until 10--20 claims, after which point those trends reverse).
+
 # Discussion
@@
 Data from other years could help confirm any of the (relatively small) differences in inequality in consumption.
-
-Surprisingly, the characteristics of the beneficiaries are fairly constant by consumption $n$-tile among those who took antibiotics. Those who took no antibiotics where more male and less white.

### 2017-03-02 states/resistance/state-report.Rmd
cleaning up Rmds

@@
 I used the 2011 Medicare data.
+To ensure coherence, I only use the bug/drug combinations that appear in antibiograms in at least 20 states.
+This is equivalent to filtering those bug/drug combinations that have at least 125 records.
 
@@
 ```
-
-## Coherence of bug/drug combinations in resistance data
-
-To ensure coherence, I only use the bug/drug combinations that appear in antibiograms in at least 20 states. This is equivalent to filtering those bug/drug combinations that have at least 125 records.
 
@@
 ```
-
-Linear models:
-
-```{r}
-safe_lm = function(data, indep, expl) {
-  try(return(lm(sprintf('%s ~ %s', indep, expl), data) %>% tidy), silent=TRUE)
-  data_frame(term=expl, estimate=NA, std.error=NA, statistic=NA, p.value=1.0)
-}
-
-linear_models = function(df, indep) {
-  lapply(c('cpkp', 'did'), function(expl) {
-    safe_lm(df, indep, expl) %>%
-      filter(term==expl) %>%
-      rename(explanatory=term) %>%
-      verify(nrow(.) == 1)
-  }) %>% bind_rows
-}
-
-linear_models_result = function(x, this_place) {
-  x %>%
-    group_by_('bug', 'drug_group', this_place) %>%
-    summarize(mean_nonsus=mean(percent_nonsusceptible, na.rm=TRUE),
-              cpkp=head(cpkp, 1),
-              did=head(did, 1)) %>%
-    group_by(bug, drug_group) %>%
-    do(linear_models(., 'mean_nonsus')) %>%
-    ungroup %>%
-    filter(!is.na(estimate)) %>%
-    mutate(adj.p.value=p.adjust(p.value, method='BH'),
-           sig=pvalue_stars(adj.p.value)) %>%
-    select(bug, drug_group,
-           explanatory, slope=estimate, p.value,
-           adj.p.value, sig)
-}
-
-linear_state_res = linear_models_result(state_dat, 'state')
-
-linear_state_res %>% kable
-```
-
-end linear models
 
@@
 
+#### State-level resistance-averaged correlations using linear models
+
+```{r}
+safe_lm = function(data, indep, expl) {
+  try(return(lm(sprintf('%s ~ %s', indep, expl), data) %>% tidy), silent=TRUE)
+  data_frame(term=expl, estimate=NA, std.error=NA, statistic=NA, p.value=1.0)
+}
+
+linear_models = function(df, indep) {
+  lapply(c('cpkp', 'did'), function(expl) {
+    safe_lm(df, indep, expl) %>%
+      filter(term==expl) %>%
+      rename(explanatory=term) %>%
+      verify(nrow(.) == 1)
+  }) %>% bind_rows
+}
+
+linear_models_result = function(x, this_place) {
+  x %>%
+    group_by_('bug', 'drug_group', this_place) %>%
+    summarize(mean_nonsus=mean(percent_nonsusceptible, na.rm=TRUE),
+              cpkp=head(cpkp, 1),
+              did=head(did, 1)) %>%
+    group_by(bug, drug_group) %>%
+    do(linear_models(., 'mean_nonsus')) %>%
+    ungroup %>%
+    filter(!is.na(estimate)) %>%
+    mutate(adj.p.value=p.adjust(p.value, method='BH'),
+           sig=pvalue_stars(adj.p.value)) %>%
+    select(bug, drug_group,
+           explanatory, slope=estimate, p.value,
+           adj.p.value, sig)
+}
+
+linear_state_res = linear_models_result(state_dat, 'state')
+
+linear_state_res %>% kable
+```
+
 #### HRR-level resistance-averaged
@@
 ```
-
-**swo adjust** The results from the HRR level differ: both identified a relationship between *E.
-coli* and cipro and MSSA and azithromycin.
-However, the state-level data  identified a negative relationship between MRSA and Bactrim while the HRR data identified a negative relationship between *K.
-pneumoniae* and Bactrim.
-The state-level data also identified *S.
-pneumoniae* and azithromycin.
 
@@
 
-#### HRR-wide consumption, all-antibiogram
-
-```{r}
-all_result %>%
-  filter(place=='hrr') %>% 
-  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
-  select(bug, drug_group, 
-         explanatory, correlation, p.value,
-         adj.p.value, sig) %>%
-  kable
-```
-
-#### Linear
+#### State-wide consumtion, all-antibiogram using linear models
 
@@
 linear_all_state_res %>% kable
+```
+
+#### HRR-wide consumption, all-antibiogram
+
+```{r}
+all_result %>%
+  filter(place=='hrr') %>% 
+  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
+  select(bug, drug_group, 
+         explanatory, correlation, p.value,
+         adj.p.value, sig) %>%
+  kable
 ```

### 2017-03-03 chronic-conditions/chronic-condition-drivers.Rmd
chronic conditions, better

@@
                       echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
+library(psych)
+library(gplots)
 ```
@@
   group_by(bene_id) %>%
-  summarize(n_claims=sum(n_claims), age=max(age)) %>%
+  summarize(n_claims=sum(n_claims), age=max(age), sex=head(sex, 1)) %>%
+  mutate(is_female=if_else(sex=='female', 1L, 0L)) %>%
+  select(-sex) %>%
   left_join(chronic_conditions, by='bene_id')
+```
+
+```{r compute_correlations}
+tetra = chronic_conditions %>%
+  select(-bene_id) %>%
+  mutate_all(function(x) x == 3) %>%
+  tetrachoric()
 ```
@@
   x %>%
-    glm(n_claims ~ has_condition + age, family='poisson', data=.) %>%
+    glm(n_claims ~ has_condition, family='poisson', data=.) %>%
     tidy %>%
@@
 
+```{r multivariate_model}
+multivariate_formula = paste(c('n_claims ~ age + is_female', conditions), collapse=' + ') %>%
+  formula
+
+multivariate_res = usage_summary %>%
+  lm(multivariate_formula, family='poisson', data=.) %>%
+  tidy
+```
+
+## Correlations
+
+Presumably many of the conditions are correlated. I computed tetrachoric correlations, which is a method designed to deal with correlations of dichotomous variables.
+
+```{r, cache=FALSE}
+tetra$rho %>% heatmap
+```
+
+Thankfully, Alzheimer's correlates strongly with Alzheimer's and dementia, and prostatic hyperplasia and prostate cancer correlate strongly negatively with breast cancer and endometrial cancer.
+Unfortunately, the negative correlations make everything else harder to see, so here's the same thing but without the male conditions.
+
+```{r fig.width=7, fig.height=7, cache=FALSE}
+(!(rownames(tetra$rho) %in% c('CNCRPRST', 'HYPERP'))) %>%
+  { tetra$rho[., .] } %>%
+  heatmap.2(trace='none')
+```
+
+COPD correlates strongly with asthma and lung cancer, but lung cancer and asthma do not correlate strongly with one another.
+There is a block of correlations focused around hypertension that includes ischemic heart disease, heart failure, anemia, chronic kidney disease, diabetes, hypertension, hyperlipidemia, and, importantly, COPD.
+
+## Regression
+
+### Univariate
+
+First, I performed univariate regressions. In each case, the explanatory variable was the presence of the chronic condition.
+
 Almost all the conditions are significantly associated with greater consumption.
@@
 
-```{r show_poisson}
+```{r show_poisson, cache=FALSE}
 res %>%
@@
 ```
+
+### Multivariate
+
+I also performed a multivariate Poisson regression, keeping all the chronic conditions as well as age and sex (coded as "is female") as explanatory variables.
+
+Not surprisingly, COPD tops the list.
+However, asthma does provide explanatory power beyond COPD, coming in second.
+(Lung cancer, however, now has a negative effect on consumption.) Strangely, depression and anemia are next.
+Age, kindney disease, and diabetes have coefficients very near zero!
+
+```{r cache=FALSE}
+multivariate_res %>%
+  mutate(pct.increase=round((exp(estimate)-1.0)*100)) %>%
+  arrange(desc(pct.increase)) %>%
+  kable
+```

### 2017-03-06 inequality-resistance/inequality-resistance.Rmd
beginning ineq res

---
title: ""
author: "Scott Olesen"
date: ""
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, 
                      echo=TRUE, warning=FALSE, message=FALSE, cache=TRUE)
```

First, load the relevant data. I need the usage and antibiograms.

```{r load_groups}
resistance_groups = read_tsv('../states/resistance/resistance_groups.tsv')
consumption_groups = read_tsv('../states/resistance/consumption_groups.tsv')
```

```{r load_antibiogram}
organisms = c('Escherichia coli')

hrr = read_tsv('../../db/hrr/hrr_2011.tsv') %>%
  select(zipcode, hrr)

abg = read_tsv('../../../antibiogram/data/abg.tsv', col_types=cols(zipcode='c')) %>%
  mutate(drug=tolower(drug), percent_nonsusceptible=100-percent_susceptible) %>%
  left_join(hrr, by='zipcode') %>%
  right_join(resistance_groups, by='drug') %>%
  filter(bug %in% organisms) %>%
  select(bug, drug_group, state, hrr, percent_nonsusceptible, n_isolates)

# Keep only those bug/drug combinations that are in at least 20 states
abg %<>%
  select(bug, drug_group, state) %>%
  distinct %>%
  count(bug, drug_group) %>% rename(n_states=n) %>%
  filter(n_states >= 20) %>%
  left_join(abg, by=c('bug', 'drug_group'))
```

```{r load_consumption}
usage = read_tsv('../bene_usage_2011.tsv') %>%
  rename(drug=antibiotic)

denom = usage %>% select(bene_id, state) %>% distinct

consumption = usage %>%
  left_join(consumption_groups, by='drug') %>%
  group_by(bene_id, drug_group) %>%
  summarize(n_claims=sum(n_claims)) %>%
  ungroup

consumption_of = function(drug_group0) {
  consumption %>%
    filter(drug_group==drug_group0) %>%
    select(bene_id, n_claims) %>%
    right_join(denom, by='bene_id') %>%
    mutate(n_claims=if_else(is.na(n_claims), 0L, n_claims))
}
```

```{r inequality_functions}
library(ineq)
is_nonzero = function(x) x > 0
nonzero = function(x) x[is_nonzero(x)]
fraction_nonzero = function(x) sum(is_nonzero(x)) / length(x)
gini = function(x) ineq(x, type='Gini')

normalize = function(x) x / max(x)
juran = function(x) {
  # better to do with percent_rank and cume_dist
  fraction_consumed = cumsum(sort(x, decreasing=TRUE)) %>% normalize
  fraction_users = 1:length(x) %>% normalize
  fraction_nonzero(fraction_consumed + fraction_users - 1)
}
```

### 2017-03-08 inequality-resistance/inequality-resistance.Rmd
many changes

@@
 consumption_groups = read_tsv('../states/resistance/consumption_groups.tsv')
+drug_groups = resistance_groups$drug_group
 ```
@@
 ```{r load_antibiogram}
-organisms = c('Escherichia coli')
+organisms = c('Escherichia coli',
+              'Proteus mirabilis', 'Enterobacter cloacae',
+              'Enterococcus faecium',
+              'Coagulase-negative staphylococci',
+              'Methicillin-resistant S. aureus (MRSA)',
+              'Methicillin-susceptible S. aureus (MSSA)',
+              'Streptococcus pneumoniae (non-meningitis)',
+              'Staphylococcus aureus')
 
@@
   summarize(n_claims=sum(n_claims)) %>%
-  ungroup
+  ungroup %>%
+  complete(bene_id, drug_group, fill=list('n_claims'=0L)) %>%
+  left_join(denom, by='bene_id')
 
@@
 }
+
+klasma = function(x, threshold) {
+  if (!between(threshold, 0.0, 1.0)) stop("bad threshold")
+  data_frame(x=sort(x, decreasing=TRUE)) %>%
+    mutate(csx=cumsum(x), cd=cume_dist(csx)) %$%
+    sum(x[cd <= 0.1]) / sum(x)
+}
 ```
+
+```{r}
+mean_abg = abg %>%
+  group_by(bug, drug_group, state) %>%
+  summarize(mpr=mean(percent_nonsusceptible),
+            wmpr=weighted.mean(percent_nonsusceptible, sqrt(n_isolates)))
+
+state_consumption = consumption %>%
+  group_by(state, drug_group) %>%
+  summarize(f_nz=fraction_nonzero(n_claims),
+            pct_nz=100*f_nz,
+            juran0=juran(n_claims),
+            juran_nz=juran(nonzero(n_claims)),
+            gini0=gini(n_claims),
+            gini_nz=gini(nonzero(n_claims)),
+            k10_nz=klasma(nonzero(n_claims), 0.1),
+            n_bene=n(),
+            n_claims=sum(n_claims),
+            cpkp=n_claims/n_bene*1000,
+            cpku=cpkp/f_nz)
+
+```
+
+```{r do_models}
+models = function(x, independent) {
+  univariate_formula = paste(independent, '~ cpkp') %>% formula
+  univariate_results = lm(univariate_formula, data=x) %>%
+    anova %>% tidy %>% mutate(model='univariate')
+  multivariate_formula = paste(independent, '~ pct_nz + cpku') %>% formula
+  multivariate_results = lm(multivariate_formula, data=x) %>%
+    anova %>% tidy %>% mutate(model='multivariate')
+  bind_rows(univariate_results, multivariate_results)
+}
+
+results = left_join(mean_abg, state_consumption, by=c('state', 'drug_group')) %>%
+  group_by(bug, drug_group) %>%
+  do(models(., 'mpr')) %>%
+  select(bug, drug_group, model, term, estimate, p.value)
+
+results2 = left_join(abg, state_consumption, by=c('state', 'drug_group')) %>%
+  group_by(bug, drug_group) %>%
+  do(models(., 'percent_nonsusceptible')) %>%
+  select(bug, drug_group, model, term, estimate, p.value)
+```
+
+# Results
+
+## Correlations between metrics
+
+In general,
+
+- CPKP correlates strongly with `f_nz` (makes sense, since most users take ~1 rx)
+- Intensive metrics correlate with one another (`gini_nz` and `juran_nz`)
+- Extensive metrics correlate (e.g., `juran` with `f_nz`)
+- Intensive and extensive metrics correlation (`juran_nz` with `f_nz`)
+
+Thus, in general, increased consumption goes with increased inequality.
+
+## Klasma 10
+
+Klasma 10 is the fraction of the drugs consumed by the top 10% of users.
+Consider bug/drug combos where `k10_nz` in the model `percent_susceptible` (or mean r) `~ pct_nz + k10_nz` is significant.
+Those are all either tmp-smx or tetracycline, and in each case, the estimate for `cpkp` in the univariate model matches the sign of `pct_nz` in the multivariate model, and `k10_nz` in the multivariate model is always negative.
+
+## Using CPKU
+
+I tried to use CPKU as an intensive measure: what's the average consump (CPK) among users (U)? I got similar results to when I was using other intensive metrics.
+
+## f_nz vs. CPKP
+
+In general, the fraction nonzero is a better predictor of resistance than CPKP, i.e., the extensive metric is better than the garbled intensive metric.

### 2017-03-08 inequality/consumption-inequality.Rmd
many changes

@@
 
+```{r inequality_by_region2}
+# This uses code from inequality-resistance
+normalize = function(x) x / max(x)
+cumrelfreq = function(df, var) {
+  n = sort(df[[var]], decreasing=FALSE)
+  data_frame(x=cume_dist(n), y=normalize(cumsum(n))) %>%
+    group_by(x) %>%
+    summarize(y=max(y))
+}
+
+m = consumption_of('macrolide')
+
+x = m %>%
+  left_join(census_regions, by='state') %>%
+  group_by(region) %>%
+  do(cumrelfreq(., 'n_claims'))
+
+p = x %>%
+  ggplot(aes(x=x, y=y, color=region, group=region)) +
+  geom_line() + geom_point() +
+  xlab('cumulative fraction of users') +
+  ylab('cumulative fraction of drugs consumed') +
+  theme_minimal()
+
+show(p)
+```
+
 Among the most popular drugs, unevenness varies only slightly.

### 2017-03-08 summarize-abx.R

@@ -1,16 +1,0 @@
-#!/usr/bin/env Rscript
-
-# Summarize the information about abx usage for each beneficiary:
-# - Use the abx PDE information
-# - Give just # of events, total days, and total cost
-
-summarize_abx_pde = function(in_fn, out_fn) {
-  read_tsv(in_fn) %>%
-    rename(bene=BENE_ID) %>%
-    group_by(bene) %>%
-    summarize(total_abx_pde=n(), total_days=sum(DAYSSPLY), total_cost=sum(TOTALCST)) %>%
-    select(bene, total_abx_pde, total_days, total_cost) %>%
-    write_tsv(out_fn)
-}
-
-summarize_abx_pde('../data/abx_pde_2011.tsv', 'summary-abx-pde-2011.tsv')

### 2017-03-08 summarize-bene-locations.R

@@ -1,33 +1,0 @@
-#!/usr/bin/env Rscript
-
-# create summary beneficiary x abx file
-
-hrr = read_tsv("../db/hrr/ZipHsaHrr11.txt") %>%
-  select(zip=zipcode11, hrr=hrrnum)
-census = read_tsv("../db/census-regions/census-regions.tsv") %>%
-    select(state, region)
-
-report = function(dat, group, out_fn) {
-  dat %>%
-    group_by_(group) %>%
-    summarize(n=n()) %>%
-    write_tsv(out_fn)
-}
-
-report_for_year = function(year) {
-  abx_fn = sprintf('bene-abx-%s.tsv', year)
-
-  abx = read_tsv(abx_fn) %>%
-    select(bene, zip, state) %>%
-    left_join(hrr, by='zip') %>%
-    left_join(census, by='state')
-
-  for (g in c('region', 'state', 'hrr')) {
-    out_fn = sprintf('summary-bene-location-%s-%s.tsv', g, year)
-    report(abx, g, out_fn)
-  }
-}
-
-for (y in c('2011', '2012', '2013', '2014')) {
-  report_for_year(y)
-}

### 2017-03-13 abx-abx-correlations/drug-drug.Rmd
tetrachoric

---
title: "Drug-drug correlations"
author: "Scott Olesen"
date: "8 March 2017"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, 
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
```

# Background

Does taking one drug make you more likely to take another one? Or less likely?

# Methods

I use the 2011 data and tetrachoric correlations. I look at only the top 10 drugs (and I group amox and amox/clav under amox).

## Tetrachoric correlation

Consider two dichotomous variables $X$ and $Y$. Each joint observation ($x_i$, $y_i$) can be one of $(0, 0)$, $(0, 1)$, $(1, 0)$, or $(1, 1)$. The tetrachoric correlation finds a correlation $\rho$ by finding the "cutpoints" $tau_X$ and $tau_Y$ such that the bivariate standard normal distribution with correlation $\rho$ has "quadrants" with masses that best match the relative frequencies of the joint observations. For example, the "lower-left" quadrant has mass

$$
\int_{-\infty}^{\tau_Y} \int_{-\infty}^{\tau_X} \mathcal{N}(x, y; \rho) \,\mathrm{d}x \,\mathrm{d}y
$$

and should match the relative frequency of observations $(0, 0)$. The $\rho$ of the best-fitting bivariate normal distribution is the $\rho$ for those two variables. For more than two variables, a multivariate normal distribution is used.

This method can be extend to a *polychoric* method where the variables are ordinal (rather than dichotomous) by adding multiple cutpoints for the various categories.

```{r load_data}
usage = read_tsv('../bene_usage_2011.tsv') %>%
  select(bene_id, drug=antibiotic)

denom = usage %>% select(bene_id) %>% distinct
```

# Results

```{r set_drug_levels}
drugs = c(
  'ciprofloxacin',
  'trimethoprim/sulfamethoxazole',
  'nitrofurantoin',
  'cephalexin',
  'metronidazole',
  'levofloxacin',
  'azithromycin',
  'clindamycin',
  'amoxicillin', 'amoxicillin/clavulanate',
  'doxycycline'
)

drugs_short = c(
  'cpro', 'tpsx', 'ntro', 'cphx', 'mtro', 'levo', 'azth', 'clnd', 'amox',
  'amox/clav', 'doxy')
```

```{r do_correlation}
library(psych)
# convert all amox/clav entries into amox
res = usage %>%
  mutate(drug=if_else(
    drug %in% c('amoxicillin', 'amoxicillin/clavulanate'),
    'amoxicillin', drug)) %>%
  filter(drug %in% drugs) %>%
  distinct %>% # only care if had at least one claim
  right_join(denom, by='bene_id') %>% # keep benes with no claims
  mutate(has=TRUE) %>% # fill with TRUEs
  spread(drug, has, fill=FALSE) %>% # but with FALSE is they don't have
  select(-`<NA>`) %>% # drop the no-having column
  select(-bene_id) %>%
  tetrachoric
```

First, a table of the results.

```{r show_table}
res$rho %>%
  as.data.frame() %>%
  mutate(drug1=rownames(.)) %>%
  mutate_at(vars(-drug1), function(x) round(x, 2)) %>%
  select(drug1,
         amox=amoxicillin,
         azth=azithromycin,
         cphx=cephalexin,
         cipr=ciprofloxacin,
         clnd=clindamycin,
         doxy=doxycycline,
         levo=levofloxacin,
         mtro=metronidazole,
         ntro=nitrofurantoin,
         tpsx=`trimethoprim/sulfamethoxazole`) %>%
  kable
```

Then, a heatmap showing the same things. I picked the order of the rows/columns by hand.

```{r show_heatmap}
drug_factor = function(x) factor(x, levels=drugs, labels=drugs_short, ordered=TRUE)

res$rho %>%
  as.data.frame %>%
  mutate(drug1=rownames(.)) %>%
  mutate_at(vars(-drug1), function(x) round(x, 2)) %>%
  gather('drug2', 'corr', -drug1) %>%
  filter(drug1 != drug2) %>%
  mutate_at(vars(drug1, drug2), drug_factor) %>%
  ggplot(aes(y=drug1, x=drug2, fill=corr)) +
  geom_tile() +
  scale_fill_distiller(palette='RdGy') +
  theme_classic()
```

# Discussion

Notably, all the correlations are positive: taking one drug means you are more likely to take another.

The strongest correlations are between ciprofloxacin and each of metronidazole, nitrofurantoin, and TMP/SMX. TMP/SMX and nitrofurantoin also correlate.

Levofloxacin has a weaker correlation with metronidazole than ciprofloxacin does, but azithromycin's strongest correlation is with azithromycin.

### 2017-03-13 sex-difference/sex-difference.Rmd
sex difference

---
title: "Drivers of sex differences in consumption"
author: "Scott Olesen"
date: "2/24/2017"
output: html_document
---

# Background

Why do females account for 25% more antibiotic claims per capita than males? I have two easy explanations:

- More females are alive at later ages, and older people consume more antibiotics.
- Females consume more of the same antibiotics.
- Females consume different antibiotics.

# Methods

I'll use the Medicare 2011 data.

# Results

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/',
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
```

```{r load_data, cache=TRUE}
usage = read_tsv('../bene_usage_2011.tsv')
```

## Consumption by age and sex

It seems that females consume more antibiotics even within each age bracket (except for those over 95).

```{r compute_consumption_by_age_and_sex}
bounds_group = function(x, upper_limits) {
  upper_limits = as.integer(upper_limits)
  levels = lapply(seq(from=1, to=length(upper_limits)), function(i) {
    if (i == 1) {
      sprintf('<= %i', upper_limits[i])
    } else {
      sprintf('> %i and <= %i', upper_limits[i - 1], upper_limits[i])
    }
  })
  top_level = sprintf('> %i', tail(upper_limits, 1))
  formula_list = mapply(function(str, lim) as.formula(sprintf('x <= %i ~ "%s"', lim, str)),
                        levels, upper_limits) %>%
    c(as.formula(sprintf('TRUE ~ "%s"', top_level)))

  levels = c(levels, top_level)

  do.call(case_when, formula_list) %>%
    factor(levels=levels, ordered=TRUE)
}

usage %>%
  mutate(age_group=bounds_group(age, c(65, 75, 85, 95))) %>%
  group_by(age_group, sex, bene_id) %>%
  summarize_at(vars(n_claims, n_days), sum) %>%
  summarize(mean_n_claims=mean(n_claims),
            mean_n_days=mean(n_days),
            n_bene=n()) %>%
  kable
```

## Consumption of drugs by sex

Looking at top drugs by sex.

```{r drugs_by_sex, cache=TRUE}
bene_by_sex = usage %>%
  #sample_frac(0.01) %>%
  select(bene_id, sex) %>%
  distinct %>%
  count(sex) %>%
  rename(n_bene=n)

dbs = usage %>%
  #sample_frac(0.01) %>%
  filter(age >= 65, sex %in% c('male', 'female')) %>%
  group_by(antibiotic, sex, bene_id) %>%
  summarize_at(vars(n_claims, n_days), sum) %>%
  summarize(n_claims=sum(n_claims), n_days=sum(n_days)) %>%
  left_join(bene_by_sex, by='sex') %>%
  mutate(cpkp=n_claims*1000/n_bene, did=n_days*1000/(365*n_bene))

dbs %>%
  filter(sex=='female') %>%
  arrange(desc(cpkp)) %>%
  head(10) %>%
  kable

dbs %>%
  filter(sex=='male') %>% 
  arrange(desc(cpkp)) %>%
  head(10) %>%
  kable
```

```{r drug_by_sex_plot}
dbs %>%
  select(antibiotic, sex, cpkp) %>%
  spread(sex, cpkp) %>%
  ggplot(aes(x=male, y=female, label=antibiotic)) +
  geom_label() +
  geom_point() +
  geom_abline() +
  xlab('claims per 1,000 male beneficiaries') +
  ylab('claims per 1,000 female beneficiaries') +
  theme_minimal()
```

For a more quantitative approach, this is a table showing, for each antibiotic, the CPKP for female and males, the difference (residual) between the two, and the residual fraction (the residual over the mean of the female and male consumptions). Positive residuals means more consumption by females.

```{r drug_by_sex_residuals}
dbs %>%
  select(antibiotic, sex, cpkp) %>%
  spread(sex, cpkp) %>%
  mutate(residual=female-male, residual_fraction=residual/(0.5 * (male + female))) %>%
  arrange(desc(abs(residual))) %>%
  head(10) %>%
  kable
```

## Azithromycin

Nitrofurantoin, ciprofloxacin, and TMP-SMX are all used as urinary anti-infectives, and UTIs more commonly affect women. This does not explain, however, the increased consumption of azithromycin.

```{r women_azithro_vs_uai}
any_in = function(lst, elts) match(elts, lst) %>% { .[!is.na(.)] } %>% { length(.) > 0 }

usage %>%
  filter(age >= 65, sex=='female') %>%
  group_by(bene_id) %>%
  mutate(used_azithro='azithromycin' %in% antibiotic,
         used_uai=any_in(antibiotic, c('nitrofurantoin', 'trimethoprim/sulfamethoxazole', 'ciprofloxacin')),
         used_both=used_azithro && used_uai,
         is_white=race=='white') %>%
  ungroup() %>%
  mutate(user_type=case_when(.$used_both ~ 'both',
                             .$used_azithro ~ 'azithro',
                             .$used_uai ~ 'UAI',
                             TRUE ~ 'none')) %>%
  group_by(bene_id) %>%
  summarize(age=age[1],
            is_white=is_white[1],
            chronic=chronic[1],
            user_type=user_type[1],
            n_claims=sum(n_claims)) %>%
  group_by(user_type) %>%
  summarize(n_beneficiaries=n(),
            mean_age=mean(age),
            percent_white=mean(is_white)*100,
            mean_n_chronic_cond=mean(chronic),
            mean_n_claims=mean(n_claims)) %>%
  arrange(desc(n_beneficiaries)) %>%
  kable
```

### 2017-03-13 inequality/consumption-inequality.Rmd
new inequality plot
updated inequality

--- 2017-03-02 inequality/consumption-inequality.Rmd
@@
 
+```{r inequality_by_region2}
+# This uses code from inequality-resistance
+normalize = function(x) x / max(x)
+cumrelfreq = function(df, var) {
+  n = sort(df[[var]], decreasing=FALSE)
+  data_frame(x=cume_dist(n), y=normalize(cumsum(n))) %>%
+    group_by(x) %>%
+    summarize(y=max(y))
+}
+
+#m = consumption_of('macrolide')
+
+#x = m %>%
+#  left_join(census_regions, by='state') %>%
+#  group_by(region) %>%
+#  do(cumrelfreq(., 'n_claims'))
+
+#p = x %>%
+#  ggplot(aes(x=x, y=y, color=region, group=region)) +
+#  geom_line() + geom_point() +
+#  xlab('cumulative fraction of users') +
+#  ylab('cumulative fraction of drugs consumed') +
+#  theme_minimal()
+
+#show(p)
+```
+
 Among the most popular drugs, unevenness varies only slightly.
@@
 ```{r chars_by_consumption}
+# load the encounters claim data
+carrier_count = read_tsv('../../data/carrier_count_2011.tsv') %>%
+  rename(bene_id=BENE_ID, n_carrier_dates=n_from_dates)
+inpatient_count = read_tsv('../../data/inpatient_count_2011.tsv') %>%
+  rename(bene_id=BENE_ID, n_inpatient_dates=n_from_dates)
+outpatient_count = read_tsv('../../data/outpatient_count_2011.tsv') %>%
+  rename(bene_id=BENE_ID, n_outpatient_dates=n_from_dates)
+
 ntile_zero = function(x, n) {
@@
 
+na_to_zero = function(x) if_else(is.na(x), 0L, x)
+
 usage %>%
-  group_by(age, sex, race, chronic, bene_id) %>%
+  group_by(bene_id) %>%
   summarize(n_claims=sum(n_claims),
             n_days=sum(n_days)) %>%
-  ungroup() %>%
+  left_join(select(usage, bene_id, sex, age, race, chronic), by='bene_id') %>%
+  left_join(carrier_count, by='bene_id') %>%
+  left_join(outpatient_count, by='bene_id') %>%
+  left_join(inpatient_count, by='bene_id') %>%
+  mutate_at(vars(n_carrier_dates, n_outpatient_dates, n_inpatient_dates), na_to_zero) %>%
   mutate(claims_group=bounds_group(n_claims, c(0, 1, 2, 5, 10, 20, 50, 100)),
@@
             mean_n_chronic_cond=mean(chronic),
-            mean_n_claims=mean(n_claims)) %>%
+            mean_n_claims=mean(n_claims),
+            mean_n_carrier_dates=mean(n_carrier_dates),
+            mean_n_outpatient_dates=mean(n_outpatient_dates),
+            mean_n_inpatient_dates=mean(n_inpatient_dates)) %>%
   mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%

### 2017-03-13 sex-difference/sex-difference.Rmd
sex differences

@@
 usage %>%
+  filter(sex %in% c('male', 'female')) %>%
   mutate(age_group=bounds_group(age, c(65, 75, 85, 95))) %>%
@@
 ```
+
+Does that mean that:
+
+- women are older and therefore use more azithromycin,
+- women have more chronic conditions (esp. COPD) and therefore use more, or
+- women have more encounters with doctors and therefore are prescribed more?
+
+### Sex by age
+
+The older people are more female: age 65 is about 60% female.
+This stays constant around 60% (or dips slightly) until age 75, when it steadily increases up to 85-90% female at age 100.
+So we'll need to account for this age difference.
+
+```{r}
+plot_dat = usage %>%
+  select(age, sex) %>%
+  filter(sex %in% c('male', 'female')) %>%
+  group_by(age) %>% count(sex) %>% ungroup() %>%
+  spread(sex, n, fill=0L) %>%
+  mutate(total=female+male, grand_total=sum(total),
+         fraction_year=total/grand_total, fraction_female=female/total)
+
+ggplot(mapping=aes(x=age, y=fraction)) +
+  geom_point(data=rename(plot_dat, fraction=fraction_female)) +
+  geom_line(data=mutate(plot_dat, fraction=fraction_year*10), col='red') +
+  ylab('fraction female') + ylim(0, 1) +
+  theme_minimal()
+```
+
+### Chronic conditions
+
+Cf. the chronic conditions report.
+
+### Encounters by sex
+
+Do women have more encounters than men? Does that explain their increased usage?
+
+```{r load_carrier_data}
+carrier = read_tsv('../../data/carrier_count_2011.tsv') %>%
+  rename(bene_id=BENE_ID, n_carrier_dates=n_from_dates)
+
+inpatient = read_tsv('../../data/inpatient_count_2011.tsv') %>%
+  rename(bene_id=BENE_ID, n_inpatient_dates=n_from_dates)
+
+outpatient = read_tsv('../../data/outpatient_count_2011.tsv') %>%
+  rename(bene_id=BENE_ID, n_outpatient_dates=n_from_dates)
+
+denom = usage %>% select(bene_id, sex, age) %>% distinct
+```
+
+```{r azithromycin_usage}
+na_to_zero = function(x) if_else(is.na(x), 0L, x)
+
+azi_usage = usage %>%
+  filter(antibiotic=='azithromycin') %>%
+  group_by(bene_id) %>%
+  summarize(n_claims=sum(n_claims)) %>%
+  left_join(denom, by='bene_id') %>%
+  filter(age >= 65, sex %in% c('male', 'female')) %>%
+  left_join(carrier, by='bene_id') %>%
+  left_join(inpatient, by='bene_id') %>%
+  left_join(outpatient, by='bene_id') %>%
+  mutate(age_group=bounds_group(age, c(65, 75, 85, 95))) %>%
+  mutate_at(vars(n_claims, n_carrier_dates, n_inpatient_dates, n_outpatient_dates), na_to_zero)
+```
+
+```{r azithromycin_usage_table}
+azi_usage %>%
+  group_by(age_group, sex) %>%
+  summarize(n_beneficiaries=n(),
+            mean_n_azithromycin_claims=mean(n_claims),
+            med_n_carrier_dates=median(n_carrier_dates),
+            med_n_outpatient_dates=median(n_outpatient_dates),
+            med_n_inpatient_dates=median(n_inpatient_dates)) %>%
+  kable
+```
+
+And I'll do a Poisson regression of number of azithromycin claims against age, being female, and number of carrier dates:
+
+```{r azithromycin_usage_model}
+library(broom)
+azi_usage %>%
+  mutate(is_female=sex=='female') %>%
+  glm(n_claims ~ age + is_female + n_carrier_dates + n_inpatient_dates + n_outpatient_dates, data=., family='poisson') %>%
+  tidy %>%
+  mutate(pct.increase=round((exp(estimate)-1.0)*100, 2)) %>%
+  kable
+```
+
+The coefficient for being female (0.05) accounts for almost the entire difference in consumption between men and women (56 per thousand). Note that age has an approximately zero correlation (i.e.,

### 2017-03-14 chronic-conditions/chronic-condition-drivers.Rmd
chronic: lasso

@@
   rename(bene_id=BENE_ID) %>%
-  select(-STRCTFLG)
+  select(-STRCTFLG) %>%
+  mutate_at(vars(AMI:HYPOTH), function(x) x==3)
 condition_names = read_tsv('names.txt', col_names=c('condition', 'condition_name'))
@@
 ```{r summarize_users}
+denom = usage %>% select(bene_id, age, sex) %>% distinct
+
 usage_summary = usage %>%
@@
   group_by(bene_id) %>%
-  summarize(n_claims=sum(n_claims), age=max(age), sex=head(sex, 1)) %>%
-  mutate(is_female=if_else(sex=='female', 1L, 0L)) %>%
-  select(-sex) %>%
+  summarize(n_claims=sum(n_claims)) %>%
+  left_join(denom, by='bene_id') %>%
+  mutate(is_female=sex=='female') %>% select(-sex) %>%
   left_join(chronic_conditions, by='bene_id')
+```
+
+```{r}
+chronic_conditions %>% head(100) %>% write_tsv('tmp')
 ```
@@
   select(-bene_id) %>%
-  mutate_all(function(x) x == 3) %>%
   tetrachoric()
@@
 
-```{r compute_models}
+```{r univariate_models}
 conditions = c("AMI", "ALZH", "ALZHDMTA", "ATRIALFB", "CATARACT", "CHRNKIDN", "COPD", "CHF", "DIABETES", "GLAUCOMA", "HIPFRAC", "ISCHMCHT", "DEPRESSN", "OSTEOPRS", "RA_OA", "STRKETIA", "CNCRBRST", "CNCRCLRC", "CNCRPRST", "CNCRLUNG", "CNCRENDM", "ANEMIA", "ASTHMA", "HYPERL", "HYPERP", "HYPERT", "HYPOTH")
@@
 summarize_condition = function(x, condition) {
-  has_condition = x[[condition]] == 3
+  univariate_formula = paste0('n_claims ~ age + is_female + ', condition) %>%
+    formula
 
   x %>%
-    glm(n_claims ~ has_condition, family='poisson', data=.) %>%
+    glm(univariate_formula, family='poisson', data=.) %>%
     tidy %>%
@@
 
-```{r multivariate_model}
+```{r lasso}
+library(glmnet)
+lasso_x = usage_summary %>%
+  select(age, is_female, AMI:HYPOTH) %>%
+  mutate_at(vars(AMI:HYPOTH), function(x) if_else(x==3, 1L, 0L)) %>%
+  as.matrix
+lasso_y = usage_summary$n_claims
+lasso_model = glmnet(lasso_x, lasso_y, family='poisson')
+```
+
+```{r multivariate_linear_model}
 multivariate_formula = paste(c('n_claims ~ age + is_female', conditions), collapse=' + ') %>%
@@
 
-multivariate_res = usage_summary %>%
-  lm(multivariate_formula, family='poisson', data=.) %>%
+multivariate_linear_res = usage_summary %>%
+  lm(multivariate_formula, data=.) %>%
   tidy
 ```
+
+```{r multivariate_poisson_model}
+multivariate_poisson_res = usage_summary %>%
+  glm(multivariate_formula, family='poisson', data=.) %>%
+  tidy
+```
+
+## Summary
+
+`r usage_summary %>% kable`
 
@@
 
-### Multivariate
+### Multivariate models
 
-I also performed a multivariate Poisson regression, keeping all the chronic conditions as well as age and sex (coded as "is female") as explanatory variables.
+I also performed multivariate regressions, keeping all the chronic conditions as well as age and sex (coded as "is female") as explanatory variables.
+
+#### Multivariate linear
+
+I don't expect a linear model to work well, since the output is count data.
 
@@
 ```{r cache=FALSE}
-multivariate_res %>%
+multivariate_linear_res %>%
   mutate(pct.increase=round((exp(estimate)-1.0)*100)) %>%
@@
 ```
+
+#### Multivariate Poisson
+
+This should be more accurate.
+
+```{r cache=FALSE}
+multivariate_poisson_res %>%
+  mutate(pct.increase=round((exp(estimate)-1.0)*100)) %>%
+  arrange(desc(pct.increase)) %>%
+  kable
+```
+
+#### Lasso
+
+I also performed a lasso regression with the Poisson model. I'm not sure how to display the results sensibly. I picked a value of $s$ that limited me to just a few nonzero model coefficients.
+
+```{r cache=FALSE}
+lasso_model %>%
+  coef(s=0.25) %>%
+  tidy %>%
+  select(term=row, coefficient=value) %>%
+  kable
+```

### 2017-03-14 chronic-conditions/chronic-condition-drivers.Rmd
more lasso

@@
 
-```{r}
-chronic_conditions %>% head(100) %>% write_tsv('tmp')
-```
-
 ```{r compute_correlations}
@@
 library(glmnet)
+require(doMC)
+registerDoMC(cores=3)
+
 lasso_x = usage_summary %>%
   select(age, is_female, AMI:HYPOTH) %>%
-  mutate_at(vars(AMI:HYPOTH), function(x) if_else(x==3, 1L, 0L)) %>%
   as.matrix
 lasso_y = usage_summary$n_claims
-lasso_model = glmnet(lasso_x, lasso_y, family='poisson')
+
+lasso_cv = cv.glmnet(lasso_x, lasso_y, parallel=TRUE, family='poisson')
+#lasso_model = glmnet(lasso_x, lasso_y, family='poisson')
 ```
@@
   tidy
-```
 
-```{r multivariate_poisson_model}
-multivariate_poisson_res = usage_summary %>%
-  glm(multivariate_formula, family='poisson', data=.) %>%
-  tidy
+# For multivariate Poisson:
+# glm(multivariate_formula, family='poisson', data=.)
 ```
@@
 
-### Multivariate models
+### Multivariate linear model
 
@@
 
-#### Multivariate linear
-
-I don't expect a linear model to work well, since the output is count data.
+I don't expect a linear model to work well, since the output is count data, but doing a Poisson regression took a really long time.
 
@@
 
-#### Multivariate Poisson
+### Lasso
 
-This should be more accurate.
+I also performed a lasso regression with the Poisson model. I started with 10-fold cross-validation to pick the $\lambda$.
 
 ```{r cache=FALSE}
-multivariate_poisson_res %>%
-  mutate(pct.increase=round((exp(estimate)-1.0)*100)) %>%
-  arrange(desc(pct.increase)) %>%
+plot(lasso_cv)
+```
+
+Although the error continues to fall with decreasing $\lambda$, it seems like there are diminishing returns after the first 8 or 10 variables are included.
+
+```{r cache=FALSE}
+lasso_df = lasso_cv$glmnet.fit %>%
+  tidy %>%
+  filter(estimate > 0) %>%
+  group_by(step) %>%
+  summarize(df=n() - 1)
+
+lasso_cv$glmnet.fit %>%
+  tidy %>%
+  select(step, lambda, dev.ratio) %>%
+  distinct() %>%
+  left_join(lasso_df, by='step') %>%
+  select(step, df, lambda) %>%
+  mutate_at(vars(lambda), function(x) signif(x, digits=2)) %>%
   kable
@@
 
-#### Lasso
-
-I also performed a lasso regression with the Poisson model. I'm not sure how to display the results sensibly. I picked a value of $s$ that limited me to just a few nonzero model coefficients.
+It looks like this corresponds to $\lambda \sim 0.1$. For that value, the coefficients are:
 
 ```{r cache=FALSE}
-lasso_model %>%
-  coef(s=0.25) %>%
+lasso_cv %>%
+  coef(s=0.1) %>%
   tidy %>%

### 2017-03-14 chronic-conditions/chronic-condition-drivers.Rmd
only lasso

@@
 
-```{r lasso}
+```{r lasso_poisson}
 library(glmnet)
@@
 lasso_cv = cv.glmnet(lasso_x, lasso_y, parallel=TRUE, family='poisson')
-#lasso_model = glmnet(lasso_x, lasso_y, family='poisson')
-```
-
-```{r multivariate_linear_model}
-multivariate_formula = paste(c('n_claims ~ age + is_female', conditions), collapse=' + ') %>%
-  formula
-
-multivariate_linear_res = usage_summary %>%
-  lm(multivariate_formula, data=.) %>%
-  tidy
-
-# For multivariate Poisson:
-# glm(multivariate_formula, family='poisson', data=.)
 ```
@@
 
-### Multivariate linear model
+### Multivariate lasso
 
-I also performed multivariate regressions, keeping all the chronic conditions as well as age and sex (coded as "is female") as explanatory variables.
-
-I don't expect a linear model to work well, since the output is count data, but doing a Poisson regression took a really long time.
-
-Not surprisingly, COPD tops the list.
-However, asthma does provide explanatory power beyond COPD, coming in second.
-(Lung cancer, however, now has a negative effect on consumption.) Strangely, depression and anemia are next.
-Age, kindney disease, and diabetes have coefficients very near zero!
-
-```{r cache=FALSE}
-multivariate_linear_res %>%
-  mutate(pct.increase=round((exp(estimate)-1.0)*100)) %>%
-  arrange(desc(pct.increase)) %>%
-  kable
-```
-
-### Lasso
-
-I also performed a lasso regression with the Poisson model. I started with 10-fold cross-validation to pick the $\lambda$.
+I also performed a multivariate lasso regression with the Poisson model, keeping all the chronic conditions as well as age and sex (coded as "is female") as explanatory variables.
+I started with 10-fold cross-validation to pick the lasso $\lambda$.
 
@@
 ```
+
+Not surprisingly, COPD tops the list.
+However, asthma does provide explanatory power beyond COPD, coming in second.
+(Lung cancer, however, now has a negative effect on consumption.) Strangely, depression and anemia are next.
+Age, kindney disease, and diabetes have coefficients very near zero!

### 2017-03-14 chronic-conditions/chronic-condition-drivers.Rmd
chronic conditions

@@
 
-```{r global_options, include=FALSE}
+```{r global_options, include=FALSE, cache=FALSE}
 knitr::opts_chunk$set(fig.width=7, fig.height=4, 
@@
 library(gplots)
+library(broom)
+library(glmnet)
+require(doMC)
+library(stringr)
 ```
@@
 ```{r load_data}
-usage = read_tsv('../bene_usage_2011.tsv')
+denom = read_tsv('../bene_2011.tsv') %>%
+  filter(age >= 65) %>%
+  mutate(is_female=sex=='female') %>%
+  select(bene_id, age, is_female)
 chronic_conditions = read_tsv('cc_2011.tsv') %>%
@@
   mutate_at(vars(AMI:HYPOTH), function(x) x==3)
+usage = read_tsv('../pde_2011.tsv') %>%
+  count(bene_id) %>% rename(n_claims=n)
+
+na_to_0 = function(x) if_else(is.na(x), 0L, x)
+
+usage_summary = denom %>%
+  left_join(chronic_conditions, by='bene_id') %>%
+  left_join(usage, by='bene_id') %>%
+  mutate(n_claims=na_to_0(n_claims))
+
 condition_names = read_tsv('names.txt', col_names=c('condition', 'condition_name'))
-```
-
-```{r summarize_users}
-denom = usage %>% select(bene_id, age, sex) %>% distinct
-
-usage_summary = usage %>%
-  filter(age >= 65) %>%
-  group_by(bene_id) %>%
-  summarize(n_claims=sum(n_claims)) %>%
-  left_join(denom, by='bene_id') %>%
-  mutate(is_female=sex=='female') %>% select(-sex) %>%
-  left_join(chronic_conditions, by='bene_id')
 ```
@@
 
-library(broom)
 summarize_condition = function(x, condition) {
-  univariate_formula = paste0('n_claims ~ age + is_female + ', condition) %>%
-    formula
-
-  x %>%
-    glm(univariate_formula, family='poisson', data=.) %>%
+  rename_(x, .dots=setNames(condition, 'has_condition')) %>%
+    glm(n_claims ~ age + is_female + has_condition, family='poisson', data=.) %>%
     tidy %>%
@@
 
+```{r multivariate_poisson}
+multivariate_formula = paste(c('n_claims ~ age + is_female', conditions), collapse=' + ') %>% formula
+
+multivariate_res = usage_summary %>%
+  select(n_claims, age, is_female, AMI:HYPOTH) %>%
+  glm(multivariate_formula, data=., family='poisson') %>%
+  tidy
+```
+
 ```{r lasso_poisson}
-library(glmnet)
-require(doMC)
 registerDoMC(cores=3)
@@
 
-`r usage_summary %>% kable`
+**XXX**
 
@@
 usage_summary %>%
-  mutate(has_copd=COPD==3) %>%
-  group_by(has_copd) %>%
+  group_by(COPD) %>%
   do(summary(.$n_claims) %>% tidy) %>%
+  kable
+```
+
+### Multivariate
+
+```{r show_multivariate, cache=FALSE}
+# Here I need to do some fancy footwork to get the condition names in
+# The terms look like (Intercept), age, is_female, AMITRUE, COPDTRUE, etc.
+# So chop off TRUE where it is, then match with the known names, and
+# keep the (Intercept), etc. if it didn't match
+
+multivariate_res %>%
+  mutate(pct.increase=round((exp(estimate)-1.0)*100)) %>%
+  mutate(condition=str_replace(term, 'TRUE$', '')) %>%
+  left_join(condition_names, by='condition') %>%
+  mutate(term=if_else(is.na(condition_name), condition, condition_name)) %>%
+  select(term, pct.increase, estimate, p.value) %>%
+  arrange(desc(abs(estimate))) %>%
   kable
@@
   mutate_at(vars(lambda), function(x) signif(x, digits=2)) %>%
+  filter(df < 12) %>%
   kable
@@
   tidy %>%
-  select(term=row, coefficient=value) %>%
+  select(condition=row, coefficient=value) %>%
+  left_join(condition_names, by='condition') %>%
+  mutate(term=if_else(is.na(condition_name), condition, condition_name)) %>%
+  select(term, coefficient) %>%
+  arrange(desc(coefficient)) %>%
   kable
@@
 
-Not surprisingly, COPD tops the list.
-However, asthma does provide explanatory power beyond COPD, coming in second.
-(Lung cancer, however, now has a negative effect on consumption.) Strangely, depression and anemia are next.
-Age, kindney disease, and diabetes have coefficients very near zero!
+# Discussion
+
+Not surprisingly, COPD tops the list of chronic conditions associated with increased consumption.
+Interestingly, asthma does provide explanatory power beyond COPD, coming in second.
+Depression and anemia are next.
+Kidney disease has a middling effect, while age and diabetes have almost no effect!
+
+The lasso results seem to be the nicest: they sum up the major findings very well but seem to avoid some of the weird noise (e.g., enlarged prostate as the #2 coefficient in the multivariate Poisson model, lung cancer as the #3 coefficient in the univariate models).

### 2017-03-14 states/resistance/state-report.Rmd
many changes
added theil-sen

@@
 ---
+
+# To do
+
+- Implement weights in the linear correlations
+- Join multiple results into one table, or do pairwise comparisons of analytical methods
 
@@
 
+#### State-level resistance-averaged robust regressions
+
+```{r}
+library(mblm)
+
+safe_mblm = function(data, indep, expl) {
+  try(return(mblm(sprintf('%s ~ %s', indep, expl), data) %>% tidy), silent=TRUE)
+  data_frame(term=expl, estimate=NA, std.error=NA, statistic=NA, p.value=1.0)
+}
+
+ts_models = function(df, indep) {
+  lapply(c('cpkp', 'did'), function(expl) {
+    safe_mblm(df, indep, expl) %>%
+      filter(term==expl) %>%
+      rename(explanatory=term) %>%
+      verify(nrow(.) == 1)
+  }) %>% bind_rows
+}
+
+ts_models_result = function(x, this_place) {
+  x %>%
+    group_by_('bug', 'drug_group', this_place) %>%
+    summarize(mean_nonsus=mean(percent_nonsusceptible, na.rm=TRUE),
+              cpkp=head(cpkp, 1),
+              did=head(did, 1)) %>%
+    group_by(bug, drug_group) %>%
+    do(linear_models(., 'mean_nonsus')) %>%
+    ungroup %>%
+    filter(!is.na(estimate)) %>%
+    mutate(adj.p.value=p.adjust(p.value, method='BH'),
+           sig=pvalue_stars(adj.p.value)) %>%
+    select(bug, drug_group,
+           explanatory, slope=estimate, p.value,
+           adj.p.value, sig)
+}
+
+ts_state_res = linear_models_result(state_dat, 'state')
+
+ts_state_res %>% kable
+```
+
 #### HRR-level resistance-averaged

### 2017-03-16 hicks-summary/hicks-summary.Rmd
fixed hicks, added summary

--- 2017-02-22 hicks-summary/new-hicks/hicks-summary2.Rmd
@@
 
-```{r setup, echo=FALSE}
+```{r global_options, include=FALSE, cache=FALSE}
+knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='fig/',
+                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
+library(knitr)
 library(assertive)
+```
 
+```{r setup}
 # Group values in x into ranges determined by upper_limits.
@@
 
-summarize_by = function(usage, bene, by) {
-  denom = group_by_(bene, by) %>%
-    summarize(n_ppl=n())
+summarize_by = function(bene, pde, by) {
+  # e.g., by='sex'
+  # count number of beneficiaries in each group
+  denom = count_(bene, by) %>% rename(n_bene=n)
 
-  group_by_(usage, by) %>%
-    summarize(n_claims=n()) %>%
-    left_join(denom, by=by) %>%
-    mutate(cpkp=n_claims*1000/n_ppl) %>%
+  pde %>%
+    left_join(select_(bene, 'bene_id', by), by='bene_id') %>% # join information about the bene's sex (e.g.)
+    count_(by) %>% # count claims
+    left_join(denom, by=by) %>% # join n_bene
+    mutate(cpkp=n*1000/n_bene) %>%
     select_(by, 'cpkp')
@@
 
-```{r load_db, echo=FALSE, message=FALSE}
-
-regions = read_tsv('../../../db/census-regions/census-regions.tsv') %>%
+```{r load_db, cache=FALSE}
+regions = read_tsv('../../db/census-regions/census-regions.tsv') %>%
   select(state, region)
-
-antibiotic_class = read_tsv('../../abx_class.tsv') %>%
-  select(antibiotic, antibiotic_class) %>%
-  distinct
-
-assert_has_no_duplicates(antibiotic_class$antibiotic)
 ```
 
-```{r analyze_data, cache=TRUE, echo=FALSE, message=FALSE}
+```{r analyze_function}
 analyze = function(this_year) {
-  bene = read_tsv(sprintf('../../bene_%i.tsv', this_year)) %>%
-    mutate_at(vars(age, plan_coverage_months), as.integer) %>%
-    filter(!is.na(state), plan_coverage_months == 12, age >= 65) %>%
+  # bene file is already filtered for plan coverage months and good geographies
+  bene = read_tsv(sprintf('../bene_%i.tsv', this_year)) %>%
+    filter(age >= 65, hmo_months==12) %>%
     left_join(regions, by='state')
@@
 
-  pde = read_tsv(sprintf('../../pde_%i.tsv', this_year)) %>%
-    mutate_at(vars(days_supply), as.integer) %>%
-    left_join(antibiotic_class, by='antibiotic')
+  pde = read_tsv(sprintf('../pde_%i.tsv', this_year)) %>%
+    semi_join(bene, by='bene_id') # keep only PDEs for known beneficiaries
 
-  usage = pde %>%
-    inner_join(bene, by='bene_id') # inner join to keep only matching rows
+  total_n_pde = nrow(pde)
 
-  # usage only has PDEs from beneficiaries in our list,
-  # so there should be fewer
-  is_greater_than_or_equal_to(nrow(pde), nrow(usage))
+  totals = data_frame(key=c('n_bene', 'n_pde'),
+                      value=c(total_n_bene, total_n_pde),
+                      year=this_year)
 
   # claims per 1,000 people by abx class
-  claims_by_class = usage %>%
-    group_by(antibiotic_class) %>%
-    summarize(n_claims=n()) %>%
-    mutate(cpkp=n_claims*1000/total_n_bene) %>%
+  claims_by_class = pde %>%
+    count(antibiotic_class) %>%
+    mutate(cpkp=n*1000/total_n_bene) %>%
     select(antibiotic_class, cpkp) %>%
@@
   # the top few inidividual antibiotics
-  top_abx = usage %>%
-    group_by(antibiotic) %>%
-    summarize(n_claims=n()) %>%
-    mutate(cpkp=n_claims*1000/total_n_bene) %>%
+  top_abx = pde %>%
+    count(antibiotic) %>%
+    mutate(cpkp=n*1000/total_n_bene) %>%
     select(antibiotic, cpkp) %>% 
@@
   # in the same way as the consumption data, so we can use the summarize_by function
-  claims_by_sex = summarize_by(usage, bene, 'sex') %>%
+  claims_by_sex = summarize_by(bene, pde, 'sex') %>%
     mutate(year=this_year)
@@
   claims_by_age = summarize_by(
-    usage %>% mutate(age_group=age_group_f(age)),
     bene %>% mutate(age_group=age_group_f(age)),
-    'age_group') %>%
+    pde, 'age_group') %>%
     mutate(year=this_year)
@@
   # then by census region
-  claims_by_region = summarize_by(usage, bene, 'region') %>%
+  claims_by_region = summarize_by(bene, pde, 'region') %>%
     mutate(year=this_year)
@@
   # claims by state (broken into sextiles)
-  claims_by_state = summarize_by(usage, bene, 'state') %>%
+  claims_by_state = summarize_by(bene, pde, 'state') %>%
     mutate(sextile=ntile(cpkp, 6), year=this_year)
 
-  list(claims_by_class=claims_by_class,
+  list(totals=totals,
+       claims_by_class=claims_by_class,
        top_abx=top_abx,
@@
 
-```{r arrange_data, echo=FALSE}
-library(knitr)
+```{r arrange_data}
 data = list("2011"=data_2011,
@@
 data_by_field = function(field) lapply(data, function(x) x[[field]]) %>% bind_rows()
+
+totals = data_by_field('totals') %>%
+  spread(key, value) %>%
+  mutate(cpkp=n_pde*1000/n_bene) %>%
+  mutate_at(vars(n_bene, n_pde), function(x) prettyNum(x, big.mark=',')) %>%
+  rename(`number of beneficiaries`=n_bene, `number of antibiotic PDEs`=n_pde, `CPKP`=cpkp) %>%
+  arrange(year) %>%
+  kable
 
@@
 
+**To do**: Rework conclusions, include references.
+
 This summary is modeled after @hicks-us-2015.
 That study, which used the IMS Health Xponent database to infer the number of prescriptions per capita, reported higher prescription rates than what we observe here.
-In that paper, the number of prescriptions per 1,000 people year (PKPY) is approximately 1,000 for people over 65.
+In that paper, the number of claims per 1,000 people year (CKPY) is approximately 1,000 for people over 65.
 In this data, we find that number to be between 600 (in 2011) and 750 (in 2012).
@@
 A third report, which used the Xponent database, reported that prescriptions fell from about 1,000 PKPY in 1996 to about 600 PKPY in 2003 [@hicks-outpatient-2011].
+
+# Summary
+
+This data set is a 20% sample of Part D beneficiaries, which I have further filtered by age (only those at least 65 years old are included), for HMO membership (beneficiaries who were part of Medicare Advantage for any part of a year were excluded from that year), and for data quality (members with incomplete zipcodes were excluded).
+
+`r totals`
+
+There is a slight jump between the number of antibiotic claims in 2012 and 2013.
+This discontinuity appears between the December 2012 and January 2013, it is present among all PDE data (not just antibiotics), and it appears to be a real feature of the data.
 
@@
 
-```{r}
-# In Hicks, they also do:
-# logistic regression on...
-# dependent: county in top quartile for prescription rate
-# independent:
-#  - prescribers per capita above median,
-#  - proportion of people obese above median,
-#  - proportion of people <2 years old above median,
-#  - proportion of people black above median,
-#  - per capita income in top 2/3s,
-#  - proportion with college education in top 2/3s,
-#  - proportion female in top 3/4s
-#
-# Univariate regressions of each independent against the dependent
-# Multivariate regression somehow using adjusted ratios with confounders
-```
+# Other analysis in Hicks *et al.*
+
+In Hicks, they also do:
+
+- Logistic regression on...
+    - dependent: county in top quartile for prescription rate
+    - independent:
+        - prescribers per capita above median,
+        - proportion of people obese above median,
+        - proportion of people <2 years old above median,
+        - proportion of people black above median,
+        - per capita income in top 2/3s,
+        - proportion with college education in top 2/3s,
+        - proportion female in top 3/4s
+- Univariate regressions of each independent against the dependent
+- Multivariate regression somehow using adjusted ratios with confounders

### 2017-03-16 verification/all-pde-days/pde-days-between-years.Rmd
updated verification

@@
     mutate(date=dmy(date)) %>%
-    mutate(week_n=week(date))
+    mutate(week_n=week(date)) %>%
+    mutate(month_n=month(date))
 }
@@
   bind_rows() %>%
-  mutate(data_year=factor(data_year)) %>%
-  group_by(data_year, year(date), week_n) %>%
-  do(sum_by_week(.)) %>%
-  ungroup
+  mutate(data_year=factor(data_year))
+
+week_dat = dat %>%
+  group_by(data_year, year(date), week(date)) %>%
+  summarize(n_days=n(), count=sum(count), date=head(date, 1)) %>%
+  filter(n_days == 7)
+
+month_dat = dat %>%
+  group_by(data_year, year(date), month(date)) %>%
+  summarize(count=sum(count), date=head(date, 1))
 ```
 
-The Medicare data on the server contains Medicare Part D data for 2011 through 2014.
-I've been focusing on analyzing this data, especially in the files ``pdesaf2011.sas7bdat`` through ``pdesaf2014.sas7bdat``.
-These files contain the Part D events (PDEs), which are individual prescription drug claim records.
-
-It seems like there is something wrong with the 2012 data. I'm hoping you can help me understand what is going on.
-
-First, I noticed that the number of PDEs is greater in the 2012 data than in the other years:
+I'm now concerned about the 2013-2014 data (or the 2011-2012 data).
+It seems like the increase in the number of PDEs between 2012 and 2013 is larger than the 2011-2012 increase or the 2013-2014 increase.
 
@@
   summarize(n=sum(count)) %>%
-  mutate(n=round(n/1e6, 1)) %>%
+  mutate(n=round(n/1e6, 1), `diff. since prev. year`=c(0, diff(n))) %>%
   rename(`data year`=data_year, `number of PDEs (millions)`=n) %>%
@@
 
-This makes it seem like there were many more claims in 2012 than in the other years. However, the 2012 data file includes claims from 2011.
+Looking week-by-week, I see a discontinuity between December 2012 and January 2013:
 
-```{r plot_days, echo=FALSE}
-dat %>%
-  filter(week_n < 53) %>%
+```{r plot_weeks, echo=FALSE}
+week_dat %>%
   ggplot(aes(x=date, y=count, group=data_year, color=data_year)) +
@@
 
-This plot shows the number of PDEs that occurred during each week, colored by the year of the data file.
-The 2012 data file (``pdesaf2012.sas7bdat``) contains claims information for both 2011 and 2011, while the other three files only contain claims from their years.
+I wondered if this was an artifact of the way the busy days (e.g., the first of the month and the first days after holidays tend to have a larger number of PDEs) fell into the weeks.
+However, I see the same pattern even grouping by month:
 
-Furthermore, the number of claims in the 2012 data doesn't match up with the number of claims from the other years.
-The 2013 and 2014 data match up nicely, but it looks like there are missing claims for 2012.
-
-```{r compute_mean_missing, echo=FALSE}
-mean_missing_percent = dat %>%
-  filter(year(date) == 2011) %>%
-  select(date, count, data_year) %>%
-  spread(data_year, count) %>%
-  mutate(f=`2011`/`2012`) %$%
-  mean(f) %>%
-  { (. - 1) * 100 } %>% round
+```{r plot_months, echo=FALSE}
+month_dat %>%
+  ggplot(aes(x=date, y=count, group=data_year, color=data_year)) +
+  geom_line() + geom_point() +
+  ylab('PDEs in that month') +
+  theme_minimal()
 ```
-
-To see how many claims are "missing", I looked at PDEs from 2011.
-In the 2011 data file, there are on average `r mean_missing_percent`% more claims in the 2011 data file than in the 2012 data file.
-I'm not sure what this means.
-Do we have a 20% sample for the 2011, 2013, and 2014 data but only a 15% sample for the 2012 data (such that the 2011 data has $20/15 = 4/3$ fold, or 33% more, claims as the 2012 data)?

### 2017-03-24 chronic-conditions/young-people/cc-young.Rmd
young chronic

--- 2017-03-14 chronic-conditions/chronic-condition-drivers.Rmd
@@
 ---
-title: "Chronic conditions and increased antibiotic consumption"
+title: "Chronic conditions and antibiotic consumption among the non-elderly"
 author: "Scott Olesen"
@@
 
-Just summing up the chronic conditions doesn't seem like a very elegant way to
-summarize people's health. Choosing which chronic conditions to consider could
-be a matter of expert opinion. It's also worth asking which ones are
-interesting, from the data's viewpoint.
+We see that chronic conditions, not age, control consumption among the elderly.
+Is the same true among the non-elderly? The Medicare people are going to be
+pretty sick.
 
@@
 
-I'll use the Medicare 2011 data with normal filters (age 65 or above, only beneficiaries on Medicare for 12 months of the year).
+I'll use the Medicare 2011 data with different filters: age below 65, on
+Medicare for 12 months of the year).
 
-For each chronic condition, I did a Poisson regression of the number of claims for a beneficiary against the beneficiary's age and their status for that condition.
-I hoped that including age in the regression would prevent a confounding between age, conditions, and consumption.
+For each chronic condition, I did a Poisson regression of the number of claims
+for a beneficiary against the beneficiary's age and their status for that
+condition. I hoped that including age in the regression would prevent a
+confounding between age, conditions, and consumption.
 
-To report the results of the regression, I show the "percent increase" due to a variable in the regression. A Poisson regression estimates that
-$E(Y|x) = \exp(\alpha + \beta x)$. Let $\beta_i$ be the term for the boolean variable $x_i$ of interest. Then the expected increase in consumption
-due to that $x_i$ being true is $\exp(\beta_i)$.
+To report the results of the regression, I show the "percent increase" due to a
+variable in the regression. A Poisson regression estimates that $E(Y|x) =
+\exp(\alpha + \beta x)$. Let $\beta_i$ be the term for the boolean variable
+$x_i$ of interest. Then the expected increase in consumption due to that $x_i$
+being true is $\exp(\beta_i)$.
 
@@
 ```{r load_data}
-denom = read_tsv('../bene_2011.tsv') %>%
-  filter(age >= 65) %>%
+bene = read_tsv('../../bene_2011.tsv') %>%
+  filter(age < 65) %>%
   mutate(is_female=sex=='female') %>%
   select(bene_id, age, is_female)
-chronic_conditions = read_tsv('cc_2011.tsv') %>%
+chronic_conditions = read_tsv('../cc_2011.tsv') %>%
   rename(bene_id=BENE_ID) %>%
@@
   mutate_at(vars(AMI:HYPOTH), function(x) x==3)
-usage = read_tsv('../pde_2011.tsv') %>%
+pde_summary = read_tsv('../../pde_2011.tsv') %>%
   count(bene_id) %>% rename(n_claims=n)
@@
 
-usage_summary = denom %>%
+usage = bene %>%
   left_join(chronic_conditions, by='bene_id') %>%
-  left_join(usage, by='bene_id') %>%
+  left_join(pde_summary, by='bene_id') %>%
   mutate(n_claims=na_to_0(n_claims))
 
-condition_names = read_tsv('names.txt', col_names=c('condition', 'condition_name'))
+condition_names = read_tsv('../names.txt', col_names=c('condition', 'condition_name'))
 ```
@@
 
-res = lapply(conditions, function(cond) summarize_condition(usage_summary, cond)) %>%
+res = lapply(conditions, function(cond) summarize_condition(usage, cond)) %>%
   bind_rows() %>%
@@
 
-```{r multivariate_poisson}
-multivariate_formula = paste(c('n_claims ~ age + is_female', conditions), collapse=' + ') %>% formula
-
-multivariate_res = usage_summary %>%
-  select(n_claims, age, is_female, AMI:HYPOTH) %>%
-  glm(multivariate_formula, data=., family='poisson') %>%
-  tidy
-```
-
 ```{r lasso_poisson}
@@
 
-lasso_x = usage_summary %>%
+lasso_x = usage %>%
   select(age, is_female, AMI:HYPOTH) %>%
   as.matrix
-lasso_y = usage_summary$n_claims
+lasso_y = usage$n_claims
 
@@
 
-## Summary
-
-**XXX**
-
 ## Correlations
-
-Presumably many of the conditions are correlated. I computed tetrachoric correlations, which is a method designed to deal with correlations of dichotomous variables.
 
@@
 
-Thankfully, Alzheimer's correlates strongly with Alzheimer's and dementia, and prostatic hyperplasia and prostate cancer correlate strongly negatively with breast cancer and endometrial cancer.
-Unfortunately, the negative correlations make everything else harder to see, so here's the same thing but without the male conditions.
+Without male conditions (prostate cancer, enlarged prostate):
 
@@
 
-COPD correlates strongly with asthma and lung cancer, but lung cancer and asthma do not correlate strongly with one another.
-There is a block of correlations focused around hypertension that includes ischemic heart disease, heart failure, anemia, chronic kidney disease, diabetes, hypertension, hyperlipidemia, and, importantly, COPD.
-
 ## Regression
@@
 
-First, I performed univariate regressions. In each case, the explanatory variable was the presence of the chronic condition.
+First, I performed univariate regressions. In each case, the explanatory
+variable was the presence of the chronic condition.
 
-Almost all the conditions are significantly associated with greater consumption.
-Here I show the results of the Poisson regression: the column `estimate` is the estimate for the parameters associated with each condition, and `p.value` measures whether the data are consistent with that parameter being zero.
-The column `pct.increase` shows that percent increase in number of claims derived from that estimate.
+Here I show the results of the Poisson regression: the column
+`estimate` is the estimate for the parameters associated with each condition,
+and `p.value` measures whether the data are consistent with that parameter
+being zero. The column `pct.increase` shows that percent increase in number of
+claims derived from that estimate.
 
@@
 
-As a verification, I'll look at the distribution of number of claims among patients who have COPD or not.
-The mean for patients with COPD is more than double than that for patients without, which makes the age-adjusted 2.23 fold increase seem reasonable.
-
-```{r copd}
-usage_summary %>%
-  group_by(COPD) %>%
-  do(summary(.$n_claims) %>% tidy) %>%
-  kable
-```
-
-### Multivariate
-
-```{r show_multivariate, cache=FALSE}
-# Here I need to do some fancy footwork to get the condition names in
-# The terms look like (Intercept), age, is_female, AMITRUE, COPDTRUE, etc.
-# So chop off TRUE where it is, then match with the known names, and
-# keep the (Intercept), etc. if it didn't match
-
-multivariate_res %>%
-  mutate(pct.increase=round((exp(estimate)-1.0)*100)) %>%
-  mutate(condition=str_replace(term, 'TRUE$', '')) %>%
-  left_join(condition_names, by='condition') %>%
-  mutate(term=if_else(is.na(condition_name), condition, condition_name)) %>%
-  select(term, pct.increase, estimate, p.value) %>%
-  arrange(desc(abs(estimate))) %>%
-  kable
-```
-
 ### Multivariate lasso
 
-I also performed a multivariate lasso regression with the Poisson model, keeping all the chronic conditions as well as age and sex (coded as "is female") as explanatory variables.
-I started with 10-fold cross-validation to pick the lasso $\lambda$.
+I also performed a multivariate lasso regression with the Poisson model,
+keeping all the chronic conditions as well as age and sex (coded as "is
+female") as explanatory variables.  I started with 10-fold cross-validation to
+pick the lasso $\lambda$.
 
@@
 
-Although the error continues to fall with decreasing $\lambda$, it seems like there are diminishing returns after the first 8 or 10 variables are included.
+Although the error continues to fall with decreasing $\lambda$, it seems like
+there are diminishing returns after the first 8 or 10 variables are included.
 
@@
 # Discussion
-
-Not surprisingly, COPD tops the list of chronic conditions associated with increased consumption.
-Interestingly, asthma does provide explanatory power beyond COPD, coming in second.
-Depression and anemia are next.
-Kidney disease has a middling effect, while age and diabetes have almost no effect!
-
-The lasso results seem to be the nicest: they sum up the major findings very well but seem to avoid some of the weird noise (e.g., enlarged prostate as the #2 coefficient in the multivariate Poisson model, lung cancer as the #3 coefficient in the univariate models).

### 2017-03-27 chronic-conditions/chronic-condition-drivers.Rmd
sex difference plots

@@
 
-I'll use the Medicare 2011 data with normal filters (age 65 or above, only beneficiaries on Medicare for 12 months of the year).
+I'll use the Medicare 2011 data with normal filters (age 65 or above, only beneficiaries on Medicare for 12 months of the year). I'll run separate analyses on those under 65.
 
@@
 ```{r load_data}
-denom = read_tsv('../bene_2011.tsv') %>%
+all_denom = read_tsv('../bene_2011.tsv') %>%
   filter(age >= 65) %>%

### 2017-03-27 drug-drug-correlations/drug-drug.Rmd
sex difference plots

--- 2017-03-13 abx-abx-correlations/drug-drug.Rmd
@@
 Each joint observation ($x_i$, $y_i$) can be one of $(0, 0)$, $(0, 1)$, $(1, 0)$, or $(1, 1)$.
-The tetrachoric correlation finds a correlation $\rho$ by finding the "cutpoints" $tau_X$ and $tau_Y$ such that the bivariate standard normal distribution with correlation $\rho$ has "quadrants" with masses that best match the relative frequencies of the joint observations.
+The tetrachoric correlation finds a correlation $\rho$ by finding the "cutpoints" $\tau_X$ and $\tau_Y$ such that the bivariate standard normal distribution with correlation $\rho$ has "quadrants" with masses that best match the relative frequencies of the joint observations.
 For example, the "lower-left" quadrant has mass
@@
 
+Yonatan has a number of interpretations and guesses:
+
+- Azithromycin and levofloxacin are used for pneumonia, and they will also go in that sequential order
+- Cipro, Bactrim, and nitro are for UTIs and will proceed in that order
+- Keflex/Bactrim is for skin and soft tissue infections
+- Cipro and metro are simulatenous
+- Levo and metro are simulatneous
+
+We should be able to compare these to what's actually happening, which is a
+statement about prescribing practice.
+
 # Discussion
@@
 Levofloxacin has a weaker correlation with metronidazole than ciprofloxacin does, but azithromycin's strongest correlation is with azithromycin.
+
+There is probably a technically-superior way to do this: I measured consumption
+as an on-off thing, and the polychoric only works well with a small number of
+categories. I may be able to reduce the number of claims into a small number of
+groups (e.g., 0, 1, 2--5, etc.). I wonder also if there is a "Poisson regression",
+where you postulate that the Poisson means are correlated.

### 2017-03-27 inequality-resistance/inequality-resistance.Rmd
sex difference plots

@@
 ```{r global_options, include=FALSE}
-knitr::opts_chunk$set(fig.width=7, fig.height=4, 
+knitr::opts_chunk$set(fig.width=7, fig.height=4, dev=c('pdf', 'png'), 
                       echo=TRUE, warning=FALSE, message=FALSE, cache=TRUE)

### 2017-03-27 sex-difference/sex-difference.Rmd
sex difference plots

@@
 ```{r global_options, include=FALSE}
-knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/',
+knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/', dev=c('png', 'pdf'),
                       echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
@@
 
-```{r}
+```{r sex_by_age}
 plot_dat = usage %>%

### 2017-03-27 states/resistance/state-report.Rmd
sex difference plots

@@
 
-```{r global_options, include=FALSE}
+```{r global_options, include=FALSE, cache=FALSE}
 knitr::opts_chunk$set(fig.width=7, fig.height=4, cache=TRUE,
                       echo=FALSE, warning=FALSE, message=FALSE)
+library(broom)
+library(pander)
+library(assertr)
 ```
@@
 ```{r setup}
-library(broom)
-library(pander)
-library(assertr)
-
 # swo> I removed Klebsiella because it's mostly in immune-compromised people
@@
   #filter(state %in% known_states)
+```
+
+```{r load_consumption_data}
+bene = read_tsv('../../bene_2011.tsv') %>%
+  left_join(hrr, by='zipcode')
+pde = read_tsv('../../pde_2011.tsv') %>%
+  select(bene_id, drug=antibiotic) %>%
+  left_join(select(bene, bene_id, state, hrr), by='bene_id')
+
+na20 = function(x) if_else(is.na(x), 0L, x)
+
+state_denom = bene %>% count(state) %>% rename(n_bene=n)
+state_consumption = pde %>%
+  group_by(state, drug) %>%
+  summarize(n_claims=n()) %>%
+  ungroup() %>%
+  left_join(state_denom, by='state')
+
+hrr_denom = bene %>% count(hrr) %>% rename(n_bene=n)
+hrr_consumption = pde %>%
+  group_by(hrr, drug) %>%
+  summarize(n_claims=n()) %>%
+  ungroup() %>%
+  left_join(hrr_denom, by='hrr')
 
 # State-level consumption data
-state_consumption = read_tsv('../bene-consumption/state_usage_2011.tsv') %>%
-  filter(!is.na(state)) %>%
-  right_join(consumption_groups, by=c('antibiotic'='drug')) %>%
-  group_by(state, drug_group) %>%
-  summarize_at(vars(cpkp, did), sum)
+#state_consumption = read_tsv('../bene-consumption/state_usage_2011.tsv') %>%
+  #filter(!is.na(state)) %>%
+  #right_join(consumption_groups, by=c('antibiotic'='drug')) %>%
+  #group_by(state, drug_group) %>%
+  #summarize_at(vars(cpkp, did), sum)
 

### 2017-03-27 carrier-outpatient/carrier-outpatient.Rmd
new analyses

---
title: "Carrier and outpatient claims"
author: "Scott Olesen"
date: "4 March 2017"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, 
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
```

# Methods

I use the 2011 data.

```{r load_data}
n_max=1e4

carrier_count = read_tsv('../../data/carrier_count_2011.tsv', n_max=n_max) %>%
  rename(bene_id=BENE_ID, carrier_count=n_from_dates)
outpatient_count = read_tsv('../../data/outpatient_count_2011.tsv', n_max=n_max) %>%
  rename(bene_id=BENE_ID, outpatient_count=n_from_dates)
inpatient_count = read_tsv('../../data/inpatient_count_2011.tsv', n_max=n_max) %>%
  rename(bene_id=BENE_ID, inpatient_count=n_from_dates)
```

## Carrier character information

Are there repetitions among the carrier data? Yes. Mostly these are for zipcodes: the same beneficiary can have claims associated to multiple zipcodes. Many of these zipcodes are nearby (e.g., one in the suburbs and one in the city), and for the most part, one of the zipcodes is the same as the zipcode that we have for that beneficiary as their home address. In a few cases, the claims zipcodes include claims from multiple states (which I guess means that they were traveling and got ill, or that they have multiple homes, or stay with family, etc.).

In a few cases, a beneficiary has multiple birth dates. I looked at a few cases, which just looked like typos or small errors (e.g., "14" and "19" for the day of the month). One of the dates matched the one in the beneficiary information.

```{r}
library(stringr)

bene_char = read_tsv('../bene_2011.tsv') %>%
  select(bene_id, zipcode)

carrier_char = read_tsv('../../data/carrier_char_2011.tsv', n_max=n_max) %>%
  rename(bene_id=BENE_ID, date_of_birth=DOB_DT, zipcode=ZIP_CD) %>%
  mutate(date_of_birth=dmy(date_of_birth), zipcode=str_sub(zipcode, 1, 5))

repeated_benes = carrier_char %>%
  filter(duplicated(bene_id)) %>%
  select(bene_id) %>%
  distinct() %>%
  left_join(carrier_char, by='bene_id') %>%
  left_join(bene_char, by='bene_id', suffix=c('.carrier', '.bene'))
```

### 2017-03-27 death-year/death-year.Rmd
new analyses

---
title: "Death years"
author: "Scott Olesen"
date: ""
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, 
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
```

Do Medicare beneficiaries use a lot of antibiotics in their last year of life? Is it bad to be excluding these people?

```{r load_data}
bene = read_tsv('bene_2011.tsv')
pde = read_tsv('pde_2011.tsv')

na20 = function(x) if_else(is.na(x), 0L, x)

usage = pde %>%
  group_by(bene_id) %>%
  summarize(n_claims=n(), n_days=sum(days_supply)) %>%
  right_join(bene, by='bene_id') %>%
  mutate_at(vars(n_claims, n_days), na20)
```

```{r histograms}
usage %>%
  filter(is.na(death_date)) %>%
  ggplot(aes(x=n_claims)) +
  geom_histogram() +
  scale_y_log10()

usage %>%
  filter(!is.na(death_date)) %>%
  mutate(death_week=week(death_date),
         scaled_n_claims=52/death_week * n_claims) %>%
  ggplot(aes(x=scaled_n_claims)) +
  geom_histogram() +
  scale_y_log10()
```

```{r tables}
nz = function(x) x[x > 0]
fnz = function(x) length(nz(x)) / length(x)

usage %>%
  mutate(died=!is.na(death_date)) %>%
  group_by(died) %>%
  summarize(n_beneficiaries=n(),
            mean_n=mean(n_claims),
            fraction_nonzero=fnz(n_claims),
            nonzero_mean=mean(nz(n_claims))) %>%
  kable
```

### 2017-03-27 top-users/top-users.Rmd
new analyses

---
title: "Top antibiotic users"
author: "Scott Olesen"
date: "27 March 2017"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, 
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
```

# Motivation

What drugs are the top users taking? Are they different from the drugs that people take in general? Are the patterns of drug-drug correlations different among the top users?

# Data

I use the 2011 Medicare data. I will sometimes exclude claims for the anti-tuberculosis drugs (ethambutol, rifampin, isoniazid, pyrazinamide) because many of the "top" users, depending on the defintion, are taking these drugs, which constitute a different problem.

```{r load_data}
bene = read_tsv('../bene_2011.tsv')
n_bene = nrow(bene)
pde = read_tsv('../pde_2011.tsv') %>%
  rename(drug=antibiotic) %>%
  group_by(bene_id, drug) %>%
  summarize(n_claims=n(), n_days=sum(days_supply)) %>%
  ungroup()
```

# Results

## Drug consumption sort-of follows a power law

That is, the *exponent* of the number of beneficiaries with a certain of claims *n* (or days' supply of drug) falls with *n*.

```{r drug_distribution}
pde_total = pde %>%
  group_by(bene_id) %>%
  summarize(n_total_claims=sum(n_claims), n_total_days=sum(n_days))

pde_total %>%
  count(n_total_claims) %>%
  ggplot(aes(x=n_total_claims, y=n)) +
  geom_point() +
  scale_y_log10() +
  xlab('number of claims') + ylab('number of beneficiaries') +
  theme_minimal()

pde_total %>%
  count(n_total_days) %>%
  ggplot(aes(x=n_total_days, y=n)) +
  geom_point() +
  scale_y_log10() +
  xlab('number of days\' supply') + ylab('number of beneficiaries') +
  theme_minimal()
```

## Top users by claim are enriched for vancomycin

- Merge amox, amox/clav
- Exclude TB drugs? Do with and without?
- Split people into top users (by # claims) or not
- Analyze them that way (with CPKP)
- Add zeros for people/drug combos? Or just find one denominator and pad out
- Then repeat the same thing but using days (and DID)
- Make plots & tables to compare both

Expected punchlines:

- TB drugs make up a lot of the top use
- Top claims enrich for vancomycin, which people are taking with ~2 day claims
- Top days are for mostly the same drugs

```{r top_claims_use}
top_claims_cutoff = 15

pad_to = function(x, n) c(x, rep(0, n - length(x)))
stopifnot(pad_to(c(1, 2, 3), 5) == c(1, 2, 3, 0, 0))

# classify beneficiaries as in top bracket or not
top_claims_bene = pde %>%
  group_by(bene_id) %>%
  summarize(top_user = sum(n_claims) >= top_claims_cutoff)

n_top_claims_bene = sum(top_claims_bene$top)

pde %>%
  left_join(top_claims_bene, by='bene_id') %>%
  group_by(drug, top) %>%
  summarize(n_claims=sum(n_claims)) %>%
  ungroup() %>%
  mutate(n_bene=if_else(top, n_top_claims_bene, n_bene - n_top_claims_bene),
         cpkp=n_claims*1000/n_bene) %>%
  select(drug, top, cpkp) %>%
  spread(top, cpkp) %>%
  arrange(desc(`TRUE`)) %>%
  head(20) %>%
  rename(`non-top user CPKP`=`FALSE`, `top user CPKP`=`TRUE`) %>%
  kable
```

Of `r nrow(bene)` total beneficiaries, there are `r n_top_bene` that had more than `r n_top_cutoff` claims. In aggregate, their consumption looks fairly different from the rest of the population:

`r top_pde_table`

```{r compute_user_overlap}
# The idea here was to see about the overlap between drug users.
# If you took drug 1, how likely was it for you to also take drug 2?
# This is actually pretty similar to the tetrachoric analysis I did previously.
# Instead, I summarized the confusion matrix with the Matthews correlations coefficient.

drugs = pde$drug %>% unique

user_overlap = function(pde, drug1_name, drug2_name) {
  full_join(pde %>% filter(drug==drug1_name) %>% rename(drug1=drug),
            pde %>% filter(drug==drug2_name) %>% rename(drug2=drug), by='bene_id') %>%
    replace_na(list(drug1=0, drug2=0)) %>%
    mutate(case=case_when(.$drug1 > 0 & .$drug2 > 0 ~ 'took_both',
                          .$drug1 > 0 ~ 'took_drug1_only',
                          .$drug2 > 0 ~ 'took_drug2_only',
                          TRUE ~ 'took_neither')) %>%
    count(case) %>%
    spread(case, n) %>%
    mutate(drug1=drug1_name, drug2=drug2_name)
}

mcc = function (tp, fp, fn, tn) {
  # tp = true positive, fp = false positive, etc.
  n = as.double(tp + fp + fn + tn)
  s = (tp + fn) / n
  p = (tp + fp) / n
  return (tp / n - (s * p)) / sqrt(p * s * (1 - s) * (1 - p))
  #return ((tp * tn) - (fp * fn)) / sqrt((tp + fp) * (tp + fn) * (tn + fp) * (tn + fn))
}

#drug_overlap = crossing(drug1=drugs, drug2=drugs) %>%
#  filter(drug1 > drug2) %>%
#  rowwise() %>% do(user_overlap(pde, .$drug1, .$drug2)) %>%
#  mutate(took_neither=n_bene-(took_both + took_drug1_only + took_drug2_only)) %>%
#  select(drug1, drug2, took_both, took_drug1_only, took_drug2_only, took_neither) %>%
#  mutate(mcc=mcc(took_both, took_drug1_only, took_drug2_only, took_neither))
```

```{r top_user_overlap}
# The idea here was to ask which users are the "top" users for that drug.
# Then I would overlap them in the same way that I did for any-users above.

pad_to = function(x, n, with=0) {
  if (length(x) > n) stop('input longer than pad to length')
  c(x, rep(0, n - length(x)))
}

top_bracket_limit = function(n, fraction) {
  n[cume_dist(n) > (1 - fraction)] %>% min
}

#drug_upper_brackets = pde %>%
#  inner_join(pde %>% group_by(bene_id) %>% summarize(n_claims=sum(n_claims)) %>% mutate(antibiotic='all'), by='bene_id') %>%
#  group_by(antibiotic) %>%
#  summarize(limit=top_bracket_limit(pad_to(n_claims, n_bene), 0.001)) %>%
#  filter(limit > 0)
```

### 2017-03-27 death-year/polish-bene-pde.R

@@ -1,0 +1,113 @@
+#!/usr/bin/env Rscript
+
+# This is mostly the same as the polish PDE file in the above directory, but
+# here I allow less than 12 months of coverage in the case that the person died
+
+state_codes = read_tsv('../../db/state-codes/state-codes.tsv') %>%
+  mutate(is_state_or_dc=as.logical(is_state_or_dc)) %>%
+  filter(is_state_or_dc)
+
+oral_injected_codes = scan('../dosage-form/oral_injected_codes.txt', what=character())
+antibiotic_names = read_tsv('../abx_class.tsv')
+
+polish_bene = function(x) {
+  rename(x, bene_id=BENE_ID,
+          state_cd=STATE_CD,
+          age=AGE,
+          hmo_months=HMO_MO,
+          plan_coverage_months=PLNCOVMO) %>%
+    mutate_at(vars(age, hmo_months, plan_coverage_months), as.integer) %>% # should be already done from read_tsv
+    left_join(state_codes, by='state_cd') %>%
+    mutate(zipcode=substr(BENE_ZIP, 1, 5)) %>%
+    mutate(sex=case_when(.$SEX == 0 ~ 'unknown',
+                          .$SEX == 1 ~ 'male',
+                          .$SEX == 2 ~ 'female')) %>%
+    mutate(race=case_when(.$RTI_RACE_CD == 0 ~ 'unknown',
+                          .$RTI_RACE_CD == 1 ~ 'white', # non-Hispanic white
+                          .$RTI_RACE_CD == 2 ~ 'black',
+                          .$RTI_RACE_CD == 3 ~ 'other',
+                          .$RTI_RACE_CD == 4 ~ 'Asian', # or Pacific Islander
+                          .$RTI_RACE_CD == 5 ~ 'Hispanic',
+                          .$RTI_RACE_CD == 6 ~ 'American Indian')) %>% # or Alaska Native
+    mutate(death_date=dmy(DEATH_DT)) %>%
+    filter(!is.na(state),
+           plan_coverage_months > 0, hmo_months==0,
+           plan_coverage_months==12 | !is.na(death_date)) %>% # filter for geography and plan
[…]

### 2017-03-27 top-users/top-users.R

@@ -1,0 +1,47 @@
+bene = read_tsv('../bene_2011.tsv')
+pde = read_tsv('../pde_2011.tsv') %>%
+  rename(drug=antibiotic) %>%
+  group_by(bene_id, drug) %>%
+  summarize(n_claims=n()) %>%
+  ungroup()
+
+drugs = unique(pde$drug)
+
+# add "all" antibiotics (the sum of usage of all drugs)
+#pde %<>%
+#  bind_rows(pde %>% group_by(bene_id) %>% summarize(n_claims=sum(n_claims)) %>% mutate(antibiotic='all'))
+
+n_bene = nrow(bene)
+
+user_overlap = function(pde, drug1_name, drug2_name) {
+  full_join(pde %>% filter(drug==drug1_name) %>% rename(drug1=drug),
+            pde %>% filter(drug==drug2_name) %>% rename(drug2=drug), by='bene_id') %>%
+    replace_na(list(drug1=0, drug2=0)) %>%
+    mutate(case=case_when(.$drug1 > 0 & .$drug2 > 0 ~ 'took_both',
+                          .$drug1 > 0 ~ 'took_drug1_only',
+                          .$drug2 > 0 ~ 'took_drug2_only',
+                          TRUE ~ 'took_neither')) %>%
+    count(case) %>%
+    spread(case, n) %>%
+    mutate(drug1=drug1_name, drug2=drug2_name)
+}
+
+drug_overlap = crossing(drug1=drugs, drug2=drugs) %>%
+  filter(drug1 > drug2) %>%
+  rowwise() %>% do(user_overlap(pde, .$drug1, .$drug2)) %>%
+  mutate(took_neither=n_bene-(took_both + took_drug1_only + took_drug2_only))
+
+pad_to = function(x, n, with=0) {
+  if (length(x) > n) stop('input longer than pad to length')
+  c(x, rep(0, n - length(x)))
+}
+
+top_bracket_limit = function(n, fraction) {
+  n[cume_dist(n) > (1 - fraction)] %>% min
+}
+
+drug_upper_brackets = pde %>%
+  inner_join(pde %>% group_by(bene_id) %>% summarize(n_claims=sum(n_claims)) %>% mutate(antibiotic='all'), by='bene_id') %>%
+  group_by(antibiotic) %>%
+  summarize(limit=top_bracket_limit(pad_to(n_claims, n_bene), 0.001)) %>%
+  filter(limit > 0)  

### 2017-03-28 top-users/top-users.Rmd
more tables

@@
 
+```{r}
+pn = function(x) prettyNum(x, big.mark=',')
+```
+
 # Motivation
@@
 
+I merged amoxicillin/clavulanate into amoxicillin.
+
 ```{r load_data}
@@
   rename(drug=antibiotic) %>%
+  mutate(drug=if_else(drug=='amoxicillin/clavulanate', 'amoxicillin', drug)) %>%
   group_by(bene_id, drug) %>%
@@
 
-## Top users by claim are enriched for vancomycin
+## Top users by claim
 
@@
 
-pad_to = function(x, n) c(x, rep(0, n - length(x)))
-stopifnot(pad_to(c(1, 2, 3), 5) == c(1, 2, 3, 0, 0))
-
 # classify beneficiaries as in top bracket or not
@@
   group_by(bene_id) %>%
-  summarize(top_user = sum(n_claims) >= top_claims_cutoff)
+  summarize(top = sum(n_claims) >= top_claims_cutoff)
 
@@
 
-pde %>%
+top_claims_table = pde %>%
   left_join(top_claims_bene, by='bene_id') %>%
@@
 
-Of `r nrow(bene)` total beneficiaries, there are `r n_top_bene` that had more than `r n_top_cutoff` claims. In aggregate, their consumption looks fairly different from the rest of the population:
+Of `r pn(n_bene)` total beneficiaries, there are `r pn(n_top_claims_bene)` that had more than `r top_claims_cutoff` claims.
+In aggregate, their consumption looks fairly different from the rest of the population:
 
-`r top_pde_table`
+`r top_claims_table`
+
+Notable differences include:
+
+- More urinary anti-infectives (Bactrim, nitro, and trimethoprim)
+- Much more vancomycin
+- More anti-TB drugs (esp. rifampin, ethambutol)
+
+## Top users by days
+
+Measuring top users by claims biases toward people taking drugs that are given with short dosages.
+
+```{r top_days_use}
+# classify beneficiaries as in top bracket or not
+top_days_bene = pde %>%
+  group_by(bene_id) %>%
+  summarize(top = sum(n_days) >= 365)
+
+n_top_days_bene = sum(top_days_bene$top)
+
+top_days_table = pde %>%
+  left_join(top_days_bene, by='bene_id') %>%
+  group_by(drug, top) %>%
+  summarize(n_days=sum(n_days)) %>%
+  ungroup() %>%
+  mutate(n_bene=if_else(top, n_top_days_bene, n_bene - n_top_days_bene),
+         did=n_days*1000/(365*n_bene)) %>%
+  select(drug, top, did) %>%
+  spread(top, did) %>%
+  arrange(desc(`TRUE`)) %>%
+  head(20) %>%
+  rename(`non-top user DID`=`FALSE`, `top user DID`=`TRUE`) %>%
+  kable
+```
+
+Of `r pn(n_bene)` total beneficiaries, there are `r pn(n_top_days_bene)` that had more than 365 days of days' supply.
+In aggregate, their consumption looks fairly different from the rest of the population:
+
+`r top_days_table`
+
+The drugs that make it into the top-days-users top list are similar to those from the top-claims-users, except that vancomycin is missing here.
+
+## Top claims, no anti-TB
+
+I worry that we're mostly selecting for people who have TB when doing these analyses, so I did a different version where I excluded the TB drugs:
+
+```{r tb_drugs}
+tb_drugs = c('isoniazid', 'rifabutin', 'rifampin', 'rifampin/isoniazid', 'rifampin/isoniazid/pyrazinamide', 'rifapentine', 'rifamixin', 'pyrazinamide', 'ethambutol')
+stopifnot(all(tb_drugs %in% unique(pde$drug)))
+
+data_frame(drug=tb_drugs) %>% kable
+```
+
+```{r top_claims_notb}
+top_notb_cutoff = 15
+
+# classify beneficiaries as in top bracket or not
+top_notb_bene = pde %>%
+  filter(!(drug %in% tb_drugs)) %>%
+  group_by(bene_id) %>%
+  summarize(top = sum(n_claims) >= top_notb_cutoff)
+
+n_top_notb_bene = sum(top_notb_bene$top)
+
+top_notb_table = pde %>%
+  left_join(top_notb_bene, by='bene_id') %>%
+  group_by(drug, top) %>%
+  summarize(n_claims=sum(n_claims)) %>%
+  ungroup() %>%
+  mutate(n_bene=if_else(top, n_top_notb_bene, n_bene - n_top_notb_bene),
+         cpkp=n_claims*1000/n_bene) %>%
+  select(drug, top, cpkp) %>%
+  spread(top, cpkp) %>%
+  arrange(desc(`TRUE`)) %>%
+  head(20) %>%
+  rename(`non-top user CPKP`=`FALSE`, `top user CPKP`=`TRUE`) %>%
+  kable
+```
 

### 2017-03-28 top-users/top-users.Rmd
top users TB

@@
 ## Top users by claim
-
-- Merge amox, amox/clav
-- Exclude TB drugs? Do with and without?
-- Split people into top users (by # claims) or not
-- Analyze them that way (with CPKP)
-- Add zeros for people/drug combos? Or just find one denominator and pad out
-- Then repeat the same thing but using days (and DID)
-- Make plots & tables to compare both
-
-Expected punchlines:
-
-- TB drugs make up a lot of the top use
-- Top claims enrich for vancomycin, which people are taking with ~2 day claims
-- Top days are for mostly the same drugs
 
@@
 ```{r tb_drugs}
-tb_drugs = c('isoniazid', 'rifabutin', 'rifampin', 'rifampin/isoniazid', 'rifampin/isoniazid/pyrazinamide', 'rifapentine', 'rifamixin', 'pyrazinamide', 'ethambutol')
+tb_drugs = c('isoniazid', 'rifabutin', 'rifampin', 'rifampin/isoniazid', 'rifampin/isoniazid/pyrazinamide', 'rifapentine', 'rifaximin', 'pyrazinamide', 'ethambutol')
 stopifnot(all(tb_drugs %in% unique(pde$drug)))
@@
 
-```{r compute_user_overlap}
-# The idea here was to see about the overlap between drug users.
-# If you took drug 1, how likely was it for you to also take drug 2?
-# This is actually pretty similar to the tetrachoric analysis I did previously.
-# Instead, I summarized the confusion matrix with the Matthews correlations coefficient.
+`r top_notb_table`
 
-drugs = pde$drug %>% unique
-
-user_overlap = function(pde, drug1_name, drug2_name) {
-  full_join(pde %>% filter(drug==drug1_name) %>% rename(drug1=drug),
-            pde %>% filter(drug==drug2_name) %>% rename(drug2=drug), by='bene_id') %>%
-    replace_na(list(drug1=0, drug2=0)) %>%
-    mutate(case=case_when(.$drug1 > 0 & .$drug2 > 0 ~ 'took_both',
-                          .$drug1 > 0 ~ 'took_drug1_only',
-                          .$drug2 > 0 ~ 'took_drug2_only',
-                          TRUE ~ 'took_neither')) %>%
-    count(case) %>%
-    spread(case, n) %>%
-    mutate(drug1=drug1_name, drug2=drug2_name)
-}
-
-mcc = function (tp, fp, fn, tn) {
-  # tp = true positive, fp = false positive, etc.
-  n = as.double(tp + fp + fn + tn)
-  s = (tp + fn) / n
-  p = (tp + fp) / n
-  return (tp / n - (s * p)) / sqrt(p * s * (1 - s) * (1 - p))
-  #return ((tp * tn) - (fp * fn)) / sqrt((tp + fp) * (tp + fn) * (tn + fp) * (tn + fn))
-}
-
-#drug_overlap = crossing(drug1=drugs, drug2=drugs) %>%
-#  filter(drug1 > drug2) %>%
-#  rowwise() %>% do(user_overlap(pde, .$drug1, .$drug2)) %>%
-#  mutate(took_neither=n_bene-(took_both + took_drug1_only + took_drug2_only)) %>%
-#  select(drug1, drug2, took_both, took_drug1_only, took_drug2_only, took_neither) %>%
-#  mutate(mcc=mcc(took_both, took_drug1_only, took_drug2_only, took_neither))
-```
-
-```{r top_user_overlap}
-# The idea here was to ask which users are the "top" users for that drug.
-# Then I would overlap them in the same way that I did for any-users above.
-
-pad_to = function(x, n, with=0) {
-  if (length(x) > n) stop('input longer than pad to length')
-  c(x, rep(0, n - length(x)))
-}
-
-top_bracket_limit = function(n, fraction) {
-  n[cume_dist(n) > (1 - fraction)] %>% min
-}
-
-#drug_upper_brackets = pde %>%
-#  inner_join(pde %>% group_by(bene_id) %>% summarize(n_claims=sum(n_claims)) %>% mutate(antibiotic='all'), by='bene_id') %>%
-#  group_by(antibiotic) %>%
-#  summarize(limit=top_bracket_limit(pad_to(n_claims, n_bene), 0.001)) %>%
-#  filter(limit > 0)
-```
+The results don't seem like they change that much.

### 2017-03-28 chronic-conditions/chronic-condition-drivers.Rmd
added mcc

@@
 ```{r load_data}
-all_denom = read_tsv('../bene_2011.tsv') %>%
+bene = read_tsv('../bene_2011.tsv') %>%
   filter(age >= 65) %>%
@@
   mutate_at(vars(AMI:HYPOTH), function(x) x==3)
-usage = read_tsv('../pde_2011.tsv') %>%
+pde = read_tsv('../pde_2011.tsv') %>%
   count(bene_id) %>% rename(n_claims=n)
 
-na_to_0 = function(x) if_else(is.na(x), 0L, x)
-
-usage_summary = denom %>%
+usage = bene %>%
   left_join(chronic_conditions, by='bene_id') %>%
-  left_join(usage, by='bene_id') %>%
-  mutate(n_claims=na_to_0(n_claims))
+  left_join(pde, by='bene_id') %>%
+  replace_na(list(n_claims=0))
 
@@
 
-multivariate_res = usage_summary %>%
+multivariate_res = usage %>%
   select(n_claims, age, is_female, AMI:HYPOTH) %>%
@@
 ```
-
-## Summary
-
-**XXX**
 
@@
 Unfortunately, the negative correlations make everything else harder to see, so here's the same thing but without the male conditions.
+I also leave out Alzheimer's and dementia, because it's extremely well-correlated with Alzheimer's and leads to some confusing setup.
 
 ```{r fig.width=7, fig.height=7, cache=FALSE}
-(!(rownames(tetra$rho) %in% c('CNCRPRST', 'HYPERP'))) %>%
+(!(rownames(tetra$rho) %in% c('CNCRPRST', 'HYPERP', 'ALZHDMTA'))) %>%
   { tetra$rho[., .] } %>%
@@
 There is a block of correlations focused around hypertension that includes ischemic heart disease, heart failure, anemia, chronic kidney disease, diabetes, hypertension, hyperlipidemia, and, importantly, COPD.
+
+### MCC correlations
+
+```{r mcc}
+mcc = function(true, pred) {
+  ut = true - mean(true)
+  up = pred - mean(pred)
+
+  co = mean(ut * up)
+  vt = mean(ut ** 2)
+  vp = mean(up ** 2)
+
+  if (vt * vp == 0) {
+    NA
+  } else {
+    co / sqrt(vt * vp)
+  }
+}
+
+mcc_cor = function(x) {
+  x = as.matrix(x)
+  labels = colnames(x)
+  n = length(labels)
+  out = matrix(nrow=n, ncol=n, dimnames=list(labels, labels))
+
+  for (i in 1:n) out[i, i] = 1.0
+
+  for (i in 1:(n-1)) {
+    ui = x[,i] - mean(x[,i])
+    vi = mean(ui ** 2)
+    for (j in (i+1):n) {
+      uj = x[,j] - mean(x[,j])
+      vj = mean(uj ** 2)
+      co = mean(ui * uj)
+      val = co / sqrt(vi * vj)
+      out[i, j] = val
+      out[j, i] = val
+    }
+  }
+  out
+}
+
+mccs = chronic_conditions %>%
+  select(-bene_id) %>%
+  mcc_cor
+```
+
+Here is a heatmap using the square roots of the Matthews correlation coefficients.
+(I used square roots because it emphasizes that there are two clusters.
+Using just raw values is not as nice to look at.) Interestingly, all the correlations are positive.
+
+```{r fig.width=7, fig.height=7, cache=FALSE}
+(!(rownames(mccs) %in% c('CNCRPRST', 'HYPERP', 'ALZHDMTA'))) %>%
+  { mccs[., .] } %>%
+  sqrt %>%
+  gplots::heatmap.2(trace='none', symm=TRUE)
+```
 
@@
 ```{r copd}
-usage_summary %>%
+usage %>%
   group_by(COPD) %>%

### 2017-03-28 chronic-conditions/chronic-condition-drivers.Rmd
negative binomial

@@
 
+### Negative binomial
+
+The claims data are distributed more like a negative binomial than a Poisson (i.e., the variance in counts exceeds the mean).
+Negative binomial regressions give similar results to the Poisson results.
+For example, the COPD coefficient is 0.811 (vs.
+0.806) and the rheumatoid arthritis coefficient is 0.466 (vs.
+0.465).
+
+```{r negative_binomial}
+usage %>%
+  mutate(age=age-65) %>%
+  MASS::glm.nb(n_claims ~ age + is_female + COPD, data=.) %>%
+  broom::tidy %>%
+  select(term, estimate)
+```
+
 # Discussion

### 2017-03-28 chronic-conditions/chronic-condition-drivers.Rmd
added young people

@@
 bene = read_tsv('../bene_2011.tsv') %>%
-  filter(age >= 65) %>%
-  mutate(is_female=sex=='female') %>%
-  select(bene_id, age, is_female)
+  filter(hmo_months==0) %>%
+  #filter(age >= 65) %>%
+  mutate(is_female=sex=='female', is_white=race=='white') %>%
+  select(bene_id, age, is_female, is_white, state)
 chronic_conditions = read_tsv('cc_2011.tsv') %>%
@@
 
+census_regions = read_tsv('../../db/census-regions/census-regions.tsv') %>% 
+  select(state, region)
+
 usage = bene %>%
+  filter(age >= 65) %>%
+  left_join(census_regions, by='state') %>%
   left_join(chronic_conditions, by='bene_id') %>%
@@
 
+### Including region and race
+
+I wanted to know if there was a confound with region or race: is it the case that, say, people of one race are consuming more antibiotics and also just happen to have more chronic conditions?
+
+```{r region_race}
+usage %>%
+  filter(age >= 65) %>%
+  mutate(region=factor(region, levels=c('Northeast', 'West', 'Midwest', 'South'))) %>%
+  glm(n_claims ~ age + is_female + is_white + region + COPD, family='poisson', data=.) %>%
+  broom::tidy() %>%
+  kable
+```
+
+This model treats the Northeast as the baseline region; all other regions have higher consumption.
+Notably, the effect of going to the south is about the same as being white (vs.
+any other race) or being female (vs.
+male).
+
+### Including young people
+
+I'm making the claim that chronic conditions explain consumption among the elderly.
+Maybe the effect of age among the elderly is weak, and we would see an effect if we included some younger people.
+I therefore tried keeping all the beneficiaries to see what would happen.
+
+```{r young_people}
+# remake the usage with young people
+bene %>%
+  left_join(chronic_conditions, by='bene_id') %>%
+  left_join(pde, by='bene_id') %>%
+  replace_na(list(n_claims=0)) %>%
+  glm(n_claims ~ age + is_female + is_white + COPD, family='poisson', data=.) %>%
+  broom::tidy() %>%
+  kable
+```
+
 # Discussion

### 2017-03-30 chronic-conditions/chronic-condition-drivers.Rmd
chronic conditions, aggregate

@@
   glm(multivariate_formula, data=., family='poisson') %>%
-  tidy
+  broom::tidy()
+```
+
+```{r multivariate_negbin}
+multivariate_negbin_res = usage %>%
+  MASS::glm.nb(multivariate_formula, data=.) %>%
+  broom::tidy()
 ```
@@
 # keep the (Intercept), etc. if it didn't match
-
 multivariate_res %>%
@@
   kable
+```
+
+#### Comparison to negative binomial
+
+The negative binomial model gave almost exactly the same results.
+Here I show a plot comparing the coefficients estimates the two models.
+The one point that's deviated from the line is the intercept, which isn't comparable between the two models anyhow.
+Notably, the Poisson model appears to slightly underestimates the coefficients.
+
+```{r compare_poisson_negbin}
+bind_rows(mutate(multivariate_res, method='Poisson'),
+          mutate(multivariate_negbin_res, method='negbin')) %>%
+  mutate(condition=str_replace(term, 'TRUE$', '')) %>%
+  left_join(condition_names, by='condition') %>%
+  mutate(term=if_else(is.na(condition_name), condition, condition_name)) %>%
+  select(term, method, estimate) %>%
+  spread(method, estimate) %>%
+  ggplot(aes(x=Poisson, y=negbin)) +
+  geom_point() +
+  geom_abline() +
+  xlim(-0.2, 0.6) +
+  xlab('Poisson multivariate model coefficient') + ylab('negative binomial multivariate model coefficient') +
+  theme_minimal()
 ```
@@
 
-### Negative binomial
-
-The claims data are distributed more like a negative binomial than a Poisson (i.e., the variance in counts exceeds the mean).
-Negative binomial regressions give similar results to the Poisson results.
-For example, the COPD coefficient is 0.811 (vs.
-0.806) and the rheumatoid arthritis coefficient is 0.466 (vs.
-0.465).
-
-```{r negative_binomial}
-usage %>%
-  mutate(age=age-65) %>%
-  MASS::glm.nb(n_claims ~ age + is_female + COPD, data=.) %>%
-  broom::tidy %>%
-  select(term, estimate)
-```
-
 ### Including region and race

### 2017-03-30 chronic-conditions/multi-year/multiyear-chronic.Rmd
chronic conditions, aggregate

---
title: ""
author: "Scott Olesen"
date: ""
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, 
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
```

```{r load_bene}
parse_multi_files = function(parse_file_f, fns, labels=NULL) {
  if (is.null(labels)) labels = fns
  stopifnot(length(fns) == length(labels))

  if (length(fns) == 1) return(mutate(parse_file_f(fns[1]), label=labels[1]))
  stopifnot(length(fns) > 1)

  rest = parse_multi_files(parse_file_f, fns[-1], labels=labels[-1])
  this = parse_file_f(fns[1]) %>% mutate(label=labels[1])

  # find the common bene IDs
  common_bene = intersect(rest$bene_id, this$bene_id)

  bind_rows(this %>% filter(bene_id %in% common_bene),
            rest %>% filter(bene_id %in% common_bene))
}

# Find only the good beneficiaries
parse_bene = function(fn) {
  read_tsv(fn) %>%
  filter(hmo_months==0, age >= 65) %>%
  mutate(is_female=sex=='female', is_white=race=='white') %>%
  select(bene_id, age, is_female, is_white, state)
}

census_regions = read_tsv('../../../db/census-regions/census-regions.tsv') %>%
  select(state, region)

# convenience function: length(unique(.)) ==
lueq = function(x, l) length(unique(x)) == l

years = 2011:2014
bene_fns = sapply(years, function(y) sprintf('../../bene_%i.tsv', y))
bene = parse_multi_files(parse_bene, bene_fns, labels=years) %>%
  left_join(census_regions, by='state') %>% # get census region
  # keep only beneficiaries that change age every year but stay the same race/sex/region
  group_by(bene_id) %>%
  filter(lueq(age, 4), lueq(is_female, 1), lueq(is_white, 1), lueq(region, 1), lueq(label, 4)) %>%
  ungroup() %>%
  filter(label==years[1]) %>% # just take the first entry for each bene
  select(-label, -state) # drop variables that are not representative of the whole
```

```{r load_cc}
parse_cc = function(fn) {
  read_tsv(fn) %>%
  rename(bene_id=BENE_ID) %>%
  select(-STRCTFLG) %>%
  mutate_at(vars(AMI:HYPOTH), function(x) x==3)
}

cc_fns = sapply(years, function(y) sprintf('../../../data/cc_%i.tsv', y))
cc = parse_multi_files(parse_cc, cc_fns, labels=years)
```

How often did beneficiaries change their chronic condition status?

```{r}
# I want to take the parallel maximum over the conditions columns
# This is not very elegant. I make a list of dummy variables x1, x2, &c.
# and then make a list that points from those dummy strings to my actual
# variables. Then I use interp to run the pmax on the actual data frame.
dummy_vars = sapply(1:length(conditions), function(i) sprintf('x%i', i)) # c('x1', 'x2')
dummy_vars_s = paste0(dummy_vars, collapse=',') # 'x1,x2'
dummy_formula = as.formula(paste0(c('~pmax(', dummy_vars_s, ')'), collapse='')) # ~pmax(x1,x2)
dummy_vars_l = lapply(conditions, as.name) %>% setNames(dummy_vars) # list(x1=AMI, x2=COPD)

# Maybe later we can worry about whether people who changed are not are generally more or less sick
dummy_sum_formula = as.formula(paste0(c('~sum(', dummy_vars_s, ')'), collapse=''))

cc_diff = cc %>%
  group_by(bene_id) %>%
  summarize_all(function(x) length(unique(x))) %>%
  mutate_(max_len_unique=interp(dummy_formula, .values=dummy_vars_l))

cc_diff %>%
  count(max_len_unique) %>%
  rename(status=max_len_unique, `number of beneficiaries`=n) %>%
  mutate(status=case_when(.$status == 1 ~ 'no conditions changed',
                          .$status == 2 ~ 'at least one condition changed',
                          TRUE ~ 'other')) %>%
  kable
```

We have 3 million people whose status for all chronic conditions did not change over the course of the four years, so let's start with just them.

```{r}
no_cc_change_benes = cc_diff %>% filter(max_len_unique==1) %$% bene_id
cc_nochange = cc %>%
  filter(bene_id %in% no_cc_change_benes, label==2011) %>%
  select(-label)

bene_nochange = inner_join(bene, cc_nochange, by='bene_id')
```

There are `r nrow(bene_nochange)` beneficiaries that pass both the demographic and chronic condition filters.

```{r}
parse_pde = function(fn) read_tsv(fn) %>% count(bene_id) %>% rename(n_claims=n)
pde_fns = sapply(years, function(y) sprintf('../../pde_%i.tsv', y))
pde = parse_multi_files(parse_pde, pde_fns, labels=years)
```

We can either run the mode using the claims from each year (treating each person as if they were independent) or by summing up their claims and looking at that.

```{r}
aggregate_pde = pde %>% group_by(bene_id) %>% summarize(n_claims=sum(n_claims))
```

```{r}
year_model = function(y) {
  left_join(bene_nochange, filter(pde, label==y), by='bene_id') %>%
  replace_na(n_claims=0) %>%
  glm(n_claims ~ age + is_female + COPD, family='poisson', data=.)
}

aggregate_model = left_join(bene_nochange, aggregate_pde, by='bene_id') %>%
  replace_na(n_claims=0) %>%
  glm(n_claims ~ age + is_female + ANEMIA, family='poisson', data=.)

yearly_models = lapply(years, year_model)
```

### 2017-03-31 chronic-conditions/chronic-condition-drivers.Rmd
added all years multivar

@@
 
+## On all years
+
+It seems like the multivariate model is the easiest one to understand and interpret. I therefore ran this one analysis on all the years.
+
+```{r}
+year_model = function(y) {
+  bene_fn = sprintf('../bene_%i.tsv', y)
+  pde_fn = sprintf('../pde_%i.tsv', y)
+  cc_fn = sprintf('../../data/cc_%i.tsv', y)
+
+  bene = read_tsv(bene_fn) %>%
+    filter(hmo_months==0, age >= 65) %>%
+    mutate(is_female=sex=='female', is_white=race=='white') %>%
+    select(bene_id, age, is_female, is_white, state)
+  chronic_conditions = read_tsv(cc_fn) %>%
+    rename(bene_id=BENE_ID) %>%
+    select(-STRCTFLG) %>%
+    mutate_at(vars(AMI:HYPOTH), function(x) x==3)
+  pde = read_tsv(pde_fn) %>%
+    count(bene_id) %>% rename(n_claims=n)
+
+  usage = bene %>%
+    left_join(census_regions, by='state') %>%
+    left_join(chronic_conditions, by='bene_id') %>%
+    left_join(pde, by='bene_id') %>%
+    replace_na(list(n_claims=0))
+
+  multivariate_res = usage %>%
+    glm(multivariate_formula, data=., family='poisson') %>%
+    broom::tidy() %>%
+    mutate(label=y)
+
+  multivariate_res
+}
+
+yearly_models = lapply(2011:2014, year_model) %>%
+  bind_rows
+```
+
+```{r}
+#yearly_models %>%
+#  select(term, label, estimate) %>%
+#  spread(label, estimate) %>%
+#  arrange(desc(`2011`)) %>%
+#  kable
+```
+
+```{r}
+term_levels = yearly_models %>%
+  filter(label==2011) %>%
+  arrange(desc(estimate)) %$%
+  rev(term)
+
+yearly_models %>%
+  mutate(term=factor(term, levels=term_levels, ordered=TRUE)) %>%
+  ggplot(aes(x=term, y=estimate, color=factor(label))) +
+  geom_point() +
+  coord_flip() +
+  theme_minimal()
+```
+
 # Discussion

### 2017-03-31 polish-cc.R
polish cc

@@ -1,0 +1,26 @@
+#!/usr/bin/env Rscript
+
+# Polish the chronic condition data, relying on the bene data already being done
+# swo> Merge this into the other polishing script
+
+polish_cc = function(x) {
+  rename(x, bene_id=BENE_ID) %>%
+    select(-STRCTFLG) %>%
+    mutate_at(vars(AMI:HYPOTH), function(x) x==3)
+}
+
+polish_year = function(y) {
+  cc_fn = sprintf('../data/cc_%i.tsv', y)
+  bene_fn = sprintf('bene_%i.tsv', y)
+  cc_out_fn = sprintf('cc_%i.tsv', y)
+
+  bene_ids = read_tsv(bene_fn) %$% bene_id
+
+  cc = read_tsv(cc_fn) %>%
+    polish_cc() %>%
+    filter(bene_id %in% bene_ids)
+
+  write_tsv(cc, cc_out_fn)
+}
+
+for (year in 2011:2014) polish_year(year)

### 2017-03-31 ms1/ms1.Rmd
ms1 initial

--- 2017-03-16 hicks-summary/hicks-summary.Rmd
@@
 ---
-title: "Summary of antibiotic consumption"
+title: "Data for ms. 1"
 author: "Scott Olesen"
-date: "2/13/2017"
 output: html_document
@@
 ```{r global_options, include=FALSE, cache=FALSE}
-knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='fig/',
+knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='fig/', dev=c('pdf', 'png'),
                       echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
-library(knitr)
+import::from(knitr, kable)
+import::from(lazyeval, interp)
+import::from(broom, tidy)
+import::from(stringr, str_replace)
 library(assertive)
@@
 }
+
+multicol_sum = function(x, new_name, target_columns) {
+  dummy_vars = sapply(1:length(target_columns), function(i) sprintf('x%i', i))
+  dummy_vars_string = paste0(dummy_vars, collapse=',')
+  dummy_vars_list = lapply(target_columns, as.name) %>% setNames(dummy_vars)
+  f = as.formula(paste0(c('~sum(', dummy_vars_string, ')'), collapse=''))
+
+  rowwise(x) %>%
+    mutate_(new_attribute=interp(f, .values=dummy_vars_list)) %>%
+    rename_(.dots=setNames('new_attribute', new_name)) %>%
+    ungroup()
+}
+
+conditions = c("AMI", "ALZH", "ALZHDMTA", "ATRIALFB", "CATARACT", "CHRNKIDN", "COPD", "CHF", "DIABETES", "GLAUCOMA", "HIPFRAC", "ISCHMCHT", "DEPRESSN", "OSTEOPRS", "RA_OA", "STRKETIA", "CNCRBRST", "CNCRCLRC", "CNCRPRST", "CNCRLUNG", "CNCRENDM", "ANEMIA", "ASTHMA", "HYPERL", "HYPERP", "HYPERT", "HYPOTH")
+
+multivariate_formula = paste(c('n_claims ~ age + is_female + ', conditions), collapse=' + ') %>% formula
+
+all_condition_model = function(usage) {
+  glm(multivariate_formula, data=usage, family='poisson') %>% tidy()
+}
 ```
@@
   select(state, region)
+condition_names = read_tsv('../chronic-conditions/names.txt',
+                           col_names=c('condition', 'condition_name'))
 ```
@@
   bene = read_tsv(sprintf('../bene_%i.tsv', this_year)) %>%
-    filter(age >= 65, hmo_months==12) %>%
+    filter(age >= 65, hmo_months==0) %>%
+    mutate(is_female=sex=='female', is_white=race=='white') %>%
     left_join(regions, by='state')
@@
 
-  # claims by state (broken into sextiles)
-  claims_by_state = summarize_by(bene, pde, 'state') %>%
-    mutate(sextile=ntile(cpkp, 6), year=this_year)
+  # characteristics by consumption
+  cc = read_tsv(sprintf('../../data/cc_%i.tsv', this_year))
+
+  usage = pde %>% count(bene_id) %>% rename(n_claims=n) %>%
+    right_join(bene, by='bene_id') %>%
+    replace_na(n_claims=0) %>%
+    left_join(cc, by='bene_id') %>%
+    multicol_sum('n_cc', conditions)
+
+  claims_group_f = function(x) bounds_group(x, c(0, 1, 5, 10, 20, Inf))
+  chars_by_consumption = usage %>%
+    mutate(claims_group=claims_group_f(n_claims)) %>%
+    group_by(claims_group) %>%
+    summarize(`no. beneficiaries`=n(),
+              `mean age`=mean(age),
+              `% female`=mean(is_female),
+              `mean no. chronic conditions`=mean(n_cc)) %>%
+    mutate(year=this_year)
+
+  condition_model = all_condition_model(usage) %>%
+    mutate(year=this_year)
 
@@
        claims_by_region=claims_by_region,
-       claims_by_state=claims_by_state)
+       chars_by_consumption=chars_by_consumption,
+       condition_model=condition_model)
 }
 
-data_2011 = analyze(2011)
-data_2012 = analyze(2012)
-data_2013 = analyze(2013)
-data_2014 = analyze(2014)
+years = 2011:2014
+results = lapply(years, analyze)
+names(results) = as.character(years)
 ```
@@
 ```{r arrange_data}
-data = list("2011"=data_2011,
-            "2012"=data_2012,
-            "2013"=data_2013,
-            "2014"=data_2014)
-
-data_by_field = function(field) lapply(data, function(x) x[[field]]) %>% bind_rows()
+data_by_field = function(field) lapply(results, function(x) x[[field]]) %>% bind_rows()
 
@@
 
-state_sextiles_2011 = data_by_field('claims_by_state') %>%
-  filter(!is.na(state), state != 'NA', year==2011) %>%
-  select(sextile, state) %>%
-  arrange(desc(sextile), state)
+chars_by_consumption = data_by_field('chars_by_consumption') %>%
+  arrange(desc(y), desc(claims_group)) %>%
+  kable
+
+condition_model = data_by_field('condition_model') %>%
+  mutate(condition=str_replace(term, 'TRUE$', '')) %>%
+  left_join(condition_names, by='condition') %>%
+  mutate(term=if_else(is.na(condition_name), condition, condition_name)) %>%
+  select(term, estimate, year) %>%
+  spread(year, estimate) %>%
+  arrange(desc(`2011`)) %>%
+  kable
 ```
 
-**To do**: Rework conclusions, include references.
-
-This summary is modeled after @hicks-us-2015.
-That study, which used the IMS Health Xponent database to infer the number of prescriptions per capita, reported higher prescription rates than what we observe here.
-In that paper, the number of claims per 1,000 people year (CKPY) is approximately 1,000 for people over 65.
-In this data, we find that number to be between 600 (in 2011) and 750 (in 2012).
-Our value is more similar to a report that used the National Ambulatory Care Survey (NAMCS) and National Hospital Ambulatory Care Survey (NHAMCS) rather than the Xponent data.
-That report estimated about 500 PKPY [@flemingdutra-prevalence-2016].
-A third report, which used the Xponent database, reported that prescriptions fell from about 1,000 PKPY in 1996 to about 600 PKPY in 2003 [@hicks-outpatient-2011].
-
-# Summary
-
-This data set is a 20% sample of Part D beneficiaries, which I have further filtered by age (only those at least 65 years old are included), for HMO membership (beneficiaries who were part of Medicare Advantage for any part of a year were excluded from that year), and for data quality (members with incomplete zipcodes were excluded).
+# Summary of data
 
@@
 
-There is a slight jump between the number of antibiotic claims in 2012 and 2013.
-This discontinuity appears between the December 2012 and January 2013, it is present among all PDE data (not just antibiotics), and it appears to be a real feature of the data.
+# Table 1: Patterns of consumption
 
-# Antibiotic consumption by class
+All consumption rates are reported as claims per thousand people per year (CPKP).
 
-Our data for antibiotic class roughly correspond to those reported in @hicks-us-2015, when allowing for the fact that their results have about 40% higher consumption overall.
-The striking difference is a in consumption of quinolones: in our 2011 data, quinolones are the most-prescribed class and account for 127 PKPY.
-In their data, quinolones are the fourth-most-prescribed class and account for 89 PKPY (adjusted to approximately $90 \times 0.6 = 54$).
-
-Across years, the consumption of drugs seems to follow similar patterns (excepting that consumption in 2012 was about 20% higher than in 2011 and consumption in 2013-2014 was about 10% higher).
-
-All consumption rates are reported as PKPY.
+## Claims by class
 
@@
 
-# Most-prescribed antibiotics
-
-Our data for the most prescribed individual antibiotics also generally accords with @hicks-us-2015 (again, allowing for the 40% discrepancy).
-The major difference is that ciprofloxacin consumption is greater in our data and amoxicillin (and augmentin) consumption are greater in their data.
-The reduced consumption of amoxicillin in our data could be explained by the fact that our population is older.
-
-Allowing for the overall differences in consumption, consumption of azithromycin appears to have declined during 2011 to 2014, while the consumption of levofloxacin increased.
+## Most-prescribed antibiotics
 
@@
 
-# Consumption by sex
-
-Our result, that the consumption of antibiotics by women and men is in a 3:2 ratio, agrees with the result in @hicks-us-2015. The ratio also remained steady across years in our dataset.
+## Consumption by sex
 
@@
 
-# Consumption by region
-
-In @hicks-us-2015, stratifying consumption in 2011 by census region led to one major divide: the South, West, and Northeast consumed more antibiotics than the West (931, 897, and 848 vs.
-647 PKPY).
-In our results, the major divide is between the South and the Midwest, West, and Northeast (641 vs.
-583, 563, and 533 PKPY).
-
-In our data, 2012 appears exceptional: consumption in the Midwest and West was relatively high, and consumption in the Northeast was relatively low.
-Results from 2013 and 2014 accord more closely with the results from @hicks-us-2015, in which the West is the region with the lowest consumption.
+## Consumption by region
 
@@
 
-# Consumption by age
-
-When stratifying by age, @hicks-us-2015 places all people 65 or older into one stratum.
-Here I stratify within this population.
-Excepting the beneficiaries under 65, older beneficiaries tend to consume more antibiotics.
-The beneficiaries under 65 qualify for Medicare because they are disabled, have end stage renal disease, or ALS, which might contribute to their increased consumption of antibiotics.
-However, this trend across age groups does not hold in 2012.
+## Consumption by age
 
@@
 
-# Other analysis in Hicks *et al.*
+# Table 2: Beneficiary characteristics by consumption
 
-In Hicks, they also do:
+`r chars_by_consumption`
 
-- Logistic regression on...
-    - dependent: county in top quartile for prescription rate
-    - independent:
-        - prescribers per capita above median,
-        - proportion of people obese above median,
-        - proportion of people <2 years old above median,
-        - proportion of people black above median,
-        - per capita income in top 2/3s,
-        - proportion with college education in top 2/3s,
-        - proportion female in top 3/4s
-- Univariate regressions of each independent against the dependent
-- Multivariate regression somehow using adjusted ratios with confounders
+# Table 3: Chronic conditions predict consumption
+
+`r condition_model`

### 2017-03-31 ms1/ms1.Rmd
fixed analysis

@@
 knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='fig/', dev=c('pdf', 'png'),
-                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
+                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE,
+                      error=FALSE)
 import::from(knitr, kable)
@@
 
-```{r setup}
+```{r setup, cache=FALSE}
 # Group values in x into ranges determined by upper_limits.
@@
   # characteristics by consumption
-  cc = read_tsv(sprintf('../../data/cc_%i.tsv', this_year))
+  cc = read_tsv(sprintf('../cc_%i.tsv', this_year))
 
@@
 
-  claims_group_f = function(x) bounds_group(x, c(0, 1, 5, 10, 20, Inf))
+  claims_group_f = function(x) bounds_group(x, c(1, 2, 5, 10, 20, Inf))
   chars_by_consumption = usage %>%
@@
               `mean age`=mean(age),
-              `% female`=mean(is_female),
+              `% female`=mean(is_female) * 100,
               `mean no. chronic conditions`=mean(n_cc)) %>%
@@
 chars_by_consumption = data_by_field('chars_by_consumption') %>%
-  arrange(desc(y), desc(claims_group)) %>%
+  arrange(desc(year), desc(claims_group)) %>%
   kable

### 2017-04-04 dosage-form/dosage-form.Rmd
streamlined dosage

@@
 
-```{r load_data, message=FALSE}
-dat = 2011:2014 %>%
-  lapply(function(y) read_tsv(sprintf('../../data/pde_%i.tsv', y)) %>%
+```{r global_options, include=FALSE}
+knitr::opts_chunk$set(fig.width=7, fig.height=4,
+                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
+```
+
+```{r load_formulary, message=FALSE}
+formulary = 2011:2014 %>%
+  lapply(function(y) read_tsv(sprintf('../../data/formulary/formulary_%i.tsv', y)) %>%
            select(brand_name=BN, generic_name=GNN, dosage_form_code=GCDF, dosage_form_desc=GCDF_DESC) %>%
@@
 
+I no longer have the raw PDE data, so I can't use that!
+```{r load_pde, message=FALSE}
+# Start with just one year
+#pde = 2011:2012 %>%
+  #lapply(function(y) read_tsv(sprintf('../pde_%i.tsv', y)) %>% mutate(year=y)) %>%
+  #bind_rows()
+```
+
 Most codes have a unique descriptions.
 ```{r code_desc}
-filter_repeated_block = function(x) {
-  if (nrow(x) == 1) {
-    data_frame()
-  } else {
-    x
-  }
-}
-
-unique_combos = dat %>%
+unique_combos = formulary %>%
   select(dosage_form_code, dosage_form_desc) %>%
@@
   group_by(dosage_form_code) %>%
-  do(filter_repeated_block(.)) %>%
+  filter(length(dosage_form_desc) > 1) %>%
   arrange(dosage_form_code, dosage_form_desc) %>%
@@
 ```{r codes_by_year}
-canonical_codes = unique_combos %>%
-  group_by(dosage_form_code) %>%
-  do(head(., 1))
-
-dat %>%
-  select(dosage_form_code, year) %>%
-  group_by(year) %>%
-  count(dosage_form_code) %>%
-  spread(year, n) %>%
-  left_join(canonical_codes, by='dosage_form_code') %>%
-  arrange(desc(`2011`)) %>%
-  kable
+#canonical_codes = unique_combos %>%
+#  group_by(dosage_form_code) %>%
+#  do(head(., 1))
+#
+#dat %>%
+#  select(dosage_form_code, year) %>%
+#  group_by(year) %>%
+#  count(dosage_form_code) %>%
+#  spread(year, n) %>%
+#  left_join(canonical_codes, by='dosage_form_code') %>%
+#  arrange(desc(`2011`)) %>%
+#  kable
 ```
@@
 ```{r drugs_by_code, message=FALSE}
-
-drugs_by_code = dat %>%
-  filter(year == 2011) %>%
-  select(generic_name, brand_name, dosage_form_code) %>%
-  group_by(dosage_form_code) %>%
-  count(generic_name) %>%
-  group_by(dosage_form_code) %>%
-  do(head(., 3))
+#drugs_by_code = dat %>%
+#  filter(year == 2011) %>%
+#  select(generic_name, brand_name, dosage_form_code) %>%
+#  group_by(dosage_form_code) %>%
+#  count(generic_name) %>%
+#  group_by(dosage_form_code) %>%
+#  do(head(., 3))
 
@@
 
-drugs_by_code %>%
-  mutate(keep_code=dosage_form_code %in% codes_to_keep) %>%
-  kable
+#drugs_by_code %>%
+#  mutate(keep_code=dosage_form_code %in% codes_to_keep) %>%
+#  kable
 ```
+
+For posterity, I export a list of the code/values that I kept.
+```{r save_codes}
+unique_combos %>%
+  filter(dosage_form_code %in% codes_to_keep) %>%
+  arrange(dosage_form_code) %>%
+  write_tsv('code_forms.tsv')
+```

### 2017-04-04 ms1/ms1.Rmd
new table

@@
 
-conditions = c("AMI", "ALZH", "ALZHDMTA", "ATRIALFB", "CATARACT", "CHRNKIDN", "COPD", "CHF", "DIABETES", "GLAUCOMA", "HIPFRAC", "ISCHMCHT", "DEPRESSN", "OSTEOPRS", "RA_OA", "STRKETIA", "CNCRBRST", "CNCRCLRC", "CNCRPRST", "CNCRLUNG", "CNCRENDM", "ANEMIA", "ASTHMA", "HYPERL", "HYPERP", "HYPERT", "HYPOTH")
+conditions = c("AMI", "ALZHDMTA", "ATRIALFB", "CATARACT", "CHRNKIDN", "COPD", "CHF", "DIABETES", "GLAUCOMA", "HIPFRAC", "ISCHMCHT", "DEPRESSN", "OSTEOPRS", "RA_OA", "STRKETIA", "CNCRBRST", "CNCRCLRC", "CNCRPRST", "CNCRLUNG", "CNCRENDM", "ANEMIA", "ASTHMA", "HYPERL", "HYPERP", "HYPERT", "HYPOTH")
 
-multivariate_formula = paste(c('n_claims ~ age + is_female + ', conditions), collapse=' + ') %>% formula
+multivariate_formula = paste(c('n_claims ~ age_over_65 + is_female + ', conditions), collapse=' + ') %>% formula
 
 all_condition_model = function(usage) {
-  glm(multivariate_formula, data=usage, family='poisson') %>% tidy()
+  usage %>%
+    mutate(age_over_65 = age - 65) %>%
+    glm(multivariate_formula, data=., family='poisson') %>%
+    tidy()
 }
@@
   bene = read_tsv(sprintf('../bene_%i.tsv', this_year)) %>%
-    filter(age >= 65, hmo_months==0) %>%
+    filter(age >= 65, age < 116, hmo_months==0) %>%
     mutate(is_female=sex=='female', is_white=race=='white') %>%
@@
 
+  # keep only PDEs for known beneficiaries (semi join)
+  # convert all amox/clav to just amox
   pde = read_tsv(sprintf('../pde_%i.tsv', this_year)) %>%
-    semi_join(bene, by='bene_id') # keep only PDEs for known beneficiaries
+    semi_join(bene, by='bene_id') %>%
+    mutate(antibiotic=if_else(antibiotic=='amoxicillin/clavulanate', 'amoxicillin', antibiotic))
 
@@
   # not just by raw age)
-  age_group_f = function(x) bounds_group(x, c(65, 75, 85, Inf))
+  age_group_f = function(x) bounds_group(x, c(65, 70, 75, 80, 85, 90, 95, 116))
   claims_by_age = summarize_by(
@@
   # characteristics by consumption
-  cc = read_tsv(sprintf('../cc_%i.tsv', this_year))
+  cc = read_tsv(sprintf('../cc_%i.tsv', this_year)) %>%
+    select(-ALZH)
 
@@
     right_join(bene, by='bene_id') %>%
-    replace_na(n_claims=0) %>%
+    replace_na(list(n_claims=0)) %>%
     left_join(cc, by='bene_id') %>%
@@
 `r condition_model`
+
+## Table 3 caption: model coefficient p-values
+
+Only $p$-values above (less significant than) $10^{-50}$ are shown. All the remaining values are more significant than the ones shown here.
+
+```{r p_values}
+data_by_field('condition_model') %>%
+  select(term, year, p.value) %>%
+  mutate(adj.p=p.adjust(p.value, method='bonferroni'),
+         log10.adj.p=log10(adj.p)) %>%
+  filter(adj.p > 10^(-50)) %>%
+  arrange(desc(log10.adj.p)) %>%
+  kable
+```

### 2017-04-04 ms1/supp1.Rmd
supplement

---
title: "Data for supplement to ms. 1"
author: "Scott Olesen"
output: html_document
---

```{r global_options, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='fig/', dev=c('png', 'pdf'),
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE,
                      error=FALSE)
import::from(knitr, kable)
import::from(lazyeval, interp)
import::from(broom, tidy)
import::from(stringr, str_replace)
import::from(doMC, registerDoMC)
import::from(gplots, heatmap.2)
library(glmnet)
library(assertive)
```

```{r load_db, cache=FALSE}
regions = read_tsv('../../db/census-regions/census-regions.tsv') %>%
  select(state, region)
condition_names = read_tsv('../chronic-conditions/names.txt',
                           col_names=c('condition', 'condition_name'))

conditions = c("AMI", "ALZHDMTA", "ATRIALFB", "CATARACT", "CHRNKIDN", "COPD", "CHF", "DIABETES", "GLAUCOMA", "HIPFRAC", "ISCHMCHT", "DEPRESSN", "OSTEOPRS", "RA_OA", "STRKETIA", "CNCRBRST", "CNCRCLRC", "CNCRPRST", "CNCRLUNG", "CNCRENDM", "ANEMIA", "ASTHMA", "HYPERL", "HYPERP", "HYPERT", "HYPOTH")
```

```{r mcc, cache=FALSE}
mcc = function(true, pred) {
  ut = true - mean(true)
  up = pred - mean(pred)

  co = mean(ut * up)
  vt = mean(ut ** 2)
  vp = mean(up ** 2)

  if (vt * vp == 0) {
    NA
  } else {
    co / sqrt(vt * vp)
  }
}

mcc_cor = function(x) {
  x = as.matrix(x)
  labels = colnames(x)
  n = length(labels)
  out = matrix(nrow=n, ncol=n, dimnames=list(labels, labels))

  for (i in 1:n) out[i, i] = 1.0

  for (i in 1:(n-1)) {
    ui = x[,i] - mean(x[,i])
    vi = mean(ui ** 2)
    for (j in (i+1):n) {
      uj = x[,j] - mean(x[,j])
      vj = mean(uj ** 2)
      co = mean(ui * uj)
      val = co / sqrt(vi * vj)
      out[i, j] = val
      out[j, i] = val
    }
  }
  out
}
```

```{r load_data}
this_year = 2011

bene = read_tsv(sprintf('../bene_%i.tsv', this_year)) %>%
  filter(age >= 65, age < 116, hmo_months==0) %>%
  mutate(is_female=sex=='female', is_white=race=='white') %>%
  left_join(regions, by='state')

pde = read_tsv(sprintf('../pde_%i.tsv', this_year)) %>%
  semi_join(bene, by='bene_id') # keep only PDEs for known beneficiaries

cc = read_tsv(sprintf('../cc_%i.tsv', this_year)) %>%
  select(-ALZH)

usage = pde %>% count(bene_id) %>% rename(n_claims=n) %>%
  right_join(bene, by='bene_id') %>%
  replace_na(list(n_claims=0)) %>%
  left_join(cc, by='bene_id') %>%
  mutate(age_over_65 = age - 65)

stopifnot(all(usage$age >= 0))
```

# All-condition negative binomial model

Each point represents a chronic condition. The coordinates are corresponding coefficients in the two models. The formula for the model is claims against age and sex.

```{r all_condition_negbin}
all_condition_formula_start = 'n_claims ~ age_over_65 + is_female + '
all_condition_formula = formula(paste(c(all_condition_formula_start, conditions), collapse=' + '))

all_condition_negbin_model = usage %>%
  MASS::glm.nb(all_condition_formula, data=.) %>%
  tidy()

all_condition_poisson_model = usage %>%
  glm(all_condition_formula, data=., family='poisson') %>%
  tidy()
```

```{r cache=FALSE}
bind_rows(mutate(all_condition_poisson_model, method='Poisson'),
          mutate(all_condition_negbin_model, method='negbin')) %>%
  mutate(condition=str_replace(term, 'TRUE$', '')) %>%
  left_join(condition_names, by='condition') %>%
  mutate(term=if_else(is.na(condition_name), condition, condition_name)) %>%
  select(term, method, estimate) %>%
  spread(method, estimate) %>%
  ggplot(aes(x=Poisson, y=negbin)) +
  geom_abline(col='gray') +
  geom_point() +
  xlim(-0.2, 0.6) + ylim(-0.2, 0.6) +
  xlab('Poisson multivariate model coefficient') +
  ylab('negative binomial multivariate model coefficient') +
  coord_fixed() +
  theme_minimal()
```

# Single-condition Poisson models

```{r single_condition}
summarize_condition = function(x, condition) {
rename_(x, .dots=setNames(condition, 'has_condition')) %>%
  glm(n_claims ~ age_over_65 + is_female + has_condition, family='poisson', data=.) %>%
  tidy() %>%
  mutate(condition=condition)
}

single_condition_models = lapply(conditions, function(cond) summarize_condition(usage, cond)) %>%
  bind_rows()
```

```{r cache=FALSE}
single_condition_models %>%
  left_join(condition_names, by='condition') %>%
  select(-condition) %>%
  rename(condition=condition_name) %>%
  filter(term=='has_conditionTRUE') %>%
  select(condition, estimate, p.value) %>%
  mutate(adj.p.value=p.adjust(p.value, method='bonferroni')) %>%
  arrange(desc(estimate)) %>%
  kable
```

# All-condition Poisson lasso model

I also performed a multivariate lasso regression with the Poisson model, keeping all the chronic conditions as well as age and sex (coded as "is female") as explanatory variables.
I started with 10-fold cross-validation to pick the lasso $\lambda$.

```{r all_condition_lasso}
registerDoMC(cores=3)

lasso_x = usage %>%
  select(age_over_65, is_female, AMI:HYPOTH) %>%
  as.matrix
lasso_y = usage$n_claims

lasso_cv = cv.glmnet(lasso_x, lasso_y, parallel=TRUE, family='poisson')
```

```{r cache=FALSE}
plot(lasso_cv)
```

Although the error continues to fall with decreasing $\lambda$, it seems like there are diminishing returns after the first 8 or 10 variables are included.

```{r cache=FALSE}
lasso_df = lasso_cv$glmnet.fit %>%
  tidy %>%
  filter(estimate > 0) %>%
  group_by(step) %>%
  summarize(df=n() - 1)

lasso_cv$glmnet.fit %>%
  tidy %>%
  select(step, lambda, dev.ratio) %>%
  distinct() %>%
  left_join(lasso_df, by='step') %>%
  select(step, df, lambda) %>%
  mutate_at(vars(lambda), function(x) signif(x, digits=2)) %>%
  filter(df < 12) %>%
  kable
```

```{r pick_lambda}
lambda_lasso = 0.22
```

It looks like this corresponds to $\lambda \sim `r lambda_lasso`$. For that value, the coefficients are:

```{r cache=FALSE}
lasso_cv %>%
  coef(s=lambda_lasso) %>%
  tidy %>%
  select(condition=row, coefficient=value) %>%
  left_join(condition_names, by='condition') %>%
  mutate(term=if_else(is.na(condition_name), condition, condition_name)) %>%
  select(term, coefficient) %>%
  arrange(desc(coefficient)) %>%
  kable
```

# Single-condition Poisson models including race and geography for top conditions

```{r race_geography}
top_conditions = single_condition_models %>%
  filter(term=='has_conditionTRUE') %>%
  arrange(desc(estimate)) %$%
  head(condition, 5)

race_geography_model_f = function(condition) {
  form = formula(paste0('n_claims ~ age_over_65 + is_female + is_white + region + ', condition))
  usage %>%
    mutate(region=factor(region, levels=c('Northeast', 'West', 'Midwest', 'South'))) %>%
    glm(form, family='poisson', data=.) %>%
    tidy() %>%
    mutate(condition=condition)
}

race_geography_model = lapply(top_conditions, race_geography_model_f) %>%
  bind_rows()
```

```{r cache=FALSE}
transpose_tibble = function(x) {
  gather(x, 'var_name', 'value', 2:ncol(x)) %>% 
    spread_(names(x)[1], 'value')
}

single_condition_estimates = single_condition_models %>%
  filter(term=='has_conditionTRUE') %>%
  select(condition, single_estimate=estimate)

race_geography_model %>%
  select(condition, term, estimate) %>%
  mutate(term=if_else(term==paste0(condition, 'TRUE'), 'racegeo_estimate', term)) %>%
  spread(term, estimate) %>%
  left_join(single_condition_estimates, by='condition') %>%
  transpose_tibble %>%
  kable
```

# Correlations between chronic conditions

```{r correlations}
correlations = cc %>%
  select(-bene_id) %>%
  mcc_cor
```

Here I show Matthews correlation coefficients for all the conditions. The heatmap uses $\sqrt{c}$, where $c$ is the correlation; the second dendrogram figure uses $\sqrt{1 - |c|}$ as the distance metric.

```{r fig.width=7, fig.height=7, cache=FALSE}
male_conditions = c('CNCRPRST', 'HYPERP')
#excluded_conditions = c()
(!(rownames(correlations) %in% male_conditions)) %>%
  { correlations[., .] } %>%
  #sqrt %>%
  gplots::heatmap.2(trace='none', symm=TRUE)
```

```{r fig.width=7, fig.height=7, cache=FALSE}
female_conditions = c('CNCRENDM', 'CNCRBRST', 'OSTEOPRS')
(!(rownames(correlations) %in% female_conditions)) %>%
  { correlations[., .] } %>%
  sqrt %>%
  gplots::heatmap.2(trace='none', symm=TRUE)
```

```{r}
plot(hclust(as.dist(sqrt(1.0 - abs(correlations)))))
```

### 2017-04-05 ms1/ms1.Rmd
shouldnt lump amox/clav

@@
   pde = read_tsv(sprintf('../pde_%i.tsv', this_year)) %>%
-    semi_join(bene, by='bene_id') %>%
-    mutate(antibiotic=if_else(antibiotic=='amoxicillin/clavulanate', 'amoxicillin', antibiotic))
+    semi_join(bene, by='bene_id') #%>%
+    #mutate(antibiotic=if_else(antibiotic=='amoxicillin/clavulanate', 'amoxicillin', antibiotic))
 

### 2017-04-05 region-consumption/regional-differences.Rmd
regional consumption

---
title: "Regional differences"
author: "Scott Olesen"
date: "2/20/2017"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/',
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
```

# Background

It's known that there are differences in antibiotic consumption between regions in the US.

- Is this true for all antibiotics? Are there drugs that are consumed against general trends? For example, are there drugs that are more consumed in the West than in the South?
- What drives the differences between regions? Previous research shows that demographics are not sufficient to explain these differences. Is it that more people take antibiotics, or that people who take antibiotics take more?

# Methods

# Results

```{r load_data}
usage = read_tsv('../bene_usage_2011.tsv')
```

```{r bene_by_region}
bene_by_region = usage %>%
  select(bene_id, region) %>%
  distinct %>%
  count(region) %>%
  rename(n_bene=n)

bene_by_region %>% kable
```

For azithromycin, I looked at the distribution of number of claims (truncating at 20 claims or 100 days).

```{r azithromycin_overall, cache=TRUE}
azo = usage %>%
  filter(antibiotic=='azithromycin') %>%
  group_by(bene_id) %>%
  summarize(n_claims=sum(n_claims), n_days=sum(n_days))

azo %>%
  ggplot(aes(x=n_claims)) +
  geom_histogram() +
  scale_y_log10() +
  xlab('number of azithromycin claims') + ylab('number of beneficiaries') +
  theme_minimal()

azo$n_claims %>%
  sort(decreasing=TRUE) %>%
  cumsum %>%
  data_frame(cum_n_bene=1:length(.), cum_n_claims=.) %>%
  ggplot(aes(x=cum_n_bene, y=cum_n_claims)) +
  geom_line() +
  xlab('cumulative number of beneficiaries') + ylab('cumulative number of azithromycin claims') +
  theme_minimal()

azo %>%
  ggplot(aes(x=n_days)) +
  geom_histogram() +
  scale_y_log10() +
  xlab('number of azithromycin days') + ylab('number of beneficiaries')
```

```{r tmpsmx_overall, cache=TRUE}
tmpsmx = usage %>%
  filter(antibiotic=='trimethoprim/sulfamethoxazole') %>%
  group_by(bene_id) %>%
  summarize(n_claims=sum(n_claims), n_days=sum(n_days))

tmpsmx %>%
  ggplot(aes(x=n_claims)) +
  geom_histogram() +
  scale_y_log10() +
  xlab('number of TMP-SMX claims') + ylab('number of beneficiaries') +
  theme_minimal()

tmpsmx$n_claims %>%
  sort(decreasing=TRUE) %>%
  cumsum %>%
  data_frame(cum_n_bene=1:length(.), cum_n_claims=.) %>%
  ggplot(aes(x=cum_n_bene, y=cum_n_claims)) +
  geom_line() +
  xlab('cumulative number of beneficiaries') + ylab('cumulative number of TMP-SMX claims') +
  theme_minimal()
```

```{r azithromycin_by_region, cache=TRUE}
az = usage %>%
  filter(antibiotic=='azithromycin') %>%
  group_by(region, bene_id) %>%
  summarize(n_claims=sum(n_claims), n_days=sum(n_days)) %>%
  ungroup %>%
  mutate(region=factor(region, levels=c('South', 'Midwest', 'West', 'Northeast'), ordered=TRUE))

az %>%
  filter(n_claims < 20) %>%
  ggplot(aes(x=factor(n_claims), fill=region)) +
  stat_count(position='dodge') +
  scale_y_log10() +
  theme_minimal() +
  xlab('number of azithromycin claims') + ylab('number of beneficiaries')

round_up_to_nearest = function(x, n) n * ceiling(x / n)
az %>%
  mutate(n_days=round_up_to_nearest(n_days, 5)) %>%
  filter(n_days <= 100) %>%
  ggplot(aes(x=factor(n_days), fill=region)) +
  stat_count(position='dodge') +
  scale_y_log10() +
  theme_minimal() +
  xlab('number of azithromycin days') + ylab('number of beneficiaries') +
  theme_minimal()
```

It looks like all regions have almost exactly the same distribution; the South simply has more people. What is we work per capita?

# Discussion

### 2017-04-05 region-consumption/first.R

@@ -1,0 +1,48 @@
+abx = read_tsv("~/grad/proj/medicare/analysis/bene-abx-2011.tsv")
+census = read_tsv("~/grad/proj/medicare/db/census-regions/census-regions.tsv")
+
+# merge in the census region data
+abx %<>%
+  rename(state_abbreviation=state) %>% 
+  left_join(census %>% select(-state))
+
+counts_table = function(x) {
+  counts = tabulate(x)
+  n = 1:length(counts)
+  data_frame(n=n, counts=counts)
+}
+
+plot_days = function(days) {
+  days %>%
+    counts_table %>% 
+    ggplot(aes(x=n, y=counts)) + geom_point() + scale_y_log10() + xlab("# days") + ylab("# beneficiaries")
+}
+
+plot_drug_region = function(abx, drug, region) {
+  abx %>% 
+    rename_(.dots=setName(drug, 'drug')) %>%
+    filter(region==region) %$%
+    plot_days(drug)
+}
+
+plot_drug_region2 = function(abx, drug, region1, region2) {
+  a = abx %>% 
+    rename_(.dots=setNames(drug, 'drug')) %>% 
+    filter(region==region1) %$%
+    counts_table(drug)
+  
+  b = abx %>% 
+    rename_(.dots=setNames(drug, 'drug')) %>% 
+    filter(region==region2) %$%
+    counts_table(drug)
+  
+  color1 = 'black'
+  color2 = 'blue'
+  title = sprintf('%s consumption in %s (%s) and %s (%s)', drug, region1, color1, region2, color2)
+  ggplot(a, aes(x=n, y=counts)) +
+    geom_point() + geom_smooth(col='black', se=FALSE) +
+    geom_point(dat=b, col='blue') + geom_smooth(dat=b, col='blue', se=FALSE) +
+    scale_y_log10() +
+    xlab('# days') + ylab('# beneficiaries') + ggtitle(title) +
+    theme_minimal()
+}

### 2017-04-05 states/resistance/ms2.Rmd
ms2 initial

--- 2017-03-14 states/resistance/state-report.Rmd
@@
 ---
-title: "Consumption and resistance across states and HRRs"
+title: "Ms. 2: Consumption and resistance across states and HRRs"
 bibliography: refs.bib
@@
 
-```{r global_options, include=FALSE}
+```{r global_options, include=FALSE, cache=FALSE}
 knitr::opts_chunk$set(fig.width=7, fig.height=4, cache=TRUE,
                       echo=FALSE, warning=FALSE, message=FALSE)
+import::from(broom, tidy)
+library(pander)
+library(assertr)
 ```
@@
 ```{r setup}
-library(broom)
-library(pander)
-library(assertr)
-
-# swo> I removed Klebsiella because it's mostly in immune-compromised people
-# and K. oxytoca because it's from neonatal septicemia?
 gram_negs = c('Escherichia coli',
@@
 ```{r load_data}
-hrr = read_tsv('../../../db/hrr/hrr_2011.tsv') %>%
+hrr = read_tsv('../../../db/hrr/hrr.tsv') %>%
+  filter(year==2011) %>%
   select(zipcode, hrr)
@@
   #filter(state %in% known_states)
+```
 
-# State-level consumption data
-state_consumption = read_tsv('../bene-consumption/state_usage_2011.tsv') %>%
-  filter(!is.na(state)) %>%
-  right_join(consumption_groups, by=c('antibiotic'='drug')) %>%
-  group_by(state, drug_group) %>%
-  summarize_at(vars(cpkp, did), sum)
-
-hrr_consumption = read_tsv('../bene-consumption/hrr_usage_2011.tsv') %>%
-  filter(!is.na(hrr)) %>%
+```{r load_consumption_data}
+state_usage = read_tsv('../bene-consumption/state_usage.tsv') %>%
+  filter(year==2011) %>%
   right_join(consumption_groups, by=c('antibiotic'='drug')) %>%
@@
 
-state_dat = left_join(abg, state_consumption, by=c('drug_group', 'state'))
-hrr_dat = left_join(abg, hrr_consumption, by=c('drug_group', 'hrr'))
+hrr_usage = read_tsv('../bene-consumption/hrr_usage.tsv') %>%
+  filter(year==2011) %>%
+  right_join(consumption_groups, by=c('antibiotic'='drug')) %>%
+  group_by(hrr, drug_group) %>%
+  summarize_at(vars(cpkp, did), sum)
+
+state_dat = left_join(abg, state_usage, by=c('drug_group', 'state'))
+hrr_dat = left_join(abg, hrr_usage, by=c('drug_group', 'hrr'))
 ```
@@
 ```{r run_models}
-safe_spearman = function(df, expl, indep) {
-  try(return(cor.test(df[[indep]], df[[expl]], method='spearman') %>% tidy), silent=TRUE)
-  data_frame(estimate=NA, statistic=NA, p.value=NA, method=NA, alternative=NA)
-}
-
-models = function(df, indep, keep=NULL) {
-  lapply(c('cpkp', 'did'), function(expl) {
-    safe_spearman(df, expl, indep) %>%
-      mutate(explanatory=expl)
-  }) %>% bind_rows
-}
-
 pvalue_stars = function(x) {

### 2017-04-05 states/resistance/ms2.Rmd
moving toward ms2 first linear

@@
 
-I used the Medicare Part D data from 2011. Only beneficiaries greater than 65 years old were used. I also used the antibiogram data. I pooled the antiobiograms from all years.
+I used the Medicare Part D data from 2011. Normal filtering was applied to beneficiaries: at least 65 years old, no HMO, 12 months of coverage.
+
+I pooled the antiobiograms from all years.
 
@@
 import::from(broom, tidy)
+import::from(mblm, mblm)
 library(pander)
@@
 
+summarize_abg = function(df, place_name) {
+  group_by_(df, 'bug', 'drug_group', place_name) %>%
+    summarize(mean_percent_nonsusceptible=weighted.mean(percent_nonsusceptible, sqrt(n_isolates)),
+              n_place_antibiograms=n()) %>%
+    mutate(place_weight=sqrt(n_place_antibiograms))
+}
+
+abg_state_dat = summarize_abg(abg, 'state')
+abg_hrr_dat = summarize_abg(abg, 'hrr')
+
 # swo> State beneficiary information could be used as control variables
@@
   right_join(consumption_groups, by=c('antibiotic'='drug')) %>%
-  group_by(hrr, drug_group) %>%
-  summarize_at(vars(cpkp, did), sum)
+  group_by(state, drug_group) %>%
+  summarize_at(vars(cpkp, did), sum) %>%
+  ungroup()
 
@@
   group_by(hrr, drug_group) %>%
-  summarize_at(vars(cpkp, did), sum)
+  summarize_at(vars(cpkp, did), sum) %>%
+  ungroup()
 
-state_dat = left_join(abg, state_usage, by=c('drug_group', 'state'))
-hrr_dat = left_join(abg, hrr_usage, by=c('drug_group', 'hrr'))
+state_dat = left_join(abg_state_dat, state_usage, by=c('drug_group', 'state')) %>%
+  rename(place=state)
+hrr_dat = left_join(abg_hrr_dat, hrr_usage, by=c('drug_group', 'hrr')) %>%
+  rename(place=hrr)
 ```
@@
 
-models_result = function(x, this_place) {
-  x %>%
-    group_by_('bug', 'drug_group', this_place) %>%
-    summarize(mean_nonsus=mean(percent_nonsusceptible, na.rm=TRUE),
-              cpkp=head(cpkp, 1),
-              did=head(did, 1)) %>%
-    group_by(bug, drug_group) %>%
-    do(models(., 'mean_nonsus')) %>%
-    ungroup %>%
-    filter(!is.na(estimate)) %>%
-    mutate(adj.p.value=p.adjust(p.value, method='BH'),
-           sig=pvalue_stars(adj.p.value)) %>%
-    mutate(record_number=1:n(),
-           plot_fn=sprintf('plots/%s_%03i.pdf', this_place, record_number),
-           plot_link=pandoc.link.return(url=plot_fn, text='plot')) %>%
-    select(bug, drug_group,
-           explanatory, correlation=estimate, p.value,
-           adj.p.value, sig,
-           plot_fn, plot_link)
+models = function(df, independent_var, explanatory_vars) {
+  f = paste0(independent_var, ' ~ ', paste0(explanatory_vars, collapse = ' + ')) %>% formula
+  lm(f, data=df) %>% tidy
 }
 
-state_res = models_result(state_dat, 'state')
-hrr_res = models_result(hrr_dat, 'hrr')
+models_result = function(df) {
+  group_by(df, bug, drug_group, place) %>%
+    summarize(mean_percent_nonsusceptible=weighted.mean(mean_percent_nonsusceptible, place_weight),
+              cpkp=unique(cpkp),
+              did=unique(did)) %>%
+    group_by(bug, drug_group) %>%
+    do(models(., 'mean_percent_nonsusceptible', 'cpkp')) %>%
+    ungroup %>%
+    filter(term=='cpkp') %>%
+    mutate(adj.p.value=p.adjust(p.value, method='BH'),
+           sig=pvalue_stars(adj.p.value))
+   # %>%
+    #mutate(record_number=1:n(),
+    #       plot_fn=sprintf('plots/%s_%03i.pdf', this_place, record_number),
+    #       plot_link=pandoc.link.return(url=plot_fn, text='plot')) %>%
+    #select(bug, drug_group,
+    #       explanatory, correlation=estimate, p.value,
+    #       adj.p.value, sig,
+    #       plot_fn, plot_link)
+}
+
+state_res = models_result(state_dat)
+hrr_res = models_result(hrr_dat)
 
 result = bind_rows(
-  state_res %>% mutate(place='state'),
-  hrr_res %>% mutate(place='hrr')
+  state_res %>% mutate(place_type='state'),
+  hrr_res %>% mutate(place_type='hrr')
 )
@@
 
-#### State-level resistance-averaged correlations
+#### State-level, resistance-averaged linear models
 
@@
 result %>%
-  filter(place=='state') %>%
+  filter(place_type=='state') %>%
   mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
   select(bug, drug_group,
-         explanatory, correlation,
-         adj.p.value, sig,
-         plot_link) %>%
+         term, estimate, std.error,
+         adj.p.value, sig
+         #,plot_link
+         ) %>%
   kable
@@
 result %>%
-  filter(place=='state') %>%
-  ggplot(aes(x=correlation, y=adj.p.value)) +
+  filter(place_type=='state') %>%
+  ggplot(aes(x=estimate, y=adj.p.value)) +
   geom_point() +
   geom_hline(yintercept=0.05, col='red') +
-  xlab('correlation coefficient') + ylab('p value (reversed)') +
+  xlab('slope') + ylab('p value (reversed)') +
   scale_y_reverse() +
@@
 
-#### State-level resistance-averaged correlations using linear models
+#### HRR-level, resistance-averaged linear models
 
-```{r}
-safe_lm = function(data, indep, expl) {
-  try(return(lm(sprintf('%s ~ %s', indep, expl), data) %>% tidy), silent=TRUE)
-  data_frame(term=expl, estimate=NA, std.error=NA, statistic=NA, p.value=1.0)
-}
+```{r hrr_average_results_table}
+result %>%
+  filter(place_type=='hrr') %>%
+  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
+  select(bug, drug_group,
+         term, estimate, std.error,
+         adj.p.value, sig
+         #,plot_link
+         ) %>%
+  kable
+```
 
-linear_models = function(df, indep) {
-  lapply(c('cpkp', 'did'), function(expl) {
-    safe_lm(df, indep, expl) %>%
-      filter(term==expl) %>%
-      rename(explanatory=term) %>%
-      verify(nrow(.) == 1)
-  }) %>% bind_rows
-}
+(red line marks $p = 0.05$).
 
-linear_models_result = function(x, this_place) {
-  x %>%
-    group_by_('bug', 'drug_group', this_place) %>%
-    summarize(mean_nonsus=mean(percent_nonsusceptible, na.rm=TRUE),
-              cpkp=head(cpkp, 1),
-              did=head(did, 1)) %>%
-    group_by(bug, drug_group) %>%
-    do(linear_models(., 'mean_nonsus')) %>%
-    ungroup %>%
-    filter(!is.na(estimate)) %>%
-    mutate(adj.p.value=p.adjust(p.value, method='BH'),
-           sig=pvalue_stars(adj.p.value)) %>%
-    select(bug, drug_group,
-           explanatory, slope=estimate, p.value,
-           adj.p.value, sig)
-}
-
-linear_state_res = linear_models_result(state_dat, 'state')
-
-linear_state_res %>% kable
+```{r hrr_average_volcano}
+result %>%
+  filter(place_type=='hrr') %>%
+  ggplot(aes(x=estimate, y=adj.p.value)) +
+  geom_point() +
+  geom_hline(yintercept=0.05, col='red') +
+  xlab('slope') + ylab('p value (reversed)') +
+  scale_y_reverse() +
+  theme_minimal()
 ```
@@
 
-```{r}
-library(mblm)
-
-safe_mblm = function(data, indep, expl) {
-  try(return(mblm(sprintf('%s ~ %s', indep, expl), data) %>% tidy), silent=TRUE)
-  data_frame(term=expl, estimate=NA, std.error=NA, statistic=NA, p.value=1.0)
-}
-
-ts_models = function(df, indep) {
-  lapply(c('cpkp', 'did'), function(expl) {
-    safe_mblm(df, indep, expl) %>%
-      filter(term==expl) %>%
-      rename(explanatory=term) %>%
-      verify(nrow(.) == 1)
-  }) %>% bind_rows
-}
-
-ts_models_result = function(x, this_place) {
-  x %>%
-    group_by_('bug', 'drug_group', this_place) %>%
-    summarize(mean_nonsus=mean(percent_nonsusceptible, na.rm=TRUE),
-              cpkp=head(cpkp, 1),
-              did=head(did, 1)) %>%
-    group_by(bug, drug_group) %>%
-    do(linear_models(., 'mean_nonsus')) %>%
-    ungroup %>%
-    filter(!is.na(estimate)) %>%
-    mutate(adj.p.value=p.adjust(p.value, method='BH'),
-           sig=pvalue_stars(adj.p.value)) %>%
-    select(bug, drug_group,
-           explanatory, slope=estimate, p.value,
-           adj.p.value, sig)
-}
-
-ts_state_res = linear_models_result(state_dat, 'state')
-
-ts_state_res %>% kable
-```
-
-#### HRR-level resistance-averaged
-
-```{r hrr_average_results_table}
-result %>%
-  filter(place=='hrr') %>%
-  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
-  select(bug, drug_group,
-         explanatory, correlation,
-         adj.p.value, sig,
-         plot_link) %>%
-  kable
-```
-
-(red line marks $p = 0.05$)
-
-```{r hrr_average_histogram}
-result %>%
-  filter(place=='hrr') %>%
-  ggplot(aes(x=correlation, y=adj.p.value)) +
-  geom_point() +
-  geom_hline(yintercept=0.05, col='red') +
-  xlab('correlation coefficient') + ylab('p value (reversed)') +
-  scale_y_reverse() +
-  theme_minimal()
-```
-
-### All-antibiogram results
-
-In these data, I don't average the resistance before running the correlation, but I do associate each antibiogram with the same geography-wide consumption.
-
-```{r}
-all_models_result = function(x, this_place, consumption) {
-  x %>%
-    mutate_at(vars(cpkp, did), function(x) if_else(is.na(x), 0, x)) %>%
-    group_by(bug, drug_group) %>%
-    do(models(., 'percent_nonsusceptible')) %>%
-    ungroup %>%
-    filter(!is.na(estimate)) %>%
-    mutate(adj.p.value=p.adjust(p.value, method='BH'),
-           sig=pvalue_stars(adj.p.value)) %>%
-    select(bug, drug_group,
-           explanatory, correlation=estimate, p.value,
-           adj.p.value, sig)
-}
-
-all_state_res = all_models_result(state_dat, 'state', state_consumption)
-all_hrr_res = all_models_result(hrr_dat, 'hrr', hrr_consumption)
-
-all_result = bind_rows(
-  all_state_res %>% mutate(place='state'),
-  all_hrr_res %>% mutate(place='hrr')
-)
-```
-
-#### State-wide consumption, all-antibiogram
-
-```{r}
-all_result %>%
-  filter(place=='state') %>% 
-  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
-  select(bug, drug_group, 
-         explanatory, correlation, p.value,
-         adj.p.value, sig) %>%
-  kable
-```
-
-#### State-wide consumtion, all-antibiogram using linear models
-
-```{r}
-linear_all_models_result = function(x, this_place) {
-  x %>%
-    group_by(bug, drug_group) %>%
-    do(linear_models(., 'percent_nonsusceptible')) %>%
-    ungroup %>%
-    filter(!is.na(estimate)) %>%
-    mutate(adj.p.value=p.adjust(p.value, method='BH'),
-           sig=pvalue_stars(adj.p.value)) %>%
-    select(bug, drug_group,
-           explanatory, slope=estimate, p.value,
-           adj.p.value, sig)
-}
-
-linear_all_state_res = linear_all_models_result(state_dat, 'state')
-linear_all_hrr_res = linear_all_models_result(hrr_dat, 'hrr')
-
-linear_all_state_res %>% kable
-```
-
-#### HRR-wide consumption, all-antibiogram
-
-```{r}
-all_result %>%
-  filter(place=='hrr') %>% 
-  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
-  select(bug, drug_group, 
-         explanatory, correlation, p.value,
-         adj.p.value, sig) %>%
-  kable
-```
+swo> Copy the code, and allow a function pointer for lm function to use
 

### 2017-04-05 states/resistance/ms2.Rmd
improved a lot

@@
 
-- Implement weights in the linear correlations
 - Join multiple results into one table, or do pairwise comparisons of analytical methods
+- How to use consumption data from multiple years?
+- Show that resistance doesn't have a pattern over years
+- more intelligently pooling data from the different *S. aureus* designations (in the antibiogram data, *S. aureus* is reported as one of *S. aureus*, MRSA, or MSSA)
+- do Spearman so we can compare with Goossens?
+- Get the right bug list
 
@@
 Here I report results of comparisons antibiotic consumption and resistance across states and HRRs.
-
-In a separate analysis, I determined that the antibiogram data in certain locales, separating the data into those locales decreased variance more than could be expected by chance.
-Thus, for some bug/drug combinations, it is sensible to break the antibiogram data from census regions down into states, while for other bug/drug combinations, it does not.
-Similarly, HRRs can be coherent or not.
 
@@
 I pooled the antiobiograms from all years.
-
+To ensure coherence, I only use the bug/drug combinations that appear in antibiograms in at least 20 states.
+This is equivalent to filtering those bug/drug combinations that have at least 125 records.
 I selected a set of organisms in the antibiogram to examine.
@@
 
-For each bug, consumption/resistance drug, and state, I ran a Spearman correlation test and adjusted the $p$-values using the Benjamini-Hochberg method.
-I also separately adjusted the $p$-values considering only the tests performed on coherent bug/drug/state combinations.
+For each bug/drug/state, I computed a weighted mean resistance: the values are $r_i/n_i$, the number of resistant isolates over the total number of isolates, and the weights are $\sqrt{n_i}$.
+I similarly computed a mean resistance for each HRR.
+
+For each bug/drug, I ran a linear model predicting the state resistance means from the state consumptions, weighted by the square root of the number of antibiograms for that bug/drug/state.
+I adjusted the $p$-values for the slope terms using the Benjamini-Hochberg method.
 I performed a similar analysis using HRRs.
-
-The explanatory variable in the correlation was either claims per 1,000 people per year (CPKP) or days' supply (my proxy for daily defined dose) per 1,000 people per day (DID).
-The dependent variable was always mean percent susceptibility of that bug to that resistance drug across antibiograms from that state/HRR.
+The explanatory variable was claims per 1,000 people per year (CPKP).
 
@@
 
-I used the 2011 Medicare data.
-To ensure coherence, I only use the bug/drug combinations that appear in antibiograms in at least 20 states.
-This is equivalent to filtering those bug/drug combinations that have at least 125 records.
-
 ```{r load_data}
@@
 abg = read_tsv('../../../../antibiogram/data/abg.tsv', col_types=cols(zipcode='c')) %>%
+  filter(!is.na(n_isolates), n_isolates > 0) %>%
   mutate(drug=tolower(drug), percent_nonsusceptible=100-percent_susceptible) %>%
@@
 
-models = function(df, independent_var, explanatory_vars) {
-  f = paste0(independent_var, ' ~ ', paste0(explanatory_vars, collapse = ' + ')) %>% formula
-  lm(f, data=df) %>% tidy
+linear_model_formula = function(independent_var, explanatory_vars) {
+  paste0(independent_var, ' ~ ', paste0(explanatory_vars, collapse = ' + ')) %>% formula
+}
+
+linear_models = function(df, independent_var, explanatory_vars) {
+  f = linear_model_formula(independent_var, explanatory_vars)
+  lm(formula=f, weights=place_weight, data=df) %>% tidy
 }
@@
 models_result = function(df) {
-  group_by(df, bug, drug_group, place) %>%
-    summarize(mean_percent_nonsusceptible=weighted.mean(mean_percent_nonsusceptible, place_weight),
-              cpkp=unique(cpkp),
-              did=unique(did)) %>%
-    group_by(bug, drug_group) %>%
-    do(models(., 'mean_percent_nonsusceptible', 'cpkp')) %>%
+  group_by(df, bug, drug_group) %>%
+    do(linear_models(., 'mean_percent_nonsusceptible', 'cpkp')) %>%
     ungroup %>%
@@
 
-```{r make_plots}
+```{r make_plots, eval=FALSE}
 unlink('plots/*.pdf')
@@
 
-swo> Copy the code, and allow a function pointer for lm function to use
+```{r}
+mblm_model = function(df, independent_var, explanatory_vars) {
+  f = linear_model_formula(independent_var, explanatory_vars)
+  mblm(f, dataframe=df) %>% tidy
+}
+
+mblm_models = function(df) {
+  group_by(df, bug, drug_group) %>%
+    do(mblm_model(., 'mean_percent_nonsusceptible', 'cpkp')) %>%
+    ungroup %>%
+    filter(term=='cpkp') %>%
+    mutate(adj.p.value=p.adjust(p.value, method='BH'),
+           sig=pvalue_stars(adj.p.value))
+}
+
+mblm_state_res = mblm_models(state_dat)
+mblm_hrr_res = mblm_models(hrr_dat)
+
+mblm_result = bind_rows(
+  mblm_models(state_dat) %>% mutate(place_type='state'),
+  mblm_models(hrr_dat) %>% mutate(place_type='hrr')
+)
+```
+
+These results are fairly well-correlated with the linear models. I'll compare the slopes:
+
+```{r}
+left_join(result, mblm_result, by=c('bug', 'drug_group', 'place_type')) %>%
+  filter(estimate.y > 1e-9) %>%
+  ggplot(aes(x=estimate.x, y=estimate.y)) +
+  geom_point() +
+  geom_abline() + geom_smooth(method='lm') +
+  xlab('weighted least squares slope') + ylab('robust regression slope') +
+  theme_minimal()
+```
+
+And the $p$-values:
+
+```{r}
+left_join(result, mblm_result, by=c('bug', 'drug_group', 'place_type')) %>%
+  ggplot(aes(x=p.value.x, y=p.value.y)) +
+  geom_point() +
+  geom_abline() + geom_hline(yintercept = 0.05) + geom_vline(xintercept = 0.05) +
+  xlab('weighted least squares p') + ylab('robust regression p') +
+  scale_x_log10() + scale_y_log10() +
+  theme_minimal()
+```
 
 # Discussion
-
-## Future work
-
-In the future, I intend to refine this analysis by:
-
-- pooling various consumption drugs (e.g., cipro and levo consumption probably have comparable effects on cipro resistance)
-- more intelligently pooling data from the different *S. aureus* designations (in the antibiogram data, *S. aureus* is reported as one of *S. aureus*, MRSA, or MSSA)
-- comparing data from various years (i.e., is there any evidence that a consumption/resistance relationship changed drastically?)
-- pooling data from various years (i.e., if the relationship doesn't change over years, we might get more power by including
 
@@
 
-- *S. aureus* and azithromycin
-- Potentially, a backward relationship for *S. aureus* and TMP-SMX
-- *S. pneumoniae* and azithromycin
-- *E. coli* and quinolones
+- CNST and quinolones (HRR only)
+- *E. faecium* and penicillin (state only)
+- *E. coli* with penicillin and quinolones (state & HRR)
+- MSSA and macrolides (state only)
+- MSSA and clindamycin (HRR only)
+- *P. mirabilis* and quinolones (HRR only)
+- Pneumococcus with macrolides (HRR; marginal in state)
+
 Interestingly, this accords with some of the conclusions from @bell-systematic-2014, namely:

### 2017-04-11 inequality-resistance/inequality-resistance.Rmd
many changes!

--- 2017-03-08 inequality-resistance/inequality-resistance.Rmd
@@
 ```{r global_options, include=FALSE}
-knitr::opts_chunk$set(fig.width=7, fig.height=4, 
+knitr::opts_chunk$set(fig.width=7, fig.height=4, dev=c('pdf', 'png'), 
                       echo=TRUE, warning=FALSE, message=FALSE, cache=TRUE)
+import::from(broom, tidy)
 ```

### 2017-04-11 ms1/ms1.Rmd
many changes!

@@
     if (i == 1) {
-      sprintf('< %.1f', upper_limits[i])
+      sprintf('<= %.1f', upper_limits[i])
     } else {
-      sprintf('>= %.1f and < %.1f', upper_limits[i - 1], upper_limits[i])
+      sprintf('> %.1f and <= %.1f', upper_limits[i - 1], upper_limits[i])
     }
   })
-  formula_list = mapply(function(str, lim) as.formula(sprintf('x < %.1f ~ "%s"', lim, str)),
+  formula_list = mapply(function(str, lim) as.formula(sprintf('x <= %.1f ~ "%s"', lim, str)),
                         levels, upper_limits)
@@
 
-multivariate_formula = paste(c('n_claims ~ age_over_65 + is_female + ', conditions), collapse=' + ') %>% formula
+multivariate_formula = paste(c('n_claims ~ age_over_66 + is_female + ', conditions), collapse=' + ') %>% formula
+start = rep(1.0, 3 + length(conditions))
 
@@
   usage %>%
-    mutate(age_over_65 = age - 65) %>%
-    glm(multivariate_formula, data=., family='poisson') %>%
+    mutate(age_over_66 = age - 66) %>%
+    glm(multivariate_formula, data=., family=quasipoisson(link='identity'), start=start) %>%
     tidy()
@@
 ```{r analyze_function}
+age_group_f = function(x) bounds_group(x, c(65, 70, 75, 80, 85, 90, 95, 100, 116))
+claims_group_f = function(x) bounds_group(x, c(0, 1, 2, 5, 10, 20, Inf))
+
 analyze = function(this_year) {
@@
   bene = read_tsv(sprintf('../bene_%i.tsv', this_year)) %>%
-    filter(age >= 65, age < 116, hmo_months==0) %>%
+    filter(between(age, 66, 116), hmo_months==0) %>%
     mutate(is_female=sex=='female', is_white=race=='white') %>%
@@
   # not just by raw age)
-  age_group_f = function(x) bounds_group(x, c(65, 70, 75, 80, 85, 90, 95, 116))
   claims_by_age = summarize_by(
       bene %>% mutate(age_group=age_group_f(age)),
-    pde, 'age_group') %>%
+      pde, 'age_group'
+    ) %>%
     mutate(year=this_year)
@@
 
-  claims_group_f = function(x) bounds_group(x, c(1, 2, 5, 10, 20, Inf))
   chars_by_consumption = usage %>%

### 2017-04-11 ms1/supp1.Rmd
many changes!

@@
 bene = read_tsv(sprintf('../bene_%i.tsv', this_year)) %>%
-  filter(age >= 65, age < 116, hmo_months==0) %>%
+  filter(between(age, 66, 100), hmo_months==0) %>%
   mutate(is_female=sex=='female', is_white=race=='white') %>%
@@
   left_join(cc, by='bene_id') %>%
-  mutate(age_over_65 = age - 65)
+  mutate(age_over_66 = age - 66)
 
 stopifnot(all(usage$age >= 0))
-```
-
-# All-condition negative binomial model
-
-Each point represents a chronic condition. The coordinates are corresponding coefficients in the two models. The formula for the model is claims against age and sex.
-
-```{r all_condition_negbin}
-all_condition_formula_start = 'n_claims ~ age_over_65 + is_female + '
-all_condition_formula = formula(paste(c(all_condition_formula_start, conditions), collapse=' + '))
-
-all_condition_negbin_model = usage %>%
-  MASS::glm.nb(all_condition_formula, data=.) %>%
-  tidy()
-
-all_condition_poisson_model = usage %>%
-  glm(all_condition_formula, data=., family='poisson') %>%
-  tidy()
-```
-
-```{r cache=FALSE}
-bind_rows(mutate(all_condition_poisson_model, method='Poisson'),
-          mutate(all_condition_negbin_model, method='negbin')) %>%
-  mutate(condition=str_replace(term, 'TRUE$', '')) %>%
-  left_join(condition_names, by='condition') %>%
-  mutate(term=if_else(is.na(condition_name), condition, condition_name)) %>%
-  select(term, method, estimate) %>%
-  spread(method, estimate) %>%
-  ggplot(aes(x=Poisson, y=negbin)) +
-  geom_abline(col='gray') +
-  geom_point() +
-  xlim(-0.2, 0.6) + ylim(-0.2, 0.6) +
-  xlab('Poisson multivariate model coefficient') +
-  ylab('negative binomial multivariate model coefficient') +
-  coord_fixed() +
-  theme_minimal()
 ```
@@
 rename_(x, .dots=setNames(condition, 'has_condition')) %>%
-  glm(n_claims ~ age_over_65 + is_female + has_condition, family='poisson', data=.) %>%
+  glm(n_claims ~ age_over_66 + is_female + has_condition, family='poisson', data=.) %>%
   tidy() %>%
@@
 lasso_x = usage %>%
-  select(age_over_65, is_female, AMI:HYPOTH) %>%
+  select(age_over_66, is_female, AMI:HYPOTH) %>%
   as.matrix
@@
 race_geography_model_f = function(condition) {
-  form = formula(paste0('n_claims ~ age_over_65 + is_female + is_white + region + ', condition))
+  form = formula(paste0('n_claims ~ age_over_66 + is_female + is_white + region + ', condition))
   usage %>%

### 2017-04-11 states/resistance/ms2.Rmd
many changes!

@@
 
+spearman_model = function(df, independent_var, explanatory_vars) {
+  cor.test(df[[independent_var]], df[[explanatory_vars]], method='spearman') %>% tidy
+}
+
 models_result = function(df) {
@@
 
+cor_models_result = function(df) {
+  group_by(df, bug, drug_group) %>%
+    do(spearman_model(., 'mean_percent_nonsusceptible', 'cpkp')) %>%
+    ungroup %>%
+    #filter(term=='cpkp') %>%
+    mutate(adj.p.value=p.adjust(p.value, method='BH'),
+           sig=pvalue_stars(adj.p.value))
+}
+
 state_res = models_result(state_dat)
@@
   hrr_res %>% mutate(place_type='hrr')
+)
+
+cor_result = bind_rows(
+  cor_models_result(state_dat) %>% mutate(place_type='state'),
+  cor_models_result(hrr_dat) %>% mutate(place_type='hrr')
 )
@@
 
+## Cor vs lm
+
+I asked whether using Spearman correlations is better/different than using linear models. For the most part, they identify the same relationships.
+
+```{r cor_vs_lm}
+f = function(x) select(x, bug, drug_group, adj.p.value, place_type)
+x = bind_rows(
+  cor_result %>% f %>% mutate(model='cor'),
+  result %>% f %>% mutate(model='lm')
+) %>% spread(model, adj.p.value)
+
+x %>% filter(place_type=='state') %>% ggplot(aes(x=lm, y=cor)) +
+  geom_point() +
+  scale_x_log10() + scale_y_log10() +
+  geom_abline() + geom_hline(yintercept = 0.05) + geom_vline(xintercept = 0.05) +
+  coord_fixed()
+
+x %>% filter(cor < 0.05, lm > 0.05) %>% kable
+```
+
 # Discussion

### 2017-04-11 polish-bene-pde.R

@@ -37,6 +37,7 @@ polish_bene = function(x) {
 polish_pde = function(x) {
   select(x, bene_id=BENE_ID,
             service_date=SRVC_DT,
+            fill_number=FILL_NUM,
             days_supply=DAYSSPLY,
             generic_name=GNN,
             dosage_form_code=GCDF,
@@ -45,7 +46,7 @@ polish_pde = function(x) {
            days_supply=as.integer(days_supply)) %>%
     filter(dosage_form_code %in% oral_injected_codes) %>%
     left_join(antibiotic_names, by='generic_name') %>%
-    select(bene_id, service_date, antibiotic, antibiotic_class, days_supply)
+    select(bene_id, service_date, fill_number, antibiotic, antibiotic_class, days_supply)
 }
 
 # bene column types

### 2017-04-12 ms1/ms1.Rmd
updated

@@
 
-summarize_by = function(bene, pde, by) {
-  # e.g., by='sex'
-  # count number of beneficiaries in each group
-  denom = count_(bene, by) %>% rename(n_bene=n)
-
-  pde %>%
-    left_join(select_(bene, 'bene_id', by), by='bene_id') %>% # join information about the bene's sex (e.g.)
-    count_(by) %>% # count claims
-    left_join(denom, by=by) %>% # join n_bene
-    mutate(cpkp=n*1000/n_bene) %>%
+summarize_by = function(usage, by) {
+  group_by_(usage, by) %>%
+  summarize(n_claims=sum(n_claims), n_bene=n()) %>%
+  mutate(cpkp=n_claims*1000/n_bene) %>%
   select_(by, 'cpkp')
-}
-
-multicol_sum = function(x, new_name, target_columns) {
-  dummy_vars = sapply(1:length(target_columns), function(i) sprintf('x%i', i))
-  dummy_vars_string = paste0(dummy_vars, collapse=',')
-  dummy_vars_list = lapply(target_columns, as.name) %>% setNames(dummy_vars)
-  f = as.formula(paste0(c('~sum(', dummy_vars_string, ')'), collapse=''))
-
-  rowwise(x) %>%
-    mutate_(new_attribute=interp(f, .values=dummy_vars_list)) %>%
-    rename_(.dots=setNames('new_attribute', new_name)) %>%
-    ungroup()
-}
-
-conditions = c("AMI", "ALZHDMTA", "ATRIALFB", "CATARACT", "CHRNKIDN", "COPD", "CHF", "DIABETES", "GLAUCOMA", "HIPFRAC", "ISCHMCHT", "DEPRESSN", "OSTEOPRS", "RA_OA", "STRKETIA", "CNCRBRST", "CNCRCLRC", "CNCRPRST", "CNCRLUNG", "CNCRENDM", "ANEMIA", "ASTHMA", "HYPERL", "HYPERP", "HYPERT", "HYPOTH")
-
-multivariate_formula = paste(c('n_claims ~ age_over_66 + is_female + ', conditions), collapse=' + ') %>% formula
-start = rep(1.0, 3 + length(conditions))
-
-all_condition_model = function(usage) {
-  usage %>%
-    mutate(age_over_66 = age - 66) %>%
-    glm(multivariate_formula, data=., family=quasipoisson(link='identity'), start=start) %>%
-    tidy()
 }
@@
 
-```{r load_db, cache=FALSE}
+```{r load_data}
 regions = read_tsv('../../db/census-regions/census-regions.tsv') %>%
@@
                            col_names=c('condition', 'condition_name'))
+
+conditions = c("AMI", "ALZHDMTA", "ATRIALFB", "CATARACT", "CHRNKIDN", "COPD", "CHF", "DIABETES", "GLAUCOMA", "HIPFRAC", "ISCHMCHT", "DEPRESSN", "OSTEOPRS", "RA_OA", "STRKETIA", "CNCRBRST", "CNCRCLRC", "CNCRPRST", "CNCRLUNG", "CNCRENDM", "ANEMIA", "ASTHMA", "HYPERL", "HYPERP", "HYPERT", "HYPOTH")
+
+load_data = function(year) {
+  usage = read_tsv(sprintf('usage_%i.tsv', year))
+  pde = read_tsv(sprintf('pde_%i.tsv', year))
+
+  list(usage=usage, pde=pde)
+}
+```
+
+```{r single_condition_model}
+single_condition_model = function(x, condition) {
+  filter(x, age >= 67) %>%
+  mutate(age_over_67 = age - 67) %>%
+  rename_(.dots=setNames(condition, 'has_condition')) %>%
+    glm(n_claims ~ age_over_67 + is_female + has_condition,
+        family=quasipoisson(link='identity'), data=.) %>%
+    tidy() %>%
+    mutate(condition=condition)
+}
+
+single_condition_models = function(usage) {
+  lapply(conditions, function(cond) single_condition_model(usage, cond)) %>%
+    bind_rows()
+}
 ```
@@
 ```{r analyze_function}
-age_group_f = function(x) bounds_group(x, c(65, 70, 75, 80, 85, 90, 95, 100, 116))
+age_group_f = function(x) bounds_group(x, c(66, 70, 75, 80, 85, 90, 95, 100))
 claims_group_f = function(x) bounds_group(x, c(0, 1, 2, 5, 10, 20, Inf))
 
-analyze = function(this_year) {
-  # bene file is already filtered for plan coverage months and good geographies
-  bene = read_tsv(sprintf('../bene_%i.tsv', this_year)) %>%
-    filter(between(age, 66, 116), hmo_months==0) %>%
-    mutate(is_female=sex=='female', is_white=race=='white') %>%
-    left_join(regions, by='state')
+analyze = function(year) {
+  dat = load_data(year)
 
-  assert_has_no_duplicates(bene$bene_id)
-  total_n_bene = nrow(bene)
-
-  # keep only PDEs for known beneficiaries (semi join)
-  # convert all amox/clav to just amox
-  pde = read_tsv(sprintf('../pde_%i.tsv', this_year)) %>%
-    semi_join(bene, by='bene_id') #%>%
-    #mutate(antibiotic=if_else(antibiotic=='amoxicillin/clavulanate', 'amoxicillin', antibiotic))
-
-  total_n_pde = nrow(pde)
+  total_n_bene = nrow(dat$usage)
+  total_n_pde = nrow(dat$pde)
 
   totals = data_frame(key=c('n_bene', 'n_pde'),
-                      value=c(total_n_bene, total_n_pde),
-                      year=this_year)
+                      value=c(total_n_bene, total_n_pde))
 
   # claims per 1,000 people by abx class
-  claims_by_class = pde %>%
+  claims_by_class = dat$pde %>%
     count(antibiotic_class) %>%
     mutate(cpkp=n*1000/total_n_bene) %>%
-    select(antibiotic_class, cpkp) %>%
-    mutate(year=this_year)
+    select(antibiotic_class, cpkp)
 
   # the top few inidividual antibiotics
-  top_abx = pde %>%
+  top_abx = dat$pde %>%
     count(antibiotic) %>%
     mutate(cpkp=n*1000/total_n_bene) %>%
-    select(antibiotic, cpkp) %>% 
-    mutate(year=this_year)
+    select(antibiotic, cpkp)
 
@@
   # in the same way as the consumption data, so we can use the summarize_by function
-  claims_by_sex = summarize_by(bene, pde, 'sex') %>%
-    mutate(year=this_year)
+  claims_by_sex = summarize_by(dat$usage, 'is_female')
 
@@
   # not just by raw age)
-  claims_by_age = summarize_by(
-      bene %>% mutate(age_group=age_group_f(age)),
-      pde, 'age_group'
-    ) %>%
-    mutate(year=this_year)
+  claims_by_age = dat$usage %>%
+    group_by(age) %>%
+    summarize(n_bene=n(),
+              mean_n_claims=mean(n_claims),
+              sd_n_claims=sd(n_claims),
+              mean_n_cc=mean(n_cc),
+              sd_n_cc=sd(n_cc)) %>%
+    mutate(sem_n_claims=sd_n_claims/sqrt(n_bene),
+           sem_n_cc=sd_n_cc/sqrt(n_bene))
 
-  # then by census region
-  claims_by_region = summarize_by(bene, pde, 'region') %>%
-    mutate(year=this_year)
+  claims_by_age_group = dat$usage %>%
+    mutate(age_group=age_group_f(age)) %>%
+    group_by(age_group) %>%
+    summarize(n_claims=sum(n_claims), n_bene=n()) %>%
+    mutate(cpkp=n_claims*1000/n_bene) %>%
+    select(age_group, n_bene, cpkp)
 
-  # characteristics by consumption
-  cc = read_tsv(sprintf('../cc_%i.tsv', this_year)) %>%
-    select(-ALZH)
+  claims_by_region = summarize_by(dat$usage, 'region')
 
-  usage = pde %>% count(bene_id) %>% rename(n_claims=n) %>%
-    right_join(bene, by='bene_id') %>%
-    replace_na(list(n_claims=0)) %>%
-    left_join(cc, by='bene_id') %>%
-    multicol_sum('n_cc', conditions)
-
-  chars_by_consumption = usage %>%
+  chars_by_consumption = dat$usage %>%
     mutate(claims_group=claims_group_f(n_claims)) %>%
@@
               `% female`=mean(is_female) * 100,
-              `mean no. chronic conditions`=mean(n_cc)) %>%
-    mutate(year=this_year)
+              `mean no. chronic conditions`=mean(n_cc))
 
-  condition_model = all_condition_model(usage) %>%
-    mutate(year=this_year)
+  condition_model = single_condition_models(dat$usage)
 
@@
        claims_by_age=claims_by_age,
+       claims_by_age_group=claims_by_age_group,
        claims_by_region=claims_by_region,
        chars_by_consumption=chars_by_consumption,
-       condition_model=condition_model)
+       condition_model=condition_model) %>%
+    lapply(function(x) mutate(x, year=year))
 }
@@
   spread(key, value) %>%
-  mutate(cpkp=n_pde*1000/n_bene) %>%
+  mutate(cpkp=round(n_pde*1000/n_bene)) %>%
   mutate_at(vars(n_bene, n_pde), function(x) prettyNum(x, big.mark=',')) %>%
   rename(`number of beneficiaries`=n_bene, `number of antibiotic PDEs`=n_pde, `CPKP`=cpkp) %>%
-  arrange(year) %>%
+  arrange(year) %T>%
+  write_tsv('tables/table1.tsv') %>%
   kable
@@
   head(9) %$%
-  antibiotic_class %>%
-  c('other')
+  antibiotic_class
+
+class_levels = c(top_classes, 'other')
 
@@
   group_by(year, antibiotic_class) %>%
-  summarize(cpkp=sum(cpkp)) %>%
+  summarize(cpkp=round(sum(cpkp))) %>%
   spread(year, cpkp) %>%
-  mutate(antibiotic_class=factor(antibiotic_class, levels=top_classes, ordered=TRUE)) %>%
-  arrange(antibiotic_class) %>%
+  mutate(antibiotic_class=factor(antibiotic_class, levels=class_levels, ordered=TRUE)) %>%
+  arrange(antibiotic_class) %T>%
+  write_tsv('tables/table2_class.tsv') %>%
   kable
@@
 top_abx = data_by_field('top_abx') %>%
+  mutate(cpkp=round(cpkp)) %>%
   spread(year, cpkp) %>%
   arrange(desc(`2011`)) %>%
-  head(11) %>%
+  head(11) %T>%
+  write_tsv('tables/table2_drug.tsv') %>%
   kable
@@
 claims_by_sex = data_by_field('claims_by_sex') %>%
-  filter(sex %in% c('male', 'female')) %>%
-  spread(year, cpkp) %>%
+  mutate(cpkp=round(cpkp)) %>%
+  spread(year, cpkp) %T>%
+  write_tsv('tables/table2_sex.tsv') %>%
   kable
 
-claims_by_age = data_by_field('claims_by_age') %>%
-  filter(!is.na(age_group)) %>%
-  spread(year, cpkp) %>%
+claims_by_age_group = data_by_field('claims_by_age_group') %>%
+  mutate(cpkp=round(cpkp)) %>%
+  select(age_group, year, cpkp) %>%
+  spread(year, cpkp) %T>%
+  write_tsv('tables/supptable1_age_group.tsv') %>%
   kable
 
+bene_by_age_group = data_by_field('claims_by_age_group') %>%
+  mutate(n_bene=prettyNum(n_bene, big.mark=',')) %>%
+  select(age_group, year, n_bene) %>%
+  spread(year, n_bene) %T>%
+  write_tsv('tables/supptable1_age_group_denom.tsv') %>%
+  kable
+
+claims_by_age = data_by_field('claims_by_age')
+
 claims_by_region = data_by_field('claims_by_region') %>%
-  filter(!is.na(region)) %>%
+  mutate(cpkp=round(cpkp)) %>%
   spread(year, cpkp) %>%
-  arrange(desc(`2011`)) %>%
+  arrange(desc(`2011`)) %T>%
+  write_tsv('tables/table2_region.tsv') %>%
   kable
@@
 chars_by_consumption = data_by_field('chars_by_consumption') %>%
-  arrange(desc(year), desc(claims_group)) %>%
+  mutate_at(vars(`mean age`, `% female`, `mean no. chronic conditions`), function(x) round(x, digits=2)) %>%
+  arrange(year, claims_group) %>%
   kable
@@
 condition_model = data_by_field('condition_model') %>%
-  mutate(condition=str_replace(term, 'TRUE$', '')) %>%
   left_join(condition_names, by='condition') %>%
-  mutate(term=if_else(is.na(condition_name), condition, condition_name)) %>%
-  select(term, estimate, year) %>%
-  spread(year, estimate) %>%
-  arrange(desc(`2011`)) %>%
+  filter(year==2011, term != '(Intercept)') %>%
+  mutate(estimate=round(estimate, digits=3)) %>%
+  select(condition=condition_name, term, estimate) %>%
+  spread(term, estimate) %>%
+  arrange(desc(has_conditionTRUE)) %T>%
+  write_tsv('tables/table3.tsv') %>%
   kable
@@
 
-`r claims_by_age`
+`r claims_by_age_group`
+
+`r bene_by_age_group`
+
+# Figure 1: Claims and conditions by age
+
+```{r fig1_claims_by_age}
+claims_by_age %>%
+  filter(year==2011) %>%
+  ggplot(aes(x=age,
+             y=mean_n_claims,
+             ymin=mean_n_claims-sem_n_claims,
+             ymax=mean_n_claims+sem_n_claims)) +
+  geom_point() +
+  geom_errorbar() +
+  theme_minimal() +
+  scale_x_continuous(breaks=c(66, 70, 80, 90, 100))
+```
+
+```{r fig1_cc_by_age}
+claims_by_age %>%
+  filter(year==2011) %>%
+  ggplot(aes(x=age,
+             y=mean_n_cc,
+             ymin=mean_n_cc-sem_n_cc,
+             ymax=mean_n_cc+sem_n_cc)) +
+  geom_point() +
+  geom_errorbar() +
+  theme_minimal() +
+  scale_x_continuous(breaks=c(66, 70, 80, 90, 100))
+```
 

### 2017-04-12 ms1/supp1.Rmd
updated

@@
 
-```{r single_condition}
-summarize_condition = function(x, condition) {
-rename_(x, .dots=setNames(condition, 'has_condition')) %>%
-  glm(n_claims ~ age_over_66 + is_female + has_condition, family='poisson', data=.) %>%
-  tidy() %>%
-  mutate(condition=condition)
+```{r all_condition_model}
+multivariate_formula = paste(c('n_claims ~ age_over_66 + is_female + ', conditions), collapse=' + ') %>% formula
+start = rep(1.0, 3 + length(conditions))
+
+all_condition_model = function(usage) {
+  usage %>%
+    mutate(age_over_66 = age - 66) %>%
+    glm(multivariate_formula, data=., family=quasipoisson(link='identity'), start=start) %>%
+    tidy()
 }
-
-single_condition_models = lapply(conditions, function(cond) summarize_condition(usage, cond)) %>%
-  bind_rows()
 ```

### 2017-04-12 ms1/polish-usage.R

@@ -1,0 +1,38 @@
+regions = read_tsv('../../db/census-regions/census-regions.tsv') %>%
+  select(state, region)
+
+antibiotic_class = read_tsv('../abx_class.tsv') %>%
+  select(antibiotic, antibiotic_class) %>%
+  distinct
+
+analyze = function(year) {
+  cc = read_tsv(sprintf('../cc_%i.tsv', year)) %>%
+    select(-ALZH)
+
+  n_cc = cc %>% select(AMI:HYPOTH) %>% rowSums
+
+  bene = read_tsv(sprintf('../bene_%i.tsv', year)) %>%
+    filter(between(age, 66, 100), hmo_months==0) %>%
+    mutate(is_female=sex=='female', is_white=race=='white') %>%
+    left_join(regions, by='state') %>%
+    left_join(cc, by='bene_id') %>%
+    select(-sex, -race, -zipcode, -hmo_months)
+
+  n_cc = bene %>% select(AMI:HYPOTH) %>% rowSums
+
+  bene %<>% mutate(n_cc=n_cc)
+
+  pde = read_tsv(sprintf('../pde_%i.tsv', year)) %>%
+    semi_join(bene, by='bene_id') %>%
+    mutate(year=year)
+
+  usage = pde %>% count(bene_id) %>% rename(n_claims=n) %>%
+    right_join(bene, by='bene_id') %>%
+    replace_na(list(n_claims=0)) %>%
+    mutate(year=year)
+
+  write_tsv(usage, sprintf('usage_%i.tsv', year))
+  write_tsv(pde, sprintf('pde_%i.tsv', year))
+}
+
+for (y in 2011:2014) analyze(y)

### 2017-04-12 ms1/ms1.Rmd
updates

@@
 
-  claims_by_age_group = dat$usage %>%
+  chars_by_age_group = dat$usage %>%
     mutate(age_group=age_group_f(age)) %>%
     group_by(age_group) %>%
-    summarize(n_claims=sum(n_claims), n_bene=n()) %>%
-    mutate(cpkp=n_claims*1000/n_bene) %>%
-    select(age_group, n_bene, cpkp)
+    summarize(mean_n_claims=mean(n_claims), mean_n_cc=mean(n_cc)) %>%
+    mutate(cpkp=mean_n_claims*1000) %>%
+    select(age_group, cpkp, mean_n_cc)
 
@@
        claims_by_age=claims_by_age,
-       claims_by_age_group=claims_by_age_group,
+       chars_by_age_group=chars_by_age_group,
        claims_by_region=claims_by_region,
@@
 
-claims_by_age_group = data_by_field('claims_by_age_group') %>%
-  mutate(cpkp=round(cpkp)) %>%
-  select(age_group, year, cpkp) %>%
-  spread(year, cpkp) %T>%
+chars_by_age_group = data_by_field('claims_by_age_group') %>%
+  mutate(cpkp=round(cpkp), mean_n_cc=round(mean_n_cc, digits=2)) %>%
+  select(age_group, year, cpkp, mean_n_cc) %>%
+  gather('variable', 'value', cpkp, mean_n_cc) %>%
+  mutate(variable=interaction(year, variable)) %>%
+  select(-year) %>%
+  spread(variable, value) %T>%
   write_tsv('tables/supptable1_age_group.tsv') %>%
@@
 
-condition_model = data_by_field('condition_model') %>%
+condition_model_all_years = data_by_field('condition_model') %>%
   left_join(condition_names, by='condition') %>%
-  filter(year==2011, term != '(Intercept)') %>%
+  filter(term != '(Intercept)') %>%
   mutate(estimate=round(estimate, digits=3)) %>%
-  select(condition=condition_name, term, estimate) %>%
+  select(year, condition=condition_name, term, estimate) %T>%
+  write_tsv('tables/supptable4_singlecondition.tsv')
+
+condition_model = condition_model_all_years %>%
+  filter(year==2011) %>% select(-year) %>%
   spread(term, estimate) %>%
@@
 ## Consumption by age
+
+`r chars_by_age_group`
 

### 2017-04-13 ms1/ms1.Rmd
updated ms1

@@
 ```{r analyze_function}
-age_group_f = function(x) bounds_group(x, c(66, 70, 75, 80, 85, 90, 95, 100))
+age_group_f = function(x) bounds_group(x, c(65, 70, 75, 80, 85, 90, 95, 100))
 claims_group_f = function(x) bounds_group(x, c(0, 1, 2, 5, 10, 20, Inf))
@@
 
-chars_by_age_group = data_by_field('claims_by_age_group') %>%
+chars_by_age_group = data_by_field('chars_by_age_group') %>%
   mutate(cpkp=round(cpkp), mean_n_cc=round(mean_n_cc, digits=2)) %>%
-  select(age_group, year, cpkp, mean_n_cc) %>%
-  gather('variable', 'value', cpkp, mean_n_cc) %>%
-  mutate(variable=interaction(year, variable)) %>%
-  select(-year) %>%
-  spread(variable, value) %T>%
-  write_tsv('tables/supptable1_age_group.tsv') %>%
-  kable
+  select(age_group, year, cpkp, mean_n_cc) %T>%
+  write_tsv('tables/data_age_group.tsv')
 

### 2017-04-13 ms2/ms2.Rmd
ms2 start

--- 2017-03-13 inequality/consumption-inequality.Rmd
@@
 author: "Scott Olesen"
-date: "2/22/2017"
 output: html_document
@@
                       echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
+import::from(ineq, Gini)
 ```
 
-# Background
-
-What is the inequality in consumption for antibiotics? How does it depend on region or the antibiotic used?
-
-# Methods
-
-I define the Juran inequality index $J$: for a random variable $X$ with a domain $\{0, 1\}$ and a cumulative distribution function $F_X$, $J(X)$ is that value such that $F(J(X)) = 1 - J(X)$.
-For example, if the top 20% of beneficiaries make 80% of the claims, then $J = 0.8$.
-
-For each group under consideration, I considered 5 different metrics:
-
-- the fraction of beneficiaries with at least one claim
-- the Juran index among all beneficiaries
-- the Juran index among beneficiaries with at least one claim
-- the Gini index among all beneficiaries
-- the Gini index among beneficiaries with at least one claim
-
 ```{r setup}
-library(ineq)
 is_nonzero = function(x) x > 0
@@
 fraction_nonzero = function(x) sum(is_nonzero(x)) / length(x)
-gini = function(x) ineq(x, type='Gini')
 
@@
 
-add_zeros = function(x, n) {
-  if (n < length(x)) {
-    stop(sprintf('cannot extend x of length %i to %i', length(x), n))
-  }
-
-  c(x, rep(0, n - length(x)))
-}
-
-summarize_claims = function(x) {
-  summarize(x, n_claims=sum(n_claims), n_days=sum(days_supply)) %>%
-    mutate(n_days=if_else(is.na(n_days), 0L, n_days))
-}
-
-summarize_inequality = function(x, ...) {
-  do(x, inequalities(., ...))
+klasma = function(x, threshold) {
+  if (!between(threshold, 0.0, 1.0)) stop("bad threshold")
+  data_frame(x=x) %>%
+    mutate(cdf=cume_dist(x)) %$%
+    { sum(x[cdf >= threshold]) / sum(x) }
 }
@@
 
-```{r load_data}
-usage = read_tsv('../bene_usage_2011.tsv')
+```{r compute_inequalities}
+bene = read_tsv('../
+
+usage = read_tsv('usage_2011.tsv')
 ```

### 2017-04-13 ms2/data/collect-data.R

@@ -1,0 +1,44 @@
+# inequality function
+import::from(ineq, Gini)
+nonzero = function(x) x[x > 0]
+fraction_nonzero = function(x) length(nonzero(x)) / length(x)
+
+pad0 = function(x, to) c(x, rep(0, to - length(x)))
+
+inequalities = function(x) {
+  data_frame(total=sum(x),
+             mean=mean(x),
+             fnz=fraction_nonzero(x),
+             nzgini=Gini(nonzero(x)))
+}
+
+consumption_groups = read_tsv('../consumption_groups.tsv')
+
+regions = read_tsv('../../../db/census-regions/census-regions.tsv') %>%
+  select(state, state_abbreviation, region)
+
+bene = read_tsv('../../bene_2011.tsv') %>%
+  filter(between(age, 66, 100), hmo_months==0) %>%
+  left_join(regions, by='state')
+
+
+# compute total claims separately from individual drug groups
+total_pde = pde %>%
+  group_by(bene_id) %>%
+  summarize(n_claims=sum(n_claims)) %>%
+  right_join(bene, by='bene_id') %>%
+  replace_na(list(n_claims=0)) %>%
+  group_by(state) %>%
+  do(inequalities(.$n_claims))
+
+write_tsv(total_pde, 'total_pde.tsv')
+
+state_denom = bene %>% count(state) %>% rename(n_bene=n)
+state_inequality = pde %>%
+  right_join(select(bene, bene_id, state), by='bene_id') %>%
+  filter(!is.na(drug_group)) %>%
+  left_join(state_denom, by='state') %>%
+  group_by(state, drug_group) %>%
+  do(inequalities(pad0(.$n_claims, unique(.$n_bene))))
+
+write_tsv(state_inequality, 'group_inequality.tsv')

### 2017-04-13 ms2/work.R

@@ -1,0 +1,34 @@
+# > names(x)
+# [1] "state"    "cpb"      "v"        "fnznc"    "nzgini"   "n_bene"   "n_claims" "fz"      
+
+x = usage %>%
+  group_by(state) %>%
+  summarize(cpb=mean(n_claims),
+            v=var(n_claims),
+            fnznc=fnz(n_claims),
+            fz=1-fnznc,
+            nzgini=Gini(nz(n_claims)),
+            n_bene=n(),
+            n_claims=sum(n_claims))
+
+# fit the p value
+m = nls(fz ~ (1-p) ** ((1-p)/p*cpb), data=x, start=list(p=0.5))
+p = coef(m)[['p']]
+
+# or, do this with the variance
+# the coefficient of the variance v is (1-p)
+# this give 0.251 -> 0.749, which is higher than p=0.7 from above
+# lm(cpb ~ v + 0, data=x)
+
+# get the r values
+r = (1-p)/p * x$cpb
+
+gini_range = function(r, n_individuals, n_trials=1e2) {
+  values = replicate(n_trials, Gini(nz(rnbinom(n_individuals, r, p))))
+  data_frame(mu=mean(values), cil=quantile(values, 0.05), ciu=quantile(values, 0.95))
+}
+
+y = data_frame(r=r, n_bene=x$n_bene) %>%
+  group_by(r) %>%
+  do(gini_range(.$r, .$n_bene)) %>%
+  mutate(nbmu=p*r/(1-p), nbf0=(1-p)**r)

### 2017-04-13 ms2/ms2.Rmd
ms2 to new content

--- 2017-03-08 inequality/consumption-inequality.Rmd
@@
 ---
-title: "Inequality in consumption"
+title: "ms2"
 author: "Scott Olesen"
-date: "2/22/2017"
 output: html_document
@@
 
-# Background
+```{r load_data}
+regions = read_tsv('../../db/census-regions/census-regions.tsv') %>%
+  select(state, state_abbreviation, region)
 
-What is the inequality in consumption for antibiotics? How does it depend on region or the antibiotic used?
-
-# Methods
-
-I define the Juran inequality index $J$: for a random variable $X$ with a domain $\{0, 1\}$ and a cumulative distribution function $F_X$, $J(X)$ is that value such that $F(J(X)) = 1 - J(X)$.
-For example, if the top 20% of beneficiaries make 80% of the claims, then $J = 0.8$.
-
-For each group under consideration, I considered 5 different metrics:
-
-- the fraction of beneficiaries with at least one claim
-- the Juran index among all beneficiaries
-- the Juran index among beneficiaries with at least one claim
-- the Gini index among all beneficiaries
-- the Gini index among beneficiaries with at least one claim
-
-```{r setup}
-library(ineq)
-is_nonzero = function(x) x > 0
-nonzero = function(x) x[is_nonzero(x)]
-fraction_nonzero = function(x) sum(is_nonzero(x)) / length(x)
-gini = function(x) ineq(x, type='Gini')
-
-normalize = function(x) x / max(x)
-juran = function(x) {
-  # better to do with percent_rank and cume_dist
-  fraction_consumed = cumsum(sort(x, decreasing=TRUE)) %>% normalize
-  fraction_users = 1:length(x) %>% normalize
-  fraction_nonzero(fraction_consumed + fraction_users - 1)
-}
-
-add_zeros = function(x, n) {
-  if (n < length(x)) {
-    stop(sprintf('cannot extend x of length %i to %i', length(x), n))
-  }
-
-  c(x, rep(0, n - length(x)))
-}
-
-summarize_claims = function(x) {
-  summarize(x, n_claims=sum(n_claims), n_days=sum(days_supply)) %>%
-    mutate(n_days=if_else(is.na(n_days), 0L, n_days))
-}
-
-summarize_inequality = function(x, ...) {
-  do(x, inequalities(., ...))
-}
-
-inequalities = function(x, what, denom=NULL) {
-  z = x[[what]]
-  if (!is.null(denom)) z = add_zeros(z, denom)
-
-  data_frame(fraction_nonzero=fraction_nonzero(z),
-             juran=juran(z),
-             juran_nonzero=juran(nonzero(z)),
-             gini=gini(z),
-             gini_nonzero=gini(nonzero(z)))
-}
-
-top_antibiotics = c('azithromycin', 'ciprofloxacin', 'levofloxacin', 'amoxicillin', 'cephalexin', 'trimethoprim/sulfamethoxazole', 'nitrofurantoin', 'doxycycline')
+total_ineq = read_tsv('data/total_pde.tsv') %>%
+  left_join(regions, by='state')
+group_ineq = read_tsv('data/group_inequality.tsv')
 ```
 
-# Results
+# Aggregate usage correlates with fraction of users
 
-```{r load_data}
-usage = read_tsv('../bene_usage_2011.tsv')
+```{r fnz-mean}
+ggplot(total_ineq, aes(x=mean, y=fnz, color=region, label=state_abbreviation)) +
+geom_point() +
+geom_label(nudge_x=0.05) +
+xlab('claims per beneficiary') +
+ylab('fraction of beneficiaries w/ >= 1 claim')
 ```
 
-Overall, about half of beneficiaries took at least one antibiotic. Overall, 20% of beneficiaries claimed 80% of the antibiotic days' supply.
+# Inequality among users does not correlate well
 
-```{r inequality_overall}
-# put in both metrics of consumption
-# run it through summarize_inequality
-# then group by region or by drug also
-usage %>%
-  group_by(bene_id) %>%
-  summarize_claims() %>%
-  gather('consumption_metric', 'amount', n_claims, n_days) %>%
-  group_by(consumption_metric) %>%
-  summarize_inequality('amount') %>%
-  kable
+```{r nzgini-mean}
+ggplot(total_ineq, aes(x=mean, y=nzgini, color=region, label=state_abbreviation)) +
+geom_point() +
+geom_label(nudge_x=0.05) +
+xlab('claims per beneficiary') +
+ylab('Gini coefficient among benes. w/ >= 1 claim')
 ```
-
-```{r inequality_hist}
-usage %>%
-  group_by(bene_id) %>%
-  summarize(n_claims=sum(n_claims)) %>%
-  ggplot(aes(x=n_claims)) +
-  geom_histogram() +
-  scale_y_log10() +
-  xlab('number of claims') + ylab('number of beneficiaries') +
-  theme_minimal()
-```
-
-```{r inequality_cumrelfreq}
-normalize = function(x) x / max(x)
-usage %>%
-  group_by(bene_id) %>%
-  summarize(n_claims=sum(n_claims)) %>%
-  arrange(n_claims) %>%
-  mutate(cdf_x=cume_dist(n_claims), cdf_y=normalize(cumsum(n_claims))) %>%
-  group_by(cdf_x) %>%
-  summarize(cdf_y=max(cdf_y)) %>%
-  ggplot(aes(x=cdf_x, y=cdf_y)) +
-  geom_point() + geom_line() +
-  xlab('cumulative fraction of beneficiaries') +
-  ylab('cumulative fraction of claims') +
-  xlim(0, 1) + ylim(0, 1) +
-  theme_minimal()
-```
-
-Breaking down these results by region shows that the South, which consumed the most antibiotic per capita, has a slightly more even usage (Juran index of 79% versus 80% in other regions).
-I'm not sure that's a meaningful difference.
-
-```{r inequality_by_region}
-usage %>%
-  group_by(region, bene_id) %>%
-  summarize_claims() %>%
-  gather('consumption_metric', 'amount', n_claims, n_days) %>%
-  group_by(region, consumption_metric) %>%
-  summarize_inequality('amount') %>%
-  kable
-```
-
-```{r inequality_by_region2}
-# This uses code from inequality-resistance
-normalize = function(x) x / max(x)
-cumrelfreq = function(df, var) {
-  n = sort(df[[var]], decreasing=FALSE)
-  data_frame(x=cume_dist(n), y=normalize(cumsum(n))) %>%
-    group_by(x) %>%
-    summarize(y=max(y))
-}
-
-m = consumption_of('macrolide')
-
-x = m %>%
-  left_join(census_regions, by='state') %>%
-  group_by(region) %>%
-  do(cumrelfreq(., 'n_claims'))
-
-p = x %>%
-  ggplot(aes(x=x, y=y, color=region, group=region)) +
-  geom_line() + geom_point() +
-  xlab('cumulative fraction of users') +
-  ylab('cumulative fraction of drugs consumed') +
-  theme_minimal()
-
-show(p)
-```
-
-Among the most popular drugs, unevenness varies only slightly.
-Among users, azithromycin consumption has an approximately 60% Juran index, and ciprofloxacin has 65%, while TMP-SMX and vancomycin have 70%.
-
-```{r inequality_by_drug}
-n_bene = usage$bene_id %>% unique %>% length
-usage %>%
-  group_by(antibiotic, bene_id) %>%
-  summarize_claims() %>%
-  gather('consumption_metric', 'amount', n_claims, n_days) %>%
-  group_by(antibiotic, consumption_metric) %>%
-  summarize_inequality('amount', n_bene) %>%
-  filter(antibiotic %in% top_antibiotics) %>%
-  kable
-```
-
-## Inequality by state
-```{r compute_inequality_by_state}
-inequality_by_state = usage %>%
-  group_by(state, bene_id) %>%
-  summarize_claims() %>%
-  gather('consumption_metric', 'amount', n_claims, n_days) %>%
-  group_by(state, consumption_metric) %>%
-  summarize(n_bene=n(),
-             fraction_nonzero=fraction_nonzero(amount),
-             juran=juran(amount),
-             juran_nonzero=juran(nonzero(amount)),
-             gini=gini(amount),
-             gini_nonzero=gini(nonzero(amount))) %>%
-  ungroup()
-```
-
-The distribution of inequalities seems small: all states have (non-zero) Juran
-indices that fall between about 0.7 and 0.75. That being said, there does seem
-to be a trend by region: the fraction nonzero is highest in the South?
-
-These results all show inequalities measured using number of days.
-
-```{r show_inequality_by_state}
-inequality_by_state %>%
-  filter(consumption_metric=='n_days') %>%
-  ggplot(aes(x=juran_nonzero)) + geom_histogram() +
-  theme_minimal()
-
-inequality_by_state %>%
-  filter(fraction_nonzero > 0) %>%
-  filter(consumption_metric=='n_days') %>%
-  select(-consumption_metric) %>%
-  arrange(state) %>%
-  kable
-```
-
-These data break down, to some degree, by region: the fraction nonzero is highest for states in the South, but the inequality among users is highest for users in the Midwest.
-
-```{r show_inequality_by_region}
-regions = read_tsv('../../db/census-regions/census-regions.tsv') %>%
-  select(state, region)
-
-inequality_by_region = inequality_by_state %>%
-  filter(consumption_metric=='n_claims') %>%
-  left_join(regions, by='state')
-
-inequality_by_region %>%
-  ggplot(aes(x=n_bene, y=fraction_nonzero, color=region)) +
-  geom_point() +
-  xlab('number of beneficiaries in state') + ylab('fraction of beneficiaries with at least one claim') +
-  theme_minimal()
-
-inequality_by_region %>%
-  ggplot(aes(x=n_bene, y=juran_nonzero, color=region)) +
-  geom_point() +
-  xlab('number of beneficiaries in state') + ylab('Juran index for days of antibiotics among users') +
-  theme_minimal()
-```
-
-## Characteristics by consumption
-
-I broke down the beneficiaries based on their consumption: those who had no antibiotic claims were in one "tile", and I separated the rest into $n$-tiles.
-
-```{r chars_by_consumption}
-ntile_zero = function(x, n) {
-  out = x
-  out[x > 0] = ntile(out[x > 0], n)
-  out
-}
-
-bounds_group = function(x, upper_limits) {
-  upper_limits = as.integer(upper_limits)
-  levels = lapply(seq(from=1, to=length(upper_limits)), function(i) {
-    if (i == 1) {
-      sprintf('<= %i', upper_limits[i])
-    } else {
-      sprintf('> %i and <= %i', upper_limits[i - 1], upper_limits[i])
-    }
-  })
-  top_level = sprintf('> %i', tail(upper_limits, 1))
-  formula_list = mapply(function(str, lim) as.formula(sprintf('x <= %i ~ "%s"', lim, str)),
-                        levels, upper_limits) %>%
-    c(as.formula(sprintf('TRUE ~ "%s"', top_level)))
-
-  levels = c(levels, top_level)
-
-  do.call(case_when, formula_list) %>%
-    factor(levels=levels, ordered=TRUE)
-}
-
-usage %>%
-  group_by(age, sex, race, chronic, bene_id) %>%
-  summarize(n_claims=sum(n_claims),
-            n_days=sum(n_days)) %>%
-  ungroup() %>%
-  mutate(claims_group=bounds_group(n_claims, c(0, 1, 2, 5, 10, 20, 50, 100)),
-         is_female=if_else(sex=='female', 1, 0),
-         is_white=if_else(race=='white', 1, 0)) %>%
-  group_by(claims_group) %>%
-  summarize(n_beneficiaries=n(),
-            mean_age=mean(age),
-            percent_female=mean(is_female)*100,
-            percent_white=mean(is_white)*100,
-            mean_n_chronic_cond=mean(chronic),
-            mean_n_claims=mean(n_claims)) %>%
-  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
-  kable
-```
-
-As expected, greater consumption is correlated with greater age and comorbidity.
-There are strange trends among gender and race (higher consumption is associated with more female and more white until 10--20 claims, after which point those trends reverse).
-
-# Discussion
-
-I used only the 2011 data in this analysis.
-I intend to extend the analysis to other years.
-Data from other years could help confirm any of the (relatively small) differences in inequality in consumption.

### 2017-04-20 ms1/ms1.Rmd
many updates

@@
               sd_n_claims=sd(n_claims),
+              mean_n_claims_0cc=mean(n_claims[n_cc==0]),
+              sd_n_claims_0cc=sd(n_claims[n_cc==0]),
               mean_n_cc=mean(n_cc),
@@
     mutate(sem_n_claims=sd_n_claims/sqrt(n_bene),
+           sem_n_claims_0cc=sd_n_claims_0cc/sqrt(n_bene),
            sem_n_cc=sd_n_cc/sqrt(n_bene))
@@
     group_by(age_group) %>%
-    summarize(mean_n_claims=mean(n_claims), mean_n_cc=mean(n_cc)) %>%
+    summarize(n_bene=n(),
+              mean_n_claims=mean(n_claims),
+              mean_n_cc=mean(n_cc)) %>%
     mutate(cpkp=mean_n_claims*1000) %>%
-    select(age_group, cpkp, mean_n_cc)
+    select(age_group, n_bene, cpkp, mean_n_cc)
 
@@
   select(age_group, year, cpkp, mean_n_cc) %T>%
-  write_tsv('tables/data_age_group.tsv')
+  write_tsv('tables/data_age_group.tsv') %>%
+  kable
 
-bene_by_age_group = data_by_field('claims_by_age_group') %>%
+bene_by_age_group = data_by_field('chars_by_age_group') %>%
   mutate(n_bene=prettyNum(n_bene, big.mark=',')) %>%
@@
 
-claims_by_age = data_by_field('claims_by_age')
+claims_by_age = data_by_field('claims_by_age') %T>%
+  write_tsv('tables/tmp_claims_by_age.tsv')
 
@@
 
-`r claims_by_age_group`
-
 `r bene_by_age_group`
@@
 
+```{r}
+f = function(df) {
+  stopifnot(nrow(df) == 1)
+  data_frame(group=c('all', '0cc'),
+             mean_n_claims=c(df$mean_n_claims, df$mean_n_claims_0cc),
+             sem_n_claims=c(df$sem_n_claims, df$sem_n_claims_0cc))
+}
+
+claims_by_age %>%
+  filter(year==2011, between(age, 66, 96)) %>%
+  select(age, mean_n_claims, mean_n_claims_0cc) %>%
+  gather('group', 'mean_n_claims', mean_n_claims, mean_n_claims_0cc) %>%
+  ggplot(aes(x=age,
+             y=mean_n_claims,
+             color=group,
+             group=group)) +
+  geom_point() +
+  geom_smooth(method='lm', se=FALSE) +
+  theme_minimal() +
+  ylab('claims per person') +
+  scale_x_continuous(breaks=c(66, 70, 80, 90, 96)) +
+  scale_y_continuous(breaks=c(0, 0.5, 1, 1.5, 2.0), limits=c(0, 2))
+```
+
 # Table 2: Beneficiary characteristics by consumption

### 2017-04-20 ms2/ms2.Rmd
many updates

@@
 ```{r global_options, include=FALSE}
-knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/',
+knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/', dev=c('png', 'pdf'),
                       echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
+import::from(pander, pander)
+import::from(broom, tidy)
 ```
+
+# Consumption data
 
@@
 
-total_ineq = read_tsv('data/total_pde.tsv') %>%
+hrr = read_tsv('../../db/hrr/hrr.tsv') %>%
+  filter(year==2011) %>%
+  select(zipcode, hrr)
+
+all_ineq = read_tsv('data/all_drug_ineq.tsv') %>%
   left_join(regions, by='state')
-group_ineq = read_tsv('data/group_inequality.tsv')
+state_ineq = read_tsv('data/drug_ineq.tsv')
+hrr_ineq = read_tsv('data/hrr_drug_ineq.tsv')
 ```
 
-# Aggregate usage correlates with fraction of users
+## Aggregate usage correlates with fraction of users
 
-```{r fnz-mean}
-ggplot(total_ineq, aes(x=mean, y=fnz, color=region, label=state_abbreviation)) +
+```{r fnz-mean, cache=FALSE}
+ggplot(all_ineq, aes(x=mean, y=fnz, color=region, label=state_abbreviation)) +
 geom_point() +
@@
 xlab('claims per beneficiary') +
-ylab('fraction of beneficiaries w/ >= 1 claim')
+ylab('fraction of beneficiaries w/ >= 1 claim') +
+theme_minimal()
 ```
 
-# Inequality among users does not correlate well
+## Inequality among users does not correlate well
 
-```{r nzgini-mean}
-ggplot(total_ineq, aes(x=mean, y=nzgini, color=region, label=state_abbreviation)) +
+```{r nzgini-mean, cache=FALSE}
+ggplot(all_ineq, aes(x=mean, y=nzgini, color=region, label=state_abbreviation)) +
 geom_point() +
@@
 xlab('claims per beneficiary') +
-ylab('Gini coefficient among benes. w/ >= 1 claim')
+ylab('Gini coefficient among benes. w/ >= 1 claim') +
+theme_minimal()
 ```
+
+# Resistance data
+
+```{r load_abg}
+resistance_groups = read_tsv('../states/resistance/resistance_groups.tsv')
+
+organisms = c('Escherichia coli',
+              'Proteus mirabilis', 'Enterobacter cloacae',
+              'Enterococcus faecium',
+              'Coagulase-negative staphylococci',
+              'Methicillin-resistant S. aureus (MRSA)',
+              'Methicillin-susceptible S. aureus (MSSA)',
+              'Streptococcus pneumoniae (non-meningitis)',
+              'Staphylococcus aureus')
+
+abg = read_tsv('../../../antibiogram/data/abg.tsv', col_types=cols(zipcode='c')) %>%
+  mutate(drug=tolower(drug), percent_nonsusceptible=100-percent_susceptible) %>%
+  left_join(hrr, by='zipcode') %>%
+  right_join(resistance_groups, by='drug') %>%
+  filter(bug %in% organisms) %>%
+  select(bug, drug_group, state, hrr, percent_nonsusceptible, n_isolates)
+
+# Keep only those bug/drug combinations that are in at least 20 states
+abg %<>%
+  select(bug, drug_group, state) %>%
+  distinct %>%
+  count(bug, drug_group) %>% rename(n_states=n) %>%
+  filter(n_states >= 20) %>%
+  left_join(abg, by=c('bug', 'drug_group'))
+
+summarize_abg = function(df, place_name, min_places=20) {
+    group_by_(df, 'bug', 'drug_group', place_name) %>%
+    summarize(mean_percent_nonsusceptible=weighted.mean(percent_nonsusceptible, sqrt(n_isolates)),
+              n_place_antibiograms=n()) %>%
+    mutate(place_weight=sqrt(n_place_antibiograms))
+}
+
+state_abg = summarize_abg(abg, 'state')
+hrr_abg = summarize_abg(abg, 'hrr')
+```
+
+I consider resistance among these organisms:
+
+`r organisms %>% as.list %>% pander`
+
+And I join the resistance drugs into classes:
+
+`r kable(resistance_groups)`
+
+I similarly joined the consumption drugs into similar classes.
+
+```{r do_models}
+pvalue_stars = function(x) {
+  case_when(x < 0.001 ~ '***',
+            x < 0.01 ~ '**',
+            x < 0.05 ~ '*',
+            x < 0.1 ~ 'o',
+            TRUE ~ '')
+}
+
+linear_model = function(df, dependent_var, explanatory_vars) {
+  f = paste0(dependent_var, ' ~ ', paste0(explanatory_vars, collapse = ' + ')) %>% formula
+  lm(formula=f, weights=place_weight, data=df) %>%
+    tidy %>%
+    select(term, estimate, p.value)
+}
+
+spearman_model = function(df, dependent_var, explanatory_var) {
+  cor.test(df[[dependent_var]], df[[explanatory_var]], method='spearman') %>%
+    tidy %>%
+    mutate(term=explanatory_var, method='spearman') %>%
+    select(term, estimate, p.value)
+}
+
+multivariate_term = 'nzgini'
+#multivariate_term = 'mean_fnz_residual'
+models = function(x) {
+  dep = 'mean_percent_nonsusceptible'
+  # swo> "mean" is a confusing name
+  bind_rows(
+    linear_model(x, dep, 'mean') %>% mutate(model='univariate'),
+    linear_model(x, dep, c('fnz', multivariate_term)) %>% mutate(model='multivariate'),
+    spearman_model(x, dep, 'mean') %>% mutate(model='spearman')
+  )
+}
+
+# swo> why do I need to redo this filtering?
+results = left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
+  group_by(bug, drug_group) %>%
+  do(models(.)) %>%
+  ungroup
+
+hrr_results = left_join(hrr_abg, hrr_ineq, by=c('hrr', 'drug_group')) %>%
+  group_by(bug, drug_group) %>%
+  do(models(.)) %>%
+  ungroup
+```
+
+## Spearman
+
+Significant results using Spearman correlations of mean percent nonsusceptible ~ claims per beneficiary:
+
+```{r spearman_models, cache=FALSE}
+results %>%
+  filter(model=='spearman') %>%
+  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
+  filter(p.adj < 0.05) %>%
+  select(bug, drug_group, estimate, p.adj) %>%
+  kable
+```
+
+And the same thing using HRRs:
+
+```{r hrr_spearman_models, cache=FALSE}
+hrr_results %>%
+  filter(model=='spearman') %>%
+  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
+  filter(p.adj < 0.05) %>%
+  select(bug, drug_group, estimate, p.adj) %>%
+  kable
+```
+
+## Univariate
+
+The model here is mean percent nonsusceptible ~ claims per beneficiary. The significant results are:
+
+```{r univariate_models, cache=FALSE}
+results %>%
+  filter(model=='univariate', term == 'mean') %>%
+  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
+  filter(p.adj < 0.05) %>%
+  select(bug, drug_group, estimate, p.adj) %>%
+  kable
+```
+
+And the same thing with HRRs:
+
+```{r hrr_univariate_models, cache=FALSE}
+hrr_results %>%
+  filter(model=='univariate', term == 'mean') %>%
+  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
+  filter(p.adj < 0.05) %>%
+  select(bug, drug_group, estimate, p.adj) %>%
+  kable
+```
+
+## Multivariate
+
+The model here is mean percent nonsusceptible ~ fraction nonzero + nonzero Gini.
+In some of the models, the nonzero Gini term was significant, but not after adjusting for multiple hypotheses.
+I show just those terms that met the unadjusted threshold:
+
+```{r multivariate_models, cache=FALSE}
+results %>%
+  filter(model=='multivariate', term==multivariate_term) %>%
+  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
+  filter(p.value < 0.05) %>%
+  select(bug, drug_group, estimate, p.value, p.adj) %>%
+  kable
+```
+
+When using the HRRs, some of the values remained significant even after correction. I show all those values that passed the unadjusted threshold (like above):
+
+```{r hrr_multivariate_models, cache=FALSE}
+hrr_results %>%
+  filter(model=='multivariate', term==multivariate_term) %>%
+  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
+  filter(p.value < 0.05) %>%
+  select(bug, drug_group, estimate, p.value, p.adj) %>%
+  kable
+```
+
+## Two metrics
+
+It turns out that the different drugs have different relationships with the two metrics. Each point shows the consumption characteristics for a drug group in a state.
+
+```{r nzgini_cpb_drug}
+cbbPalette <- c("#000000", "#D55E00", "#CC79A7", "#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2")
+state_ineq %>%
+  ggplot(aes(x=nzgini, y=mean, color=drug_group)) +
+  geom_point() +
+  xlab('nonzero Gini') + ylab('claims per beneficiary') +
+  theme_minimal() +
+  scale_color_manual(values=cbbPalette)
+```
+
+## Specific plots
+
+Quinolone consumption predicts *E. coli* quinolone resistance:
+
+```{r ecoli_quin, cache=FALSE}
+left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
+  left_join(regions, by='state') %>%
+  filter(bug=='Escherichia coli', drug_group=='quinolone') %>%
+  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region)) +
+  geom_smooth(aes(x=mean, y=mean_percent_nonsusceptible, color=NULL), method='lm', col='black') +
+  geom_point() +
+  xlab('quinolone claims per beneficiary') + ylab('mean percent nonsusceptible E. coli isolates') +
+  theme_minimal()
+```

### 2017-04-20 ms2/data/collect-data.R

@@ -5,6 +5,8 @@ fraction_nonzero = function(x) length(nonzero(x)) / length(x)
 
 pad0 = function(x, to) c(x, rep(0, to - length(x)))
 
+lm_residuals = function(y, x) residuals(lm(y ~ x))
+
 inequalities = function(x) {
   data_frame(total=sum(x),
              mean=mean(x),
@@ -17,10 +19,21 @@ consumption_groups = read_tsv('../consumption_groups.tsv')
 regions = read_tsv('../../../db/census-regions/census-regions.tsv') %>%
   select(state, state_abbreviation, region)
 
+hrr = read_tsv('../../../db/hrr/hrr.tsv') %>%
+  filter(year==2011) %>%
+  select(zipcode, hrr)
+
 bene = read_tsv('../../bene_2011.tsv') %>%
   filter(between(age, 66, 100), hmo_months==0) %>%
-  left_join(regions, by='state')
+  left_join(regions, by='state') %>%
+  left_join(hrr, by='zipcode')
 
+pde = read_tsv('../../pde_2011.tsv') %>%
+  rename(drug=antibiotic) %>%
+  left_join(consumption_groups, by='drug') %>%
+  group_by(bene_id, drug_group) %>%
+  summarize(n_claims=n()) %>%
+  ungroup
 
 # compute total claims separately from individual drug groups
 total_pde = pde %>%
@@ -29,9 +42,11 @@ total_pde = pde %>%
   right_join(bene, by='bene_id') %>%
   replace_na(list(n_claims=0)) %>%
   group_by(state) %>%
-  do(inequalities(.$n_claims))
+  do(inequalities(.$n_claims)) %>%
+  ungroup() %>%
+  mutate(mean_fnz_residual=lm_residuals(mean, fnz))
 
-write_tsv(total_pde, 'total_pde.tsv')
+write_tsv(total_pde, 'all_drug_ineq.tsv')
 
 state_denom = bene %>% count(state) %>% rename(n_bene=n)
 state_inequality = pde %>%
@@ -39,6 +54,33 @@ state_inequality = pde %>%
   filter(!is.na(drug_group)) %>%
   left_join(state_denom, by='state') %>%
   group_by(state, drug_group) %>%
-  do(inequalities(pad0(.$n_claims, unique(.$n_bene))))
[…]

### 2017-04-28 ms1/ms1.Rmd
ms1

@@
 
-summarize_by = function(usage, by) {
-  group_by_(usage, by) %>%
+summarize_by = function(bene, by) {
+  group_by_(bene, by) %>%
   summarize(n_claims=sum(n_claims), n_bene=n()) %>%
@@
 load_data = function(year) {
-  usage = read_tsv(sprintf('usage_%i.tsv', year))
-  pde = read_tsv(sprintf('pde_%i.tsv', year))
+  bene = read_tsv(sprintf('../bene_%i.tsv', year)) %>%
+    filter(between(age, 66, 96), hmo_months==0) %>%
+    mutate(is_female=sex=='female', is_white=race=='white') %>%
+    left_join(regions, by='state')
 
-  list(usage=usage, pde=pde)
+  pde = read_tsv(sprintf('../pde_%i.tsv', year))
+
+  # get total numbers of PDEs
+  n_claims = pde %>% count(bene_id) %>% rename(n_claims=n)
+  bene %<>%
+    left_join(n_claims, by='bene_id') %>%
+    replace_na(list(n_claims=0))
+
+  cc = read_feather(sprintf('../cc_%i.feather', year)) %>%
+    mutate(n_cc=rowSums(select_(., '-bene_id')))
+
+  bene %<>% left_join(cc, by='bene_id')
+
+  cu = read_tsv('../cu_%i.tsv') %>%
+    select(bene_id, n_visits=hospital_op_visits)
+
+  bene %<>% left_join(cu, by='bene_id')
+
+  list(bene=bene, pde=pde)
 }
@@
   filter(x, age >= 67) %>%
-  mutate(age_over_67 = age - 67) %>%
   rename_(.dots=setNames(condition, 'has_condition')) %>%
-    glm(n_claims ~ age_over_67 + is_female + has_condition,
+    glm(n_claims ~ age + is_female + has_condition,
         family=quasipoisson(link='identity'), data=.) %>%
@@
 
-single_condition_models = function(usage) {
-  lapply(conditions, function(cond) single_condition_model(usage, cond)) %>%
+single_condition_models = function(bene) {
+  lapply(conditions, function(cond) single_condition_model(bene, cond)) %>%
     bind_rows()
@@
 ```{r analyze_function}
-age_group_f = function(x) bounds_group(x, c(65, 70, 75, 80, 85, 90, 95, 100))
+age_group_f = function(x) bounds_group(x, c(65, 70, 75, 80, 85, 90, 96))
 claims_group_f = function(x) bounds_group(x, c(0, 1, 2, 5, 10, 20, Inf))
@@
 
-  total_n_bene = nrow(dat$usage)
+  total_n_bene = nrow(dat$bene)
   total_n_pde = nrow(dat$pde)
@@
   # in the same way as the consumption data, so we can use the summarize_by function
-  claims_by_sex = summarize_by(dat$usage, 'is_female')
+  claims_by_sex = summarize_by(dat$bene, 'is_female')
 
@@
   # not just by raw age)
-  claims_by_age = dat$usage %>%
-    group_by(age) %>%
+  claims_by_age = dat$bene %>%
+    bind_rows(. %>% mutate(group='all'),
+              . %>% filter(n_cc==0) %>% mutate(group='no_cc'),
+              . %>% filter(n_visits==0) %>% mutate(group='no_visits')) %>% 
+    group_by(group, age) %>%
     summarize(n_bene=n(),
@@
               sd_n_claims=sd(n_claims),
-              mean_n_claims_0cc=mean(n_claims[n_cc==0]),
-              sd_n_claims_0cc=sd(n_claims[n_cc==0]),
               mean_n_cc=mean(n_cc),
-              sd_n_cc=sd(n_cc)) %>%
+              sd_n_cc=sd(n_cc),
+              mean_n_visits=mean(n_visits),
+              sd_n_visits=sd(n_visits)) %>%
     mutate(sem_n_claims=sd_n_claims/sqrt(n_bene),
-           sem_n_claims_0cc=sd_n_claims_0cc/sqrt(n_bene),
-           sem_n_cc=sd_n_cc/sqrt(n_bene))
+           sem_n_cc=sd_n_cc/sqrt(n_bene),
+           sem_n_visits=sd_n_visits/sqrt(n_bene))
 
-  chars_by_age_group = dat$usage %>%
+  chars_by_age_group = dat$bene %>%
     mutate(age_group=age_group_f(age)) %>%
@@
 
-  claims_by_region = summarize_by(dat$usage, 'region')
+  claims_by_region = summarize_by(dat$bene, 'region')
 
-  chars_by_consumption = dat$usage %>%
+  claims_by_race = dat$bene %>%
+    mutate(race=if_else(race %in% c('white', 'black', 'Hispanic'), race, 'other')) %>%
+    summarize_by('race')
+
+  chars_by_consumption = dat$bene %>%
     mutate(claims_group=claims_group_f(n_claims)) %>%
@@
 
-  condition_model = single_condition_models(dat$usage)
+  condition_model = single_condition_models(dat$bene)
 
@@
        claims_by_sex=claims_by_sex,
+       claims_by_race=claims_by_race,
        claims_by_age=claims_by_age,
@@
   write_tsv('tables/table2_sex.tsv') %>%
+  kable
+
+claims_by_race = data_by_field('claims_by_race') %>%
+  mutate(cpkp=round(cpkp)) %>%
+  spread(year, cpkp) %T>%
+  write_tsv('tables/tmp_race.tsv') %>%
   kable
@@
 
+## Consumption by race
+
+`r claims_by_race`
+
 ## Consumption by age
@@
              ymin=mean_n_claims-sem_n_claims,
-             ymax=mean_n_claims+sem_n_claims)) +
+             ymax=mean_n_claims+sem_n_claims,
+             color=group)) +
   geom_point() +
-  geom_errorbar() +
+  geom_linerange() +
   theme_minimal() +
-  scale_x_continuous(breaks=c(66, 70, 80, 90, 100))
+  scale_x_continuous(breaks=c(66, 70, 80, 90, 96))
 ```
@@
 
-```{r}
-f = function(df) {
-  stopifnot(nrow(df) == 1)
-  data_frame(group=c('all', '0cc'),
-             mean_n_claims=c(df$mean_n_claims, df$mean_n_claims_0cc),
-             sem_n_claims=c(df$sem_n_claims, df$sem_n_claims_0cc))
-}
-
+```{r stratified_claims_by_age}
 claims_by_age %>%
@@
              group=group)) +
-  geom_point() +
+  geom_point(shape=1, cex=5) +
   geom_smooth(method='lm', se=FALSE) +
-  theme_minimal() +
+  theme_minimal(base_size=24) +
   ylab('claims per person') +
@@
   scale_y_continuous(breaks=c(0, 0.5, 1, 1.5, 2.0), limits=c(0, 2))
+
+claims_by_age %>%
+  filter(year==2011, between(age, 67, 96)) %>%
+  ggplot(aes(x=age, y=mean_n_cc)) +
+  geom_point(shape=1, size=6) +
+  geom_smooth(method='lm', se=FALSE, col='black') +
+  theme_minimal(base_size=24) +
+  scale_x_continuous(breaks=c(67, 70, 80, 90, 96)) +
+  ylab('mean no. chronic conditions') +
+  ylim(0, 6)
 ```

### 2017-04-28 ms1/ms1_work.R

@@ -1,0 +1,83 @@
+qplm = function(df, indep, expl) {
+  f = formula(paste0(indep, ' ~ ', paste0(expl, collapse=' + ')))
+  glm(f, data=df, family=quasipoisson(link='identity'), start=rep(0.1, length(labels(terms(f))) + 1))
+}
+
+# Quasi-Poisson (w/ identity link) partial correlation
+qppc = function(df, y, x, controls=NULL) {
+  stopifnot(is.character(y) && is.character(x))
+  stopifnot(is.null(controls) || is.vector(controls, mode='character'))
+  if (is.null(controls) || length(controls) == 0) {
+    # this is just a normal correlation
+    cor.test(df[[x]], df[[y]]) %>% tidy
+  } else {
+    ry = qplm(df, y, controls)$residuals
+    rx = qplm(df, x, controls)$residuals
+    cor.test(rx, ry) %>% tidy
+  }
+}
+
+qppr = function(df, y, x, controls) {
+  stopifnot(is.character(y) && is.character(x))
+  stopifnot(is.vector(controls, mode='character'))
+  
+  ry = qplm(df, y, controls)$residuals
+  rx = qplm(df, x, controls)$residuals
+  lm(ry ~ rx) %>% tidy
+}
+
+bene = read_tsv('../bene_2011.tsv') %>%
+  filter(between(age, 66, 96), hmo_months==0) %>%
+  mutate(is_female=sex=='female', is_white=race=='white') %>%
+  select(-zipcode, -hmo_months, -sex, -race)
+
+cu = read_tsv('../cu_2011.tsv') %>%
+  filter(bene_id %in% bene$bene_id)
+
+cc = read_feather('../cc_2011.feather') %>%
+  filter(bene_id %in% bene$bene_id)
+
+pde = read_tsv('../pde_2011.tsv') %>%
+  filter(bene_id %in% bene$bene_id) %>%
+  count(bene_id) %>%
+  rename(abx_pde=n)
+
+dat = bene %>%
+  left_join(pde, by='bene_id') %>%
+  left_join(cu, by='bene_id') %>%
+  left_join(cc, by='bene_id') %>%
+  replace_na(list(abx_pde=0))
+
+ncc = dat %>% select(AMI:HYPOTH) %>% rowSums
+dat %<>% mutate(ncc=ncc)
+
+f = function(df, group_name) {
+  group_by(df, age) %>%
[…]

### 2017-04-28 ms2/ms2.Rmd
updated

@@
 knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/', dev=c('png', 'pdf'),
+                      pdf.options(useDingbats=FALSE),
                       echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
@@
 geom_point() +
-geom_label(nudge_x=0.05) +
+geom_text(nudge_x=0.04) +
 xlab('claims per beneficiary') +
@@
 
+And what about looking at multivariate only for ones that passed the threshold for univariate? Doesn't seem like it works for any of them!
+
+```{r}
+sig_bug_drug = results %>%
+  filter(model=='univariate', term=='mean') %>%
+  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
+  filter(p.adj < 0.05) %>%
+  select(bug, drug_group)
+
+results %>%
+  filter(model=='multivariate', term==multivariate_term) %>%
+  right_join(sig_bug_drug, by=c('bug', 'drug_group')) %>%
+  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
+  select(bug, drug_group, estimate, p.value, p.adj) %>%
+  kable
+```
+
 When using the HRRs, some of the values remained significant even after correction. I show all those values that passed the unadjusted threshold (like above):
@@
   theme_minimal()
+
+left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
+  left_join(regions, by='state') %>%
+  filter(bug=='Escherichia coli', drug_group=='quinolone') %>%
+  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, label=state)) +
+  geom_smooth(method='lm', col='black', se=FALSE) +
+  geom_point() +
+  geom_text() +
+  xlim(0.2, 0.45) + ylim(0, 50) +
+  xlab('quinolone claims per beneficiary') + ylab('mean percent nonsusceptible E. coli isolates') +
+  theme_minimal()
 ```
+
+Bactrim resistance in *E. coli*. First, the aggregate relationship, then the relationship between "excess" resistance (the residual of resistance against aggregate consumption) against nonzero Gini.
+
+```{r ecoli_bactrim}
+ecoli_bactrim_dat = left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
+  left_join(regions, by='state') %>%
+  filter(bug=='Escherichia coli', drug_group=='tmp_smx') %>%
+  filter(!is.na(mean_percent_nonsusceptible))
+
+ecoli_bactrim_dat %>%
+  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, label=state)) +
+  geom_smooth(method='lm', col='gray', se=FALSE) +
+  geom_point() +
+  geom_text() +
+  scale_x_continuous(limits=c(0.08, 0.2), breaks=c(0.08, 0.1, 0.15, 0.2)) +
+  ylim(10, 45) +
+  xlab('TMP/SMX claims per beneficiary') + ylab('mean percent nonsusceptible E. coli isolates') +
+  theme_minimal()
+
+ecoli_bactrim_dat %>%
+  { mutate(., excess_resistance=residuals(lm(mean_percent_nonsusceptible ~ mean, data=.))) } %>%
+  ggplot(aes(x=nzgini, y=excess_resistance)) +
+  geom_smooth(method='lm', col='gray', se=FALSE) +
+  geom_point() +
+  scale_x_continuous(limits=c(0.25, 0.375)) +
+  scale_y_continuous(limits=c(-12, 20), breaks=c(-12, 0, 10, 20)) +
+  xlab('Skew towards high consumers (Gini coefficient)') +
+  ylab('"Excess" E. coli resistance (see caption)') +
+  theme_minimal()
+```

### 2017-04-28 polish-cc.R

@@ -12,7 +12,7 @@ polish_cc = function(x) {
 polish_year = function(y) {
   cc_fn = sprintf('../data/cc_%i.tsv', y)
   bene_fn = sprintf('bene_%i.tsv', y)
-  cc_out_fn = sprintf('cc_%i.tsv', y)
+  cc_out_fn = sprintf('cc_%i.feather', y)
 
   bene_ids = read_tsv(bene_fn) %$% bene_id
 
@@ -20,7 +20,7 @@ polish_year = function(y) {
     polish_cc() %>%
     filter(bene_id %in% bene_ids)
 
-  write_tsv(cc, cc_out_fn)
+  write_feather(cc, cc_out_fn)
 }
 
 for (year in 2011:2014) polish_year(year)

### 2017-05-01 ms1/ms1.Rmd
new model, fig, table

@@
   cc = read_feather(sprintf('../cc_%i.feather', year)) %>%
+    select(-ALZH) %>%
     mutate(n_cc=rowSums(select_(., '-bene_id')))
@@
 
-  cu = read_tsv('../cu_%i.tsv') %>%
+  cu = read_tsv(sprintf('../cu_%i.tsv', year)) %>%
     select(bene_id, n_visits=hospital_op_visits)
@@
 ```{r single_condition_model}
-single_condition_model = function(x, condition) {
+single_condition_model = function(x, condition_code) {
   filter(x, age >= 67) %>%
-  rename_(.dots=setNames(condition, 'has_condition')) %>%
-    glm(n_claims ~ age + is_female + has_condition,
+  rename_(.dots=setNames(condition_code, 'has_condition')) %>%
+    glm(n_claims ~ age + is_female + has_condition + n_visits,
         family=quasipoisson(link='identity'), data=.) %>%
     tidy() %>%
-    mutate(condition=condition)
+    mutate(condition=condition_code)
 }
@@
 }
+
+all_condition_model = function(bene) {
+  f = formula(paste0('n_claims ~ age + is_female + n_visits + ', paste0(conditions, collapse=' + ')))
+  start = rep(0.1, length(all.vars(f)))
+  filter(bene, age >= 67) %>%
+    glm(f, family=quasipoisson(link='identity'), start=start, data=.) %>%
+    tidy
+}
+
 ```
@@
   claims_by_age = dat$bene %>%
-    bind_rows(. %>% mutate(group='all'),
-              . %>% filter(n_cc==0) %>% mutate(group='no_cc'),
-              . %>% filter(n_visits==0) %>% mutate(group='no_visits')) %>% 
+    mutate(group='all') %>%
+    bind_rows(filter(., n_cc==0) %>% mutate(group='no_cc'),
+              filter(., n_visits==0) %>% mutate(group='no_visits')) %>% 
     group_by(group, age) %>%
@@
 
+  big_model = all_condition_model(dat$bene)
+
   list(totals=totals,
@@
        chars_by_consumption=chars_by_consumption,
-       condition_model=condition_model) %>%
+       condition_model=condition_model,
+       big_model=big_model) %>%
     lapply(function(x) mutate(x, year=year))
@@
 
-```{r arrange_data}
+```{r arrange_data1}
 data_by_field = function(field) lapply(results, function(x) x[[field]]) %>% bind_rows()
@@
   mutate(cpkp=round(cpkp)) %>%
-  spread(year, cpkp) %T>%
+  spread(year, cpkp) %>%
+  arrange(desc(`2011`)) %T>%
   write_tsv('tables/tmp_race.tsv') %>%
@@
   kable
+```
 
+```{r arrange_data2}
 condition_model_all_years = data_by_field('condition_model') %>%
@@
   kable
+
+big_model = data_by_field('big_model') %>%
+  mutate(term=stringr::str_replace(term, 'TRUE$', '')) %>%
+  left_join(rename(condition_names, term=condition, condition=condition_name), by='term') %>%
+  filter(term != '(Intercept)') %>%
+  mutate(term=if_else(is.na(condition), term, condition)) %>%
+  select(year, term, estimate, std.error) %T>%
+  write_tsv('tables/tmp_big_model.tsv')
 ```
@@
 ```{r fig1_claims_by_age}
+# Show three lines: claims among all beneficiaries, among those with no ccs,
+# and those with no OP visits. Exclude the no_cc, age=66 datapoint, which is
+# artificially high.
 claims_by_age %>%
-  filter(year==2011) %>%
+  filter(year==2011, !(group=='no_cc' & age < 67)) %>%
   ggplot(aes(x=age,
@@
 claims_by_age %>%
-  filter(year==2011) %>%
+  filter(year==2011, age >= 67) %>%
   ggplot(aes(x=age,
@@
   geom_point() +
-  geom_errorbar() +
+  geom_linerange() +
   theme_minimal() +
-  scale_x_continuous(breaks=c(66, 70, 80, 90, 100))
+  scale_x_continuous(breaks=c(67, 70, 80, 90, 96)) +
+  ylim(0, 6)
 ```
 
-```{r stratified_claims_by_age}
+```{r visits_by_age}
 claims_by_age %>%
-  filter(year==2011, between(age, 66, 96)) %>%
-  select(age, mean_n_claims, mean_n_claims_0cc) %>%
-  gather('group', 'mean_n_claims', mean_n_claims, mean_n_claims_0cc) %>%
+  filter(year==2011) %>%
   ggplot(aes(x=age,
-             y=mean_n_claims,
-             color=group,
-             group=group)) +
-  geom_point(shape=1, cex=5) +
-  geom_smooth(method='lm', se=FALSE) +
-  theme_minimal(base_size=24) +
-  ylab('claims per person') +
-  scale_x_continuous(breaks=c(66, 70, 80, 90, 96)) +
-  scale_y_continuous(breaks=c(0, 0.5, 1, 1.5, 2.0), limits=c(0, 2))
-
-claims_by_age %>%
-  filter(year==2011, between(age, 67, 96)) %>%
-  ggplot(aes(x=age, y=mean_n_cc)) +
-  geom_point(shape=1, size=6) +
-  geom_smooth(method='lm', se=FALSE, col='black') +
-  theme_minimal(base_size=24) +
-  scale_x_continuous(breaks=c(67, 70, 80, 90, 96)) +
-  ylab('mean no. chronic conditions') +
-  ylim(0, 6)
+             y=mean_n_visits,
+             ymin=mean_n_visits-sem_n_visits,
+             ymax=mean_n_visits+sem_n_visits)) +
+  geom_point() +
+  geom_linerange() +
+  theme_minimal() +
+  scale_x_continuous(breaks=c(66, 70, 80, 90, 96))
 ```

### 2017-05-02 ms1/ms1.Rmd
plr

@@
 }
-
 ```
@@
              color=group)) +
+  geom_smooth(method='lm', se=F, size=0.5) +
   geom_point() +
@@
 claims_by_age %>%
-  filter(year==2011, age >= 67) %>%
+  filter(year==2011, age >= 67, group != 'no_cc') %>%
   ggplot(aes(x=age,
@@
              ymin=mean_n_cc-sem_n_cc,
-             ymax=mean_n_cc+sem_n_cc)) +
+             ymax=mean_n_cc+sem_n_cc,
+             color=group)) +
   geom_point() +
@@
 claims_by_age %>%
-  filter(year==2011) %>%
+  filter(year==2011, group != 'no_visits') %>%
   ggplot(aes(x=age,
@@
              ymin=mean_n_visits-sem_n_visits,
-             ymax=mean_n_visits+sem_n_visits)) +
+             ymax=mean_n_visits+sem_n_visits,
+             color=group)) +
   geom_point() +

### 2017-05-02 ms1/supp/supp_plr.R

@@ -1,0 +1,67 @@
+plr = function(test, disease) {
+  # positive likelihood ratio:
+  # P(T+ | D+) / P(T+ | D-)
+  # = N_true_pos * N_false / (N_false_pos * N_true)
+  
+  nsum = function(x) as.numeric(sum(x)) # to prevent overflow
+  n_true_pos = nsum(test & disease)
+  n_false_pos = nsum(test & !disease)
+  n_true = nsum(disease)
+  n_false = nsum(!disease)
+  
+  #n_true_pos * n_true / (n_false_pos * n_false)
+  n_true_pos * n_false / (n_false_pos * n_true)
+}
+
+plr_matrix = function(df) {
+  crossing(test_condition=names(df), true_condition=names(df)) %>%
+    filter(test_condition != true_condition) %>%
+    rowwise %>%
+    mutate(PLR=plr(df[[test_condition]], df[[true_condition]]))
+}
+
+condition_names = read_tsv('../../chronic-conditions/names.txt',
+                           col_names=c('condition', 'condition_name'))
+
+# load data, keeping only the CC data for the qualifying beneficiaries
+bene_ids = read_tsv('../../bene_2011.tsv') %>%
+  filter(between(age, 67, 96), hmo_months==0) %$%
+  bene_id
+
+cc = read_feather('../../cc_2011.feather') %>%
+  filter(bene_id %in% bene_ids) %>%
+  select(-ALZH)
+
+m = cc %>% select(-bene_id) %>% plr_matrix
+
+conditions = c("AMI", "ISCHMCHT", "CHF", "HYPERT", "CHRNKIDN", "STRKETIA", "HYPERL", 
+               "COPD", "ASTHMA", "CNCRLUNG",
+               "HIPFRAC", "OSTEOPRS", "ANEMIA", "DEPRESSN", 
+               "ALZHDMTA", "DIABETES", "ATRIALFB", "RA_OA", 
+               "HYPOTH", "CNCRBRST",  "CNCRENDM",  
+               "CNCRCLRC", "CNCRPRST", "HYPERP",
+               "CATARACT", "GLAUCOMA")
+male_conditions = c('CNCRPRST', 'HYPERP')
+female_conditions = c('CNCRENDM', 'CNCRBRST', 'OSTEOPRS', 'HYPOTH')
+
+plot_f = function(cc, excluded_conditions) {
+  levels = setdiff(conditions, excluded_conditions)
[…]

### 2017-05-02 ms1/ms1.Rmd
ignore feather

@@
 
-  pde = read_tsv(sprintf('../pde_%i.tsv', year))
+  pde = read_tsv(sprintf('../pde_%i.tsv', year)) %>%
+    filter(bene_id %in% bene$bene_id)
 
@@
 
+add_ci = function(df) mutate(df, ci=1.96*std.error, cil=estimate-ci, ciu=estimate+ci)
+
 big_model = data_by_field('big_model') %>%
@@
   mutate(term=if_else(is.na(condition), term, condition)) %>%
-  select(year, term, estimate, std.error) %T>%
-  write_tsv('tables/tmp_big_model.tsv')
+  mutate(adj.p=p.adjust(p.value, method='BH')) %>%
+  select(year, term, estimate, std.error, adj.p) %T>%
+  write_tsv('tables/tmp_big_model.tsv') %>%
+  add_ci() %>%
+  filter(year==2011) %>%
+  select(-year, -ci) %>%
+  arrange(desc(estimate)) %>%
+  kable
+```
+
+# Abstract's linear model results
+
+```{r ci}
+extract_ci = function(y) {
+  f = formula(paste0(y, ' ~ age'))
+  claims_by_age %>%
+    filter(group=='all') %>%
+    lm(f, data=.) %>%
+    tidy %>%
+    filter(term=='age') %>%
+    add_ci() %>%
+    mutate(model=y) %>%
+    select(model, estimate, cil, ciu) %>%
+    mutate_if(is.numeric, function(x) round(x, digits=4))
+}
+
+bind_rows(lapply(c('mean_n_claims', 'mean_n_cc', 'mean_n_visits'), extract_ci)) %>% kable
 ```
@@
 
+`r big_model`
+
 ## Table 3 caption: model coefficient p-values

### 2017-05-02 ms1/polish-usage.R

@@ -1,38 +1,0 @@
-regions = read_tsv('../../db/census-regions/census-regions.tsv') %>%
-  select(state, region)
-
-antibiotic_class = read_tsv('../abx_class.tsv') %>%
-  select(antibiotic, antibiotic_class) %>%
-  distinct
-
-analyze = function(year) {
-  cc = read_tsv(sprintf('../cc_%i.tsv', year)) %>%
-    select(-ALZH)
-
-  n_cc = cc %>% select(AMI:HYPOTH) %>% rowSums
-
-  bene = read_tsv(sprintf('../bene_%i.tsv', year)) %>%
-    filter(between(age, 66, 100), hmo_months==0) %>%
-    mutate(is_female=sex=='female', is_white=race=='white') %>%
-    left_join(regions, by='state') %>%
-    left_join(cc, by='bene_id') %>%
-    select(-sex, -race, -zipcode, -hmo_months)
-
-  n_cc = bene %>% select(AMI:HYPOTH) %>% rowSums
-
-  bene %<>% mutate(n_cc=n_cc)
-
-  pde = read_tsv(sprintf('../pde_%i.tsv', year)) %>%
-    semi_join(bene, by='bene_id') %>%
-    mutate(year=year)
-
-  usage = pde %>% count(bene_id) %>% rename(n_claims=n) %>%
-    right_join(bene, by='bene_id') %>%
-    replace_na(list(n_claims=0)) %>%
-    mutate(year=year)
-
-  write_tsv(usage, sprintf('usage_%i.tsv', year))
-  write_tsv(pde, sprintf('pde_%i.tsv', year))
-}
-
-for (y in 2011:2014) analyze(y)

### 2017-05-02 ms2/ms2.Rmd
moving to hrr

@@
 
+zipcodes = read_tsv('../../db/zipcode/zipcode.tsv') %>%
+  select(zipcode=zip, state)
+
 hrr = read_tsv('../../db/hrr/hrr.tsv') %>%
   filter(year==2011) %>%
-  select(zipcode, hrr)
+  select(zipcode, hrr) %>%
+  left_join(zipcodes, by='zipcode') %>%
+  left_join(regions, by='state')
 
@@
   mutate(drug=tolower(drug), percent_nonsusceptible=100-percent_susceptible) %>%
-  left_join(hrr, by='zipcode') %>%
+  left_join(hrr, by=c('zipcode', 'state')) %>%
   right_join(resistance_groups, by='drug') %>%
@@
 
-# Keep only those bug/drug combinations that are in at least 20 states
-abg %<>%
-  select(bug, drug_group, state) %>%
+possible_bug_drug = read_tsv('bug_drug.tsv')
+
+# Keep only bug/drug combinations that are in at least 75 HRRs
+good_combos = possible_bug_drug %>%
+  left_join(abg, by=c('bug', 'drug_group')) %>%
+  select(bug, drug_group, hrr) %>%
   distinct %>%
-  count(bug, drug_group) %>% rename(n_states=n) %>%
-  filter(n_states >= 20) %>%
-  left_join(abg, by=c('bug', 'drug_group'))
+  count(bug, drug_group) %>% rename(n_places=n) %>%
+  filter(n_places >= 75)
 
-summarize_abg = function(df, place_name, min_places=20) {
+abg %<>% right_join(good_combos, by=c('bug', 'drug_group'))
+
+summarize_abg = function(df, place_name) {
     group_by_(df, 'bug', 'drug_group', place_name) %>%
@@
 
-I consider resistance among these organisms:
-
-`r organisms %>% as.list %>% pander`
-
 And I join the resistance drugs into classes:
@@
 
-I similarly joined the consumption drugs into similar classes.
+I similarly joined the consumption drugs into similar classes. I'm looking at these `r nrow(good_combos)` bug/drug combinations:
+
+`r good_combos %>% arrange(desc(n_places)) %>% kable`
 
 ```{r do_models}
-pvalue_stars = function(x) {
-  case_when(x < 0.001 ~ '***',
-            x < 0.01 ~ '**',
-            x < 0.05 ~ '*',
-            x < 0.1 ~ 'o',
-            TRUE ~ '')
-}
-
 linear_model = function(df, dependent_var, explanatory_vars) {
@@
 multivariate_term = 'nzgini'
-#multivariate_term = 'mean_fnz_residual'
 models = function(x) {
@@
     spearman_model(x, dep, 'mean') %>% mutate(model='spearman')
-  )
+  ) %>%
+    mutate(n_data=nrow(x))
 }
@@
 
-```{r spearman_models, cache=FALSE}
+```{r spearman_models, cache=FALSE, eval=FALSE}
 results %>%
@@
   mutate(p.adj=p.adjust(p.value, method='BH')) %>%
-  filter(p.adj < 0.05) %>%
-  select(bug, drug_group, estimate, p.adj) %>%
+  #filter(p.adj < 0.05) %>%
+  arrange(p.adj) %>%
+  select(bug, drug_group, estimate, p.adj, n_data) %>%
   kable
@@
 
-```{r univariate_models, cache=FALSE}
+```{r univariate_models, cache=FALSE, eval=FALSE}
 results %>%
@@
   mutate(p.adj=p.adjust(p.value, method='BH')) %>%
-  filter(p.adj < 0.05) %>%
+  #filter(p.adj < 0.05) %>%
+  arrange(p.adj) %>%
   select(bug, drug_group, estimate, p.adj) %>%
@@
 
-```{r multivariate_models, cache=FALSE}
+```{r multivariate_models, cache=FALSE, eval=FALSE}
 results %>%
@@
   mutate(p.adj=p.adjust(p.value, method='BH')) %>%
-  filter(p.value < 0.05) %>%
+  #filter(p.value < 0.05) %>%
+  arrange(p.adj) %>%
   select(bug, drug_group, estimate, p.value, p.adj) %>%
@@
 
-```{r}
+```{r eval=FALSE}
 sig_bug_drug = results %>%
@@
 ```{r hrr_multivariate_models, cache=FALSE}
+sig_bug_drug = hrr_results %>%
+  filter(model=='univariate', term=='mean') %>%
+  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
+  filter(p.adj < 0.05) %>%
+  select(bug, drug_group)
+
 hrr_results %>%
   filter(model=='multivariate', term==multivariate_term) %>%
+  #right_join(sig_bug_drug, by=c('bug', 'drug_group')) %>%
   mutate(p.adj=p.adjust(p.value, method='BH')) %>%
-  filter(p.value < 0.05) %>%
+  #filter(p.value < 0.05) %>%
+  arrange(p.value) %>%
   select(bug, drug_group, estimate, p.value, p.adj) %>%
@@
 
+And for HRRs:
+
+```{r}
+left_join(hrr_abg, hrr_ineq, by=c('hrr', 'drug_group')) %>%
+  left_join(hrr, by='hrr') %>%
+  filter(bug=='Escherichia coli', drug_group=='quinolone') %>%
+  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region, size=sqrt(total))) +
+  geom_smooth(aes(x=mean, y=mean_percent_nonsusceptible, color=NULL), method='lm', col='black') +
+  geom_point() +
+  xlab('quinolone claims per beneficiary') + ylab('mean percent nonsusceptible E. coli isolates') +
+  theme_minimal()
+```
+
 Bactrim resistance in *E. coli*. First, the aggregate relationship, then the relationship between "excess" resistance (the residual of resistance against aggregate consumption) against nonzero Gini.

### 2017-05-02 ms2/work.R

@@ -1,34 +1,0 @@
-# > names(x)
-# [1] "state"    "cpb"      "v"        "fnznc"    "nzgini"   "n_bene"   "n_claims" "fz"      
-
-x = usage %>%
-  group_by(state) %>%
-  summarize(cpb=mean(n_claims),
-            v=var(n_claims),
-            fnznc=fnz(n_claims),
-            fz=1-fnznc,
-            nzgini=Gini(nz(n_claims)),
-            n_bene=n(),
-            n_claims=sum(n_claims))
-
-# fit the p value
-m = nls(fz ~ (1-p) ** ((1-p)/p*cpb), data=x, start=list(p=0.5))
-p = coef(m)[['p']]
-
-# or, do this with the variance
-# the coefficient of the variance v is (1-p)
-# this give 0.251 -> 0.749, which is higher than p=0.7 from above
-# lm(cpb ~ v + 0, data=x)
-
-# get the r values
-r = (1-p)/p * x$cpb
-
-gini_range = function(r, n_individuals, n_trials=1e2) {
-  values = replicate(n_trials, Gini(nz(rnbinom(n_individuals, r, p))))
-  data_frame(mu=mean(values), cil=quantile(values, 0.05), ciu=quantile(values, 0.95))
-}
-
-y = data_frame(r=r, n_bene=x$n_bene) %>%
-  group_by(r) %>%
-  do(gini_range(.$r, .$n_bene)) %>%
-  mutate(nbmu=p*r/(1-p), nbf0=(1-p)**r)

### 2017-05-03 ms1/ms1.Rmd
ms1

@@
 ```{r p_values}
-data_by_field('condition_model') %>%
+data_by_field('big_model') %>%
   select(term, year, p.value) %>%

### 2017-05-03 ms1/ms1_work.R

@@ -81,3 +81,26 @@ p = ggplot(x, aes(x=age, y=mean_abx_pde, group=group)) +
            size=5)
 
 show(p)
+
+conditions = c("AMI", "ISCHMCHT", "CHF", "HYPERT", "CHRNKIDN", "STRKETIA", "HYPERL", 
+               "COPD", "ASTHMA", "CNCRLUNG",
+               "HIPFRAC", "OSTEOPRS", "ANEMIA", "DEPRESSN", 
+               "ALZHDMTA", "DIABETES", "ATRIALFB", "RA_OA", 
+               "HYPOTH", "CNCRBRST",  "CNCRENDM",  
+               "CNCRCLRC", "CNCRPRST", "HYPERP",
+               "CATARACT", "GLAUCOMA")
+bene_ids = read_tsv('../bene_2011.tsv')$bene_id
+cc = read_feather('../cc_2011.feather') %>%
+  filter(bene_id %in% bene_ids)
+
+f = function(c1, c2) {
+  cc %>% group_by_(c1, c2) %>% summarize(n=n()) %>% ungroup %$% min(n)
+}
+
+g = function(c1, c2) sum(cc[[c1]] & cc[[c2]]) / sum(xor(cc[[c1]], cc[[c2]]))
+
+x = crossing(c1=conditions, c2=conditions) %>%
+  filter(c1 < c2) %>%
+  rowwise() %>%
+  mutate(diag_n=g(c1, c2)) %>%
+  ungroup()

### 2017-05-03 ms2/ms2.Rmd
modernizing ms2
- Use states mainly (HRRs are more noisy)
- Use a short-list of bug/drug candidates
- Collect data from all years

@@
 ```{r load_abg}
-resistance_groups = read_tsv('../states/resistance/resistance_groups.tsv')
+resistance_groups = read_tsv('resistance_groups.tsv')
 
@@
 
-possible_bug_drug = read_tsv('bug_drug.tsv')
+possible_bug_drug = read_tsv('bug_drug2.tsv')
 
 # Keep only bug/drug combinations that are in at least 75 HRRs
+min_n_states = 35
 good_combos = possible_bug_drug %>%
   left_join(abg, by=c('bug', 'drug_group')) %>%
-  select(bug, drug_group, hrr) %>%
+  select(bug, drug_group, state) %>%
   distinct %>%
   count(bug, drug_group) %>% rename(n_places=n) %>%
-  filter(n_places >= 75)
+  filter(n_places >= min_n_states)
 
@@
 
-I similarly joined the consumption drugs into similar classes. I'm looking at these `r nrow(good_combos)` bug/drug combinations:
+I similarly joined the consumption drugs into similar classes.
+
+I started by considering these bug/drug combinations:
+
+`r kable(possible_bug_drug)`
+
+Then I filtered to keep only those bug/drug combinations that were present in `r min_n_states` states. That means I'm looking at these `r nrow(good_combos)` bug/drug combinations:
 
@@
 
-# swo> why do I need to redo this filtering?
 results = left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
@@
 
-```{r spearman_models, cache=FALSE, eval=FALSE}
+```{r spearman_models, cache=FALSE}
 results %>%
@@
   arrange(p.adj) %>%
-  select(bug, drug_group, estimate, p.adj, n_data) %>%
+  select(bug, drug_group, estimate, p.value, p.adj, n_data) %>%
   kable
+```
+
+```{r}
+results %>%
+  filter(model=='spearman') %>%
+  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
+  ggplot(aes(x=estimate, y=-log10(p.adj))) + geom_point() +
+  geom_hline(yintercept=-log10(0.05), col='red')
 ```
@@
   arrange(p.adj) %>%
-  select(bug, drug_group, estimate, p.adj, n_data) %>%
+  select(bug, drug_group, estimate, p.value, p.adj, n_data) %>%
   kable
@@
 
-```{r univariate_models, cache=FALSE, eval=FALSE}
+```{r univariate_models, cache=FALSE}
 results %>%
@@
   arrange(p.adj) %>%
-  select(bug, drug_group, estimate, p.adj) %>%
+  select(bug, drug_group, estimate, p.value, p.adj) %>%
   kable
+```
+
+```{r}
+results %>%
+  filter(model=='univariate') %>%
+  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
+  ggplot(aes(x=estimate, y=-log10(p.adj))) + geom_point() +
+  geom_hline(yintercept=-log10(0.05), col='red')
 ```
@@
 
-```{r multivariate_models, cache=FALSE, eval=FALSE}
+```{r multivariate_models, cache=FALSE}
 results %>%

### 2017-05-03 ms2/data/collect-data.R

@@ -19,68 +19,81 @@ consumption_groups = read_tsv('../consumption_groups.tsv')
 regions = read_tsv('../../../db/census-regions/census-regions.tsv') %>%
   select(state, state_abbreviation, region)
 
-hrr = read_tsv('../../../db/hrr/hrr.tsv') %>%
-  filter(year==2011) %>%
-  select(zipcode, hrr)
+hrr_all = read_tsv('../../../db/hrr/hrr.tsv')
 
-bene = read_tsv('../../bene_2011.tsv') %>%
-  filter(between(age, 66, 100), hmo_months==0) %>%
-  left_join(regions, by='state') %>%
-  left_join(hrr, by='zipcode')
+summarize_inequality = function(year) {
+  hrr = hrr_all %>%
+    filter(year==year) %>%
+    select(zipcode, hrr)
 
-pde = read_tsv('../../pde_2011.tsv') %>%
-  rename(drug=antibiotic) %>%
-  left_join(consumption_groups, by='drug') %>%
-  group_by(bene_id, drug_group) %>%
-  summarize(n_claims=n()) %>%
-  ungroup
+  bene_fn = sprintf('../../bene_%i.tsv', year)
+  pde_fn = sprintf('../../pde_%i.tsv', year)
+  state_all_fn = sprintf('state_ineq_all_%i.tsv', year)
+  state_fn = sprintf('state_ineq_%i.tsv', year)
+  hrr_all_fn = sprintf('hrr_all_%i.tsv', year)
+  hrr_fn = sprintf('hrr_ineq_%i.tsv', year)
 
-# compute total claims separately from individual drug groups
-total_pde = pde %>%
-  group_by(bene_id) %>%
-  summarize(n_claims=sum(n_claims)) %>%
-  right_join(bene, by='bene_id') %>%
-  replace_na(list(n_claims=0)) %>%
-  group_by(state) %>%
-  do(inequalities(.$n_claims)) %>%
-  ungroup() %>%
-  mutate(mean_fnz_residual=lm_residuals(mean, fnz))
+  bene = read_tsv(bene_fn) %>%
+    filter(between(age, 66, 96), hmo_months==0) %>%
+    left_join(regions, by='state') %>%
+    left_join(hrr, by='zipcode')
 
-write_tsv(total_pde, 'all_drug_ineq.tsv')
+  pde = read_tsv(pde_fn) %>%
+    rename(drug=antibiotic) %>%
[…]

### 2017-05-03 ms2/ms2.Rmd
updating

@@
 ---
+
+**swo**: filter the NAs in the abg earlier
 
@@
 
-all_ineq = read_tsv('data/all_drug_ineq.tsv') %>%
+all_ineq = read_tsv('data/state_ineq_all_2011.tsv') %>%
   left_join(regions, by='state')
-state_ineq = read_tsv('data/drug_ineq.tsv')
-hrr_ineq = read_tsv('data/hrr_drug_ineq.tsv')
+state_ineq = read_tsv('data/state_ineq_2011.tsv')
+hrr_ineq = read_tsv('data/hrr_ineq_2011.tsv')
 ```
@@
 resistance_groups = read_tsv('resistance_groups.tsv')
-
-organisms = c('Escherichia coli',
-              'Proteus mirabilis', 'Enterobacter cloacae',
-              'Enterococcus faecium',
-              'Coagulase-negative staphylococci',
-              'Methicillin-resistant S. aureus (MRSA)',
-              'Methicillin-susceptible S. aureus (MSSA)',
-              'Streptococcus pneumoniae (non-meningitis)',
-              'Staphylococcus aureus')
+bug_names = read_tsv('bug_names.tsv')
 
@@
   right_join(resistance_groups, by='drug') %>%
-  filter(bug %in% organisms) %>%
   select(bug, drug_group, state, hrr, percent_nonsusceptible, n_isolates)
 
-possible_bug_drug = read_tsv('bug_drug2.tsv')
+possible_bug_drug = read_tsv('bug_drug.tsv')
 
@@
 
-abg %<>% right_join(good_combos, by=c('bug', 'drug_group'))
+abg %<>% right_join(good_combos, by=c('bug', 'drug_group')) %>%
+  rename(bug_long=bug) %>% left_join(bug_names, by='bug_long') %>% select(-bug_long)
+
+# check that all the short names made it in
+if (any(is.na(abg$bug))) stop('missing bug names')
 
@@
   left_join(regions, by='state') %>%
-  filter(bug=='Escherichia coli', drug_group=='quinolone') %>%
+  filter(bug=='E. coli', drug_group=='quinolone') %>%
   ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region)) +
@@
   left_join(regions, by='state') %>%
-  filter(bug=='Escherichia coli', drug_group=='quinolone') %>%
+  filter(bug=='E. coli', drug_group=='quinolone') %>%
   ggplot(aes(x=mean, y=mean_percent_nonsusceptible, label=state)) +
@@
   left_join(hrr, by='hrr') %>%
-  filter(bug=='Escherichia coli', drug_group=='quinolone') %>%
+  filter(bug=='E. coli', drug_group=='quinolone') %>%
   ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region, size=sqrt(total))) +
@@
   left_join(regions, by='state') %>%
-  filter(bug=='Escherichia coli', drug_group=='tmp_smx') %>%
+  filter(bug=='E. coli', drug_group=='tmp_smx') %>%
   filter(!is.na(mean_percent_nonsusceptible))
@@
 ```
+
+## E.coli and penicillin
+
+```{r}
+ecoli_pen_dat = left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
+  left_join(regions, by='state') %>%
+  filter(bug=='E. coli', drug_group=='penicillin') %>%
+  filter(!is.na(mean_percent_nonsusceptible))
+
+ecoli_bactrim_dat %>%
+  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, label=state)) +
+  geom_smooth(method='lm', col='gray', se=FALSE) +
+  geom_point() +
+  geom_text() +
+  #scale_x_continuous(limits=c(0.08, 0.2), breaks=c(0.08, 0.1, 0.15, 0.2)) +
+  #ylim(10, 45) +
+  #xlab('TMP/SMX claims per beneficiary') + ylab('mean percent nonsusceptible E. coli isolates') +
+  theme_minimal()
+```

### 2017-05-03 ms2/ms2.Rmd
streamlined

@@
 
-And for HRRs:
-
-```{r}
-left_join(hrr_abg, hrr_ineq, by=c('hrr', 'drug_group')) %>%
-  left_join(hrr, by='hrr') %>%
-  filter(bug=='E. coli', drug_group=='quinolone') %>%
-  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region, size=sqrt(total))) +
-  geom_smooth(aes(x=mean, y=mean_percent_nonsusceptible, color=NULL), method='lm', col='black') +
-  geom_point() +
-  xlab('quinolone claims per beneficiary') + ylab('mean percent nonsusceptible E. coli isolates') +
-  theme_minimal()
-```
+## E.coli and bactrim
 
@@
 
-## E.coli and penicillin
+## Pneumococcus and macrolides
 
 ```{r}
-ecoli_pen_dat = left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
+left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
   left_join(regions, by='state') %>%
-  filter(bug=='E. coli', drug_group=='penicillin') %>%
-  filter(!is.na(mean_percent_nonsusceptible))
-
-ecoli_bactrim_dat %>%
+  filter(bug=='S. pneumoniae', drug_group=='macrolide') %>%
+  filter(!is.na(mean_percent_nonsusceptible)) %>% 
   ggplot(aes(x=mean, y=mean_percent_nonsusceptible, label=state)) +
@@
   geom_text() +
-  #scale_x_continuous(limits=c(0.08, 0.2), breaks=c(0.08, 0.1, 0.15, 0.2)) +
-  #ylim(10, 45) +
-  #xlab('TMP/SMX claims per beneficiary') + ylab('mean percent nonsusceptible E. coli isolates') +
   theme_minimal()
 ```
+
+## E. cloacae and quinolones
+
+```{r}
+left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
+  left_join(regions, by='state') %>%
+  filter(bug=='Eb. cloacae', drug_group=='quinolone') %>%
+  filter(!is.na(mean_percent_nonsusceptible)) %>%
+  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, label=state)) +
+  geom_smooth(method='lm', col='gray', se=FALSE) +
+  geom_point() +
+  geom_text() +
+  theme_minimal()
+```

### 2017-05-04 ms2/ms2.Rmd
new plots ,etc.

@@
 
-```{r fnz-mean, cache=FALSE}
+```{r fnz_mean, cache=FALSE}
 ggplot(all_ineq, aes(x=mean, y=fnz, color=region, label=state_abbreviation)) +
@@
 
-```{r nzgini-mean, cache=FALSE}
+```{r nzgini_mean, cache=FALSE}
 ggplot(all_ineq, aes(x=mean, y=nzgini, color=region, label=state_abbreviation)) +
@@
 abg %<>% right_join(good_combos, by=c('bug', 'drug_group')) %>%
-  rename(bug_long=bug) %>% left_join(bug_names, by='bug_long') %>% select(-bug_long)
+  rename(bug_long=bug) %>% left_join(bug_names, by='bug_long') %>% select(-bug_long) %>%
+  filter(!is.na(percent_nonsusceptible), !is.na(n_isolates))
 
@@
               n_place_antibiograms=n()) %>%
+    #filter(!is.na(mean_percent_nonsusceptible)) %>%
     mutate(place_weight=sqrt(n_place_antibiograms))
@@
   mutate(p.adj=p.adjust(p.value, method='BH')) %>%
-  #filter(p.adj < 0.05) %>%
+  filter(p.value < 0.05) %>%
   arrange(p.adj) %>%
-  select(bug, drug_group, estimate, p.value, p.adj, n_data) %>%
+  select(bug, drug_group, estimate, p.value, p.adj) %>%
   kable
@@
   mutate(p.adj=p.adjust(p.value, method='BH')) %>%
-  #filter(p.adj < 0.05) %>%
-  arrange(p.adj) %>%
-  select(bug, drug_group, estimate, p.value, p.adj, n_data) %>%
+  filter(p.value < 0.05) %>%
+  arrange(p.value) %>%
+  select(bug, drug_group, estimate, p.value, p.adj) %>%
   kable
@@
   mutate(p.adj=p.adjust(p.value, method='BH')) %>%
-  #filter(p.adj < 0.05) %>%
-  arrange(p.adj) %>%
+  filter(p.value < 0.05) %>%
+  arrange(p.value) %>%
   select(bug, drug_group, estimate, p.value, p.adj) %>%
@@
   mutate(p.adj=p.adjust(p.value, method='BH')) %>%
-  #filter(p.adj < 0.05) %>%
-  arrange(p.adj) %>%
-  select(bug, drug_group, estimate, p.adj) %>%
+  filter(p.value < 0.05) %>%
+  arrange(p.value) %>%
+  select(bug, drug_group, estimate, p.value, p.adj) %>%
   kable
@@
   mutate(p.adj=p.adjust(p.value, method='BH')) %>%
-  #filter(p.value < 0.05) %>%
-  arrange(p.adj) %>%
+  filter(p.value < 0.05) %>%
+  arrange(p.value) %>%
   select(bug, drug_group, estimate, p.value, p.adj) %>%
@@
 
-And what about looking at multivariate only for ones that passed the threshold for univariate? Doesn't seem like it works for any of them!
+And what about looking at multivariate only for ones that passed the threshold for univariate?
 
@@
 
-When using the HRRs, some of the values remained significant even after correction. I show all those values that passed the unadjusted threshold (like above):
+Using HRRs, adjusting $p$ with all possibilities:
+
+```{r cache=FALSE}
+hrr_results %>%
+  filter(model=='multivariate', term==multivariate_term) %>%
+  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
+  filter(p.value < 0.05) %>%
+  arrange(p.value) %>%
+  select(bug, drug_group, estimate, p.value, p.adj) %>%
+  kable
+```
+
+And then doing multiple hypothesis correction only for those that made it through the first model:
 
@@
   filter(model=='multivariate', term==multivariate_term) %>%
-  #right_join(sig_bug_drug, by=c('bug', 'drug_group')) %>%
+  right_join(sig_bug_drug, by=c('bug', 'drug_group')) %>%
   mutate(p.adj=p.adjust(p.value, method='BH')) %>%
-  #filter(p.value < 0.05) %>%
   arrange(p.value) %>%
@@
 state_ineq %>%
-  ggplot(aes(x=nzgini, y=mean, color=drug_group)) +
+  ggplot(aes(x=mean, y=nzgini, color=drug_group)) +
   geom_point() +
-  xlab('nonzero Gini') + ylab('claims per beneficiary') +
+  ylab('Gini coefficient among users') + xlab('claims per beneficiary') +
   theme_minimal() +
@@
   left_join(regions, by='state') %>%
-  filter(bug=='E. coli', drug_group=='tmp_smx') %>%
-  filter(!is.na(mean_percent_nonsusceptible))
+  filter(bug=='E. coli', drug_group=='tmp_smx') #%>%
+  #filter(!is.na(mean_percent_nonsusceptible))
 
@@
 ```
+
+## MSSA and clindamycin
+
+```{r mssa_clinda}
+dat = left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
+  left_join(regions, by='state') %>%
+  filter(bug=='MSSA', drug_group=='clindamycin') #%>%
+  #filter(!is.na(mean_percent_nonsusceptible))
+
+dat %>%
+  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, label=state)) +
+  geom_smooth(method='lm', col='gray', se=FALSE) +
+  geom_point() +
+  geom_text() +
+  #scale_x_continuous(limits=c(0.08, 0.2), breaks=c(0.08, 0.1, 0.15, 0.2)) +
+  #ylim(10, 45) +
+  xlab('clindamycin claims per beneficiary') + ylab('mean percent nonsusceptible MSSA isolates') +
+  theme_minimal()
+
+dat %>%
+  { mutate(., excess_resistance=residuals(lm(mean_percent_nonsusceptible ~ mean, data=.))) } %>%
+  ggplot(aes(x=nzgini, y=excess_resistance)) +
+  geom_smooth(method='lm', col='gray', se=FALSE) +
+  geom_point() +
+  #scale_x_continuous(limits=c(0.25, 0.375)) +
+  #scale_y_continuous(limits=c(-12, 20), breaks=c(-12, 0, 10, 20)) +
+  xlab('Gini coefficient') +
+  ylab('"Excess" MSSA resistance') +
+  theme_minimal()
+```

### 2017-05-04 ms2/ms2.Rmd
fixed merging

@@
 
-**swo**: filter the NAs in the abg earlier
-
 ```{r global_options, include=FALSE}
@@
                       echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
-import::from(pander, pander)
 import::from(broom, tidy)
@@
 min_n_states = 35
-good_combos = possible_bug_drug %>%
-  left_join(abg, by=c('bug', 'drug_group')) %>%
+good_combos = semi_join(abg, possible_bug_drug, by=c('bug', 'drug_group')) %>%
   select(bug, drug_group, state) %>%

### 2017-05-10 ms1/ms1.Rmd
many updates

@@
                       echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE,
-                      error=FALSE)
+                      error=FALSE, autodep=TRUE)
 import::from(knitr, kable)

### 2017-05-10 ms2/power_calculation/spearman_power.py

@@ -9,7 +9,7 @@ import scipy.stats
 
 # Shouldn't matter what the exact sd's are
 sd1 = 1.0
-sd2 = 1.0
+sd2 = 5.0
 
 def power(size, rho, n_trials=1000):
     n_sig = 0

### 2017-05-10 ms2/power_calculation/spearman_power2.py

@@ -27,6 +27,6 @@ def p_value(sdi2, n):
     return scipy.stats.norm.sf(z) + scipy.stats.norm.cdf(-z)
 
 print('n', 'sdi2', 'r', 'z', 'p', sep='\t')
-for n in [30, 35, 40]:
+for n in [19, 30, 35, 40]:
     for sdi2 in range(0, max_sdi2(n) + 1, 100):
         print(n, sdi2, r(sdi2, n), z_score(sdi2, n), p_value(sdi2, n), sep='\t')

### 2017-05-10 temporal/temporal.R

@@ -1,14 +1,5 @@
 #!/usr/bin/env Rscript
 
-# create summary beneficiary x abx file
-
-# - load map from NDC codes to abx names
-# - load beneficiary data; keep bene_id -> state mapping
-# - load antibiotic PDEs; keep bene, date, abx, days
-# - merge abx and bene; keep date, abx, days, state
-# - summarize information by drug, state, and year
-# - summarize also by class?
-
 library(lubridate)
 
 summarize_weeks = function(x) {
@@ -20,10 +11,13 @@ summarize_weeks = function(x) {
 
 for (y in 2011:2014) {
   pde = sprintf('../pde_%i.tsv', y) %>% read_tsv
-  bene = sprintf('../bene_%i.tsv', y) %>% read_tsv
+  bene = sprintf('../bene_%i.tsv', y) %>%
+    read_tsv %>%
+    filter(between(age, 66, 96), sex %in% c('male', 'female'), hmo_months==0)
 
   pde %>%
     left_join(select(bene, bene_id, state), by='bene_id') %>%
+    filter(!is.na(state)) %>%
     summarize_weeks %>%
     write_tsv(sprintf('temporal_abx_%i.tsv', y))
 }

### 2017-05-10 temporal/seasonal-traces/drug_seasonality.Rmd
GISP support analysis

---
title: "Antibiotic seasonalities"
author: "Scott Olesen"
date: "5/10/2017"
output: html_document
---

# Background

For the GISP project, we are trying to show that we can infer bystander selection pressure on gonococcus using the seasonality of consumption of drugs that are consumed for reasons other than gonorrhea treatment.

# Methods

I use the 2011--2014 Medicare data, filtered for:

- beneficiaries between 66 and 96
- beneficiaries without any HMO months
- beneficiaries with a non-NA state
- with a sex in male/female

I use the `spec.pgram` function with `detrend=TRUE`, which removes a linear trend from the whole data (and also removes the mean).

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/', dev=c('png', 'pdf'),
                      pdf.options(useDingbats=FALSE),
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE, autodep=TRUE)
import::from(broom, tidy)
```

```{r load_data}
dat = lapply(2011:2014, function(y) {
  sprintf('../temporal_abx_%i.tsv', y) %>%
    read_tsv %>%
    filter(!is.na(state), week <= 52) %>%
    mutate(year=y)
  }) %>%
  bind_rows %>%
  rename(week_of_year=week) %>%
  mutate(week=(year - 2011)*52 + week_of_year)
```

```{r spectrogram functions}
strongest_period = function(pg) 1.0 / (pg$freq[which.max(pg$spec)])
drug_data = function(drugs) {
  filter(dat, antibiotic %in% drugs) %>%
    group_by(week) %>%
    summarize(n_claims=sum(n_claims))
}

drug_plot = function(drugs) {
  drug_data(drugs) %>%
    ggplot(aes(x=week, y=n_claims)) + geom_point()
}

drug_spectrogram = function(drugs) drug_data(drugs) %$% spec.pgram(n_claims)
drug_period = function(drugs) strongest_period(drug_spectrogram(drugs))
```

# Results

Azithromycin and quinolones are strongly seasonal. Tetracyclines are moderately seasonal. Penicillins and ceftriaxone are not seasonal.

## Azithromycin

```{r azithromycin}
drugs = c('azithromycin')
drug_plot(drugs)
plot(drug_spectrogram(drugs))
```

The strongest period is `r drug_period(drugs)` weeks.

## Cipro/levo

```{r quinolone}
drugs = c('levofloxacin', 'ciprofloxacin')
drug_plot(drugs)
plot(drug_spectrogram(drugs))
```

## Tetracycline/doxycycline

The strongest period is `r drug_period(drugs)` weeks.

```{r tetra}
drugs = c('tetracycline', 'doxycycline')
drug_plot(drugs)
plot(drug_spectrogram(drugs))
```

The strongest period is `r drug_period(drugs)` weeks.

## Ceftriaxone

```{r ceftriaxone}
drugs = c('ceftriaxone')
drug_plot(drugs)
plot(drug_spectrogram(drugs))
```

The strongest period is `r drug_period(drugs)` weeks.

## Pencillin/ampicillin

```{r penicillin}
drugs = c('penicillin', 'ampicillin')
drug_plot(drugs)
plot(drug_spectrogram(drugs))
```

The strongest period is `r drug_period(drugs)` weeks.

### 2017-05-10 temporal/seasonal-traces/drug_seasonality.Rmd
improved seasonality

@@
 
+## Data
+
 I use the 2011--2014 Medicare data, filtered for:
@@
 
+## Analysis
+
 I use the `spec.pgram` function with `detrend=TRUE`, which removes a linear trend from the whole data (and also removes the mean).
+
+I also use `nls` to fit the curves.
+The rough equation is $y = A + Bt + C\sin(2\pi t/D + E\pi)$.
+The time $t$ is just weeks, and $y$ is the number of claims in that week.
+This function works only if you give good starting values:
+
+- vertical offset $A$ as the mean number of claims
+- secular trend slope $B \approx 0$
+- amplitude $C$ as the standard deviation of the number of claims
+- period $D \approx 52$ (because we expect a yearly period)
+- phase offset $E \approx \tfrac{1}{2}$ (because we expect consumption to be high in the winter when $t \approx 0$)
+
+For the plots, I show a fit that includes a linear
 
@@
 
-```{r spectrogram functions}
+```{r spectrogram_functions, echo=TRUE}
 strongest_period = function(pg) 1.0 / (pg$freq[which.max(pg$spec)])
@@
 drug_plot = function(drugs) {
+  pf = periodic_fit(drugs)
+  p = pf$m$getPars()
+  f = function(t) p['A'] + p['B'] * t + p['C'] * sin(2 * pi * t / p['D'] + pi * p['E'])
   drug_data(drugs) %>%
-    ggplot(aes(x=week, y=n_claims)) + geom_point()
+    ggplot(aes(x=week, y=n_claims)) + geom_point() + stat_function(fun=f)
 }
@@
 drug_period = function(drugs) strongest_period(drug_spectrogram(drugs))
+
+periodic_fit = function(drugs) {
+  d = drug_data(drugs) %>% mutate(x=week, y=n_claims)
+  nls(y ~ A + B*x + C*sin(2*pi*x/D + pi*E), start=list(A=mean(d$y), B=0, C=sd(d$y), D=52, E=0.5), data=d)
+}
+
+periodic_fit_table = function(drugs) periodic_fit(drugs) %>% tidy %>% select(term, estimate) %>% kable
 ```
@@
 
-Azithromycin and quinolones are strongly seasonal. Tetracyclines are moderately seasonal. Penicillins and ceftriaxone are not seasonal.
+Azithromycin and quinolones are strongly seasonal: the spectrogram has a clear peak at approximately 52 week period, and the curve fit gives a strong amplitude and a period of almost exaclty 52.
+The quinolones are also seasonal---there is a peak in the spectrogram and the curve fit works nicely---but the amplitude is not as strong.
+
+The tetracyclines are moderately seasonal: the spectrogram and fit both work, but the amplitudes are again smaller.
+Ceftriaxone is weakly seasonal: the spectrogram does find that approximately 52 weeks is the strongest period, but the curve fit surprises me.
+The penicillins are not seasonal at all.
 
@@
 plot(drug_spectrogram(drugs))
+periodic_fit_table(drugs)
 ```
 
-The strongest period is `r drug_period(drugs)` weeks.
+The strongest period in the spectrogram is `r drug_period(drugs)` weeks.
 
@@
 plot(drug_spectrogram(drugs))
+periodic_fit_table(drugs)
 ```
 
+The strongest period in the spectrogram is `r drug_period(drugs)` weeks.
+
 ## Tetracycline/doxycycline
-
-The strongest period is `r drug_period(drugs)` weeks.
 
@@
 plot(drug_spectrogram(drugs))
+periodic_fit_table(drugs)
 ```
 
-The strongest period is `r drug_period(drugs)` weeks.
+The strongest period in the spectrogram is `r drug_period(drugs)` weeks.

### 2017-05-12 sex-difference/sex-difference.Rmd
sex diff

@@
 knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/', dev=c('png', 'pdf'),
-                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
+                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE, autodep=TRUE)
 ```
 
-```{r load_data, cache=TRUE}
-usage = read_tsv('../bene_usage_2011.tsv')
+```{r load_data}
+year = 2011
+bene = read_tsv(sprintf('../bene_%i.tsv', year)) %>%
+  filter(between(age, 66, 96), hmo_months==0, sex %in% c('male', 'female')) %>%
+  mutate(is_female=sex=='female', is_white=race=='white')
+
+pde = read_tsv(sprintf('../pde_%i.tsv', year)) %>%
+  filter(bene_id %in% bene$bene_id)
+
+total_pde = pde %>% count(bene_id) %>% rename(n_claims=n)
 ```
@@
 
-```{r compute_consumption_by_age_and_sex}
-bounds_group = function(x, upper_limits) {
-  upper_limits = as.integer(upper_limits)
-  levels = lapply(seq(from=1, to=length(upper_limits)), function(i) {
-    if (i == 1) {
-      sprintf('<= %i', upper_limits[i])
-    } else {
-      sprintf('> %i and <= %i', upper_limits[i - 1], upper_limits[i])
-    }
-  })
-  top_level = sprintf('> %i', tail(upper_limits, 1))
-  formula_list = mapply(function(str, lim) as.formula(sprintf('x <= %i ~ "%s"', lim, str)),
-                        levels, upper_limits) %>%
-    c(as.formula(sprintf('TRUE ~ "%s"', top_level)))
-
-  levels = c(levels, top_level)
-
-  do.call(case_when, formula_list) %>%
-    factor(levels=levels, ordered=TRUE)
-}
-
-usage %>%
-  filter(sex %in% c('male', 'female')) %>%
-  mutate(age_group=bounds_group(age, c(65, 75, 85, 95))) %>%
-  group_by(age_group, sex, bene_id) %>%
-  summarize_at(vars(n_claims, n_days), sum) %>%
-  summarize(mean_n_claims=mean(n_claims),
-            mean_n_days=mean(n_days),
-            n_bene=n()) %>%
-  kable
+```{r consumption_by_age_and_sex}
+bene %>%
+  left_join(total_pde, by='bene_id') %>%
+  replace_na(list(n_claims=0)) %>%
+  group_by(sex, age) %>%
+  summarize(mean_n_claims=mean(n_claims)) %>%
+  ungroup() %>%
+  ggplot(aes(x=age, y=mean_n_claims, color=sex)) +
+  geom_point()
 ```
@@
 
-```{r drugs_by_sex, cache=TRUE}
-bene_by_sex = usage %>%
-  #sample_frac(0.01) %>%
-  select(bene_id, sex) %>%
-  distinct %>%
-  count(sex) %>%
-  rename(n_bene=n)
+```{r drugs_by_sex}
+usage_by_sex = function(drug) {
+  bene %>%
+    left_join(pde %>% filter(antibiotic==drug) %>% count(bene_id) %>% rename(n_claims=n), by='bene_id') %>%
+    replace_na(list(n_claims=0)) %>%
+    group_by(sex) %>%
+    summarize(mean_n_claims=mean(n_claims)) %>%
+    ungroup() %>%
+    spread(sex, mean_n_claims) %>%
+    mutate(drug=drug)
+}
 
-dbs = usage %>%
-  #sample_frac(0.01) %>%
-  filter(age >= 65, sex %in% c('male', 'female')) %>%
-  group_by(antibiotic, sex, bene_id) %>%
-  summarize_at(vars(n_claims, n_days), sum) %>%
-  summarize(n_claims=sum(n_claims), n_days=sum(n_days)) %>%
-  left_join(bene_by_sex, by='sex') %>%
-  mutate(cpkp=n_claims*1000/n_bene, did=n_days*1000/(365*n_bene))
+dbs = lapply(unique(pde$antibiotic), usage_by_sex) %>% bind_rows
+```
 
-dbs %>%
-  filter(sex=='female') %>%
-  arrange(desc(cpkp)) %>%
-  head(10) %>%
-  kable
-
-dbs %>%
-  filter(sex=='male') %>% 
-  arrange(desc(cpkp)) %>%
-  head(10) %>%
-  kable
+```{r}
+denom = bene %>% count(sex) %>% rename(n_denom=n)
+ubs = pde %>%
+  select(bene_id, antibiotic) %>%
+  distinct() %>%
+  left_join(select(bene, bene_id, sex), by='bene_id') %>%
+  count(sex, antibiotic) %>%
+  ungroup() %>%
+  rename(n_user=n) %>%
+  left_join(denom, by='sex') %>%
+  mutate(frac_using=n_user/n_denom) %>%
+  select(antibiotic, sex, frac_using) %>%
+  spread(sex, frac_using)
 ```
@@
 ```{r drug_by_sex_plot}
-dbs %>%
+ubs %>%
   select(antibiotic, sex, cpkp) %>%

### 2017-05-12 sex-difference/sex-difference.Rmd
sex difference todo

--- 2017-03-27 sex-difference/sex-difference.Rmd
@@
 
-I'll use the Medicare 2011 data.
+I'll use the Medicare 2011 data. Or, use all years, pool that data, and include calendar year as part of the regression.
+
+*Outcome*: At least one claim for azithromycin, Bactrim, TMP/SMX, or nitrofurantoin. *Exposure*: A diagnosis with a UTI or an acute RTI.
+
+- In the first pass, look for any claim and any diagnosis.
+- In a second pass, look for a diagnosis that came at most two weeks before a claim.
+- Maybe, in a third pass, look for diagnosis that came at most two weeks before the first claim for that drug for the beneficiary.
+(Imagine someone is on year-long prophylaxis.
+Then a diagnosis at any time of the year will make it look like the diagnosis caused the claims.)
+
+Before a model, I should just stratify by sex and exposure. Do we already see the signal (or lack of it)? Then we can go onto including age and calendar year as covariates.
 

### 2017-05-12 ms1/ms1.Rmd
table of cc prevalence

@@
 ---
+
+# To do
+
+- Remove visits from the models, since we expect that visits are casually in between disease and antibiotics
+- Put all years together, put calendar year into the model, and regress (rather than doing separate regressions on each year)
 
@@
 
+  condition_prevalence = dat$bene %>%
+    select(AMI:HYPOTH) %>%
+    summarize_all(mean) %>%
+    gather('condition', 'prevalence', everything())
+
   condition_model = single_condition_models(dat$bene)
@@
        chars_by_consumption=chars_by_consumption,
+       condition_prevalence=condition_prevalence,
        condition_model=condition_model,
@@
 
+```{r arrange_condition_prevalence}
+condition_prevalence = data_by_field('condition_prevalence') %>%
+  spread(year, prevalence) %>%
+  arrange(desc(`2011`)) %>%
+  kable
+```
+
 ```{r arrange_data2}
@@
 
+# Prevalence of conditions
+
+`r condition_prevalence`
+
 # Figure 1: Claims and conditions by age

### 2017-05-18 ms2/ms2.Rmd
confidence intervals

@@
                       pdf.options(useDingbats=FALSE),
-                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
+                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE, autodep=TRUE)
 import::from(broom, tidy)
@@
 ```{r do_models}
+linear_interval = function(model, level) {
+  confint(model, level=level) %>%
+    tidy %>%
+    setNames(c('term', 'ci_low', 'ci_high'))
+}
+
 linear_model = function(df, dependent_var, explanatory_vars) {
   f = paste0(dependent_var, ' ~ ', paste0(explanatory_vars, collapse = ' + ')) %>% formula
-  lm(formula=f, weights=place_weight, data=df) %>%
+  m = lm(formula=f, weights=place_weight, data=df)
+  tidy(m) %>%
+    select(term, estimate, p.value) %>%
+    mutate(n=nrow(df)) %>%
+    left_join(linear_interval(m, 0.95), by='term')
+}
+
+spearman_interval = function(df, rho_name, n_name, level) {
+  mada::CIrho(df[[rho_name]], df[[n_name]], level=level) %>%
     tidy %>%
-    select(term, estimate, p.value)
+    setNames(c('rho', 'ci_low', 'ci_high')) %>%
+    select(-rho)
 }
@@
     mutate(term=explanatory_var, method='spearman') %>%
-    select(term, estimate, p.value)
+    mutate(n=nrow(df)) %>%
+    group_by(term, estimate, p.value, n) %>%
+    do(spearman_interval(., 'estimate', 'n', 0.95)) %>%
+    ungroup()
 }
@@
   arrange(p.adj) %>%
-  select(bug, drug_group, estimate, p.value, p.adj) %>%
+  select(bug, drug_group, estimate, p.value, p.adj, n) %>%
   kable
+```
+
+```{r}
+results %>%
+  filter(model=='spearman') %>%
+  mutate(bug_drug=interaction(bug, drug_group)) %>%
+  ggplot(aes(x=bug_drug, y=estimate, ymin=ci_low, ymax=ci_high)) +
+  geom_point() +
+  geom_linerange() +
+  coord_flip()
 ```
@@
   mutate(p.adj=p.adjust(p.value, method='BH')) %>%
-  filter(p.value < 0.05) %>%
+  #filter(p.value < 0.05) %>%
   arrange(p.value) %>%
-  select(bug, drug_group, estimate, p.value, p.adj) %>%
+  select(bug, drug_group, estimate, p.value, p.adj, n) %>%
   kable
+```
+
+```{r}
+results %>%
+  filter(model=='univariate', term=='mean') %>%
+  arrange(estimate) %>%
+  mutate(bug_drug=interaction(bug, drug_group),
+         bug_drug=factor(bug_drug, levels=bug_drug, ordered=TRUE)) %>%
+  ggplot(aes(x=bug_drug, y=estimate, ymin=ci_low, ymax=ci_high)) +
+  geom_point() +
+  geom_linerange() +
+  coord_flip()
 ```
@@
 
-```{r hrr_univariate_models, cache=FALSE}
+```{r hrr_univariate_models}
 hrr_results %>%
@@
 
-```{r multivariate_models, cache=FALSE}
+```{r multivariate_models}
 results %>%
@@
   kable
+```
+
+```{r}
+results %>%
+  filter(model=='multivariate', term==multivariate_term) %>%
+  arrange(estimate) %>%
+  mutate(bug_drug=interaction(bug, drug_group),
+         bug_drug=factor(bug_drug, levels=bug_drug, ordered=TRUE)) %>%
+  ggplot(aes(x=bug_drug, y=estimate, ymin=ci_low, ymax=ci_high)) +
+  geom_point() +
+  geom_linerange() +
+  coord_flip()
 ```
@@
 ```
+
+## MRSA clinda
+
+```{r}
+left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
+  left_join(regions, by='state') %>%
+  filter(bug=='MRSA', drug_group=='clindamycin') %>%
+  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, label=state)) +
+  geom_smooth(method='lm', col='gray', se=FALSE) +
+  geom_point() +
+  geom_text() +
+  #scale_x_continuous(limits=c(0.08, 0.2), breaks=c(0.08, 0.1, 0.15, 0.2)) +
+  #ylim(10, 45) +
+  xlab('clindamycin claims per beneficiary') + ylab('mean percent nonsusceptible MRSA isolates') +
+  theme_minimal()
+```

### 2017-05-19 ms1/ms1_display.Rmd

---
title: "Ms 1"
author: "Scott Olesen"
output: html_document
---

```{r global_options, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='fig/', dev=c('pdf', 'png'),
                      echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE,
                      error=FALSE, autodep=TRUE)
import::from(knitr, kable)
import::from(lazyeval, interp)
import::from(broom, tidy)
import::from(stringr, str_replace)
```

# Summary of data

```{r}
read_tsv('tables/tbl_totals.tsv') %>%
  mutate_at(vars(n_bene, n_pde), function(x) prettyNum(x, big.mark=',')) %>%
  mutate_at(vars(cpkp), function(x) round(x)) %>%
  kable
```

# Claims against age

```{r claims_by_age}
claims_by_age = read_tsv('tables/tbl_claims_by_age.tsv')

claims_by_age %>%
  filter(year==2011, !(group=='no_cc' & age < 67)) %>%
  ggplot(aes(x=age,
             y=mean_n_claims,
             ymin=mean_n_claims-sem_n_claims,
             ymax=mean_n_claims+sem_n_claims,
             color=group)) +
  geom_smooth(method='lm', se=F, size=0.5) +
  geom_point() +
  geom_linerange() +
  scale_x_continuous(breaks=c(66, 70, 80, 90, 96)) +
  scale_y_continuous(breaks=c(0.0, 1.0, 1.5, 2.0)) +
  theme_minimal()

lm_ci = function(df, formula) {
  lm(formula, data=df) %>%
    tidy %>%
    mutate(ci=1.96*std.error, cil=estimate-ci, ciu=estimate+ci)
}

claims_by_age %>%
  group_by(group) %>%
  do(lm_ci(., mean_n_claims ~ age)) %>%
  filter(term=='age') %>%
  select(group, estimate, cil, ciu) %>%
  kable
```

# Patterns of consumption

All consumption rates are reported as claims per thousand people per year (CPKP).

## Claims by class

```{r}
read_tsv('tables/tbl_claims_by_class.tsv') %>%
  select(antibiotic_class, year, cpkp) %>%
  spread(year, cpkp) %>%
  mutate(antibiotic_class=if_else(row_number(desc(`2011`)) < 10, antibiotic_class, 'other')) %>%
  group_by(antibiotic_class) %>%
  summarize_all(sum) %>%
  mutate_at(vars(-antibiotic_class), round) %>%
  arrange(desc(`2011`)) %>%
  kable
```

## Most-prescribed antibiotics

```{r}
claims_by_abx = read_tsv('tables/tbl_claims_by_abx.tsv')

top_abx = claims_by_abx %>%
  filter(year==2011) %>%
  filter(row_number(desc(cpkp)) < 10) %$%
  antibiotic

claims_by_abx %>%
  filter(antibiotic %in% top_abx) %>%
  select(antibiotic, year, cpkp) %>%
  spread(year, cpkp) %>%
  mutate_at(vars(-antibiotic), round) %>%
  arrange(desc(`2011`)) %>%
  kable
```

## Consumption by sex

```{r}
read_tsv('tables/tbl_claims_by_sex.tsv') %>%
  select(sex, year, cpkp) %>%
  spread(year, cpkp) %>%
  arrange(desc(`2011`)) %>% 
  mutate_at(vars(-sex), round) %>%
  kable
```

## Consumption by region

```{r}
read_tsv('tables/tbl_claims_by_region.tsv') %>%
  select(region, year, cpkp) %>%
  spread(year, cpkp) %>%
  arrange(desc(`2011`)) %>% 
  mutate_at(vars(-region), round) %>%
  kable
```

## Consumption by race

```{r}
read_tsv('tables/tbl_claims_by_race.tsv') %>%
  select(race, year, cpkp) %>%
  spread(year, cpkp) %>%
  arrange(desc(`2011`)) %>% 
  mutate_at(vars(-race), round) %>%
  kable
```

# Prevalence of conditions

```{r}
condition_names = read_tsv('../chronic-conditions/names.txt',
                           col_names=c('condition', 'condition_name'))

read_tsv('tables/tbl_cc_prevalence.tsv') %>%
  left_join(condition_names, by='condition') %>%
  select(condition=condition_name, year, prevalence) %>%
  spread(year, prevalence) %>%
  mutate(change=(`2014`-`2011`)/`2011`) %>%
  mutate_at(vars(-condition), function(x) round(x * 100, 1)) %>%
  arrange(desc(change)) %>%
  kable
```

# All-condition model

```{r}
read_tsv('tables/tbl_all_condition_model.tsv') %>%
  mutate(term=stringr::str_replace(term, 'TRUE$', '')) %>%
  left_join(condition_names, by=c('term'='condition')) %>%
  mutate(term=if_else(is.na(condition_name), term, condition_name),
         log10.adj.p.value=log10(p.adjust(p.value, method='BH'))) %>%
  select(term, estimate, p.value, log10.adj.p.value) %>%
  filter(term != '(Intercept)') %>%
  arrange(desc(estimate)) %>%
  kable
```

# Single-condition models

```{r}
read_tsv('tables/tbl_single_condition_models.tsv') %>%
  select(condition, term, estimate) %>%
  spread(term, estimate) %>%
  kable
```

### 2017-05-19 ms2/ms2.Rmd

@@
 ```{r}
-results %>%
-  filter(model=='spearman') %>%
-  mutate(bug_drug=interaction(bug, drug_group)) %>%
+interval_plot = function(df) {
+  arrange(df, estimate) %>%
+  mutate(bug_drug=interaction(bug, drug_group),
+         bug_drug=factor(bug_drug, levels=bug_drug, ordered=TRUE)) %>% 
     ggplot(aes(x=bug_drug, y=estimate, ymin=ci_low, ymax=ci_high)) +
@@
     coord_flip()
+}
+
+results %>%
+  filter(model=='spearman') %>%
+  interval_plot()
 ```
@@
 
-```{r hrr_spearman_models, cache=FALSE}
+```{r hrr_spearman_models}
 hrr_results %>%
@@
 
-```{r univariate_models, cache=FALSE}
+```{r univariate_models}
 results %>%
@@
   filter(model=='univariate', term=='mean') %>%
-  arrange(estimate) %>%
-  mutate(bug_drug=interaction(bug, drug_group),
-         bug_drug=factor(bug_drug, levels=bug_drug, ordered=TRUE)) %>%
-  ggplot(aes(x=bug_drug, y=estimate, ymin=ci_low, ymax=ci_high)) +
-  geom_point() +
-  geom_linerange() +
-  coord_flip()
+  interval_plot()
 ```
@@
   filter(model=='multivariate', term==multivariate_term) %>%
-  arrange(estimate) %>%
-  mutate(bug_drug=interaction(bug, drug_group),
-         bug_drug=factor(bug_drug, levels=bug_drug, ordered=TRUE)) %>%
-  ggplot(aes(x=bug_drug, y=estimate, ymin=ci_low, ymax=ci_high)) +
-  geom_point() +
-  geom_linerange() +
-  coord_flip()
+  interval_plot()
 ```
@@
 
-```{r cache=FALSE}
+```{r}
 hrr_results %>%
@@
 
-```{r hrr_multivariate_models, cache=FALSE}
+```{r hrr_multivariate_models}
 sig_bug_drug = hrr_results %>%
@@
   scale_color_manual(values=cbbPalette)
+```
+
+```{r}
+left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
+  filter(bug %in% c('E. coli', 'Eb. cloacae', 'CNST', 'MSSA', 'MRSA', 'S. pneumoniae')) %>%
+  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=drug_group, shape=bug)) +
+  geom_point() +
+  xlab('claims per beneficiary')
 ```

### 2017-05-19 drug-diagnoses/drug_diagnoses.Rmd
updates

---
title: "Diagnoses associated consumption of specific antibiotics"
author: "Scott Olesen"
date: ""
output: pdf_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/', dev=c('png', 'pdf'),
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE, autodep=TRUE)
```

# Background

We see changing levels of consumption of a few antibiotics:

- Ciprofloxacin going down
- Levofloxacin going up
- Amoxicillin going up
- Azithromycin going down

These decreases could be accompanied by changes in their usage with respect to patient indications. For example, it could be that amoxicillin is being uniformly prescribed more for all indications for which it was previously prescribed, or it could be that there are indications for which it is more often being prescribed.

# Methods

I used the 2011 and 2014 Medicare data. I looked for non-hospital outpatient claims (i.e., carrier events) associated with antibiotic prescriptions restricted by:

- the beneficiary passed the regular exclusion criteria (66 to 96 years old, not on HMO)
- the drug claims (PDE) was for one of the four drugs above (with amox/clav included with amox)
- the outpatient claim was 0 to 3 days before the drug claim

There are uniformly more claims in 2014 than in 2011, so I look at ratios of proportions of diagnoses. Specifically, for each drug and year, I looked at the primary diagnosis for the associated outpatient claims, computed the fraction of outpatient claims corresponding to that diagnosis, and computed the ratio of the two proportions. If a diagnosis has a ratio near 1, then diagnosis was associated with an equal fraction of claims for that drug. If the ratio is higher than 1, then that diagnosis was better represented in drug claims from 2014 than 2011.

```{r load_data}
icd = read_tsv('../../db/icd/icd.tsv') %>% rename(dx_code=code)

dat = lapply(c(2011, 2014), function(y) {
  read_tsv(sprintf('../../data/tmp_abx_dx_%i.tsv', y)) %>% mutate(year=y)
}) %>%
  bind_rows %>%
  count(year, antibiotic, dx_code) %>%
  group_by(year, antibiotic) %>%
  mutate(p=n/sum(n)) %>%
  ungroup()
```

```{r}
summarize_abx = function(abx) {
  dat %>%
    filter(antibiotic==abx) %>%
    select(-n) %>%
    spread(year, p) %>%
    mutate(pr=`2014`/`2011`) %>%
    mutate_at(vars(pr, `2011`, `2014`), function(x) round(x, digits=3)) %>%
    arrange(desc(`2011`)) %>%
    head(20) %>%
    left_join(icd, by='dx_code') %>%
    mutate(desc=stringr::str_sub(desc, end=40)) %>%
    select(Diagnosis=desc, `2011`, `2014`, `Ratio of proportions`=pr, `ICD code`=dx_code) %>%
    kable
}
```

# Results

Each table shows the 20 diagnoses with the greatest number of outpatient claims associated with that drug, the proportion of outpatient claims with that diagnosis in each year, and the ratio of the proportions between the two years.

## Amoxicillin

The greatest increase was for unspecified, chronic sinusitis (ratio 1.45); the greatest credible decrease was for bronchitis (0.82).

```{r}
summarize_abx('amoxicillin')
```

## Ciprofloxacin

The greatest increase was for dysuria (1.37); the greatest credible decrease was for elevated PSA (0.81). The fact that elevated PSA is here as one of the top 20 diagnoses associated with ciprofloxacin prescribing is itself interesting. This makes me think that: a doctor sees elevated PSA, so the chronic condition "enlarged prostate" gets recorded, and in the attempt to rule out prostatitis (before concluding prostate cancer), the doctor prescribes cipro.

```{r}
summarize_abx('ciprofloxacin')
```

## Levofloxacin

The greatest increases for levofloxacin were for bronchitis (1.38) and acute sinusitis (1.43). No super-strong decreases.

```{r}
summarize_abx('levofloxacin')
```

## Azithromycin

Azithromycin looks pretty steady. (I suspect that diabetes is just a common diagnosis that's not associated with azithromycin consumption.)

```{r}
summarize_abx('azithromycin')
```

# Discussion

First, I'm curious if these are the sort of results to be expected. The distribution of diagnoses is pretty flat: the top diagnosis for cipro (UTI), takes up 15% of all the outpatient claims, but the top diagnoses for the others take up more in the 5-10% range. There is a really long tail of diagnoses, which I think shows that we are not very specifically catching the diagnoses that led to the antibiotic prescriptions.

That being said, there are some individual strong effects. I speculate that azithromycin is being uniformly less used, while it seems like levo is being used more for acute respiratory complaint. Cipro was weird, apparently getting more concentrated in UTI treatment, while amoxicillin was a real puzzle, changing from bronchitis to sinusitis. Amoxicillin is apparently a first choice for bacterial sinusitis.

### 2017-05-19 polish-bene-pde.R

@@ -1,112 +1,32 @@
 #!/usr/bin/env Rscript
 
-# Polish the beneficiary and antibiotic PDE files at the same time:
-#  - Keep all the beneficiaries that match criteria (good geography, 12 months coverage)
-#  - Keep PDEs that match known beneficiaries
+abx_class = read_tsv('abx_class.tsv')
 
-state_codes = read_tsv('../db/state-codes/state-codes.tsv') %>%
-  mutate(is_state_or_dc=as.logical(is_state_or_dc)) %>%
-  filter(is_state_or_dc)
-
-oral_injected_codes = scan('dosage-form/oral_injected_codes.txt', what=character())
-antibiotic_names = read_tsv('abx_class.tsv')
-
-polish_bene = function(x) {
-  rename(x, bene_id=BENE_ID,
-          state_cd=STATE_CD,
-          age=AGE,
-          hmo_months=HMO_MO,
-          plan_coverage_months=PLNCOVMO) %>%
-    mutate_at(vars(age, hmo_months, plan_coverage_months), as.integer) %>% # should be already done from read_tsv
-    left_join(state_codes, by='state_cd') %>%
-    filter(!is.na(state), plan_coverage_months==12) %>% # filter for geography and plan
-    mutate(zipcode=substr(BENE_ZIP, 1, 5)) %>%
-    mutate(sex=case_when(.$SEX == 0 ~ 'unknown',
-                          .$SEX == 1 ~ 'male',
-                          .$SEX == 2 ~ 'female')) %>%
-    mutate(race=case_when(.$RTI_RACE_CD == 0 ~ 'unknown',
-                          .$RTI_RACE_CD == 1 ~ 'white', # non-Hispanic white
-                          .$RTI_RACE_CD == 2 ~ 'black',
-                          .$RTI_RACE_CD == 3 ~ 'other',
-                          .$RTI_RACE_CD == 4 ~ 'Asian', # or Pacific Islander
-                          .$RTI_RACE_CD == 5 ~ 'Hispanic',
-                          .$RTI_RACE_CD == 6 ~ 'American Indian')) %>% # or Alaska Native
[…]

### 2017-05-23 ms1/ms1_display.Rmd
updated to new tables

@@
 knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='fig/', dev=c('pdf', 'png'),
-                      echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE,
+                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE,
                       error=FALSE, autodep=TRUE)
@@
 import::from(broom, tidy)
-import::from(stringr, str_replace)
+import::from(stringr, str_replace, str_detect)
+
+pctize = function(df, denom_name) {
+  for (n_name in Filter(function(x) str_detect(x, '^n_'), names(df))) {
+    pct_name = str_replace(n_name, '^n_', 'pct_')
+    df[[pct_name]] = df[[n_name]] / df[[denom_name]] * 100
+  }
+  df
+}
 ```
@@
 read_tsv('tables/tbl_totals.tsv') %>%
-  mutate_at(vars(n_bene, n_pde), function(x) prettyNum(x, big.mark=',')) %>%
-  mutate_at(vars(cpkp), function(x) round(x)) %>%
-  kable
-```
-
-# Claims against age
-
-```{r claims_by_age}
-claims_by_age = read_tsv('tables/tbl_claims_by_age.tsv')
-
-claims_by_age %>%
-  filter(year==2011, !(group=='no_cc' & age < 67)) %>%
-  ggplot(aes(x=age,
-             y=mean_n_claims,
-             ymin=mean_n_claims-sem_n_claims,
-             ymax=mean_n_claims+sem_n_claims,
-             color=group)) +
-  geom_smooth(method='lm', se=F, size=0.5) +
-  geom_point() +
-  geom_linerange() +
-  scale_x_continuous(breaks=c(66, 70, 80, 90, 96)) +
-  scale_y_continuous(breaks=c(0.0, 1.0, 1.5, 2.0)) +
-  theme_minimal()
-
-lm_ci = function(df, formula) {
-  lm(formula, data=df) %>%
-    tidy %>%
-    mutate(ci=1.96*std.error, cil=estimate-ci, ciu=estimate+ci)
-}
-
-claims_by_age %>%
-  group_by(group) %>%
-  do(lm_ci(., mean_n_claims ~ age)) %>%
-  filter(term=='age') %>%
-  select(group, estimate, cil, ciu) %>%
+  pctize('n_bene') %>%
+  select(year, n_bene, mean_age, mean_n_cc,
+         pct_female, pct_white, pct_dual,
+         pct_region_south, pct_region_midwest, pct_region_northeast, pct_region_west) %>%
+  mutate_at(vars(starts_with('pct_')), function(x) round(x, digits=1)) %>%
+  mutate(n_bene=prettyNum(n_bene, big.mark=','),
+         mean_age=round(mean_age, digits=1),
+         mean_n_cc=round(mean_n_cc, digits=2)) %>%
+  gather('var', 'val', -year) %>% spread(year, val) %>% # transpose
   kable
@@
 
-## Claims by class
+## Claims by abx
 
 ```{r}
-read_tsv('tables/tbl_claims_by_class.tsv') %>%
-  select(antibiotic_class, year, cpkp) %>%
+dat = read_tsv('tables/tbl_claims_by_abx.tsv')
+totals = dat %>% group_by(year) %>% summarize(cpkp=sum(cpkp)) %>% mutate(antibiotic='all')
+top_abx = dat %>%
+  group_by(antibiotic) %>%
+  summarize(cpkp=mean(cpkp)) %>%
+  arrange(desc(cpkp)) %$%
+  head(antibiotic, 10)
+
+x = dat %>%
+  select(antibiotic, year, cpkp) %>%
+  filter(antibiotic %in% top_abx) %>%
+  bind_rows(totals) %>%
+  mutate(antibiotic=factor(antibiotic, levels=c('all', top_abx)))
+
+x %>%
+  mutate(cpkp=round(cpkp)) %>%
   spread(year, cpkp) %>%
-  mutate(antibiotic_class=if_else(row_number(desc(`2011`)) < 10, antibiotic_class, 'other')) %>%
-  group_by(antibiotic_class) %>%
-  summarize_all(sum) %>%
-  mutate_at(vars(-antibiotic_class), round) %>%
   arrange(desc(`2011`)) %>%
   kable
+
+x %>% ggplot(aes(x=year, y=cpkp)) +
+  facet_wrap(~ antibiotic, ncol=3, scales='free_y') +
+  geom_point() + geom_line() +
+  theme_minimal()
 ```
 
-## Most-prescribed antibiotics
+## Claims by age group
 
-```{r}
-claims_by_abx = read_tsv('tables/tbl_claims_by_abx.tsv')
-
-top_abx = claims_by_abx %>%
-  filter(year==2011) %>%
-  filter(row_number(desc(cpkp)) < 10) %$%
-  antibiotic
-
-claims_by_abx %>%
-  filter(antibiotic %in% top_abx) %>%
-  select(antibiotic, year, cpkp) %>%
+```{r claims_by_age}
+read_tsv('tables/tbl_claims_by_age.tsv') %>%
+  mutate(cpkp=round(cpkp)) %>%
+  select(year, age_group, cpkp) %>%
   spread(year, cpkp) %>%
-  mutate_at(vars(-antibiotic), round) %>%
-  arrange(desc(`2011`)) %>%
   kable
@@
 
-```{r}
+```{r, eval=FALSE}
 read_tsv('tables/tbl_all_condition_model.tsv') %>%
@@
 ```
-
-# Single-condition models
-
-```{r}
-read_tsv('tables/tbl_single_condition_models.tsv') %>%
-  select(condition, term, estimate) %>%
-  spread(term, estimate) %>%
-  kable
-```

### 2017-05-23 ms1/ms1_work.R

@@ -1,23 +1,26 @@
+# Do all the heavy-lifting analyses so that the Rmd can just show the results
+
 import::from(lazyeval, interp)
 import::from(broom, tidy)
 import::from(stringr, str_replace)
 
+# Load the census regions. Code them as factors so that Northeast is taken as
+# the baseline in the linear models.
+# NB: I put DC into the South
 regions = read_tsv('../../db/census-regions/census-regions.tsv') %>%
   select(state, region)
+
 condition_names = read_tsv('../chronic-conditions/names.txt',
                            col_names=c('condition', 'condition_name'))
 
 load_data = function(year) {
   bene = read_tsv(sprintf('../bene_%i.tsv', year)) %>%
-    filter(between(age, 66, 96), hmo_months==0, sex %in% c('male', 'female')) %>%
-    select(-hmo_months) %>%
-    mutate(is_female=sex=='female', is_white=race=='white') %>%
+    mutate(is_female=sex=='female', is_white=race=='white', is_dual=buyin_months>0) %>%
     left_join(regions, by='state') %>%
     mutate(year=year)
   
   pde = read_tsv(sprintf('../pde_%i.tsv', year)) %>%
-    select(-service_date, -days_supply, -fill_number) %>%
-    filter(bene_id %in% bene$bene_id) %>%
+    select(bene_id, antibiotic, antibiotic_class) %>%
     mutate(year=year)
   
   # get total numbers of PDEs
@@ -41,30 +44,40 @@ pde = lapply(dat, function(df) df$pde) %>% bind_rows
 rm(dat)
 
 output_table = function(x, base) write_tsv(x, sprintf('tables/tbl_%s.tsv', base))
+sem = function(x) sd(x) / length(x)
 
-# total number of beneficiaries and PDEs
-totals = left_join(
-  bene %>% count(year) %>% rename(n_bene=n),
-  pde %>% count(year) %>% rename(n_pde=n),
-  by='year') %>%
-  mutate(cpkp=n_pde*1000/n_bene) %T>%
+# summary characteristics
+totals = bene %>%
+  group_by(year) %>%
+  summarize(n_bene=n(),
+            mean_age=mean(age),
+            sem_age=sem(age),
+            mean_n_cc=mean(n_cc),
+            sem_n_cc=sem(n_cc),
+            n_female=sum(is_female),
[…]

### 2017-05-23 ms2/data/collect-data.R

@@ -42,7 +42,6 @@ summarize_inequality = function(year) {
   hrr_fn = sprintf('hrr_ineq_%i.tsv', year)
 
   bene = read_tsv(bene_fn) %>%
-    filter(between(age, 66, 96), hmo_months==0) %>%
     left_join(regions, by='state') %>%
     left_join(hrr, by='zipcode')
 

### 2017-05-25 ms2/ms2.Rmd
ms2

@@
 
-all_ineq = read_tsv('data/state_ineq_all_2011.tsv') %>%
+target_year = 2013
+all_ineq = read_tsv(sprintf('data/state_ineq_all_%i.tsv', target_year)) %>%
   left_join(regions, by='state')
-state_ineq = read_tsv('data/state_ineq_2011.tsv')
-hrr_ineq = read_tsv('data/hrr_ineq_2011.tsv')
+state_ineq = read_tsv(sprintf('data/state_ineq_%i.tsv', target_year))
+hrr_ineq = read_tsv(sprintf('data/hrr_ineq_%i.tsv', target_year))
 ```
@@
 xlab('claims per beneficiary') +
-ylab('Gini coefficient among benes. w/ >= 1 claim') +
+ylab('Gini coefficient') +
 theme_minimal()
@@
     linear_model(x, dep, c('fnz', multivariate_term)) %>% mutate(model='multivariate'),
+    linear_model(x, dep, multivariate_term) %>% mutate(model='uni_nzg'),
     spearman_model(x, dep, 'mean') %>% mutate(model='spearman')
@@
          bug_drug=factor(bug_drug, levels=bug_drug, ordered=TRUE)) %>% 
-    ggplot(aes(x=bug_drug, y=estimate, ymin=ci_low, ymax=ci_high)) +
+    ggplot(aes(x=bug_drug, y=estimate, ymin=ci_low, ymax=ci_high, color=drug_group)) +
     geom_point() +
@@
   select(bug, drug_group, estimate, p.value, p.adj) %>%
+  kable
+```
+
+## Univariate Gini
+
+```{r}
+filter(results, model=='uni_nzg', term=='nzgini') %>%
+  arrange(p.value) %>%
+  mutate(adj.p=p.adjust(p.value, method='bonferroni')) %>%
+  select(bug, drug_group, estimate, p.value, adj.p, n) %>%
   kable
@@
   interval_plot()
+```
+
+```{r}
+results %>%
+  filter(model=='multivariate') %>%
+  filter(term == multivariate_term) %>%
+  mutate(bug_drug=interaction(bug, drug_group)) %>%
+  ggplot(aes(x=p.value, y=estimate, color=drug_group)) +
+  geom_point() +
+  scale_color_brewer(palette='Set1')
 ```
@@
   theme_minimal()
+
+ecoli_bactrim_dat %>%
+  ggplot(aes(x=nzgini, y=mean_percent_nonsusceptible, label=state)) +
+  geom_smooth(method='lm', col='gray', se=FALSE) +
+  geom_point() +
+  geom_text() +
+  xlab('Gini coefficient)') +
+  ylab('E. coli resistance') +
+  theme_minimal()
 ```

### 2017-05-25 ms2/ms2.Rmd
moving toward years

@@
 hrr = read_tsv('../../db/hrr/hrr.tsv') %>%
-  filter(year==2011) %>%
   select(zipcode, hrr) %>%
@@
 
-target_year = 2013
-all_ineq = read_tsv(sprintf('data/state_ineq_all_%i.tsv', target_year)) %>%
+years = 2011:2014
+read_years = function(template_fn) {
+  lapply(years, function(y) read_tsv(sprintf(template_fn, y)) %>% mutate(year=y)) %>%
+    bind_rows()
+}
+
+all_ineq = read_years('data/state_ineq_all_%i.tsv') %>%
+  group_by(state) %>%
+  summarize(mean_cpb=mean(mean),
+            sem_cpb=sd(mean)/sqrt(n()),
+            mean_fnz=mean(fnz),
+            sem_fnz=sd(fnz)/sqrt(n()),
+            mean_nzgini=mean(nzgini),
+            sem_nzgini=sd(nzgini)/sqrt(n())) %>%
   left_join(regions, by='state')
-state_ineq = read_tsv(sprintf('data/state_ineq_%i.tsv', target_year))
-hrr_ineq = read_tsv(sprintf('data/hrr_ineq_%i.tsv', target_year))
+
+state_ineq = read_years('data/state_ineq_%i.tsv')
+hrr_ineq = read_years('data/hrr_ineq_%i.tsv')
 ```
@@
 
-```{r fnz_mean, cache=FALSE}
-ggplot(all_ineq, aes(x=mean, y=fnz, color=region, label=state_abbreviation)) +
+Error bars show the s.e.m. across the four data years.
+
+```{r fnz_mean}
+all_ineq %>%
+  mutate_at(vars(matches('fnz')), function(x) x*100) %>%
+  ggplot(aes(x=mean_cpb, xmin=mean_cpb-sem_cpb, xmax=mean_cpb+sem_cpb,
+             y=mean_fnz, ymin=mean_fnz-sem_fnz, ymax=mean_fnz+sem_fnz,
+             color=region, label=state_abbreviation)) +
   geom_point() +
+  geom_linerange() +
+  geom_errorbarh(height=0) +
   geom_text(nudge_x=0.04) +
-xlab('claims per beneficiary') +
-ylab('fraction of beneficiaries w/ >= 1 claim') +
+  xlab('mean claims per beneficiary') +
+  ylab('mean % beneficiaries with at least one claim') +
   theme_minimal()
@@
 
-```{r nzgini_mean, cache=FALSE}
-ggplot(all_ineq, aes(x=mean, y=nzgini, color=region, label=state_abbreviation)) +
+Error bars show the s.e.m. across the four data years.
+
+```{r nzgini_mean}
+all_ineq %>%
+  ggplot(aes(x=mean_cpb, xmin=mean_cpb-sem_cpb, xmax=mean_cpb+sem_cpb,
+             y=mean_nzgini, ymin=mean_nzgini-sem_nzgini, ymax=mean_nzgini+sem_nzgini,
+             color=region, label=state_abbreviation)) +
   geom_point() +
+  geom_linerange() +
+  geom_errorbarh(height=0) +
   geom_text(nudge_x=0.04) +
-xlab('claims per beneficiary') +
-ylab('Gini coefficient') +
+  xlab('mean claims per beneficiary') +
+  ylab('mean Gini coefficient') +
   theme_minimal()
@@
 
-# Keep only bug/drug combinations that are in at least 75 HRRs
+# Keep only the abundant bug/drug combinations
 min_n_states = 35
@@
               n_place_antibiograms=n()) %>%
-    #filter(!is.na(mean_percent_nonsusceptible)) %>%
     mutate(place_weight=sqrt(n_place_antibiograms))

### 2017-05-25 ms2/ms2.Rmd
ms2 hrr

@@
 
+# Use the zipcode-HRR mapping from 2011
 hrr = read_tsv('../../db/hrr/hrr.tsv') %>%
+  filter(year==2011) %>%
   select(zipcode, hrr) %>%
@@
 
-state_ineq = read_years('data/state_ineq_%i.tsv')
-hrr_ineq = read_years('data/hrr_ineq_%i.tsv')
+state_ineq = read_years('data/state_ineq_%i.tsv') %>%
+  rename(cpb=mean) %>%
+  group_by(state, drug_group) %>%
+  summarize(cpb=mean(cpb), fnz=mean(fnz), nzg=mean(nzgini)) %>%
+  ungroup()
+
+hrr_ineq = read_years('data/hrr_ineq_%i.tsv') %>%
+  rename(cpb=mean) %>%
+  group_by(hrr, drug_group) %>%
+  summarize(cpb=mean(cpb), fnz=mean(fnz), nzg=mean(nzgini)) %>%
+  ungroup()
 ```
@@
 
-## Inequality among users does not correlate well
+```{r}
+all_ineq %$%
+  cor.test(mean_cpb, mean_fnz, method='spearman')
+```
+
+## The two inequality metrics not correlate well
 
@@
 all_ineq %>%
-  ggplot(aes(x=mean_cpb, xmin=mean_cpb-sem_cpb, xmax=mean_cpb+sem_cpb,
+  ggplot(aes(x=mean_fnz, xmin=mean_fnz-sem_fnz, xmax=mean_fnz+sem_fnz,
              y=mean_nzgini, ymin=mean_nzgini-sem_nzgini, ymax=mean_nzgini+sem_nzgini,
@@
   geom_text(nudge_x=0.04) +
-  xlab('mean claims per beneficiary') +
+  xlab('mean fraction with at least one claim') +
   ylab('mean Gini coefficient') +
   theme_minimal()
+```
+
+The correlation exists but it's much less strong.
+
+```{r}
+all_ineq %$%
+  cor.test(mean_fnz, mean_nzgini, method='spearman')
+```
+
+There is an intermediate correlation between claims per beneficiary and the Gini coefficient (see below), so in multivariable regressions I will use those two (rather than CPB and Gini).
+
+```{r}
+all_ineq %$% cor.test(mean_cpb, mean_nzgini, method='spearman')
 ```
@@
 
-multivariate_term = 'nzgini'
+uni_term = 'cpb' # univariable explanatory term
+multi_term = 'nzg' # multivariable explanatory term (along with fnz)
 models = function(x) {
   dep = 'mean_percent_nonsusceptible'
-  # swo> "mean" is a confusing name
   bind_rows(
-    linear_model(x, dep, 'mean') %>% mutate(model='univariate'),
-    linear_model(x, dep, c('fnz', multivariate_term)) %>% mutate(model='multivariate'),
-    linear_model(x, dep, multivariate_term) %>% mutate(model='uni_nzg'),
-    spearman_model(x, dep, 'mean') %>% mutate(model='spearman')
+    linear_model(x, dep, uni_term) %>% mutate(model='uni'),
+    linear_model(x, dep, c('fnz', multi_term)) %>% mutate(model='multi'),
+    spearman_model(x, dep, uni_term) %>% mutate(model='spearman')
   ) %>%
@@
 
-## Univariate Gini
-
-```{r}
-filter(results, model=='uni_nzg', term=='nzgini') %>%
-  arrange(p.value) %>%
-  mutate(adj.p=p.adjust(p.value, method='bonferroni')) %>%
-  select(bug, drug_group, estimate, p.value, adj.p, n) %>%
-  kable
-```
-
-## Univariate
+## Univariable
 
@@
 
-```{r univariate_models}
+```{r uni_models}
 results %>%
-  filter(model=='univariate', term == 'mean') %>%
+  filter(model=='uni', term==uni_term) %>%
   mutate(p.adj=p.adjust(p.value, method='BH')) %>%
@@
 results %>%
-  filter(model=='univariate', term=='mean') %>%
+  filter(model=='uni', term==uni_term) %>%
   interval_plot()
@@
 results %>%
-  filter(model=='univariate') %>%
+  filter(model=='uni') %>%
   mutate(p.adj=p.adjust(p.value, method='BH')) %>%
@@
 
-```{r hrr_univariate_models}
+```{r hrr_uni_models}
 hrr_results %>%
-  filter(model=='univariate', term == 'mean') %>%
+  filter(model=='uni', term==uni_term) %>%
   mutate(p.adj=p.adjust(p.value, method='BH')) %>%
@@
 
-## Multivariate
+## Multivariable
 
@@
 
-```{r multivariate_models}
+```{r multi_models}
 results %>%
-  filter(model=='multivariate', term==multivariate_term) %>%
+  filter(model=='multi', term==multi_term) %>%
   mutate(p.adj=p.adjust(p.value, method='BH')) %>%
@@
 results %>%
-  filter(model=='multivariate', term==multivariate_term) %>%
+  filter(model=='multi', term==multi_term) %>%
   interval_plot()
@@
 results %>%
-  filter(model=='multivariate') %>%
-  filter(term == multivariate_term) %>%
+  filter(model=='multi', term==multi_term) %>%
   mutate(bug_drug=interaction(bug, drug_group)) %>%
@@
 
-And what about looking at multivariate only for ones that passed the threshold for univariate?
+And what about looking at multivariate only for ones that passed the threshold for univariable?
 
@@
 sig_bug_drug = results %>%
-  filter(model=='univariate', term=='mean') %>%
+  filter(model=='uni', term==uni_term) %>%
   mutate(p.adj=p.adjust(p.value, method='BH')) %>%
@@
 results %>%
-  filter(model=='multivariate', term==multivariate_term) %>%
+  filter(model=='multi', term==multi_term) %>%
   right_join(sig_bug_drug, by=c('bug', 'drug_group')) %>%
@@
 hrr_results %>%
-  filter(model=='multivariate', term==multivariate_term) %>%
+  filter(model=='multi', term==multi_term) %>%
   mutate(p.adj=p.adjust(p.value, method='BH')) %>%
@@
 
-```{r hrr_multivariate_models}
+```{r hrr_multi_models}
 sig_bug_drug = hrr_results %>%
-  filter(model=='univariate', term=='mean') %>%
+  filter(model=='uni', term==uni_term) %>%
   mutate(p.adj=p.adjust(p.value, method='BH')) %>%
@@
 hrr_results %>%
-  filter(model=='multivariate', term==multivariate_term) %>%
+  filter(model=='multi', term==multi_term) %>%
   right_join(sig_bug_drug, by=c('bug', 'drug_group')) %>%
@@
 state_ineq %>%
-  ggplot(aes(x=mean, y=nzgini, color=drug_group)) +
+  ggplot(aes(x=cpb, y=nzg, color=drug_group)) +
   geom_point() +
@@
   filter(bug %in% c('E. coli', 'Eb. cloacae', 'CNST', 'MSSA', 'MRSA', 'S. pneumoniae')) %>%
-  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=drug_group, shape=bug)) +
+  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, color=drug_group, shape=bug)) +
   geom_point() +
@@
 
-```{r ecoli_quin, cache=FALSE}
+```{r ecoli_quin, eval=FALSE}
 left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
@@
 
-```{r ecoli_bactrim}
+```{r ecoli_bactrim, eval=FALSE}
 ecoli_bactrim_dat = left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
@@
 
-```{r}
+```{r, eval=FALSE}
 left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
@@
 
-```{r mssa_clinda}
+```{r mssa_clinda, eval=FALSE}
 dat = left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%

### 2017-05-29 ms1/ms1_display.Rmd
cohort, updates

@@
   kable
+```
 
+```{r}
 x %>% ggplot(aes(x=year, y=cpkp)) +
@@
   theme_minimal()
+ggsave('~/tmp.pdf')
 ```

### 2017-05-29 ms1/ms1_cohort.R

@@ -1,6 +1,7 @@
 # Do all the heavy-lifting analyses so that the Rmd can just show the results
 
 import::from(broom, tidy)
+import::from(stringr, str_replace)
 
 # Load the census regions. Code them as factors so that Northeast is taken as
 # the baseline in the linear models.
@@ -18,7 +19,7 @@ load_data = function(year) {
     mutate(year=year)
 
   pde = read_tsv(sprintf('../pde_%i.tsv', year)) %>%
-    select(bene_id, antibiotic, days_supply, antibiotic_class) %>%
+    select(bene_id, antibiotic, antibiotic_class) %>%
     mutate(year=year)
 
   # get total numbers of PDEs
@@ -37,23 +38,33 @@ load_data = function(year) {
 }
 
 dat = lapply(2011:2014, load_data)
-bene = lapply(dat, function(df) df$bene) %>% bind_rows %>%
-  group_by(bene_id) %>% mutate(min_age=min(age)) %>%
-  ungroup()
+bene = lapply(dat, function(df) df$bene) %>% bind_rows
 pde = lapply(dat, function(df) df$pde) %>% bind_rows
 rm(dat)
 
-output_table = function(x, base) write_tsv(x, sprintf('tables/tbl_%s.tsv', base))
-sem = function(x) sd(x) / length(x)
+output_table = function(x, base) write_tsv(x, sprintf('tables/cohort/tbl_%s.tsv', base))
+
+# make a table that shows when benes are in the data set
+#x = bene %>%
+  #select(bene_id, year) %>%
+  #group_by(bene_id) %>%
+  #summarize(years=paste0(year, collapse=' ')) %>%
+  #count(years) %>%
+  #output_table('cohorts')
+
+# keep only the bene's that have data in all four years
+bene %<>% group_by(bene_id) %>%
+  mutate(start_age=min(age)) %>%
+  filter(n() == 4, start_age >= 68) %>%
+  ungroup()
+pde %<>% filter(bene_id %in% bene$bene_id)
 
 # summary characteristics
 totals = bene %>%
   group_by(year) %>%
   summarize(n_bene=n(),
             mean_age=mean(age),
-            sem_age=sem(age),
             mean_n_cc=mean(n_cc),
-            sem_n_cc=sem(n_cc),
             n_female=sum(is_female),
             n_white=sum(is_white),
             n_dual=sum(is_dual),
[…]

### 2017-05-29 ms2/ms2.Rmd
ms2 year average

@@
 ---
+
+# To do
+
+- filter by HRRs, not states
 
@@
 ```
-
-# Consumption data
 
@@
 
-## Aggregate usage correlates with fraction of users
-
-Error bars show the s.e.m. across the four data years.
-
-```{r fnz_mean}
-all_ineq %>%
-  mutate_at(vars(matches('fnz')), function(x) x*100) %>%
-  ggplot(aes(x=mean_cpb, xmin=mean_cpb-sem_cpb, xmax=mean_cpb+sem_cpb,
-             y=mean_fnz, ymin=mean_fnz-sem_fnz, ymax=mean_fnz+sem_fnz,
-             color=region, label=state_abbreviation)) +
-  geom_point() +
-  geom_linerange() +
-  geom_errorbarh(height=0) +
-  geom_text(nudge_x=0.04) +
-  xlab('mean claims per beneficiary') +
-  ylab('mean % beneficiaries with at least one claim') +
-  theme_minimal()
-```
-
-```{r}
-all_ineq %$%
-  cor.test(mean_cpb, mean_fnz, method='spearman')
-```
-
-## The two inequality metrics not correlate well
-
-Error bars show the s.e.m. across the four data years.
-
-```{r nzgini_mean}
-all_ineq %>%
-  ggplot(aes(x=mean_fnz, xmin=mean_fnz-sem_fnz, xmax=mean_fnz+sem_fnz,
-             y=mean_nzgini, ymin=mean_nzgini-sem_nzgini, ymax=mean_nzgini+sem_nzgini,
-             color=region, label=state_abbreviation)) +
-  geom_point() +
-  geom_linerange() +
-  geom_errorbarh(height=0) +
-  geom_text(nudge_x=0.04) +
-  xlab('mean fraction with at least one claim') +
-  ylab('mean Gini coefficient') +
-  theme_minimal()
-```
-
-The correlation exists but it's much less strong.
-
-```{r}
-all_ineq %$%
-  cor.test(mean_fnz, mean_nzgini, method='spearman')
-```
-
-There is an intermediate correlation between claims per beneficiary and the Gini coefficient (see below), so in multivariable regressions I will use those two (rather than CPB and Gini).
-
-```{r}
-all_ineq %$% cor.test(mean_cpb, mean_nzgini, method='spearman')
-```
-
-# Resistance data
-
 ```{r load_abg}
@@
 ```
-
-And I join the resistance drugs into classes:
-
-`r kable(resistance_groups)`
-
-I similarly joined the consumption drugs into similar classes.
-
-I started by considering these bug/drug combinations:
-
-`r kable(possible_bug_drug)`
-
-Then I filtered to keep only those bug/drug combinations that were present in `r min_n_states` states. That means I'm looking at these `r nrow(good_combos)` bug/drug combinations:
-
-`r good_combos %>% arrange(desc(n_places)) %>% kable`
 
@@
 
+# Landscape of bug/drug correlations
+
+## Selecting correlations to test
+
+I join the resistance drugs into classes:
+
+`r kable(resistance_groups)`
+
+I similarly joined the consumption drugs into similar classes.
+
+I started by considering these bug/drug combinations:
+
+`r kable(possible_bug_drug)`
+
+Then I filtered to keep only those bug/drug combinations that were present in `r min_n_states` states. That means I'm looking at these `r nrow(good_combos)` bug/drug combinations:
+
+`r good_combos %>% arrange(desc(n_places)) %>% kable`
+
+## Preparing the consumption data
+
+Unlike previous analyses, this time I treated the consumption and resistance data similarly: I averaged both across all data years.
+Thus, to get a state's Gini coefficient, I computed that number each year and then took the average.
+
 ## Spearman
 
-Significant results using Spearman correlations of mean percent nonsusceptible ~ claims per beneficiary:
+Significant results using Spearman correlations of mean percent nonsusceptible ~ claims per beneficiary.
+A star after the bug/drug combination implies that the adjusted $p < 0.05$, and a dot means the unadjusted $p < 0.05$.
 
-```{r spearman_models, cache=FALSE}
+```{r spearman_models}
 results %>%
@@
 
-```{r}
+Aside from individual significant results, we can look at the landscape of correlations.
+They are almost all positive, with MRSA/Bactrim making an interesting exception.
+Lines show 95% confidence intervals.
+
+```{r spearman_interval}
 interval_plot = function(df) {
   arrange(df, estimate) %>%
-  mutate(bug_drug=interaction(bug, drug_group),
+  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
+  mutate(sig=case_when(.$p.adj < 0.05 ~ '*',
+                       .$p.value < 0.05 ~ '.',
+                       TRUE ~ ''),
+         bug_drug=interaction(bug, drug_group, sig),
          bug_drug=factor(bug_drug, levels=bug_drug, ordered=TRUE)) %>% 
@@
   filter(model=='spearman') %>%
+  interval_plot() +
+  ylab("Spearman's rho")
+```
+
+## Differences in consumption/resistance patterns explain some of these results
+
+Different bug/drug combinations have different spreads along the consumption and resistance axes, so we would expect to see different strengths of association for each one.
+
+```{r}
+left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
+  filter(bug %in% c('E. coli', 'Eb. cloacae', 'CNST', 'MSSA', 'MRSA', 'S. pneumoniae')) %>%
+  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, color=drug_group, shape=bug)) +
+  geom_point() +
+  xlab('claims per beneficiary')
+```
+
+# Unevenness of consumption and antibiotic resistance
+
+## Drugs are consumed differently in different places
+
+```{r nzgini_cpb_drug}
+cbbPalette <- c("#000000", "#D55E00", "#CC79A7", "#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2")
+state_ineq %>%
+  ggplot(aes(x=cpb, y=nzg, color=drug_group)) +
+  geom_point() +
+  ylab('Gini coefficient among users') + xlab('claims per beneficiary') +
+  theme_minimal() +
+  scale_color_manual(values=cbbPalette)
+```
+
+## Effect of unevenness on antibiotic resistance
+
+Here we use the same bug/drug combinations, but instead we test the multivariable model mean percent nonsusceptible ~ fraction nonzero + nonzero Gini.
+
+There is only one significant and one marginally significant result:
+
+```{r multi_models}
+results %>%
+  filter(model=='multi', term==multi_term) %>%
+  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
+  filter(p.value < 0.05) %>%
+  arrange(p.value) %>%
+  select(bug, drug_group, estimate, p.value, p.adj) %>%
+  kable
+```
+
+```{r}
+results %>%
+  filter(model=='multi', term==multi_term) %>%
+  interval_plot() +
+  ylab('Gini coefficient')
+```
+
+Using HRRs, one of the answers remains the same, but a different marginally-significant result pops up.
+
+```{r}
+hrr_results %>%
+  filter(model=='multi', term==multi_term) %>%
+  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
+  filter(p.value < 0.05) %>%
+  arrange(p.value) %>%
+  select(bug, drug_group, estimate, p.value, p.adj) %>%
+  kable
+```
+
+```{r}
+hrr_results %>%
+  filter(model=='multi', term==multi_term) %>%
   interval_plot()
 ```
+
+## Specific plots
+
+Quinolone consumption predicts *E. coli* quinolone resistance:
+
+```{r ecoli_quin}
+left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
+  left_join(regions, by='state') %>%
+  filter(bug=='E. coli', drug_group=='quinolone') %>%
+  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, label=state)) +
+  geom_smooth(method='lm', col='gray', se=FALSE) +
+  geom_point() +
+  geom_text() +
+  xlim(0.2, 0.45) + ylim(0, 50) +
+  xlab('quinolone claims per beneficiary') + ylab('mean percent nonsusceptible E. coli isolates') +
+  theme_minimal()
+```
+
+## E.coli and bactrim
+
+```{r ecoli_bactrim}
+ecoli_bactrim_dat = left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
+  left_join(regions, by='state') %>%
+  filter(bug=='E. coli', drug_group=='tmp_smx') #%>%
+  #filter(!is.na(mean_percent_nonsusceptible))
+
+ecoli_bactrim_dat %>%
+  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, label=state)) +
+  geom_smooth(method='lm', col='gray', se=FALSE) +
+  geom_point() +
+  geom_text() +
+  scale_x_continuous(limits=c(0.08, 0.2), breaks=c(0.08, 0.1, 0.15, 0.2)) +
+  ylim(10, 45) +
+  xlab('TMP/SMX claims per beneficiary') + ylab('mean percent nonsusceptible E. coli isolates') +
+  theme_minimal()
+```
+
+```{r}
+ecoli_bactrim_dat %>%
+  ggplot(aes(x=nzg, y=mean_percent_nonsusceptible, label=state)) +
+  geom_smooth(method='lm', col='gray', se=FALSE) +
+  geom_point() +
+  geom_text() +
+  xlab('Gini coefficient)') +
+  ylab('E. coli resistance') +
+  theme_minimal()
+```
+
+## Pneumococcus and macrolides
+
+```{r}
+left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
+  left_join(regions, by='state') %>%
+  filter(bug=='S. pneumoniae', drug_group=='macrolide') %>%
+  filter(!is.na(mean_percent_nonsusceptible)) %>% 
+  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, label=state)) +
+  geom_smooth(method='lm', col='gray', se=FALSE) +
+  geom_point() +
+  geom_text() +
+  theme_minimal()
+```
+
+## E. cloacae and quinolones
+
+```{r}
+left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
+  left_join(regions, by='state') %>%
+  filter(bug=='Eb. cloacae', drug_group=='quinolone') %>%
+  filter(!is.na(mean_percent_nonsusceptible)) %>%
+  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, label=state)) +
+  geom_smooth(method='lm', col='gray', se=FALSE) +
+  geom_point() +
+  geom_text() +
+  theme_minimal()
+```
+
+## MSSA and clindamycin
+
+```{r mssa_clinda}
+left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
+  left_join(regions, by='state') %>%
+  filter(bug=='MSSA', drug_group=='clindamycin') %>%
+  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, label=state)) +
+  geom_smooth(method='lm', col='gray', se=FALSE) +
+  geom_point() +
+  geom_text() +
+  xlab('clindamycin claims per beneficiary') + ylab('mean percent nonsusceptible MSSA isolates') +
+  theme_minimal()
+```
+
+# Supplementary results
+
+## Metrics of unevenness
+
+### Aggregate usage correlates with fraction of users
+
+Error bars show the s.e.m. across the four data years.
+
+```{r fnz_mean}
+all_ineq %>%
+  mutate_at(vars(matches('fnz')), function(x) x*100) %>%
+  ggplot(aes(x=mean_cpb, xmin=mean_cpb-sem_cpb, xmax=mean_cpb+sem_cpb,
+             y=mean_fnz, ymin=mean_fnz-sem_fnz, ymax=mean_fnz+sem_fnz,
+             color=region, label=state_abbreviation)) +
+  geom_point() +
+  geom_linerange() +
+  geom_errorbarh(height=0) +
+  geom_text(nudge_x=0.04) +
+  xlab('mean claims per beneficiary') +
+  ylab('mean % beneficiaries with at least one claim') +
+  theme_minimal()
+```
+
+```{r}
+all_ineq %$%
+  cor.test(mean_cpb, mean_fnz, method='spearman')
+```
+
+### The two inequality metrics not correlate well
+
+Error bars show the s.e.m. across the four data years.
+
+```{r nzgini_mean}
+all_ineq %>%
+  ggplot(aes(x=mean_fnz, xmin=mean_fnz-sem_fnz, xmax=mean_fnz+sem_fnz,
+             y=mean_nzgini, ymin=mean_nzgini-sem_nzgini, ymax=mean_nzgini+sem_nzgini,
+             color=region, label=state_abbreviation)) +
+  geom_point() +
+  geom_linerange() +
+  geom_errorbarh(height=0) +
+  geom_text(nudge_x=0.00) +
+  xlab('mean fraction with at least one claim') +
+  ylab('mean Gini coefficient') +
+  theme_minimal()
+```
+
+The correlation exists but it's much less strong.
+
+```{r}
+all_ineq %$%
+  cor.test(mean_fnz, mean_nzgini, method='spearman')
+```
+
+There is an intermediate correlation between claims per beneficiary and the Gini coefficient (see below), so in multivariable regressions I will use those two (rather than CPB and Gini).
+
+```{r}
+all_ineq %$% cor.test(mean_cpb, mean_nzgini, method='spearman')
+```
+
+## Spearman models
+
+### State-level volcano plot
 
@@
 
-And the same thing using HRRs:
+### HRR-level correlations
 
@@
 
-## Univariable
+## Univariable regressions
 
-The model here is mean percent nonsusceptible ~ claims per beneficiary. The significant results are:
+Like the Spearman models, the model here is mean percent nonsusceptible ~ claims per beneficiary, only this time it's a normal linear regression.
+
+### State-level
 
@@
 
-And the same thing with HRRs:
+### HRR-level
 
@@
 ```
-
-## Multivariable
-
-The model here is mean percent nonsusceptible ~ fraction nonzero + nonzero Gini.
-In some of the models, the nonzero Gini term was significant, but not after adjusting for multiple hypotheses.
-I show just those terms that met the unadjusted threshold:
-
-```{r multi_models}
-results %>%
-  filter(model=='multi', term==multi_term) %>%
-  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
-  filter(p.value < 0.05) %>%
-  arrange(p.value) %>%
-  select(bug, drug_group, estimate, p.value, p.adj) %>%
-  kable
-```
-
-```{r}
-results %>%
-  filter(model=='multi', term==multi_term) %>%
-  interval_plot()
-```
-
-```{r}
-results %>%
-  filter(model=='multi', term==multi_term) %>%
-  mutate(bug_drug=interaction(bug, drug_group)) %>%
-  ggplot(aes(x=p.value, y=estimate, color=drug_group)) +
-  geom_point() +
-  scale_color_brewer(palette='Set1')
-```
-
-And what about looking at multivariate only for ones that passed the threshold for univariable?
-
-```{r eval=FALSE}
-sig_bug_drug = results %>%
-  filter(model=='uni', term==uni_term) %>%
-  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
-  filter(p.adj < 0.05) %>%
-  select(bug, drug_group)
-
-results %>%
-  filter(model=='multi', term==multi_term) %>%
-  right_join(sig_bug_drug, by=c('bug', 'drug_group')) %>%
-  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
-  select(bug, drug_group, estimate, p.value, p.adj) %>%
-  kable
-```
-
-Using HRRs, adjusting $p$ with all possibilities:
-
-```{r}
-hrr_results %>%
-  filter(model=='multi', term==multi_term) %>%
-  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
-  filter(p.value < 0.05) %>%
-  arrange(p.value) %>%
-  select(bug, drug_group, estimate, p.value, p.adj) %>%
-  kable
-```
-
-And then doing multiple hypothesis correction only for those that made it through the first model:
-
-```{r hrr_multi_models}
-sig_bug_drug = hrr_results %>%
-  filter(model=='uni', term==uni_term) %>%
-  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
-  filter(p.adj < 0.05) %>%
-  select(bug, drug_group)
-
-hrr_results %>%
-  filter(model=='multi', term==multi_term) %>%
-  right_join(sig_bug_drug, by=c('bug', 'drug_group')) %>%
-  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
-  arrange(p.value) %>%
-  select(bug, drug_group, estimate, p.value, p.adj) %>%
-  kable
-```
-
-## Two metrics
-
-It turns out that the different drugs have different relationships with the two metrics. Each point shows the consumption characteristics for a drug group in a state.
-
-```{r nzgini_cpb_drug}
-cbbPalette <- c("#000000", "#D55E00", "#CC79A7", "#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2")
-state_ineq %>%
-  ggplot(aes(x=cpb, y=nzg, color=drug_group)) +
-  geom_point() +
-  ylab('Gini coefficient among users') + xlab('claims per beneficiary') +
-  theme_minimal() +
-  scale_color_manual(values=cbbPalette)
-```
-
-```{r}
-left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
-  filter(bug %in% c('E. coli', 'Eb. cloacae', 'CNST', 'MSSA', 'MRSA', 'S. pneumoniae')) %>%
-  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, color=drug_group, shape=bug)) +
-  geom_point() +
-  xlab('claims per beneficiary')
-```
-
-## Specific plots
-
-Quinolone consumption predicts *E. coli* quinolone resistance:
-
-```{r ecoli_quin, eval=FALSE}
-left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
-  left_join(regions, by='state') %>%
-  filter(bug=='E. coli', drug_group=='quinolone') %>%
-  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region)) +
-  geom_smooth(aes(x=mean, y=mean_percent_nonsusceptible, color=NULL), method='lm', col='black') +
-  geom_point() +
-  xlab('quinolone claims per beneficiary') + ylab('mean percent nonsusceptible E. coli isolates') +
-  theme_minimal()
-
-left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
-  left_join(regions, by='state') %>%
-  filter(bug=='E. coli', drug_group=='quinolone') %>%
-  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, label=state)) +
-  geom_smooth(method='lm', col='black', se=FALSE) +
-  geom_point() +
-  geom_text() +
-  xlim(0.2, 0.45) + ylim(0, 50) +
-  xlab('quinolone claims per beneficiary') + ylab('mean percent nonsusceptible E. coli isolates') +
-  theme_minimal()
-```
-
-## E.coli and bactrim
-
-Bactrim resistance in *E. coli*. First, the aggregate relationship, then the relationship between "excess" resistance (the residual of resistance against aggregate consumption) against nonzero Gini.
-
-```{r ecoli_bactrim, eval=FALSE}
-ecoli_bactrim_dat = left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
-  left_join(regions, by='state') %>%
-  filter(bug=='E. coli', drug_group=='tmp_smx') #%>%
-  #filter(!is.na(mean_percent_nonsusceptible))
-
-ecoli_bactrim_dat %>%
-  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, label=state)) +
-  geom_smooth(method='lm', col='gray', se=FALSE) +
-  geom_point() +
-  geom_text() +
-  scale_x_continuous(limits=c(0.08, 0.2), breaks=c(0.08, 0.1, 0.15, 0.2)) +
-  ylim(10, 45) +
-  xlab('TMP/SMX claims per beneficiary') + ylab('mean percent nonsusceptible E. coli isolates') +
-  theme_minimal()
-
-ecoli_bactrim_dat %>%
-  { mutate(., excess_resistance=residuals(lm(mean_percent_nonsusceptible ~ mean, data=.))) } %>%
-  ggplot(aes(x=nzgini, y=excess_resistance)) +
-  geom_smooth(method='lm', col='gray', se=FALSE) +
-  geom_point() +
-  scale_x_continuous(limits=c(0.25, 0.375)) +
-  scale_y_continuous(limits=c(-12, 20), breaks=c(-12, 0, 10, 20)) +
-  xlab('Skew towards high consumers (Gini coefficient)') +
-  ylab('"Excess" E. coli resistance (see caption)') +
-  theme_minimal()
-
-ecoli_bactrim_dat %>%
-  ggplot(aes(x=nzgini, y=mean_percent_nonsusceptible, label=state)) +
-  geom_smooth(method='lm', col='gray', se=FALSE) +
-  geom_point() +
-  geom_text() +
-  xlab('Gini coefficient)') +
-  ylab('E. coli resistance') +
-  theme_minimal()
-```
-
-## Pneumococcus and macrolides
-
-```{r, eval=FALSE}
-left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
-  left_join(regions, by='state') %>%
-  filter(bug=='S. pneumoniae', drug_group=='macrolide') %>%
-  filter(!is.na(mean_percent_nonsusceptible)) %>% 
-  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, label=state)) +
-  geom_smooth(method='lm', col='gray', se=FALSE) +
-  geom_point() +
-  geom_text() +
-  theme_minimal()
-```
-
-## E. cloacae and quinolones
-
-```{r, eval=FALSE}
-left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
-  left_join(regions, by='state') %>%
-  filter(bug=='Eb. cloacae', drug_group=='quinolone') %>%
-  filter(!is.na(mean_percent_nonsusceptible)) %>%
-  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, label=state)) +
-  geom_smooth(method='lm', col='gray', se=FALSE) +
-  geom_point() +
-  geom_text() +
-  theme_minimal()
-```
-
-## MSSA and clindamycin
-
-```{r mssa_clinda, eval=FALSE}
-dat = left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
-  left_join(regions, by='state') %>%
-  filter(bug=='MSSA', drug_group=='clindamycin') #%>%
-  #filter(!is.na(mean_percent_nonsusceptible))
-
-dat %>%
-  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, label=state)) +
-  geom_smooth(method='lm', col='gray', se=FALSE) +
-  geom_point() +
-  geom_text() +
-  #scale_x_continuous(limits=c(0.08, 0.2), breaks=c(0.08, 0.1, 0.15, 0.2)) +
-  #ylim(10, 45) +
-  xlab('clindamycin claims per beneficiary') + ylab('mean percent nonsusceptible MSSA isolates') +
-  theme_minimal()
-
-dat %>%
-  { mutate(., excess_resistance=residuals(lm(mean_percent_nonsusceptible ~ mean, data=.))) } %>%
-  ggplot(aes(x=nzgini, y=excess_resistance)) +
-  geom_smooth(method='lm', col='gray', se=FALSE) +
-  geom_point() +
-  #scale_x_continuous(limits=c(0.25, 0.375)) +
-  #scale_y_continuous(limits=c(-12, 20), breaks=c(-12, 0, 10, 20)) +
-  xlab('Gini coefficient') +
-  ylab('"Excess" MSSA resistance') +
-  theme_minimal()
-```
-
-## MRSA clinda
-
-```{r, eval=FALSE}
-left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
-  left_join(regions, by='state') %>%
-  filter(bug=='MRSA', drug_group=='clindamycin') %>%
-  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, label=state)) +
-  geom_smooth(method='lm', col='gray', se=FALSE) +
-  geom_point() +
-  geom_text() +
-  #scale_x_continuous(limits=c(0.08, 0.2), breaks=c(0.08, 0.1, 0.15, 0.2)) +
-  #ylim(10, 45) +
-  xlab('clindamycin claims per beneficiary') + ylab('mean percent nonsusceptible MRSA isolates') +
-  theme_minimal()
-```

### 2017-05-29 sex-difference/sex-difference.Rmd
sex difference

@@
 bene = read_tsv(sprintf('../bene_%i.tsv', year)) %>%
-  filter(between(age, 66, 96), hmo_months==0, sex %in% c('male', 'female')) %>%
   mutate(is_female=sex=='female', is_white=race=='white')
 
-pde = read_tsv(sprintf('../pde_%i.tsv', year)) %>%
-  filter(bene_id %in% bene$bene_id)
+pde = read_tsv(sprintf('../pde_%i.tsv', year))
 
@@
 ubs %>%
-  select(antibiotic, sex, cpkp) %>%
-  spread(sex, cpkp) %>%
   ggplot(aes(x=male, y=female, label=antibiotic)) +
-  geom_label() +
+  geom_text() +
   geom_point() +
@@
   ylab('claims per 1,000 female beneficiaries') +
-  theme_minimal()
+  theme_minimal() +
+  coord_fixed()
 ```

### 2017-06-01 sex-difference/sex-difference.Rmd
updating sex codes

@@
 ---
+
+```{r global_options, include=FALSE}
+knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/', dev=c('png', 'pdf'),
+                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE, autodep=TRUE)
+```
 
@@
 
+## Diagnosis codes
+
+For COPD, I use the codes recommended by the Chronic Condition Warehouse, but I exclude 490 ("Bronchitis, not specified as acute or chronic") because acute bronchitis seems more common than
+
 # Results
-
-```{r global_options, include=FALSE}
-knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/', dev=c('png', 'pdf'),
-                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE, autodep=TRUE)
-```
 
@@
 
+icd = read_tsv('../../db/icd/icd.tsv') %>%
+  rename(dx_code=code, dx_desc=desc)
+
+sex_icd = read_tsv('sex_codes.tsv', col_types=cols(dx_code='c'))
+```
+
+```{r load_dx}
+rc_dxs = c('4610', '4619', '462', '4659', '4660', '4739', '486', '490', '7862')
+uti_dxs = c('5950', '5959', '5989', '5990', '6235', '7881')
+
+dx = read_tsv(sprintf('../../data/dx/dx_car_claims_%i.tsv', year), col_types=cols(diagnosis='c')) %>%
+  select(bene_id=BENE_ID, dx_code=diagnosis)
+
+dx_counts = dx %>% count(dx_code) %>% left_join(icd, by='dx_code') %>% arrange(desc(n))
+
+stopifnot(all(dx_counts$dx_code %in% c(rc_dxs, uti_dxs)))
+
 total_pde = pde %>% count(bene_id) %>% rename(n_claims=n)
+```
+
+## Diagnostic codes
+
+For UTIs:
+
+```{r}
+sex_icd %>%
+  left_join(dx_counts, by='dx_code') %>%
+  arrange(dx_group, desc(n)) %>%
+  kable
+```
+
+For RCs:
+
+```{r}
+dx_counts %>% filter(dx_code %in% rc_dxs) %>% kable
 ```

### 2017-06-04 ms1/ms1_display.Rmd
ms1 cohort

--- 2017-05-23 ms1/ms1_display.Rmd
@@
 import::from(stringr, str_replace, str_detect)
+import::from(forcats, as_factor)
 
@@
 
+And the same in the cohort:
+
+```{r}
+read_tsv('tables/tbl_cohort_totals.tsv') %>%
+  pctize('n_bene') %>%
+  select(year, n_bene, mean_age, mean_n_cc,
+         pct_female, pct_white, pct_dual,
+         pct_region_south, pct_region_midwest, pct_region_northeast, pct_region_west) %>%
+  mutate_at(vars(starts_with('pct_')), function(x) round(x, digits=1)) %>%
+  mutate(n_bene=prettyNum(n_bene, big.mark=','),
+         mean_age=round(mean_age, digits=1),
+         mean_n_cc=round(mean_n_cc, digits=2)) %>%
+  gather('var', 'val', -year) %>% spread(year, val) %>% # transpose
+  kable
+```
+
 # Patterns of consumption
@@
 dat = read_tsv('tables/tbl_claims_by_abx.tsv')
-totals = dat %>% group_by(year) %>% summarize(cpkp=sum(cpkp)) %>% mutate(antibiotic='all')
+cohort_dat = read_tsv('tables/tbl_cohort_claims_by_abx.tsv')
+
 top_abx = dat %>%
@@
 
-x = dat %>%
+top_abx_f = function(df) {
+  totals = group_by(df, year) %>%
+    summarize(cpkp=sum(cpkp)) %>%
+    mutate(antibiotic='all')
+
+  df %>%
     select(antibiotic, year, cpkp) %>%
@@
     mutate(antibiotic=factor(antibiotic, levels=c('all', top_abx)))
+}
 
-x %>%
-  mutate(cpkp=round(cpkp)) %>%
-  spread(year, cpkp) %>%
-  arrange(desc(`2011`)) %>%
-  kable
+#top_abx_f(dat) %>%
+#  mutate(cpkp=round(cpkp)) %>%
+#  spread(year, cpkp) %>%
+#  arrange(desc(`2011`)) %>%
+#  kable
 
-x %>% ggplot(aes(x=year, y=cpkp)) +
+all_dat = bind_rows(
+  top_abx_f(dat) %>% mutate(group='all'),
+  top_abx_f(cohort_dat) %>% mutate(group='cohort')
+)
+
+all_dat %>%
+  ggplot(aes(x=year, y=cpkp, color=group)) +
   facet_wrap(~ antibiotic, ncol=3, scales='free_y') +
   geom_point() + geom_line() +
-  theme_minimal()
+  theme_minimal() +
+  ggtitle('Unadjusted consumption, overall or in cohort')
+```
+
+This shows a subset of the above data, but using percent changes from 2011 to 2014:
+
+```{r}
+all_dat %>%
+  arrange(year) %>%
+  group_by(group, antibiotic) %>%
+  summarize(mean_cpkp=mean(cpkp),
+            pct_change=(tail(cpkp, 1)/head(cpkp, 1) - 1)*100) %>%
+  ungroup() %>%
+  filter(antibiotic %in% top_abx) %>%
+  { levels = filter(., group=='cohort') %>% arrange(pct_change) %$% antibiotic;
+    mutate(., antibiotic=factor(.$antibiotic, levels=levels)) } %>%
+  ggplot(aes(x=antibiotic, y=pct_change, color=group)) +
+  geom_point() +
+  coord_flip()
 ```
@@
 
+# Adjusted consumption changes
+
+Only in the cohort:
+
+```{r}
+read_tsv('tables/tbl_cohort_model_abx.tsv') %>%
+  filter(term=='year', model %in% top_abx) %>%
+  arrange(estimate) %>%
+  mutate(cpkp_per_year=estimate*1000,
+         antibiotic=as_factor(model)) %>%
+  select(antibiotic, cpkp_per_year) %>%
+  ggplot(aes(x=antibiotic, y=cpkp_per_year)) +
+  geom_point() +
+  coord_flip() +
+  ggtitle('Adjusted changes in consumption') +
+  theme_minimal()
+```
+
 # Prevalence of conditions
 
-```{r}
-condition_names = read_tsv('../chronic-conditions/names.txt',
-                           col_names=c('condition', 'condition_name'))
+```{r, eval=FALSE}
+condition_names = read_tsv('../chronic-conditions/names.tsv')
 

### 2017-06-04 ms1/ms1_cohort.R

@@ -3,14 +3,22 @@
 import::from(broom, tidy)
 import::from(stringr, str_replace)
 
+sum_over = function(target_columns) {
+  dummy_vars = sapply(1:length(target_columns), function(i) sprintf('x%i', i))
+  dummy_vars_string = paste0(dummy_vars, collapse=',')
+  dummy_vars_list = lapply(target_columns, as.name) %>% setNames(dummy_vars)
+  f = as.formula(paste0(c('~sum(', dummy_vars_string, ')'), collapse=''))
+  lazyeval::interp(f, .values=dummy_vars_list)
+}
+
 # Load the census regions. Code them as factors so that Northeast is taken as
 # the baseline in the linear models.
 # NB: I put DC into the South
 regions = read_tsv('../../db/census-regions/census-regions.tsv') %>%
   select(state, region)
 
-condition_names = read_tsv('../chronic-conditions/names.txt',
-                           col_names=c('condition', 'condition_name'))
+condition_names = read_tsv('../chronic-conditions/names.tsv')
+names_1yr_ccs = condition_names %>% filter(condition_ref_years==1) %$% condition_code
 
 load_data = function(year) {
   bene = read_tsv(sprintf('../bene_%i.tsv', year)) %>%
@@ -30,7 +38,10 @@ load_data = function(year) {
 
   cc = read_feather(sprintf('../cc_%i.feather', year)) %>%
     select(-ALZH) %>%
-    mutate(n_cc=rowSums(select_(., '-bene_id')))
+    rowwise() %>%
+    mutate_(n_1yr_cc=sum_over(names_1yr_ccs),
+            n_all_cc=sum_over(condition_names$condition_code)) %>%
+    ungroup()
 
   bene %<>% left_join(cc, by='bene_id')
 

### 2017-06-05 sex-difference/sex-difference.Rmd
sex

@@
 ubs %>%
-  ggplot(aes(x=male, y=female, label=antibiotic)) +
+  mutate(highlight=antibiotic %in% c('azithromycin', 'ciprofloxacin', 'trimethoprim/sulfamethoxazole', 'fosfomycin', 'nitrofurantoin')) %>%
+  ggplot(aes(x=male, y=female, label=antibiotic, color=highlight)) +
   geom_text() +

### 2017-06-13 ms1/ms1_display.Rmd
updating ms1

@@
 }
+
+trend_p = function(x) t.test(diff(x))$p.value
 ```
@@
 
-And the same in the cohort:
+In the cohort, all the values remain fixed, except for:
+
+- age
+- chronic conditions
+- to a lesser degree, dual enrollment
+- to a very small degree, geographical location
 
@@
 
-#top_abx_f(dat) %>%
-#  mutate(cpkp=round(cpkp)) %>%
-#  spread(year, cpkp) %>%
-#  arrange(desc(`2011`)) %>%
-#  kable
-
 all_dat = bind_rows(
@@
 
-This shows a subset of the above data, but using percent changes from 2011 to 2014:
+P-values for trends in the data. (These are $t$-tests on the differences between subsequent years, comparing their mean to zero.)
+
+```{r}
+all_dat %>%
+  arrange(group, antibiotic, year) %>%
+  group_by(group, antibiotic) %>%
+  summarize(p.value=trend_p(cpkp)) %>%
+  ungroup() %>%
+  mutate(adj.p.value=p.adjust(p.value, method='BH')) %>%
+  select(antibiotic, group, adj.p.value) %>%
+  spread(group, adj.p.value) %>%
+  mutate(sig=if_else(all < 0.05 | cohort < 0.05, '*', '')) %>%
+  kable
+```
+
+This shows a summary of the above data, using percent changes from 2011 to 2014:
 
@@
 
+"Adjusted" means the `year` coefficient in `y ~ year + age*n_cc + is_female + is_dual + is_white + region`.
+
 Only in the cohort:
@@
 
-# Prevalence of conditions
+# Azithro and levo
 
-```{r, eval=FALSE}
-condition_names = read_tsv('../chronic-conditions/names.tsv')
+I used the encounter data to anotate which beneficiaries had COPD or an ARTI in that year. I used these primary diagnoses:
 
-read_tsv('tables/tbl_cc_prevalence.tsv') %>%
-  left_join(condition_names, by='condition') %>%
-  select(condition=condition_name, year, prevalence) %>%
-  spread(year, prevalence) %>%
-  mutate(change=(`2014`-`2011`)/`2011`) %>%
-  mutate_at(vars(-condition), function(x) round(x * 100, 1)) %>%
-  arrange(desc(change)) %>%
+```{r}
+read_tsv('../sex-difference/sex_codes.tsv') %>%
+  filter(diagnosis_type %in% c('acute_rc', 'copd')) %>%
   kable
@@
 
-# All-condition model
+Then I ran models predicting azithromycin and levofloxacin consumption using these some standard demographic variables and a narrowly-defined health status (COPD and ARTI).
 
-```{r, eval=FALSE}
-read_tsv('tables/tbl_all_condition_model.tsv') %>%
-  mutate(term=stringr::str_replace(term, 'TRUE$', '')) %>%
-  left_join(condition_names, by=c('term'='condition')) %>%
-  mutate(term=if_else(is.na(condition_name), term, condition_name),
-         log10.adj.p.value=log10(p.adjust(p.value, method='BH'))) %>%
-  select(term, estimate, p.value, log10.adj.p.value) %>%
-  filter(term != '(Intercept)') %>%
-  arrange(desc(estimate)) %>%
-  kable
+## Azithromycin
+
+`y ~ year + start_age + is_female + is_dual + is_white + region + copd + copd:year + copd:start_age + acute_rc + acute_rc:year + acute_rc:start_age`
+
+That is, add COPD, ARTI, and the interactions of those terms with beneficiary age (in 2011) and calendar year.
+
+```{r}
+read_tsv('tables/tbl_cohort_model_azithro.tsv') %>% kable
 ```
+
+Azithromycin prescribing decreases with calendar year and with age, even for those who have COPD or an ARTI.
+
+## Levofloxacin
+
+`y ~ year + start_age + is_female + is_dual + is_white + region + acute_rc + acute_rc:year + acute_rc:start_age`
+
+That is, add ARTI and the interactions with beneficiary age and calendar year.
+
+```{r}
+read_tsv('tables/tbl_cohort_model_levo.tsv') %>% kable
+```
+
+By contrast, levofloxacin prescribing increases with calendar year and with age, including for those who have and ARTI.
+The overall increase with time is about half the size of the increase with time for those who have an ARTI, meaning that levofloxacin, aside from being prescribed more overall, is being prescribed more especially for those with ARTIs.
+
+Interestingly, the effect of dual-enrollment is much higher for levofloxacin than azithromycin (about triple). Why would being poorer make you more likely to take levofloxacin?

### 2017-06-13 sex-difference/sex-difference.Rmd
updating ms1

--- 2017-05-29 sex-difference/sex-difference.Rmd
@@
 
-*Outcome*: At least one claim for azithromycin, Bactrim, TMP/SMX, or nitrofurantoin. *Exposure*: A diagnosis with a UTI or an acute RTI.
+*Outcome*: At least one claim for azithromycin, Bactrim, TMP/SMX, or nitrofurantoin. *Exposure*: A diagnosis with a UTI, COPD, or an acute RTI.
 
@@
 
-# Results
-
-```{r global_options, include=FALSE}
-knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/', dev=c('png', 'pdf'),
-                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE, autodep=TRUE)
-```
-
 ```{r load_data}
@@
 
+sex_icd = read_tsv('sex_codes.tsv', col_types=cols(code='c')) %>%
+  rename(dx_code=code, dx_desc=short_desc)
+```
+
+```{r load_dx}
+dx = read_tsv(sprintf('../../data/dx/dx_car_claims_%i.tsv', year), col_types=cols(diagnosis='c')) %>%
+  select(bene_id=BENE_ID, dx_code=diagnosis) %>%
+  left_join(sex_icd, by='dx_code')
+
+dx_counts = dx %>%
+  count(dx_code)
+
 total_pde = pde %>% count(bene_id) %>% rename(n_claims=n)
 ```
+
+## Diagnostic codes and counts
+
+```{r}
+sex_icd %>%
+  left_join(dx_counts, by='dx_code') %>%
+  kable
+```
+
+# Results
 
@@
 
-```{r drug_by_sex_residuals}
+```{r}
+dbs %>% head %>% kable
+```
+
+```{r drug_by_sex_residuals, eval=FALSE}
 dbs %>%
@@
 
-```{r women_azithro_vs_uai}
+```{r women_azithro_vs_uai, eval=FALSE}
 any_in = function(lst, elts) match(elts, lst) %>% { .[!is.na(.)] } %>% { length(.) > 0 }
@@
 
-```{r sex_by_age}
+```{r sex_by_age, eval=FALSE}
 plot_dat = usage %>%
@@
 
-```{r load_carrier_data}
+```{r load_carrier_data, eval=FALSE}
 carrier = read_tsv('../../data/carrier_count_2011.tsv') %>%

### 2017-06-13 ms1/ms1_work.R

@@ -19,6 +19,9 @@ regions = read_tsv('../../db/census-regions/census-regions.tsv') %>%
 condition_names = read_tsv('../chronic-conditions/names.tsv') %>% filter(condition_code != 'ALZH')
 names_1yr_ccs = condition_names %>% filter(condition_ref_years==1) %$% condition_code
 
+dx_codes = read_tsv('../sex-difference/sex_codes.tsv') %>%
+  select(code, diagnosis_type)
+
 load_data = function(year) {
   bene = read_tsv(sprintf('../bene_%i.tsv', year)) %>%
     mutate(is_female=sex=='female', is_white=race=='white', is_dual=buyin_months>0) %>%
@@ -33,12 +36,17 @@ load_data = function(year) {
     summarize(n_claims=n(), days_supply=sum(days_supply)) %>%
     ungroup() %>%
     mutate(year=year)
+  
+  # diagnosis claims
+  dx = read_tsv(sprintf('../../data/dx_car_claims_%i.tsv', year)) %>%
+    select(bene_id=BENE_ID, code=diagnosis) %>%
+    left_join(dx_codes, by='code') %>%
+    distinct %>%
+    mutate(dummy=TRUE) %>%
+    spread(diagnosis_type, dummy)
 
   # get total numbers of PDEs
   n_claims = pde %>% group_by(bene_id) %>% summarize(n_claims=sum(n_claims))
-  bene %<>%
-    left_join(n_claims, by='bene_id') %>%
-    replace_na(list(n_claims=0))
 
   # count chronic conditions. 'n_cc' means one-year cc's
   cc = read_feather(sprintf('../cc_%i.feather', year)) %>%
@@ -46,7 +54,13 @@ load_data = function(year) {
     mutate_(n_cc=sum_over(names_1yr_ccs)) %>%
     ungroup()
 
-  bene %<>% left_join(cc, by='bene_id')
+  # join the cc, summary PDE, and dx data into bene
+  bene %<>%
+    left_join(cc, by='bene_id') %>%
+    left_join(n_claims, by='bene_id') %>% replace_na(list(n_claims=0)) %>%
+    left_join(dx, by='bene_id') %>% replace_na(list(acute_rc=FALSE,
+                                                    copd=FALSE,
+                                                    uti=FALSE))
 
   list(bene=bene, pde=pde)
 }
@@ -193,9 +207,9 @@ lapply(top_abx, f) %>%
   output_table('cohort_model_abx')
 
 # special levo & azithro models
[…]

### 2017-06-16 sex-difference/sex-difference.Rmd
sex difference present

@@
 ```
+
+# To do
+
+- Separate beneficiaries into "high users" and "low users" based on the median number of outpatient claim dates
+- Try segmenting azithromycin consumption by low/high in wo/men and by RC+/- in wo/men.
+We want to say that RC, more than low/high, accounts for differences in azithromycin risk, i.e., it's that women get more respiratory complaint diagnoses, not that women just have more complaints.
+Or vice versa.
 
@@
 year = 2011
-bene = read_tsv(sprintf('../bene_%i.tsv', year)) %>%
-  mutate(is_female=sex=='female', is_white=race=='white')
 
@@
   rename(dx_code=code, dx_desc=short_desc)
-```
 
-```{r load_dx}
-dx = read_tsv(sprintf('../../data/dx/dx_car_claims_%i.tsv', year), col_types=cols(diagnosis='c')) %>%
+dx = read_tsv(sprintf('../../data/dx_car_claims_%i.tsv', year), col_types=cols(diagnosis='c')) %>%
   select(bene_id=BENE_ID, dx_code=diagnosis) %>%
@@
 
+bene_dx = dx %>%
+  select(bene_id, diagnosis_type) %>%
+  distinct() %>%
+  mutate(val=TRUE) %>%
+  spread(diagnosis_type, val)
+
 total_pde = pde %>% count(bene_id) %>% rename(n_claims=n)
+
+bene = read_tsv(sprintf('../bene_%i.tsv', year)) %>%
+  mutate(is_female=sex=='female', is_white=race=='white') %>%
+  left_join(total_pde, by='bene_id') %>% replace_na(list(n_claims=0)) %>%
+  left_join(bene_dx, by='bene_id') %>% replace_na(list(acute_rc=FALSE, copd=FALSE, uti=FALSE))
 ```
@@
   left_join(dx_counts, by='dx_code') %>%
-  kable
+  arrange(diagnosis_type, desc(n)) %>%
+  mutate(n=format(n, big.mark=',')) %>%
+  kable(align='lllr')
 ```
@@
 # Results
-
-## Consumption by age and sex
-
-It seems that females consume more antibiotics even within each age bracket (except for those over 95).
-
-```{r consumption_by_age_and_sex}
-bene %>%
-  left_join(total_pde, by='bene_id') %>%
-  replace_na(list(n_claims=0)) %>%
-  group_by(sex, age) %>%
-  summarize(mean_n_claims=mean(n_claims)) %>%
-  ungroup() %>%
-  ggplot(aes(x=age, y=mean_n_claims, color=sex)) +
-  geom_point()
-```
 
@@
   bene %>%
+    select(-n_claims) %>%
     left_join(pde %>% filter(antibiotic==drug) %>% count(bene_id) %>% rename(n_claims=n), by='bene_id') %>%
@@
   ggplot(aes(x=male, y=female, label=antibiotic, color=highlight)) +
+  scale_color_manual(values=c('black', 'red')) +
   geom_text() +
@@
 
-For a more quantitative approach, this is a table showing, for each antibiotic, the CPKP for female and males, the difference (residual) between the two, and the residual fraction (the residual over the mean of the female and male consumptions).
-Positive residuals means more consumption by females.
+Alternatively, I look at the mean usage (across men and women), the residual (the difference between the fraction of female beneficiaries taking the drug and male beneficiaries), and the fractional residual (the residual over the mean).
+I keep only those drugs with mean fraction at least 1%, and I also keep fosfomycin.
 
 ```{r}
-dbs %>% head %>% kable
+ubs %>%
+  mutate(mean_fraction=0.5*(male+female),
+         residual=female-male,
+         fractional_residual=residual/mean_fraction) %>%
+  filter(mean_fraction > 0.01 | antibiotic=='fosfomycin') %>%
+  ggplot(aes(x=mean_fraction, y=fractional_residual, label=antibiotic)) +
+  geom_point() +
+  geom_text()
 ```
 
-```{r drug_by_sex_residuals, eval=FALSE}
-dbs %>%
-  select(antibiotic, sex, cpkp) %>%
-  spread(sex, cpkp) %>%
-  mutate(residual=female-male, residual_fraction=residual/(0.5 * (male + female))) %>%
-  arrange(desc(abs(residual))) %>%
-  head(10) %>%
-  kable
+This makes the same four look interesting; maybe levofloxacin and metronidazole would be next. Levo fits in with the ARTI story; metro does not.
+
+## UTIs
+
+Are the increased consumptions of cipro and TMP/SMX explained by UTIs? That is, do men and women who have UTIs get cipro and TMP/SMX at similar rates, and it's just that women get more UTIs than men?
+
+### Cipro
+
+```{r}
+outcome_bene = pde %>% filter(antibiotic=='ciprofloxacin') %$% bene_id %>% unique
+epi = bene %>%
+  mutate(outcome=bene_id %in% outcome_bene) %>%
+  count(sex, uti, outcome) %>%
+  ungroup()
+```
+
+```{r}
+group_by_all_ = function(df, group_name) {
+  bind_rows(df, mutate_(df, .dots=setNames(list(~'all'), group_name))) %>%
+    group_by_(group_name)
+}
+
+epi %>%
+  group_by_all_('sex') %>%
+  summarize(n_cases=sum(n)) %>%
+
 ```
@@
 
-```{r azithromycin_usage}
+```{r azithromycin_usage, eval=FALSE}
 na_to_zero = function(x) if_else(is.na(x), 0L, x)
@@
 
-```{r azithromycin_usage_table}
+```{r azithromycin_usage_table, eval=FALSE}
 azi_usage %>%
@@
 
-```{r azithromycin_usage_model}
+```{r azithromycin_usage_model, eval=FALSE}
 library(broom)
@@
 The coefficient for being female (0.05) accounts for almost the entire difference in consumption between men and women (56 per thousand). Note that age has an approximately zero correlation (i.e.,
+
+# Supplement
+
+## Consumption by age and sex
+
+At every age, women consume more antibiotics than men.
+
+```{r consumption_by_age_and_sex}
+bene %>%
+  left_join(total_pde, by='bene_id') %>%
+  replace_na(list(n_claims=0)) %>%
+  group_by(sex, age) %>%
+  summarize(mean_n_claims=mean(n_claims)) %>%
+  ungroup() %>%
+  ggplot(aes(x=age, y=mean_n_claims, color=sex)) +
+  geom_point()
+```

### 2017-06-16 sex-difference/risk_ratio_ci.py

@@ -55,3 +55,6 @@ print()
 # Azithromycin uses respiratory complaints as denominators
 print('women', 'azithro', rrci_total(177843, 402523, 109681, 1153928, 0.95))
 print('men', 'azithro', rrci_total(84145, 208697, 54134, 713150, 0.95))
+
+# Cipro, with the new data
+print('women', 'cipro', 'newdata', rrci(122298, 127054, 138795, 1168304, 0.95))

### 2017-06-16 ms1/ms1_work.R
fixed doubling problems

@@ -23,13 +23,7 @@ dx_codes = read_tsv('../sex-difference/sex_codes.tsv') %>%
   select(code, diagnosis_type)
 
 load_data = function(year) {
-  bene = read_tsv(sprintf('../bene_%i.tsv', year)) %>%
-    mutate(is_female=sex=='female', is_white=race=='white', is_dual=buyin_months>0) %>%
-    left_join(regions, by='state') %>%
-    mutate(year=year)
-
-  # get the PDE data. record the number of claims (and days) for each bene
-  # and drug
+  # pde data
   pde = read_tsv(sprintf('../pde_%i.tsv', year)) %>%
     select(bene_id, antibiotic, days_supply) %>%
     group_by(bene_id, antibiotic) %>%
@@ -46,7 +40,7 @@ load_data = function(year) {
     mutate(dummy=TRUE) %>%
     spread(diagnosis_type, dummy)
 
-  # get total numbers of PDEs
+  # total numbers of PDEs
   n_claims = pde %>% group_by(bene_id) %>% summarize(n_claims=sum(n_claims))
 
   # count chronic conditions. 'n_cc' means one-year cc's
@@ -56,7 +50,10 @@ load_data = function(year) {
     ungroup()
 
   # join the cc, summary PDE, and dx data into bene
-  bene %<>%
+  bene = read_tsv(sprintf('../bene_%i.tsv', year)) %>%
+    mutate(is_female=sex=='female', is_white=race=='white', is_dual=buyin_months>0) %>%
+    left_join(regions, by='state') %>%
+    mutate(year=year) %>%
     left_join(cc, by='bene_id') %>%
     left_join(n_claims, by='bene_id') %>% replace_na(list(n_claims=0)) %>%
     left_join(dx, by='bene_id') %>% replace_na(list(acute_rc=FALSE,
@@ -76,9 +73,9 @@ bene = lapply(dat, function(df) df$bene) %>% bind_rows %>%
   ungroup()
 
 pde = lapply(dat, function(df) df$pde) %>% bind_rows %>%
-  left_join(select(bene, bene_id, in_cohort), by='bene_id')
+  left_join(distinct(select(bene, bene_id, in_cohort)), by='bene_id')
 
-rm(dat)
+#rm(dat)
 
 output_table = function(x, base) write_tsv(x, sprintf('tables/tbl_%s.tsv', base))
 sem = function(x) sd(x) / length(x)
[…]

### 2017-06-16 ms1/ms1_display.Rmd
updated display

@@
 
-```{r global_options, include=FALSE, cache=FALSE}
+```{r global_options, include=FALSE}
 knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='fig/', dev=c('pdf', 'png'),
@@
 
-```{r}
-read_tsv('tables/tbl_totals.tsv') %>%
-  pctize('n_bene') %>%
-  select(year, n_bene, mean_age, mean_n_cc,
-         pct_female, pct_white, pct_dual,
-         pct_region_south, pct_region_midwest, pct_region_northeast, pct_region_west) %>%
-  mutate_at(vars(starts_with('pct_')), function(x) round(x, digits=1)) %>%
-  mutate(n_bene=prettyNum(n_bene, big.mark=','),
-         mean_age=round(mean_age, digits=1),
-         mean_n_cc=round(mean_n_cc, digits=2)) %>%
-  gather('var', 'val', -year) %>% spread(year, val) %>% # transpose
-  kable
-```
-
 In the cohort, all the values remain fixed, except for:
@@
 ```{r}
-read_tsv('tables/tbl_cohort_totals.tsv') %>%
+totals_dat = read_tsv('tables/tbl_cohort_totals.tsv') %>%
   pctize('n_bene') %>%
-  select(year, n_bene, mean_age, mean_n_cc,
+  select(year, n_bene, mean_age, sd_age, mean_n_cc, sd_n_cc,
          pct_female, pct_white, pct_dual,
-         pct_region_south, pct_region_midwest, pct_region_northeast, pct_region_west) %>%
+         pct_region_south, pct_region_midwest, pct_region_northeast, pct_region_west)
+
+totals_dat %>%
   mutate_at(vars(starts_with('pct_')), function(x) round(x, digits=1)) %>%
   mutate(n_bene=prettyNum(n_bene, big.mark=','),
-         mean_age=round(mean_age, digits=1),
-         mean_n_cc=round(mean_n_cc, digits=2)) %>%
+         mean_age=round(mean_age, digits=1)) %>%
+  mutate_at(vars(mean_n_cc, sd_age, sd_n_cc),
+            function(x) round(x, digits=2)) %>% 
   gather('var', 'val', -year) %>% spread(year, val) %>% # transpose
+  kable
+```
+
+```{r}
+linear_trend_p = function(x) tidy(t.test(diff(x)))$p.value
+
+totals_dat %>%
+  gather('var', 'val', -year) %>%
+  arrange(var, year, val) %>%
+  group_by(var) %>%
+  summarize(p=linear_trend_p(val)) %>%
+  mutate(adj.p=p.adjust(p, method='BH'),
+         sig=if_else(adj.p<0.05, '*', '')) %>%
   kable
@@
 ```{r}
-dat = read_tsv('tables/tbl_claims_by_abx.tsv')
 cohort_dat = read_tsv('tables/tbl_cohort_claims_by_abx.tsv')
 
-top_abx = dat %>%
+top_abx = cohort_dat %>%
   group_by(antibiotic) %>%
@@
 }
+```
 
-all_dat = bind_rows(
-  top_abx_f(dat) %>% mutate(group='all'),
-  top_abx_f(cohort_dat) %>% mutate(group='cohort')
-)
-
-all_dat %>%
-  ggplot(aes(x=year, y=cpkp, color=group)) +
+```{r}
+cohort_dat %>%
+  top_abx_f() %>%
+  ggplot(aes(x=year, y=cpkp)) +
   facet_wrap(~ antibiotic, ncol=3, scales='free_y') +
@@
   theme_minimal() +
-  ggtitle('Unadjusted consumption, overall or in cohort')
+  ggtitle('Unadjusted consumption, in cohort')
+
+cohort_dat %>%
+  top_abx_f() %>%
+  filter(year %in% c(2011, 2014)) %>%
+  spread(year, cpkp) %>%
+  mutate(pct_change=(`2014`-`2011`)/`2011`*100) %>%
+  kable
 ```
@@
 ```{r}
-all_dat %>%
-  arrange(group, antibiotic, year) %>%
-  group_by(group, antibiotic) %>%
+cohort_dat %>%
+  top_abx_f() %>%
+  arrange(antibiotic, year) %>%
+  group_by(antibiotic) %>%
   summarize(p.value=trend_p(cpkp)) %>%
@@
   mutate(adj.p.value=p.adjust(p.value, method='BH')) %>%
-  select(antibiotic, group, adj.p.value) %>%
-  spread(group, adj.p.value) %>%
-  mutate(sig=if_else(all < 0.05 | cohort < 0.05, '*', '')) %>%
+  select(antibiotic, adj.p.value) %>%
+  mutate(sig=if_else(adj.p.value < 0.05, '*', '')) %>%
   kable
-```
-
-This shows a summary of the above data, using percent changes from 2011 to 2014:
-
-```{r}
-all_dat %>%
-  arrange(year) %>%
-  group_by(group, antibiotic) %>%
-  summarize(mean_cpkp=mean(cpkp),
-            pct_change=(tail(cpkp, 1)/head(cpkp, 1) - 1)*100) %>%
-  ungroup() %>%
-  filter(antibiotic %in% top_abx) %>%
-  { levels = filter(., group=='cohort') %>% arrange(pct_change) %$% antibiotic;
-    mutate(., antibiotic=factor(.$antibiotic, levels=levels)) } %>%
-  ggplot(aes(x=antibiotic, y=pct_change, color=group)) +
-  geom_point() +
-  coord_flip()
 ```
@@
 ```{r claims_by_age}
-read_tsv('tables/tbl_claims_by_age.tsv') %>%
+read_tsv('tables/tbl_cohort_claims_by_age.tsv') %>%
   mutate(cpkp=round(cpkp)) %>%
@@
 ```{r}
-read_tsv('tables/tbl_claims_by_sex.tsv') %>%
+read_tsv('tables/tbl_cohort_claims_by_sex.tsv') %>%
   select(sex, year, cpkp) %>%
@@
 ```{r}
-read_tsv('tables/tbl_claims_by_region.tsv') %>%
+read_tsv('tables/tbl_cohort_claims_by_region.tsv') %>%
   select(region, year, cpkp) %>%
@@
 ```{r}
-read_tsv('tables/tbl_claims_by_race.tsv') %>%
+read_tsv('tables/tbl_cohort_claims_by_race.tsv') %>%
   select(race, year, cpkp) %>%
@@
 
-"Adjusted" means the `year` coefficient in `y ~ year + age*n_cc + is_female + is_dual + is_white + region`.
+"Adjusted" means the `year` coefficient in `y ~ year + age + n_cc + is_female + is_dual + is_white + region`.
 
@@
 ```{r}
-read_tsv('tables/tbl_cohort_model_azithro.tsv') %>% kable
+read_tsv('tables/tbl_cohort_model_azithro.tsv') %>%
+  mutate(cpkp=estimate*1000) %>%
+  filter(term != '(Intercept)') %>%
+  select(term, cpkp, p.value) %>%
+  kable
 ```
@@
 ```{r}
-read_tsv('tables/tbl_cohort_model_levo.tsv') %>% kable
+read_tsv('tables/tbl_cohort_model_levo.tsv') %>%
+  mutate(cpkp=estimate*1000) %>%
+  filter(term != '(Intercept)') %>%
+  select(term, cpkp, p.value) %>%
+  kable
 ```
@@
 Interestingly, the effect of dual-enrollment is much higher for levofloxacin than azithromycin (about triple). Why would being poorer make you more likely to take levofloxacin?
+
+# Supplemental results
+
+## Summary of data for the entire group
+
+```{r}
+read_tsv('tables/tbl_totals.tsv') %>%
+  pctize('n_bene') %>%
+  select(year, n_bene, mean_age, sem_age, mean_n_cc, sem_n_cc,
+         pct_female, pct_white, pct_dual,
+         pct_region_south, pct_region_midwest, pct_region_northeast, pct_region_west) %>%
+  mutate_at(vars(starts_with('pct_')), function(x) round(x, digits=1)) %>%
+  mutate(n_bene=prettyNum(n_bene, big.mark=','),
+         mean_age=round(mean_age, digits=1),
+         mean_n_cc=round(mean_n_cc, digits=2)) %>%
+  gather('var', 'val', -year) %>% spread(year, val) %>% # transpose
+  kable
+```
+
+## Consumption trends in both groups
+
+```{r}
+dat = read_tsv('tables/tbl_claims_by_abx.tsv')
+
+all_dat = bind_rows(
+  top_abx_f(dat) %>% mutate(group='all'),
+  top_abx_f(cohort_dat) %>% mutate(group='cohort')
+)
+
+all_dat %>%
+  ggplot(aes(x=year, y=cpkp, color=group)) +
+  facet_wrap(~ antibiotic, ncol=3, scales='free_y') +
+  geom_point() + geom_line() +
+  theme_minimal() +
+  ggtitle('Unadjusted consumption, overall or in cohort')
+```

### 2017-06-19 ms1/ms1_display.Rmd
ms1 models, display

@@
 import::from(forcats, as_factor)
+import::from(tools, md5sum)
 
@@
 trend_p = function(x) t.test(diff(x))$p.value
+```
+
+# Abstract
+
+```{r}
+model_ci = function(df) {
+  mutate(df, ci_range=1.96*std.error,
+         ci_low=estimate-ci_range,
+         ci_high=estimate+ci_range)
+}
+
+model_ci_cpkp = function(df) {
+  model_ci(df) %>%
+    mutate_at(vars(ci_range, ci_low, ci_high, estimate), function(x) x * 1000) %>%
+    rename(cpkp=estimate)
+}
+
+read_tsv('tables/tbl_cohort_model2.tsv') %>%
+  filter(term=='year') %>%
+  model_ci_cpkp() %>%
+  kable
+
+read_tsv('tables/tbl_cohort_model_abx.tsv') %>%
+  filter(model %in% c('azithromycin', 'levofloxacin'), term=='year') %>%
+  select(term, estimate, std.error, model) %>%
+  model_ci_cpkp() %>%
+  kable
 ```
@@
 
-Only in the cohort:
-
-```{r}
-read_tsv('tables/tbl_cohort_model_abx.tsv') %>%
+```{r adjusted_changes, cache.extra=md5sum('tables/tbl_cohort_model_abx.tsv')}
+x = read_tsv('tables/tbl_cohort_model_abx.tsv') %>%
   filter(term=='year', model %in% top_abx) %>%
@@
          antibiotic=as_factor(model)) %>%
-  select(antibiotic, cpkp_per_year) %>%
+  select(antibiotic, cpkp_per_year)
+
+p = x %>%
   ggplot(aes(x=antibiotic, y=cpkp_per_year)) +
   geom_point() +
+  geom_linerange(aes(x=antibiotic, ymin=0, ymax=cpkp_per_year)) +
   coord_flip() +
   ggtitle('Adjusted changes in consumption') +
+  xlab('') +
+  ylab('adjusted yearly change in claims per 1,000 beneficiaries per year') +
+  scale_y_continuous(limits=c(-16, 17), breaks=c(-15, -10, -5, 0, 5, 10, 15)) +
   theme_minimal()
+
+show(p)
+ggsave('fig/fig_adjusted_changes.pdf', plot=p)
+
+x %>% kable
 ```
@@
 
-```{r}
+```{r, cache.extra=md5sum('tables/tbl_cohort_model_azithro.tsv')}
 read_tsv('tables/tbl_cohort_model_azithro.tsv') %>%
-  mutate(cpkp=estimate*1000) %>%
+  model_ci_cpkp() %>%
   filter(term != '(Intercept)') %>%
-  select(term, cpkp, p.value) %>%
+  select(term, cpkp, ci_low, ci_high, p.value) %>%
+  mutate_at(vars(cpkp, ci_low, ci_high), function(x) round(x, 1)) %>%
   kable
@@
 
-```{r}
+```{r, cache.extra=md5sum('tables/tbl_cohort_model_levo.tsv')}
 read_tsv('tables/tbl_cohort_model_levo.tsv') %>%
-  mutate(cpkp=estimate*1000) %>%
+  model_ci_cpkp() %>%
   filter(term != '(Intercept)') %>%
-  select(term, cpkp, p.value) %>%
+  select(term, cpkp, ci_low, ci_high, p.value) %>%
+  mutate_at(vars(cpkp, ci_low, ci_high), function(x) round(x, 1)) %>%
   kable

### 2017-06-19 ms1/ms1_work.R

@@ -32,13 +32,18 @@ load_data = function(year) {
     mutate(year=year)
   
   # diagnosis claims
-  dx = read_tsv(sprintf('../../data/dx_car_claims_%i.tsv', year)) %>%
-    select(bene_id=BENE_ID, code=diagnosis) %>%
+  dx_car_fn = sprintf('../../data/dx_car_claims_%i.tsv', year)
+  dx_op_fn = sprintf('../../data/dx_op_claims_%i.tsv', year)
+  
+  dx = bind_rows(read_tsv(dx_car_fn),
+                 read_tsv(dx_op_fn)) %>%
+    mutate(date=dmy(from_date)) %>%
+    rename(bene_id=BENE_ID, code=diagnosis) %>%
     left_join(dx_codes, by='code') %>%
-    select(bene_id, diagnosis_type) %>%
-    distinct %>%
-    mutate(dummy=TRUE) %>%
-    spread(diagnosis_type, dummy)
+    filter(diagnosis_type=='acute_rc') %>%
+    select(bene_id, date) %>%
+    group_by(bene_id) %>%
+    summarize(n_acute_rc_dx=length(unique(date)))
 
   # total numbers of PDEs
   n_claims = pde %>% group_by(bene_id) %>% summarize(n_claims=sum(n_claims))
@@ -56,9 +61,7 @@ load_data = function(year) {
     mutate(year=year) %>%
     left_join(cc, by='bene_id') %>%
     left_join(n_claims, by='bene_id') %>% replace_na(list(n_claims=0)) %>%
-    left_join(dx, by='bene_id') %>% replace_na(list(acute_rc=FALSE,
-                                                    copd=FALSE,
-                                                    uti=FALSE))
+    left_join(dx, by='bene_id') %>% replace_na(list(n_acute_rc_dx=0))
 
   list(bene=bene, pde=pde)
 }
@@ -171,13 +174,13 @@ model_f = function(bene, frmla, table_name) {
 # models predicting consumption
 # models with all beneficiaries
 model_f(bene, n_claims ~ year, 'model1')
-model_f(bene, n_claims ~ year + age + n_cc + is_female + is_dual + is_white + region, 'model2')
+model_f(bene, n_claims ~ year + age*n_cc + is_female + is_dual + is_white + region, 'model2')
 
 # models using just the cohort
 lm_cohort_bene = bene %>% filter(in_cohort)
 
 model_f(lm_cohort_bene, n_claims ~ year, 'cohort_model1')
[…]

### 2017-06-20 ms2/ms2.Rmd
cnst rename
cleaning up

@@
 
-# To do
-
-- filter by HRRs, not states
-
 ```{r global_options, include=FALSE}
@@
 import::from(broom, tidy)
+import::from(tools, md5sum)
 ```
@@
 
-```{r load_abg}
+```{r load_abg, cache.extra=md5sum('bug_drug.tsv')}
 resistance_groups = read_tsv('resistance_groups.tsv')
@@
 
-abg = read_tsv('../../../antibiogram/data/abg.tsv', col_types=cols(zipcode='c')) %>%
+raw_abg = read_tsv('../../../antibiogram/data/abg.tsv', col_types=cols(zipcode='c')) %>%
   mutate(drug=tolower(drug), percent_nonsusceptible=100-percent_susceptible) %>%
@@
 # Keep only the abundant bug/drug combinations
-min_n_states = 35
-good_combos = semi_join(abg, possible_bug_drug, by=c('bug', 'drug_group')) %>%
-  select(bug, drug_group, state) %>%
-  distinct %>%
-  count(bug, drug_group) %>% rename(n_places=n) %>%
-  filter(n_places >= min_n_states)
+#min_n_states = 25
+#good_combos = semi_join(raw_abg, possible_bug_drug, by=c('bug', 'drug_group')) %>%
+#  select(bug, drug_group, state) %>%
+#  distinct %>%
+#  count(bug, drug_group) %>% ungroup() %>% 
+#  rename(n_places=n) %>%
+#  filter(n_places >= min_n_states)
 
-abg %<>% right_join(good_combos, by=c('bug', 'drug_group')) %>%
+abg = raw_abg %>%
+  #right_join(good_combos, by=c('bug', 'drug_group')) %>%
+  right_join(possible_bug_drug, by=c('bug', 'drug_group')) %>%
   rename(bug_long=bug) %>% left_join(bug_names, by='bug_long') %>% select(-bug_long) %>%
@@
 
-I started by considering these bug/drug combinations:
+I started by considering these bug/drug combinations on the basis of their clinical relevance and plausibility:
 
@@
 
-Then I filtered to keep only those bug/drug combinations that were present in `r min_n_states` states. That means I'm looking at these `r nrow(good_combos)` bug/drug combinations:
+In previous iterations, I filtered by the number of states in which each bug/drug combination appeared.
+Increasing that threshold ends up testing the relationships for which we have the best data, and therefore we are the best powered for.
+Unfortunately, my simulations mostly make it seem like we should, to get the best power, use all the bug/drug combinations.
 
-`r good_combos %>% arrange(desc(n_places)) %>% kable`
+```{r, eval=FALSE}
+#Then I filtered to keep only those bug/drug combinations that were present in `r min_n_states` states. That means I'm looking at these `r nrow(good_combos)` bug/drug combinations:
+
+#`r good_combos %>% arrange(desc(n_places)) %>% kable`
+```
 
@@
 left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
-  filter(bug %in% c('E. coli', 'Eb. cloacae', 'CNST', 'MSSA', 'MRSA', 'S. pneumoniae')) %>%
+  filter(bug %in% c('E. coli', 'Eb. cloacae', 'CoNS', 'MSSA', 'MRSA', 'S. pneumoniae')) %>%
   ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, color=drug_group, shape=bug)) +
@@
 
+### Comparing state and HRR
+
+HRR-level Spearman correlations between consumption and resistance (almost) always underestimate the correlation reported at the state level:
+
+```{r}
+bind_rows(mutate(results, group='state'),
+          mutate(hrr_results, group='hrr')) %>%
+  filter(model=='spearman') %>%
+  arrange(estimate) %>%
+  mutate(bug_drug=interaction(bug, drug_group)) %>%
+  select(bug_drug, group, estimate) %>%
+  spread(group, estimate) %>%
+  ggplot(aes(x=state, y=hrr, label=bug_drug)) +
+  geom_point() +
+  geom_text() +
+  geom_abline()
+```
+
 ### State-level volcano plot

### 2017-06-20 sex-difference/sex-difference.Rmd
cleaning up

@@
 
-sex_icd = read_tsv('sex_codes.tsv', col_types=cols(code='c')) %>%
-  rename(dx_code=code, dx_desc=short_desc)
+sex_icd = read_tsv('sex_codes.tsv', col_types=cols(code='c'))
 
-dx = read_tsv(sprintf('../../data/dx_car_claims_%i.tsv', year), col_types=cols(diagnosis='c')) %>%
-  select(bene_id=BENE_ID, dx_code=diagnosis) %>%
-  left_join(sex_icd, by='dx_code')
+dx_car = sprintf('../../data/dx_car_claims_%i.tsv', year) %>%
+  read_tsv(col_types=cols(diagnosis='c'))
+dx_op = sprintf('../../data/dx_op_claims_%i.tsv', year) %>%
+  read_tsv(col_types=cols(diagnosis='c'))
+
+dx = bind_rows(dx_car, dx_op) %>%
+  mutate(date=dmy(from_date)) %>%
+  select(bene_id=BENE_ID, date, code=diagnosis) %>%
+  left_join(sex_icd, by='code')
 
 dx_counts = dx %>%
-  count(dx_code)
+  count(code)
 
@@
   distinct() %>%
-  mutate(val=TRUE) %>%
-  spread(diagnosis_type, val)
+  mutate(dummy=TRUE) %>%
+  spread(diagnosis_type, dummy)
+
+rc_dx = dx %>%
+  filter(diagnosis_type=='acute_rc') %>%
+  select(bene_id, date) %>%
+  group_by(bene_id) %>%
+  summarize(n_acute_rc=length(unique(date)))
 
 total_pde = pde %>% count(bene_id) %>% rename(n_claims=n)
+
+cc = read_feather(sprintf('../cc_%i.feather', year))
 
@@
   left_join(total_pde, by='bene_id') %>% replace_na(list(n_claims=0)) %>%
-  left_join(bene_dx, by='bene_id') %>% replace_na(list(acute_rc=FALSE, copd=FALSE, uti=FALSE))
+  left_join(bene_dx, by='bene_id') %>% replace_na(list(acute_rc=FALSE, copd=FALSE, uti=FALSE)) %>%
+  left_join(select(cc, bene_id, COPD), by='bene_id') %>%
+  left_join(rc_dx, by='bene_id') %>% replace_na(list(n_acute_rc=0))
 ```
@@
 sex_icd %>%
-  left_join(dx_counts, by='dx_code') %>%
+  left_join(dx_counts, by='code') %>%
   arrange(diagnosis_type, desc(n)) %>%
@@
 
-### Sex by age
-
-The older people are more female: age 65 is about 60% female.
-This stays constant around 60% (or dips slightly) until age 75, when it steadily increases up to 85-90% female at age 100.
-So we'll need to account for this age difference.
-
-```{r sex_by_age, eval=FALSE}
-plot_dat = usage %>%
-  select(age, sex) %>%
-  filter(sex %in% c('male', 'female')) %>%
-  group_by(age) %>% count(sex) %>% ungroup() %>%
-  spread(sex, n, fill=0L) %>%
-  mutate(total=female+male, grand_total=sum(total),
-         fraction_year=total/grand_total, fraction_female=female/total)
-
-ggplot(mapping=aes(x=age, y=fraction)) +
-  geom_point(data=rename(plot_dat, fraction=fraction_female)) +
-  geom_line(data=mutate(plot_dat, fraction=fraction_year*10), col='red') +
-  ylab('fraction female') + ylim(0, 1) +
-  theme_minimal()
-```
-
 ### Chronic conditions
@@
 ```
+
+### Sex by age
+
+The older people are more female: age 65 is about 60% female.
+This stays constant around 60% (or dips slightly) until age 75, when it steadily increases up to 85-90% female at age 100.
+So we'll need to account for this age difference.
+
+```{r sex_by_age, eval=FALSE}
+plot_dat = usage %>%
+  select(age, sex) %>%
+  filter(sex %in% c('male', 'female')) %>%
+  group_by(age) %>% count(sex) %>% ungroup() %>%
+  spread(sex, n, fill=0L) %>%
+  mutate(total=female+male, grand_total=sum(total),
+         fraction_year=total/grand_total, fraction_female=female/total)
+
+ggplot(mapping=aes(x=age, y=fraction)) +
+  geom_point(data=rename(plot_dat, fraction=fraction_female)) +
+  geom_line(data=mutate(plot_dat, fraction=fraction_year*10), col='red') +
+  ylab('fraction female') + ylim(0, 1) +
+  theme_minimal()
+```

### 2017-06-20 ms2/ms2.Rmd
merging Staph, filtering n

@@
 bug_names = read_tsv('bug_names.tsv')
+bug_groups = read_tsv('bug_groups.tsv')
+possible_bug_drug = read_tsv('bug_drug.tsv')
 
@@
   right_join(resistance_groups, by='drug') %>%
-  select(bug, drug_group, state, hrr, percent_nonsusceptible, n_isolates)
-
-possible_bug_drug = read_tsv('bug_drug.tsv')
+  select(bug, drug, drug_group, state, hrr, percent_nonsusceptible, n_isolates)
 
 # Keep only the abundant bug/drug combinations
-#min_n_states = 25
-#good_combos = semi_join(raw_abg, possible_bug_drug, by=c('bug', 'drug_group')) %>%
-#  select(bug, drug_group, state) %>%
-#  distinct %>%
-#  count(bug, drug_group) %>% ungroup() %>% 
-#  rename(n_places=n) %>%
-#  filter(n_places >= min_n_states)
+min_n_states = 35
+good_combos = semi_join(raw_abg, possible_bug_drug, by=c('bug', 'drug_group')) %>%
+  select(bug, drug_group, state) %>%
+  distinct %>%
+  count(bug, drug_group) %>% ungroup() %>% 
+  rename(n_places=n) %>%
+  filter(n_places >= min_n_states)
 
 abg = raw_abg %>%
-  #right_join(good_combos, by=c('bug', 'drug_group')) %>%
-  right_join(possible_bug_drug, by=c('bug', 'drug_group')) %>%
-  rename(bug_long=bug) %>% left_join(bug_names, by='bug_long') %>% select(-bug_long) %>%
-  filter(!is.na(percent_nonsusceptible), !is.na(n_isolates))
+  #mutate(raw_bug=bug) %>%
+  # shorten the bug names
+  left_join(bug_names, by='bug') %>%
+  select(-bug) %>% rename(bug=bug_short) %>%
+  # group the different S. aureus into meta-groups depending on the drug
+  left_join(bug_groups, by=c('bug', 'drug_group')) %>%
+  mutate(bug_group=if_else(is.na(bug_group), bug, bug_group)) %>%
+  select(-bug) %>% rename(bug=bug_group) %>%
+  right_join(good_combos, by=c('bug', 'drug_group')) %>%
+  # keep only particular bug/drug combos
+  right_join(possible_bug_drug, by=c('bug', 'drug_group')) #%>%
+  #filter(!is.na(percent_nonsusceptible), !is.na(n_isolates))
+
+# get the median number of isolates for each bug/drug
+median_n_isolates = abg %>%
+  filter(!is.na(n_isolates)) %>%
+  group_by(bug, drug_group) %>%
+  summarize(median_n_isolates=as.integer(median(n_isolates))) %>%
+  ungroup()
+
+# replace the NA number of isolates with the median for that bug/drug
+abg %<>%
+  left_join(median_n_isolates) %>%
+  mutate(n_isolates=if_else(is.na(n_isolates), median_n_isolates, n_isolates)) %>%
+  select(-median_n_isolates)
 
 # check that all the short names made it in
-if (any(is.na(abg$bug))) stop('missing bug names')
+for (x in c('bug', 'n_isolates', 'percent_nonsusceptible')) {
+  if (any(is.na(abg[[x]]))) stop(sprintf('NAs in field "%s"', x))
+}
 
@@
               n_place_antibiograms=n()) %>%
+    ungroup() %>%
     mutate(place_weight=sqrt(n_place_antibiograms))
@@
 
-I started by considering these bug/drug combinations on the basis of their clinical relevance and plausibility:
+I started by considering these `r nrow(possible_bug_drug)` bug/drug combinations on the basis of their clinical relevance and plausibility:
 
@@
   ylab("Spearman's rho")
+```
+
+```{r}
+x = abg %>% 
+  select(bug, drug_group, state) %>%
+  distinct() %>%
+  count(bug, drug_group) %>% ungroup() %>% rename(n_places=n)
+results %>% filter(model=='spearman') %>%
+  left_join(x, by=c('bug', 'drug_group')) %>%
+  ggplot(aes(x=n_places, y=estimate)) +
+  geom_point()
 ```

### 2017-06-20 ms2/ms2.Rmd
updating filtering

@@
   right_join(resistance_groups, by='drug') %>%
-  select(bug, drug, drug_group, state, hrr, percent_nonsusceptible, n_isolates)
+  select(bug, raw_drug=drug, drug_group, state, hrr, percent_nonsusceptible, n_isolates)
 
 # Keep only the abundant bug/drug combinations
-min_n_states = 35
-good_combos = semi_join(raw_abg, possible_bug_drug, by=c('bug', 'drug_group')) %>%
-  select(bug, drug_group, state) %>%
-  distinct %>%
-  count(bug, drug_group) %>% ungroup() %>% 
-  rename(n_places=n) %>%
-  filter(n_places >= min_n_states)
+# min_n_states = 35
+# good_combos = semi_join(raw_abg, possible_bug_drug, by=c('bug', 'drug_group')) %>%
+#   select(bug, drug_group, state) %>%
+#   distinct %>%
+#   count(bug, drug_group) %>% ungroup() %>% 
+#   rename(n_places=n) %>%
+#   filter(n_places >= min_n_states)
 
@@
   select(-bug) %>% rename(bug=bug_group) %>%
-  right_join(good_combos, by=c('bug', 'drug_group')) %>%
+  #right_join(good_combos, by=c('bug', 'drug_group')) %>%
   # keep only particular bug/drug combos
-  right_join(possible_bug_drug, by=c('bug', 'drug_group')) #%>%
+  right_join(possible_bug_drug, by=c('bug', 'drug_group'))
   #filter(!is.na(percent_nonsusceptible), !is.na(n_isolates))
@@
 
+# get the number of states in which each bug/drug appears
+min_n_states = 37
+abg_n_states = abg %>%
+  select(bug, drug_group, state) %>%
+  distinct() %>%
+  count(bug, drug_group) %>%
+  ungroup() %>%
+  rename(n_states=n)
+abg_n_hrr = abg %>%
+  select(bug, drug_group, hrr) %>%
+  distinct() %>%
+  count(bug, drug_group) %>%
+  ungroup() %>%
+  rename(n_hrr=n)
+
 # replace the NA number of isolates with the median for that bug/drug
@@
   mutate(n_isolates=if_else(is.na(n_isolates), median_n_isolates, n_isolates)) %>%
-  select(-median_n_isolates)
+  select(-median_n_isolates) %>%
+  left_join(abg_n_states, by=c('bug', 'drug_group')) %>%
+  left_join(abg_n_hrr, by=c('bug', 'drug_group'))
+  #filter(n_places >= min_n_states)
 

### 2017-06-21 ms2/ms2.Rmd
bug groups

@@
 state_ineq = read_years('data/state_ineq_%i.tsv') %>%
-  rename(cpb=mean) %>%
+  rename(cpb=mean, nzg=nzgini, nzr=mean_fnz_residual) %>%
   group_by(state, drug_group) %>%
-  summarize(cpb=mean(cpb), fnz=mean(fnz), nzg=mean(nzgini)) %>%
+  summarize_at(vars(-state, -drug_group, -total), mean) %>%
   ungroup()
@@
 hrr_ineq = read_years('data/hrr_ineq_%i.tsv') %>%
-  rename(cpb=mean) %>%
+  rename(cpb=mean, nzg=nzgini, nzr=mean_fnz_residual) %>%
   group_by(hrr, drug_group) %>%
-  summarize(cpb=mean(cpb), fnz=mean(fnz), nzg=mean(nzgini)) %>%
+  summarize_at(vars(-hrr, -drug_group, -total), mean) %>%
   ungroup()
@@
 uni_term = 'cpb' # univariable explanatory term
-multi_term = 'nzg' # multivariable explanatory term (along with fnz)
+multi_term_adjuster = 'cpb'
+multi_term = 'nb_prob' # multivariable explanatory term (along with fnz)
 models = function(x) {
@@
     linear_model(x, dep, uni_term) %>% mutate(model='uni'),
-    linear_model(x, dep, c('fnz', multi_term)) %>% mutate(model='multi'),
+    linear_model(x, dep, c(multi_term_adjuster, multi_term)) %>% mutate(model='multi'),
     spearman_model(x, dep, uni_term) %>% mutate(model='spearman')

### 2017-06-21 ms2/ms2.Rmd
moving to HRR

@@
 
-```{r}
+HRR spearman:
+
+```{r hrr_spearman_interval}
+hrr_results %>%
+  filter(model=='spearman') %>%
+  interval_plot() +
+  ylab("Spearman's rho")
+```
+
+```{r, eval=FALSE}
+# Used this to show that there's a correlation between
+# strength of correlation and amount of data
 x = abg %>% 
@@
 
-## MSSA and clindamycin
+## S. aureus and clindamycin
 
-```{r mssa_clinda}
+```{r staph_clinda}
 left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
   left_join(regions, by='state') %>%
-  filter(bug=='MSSA', drug_group=='clindamycin') %>%
+  filter(bug=='S. aureus', drug_group=='clindamycin') %>%
   ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, label=state)) +
@@
   geom_text() +
-  xlab('clindamycin claims per beneficiary') + ylab('mean percent nonsusceptible MSSA isolates') +
+  xlab('clindamycin claims per beneficiary') + ylab('mean percent nonsusceptible S. aureus isolates') +
   theme_minimal()

### 2017-06-22 sex-difference/namcs.Rmd
namcs sex diff

---
title: "Drivers of sex differences in consumption (NAMCS)"
author: "Scott Olesen"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/', dev=c('png', 'pdf'),
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE, autodep=TRUE)

import::from(survey, svyglm)
import::from(survey, svydesign)
```

```{r load_data}
sex_codes = read_tsv('namcs_sex_codes.tsv') %>%
  rename(dx=code)

dx = read_tsv('../../db/namcs/data/namcs_2013_dx.tsv') %>%
  left_join(sex_codes, by='dx')

dx_visit = dx %>%
  group_by(record_id) %>%
  summarize(uti='uti' %in% diagnosis_type,
            skin='skin' %in% diagnosis_type,
            vaginitis='vaginitis' %in% diagnosis_type)

abx_drug_codes = read_tsv('../../db/namcs/drug-codes/abx-codes.tsv') %>%
  select(drug_code, antibiotic) %>%
  mutate(antibiotic=tolower(antibiotic))

drugs = read_tsv('../../db/namcs/data/namcs_2013_drugs.tsv')

abx = drugs %>% left_join(abx_drug_codes, by='drug_code') %>% filter(!is.na(antibiotic))

abx_visits = abx %>%
  select(antibiotic, record_id) %>%
  distinct() %>%
  mutate(value=TRUE) %>% spread(antibiotic, value)

visits = read_tsv('../../db/namcs/data/namcs_2013_visits.tsv') %>%
  left_join(dx_visit, by='record_id') %>%
  replace_na(list(uti=FALSE, skin=FALSE, vaginitis=FALSE)) %>%
  left_join(abx_visits, by='record_id') %>%
  mutate_at(vars(amoxclav:tmp_smx), function(x) !is.na(x))

abx_names = visits %>% select(amoxclav:tmp_smx) %>% names
stopifnot(length(abx_names) == length(unique(abx_drug_codes$antibiotic)))
```

Most antibiotics are prescribed in the context of a new (less than 3 month onset) problem, but the people who get antibiotics in the context of a chronic problem tend to be older:

```{r}
abx %>%
  select(antibiotic, record_id) %>% distinct %>%
  left_join(visits, by='record_id') %>%
  group_by(major_reason) %>%
  summarize(n=n(), mean_age=mean(age), sd_age=sd(age)) %>%
  arrange(desc(n)) %>%
  kable
```

# Overall

```{r}
visits %>%
  right_join(abx, by='record_id') %>%
  group_by(sex, antibiotic) %>%
  summarize(n_rx=sum(weight)) %>%
  ungroup() %>%
  spread(sex, n_rx) %>%
  ggplot(aes(x=male, y=female, label=antibiotic)) +
  geom_point() +
  geom_text() +
  coord_fixed() +
  geom_abline()
```

```{r models}
des = visits %>%
  mutate(sex=factor(sex, levels=c('male', 'female'))) %>%
  svydesign(data=., id=~0, weights=~weight)

f = function(drug) {
  svyglm(as.formula(sprintf('%s ~ sex', drug)), des, family=binomial(link='log'), start=c(-5, 0)) %>%
    tidy %>%
    filter(term=='sexfemale') %>%
    mutate(drug=drug)
}

sex_diff_by_drug = lapply(abx_names, f) %>%
  bind_rows() %>%
  mutate(ci=1.96*std.error,
         ci_low=estimate-ci,
         ci_hi=estimate+ci,
         p.adj=p.adjust(p.value, method='BH')) %>%
  mutate(sig=case_when(.$p.adj < 0.05 ~ '**',
                       .$p.value < 0.05 ~ '*',
                       TRUE ~ '')) %>%
  select(-ci) %>%
  mutate_at(vars(estimate, ci_low, ci_hi), function(x) exp(x))
```

```{r diff_plot}
sex_diff_by_drug %>%
  mutate(drug=forcats::fct_reorder(factor(drug), estimate)) %>%
  ggplot(aes(x=drug, y=estimate, ymin=ci_low, ymax=ci_hi)) +
  geom_point() + geom_linerange() +
  geom_text(aes(y=1, label=sig), cex=6) +
  coord_flip() +
  scale_y_continuous(breaks=c(0, 1, 2, 5, 10), limits=c(0, 10)) +
  ylab('relative risk of being female for getting drug')
```

```{r diff_table}
sex_diff_by_drug %>%
  select(drug, ci_low, estimate, ci_hi, p.value, p.adj, sig) %>%
  mutate_at(vars(ci_low:ci_hi), function(x) round(x, digits=2)) %>%
  mutate_at(vars(p.value, p.adj), function(x) round(x, digits=3)) %>%
  arrange(desc(estimate)) %>%
  kable
```

```{r model_helpers}
m = function(frmla) svyglm(frmla, design=des, family=binomial(link='log'))

# explained excess risk ratio
eerr = function(m1, m2, coeff='sexfemale') {
  (1 - (exp(m2$coefficients[coeff]) - 1)/(exp(m1$coefficients[coeff]) - 1)) %>% unname
}
report_eerr = function(m1, m2) round(100 * eerr(m1, m2), 1)
```

# Nitrofurantoin

```{r nitro_models}
nitro1 = m(nitrofurantoin ~ sex)
nitro2 = m(nitrofurantoin ~ sex + uti)
```

This should be easy to explain by UTIs. They explain `r report_eerr(nitro1, nitro2)`% of the relative risk.

# Metronidazole

```{r metro_models}
metro1 = m(metronidazole ~ sex)
metro2 = m(metronidazole ~ sex + vaginitis)
```

Metronidazole consumption is increased in women. This is probably vaginitis (`61610`). So vaginitis explains `r report_eerr(metro1, metro2)`% of the relative risk. Not stupendous.

# Clindamycin

Clinda is only marginally significant in the models.

```{r clinda_models}
clinda1 = m(clindamycin ~ sex)
clinda2 = m(clindamycin ~ sex + skin)
```

Skin conditions explain `r report_eerr(clinda1, clinda2)`% of the relative risk. Also not stupendous.

### 2017-06-22 sex-difference/sex-difference.Rmd
sex diff

@@
 ```
+
+## Margins of azithromycin consumption
+
+Maybe women take more azithromycin because usage is more concentrated in women than in men.
+Then we could explain the increased consumption by a small number of people taking more drug.
+(Yonatan points to Lady Windmere syndrome.) In fact, consumption is more even among women.
+As always, this is driven by the extensive fraction: more women took at least one dose:
+
+```{r}
+pde %>%
+  filter(drug=='azithromycin') %>%
+  count(bene_id) %>% rename(n=n_azi) %>%
+  right_join(bene, by='bene_id') %>% replace_na(list(n_azi=0)) %>%
+  count(sex, n_azi) %>%
+  group_by(sex) %>%
+  mutate(f_ppl=n/sum(n), cum_f_ppl=cumsum(f_ppl),
+         cum_cons=cumsum(n_azi*n), cum_f_cons=cum_cons/max(cum_cons)) %>%
+  ggplot(aes(x=cum_f_ppl, y=cum_f_cons, color=sex)) +
+  geom_point() +
+  geom_line() +
+  xlab('Fraction of consumers') + ylab("Fraction of that sex's azithromycin consumed")
+```

### 2017-06-22 ms2/ms2.Rmd
ms2 new mtric, cleanup

@@
 
-# Keep only the abundant bug/drug combinations
-# min_n_states = 35
-# good_combos = semi_join(raw_abg, possible_bug_drug, by=c('bug', 'drug_group')) %>%
-#   select(bug, drug_group, state) %>%
-#   distinct %>%
-#   count(bug, drug_group) %>% ungroup() %>% 
-#   rename(n_places=n) %>%
-#   filter(n_places >= min_n_states)
-
 abg = raw_abg %>%
-  #mutate(raw_bug=bug) %>%
   # shorten the bug names
@@
   select(-bug) %>% rename(bug=bug_group) %>%
-  #right_join(good_combos, by=c('bug', 'drug_group')) %>%
   # keep only particular bug/drug combos
   right_join(possible_bug_drug, by=c('bug', 'drug_group'))
-  #filter(!is.na(percent_nonsusceptible), !is.na(n_isolates))
 
@@
 
-# get the number of states in which each bug/drug appears
-min_n_states = 37
-abg_n_states = abg %>%
-  select(bug, drug_group, state) %>%
-  distinct() %>%
-  count(bug, drug_group) %>%
-  ungroup() %>%
-  rename(n_states=n)
-abg_n_hrr = abg %>%
-  select(bug, drug_group, hrr) %>%
-  distinct() %>%
-  count(bug, drug_group) %>%
-  ungroup() %>%
-  rename(n_hrr=n)
-
 # replace the NA number of isolates with the median for that bug/drug
@@
   mutate(n_isolates=if_else(is.na(n_isolates), median_n_isolates, n_isolates)) %>%
-  select(-median_n_isolates) %>%
-  left_join(abg_n_states, by=c('bug', 'drug_group')) %>%
-  left_join(abg_n_hrr, by=c('bug', 'drug_group'))
-  #filter(n_places >= min_n_states)
+  select(-median_n_isolates)
 
@@
 uni_term = 'cpb' # univariable explanatory term
-multi_term_adjuster = 'cpb'
-multi_term = 'nb_prob' # multivariable explanatory term (along with fnz)
+multi_term = 'nzg' #'nb_prob' # multivariable explanatory term
+multi_term_adjuster = 'cpb' # other multivarible explanatory term
+
 models = function(x) {
@@
 
-I join the resistance drugs into classes:
+I join the resistance drugs into the classes shown below. (I also joined the consumption drugs into similar classes.)
 
@@
 
-I similarly joined the consumption drugs into similar classes.
+The *S.
+aureus* antibiograms are listed as MSSA, MRSA, or plain old *S.
+aureus*.
+For tetracycline, Bactrim, and clindamycin, I didn't find any big difference in resistance patterns about the three groups, so I merged them all into one group for purposes of the analysis.
+For the cephalosporins, I joined MSSA and plain old *S.
+aureus*.
+(MRSA, on the other hand, had clear indications that the data were artificially set to 100% resistance.
+Intriguingly, it seemed like this was also the case for MSSA and the beta-lactams [penicillin and ampicillin], so I did not include those.)
 
@@
 
-In previous iterations, I filtered by the number of states in which each bug/drug combination appeared.
-Increasing that threshold ends up testing the relationships for which we have the best data, and therefore we are the best powered for.
-Unfortunately, my simulations mostly make it seem like we should, to get the best power, use all the bug/drug combinations.
-
-```{r, eval=FALSE}
-#Then I filtered to keep only those bug/drug combinations that were present in `r min_n_states` states. That means I'm looking at these `r nrow(good_combos)` bug/drug combinations:
-
-#`r good_combos %>% arrange(desc(n_places)) %>% kable`
-```
+(In previous iterations, I filtered by the number of states/HRRs in which each bug/drug combination appeared.
+Weirdly, there is a relationship in the data, that better-represented bug/drug combinations are the ones with the stronger relationships [i.e., the higher $\rho$, not necessarily the lower $p$].
+When using the HRR data, that kind of fancy footwork wasn't necessary, so I just didn't do it.)
 
@@
 
-Significant results using Spearman correlations of mean percent nonsusceptible ~ claims per beneficiary.
+For each bug/drug combination, I computed `mean percent nonsusceptible ~ claims per beneficiary`.
 A star after the bug/drug combination implies that the adjusted $p < 0.05$, and a dot means the unadjusted $p < 0.05$.
+Intervals show 95% confidence intervals.
 
-```{r spearman_models}
-results %>%
-  filter(model=='spearman') %>%
-  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
-  filter(p.value < 0.05) %>%
-  arrange(p.adj) %>%
-  select(bug, drug_group, estimate, p.value, p.adj, n) %>%
-  kable
-```
-
-Aside from individual significant results, we can look at the landscape of correlations.
-They are almost all positive, with MRSA/Bactrim making an interesting exception.
-Lines show 95% confidence intervals.
-
-```{r spearman_interval}
+```{r hrr_spearman_interval}
 interval_plot = function(df) {
@@
 
-results %>%
-  filter(model=='spearman') %>%
-  interval_plot() +
-  ylab("Spearman's rho")
-```
-
-HRR spearman:
-
-```{r hrr_spearman_interval}
 hrr_results %>%
@@
 
-```{r, eval=FALSE}
-# Used this to show that there's a correlation between
-# strength of correlation and amount of data
-x = abg %>% 
-  select(bug, drug_group, state) %>%
-  distinct() %>%
-  count(bug, drug_group) %>% ungroup() %>% rename(n_places=n)
-results %>% filter(model=='spearman') %>%
-  left_join(x, by=c('bug', 'drug_group')) %>%
-  ggplot(aes(x=n_places, y=estimate)) +
-  geom_point()
+All the demonstrably-different-from-zero relationships are positive.
+This jives with our general understanding that more consumption means more resistance.
+Note also the bunching by drug type, rather than organism.
+
+This table has the significant and marginally significant results from the previous figure.
+
+```{r hrr_spearman_table}
+hrr_results %>%
+  filter(model=='spearman') %>%
+  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
+  filter(p.value < 0.05) %>%
+  select(bug, drug_group, p.adj, p.value, rho=estimate, ci_low, ci_high) %>%
+  arrange(p.value) %>%
+  kable
 ```
 
-## Differences in consumption/resistance patterns explain some of these results
-
-Different bug/drug combinations have different spreads along the consumption and resistance axes, so we would expect to see different strengths of association for each one.
-
-```{r}
-left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
-  filter(bug %in% c('E. coli', 'Eb. cloacae', 'CoNS', 'MSSA', 'MRSA', 'S. pneumoniae')) %>%
-  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, color=drug_group, shape=bug)) +
-  geom_point() +
-  xlab('claims per beneficiary')
-```
+Univariable linear regressions are in the Supplement. A crazy visualization of all the bug/drug combinations and resistances is also in the Supplement.
 
@@
 
+## Amount of data vs. strength of correlation
+
+There is a correlation between the amount of data (i.e., the number of HRRs in which there is an antibiogram for a bug/drug combination) and the correlation coefficient for that relationship.
+The bug/drug combinations with more data tend to be the ones with more positive $\rho$ values.
+This makes it advantageous to the study's power to filter the poorly-represented combinations.
+
+```{r}
+abg %>% 
+  select(bug, drug_group, hrr) %>%
+  distinct() %>%
+  count(bug, drug_group) %>% ungroup() %>% rename(n_places=n) %>%
+  right_join(filter(hrr_results, model=='spearman')) %>%
+  ggplot(aes(x=n_places, y=estimate)) +
+  geom_point()
+```
+
 ## Metrics of unevenness
@@
 
-## Spearman models
+## State-level Spearman models
 
-### Comparing state and HRR
+```{r state_spearman_table}
+results %>%
+  filter(model=='spearman') %>%
+  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
+  filter(p.value < 0.05) %>%
+  arrange(p.adj) %>%
+  select(bug, drug_group, estimate, p.value, p.adj, n) %>%
+  kable
+```
+
+```{r state_spearman_interval}
+results %>%
+  filter(model=='spearman') %>%
+  interval_plot() +
+  ylab("Spearman's rho")
+```
+
+## Comparing Spearman models state and HRR
 
@@
 
-### State-level volcano plot
-
-```{r}
-results %>%
-  filter(model=='spearman') %>%
-  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
-  ggplot(aes(x=estimate, y=-log10(p.adj))) + geom_point() +
-  geom_hline(yintercept=-log10(0.05), col='red')
-```
-
-### HRR-level correlations
-
-```{r hrr_spearman_models}
-hrr_results %>%
-  filter(model=='spearman') %>%
-  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
-  filter(p.value < 0.05) %>%
-  arrange(p.value) %>%
-  select(bug, drug_group, estimate, p.value, p.adj) %>%
-  kable
-```
-
 ## Univariable regressions
 
-Like the Spearman models, the model here is mean percent nonsusceptible ~ claims per beneficiary, only this time it's a normal linear regression.
+Like the Spearman models, the model here is `mean percent nonsusceptible ~ claims per beneficiary`, only this time it's a normal linear regression. At the HRR level:
 
-### State-level
-
-```{r uni_models}
-results %>%
-  filter(model=='uni', term==uni_term) %>%
-  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
-  #filter(p.value < 0.05) %>%
-  arrange(p.value) %>%
-  select(bug, drug_group, estimate, p.value, p.adj, n) %>%
-  kable
-```
-
-```{r}
-results %>%
-  filter(model=='uni', term==uni_term) %>%
-  interval_plot()
-```
-
-```{r}
-results %>%
-  filter(model=='uni') %>%
-  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
-  ggplot(aes(x=estimate, y=-log10(p.adj))) + geom_point() +
-  geom_hline(yintercept=-log10(0.05), col='red')
-```
-
-### HRR-level
-
-```{r hrr_uni_models}
+```{r hrr_uni_table}
 hrr_results %>%
@@
 ```
+
+```{r hrr_uni_interval}
+hrr_results %>%
+  filter(model=='uni', term==uni_term) %>%
+  interval_plot()
+```
+
+And at the state level:
+
+```{r state_uni_interval}
+results %>%
+  filter(model=='uni', term==uni_term) %>%
+  interval_plot()
+```
+
+```{r state_uni_table}
+results %>%
+  filter(model=='uni', term==uni_term) %>%
+  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
+  filter(p.value < 0.05) %>%
+  arrange(p.value) %>%
+  select(bug, drug_group, estimate, p.value, p.adj, n) %>%
+  kable
+```
+
+## Consumption/resistance patterns
+
+Different bug/drug combinations have different spreads along the consumption and resistance axes, so we would expect to see different strengths of association for each one.
+
+```{r consumption_resistance_patterns}
+left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
+  filter(bug %in% c('E. coli', 'Eb. cloacae', 'CoNS', 'MSSA', 'MRSA', 'S. pneumoniae')) %>%
+  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, color=drug_group, shape=bug)) +
+  geom_point() +
+  xlab('claims per beneficiary')
+```

### 2017-06-22 ms2/data/collect-data.R

@@ -1,13 +1,13 @@
 #!/usr/bin/env Rscript
 
 # inequality function
-import::from(ineq, Gini)
+import::from(ineq, ineq)
 nonzero = function(x) x[x > 0]
 fraction_nonzero = function(x) length(nonzero(x)) / length(x)
 
 pad0 = function(x, to) c(x, rep(0, to - length(x)))
 
-lm_residuals = function(y, x) residuals(lm(y ~ x))
+hoover = function(x) 0.5 * sum(abs(x - mean(x))) / sum(x)
 
 inequalities = function(x) {
   nb_par = fitdistrplus::mledist(x, 'nbinom')$estimate
@@ -16,9 +16,14 @@ inequalities = function(x) {
   data_frame(total=sum(x),
              mean=mean(x),
              fnz=fraction_nonzero(x),
-             nzgini=Gini(nonzero(x)),
+             gini=ineq(x, type='Gini'),
+             nzgini=ineq(nonzero(x), type='Gini'),
              nb_size=nb_size,
-             nb_prob=nb_prob
+             nb_prob=nb_prob,
+             pietra=ineq(x, type='RS'),
+             nzpieta=ineq(nonzero(x), type='RS'),
+             hoover=hoover(x),
+             nzhoover=hoover(nonzero(x))
              )
 }
 
@@ -60,8 +65,7 @@ summarize_inequality = function(year) {
     replace_na(list(n_claims=0)) %>%
     group_by(state) %>%
     do(inequalities(.$n_claims)) %>%
-    ungroup() %>%
-    mutate(mean_fnz_residual=lm_residuals(mean, fnz))
+    ungroup()
 
   write_tsv(total_pde, state_all_fn)
 
@@ -72,8 +76,7 @@ summarize_inequality = function(year) {
     left_join(state_denom, by='state') %>%
     group_by(state, drug_group) %>%
     do(inequalities(pad0(.$n_claims, unique(.$n_bene)))) %>%
-    ungroup() %>%
-    mutate(mean_fnz_residual=lm_residuals(mean, fnz))
+    ungroup()
 
   write_tsv(state_inequality, state_fn)
 
@@ -85,8 +88,7 @@ summarize_inequality = function(year) {
     replace_na(list(n_claims=0)) %>%
     group_by(hrr) %>%
     do(inequalities(.$n_claims)) %>%
-    ungroup() %>%
[…]

### 2017-06-26 sex-difference/namcs.Rmd
updated namcs

@@
 
-dx = read_tsv('../../db/namcs/data/namcs_2013_dx.tsv') %>%
+dx = read_tsv('../../db/namcs/data/namcs_2014_dx.tsv') %>%
   left_join(sex_codes, by='dx')
@@
 dx_visit = dx %>%
-  group_by(record_id) %>%
+  group_by(visit_id) %>%
   summarize(uti='uti' %in% diagnosis_type,
@@
 
-abx_drug_codes = read_tsv('../../db/namcs/drug-codes/abx-codes.tsv') %>%
-  select(drug_code, antibiotic) %>%
-  mutate(antibiotic=tolower(antibiotic))
+top_abx = read_tsv('../../db/namcs/analysis/top_abx.tsv')
 
-drugs = read_tsv('../../db/namcs/data/namcs_2013_drugs.tsv')
+# abx_drug_codes = read_tsv('../../db/namcs/drug-codes/abx-codes.tsv') %>%
+#   select(drug_code, antibiotic) %>%
+#   mutate(antibiotic=tolower(antibiotic))
 
-abx = drugs %>% left_join(abx_drug_codes, by='drug_code') %>% filter(!is.na(antibiotic))
+drug = read_tsv('../../db/namcs/data/namcs_2014_drugs.tsv')
 
-abx_visits = abx %>%
-  select(antibiotic, record_id) %>%
+abx = top_abx %>% left_join(drug, by='desc')
+
+abx_visit = abx %>%
+  select(visit_id, abx) %>%
   distinct() %>%
-  mutate(value=TRUE) %>% spread(antibiotic, value)
+  mutate(value=TRUE) %>% spread(abx, value) %>%
+  mutate(any_abx=TRUE)
 
-visits = read_tsv('../../db/namcs/data/namcs_2013_visits.tsv') %>%
-  left_join(dx_visit, by='record_id') %>%
+visit = read_tsv('../../db/namcs/data/namcs_2014_visits.tsv') %>%
+  left_join(dx_visit, by='visit_id') %>%
   replace_na(list(uti=FALSE, skin=FALSE, vaginitis=FALSE)) %>%
-  left_join(abx_visits, by='record_id') %>%
-  mutate_at(vars(amoxclav:tmp_smx), function(x) !is.na(x))
-
-abx_names = visits %>% select(amoxclav:tmp_smx) %>% names
-stopifnot(length(abx_names) == length(unique(abx_drug_codes$antibiotic)))
+  left_join(abx_visit, by='visit_id') %>%
+  mutate_at(c('any_abx', top_abx$abx), function(x) !is.na(x))
 ```
 
-Most antibiotics are prescribed in the context of a new (less than 3 month onset) problem, but the people who get antibiotics in the context of a chronic problem tend to be older:
+# Methods
 
-```{r}
-abx %>%
-  select(antibiotic, record_id) %>% distinct %>%
-  left_join(visits, by='record_id') %>%
-  group_by(major_reason) %>%
-  summarize(n=n(), mean_age=mean(age), sd_age=sd(age)) %>%
-  arrange(desc(n)) %>%
+## The most-consumed antibiotics
+
+I wanted to focus my analysis on the top few most important antibiotics.
+The NAMCS data includes information about the therapeutic category of each drug.
+I looked through those categories and identified the ones that could refer to antibiotics.
+
+```{r abx_cat_table}
+abx_cat = read_tsv('../../db/namcs/data/parse/drug_codes/abx_codes.tsv')
+abx_cat %>% kable
+```
+
+Then I looked through all the drugs that fell into one of these categories, summed up their usage, and then looked at the names of the top 20 drugs:
+
+```{r find_top_abx}
+drug %>%
+  semi_join(abx_cat, by=c('cat1'='category')) %>%
+  select(visit_id, desc) %>%
+  left_join(select(visit, visit_id, survey_id, stratum, weight), by='visit_id') %>%
+  mutate(dummy=1) %>%
+  svydesign(data=., ids=~survey_id, strata=~stratum, weights=~weight) %>%
+  svyby(~dummy, ~desc, design=., svytotal) %>%
+  arrange(desc(dummy)) %>%
+  select(abx=desc, n_rx=dummy, se) %>%
+  mutate(f=100*round(n_rx/sum(n_rx), 2),
+         cum_f_rx=100*round(cumsum(n_rx/sum(n_rx)), 2),
+         relative_se=100*round(se/n_rx, 2)) %>%
+  select(drug=abx, `fraction (%)`=f, `relative SE (%)`=relative_se, `cumulative fraction (%)`=cum_f_rx) %>%
+  head(15) %>%
   kable
@@
 
-# Overall
+I kept the drugs from azithromycin through clindamycin.
+I skipped moxifloxacin because it has a high relative SE and mostly won't make it through future analyses.
+I keep nitrofurantoin because it's interesting from the Medicare data.
 
-```{r}
-visits %>%
-  right_join(abx, by='record_id') %>%
-  group_by(sex, antibiotic) %>%
-  summarize(n_rx=sum(weight)) %>%
-  ungroup() %>%
-  spread(sex, n_rx) %>%
-  ggplot(aes(x=male, y=female, label=antibiotic)) +
-  geom_point() +
-  geom_text() +
-  coord_fixed() +
-  geom_abline()
+# Results
+
+## Overall consumption by sex
+
+Overall, more of the antibiotic claims were made by women. Women consume roughly 21% more antibiotics than men.
+
+```{r sex_total_odds}
+drug %>%
+  semi_join(abx_cat, by=c('cat1'='category')) %>%
+  select(visit_id, desc) %>%
+  left_join(select(visit, visit_id, sex, survey_id, stratum, weight), by='visit_id') %>%
+  mutate(dummy=1) %>%
+  svydesign(data=., ids=~survey_id, strata=~stratum, weights=~weight) %>%
+  svyby(~dummy, ~sex, design=., svytotal) %>%
+  mutate_at(vars(-sex), function(x) x/1e6) %>%
+  mutate(fraction=dummy/sum(dummy),
+         odds=fraction/min(fraction)) %>%
+  select(sex, `no. rx. (millions)`=dummy, `SE (millions)`=se, odds) %>%
+  kable
 ```
@@
 ```{r models}
-des = visits %>%
+des = visit %>%
   mutate(sex=factor(sex, levels=c('male', 'female'))) %>%
-  svydesign(data=., id=~0, weights=~weight)
+  as.data.frame() %>%
+  svydesign(data=., id=~survey_id, strata=~stratum, weights=~weight)
 
-f = function(drug) {
-  svyglm(as.formula(sprintf('%s ~ sex', drug)), des, family=binomial(link='log'), start=c(-5, 0)) %>%
-    tidy %>%
-    filter(term=='sexfemale') %>%
-    mutate(drug=drug)
+abx_total_f = function(abx) {
+  svyby(as.formula(paste0('~', abx)), ~sex, des, svytotal) %>%
+    rename_(.dots=setNames(list(paste0(abx, 'TRUE'), paste0('se.', abx, 'TRUE')), c('total', 'se'))) %>%
+    select(sex, total, se) %>%
+    mutate(abx=abx)
 }
 
-sex_diff_by_drug = lapply(abx_names, f) %>%
-  bind_rows() %>%
-  mutate(ci=1.96*std.error,
-         ci_low=estimate-ci,
-         ci_hi=estimate+ci,
-         p.adj=p.adjust(p.value, method='BH')) %>%
-  mutate(sig=case_when(.$p.adj < 0.05 ~ '**',
-                       .$p.value < 0.05 ~ '*',
-                       TRUE ~ '')) %>%
-  select(-ci) %>%
-  mutate_at(vars(estimate, ci_low, ci_hi), function(x) exp(x))
+chisq_f = function(abx) {
+  svychisq(as.formula(paste0('~sex+', abx)), des) %>%
+    tidy %>%
+    mutate(abx=abx) %>%
+    select(abx, p.value)
+}
+
+ttest_f = function(abx) {
+  svyttest(as.formula(paste0(abx, '~sex')), des) %>%
+    tidy %>%
+    mutate(abx=abx) %>% select(abx, estimate, p.value)
+}
+
+abx_totals = lapply(top_abx$abx, abx_total_f) %>% bind_rows()
+abx_chisq = lapply(top_abx$abx, chisq_f) %>% bind_rows()
+abx_ttest = lapply(top_abx$abx, ttest_f) %>% bind_rows()
 ```
 
-```{r diff_plot}
-sex_diff_by_drug %>%
-  mutate(drug=forcats::fct_reorder(factor(drug), estimate)) %>%
-  ggplot(aes(x=drug, y=estimate, ymin=ci_low, ymax=ci_hi)) +
-  geom_point() + geom_linerange() +
-  geom_text(aes(y=1, label=sig), cex=6) +
-  coord_flip() +
-  scale_y_continuous(breaks=c(0, 1, 2, 5, 10), limits=c(0, 10)) +
-  ylab('relative risk of being female for getting drug')
+For each drug, I asked how many visits were made, stratified by sex.
+I show two lines on the graph.
+The lower lines is exactly $y=x$ and shows the line of equal consumption.
+The upper line has a slope equal to the odds of a visit belong to a female; this is the line of equal consumption per visit.
+
+```{r total_plot}
+abx_totals %>%
+  mutate_at(vars(total, se), function(x) x/1e6) %>%
+  mutate(ci=se*1.96,
+         ci_low=total-ci,
+         ci_high=total+ci) %>%
+  select(abx, sex, total, ci_low, ci_high) %>%
+  gather('key', 'value', total:ci_high) %>%
+  mutate(key=case_when(.$sex=='male' & .$key=='total' ~ 'x',
+                       .$sex=='male' & .$key=='ci_low' ~ 'xmin',
+                       .$sex=='male' & .$key=='ci_high' ~ 'xmax',
+                       .$sex=='female' & .$key=='total' ~ 'y',
+                       .$sex=='female' & .$key=='ci_low' ~ 'ymin',
+                       .$sex=='female' & .$key=='ci_high' ~ 'ymax')) %>%
+  select(-sex) %>%
+  spread(key, value) %>%
+  mutate(ymin=if_else(ymin<0, 0, ymin)) %>% # fix the negative ymin for moxi
+  mutate(highlight=abx %in% c('nitro')) %>%
+  ggplot(aes(x=x, xmin=xmin, xmax=xmax, y=y, ymin=ymin, ymax=ymax, label=abx, color=highlight)) +
+  geom_abline(col='gray') + geom_abline(slope=1.21277, col='gray') +
+  geom_point() + geom_errorbar() + geom_errorbarh() +
+  scale_color_manual(values=c('black', 'red')) +
+  geom_text(nudge_x=1, nudge_y=0.5) +
+  scale_x_continuous(breaks=c(0, 5, 10), limits=c(0, 10)) + xlab('No. rx. for men (millions)') +
+  scale_y_continuous(breaks=c(0, 5, 10), limits=c(0, 12)) + ylab('No. rx. for women (millions)') +
+  coord_fixed() +
+  theme_minimal()
 ```
 
-```{r diff_table}
-sex_diff_by_drug %>%
-  select(drug, ci_low, estimate, ci_hi, p.value, p.adj, sig) %>%
-  mutate_at(vars(ci_low:ci_hi), function(x) round(x, digits=2)) %>%
-  mutate_at(vars(p.value, p.adj), function(x) round(x, digits=3)) %>%
-  arrange(desc(estimate)) %>%
+Caveat: As per NAMCS recommendations, the male/nitro combination is too small (20 records) and too disperse (relative standard error 31%) to be considered accurate, and the female/moxi combination is too disperse (RSE 56%) to be considered accurate.
+(The female/moxi error bar actually doesn't even fit on the page!)
+
+## Claims per visit
+
+To examine claims per visit (the upper line), I run $\chi^2$ tests of independence on the 2x2 contingency tables: female vs.
+male, this drug prescribed vs.
+not.
+Nitrofurantoin is definitely distributed differently (minding the caveat mentioned above), and amoxicillin and amox/clav only marginally so (they are the two crosses that could maybe be falling *below* the line of equal consumption per visit; i.e., men getting more of those drugs per visit).
+
+```{r chisq_table}
+scientific_round = function(x, digits) {
+  if_else(round(x, digits) == 0, sprintf(paste0('%.', as.character(digits), 'g'), x), as.character(round(x, digits)))
+}
+
+abx_chisq %>%
+  arrange(p.value) %>%
+  mutate(p.value.adj=p.adjust(p.value, method='BH')) %>%
+  mutate_at(vars(-abx), function(x) scientific_round(x, 3)) %>%
   kable
@@
 
-```{r model_helpers}
-m = function(frmla) svyglm(frmla, design=des, family=binomial(link='log'))
+## Claims overall
 
-# explained excess risk ratio
-eerr = function(m1, m2, coeff='sexfemale') {
-  (1 - (exp(m2$coefficients[coeff]) - 1)/(exp(m1$coefficients[coeff]) - 1)) %>% unname
-}
-report_eerr = function(m1, m2) round(100 * eerr(m1, m2), 1)
+To evaluate the difference in overall numbers of claims by sex, I use *t*-tests. Again, nitro is the only clear deviation.
+
+```{r ttest_table}
+abx_ttest %>%
+  arrange(p.value) %>%
+  mutate(p.value.adj=p.adjust(p.value, method='BH')) %>%
+  mutate_at(vars(-abx), function(x) scientific_round(x, 3)) %>%
+  kable
 ```
 
-# Nitrofurantoin
+# Conclusions
 
-```{r nitro_models}
-nitro1 = m(nitrofurantoin ~ sex)
-nitro2 = m(nitrofurantoin ~ sex + uti)
-```
+Women definitely have more prescriptions overall than men, and definitely more nitrofurantoin prescriptions.
+However, the difference in nitrofurantoin prescriptions (slightly less than 2 million) is a small fraction of the total difference in prescriptions (about 11 million).
 
-This should be easy to explain by UTIs. They explain `r report_eerr(nitro1, nitro2)`% of the relative risk.
+Broadly speaking, it seems like the increased consumption of antibiotics among women is mostly due to increased healthcaure utilization.
+Unfortunately, NAMCS doesn't have enough data (at least from a single year) to be able to show that any particular drug beyond nitrofurantoin is definitely consumed more among women than men.
 
-# Metronidazole
+Interestingly, there are some hints that amox and amox/clav, though consumed at nearly-equal rates overall, are prescribed at a lower rate *per visit* for women.
 
-```{r metro_models}
-metro1 = m(metronidazole ~ sex)
-metro2 = m(metronidazole ~ sex + vaginitis)
-```
-
-Metronidazole consumption is increased in women. This is probably vaginitis (`61610`). So vaginitis explains `r report_eerr(metro1, metro2)`% of the relative risk. Not stupendous.
-
-# Clindamycin
-
-Clinda is only marginally significant in the models.
-
-```{r clinda_models}
-clinda1 = m(clindamycin ~ sex)
-clinda2 = m(clindamycin ~ sex + skin)
-```
-
-Skin conditions explain `r report_eerr(clinda1, clinda2)`% of the relative risk. Also not stupendous.
+The lack of power in this single year motivates either adding more years and/or moving toward Medicare.

### 2017-06-26 sex-difference/namcs.Rmd
methods

@@
 # Methods
+
+## Data & overall methods
+
+I used the 2014 NAMCS data.
+Because this is a survey, you can't just look at the individual records; they all need to be weighted, and the variances need to be adjusted by the stratum sampling used.
+I rely on the `survey` package, to which the entry point is the `svydesign` function, for which I use:
+
+- NAMCS's `CPSUM` variable for `ids`
+- `CSTRATM` for `strata`
+- `PATWT` for `weights`
 

### 2017-06-26 sex-difference/namcs.Rmd
small fixes

@@
 
-import::from(survey, svyglm)
-import::from(survey, svydesign)
+library(survey)
 ```
@@
   mutate(ymin=if_else(ymin<0, 0, ymin)) %>% # fix the negative ymin for moxi
-  mutate(highlight=abx %in% c('nitro')) %>%
+  mutate(highlight=abx %in% c('nitro', 'amox', 'amoxclav')) %>%
   ggplot(aes(x=x, xmin=xmin, xmax=xmax, y=y, ymin=ymin, ymax=ymax, label=abx, color=highlight)) +

### 2017-06-28 sex-difference/namcs.Rmd
cleaned up sex diff

@@
 ```{r load_data}
+# Caveat: I removed a stratum from the 2011 data that had only one PSU in it.
+# Also: I'm not sure that combining years like this is going to be kosher.
+
+years = 2011:2014
+n_years = length(years)
+mpy = function(x) x/(1e6 * n_years) # millions per year
+
+read_tsv_years = function(fn_template, ...) {
+  f = function(y) sprintf(fn_template, y) %>% read_tsv(..., guess_max=1e6) %>% mutate(year=y, visit_id=paste0(y, '_', visit_id))
+  lapply(years, f) %>% bind_rows
+}
+
 sex_codes = read_tsv('namcs_sex_codes.tsv') %>%
@@
 
-dx = read_tsv('../../db/namcs/data/namcs_2014_dx.tsv') %>%
+dx = read_tsv_years('../../db/namcs/data/namcs_dx_%i.tsv') %>%
   left_join(sex_codes, by='dx')
@@
   summarize(uti='uti' %in% diagnosis_type,
-            skin='skin' %in% diagnosis_type,
-            vaginitis='vaginitis' %in% diagnosis_type)
+            acne='acne' %in% diagnosis_type)
 
 top_abx = read_tsv('../../db/namcs/analysis/top_abx.tsv')
+abx_cat = read_tsv('../../db/namcs/data/parse/drug_codes/abx_codes.tsv')
 
-# abx_drug_codes = read_tsv('../../db/namcs/drug-codes/abx-codes.tsv') %>%
-#   select(drug_code, antibiotic) %>%
-#   mutate(antibiotic=tolower(antibiotic))
+drugs = read_tsv_years('../../db/namcs/data/namcs_drugs_%i.tsv') %>%
+  mutate(is_abx=cat1 %in% abx_cat$category)
 
-drug = read_tsv('../../db/namcs/data/namcs_2014_drugs.tsv')
+n_abx = drugs %>%
+  filter(is_abx) %>%
+  group_by(visit_id) %>%
+  summarize(n_abx=n())
 
-abx = top_abx %>% left_join(drug, by='desc')
+abx_visit = left_join(top_abx, drugs, by='desc') %>%
+  count(visit_id, abx) %>% ungroup() %>%
+  spread(abx, n)
 
-abx_visit = abx %>%
-  select(visit_id, abx) %>%
-  distinct() %>%
-  mutate(value=TRUE) %>% spread(abx, value) %>%
-  mutate(any_abx=TRUE)
+visits = read_tsv_years('../../db/namcs/data/namcs_visits_%i.tsv') %>%
+  mutate(age_group=case_when(.$age <= 19 ~ 'child',
+                             .$age <= 64 ~ 'adult',
+                             .$age <= 92 ~ 'elderly')) %>%
+  left_join(dx_visit, by='visit_id') %>%
+  replace_na(list(uti=FALSE, acne=FALSE)) %>%
+  left_join(n_abx, by='visit_id') %>% replace_na(list(n_abx=0)) %>%
+  left_join(abx_visit, by='visit_id') %>%
+  mutate_at(top_abx$abx, function(x) if_else(is.na(x), 0L, x))
 
-visit = read_tsv('../../db/namcs/data/namcs_2014_visits.tsv') %>%
-  left_join(dx_visit, by='visit_id') %>%
-  replace_na(list(uti=FALSE, skin=FALSE, vaginitis=FALSE)) %>%
-  left_join(abx_visit, by='visit_id') %>%
-  mutate_at(c('any_abx', top_abx$abx), function(x) !is.na(x))
+filter_strata = function(df) {
+  bad_strata = select(df, survey_id, stratum) %>% distinct() %>% count(stratum) %>% filter(n==1) %$% stratum
+  filter(df, !(stratum %in% bad_strata))
+}
+
+design_f = function(df) {
+  filter_strata(df) %>%
+    as.data.frame() %>%
+    svydesign(data=., ids=~survey_id, strata=~stratum, weights=~weight, nest=TRUE)
+}
+
+design = design_f(visits)
+adult_design = subset(design, age_group != 'child')
 ```
@@
 
-## The most-consumed antibiotics
+All numbers of prescriptions will be presented in millions per year.
 
-I wanted to focus my analysis on the top few most important antibiotics.
-The NAMCS data includes information about the therapeutic category of each drug.
-I looked through those categories and identified the ones that could refer to antibiotics.
+# Results
+
+## In children, overall consumption is similar between the sexes
+
+I show total consumption (in millions of prescriptions per year) for the three age groups (children 0-19, adults 20-64, elderly 64-92+).
+Consumption is similar between the sexes for children.
+(This probably isn't a surprise, since we don't expect that children have many UTIs or display large differences in healthcare seeking.) Therefore, I'll consider only adults for the rest of this analysis.
+
+```{r children}
+svyby(~n_abx, ~sex+age_group, design=design, svytotal) %>%
+  mutate_at(vars(n_abx, se), mpy) %>%
+  group_by(age_group) %>%
+  mutate(odds=n_abx/min(n_abx),
+         ci=1.96*se) %>%
+  select(age_group, sex, n_abx, `95% CI (one side)`=ci, odds) %>%
+  kable
+```
+
+## Most-consumed drugs
+
+The common explanation about differences in consumption is that women experience more UTIs than men.
+We therefore expect that the difference between the sexes will be mostly due to consumption of a few drugs (i.e., ciprofloxacin, TMP/SMX, nitrofurantoin).
+To simplify analysis, I'll focus on a few of the most-consumed drugs.
+
+The NAMCS data includes information about the therapeutic category of each drug. I looked through those categories and identified the ones that could refer to antibiotics.
 
 ```{r abx_cat_table}
-abx_cat = read_tsv('../../db/namcs/data/parse/drug_codes/abx_codes.tsv')
 abx_cat %>% kable
@@
 ```{r find_top_abx}
-drug %>%
-  semi_join(abx_cat, by=c('cat1'='category')) %>%
-  select(visit_id, desc) %>%
-  left_join(select(visit, visit_id, survey_id, stratum, weight), by='visit_id') %>%
-  mutate(dummy=1) %>%
-  svydesign(data=., ids=~survey_id, strata=~stratum, weights=~weight) %>%
-  svyby(~dummy, ~desc, design=., svytotal) %>%
-  arrange(desc(dummy)) %>%
-  select(abx=desc, n_rx=dummy, se) %>%
-  mutate(f=100*round(n_rx/sum(n_rx), 2),
-         cum_f_rx=100*round(cumsum(n_rx/sum(n_rx)), 2),
-         relative_se=100*round(se/n_rx, 2)) %>%
-  select(drug=abx, `fraction (%)`=f, `relative SE (%)`=relative_se, `cumulative fraction (%)`=cum_f_rx) %>%
+# get a list of all the antibiotic names
+abx_names = drugs %>%
+  semi_join(abx_cat, by=c('cat1'='category')) %$%
+  unique(desc)
+
+# for each name, count the visits in which it appears
+f = function(name) {
+  drugs %>%
+    filter(desc==name) %>%
+    count(visit_id) %>%
+    right_join(select(visits, visit_id, age_group, survey_id, stratum, weight), by='visit_id') %>%
+    replace_na(list(n=0)) %>%
+    design_f() %>%
+    subset(age_group != 'child') %>%
+    svytotal(~n, design=.) %>%
+    as_data_frame() %>%
+    mutate(abx=name) %>%
+    select(abx, total, se=n)
+}
+
+abx_counts = lapply(abx_names, f) %>% bind_rows()
+```
+
+```{r top_abx_table}
+abx_counts %>%
+  arrange(desc(total)) %>%
+  mutate(f=100*round(total/sum(total), 2),
+         cum_f_rx=100*round(cumsum(total/sum(total)), 2),
+         n=1:nrow(.)) %>%
+  select(n, drug=abx, `fraction (%)`=f, `cumulative fraction (%)`=cum_f_rx) %>%
   head(15) %>%
@@
 
-I kept the drugs from azithromycin through clindamycin.
-I skipped moxifloxacin because it has a high relative SE and mostly won't make it through future analyses.
-I keep nitrofurantoin because it's interesting from the Medicare data.
+I kept the drugs from azithromycin through clarithromycin, accounting for 80% of all prescriptions.
 
-# Results
+## Overall healthcare usage by sex
 
-## Overall consumption by sex
+```{r visit_total_odds}
+x = svymean(~sex, design=adult_design) %>%
+  as_data_frame() %>%
+  mutate(sex=c('female', 'male'),
+         odds=mean/min(mean))
 
-Overall, more of the antibiotic claims were made by women. Women consume roughly 21% more antibiotics than men.
+female_visit_odds = x %>% filter(sex=='female') %$% odds
 
-```{r sex_total_odds}
-drug %>%
-  semi_join(abx_cat, by=c('cat1'='category')) %>%
-  select(visit_id, desc) %>%
-  left_join(select(visit, visit_id, sex, survey_id, stratum, weight), by='visit_id') %>%
-  mutate(dummy=1) %>%
-  svydesign(data=., ids=~survey_id, strata=~stratum, weights=~weight) %>%
-  svyby(~dummy, ~sex, design=., svytotal) %>%
-  mutate_at(vars(-sex), function(x) x/1e6) %>%
-  mutate(fraction=dummy/sum(dummy),
-         odds=fraction/min(fraction)) %>%
-  select(sex, `no. rx. (millions)`=dummy, `SE (millions)`=se, odds) %>%
+x %>%
+  select(sex, `fraction of visits`=mean, SE, odds) %>%
   kable
@@
 
+Overall, more of the visits were made by women.
+Women make `r round(female_visit_odds - 1, 2)*100`% more visits than men.
+I'll therefore evaluate consumption by two standards: first, do women consume more antibiotics overall; second, do women consume more antibiotics *per visit*.
+
+## Consumption by drug
+
 ```{r models}
-des = visit %>%
-  mutate(sex=factor(sex, levels=c('male', 'female'))) %>%
-  as.data.frame() %>%
-  svydesign(data=., id=~survey_id, strata=~stratum, weights=~weight)
+abx_diff_f = function(abx) {
+  visits %>%
+    mutate(sex_abx=case_when(.$sex == 'female' ~ .[[abx]],
+                             .$sex == 'male' ~ -.[[abx]])) %>%
+    design_f() %>%
+    subset(age_group != 'child') %>%
+    svyttest(sex_abx~0, design=.) %$%
+    data_frame(abx=abx, estimate=estimate, p.value=p.value)
+}
 
 abx_total_f = function(abx) {
-  svyby(as.formula(paste0('~', abx)), ~sex, des, svytotal) %>%
-    rename_(.dots=setNames(list(paste0(abx, 'TRUE'), paste0('se.', abx, 'TRUE')), c('total', 'se'))) %>%
+  svyby(as.formula(paste0('~', abx)), ~sex, adult_design, svytotal) %>%
+    rename_(.dots=setNames(list(abx), c('total'))) %>%
     select(sex, total, se) %>%
@@
 
-chisq_f = function(abx) {
-  svychisq(as.formula(paste0('~sex+', abx)), des) %>%
+poisson_f = function(abx) {
+  svyglm(as.formula(paste0(abx, '~sex')), adult_design) %>%
     tidy %>%
+    filter(term != '(Intercept)') %>%
     mutate(abx=abx) %>%
-    select(abx, p.value)
+    select(abx, term, estimate, p.value)
 }
 
-ttest_f = function(abx) {
-  svyttest(as.formula(paste0(abx, '~sex')), des) %>%
-    tidy %>%
-    mutate(abx=abx) %>% select(abx, estimate, p.value)
-}
-
+abx_diff = lapply(top_abx$abx, abx_diff_f) %>% bind_rows()
 abx_totals = lapply(top_abx$abx, abx_total_f) %>% bind_rows()
-abx_chisq = lapply(top_abx$abx, chisq_f) %>% bind_rows()
-abx_ttest = lapply(top_abx$abx, ttest_f) %>% bind_rows()
+abx_pois = lapply(top_abx$abx, poisson_f) %>% bind_rows()
 ```
@@
 I show two lines on the graph.
-The lower lines is exactly $y=x$ and shows the line of equal consumption.
+The lower line is exactly $y=x$ and shows the line of equal consumption.
 The upper line has a slope equal to the odds of a visit belong to a female; this is the line of equal consumption per visit.
@@
 abx_totals %>%
-  mutate_at(vars(total, se), function(x) x/1e6) %>%
+  mutate_at(vars(total, se), mpy) %>%
   mutate(ci=se*1.96,
@@
   spread(key, value) %>%
-  mutate(ymin=if_else(ymin<0, 0, ymin)) %>% # fix the negative ymin for moxi
-  mutate(highlight=abx %in% c('nitro', 'amox', 'amoxclav')) %>%
+  mutate(highlight=abx %in% c('nitro', 'clinda')) %>%
   ggplot(aes(x=x, xmin=xmin, xmax=xmax, y=y, ymin=ymin, ymax=ymax, label=abx, color=highlight)) +
-  geom_abline(col='gray') + geom_abline(slope=1.21277, col='gray') +
+  geom_abline(col='gray') + geom_abline(slope=female_visit_odds, col='gray') +
   geom_point() + geom_errorbar() + geom_errorbarh() +
   scale_color_manual(values=c('black', 'red')) +
-  geom_text(nudge_x=1, nudge_y=0.5) +
-  scale_x_continuous(breaks=c(0, 5, 10), limits=c(0, 10)) + xlab('No. rx. for men (millions)') +
-  scale_y_continuous(breaks=c(0, 5, 10), limits=c(0, 12)) + ylab('No. rx. for women (millions)') +
+  geom_text() +
+  #geom_text(nudge_x=1, nudge_y=0.5) +
+  #scale_x_continuous(breaks=c(0, 5, 10), limits=c(0, 12)) + xlab('No. rx. for men (millions)') +
+  #scale_y_continuous(breaks=c(0, 5, 10), limits=c(0, 14)) + ylab('No. rx. for women (millions)') +
   coord_fixed() +
@@
 
-Caveat: As per NAMCS recommendations, the male/nitro combination is too small (20 records) and too disperse (relative standard error 31%) to be considered accurate, and the female/moxi combination is too disperse (RSE 56%) to be considered accurate.
-(The female/moxi error bar actually doesn't even fit on the page!)
+## Prescriptions overall
 
-## Claims per visit
+To evaluate the difference in overall numbers of prescriptions by sex (i.e., deviations from the $y=x$ line in the plot), I performed a one-sample *t*-test on a kludge variable.
+This variable is the number of antibiotic prescriptions if the visit was made by a woman and *negative* the number of prescriptions if made by a man.
+The test asks if this value has mean zero.
 
-To examine claims per visit (the upper line), I run $\chi^2$ tests of independence on the 2x2 contingency tables: female vs.
-male, this drug prescribed vs.
-not.
-Nitrofurantoin is definitely distributed differently (minding the caveat mentioned above), and amoxicillin and amox/clav only marginally so (they are the two crosses that could maybe be falling *below* the line of equal consumption per visit; i.e., men getting more of those drugs per visit).
+As we would expect from examining the chart, almost all the drugs have a mean greater than zero, i.e., consumption is greater among women.
 
-```{r chisq_table}
-scientific_round = function(x, digits) {
+```{r abx_diff_table}
+scientific_round = function(x, digits=3) {
   if_else(round(x, digits) == 0, sprintf(paste0('%.', as.character(digits), 'g'), x), as.character(round(x, digits)))
@@
 
-abx_chisq %>%
-  arrange(p.value) %>%
-  mutate(p.value.adj=p.adjust(p.value, method='BH')) %>%
-  mutate_at(vars(-abx), function(x) scientific_round(x, 3)) %>%
+abx_diff %>%
+  mutate(adj.p.value=p.adjust(p.value, method='BH')) %>%
+  mutate(sig=case_when(.$adj.p.value < 0.05 ~ '**',
+                       .$p.value < 0.05 ~ '*',
+                       TRUE ~ '')) %>%
+  mutate_if(is.numeric, scientific_round) %>%
   kable
@@
 
-## Claims overall
+## Prescriptions per visit
 
-To evaluate the difference in overall numbers of claims by sex, I use *t*-tests. Again, nitro is the only clear deviation.
+To examine claims per visit (the upper line), I run Poisson regressions of the number of prescriptions for each drug per visit against sex.
+The *p*-values here are for the significance of the sex term.
+Nitro and clinda are enriched among women *per visit*.
 
-```{r ttest_table}
-abx_ttest %>%
+```{r pois_table}
+abx_pois %>%
+  select(abx, estimate, p.value) %>%
   arrange(p.value) %>%
   mutate(p.value.adj=p.adjust(p.value, method='BH')) %>%
-  mutate_at(vars(-abx), function(x) scientific_round(x, 3)) %>%
+  mutate_at(vars(-abx), scientific_round) %>%
   kable
@@
 
+## Nitrofurantoin
+
+Nitrofurantoin prescriptions among women are associated with a number of diagnoses, but only two of them---UTI (`5990-`) and dysuria (`7881-`)---meet the basic requirement for data validity (at least 30 counts).
+Among men, UTI is the leading diagnosis, but it doesn't meet the data validity requirement (there are only 10 records for nitro prescriptions to a male with a UTI diagnosis in the 4 years of data).
+
+## Clindamycin
+
+```{r clinda}
+m1 = svyglm(clinda~sex, design=adult_design, family='poisson')
+m2 = svyglm(clinda~sex+acne, design=adult_design, family='poisson')
+clinda_f_expl = 1 - exp(-m2$coefficients['sexmale']+m1$coefficients['sexmale'])
+```
+
+The only clindamycin-associated diagnosis that meets (maybe) the data validity criterion is acne (`7061-`).
+However, acne explains only a small fraction (`r 100*round(clinda_f_expl, digits=3)`%) of the difference between the sexes.
+
 # Conclusions
 
-Women definitely have more prescriptions overall than men, and definitely more nitrofurantoin prescriptions.
-However, the difference in nitrofurantoin prescriptions (slightly less than 2 million) is a small fraction of the total difference in prescriptions (about 11 million).
-
-Broadly speaking, it seems like the increased consumption of antibiotics among women is mostly due to increased healthcaure utilization.
-Unfortunately, NAMCS doesn't have enough data (at least from a single year) to be able to show that any particular drug beyond nitrofurantoin is definitely consumed more among women than men.
-
-Interestingly, there are some hints that amox and amox/clav, though consumed at nearly-equal rates overall, are prescribed at a lower rate *per visit* for women.
-
-The lack of power in this single year motivates either adding more years and/or moving toward Medicare.
+Women definitely have more prescriptions overall than men, but that consumption seems to be distributed broadly across drugs.
+For only a few drugs---nitro and clinda---do these data suggest that consumption per visit is greater among women, and those drugs contribute to a minority of the difference in total consumption.

### 2017-06-28 sex-difference/namcs.Rmd
sex diff

@@
   as_data_frame() %>%
-  mutate(sex=c('female', 'male'),
-         odds=mean/min(mean))
+  mutate(sex=c('female', 'male'))
 
-female_visit_odds = x %>% filter(sex=='female') %$% odds
+odds = function(x) x/(1-x)
+
+p = x %>% filter(sex=='female') %$% mean
+se = x %>% filter(sex=='female') %$% SE
+female_visit_odds = odds(c(p-se, p, p+se))
 
 x %>%
-  select(sex, `fraction of visits`=mean, SE, odds) %>%
+  select(sex, `fraction of visits`=mean, SE) %>%
   kable
@@
 Overall, more of the visits were made by women.
-Women make `r round(female_visit_odds - 1, 2)*100`% more visits than men.
+Women make `r round(female_visit_odds[2], 3)*100`% more visits than men.
 I'll therefore evaluate consumption by two standards: first, do women consume more antibiotics overall; second, do women consume more antibiotics *per visit*.
@@
 
-abx_total_f = function(abx) {
-  svyby(as.formula(paste0('~', abx)), ~sex, adult_design, svytotal) %>%
+abx_total_f = function(abx, design=adult_design) {
+  svyby(as.formula(paste0('~', abx)), ~sex, design, svytotal) %>%
     rename_(.dots=setNames(list(abx), c('total'))) %>%
@@
 
-poisson_f = function(abx) {
-  svyglm(as.formula(paste0(abx, '~sex')), adult_design) %>%
+poisson_f = function(abx, design=adult_design) {
+  svyglm(as.formula(paste0(abx, '~sex')), design) %>%
     tidy %>%
@@
 ```{r total_plot}
-abx_totals %>%
+total_plot_f = function(df, highlight=character(0)) {
+  df %>%
     mutate_at(vars(total, se), mpy) %>%
@@
     spread(key, value) %>%
-  mutate(highlight=abx %in% c('nitro', 'clinda')) %>%
+    mutate(highlight=abx %in% highlight) %>%
     ggplot(aes(x=x, xmin=xmin, xmax=xmax, y=y, ymin=ymin, ymax=ymax, label=abx, color=highlight)) +
-  geom_abline(col='gray') + geom_abline(slope=female_visit_odds, col='gray') +
+    geom_abline(col='gray') +
     geom_point() + geom_errorbar() + geom_errorbarh() +
@@
     #geom_text(nudge_x=1, nudge_y=0.5) +
-  #scale_x_continuous(breaks=c(0, 5, 10), limits=c(0, 12)) + xlab('No. rx. for men (millions)') +
-  #scale_y_continuous(breaks=c(0, 5, 10), limits=c(0, 14)) + ylab('No. rx. for women (millions)') +
+    #scale_x_continuous(breaks=c(0, 5, 10), limits=c(0, 12)) +
+    #scale_y_continuous(breaks=c(0, 5, 10), limits=c(0, 14)) +
+    xlab('No. rx. for men (millions)') + ylab('No. rx. for women (millions)') +
     coord_fixed() +
     theme_minimal()
+}
+
+total_plot_f(abx_totals, highlight=c('nitro', 'clinda')) +
+  geom_abline(slope=female_visit_odds[2], col='gray') +
+  geom_abline(slope=female_visit_odds[-2], col='gray', linetype='dotted')
 ```
@@
 
+sig_f = function(df) {
+  stopifnot('adj.p.value' %in% names(df) && 'p.value' %in% names(df))
+  case_when(df$adj.p.value < 0.05 ~ '**',
+            df$p.value < 0.05 ~ '*',
+            TRUE ~ '')
+}
+
 abx_diff %>%
   mutate(adj.p.value=p.adjust(p.value, method='BH')) %>%
-  mutate(sig=case_when(.$adj.p.value < 0.05 ~ '**',
-                       .$p.value < 0.05 ~ '*',
-                       TRUE ~ '')) %>%
+  mutate(sig=sig_f(.)) %>%
   mutate_if(is.numeric, scientific_round) %>%
@@
   arrange(p.value) %>%
-  mutate(p.value.adj=p.adjust(p.value, method='BH')) %>%
-  mutate_at(vars(-abx), scientific_round) %>%
+  mutate(adj.p.value=p.adjust(p.value, method='BH')) %>%
+  mutate(sig=sig_f(.)) %>%
+  mutate_if(is.numeric, scientific_round) %>%
   kable
@@
 
+## Just among the elderly
+
+```{r elderly}
+elderly_design = subset(design, age_group == 'elderly')
+
+elderly_usage = svymean(~sex, design=elderly_design) %>%
+  as_data_frame() %>%
+  mutate(sex=c('female', 'male'))
+
+elderly_female_odds = elderly_usage %>% filter(sex=='female') %$% mean %>% odds
+
+elderly_totals = abx_totals = lapply(top_abx$abx, function(x) abx_total_f(x, design=elderly_design)) %>% bind_rows()
+total_plot_f(elderly_totals) +
+  geom_abline(slope=elderly_female_odds, color='gray')
+```
+
+Usage among the elderly is more even: women make `r 100*round(elderly_female_odds - 1, digits=3)`% more visits than men.
+Patterns of consumption appear to be a little difference: more ciprofloxacin, azithro is more concentrated among women, and some drugs (e.g., doxy) appear to approach the line of equal consumption.
+
 # Conclusions
@@
 For only a few drugs---nitro and clinda---do these data suggest that consumption per visit is greater among women, and those drugs contribute to a minority of the difference in total consumption.
+
+Among the elderly, patterns of consumption appear to be somewhat different, and those patterns are not exactly what we see in the adult NAMCS data or even in the Medicare data I have.

### 2017-06-28 sex-difference/sex-difference.Rmd
sex diff

--- 2017-06-20 sex-difference/sex-difference.Rmd
@@
 
+## Top drugs
+
+Which drugs are the most used?
+
+```{r top_drugs}
+top_abx = pde %>%
+  count(antibiotic) %>%
+  arrange(desc(n)) %>%
+  head(20)
+
+kable(top_abx)
+```
+
 ## Consumption of drugs by sex
@@
   bene %>%
-    select(-n_claims) %>%
-    left_join(pde %>% filter(antibiotic==drug) %>% count(bene_id) %>% rename(n_claims=n), by='bene_id') %>%
-    replace_na(list(n_claims=0)) %>%
+    left_join(pde %>% filter(antibiotic==drug) %>% count(bene_id), by='bene_id') %>%
+    replace_na(list(n=0)) %>%
     group_by(sex) %>%
-    summarize(mean_n_claims=mean(n_claims)) %>%
-    ungroup() %>%
-    spread(sex, mean_n_claims) %>%
-    mutate(drug=drug)
+    summarize(n_bene=n(),
+              n_claims=sum(n),
+              mean_n_claims=mean(n),
+              sd_n_claims=sd(n)) %>%
+    mutate(se_n_claims=sd_n_claims/sqrt(n_bene),
+           drug=drug)
 }
 
-dbs = lapply(unique(pde$antibiotic), usage_by_sex) %>% bind_rows
+ubs = lapply(top_abx$antibiotic, usage_by_sex) %>% bind_rows
 ```
@@
   scale_color_manual(values=c('black', 'red')) +
+  geom_abline() +
+  geom_abline(slope=1.28, col='gray') +
   geom_text() +
   geom_point() +
-  geom_abline() +
   xlab('claims per 1,000 male beneficiaries') +
@@
   coord_fixed()
+```
+
+```{r drug_by_sex_plot2}
+total_plot_f = function(df, highlight=character(0)) {
+  df %>%
+    #mutate_at(vars(total, se), mpy) %>%
+    mutate(ci=se*1.96,
+           ci_low=total-ci,
+           ci_high=total+ci) %>%
+    select(drug, sex, total, ci_low, ci_high) %>%
+    gather('key', 'value', total, ci_low, ci_high) %>%
+    mutate(key=case_when(.$sex=='male' & .$key=='total' ~ 'x',
+                         .$sex=='male' & .$key=='ci_low' ~ 'xmin',
+                         .$sex=='male' & .$key=='ci_high' ~ 'xmax',
+                         .$sex=='female' & .$key=='total' ~ 'y',
+                         .$sex=='female' & .$key=='ci_low' ~ 'ymin',
+                         .$sex=='female' & .$key=='ci_high' ~ 'ymax')) %>%
+    select(-sex) %>%
+    spread(key, value) %>%
+    mutate(highlight=drug %in% highlight) %>%
+    ggplot(aes(x=x, xmin=xmin, xmax=xmax, y=y, ymin=ymin, ymax=ymax, label=drug, color=highlight)) +
+    geom_abline(col='gray') +
+    geom_point() + geom_errorbar() + geom_errorbarh() +
+    scale_color_manual(values=c('black', 'red')) +
+    geom_text() +
+    #geom_text(nudge_x=1, nudge_y=0.5) +
+    #scale_x_continuous(breaks=c(0, 5, 10), limits=c(0, 12)) +
+    #scale_y_continuous(breaks=c(0, 5, 10), limits=c(0, 14)) +
+    xlab('No. rx. for men (millions)') + ylab('No. rx. for women (millions)') +
+    coord_fixed() +
+    theme_minimal()
+}
+
+ubs %>%
+  rename(total=mean_n_claims, se=se_n_claims) %>%
+  total_plot_f()
 ```
@@
 
-```{r}
-group_by_all_ = function(df, group_name) {
-  bind_rows(df, mutate_(df, .dots=setNames(list(~'all'), group_name))) %>%
-    group_by_(group_name)
-}
-
-epi %>%
-  group_by_all_('sex') %>%
-  summarize(n_cases=sum(n)) %>%
+```{r, eval=FALSE}
+# group_by_all_ = function(df, group_name) {
+#   bind_rows(df, mutate_(df, .dots=setNames(list(~'all'), group_name))) %>%
+#     group_by_(group_name)
+# }
+# 
+# epi %>%
+#   group_by_all_('sex') %>%
+#   summarize(n_cases=sum(n)) %>%
 
@@
 
-```{r consumption_by_age_and_sex}
+```{r consumption_by_age_and_sex, eval=FALSE}
 bene %>%
@@
 ```
+
+## Margins of azithromycin consumption
+
+Maybe women take more azithromycin because usage is more concentrated in women than in men.
+Then we could explain the increased consumption by a small number of people taking more drug.
+(Yonatan points to Lady Windmere syndrome.) In fact, consumption is more even among women.
+As always, this is driven by the extensive fraction: more women took at least one dose:
+
+```{r, eval=FALSE}
+pde %>%
+  filter(drug=='azithromycin') %>%
+  count(bene_id) %>% rename(n=n_azi) %>%
+  right_join(bene, by='bene_id') %>% replace_na(list(n_azi=0)) %>%
+  count(sex, n_azi) %>%
+  group_by(sex) %>%
+  mutate(f_ppl=n/sum(n), cum_f_ppl=cumsum(f_ppl),
+         cum_cons=cumsum(n_azi*n), cum_f_cons=cum_cons/max(cum_cons)) %>%
+  ggplot(aes(x=cum_f_ppl, y=cum_f_cons, color=sex)) +
+  geom_point() +
+  geom_line() +
+  xlab('Fraction of consumers') + ylab("Fraction of that sex's azithromycin consumed")
+```

### 2017-07-03 ms1/namcs_trends.Rmd
fixes

---
title: "Temporal trends (NAMCS)"
author: "Scott Olesen"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/', dev=c('png', 'pdf'),
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE, autodep=TRUE)

library(survey)
```

```{r load_data}
# Caveat: I removed a stratum from the 2011 data that had only one PSU in it.
# Also: I'm not sure that combining years like this is going to be kosher.

years = 2011:2014
n_years = length(years)
mpy = function(x) x/(1e6 * n_years) # millions per year

read_tsv_years = function(fn_template, ...) {
  f = function(y) sprintf(fn_template, y) %>% read_tsv(..., guess_max=1e6) %>% mutate(year=y, visit_id=paste0(y, '_', visit_id))
  lapply(years, f) %>% bind_rows
}

sex_codes = read_tsv('../sex-difference/namcs_sex_codes.tsv') %>%
  rename(dx=code)

dx = read_tsv_years('../../db/namcs/data/namcs_dx_%i.tsv') %>%
  left_join(sex_codes, by='dx')

dx_visit = dx %>%
  group_by(visit_id) %>%
  summarize(acute_rc='acute_rc' %in% diagnosis_type)

top_abx = read_tsv('../../db/namcs/analysis/top_abx.tsv')
abx_cat = read_tsv('../../db/namcs/data/parse/drug_codes/abx_codes.tsv')

drugs = read_tsv_years('../../db/namcs/data/namcs_drugs_%i.tsv') %>%
  mutate(is_abx=cat1 %in% abx_cat$category)

n_abx = drugs %>%
  filter(is_abx) %>%
  group_by(visit_id) %>%
  summarize(n_abx=n())

abx_visit = left_join(top_abx, drugs, by='desc') %>%
  count(visit_id, abx) %>% ungroup() %>%
  spread(abx, n)

visits = read_tsv_years('../../db/namcs/data/namcs_visits_%i.tsv') %>%
  mutate(age_group=case_when(.$age <= 19 ~ 'child',
                             .$age <= 64 ~ 'adult',
                             .$age <= 92 ~ 'elderly')) %>%
  left_join(dx_visit, by='visit_id') %>%
  replace_na(list(acute_rc=FALSE)) %>%
  left_join(n_abx, by='visit_id') %>% replace_na(list(n_abx=0)) %>%
  left_join(abx_visit, by='visit_id') %>%
  mutate_at(top_abx$abx, function(x) if_else(is.na(x), 0L, x))

filter_strata = function(df) {
  bad_strata = select(df, survey_id, stratum) %>% distinct() %>% count(stratum) %>% filter(n==1) %$% stratum
  filter(df, !(stratum %in% bad_strata))
}

design_f = function(df) {
  filter_strata(df) %>%
    as.data.frame() %>%
    svydesign(data=., ids=~survey_id, strata=~stratum, weights=~weight, nest=TRUE)
}

design = design_f(visits)
adult_design = subset(design, age_group != 'child')
```

# Methods

## Data & overall methods

I used the 2011-2014 NAMCS data. All numbers of prescriptions will be presented in millions per year. I'll mostly consider only the adults.

## Drugs

I consider the same drugs as last time:

```{r}
kable(top_abx)
```

# Results

## Summary of trends

```{r drugs_time}
f = function(design, y, d) {
  subset(design, year==y) %>%
    svytotal(as.formula(paste0('~', d)), design=.) %>%
    data_frame(year=y, abx=d, total=., se=as.numeric(SE(.)))
}

g = function(design) {
  crossing(year=2011:2014, desc=top_abx$abx) %>%
    rowwise() %>% do(f(design, .$year, .$desc)) %>% ungroup()
}

abx_time = g(design)
abx_time_adult = g(adult_design)
```

```{r drugs_time_plot}
abx_time %>%
  mutate(ci=1.96*se, ymin=total-ci, ymax=total+ci) %>%
  ggplot(aes(x=year, y=total, ymin=ymin, ymax=ymax)) +
  facet_wrap(~abx, ncol=3, scales='free_y') +
  geom_point() + geom_line() + geom_linerange() +
  theme_minimal()
```

It doesn't look by eye that any trends are particularly strong. I used Poisson regressions to see if anything was changing with time. Only levo:

```{r drugs_time_model}
qpglm = function(design, frmla) svyglm(frmla, design=design, family=quasipoisson(link='log'))

f = function(design, d) {
  qpglm(design, as.formula(paste0(d, '~ year'))) %>%
    tidy %>%
    mutate(abx=d) %>%
    filter(term != '(Intercept)')
}

g = function(design) lapply(top_abx$abx, function(x) f(design, x)) %>% bind_rows()

abx_time_model = g(design)
abx_time_model_adult = g(adult_design)
```

```{r drugs_time_model_table}
abx_time_model %>%
  select(abx, estimate, p.value) %>%
  mutate(adj.p.value=p.adjust(p.value, method='BH')) %>%
  arrange(p.value) %>%
  kable
```

```{r levo_plot}
american_population = data_frame(year=2011:2014, n_ppl=c(312.92, 315.22, 317.47, 319.85) * 1e6)

levo_time = abx_time %>%
  filter(abx=='levo') %>%
  left_join(american_population, by='year') %>%
  mutate(total=total/n_ppl*1000, se=se/n_ppl*1000)

levo_time %>%
  mutate(ci=1.96*se, ymin=total-ci, ymax=total+ci) %>%
  ggplot(aes(x=year, y=total, ymin=ymin, ymax=ymax)) +
  geom_point() + geom_line() + geom_linerange() +
  ylim(5, 21) +
  ylab('Levo claims per 1,000 Americans') +
  theme_minimal()
```

```{r levo_ratio}
t1 = levo_time %>% filter(year==2011) %$% total
v1 = levo_time %>% filter(year==2011) %$% se
t2 = levo_time %>% filter(year==2014) %$% total
v2 = levo_time %>% filter(year==2014) %$% se

tr = t2 / t1
vr = v1 / t2^2 + t1^1 / t2^4 * v2
sdr = sqrt(vr)

show_tot = function(tot, v) sprintf('%g (±%g)', round(tot, 2), round(v, 2))
show_pct = function(tot, v) sprintf('%g%% (±%g%%)', round((tot - 1)*100, digits=1), round(sqrt(v)*100, digits=1))
```

Levo consumption increased from `r show_tot(t1, v1)` prescriptions per 1,000 Americans per year to `r show_tot(t2, v2)`. That is an increase of `r show_pct(tr, vr)`, i.e., doubled.

## Presence of this trend in different age groups

I can confirm that levo is the only drug with a strong trend among adults (rather than everyone):

```{r drugs_time_model_adult_table}
abx_time_model_adult %>%
  select(abx, estimate, p.value) %>%
  mutate(adj.p.value=p.adjust(p.value, method='BH')) %>%
  arrange(p.value) %>%
  kable
```

And, in fact, this trend exists just among the elderly, although it's only marginally significant if you tested for everything.

```{r levo_elderly}
design %>% subset(age_group=='elderly') %>%
  qpglm(levo~year) %>% tidy
```

Levo may be on the decline for children, however:

```{r levo_children}
design %>% subset(age_group=='child') %>%
  qpglm(levo~year) %>% tidy
```

# Conclusions

Levo is definitely going up.

### 2017-07-03 sex-difference/sex-difference.Rmd
fixes

@@
 ```{r load_data}
-year = 2011
+year = 2014
 
@@
 
-```{r}
+```{r compute_ubs}
 denom = bene %>% count(sex) %>% rename(n_denom=n)
@@
   geom_abline() +
-  geom_abline(slope=1.28, col='gray') +
+  #geom_abline(slope=c(1.06, 1.109, 1.366, 1.41, 2.202), col='gray') +
+  #geom_abline(slope=c(1.062406, 1.0976, 1+0.26*1.268177), col='blue') +
+  geom_abline(slope=exp(3)*(1.06)) +
+  # NAMCS: all adult usage 1.495, UTI: 2.164, ARC: 1.0976
+  # Medicare: nabxclaims 1.24, n OP days 1.062406, arc 1.108944, uti 2.268177, copd MALE! 1.123728
   geom_text() +
@@
 
-```{r drug_by_sex_plot2}
+```{r}
+dx %>%
+  count(bene_id, diagnosis_type) %>%
+  ungroup() %>%
+  right_join(bene, by='bene_id') %>%
+  group_by(sex) %>%
+  summarize(mn=mean(n), n_bene=n())
+```
+
+```{r drug_by_sex_plot2, eval=FALSE}
 total_plot_f = function(df, highlight=character(0)) {
@@
 ubs %>%
-  rename(total=mean_n_claims, se=se_n_claims) %>%
-  total_plot_f()
+  gather('sex', 'total', female, male) %>%
+  left_join(bene %>% count(sex) %>% rename(n_bene=n), by='sex') %>%
+  mutate(se=total/sqrt(n_bene)) %>%
+  rename(drug=antibiotic) %>%
+  total_plot_f() +
+  geom_abline(slope=c(1.06, 1.306122), col='gray')
 ```

### 2017-07-03 sex-difference/sex-difference.Rmd
many updates

--- 2017-06-28 sex-difference/sex-difference.Rmd
@@
 
-# To do
-
-- Separate beneficiaries into "high users" and "low users" based on the median number of outpatient claim dates
-- Try segmenting azithromycin consumption by low/high in wo/men and by RC+/- in wo/men.
-We want to say that RC, more than low/high, accounts for differences in azithromycin risk, i.e., it's that women get more respiratory complaint diagnoses, not that women just have more complaints.
-Or vice versa.
-
 # Background
@@
 
-I'll use the Medicare 2011 data. Or, use all years, pool that data, and include calendar year as part of the regression.
+*Data*: Medicare 2011-2014 in the cohort.
 
-*Outcome*: At least one claim for azithromycin, Bactrim, TMP/SMX, or nitrofurantoin. *Exposure*: A diagnosis with a UTI, COPD, or an acute RTI.
+*Outcome*: Antibiotic claims overall and for the top 10 or so individual drugs.
 
-- In the first pass, look for any claim and any diagnosis.
-- In a second pass, look for a diagnosis that came at most two weeks before a claim.
-- Maybe, in a third pass, look for diagnosis that came at most two weeks before the first claim for that drug for the beneficiary.
-(Imagine someone is on year-long prophylaxis.
-Then a diagnosis at any time of the year will make it look like the diagnosis caused the claims.)
-
-Before a model, I should just stratify by sex and exposure. Do we already see the signal (or lack of it)? Then we can go onto including age and calendar year as covariates.
+*Exposure*: A diagnosis with a UTI, COPD, or an acute RTI.
 
 ```{r load_data}
-year = 2011
+sum_over = function(target_columns) {
+  dummy_vars = sapply(1:length(target_columns), function(i) sprintf('x%i', i))
+  dummy_vars_string = paste0(dummy_vars, collapse=',')
+  dummy_vars_list = lapply(target_columns, as.name) %>% setNames(dummy_vars)
+  f = as.formula(paste0(c('~sum(', dummy_vars_string, ')'), collapse=''))
+  lazyeval::interp(f, .values=dummy_vars_list)
+}
 
-pde = read_tsv(sprintf('../pde_%i.tsv', year))
+regions = read_tsv('../../db/census-regions/census-regions.tsv') %>%
+  select(state, region)
 
-sex_icd = read_tsv('sex_codes.tsv', col_types=cols(code='c'))
+condition_names = read_tsv('../chronic-conditions/names.tsv') %>% filter(condition_code != 'ALZH')
+names_1yr_ccs = condition_names %>% filter(condition_ref_years==1) %$% condition_code
 
-dx_car = sprintf('../../data/dx_car_claims_%i.tsv', year) %>%
-  read_tsv(col_types=cols(diagnosis='c'))
-dx_op = sprintf('../../data/dx_op_claims_%i.tsv', year) %>%
-  read_tsv(col_types=cols(diagnosis='c'))
+dx_codes = read_tsv('../sex-difference/sex_codes.tsv')
 
-dx = bind_rows(dx_car, dx_op) %>%
+load_data = function(year) {
+  # pde data
+  pde = read_tsv(sprintf('../pde_%i.tsv', year)) %>%
+    select(bene_id, antibiotic, days_supply) %>%
+    group_by(bene_id, antibiotic) %>%
+    summarize(n_claims=n(), days_supply=sum(days_supply)) %>%
+    ungroup() %>%
+    mutate(year=year)
+
+  # diagnosis claims
+  dx_car_fn = sprintf('../../data/dx_car_claims_%i.tsv', year)
+  dx_op_fn = sprintf('../../data/dx_op_claims_%i.tsv', year)
+
+  dx = bind_rows(read_tsv(dx_car_fn),
+                 read_tsv(dx_op_fn)) %>%
     mutate(date=dmy(from_date)) %>%
-  select(bene_id=BENE_ID, date, code=diagnosis) %>%
-  left_join(sex_icd, by='code')
+    rename(bene_id=BENE_ID, code=diagnosis) %>%
+    left_join(dx_codes, by='code')
 
-dx_counts = dx %>%
-  count(code)
+  dx_presence = dx %>%
+    group_by(bene_id) %>%
+    summarize(rc='acute_rc' %in% diagnosis_type,
+              uti='uti' %in% diagnosis_type)
 
-bene_dx = dx %>%
-  select(bene_id, diagnosis_type) %>%
-  distinct() %>%
-  mutate(dummy=TRUE) %>%
-  spread(diagnosis_type, dummy)
+  count_car_fn = sprintf('../../data/count_car_claims_%i.tsv', year)
+  count_op_fn = sprintf('../../data/count_op_claims_%i.tsv', year)
+  n_op_days = bind_rows(read_tsv(count_car_fn), read_tsv(count_op_fn)) %>%
+    rename(bene_id=BENE_ID) %>%
+    group_by(bene_id) %>%
+    summarize(n_op_days=sum(n_from_date))
 
-rc_dx = dx %>%
-  filter(diagnosis_type=='acute_rc') %>%
-  select(bene_id, date) %>%
+  # total numbers of PDEs
+  n_claims = pde %>% group_by(bene_id) %>% summarize(n_claims=sum(n_claims))
+
+  # count chronic conditions. 'n_cc' means one-year cc's
+  cc = read_feather(sprintf('../cc_%i.feather', year)) %>%
+    rowwise() %>%
+    mutate_(n_cc=sum_over(names_1yr_ccs)) %>%
+    ungroup()
+
+  # join the cc, summary PDE, and dx data into bene
+  bene = read_tsv(sprintf('../bene_%i.tsv', year)) %>%
+    mutate(is_female=sex=='female', is_white=race=='white', is_dual=buyin_months>0) %>%
+    left_join(regions, by='state') %>%
+    mutate(year=year) %>%
+    left_join(cc, by='bene_id') %>%
+    left_join(n_op_days, by='bene_id') %>% replace_na(list(n_op_days=0)) %>%
+    left_join(n_claims, by='bene_id') %>% replace_na(list(n_claims=0)) %>%
+    left_join(dx_presence, by='bene_id') %>% replace_na(list(rc=FALSE, uti=FALSE))
+
+  list(bene=bene, pde=pde, dx=dx)
+}
+
+years = 2011:2014
+n_years = length(years)
+dat = lapply(years, load_data)
+
+bene = lapply(dat, function(df) df$bene) %>% bind_rows %>%
   group_by(bene_id) %>%
-  summarize(n_acute_rc=length(unique(date)))
+  mutate(n_years=n(),
+         start_age=min(age),
+         in_cohort=n_years==4) %>%
+  ungroup()
 
-total_pde = pde %>% count(bene_id) %>% rename(n_claims=n)
+pde = lapply(dat, function(df) df$pde) %>% bind_rows %>%
+  left_join(distinct(select(bene, bene_id, in_cohort)), by='bene_id')
 
-cc = read_feather(sprintf('../cc_%i.feather', year))
-
-bene = read_tsv(sprintf('../bene_%i.tsv', year)) %>%
-  mutate(is_female=sex=='female', is_white=race=='white') %>%
-  left_join(total_pde, by='bene_id') %>% replace_na(list(n_claims=0)) %>%
-  left_join(bene_dx, by='bene_id') %>% replace_na(list(acute_rc=FALSE, copd=FALSE, uti=FALSE)) %>%
-  left_join(select(cc, bene_id, COPD), by='bene_id') %>%
-  left_join(rc_dx, by='bene_id') %>% replace_na(list(n_acute_rc=0))
+dx = lapply(dat, function(df) df$dx) %>% bind_rows() %>%
+  left_join(distinct(select(bene, bene_id, in_cohort)), by='bene_id')
 ```
@@
 
-```{r}
-sex_icd %>%
-  left_join(dx_counts, by='code') %>%
+```{r dx_counts}
+dx %>%
+  filter(in_cohort) %>%
+  count(code) %>%
+  left_join(dx_codes, by='code') %>%
   arrange(diagnosis_type, desc(n)) %>%
   mutate(n=format(n, big.mark=',')) %>%
-  kable(align='lllr')
+  kable()
 ```
@@
   arrange(desc(n)) %>%
-  head(20)
-
-kable(top_abx)
+  head(10) %T>%
+  kable %$%
+  antibiotic
 ```
@@
 
-Looking at top drugs by sex.
-
 ```{r drugs_by_sex}
@@
   bene %>%
+    filter(in_cohort) %>%
     left_join(pde %>% filter(antibiotic==drug) %>% count(bene_id), by='bene_id') %>%
@@
     summarize(n_bene=n(),
+              n_using=sum(n>0),
               n_claims=sum(n),
@@
 
-ubs = lapply(top_abx$antibiotic, usage_by_sex) %>% bind_rows
+ubs = lapply(top_abx, usage_by_sex) %>% bind_rows
 ```
 
-```{r}
-denom = bene %>% count(sex) %>% rename(n_denom=n)
-ubs = pde %>%
-  select(bene_id, antibiotic) %>%
-  distinct() %>%
-  left_join(select(bene, bene_id, sex), by='bene_id') %>%
-  count(sex, antibiotic) %>%
+```{r healthcare_usage_by_sex}
+hc_by_sex = dx %>%
+  count(bene_id, diagnosis_type) %>%
   ungroup() %>%
-  rename(n_user=n) %>%
-  left_join(denom, by='sex') %>%
-  mutate(frac_using=n_user/n_denom) %>%
-  select(antibiotic, sex, frac_using) %>%
-  spread(sex, frac_using)
+  right_join(bene, by='bene_id') %>%
+  replace_na(list(n=0)) %>%
+  group_by(sex) %>%
+  summarize(total_n_days=sum(n),
+            mean_n_days=mean(n),
+            n_using=sum(n>0),
+            n_bene=n())
+
+hc_use_ratio = hc_by_sex %>% mutate(ratio=mean_n_days/min(mean_n_days)) %$% max(ratio)
 ```
+
+Looking at top drugs by sex.
+I compare total consumption and consumption normalized by healthcare usage.
+I measure healthcare usage as the number of days on which the beneficiary generated an outpatient (i.e., Carrier or Outpatient) claim.
+The ratio between the mean healthcare usage of the two sexes is `r hc_use_ratio`.
 
@@
 ubs %>%
-  mutate(highlight=antibiotic %in% c('azithromycin', 'ciprofloxacin', 'trimethoprim/sulfamethoxazole', 'fosfomycin', 'nitrofurantoin')) %>%
-  ggplot(aes(x=male, y=female, label=antibiotic, color=highlight)) +
+  select(drug, sex, mean_n_claims) %>%
+  mutate(mean_n_claims=mean_n_claims/n_years*1000) %>%
+  spread(sex, mean_n_claims) %>%
+  mutate(highlight=drug %in% c('azithromycin', 'ciprofloxacin', 'trimethoprim/sulfamethoxazole', 'fosfomycin', 'nitrofurantoin')) %>%
+  ggplot(aes(x=male, y=female, label=drug, color=highlight)) +
   scale_color_manual(values=c('black', 'red')) +
   geom_abline() +
-  geom_abline(slope=1.28, col='gray') +
+  geom_abline(slope=hc_use_ratio, col='gray') +
   geom_text() +
   geom_point() +
-  xlab('claims per 1,000 male beneficiaries') +
-  ylab('claims per 1,000 female beneficiaries') +
+  xlab('male claims per 1,000 male beneficiaries per year') +
+  ylab('female claims per 1,000 female beneficiaries per year') +
   theme_minimal() +
@@
 
-```{r drug_by_sex_plot2}
-total_plot_f = function(df, highlight=character(0)) {
-  df %>%
-    #mutate_at(vars(total, se), mpy) %>%
-    mutate(ci=se*1.96,
-           ci_low=total-ci,
-           ci_high=total+ci) %>%
-    select(drug, sex, total, ci_low, ci_high) %>%
-    gather('key', 'value', total, ci_low, ci_high) %>%
-    mutate(key=case_when(.$sex=='male' & .$key=='total' ~ 'x',
-                         .$sex=='male' & .$key=='ci_low' ~ 'xmin',
-                         .$sex=='male' & .$key=='ci_high' ~ 'xmax',
-                         .$sex=='female' & .$key=='total' ~ 'y',
-                         .$sex=='female' & .$key=='ci_low' ~ 'ymin',
-                         .$sex=='female' & .$key=='ci_high' ~ 'ymax')) %>%
-    select(-sex) %>%
-    spread(key, value) %>%
-    mutate(highlight=drug %in% highlight) %>%
-    ggplot(aes(x=x, xmin=xmin, xmax=xmax, y=y, ymin=ymin, ymax=ymax, label=drug, color=highlight)) +
-    geom_abline(col='gray') +
-    geom_point() + geom_errorbar() + geom_errorbarh() +
-    scale_color_manual(values=c('black', 'red')) +
-    geom_text() +
-    #geom_text(nudge_x=1, nudge_y=0.5) +
-    #scale_x_continuous(breaks=c(0, 5, 10), limits=c(0, 12)) +
-    #scale_y_continuous(breaks=c(0, 5, 10), limits=c(0, 14)) +
-    xlab('No. rx. for men (millions)') + ylab('No. rx. for women (millions)') +
-    coord_fixed() +
-    theme_minimal()
-}
-
-ubs %>%
-  rename(total=mean_n_claims, se=se_n_claims) %>%
-  total_plot_f()
-```
-
-Alternatively, I look at the mean usage (across men and women), the residual (the difference between the fraction of female beneficiaries taking the drug and male beneficiaries), and the fractional residual (the residual over the mean).
-I keep only those drugs with mean fraction at least 1%, and I also keep fosfomycin.
-
-```{r}
-ubs %>%
-  mutate(mean_fraction=0.5*(male+female),
-         residual=female-male,
-         fractional_residual=residual/mean_fraction) %>%
-  filter(mean_fraction > 0.01 | antibiotic=='fosfomycin') %>%
-  ggplot(aes(x=mean_fraction, y=fractional_residual, label=antibiotic)) +
-  geom_point() +
-  geom_text()
-```
-
-This makes the same four look interesting; maybe levofloxacin and metronidazole would be next. Levo fits in with the ARTI story; metro does not.
+The $N$ is so big that the standard errors are so small that I don't even bother to put them on there.
 
@@
 
-```{r}
-outcome_bene = pde %>% filter(antibiotic=='ciprofloxacin') %$% bene_id %>% unique
-epi = bene %>%
-  mutate(outcome=bene_id %in% outcome_bene) %>%
-  count(sex, uti, outcome) %>%
-  ungroup()
+```{r cipro}
+cipro_dat = pde %>%
+  filter(in_cohort, antibiotic=='ciprofloxacin') %>%
+  select(bene_id, year, cipro=n_claims) %>%
+  right_join(filter(bene, in_cohort), by=c('bene_id', 'year')) %>%
+  replace_na(list(cipro=0)) %>%
+  group_by(bene_id) %>%
+  summarize(is_female=any(is_female),
+            uti=any(uti),
+            cipro=sum(cipro))
+
+model_f = function(df, f) glm(f, data=df, family='poisson') %>% tidy
+
+cipro_results = bind_rows(model_f(cipro_dat, cipro ~ is_female) %>% mutate(model='base'),
+                          model_f(cipro_dat, cipro ~ is_female + uti) %>% mutate(model='expl'))
 ```
 
-```{r, eval=FALSE}
-# group_by_all_ = function(df, group_name) {
-#   bind_rows(df, mutate_(df, .dots=setNames(list(~'all'), group_name))) %>%
-#     group_by_(group_name)
-# }
-# 
-# epi %>%
-#   group_by_all_('sex') %>%
-#   summarize(n_cases=sum(n)) %>%
+```{r cipro_results}
+# fraction of risk ratio explained
+frre = function(res, tm='is_femaleTRUE') {
+  rr0 = res %>% filter(model=='base', term==tm) %$% exp(estimate)
+  rr1 = res %>% filter(model=='expl', term==tm) %$% exp(estimate)
+  (rr0 - rr1) / (rr0 - 1)
+}
 
+pct_show = function(x) sprintf('%g%%', round(100*x, 2))
+
+cipro_results %>% kable
 ```
+
+I ran two models.
+In the first, I try to predict the number of ciprofloxacin claims that a beneficiary in the cohort will have over the entire study period using just their sex.
+Next, I use their sex and whether or not they had a diagnosis for a UTI at any point during the study period.
+
+Adding UTI to the model explained `r pct_show(frre(cipro_results))` of the increased mean ciprofloxacin consumption associated with being female.
+
+### Bactrim
+
+```{r bactrim}
+bactrim_dat = pde %>%
+  filter(in_cohort, antibiotic=='trimethoprim/sulfamethoxazole') %>%
+  select(bene_id, year, bactrim=n_claims) %>%
+  right_join(filter(bene, in_cohort), by=c('bene_id', 'year')) %>%
+  replace_na(list(bactrim=0)) %>%
+  group_by(bene_id) %>%
+  summarize(is_female=any(is_female),
+            uti=any(uti),
+            bactrim=sum(bactrim))
+
+bactrim_results = bind_rows(model_f(bactrim_dat, bactrim ~ is_female) %>% mutate(model='base'),
+                            model_f(bactrim_dat, bactrim ~ is_female + uti) %>% mutate(model='expl'))
+
+bactrim_results %>% kable
+```
+
+So, when adding in UTIs, the effect of sex appears to invert, meaning that UTIs explained *more* than all of the difference.
+
+### Nitro
+
+```{r nitro}
+nitro_dat = pde %>%
+  filter(in_cohort, antibiotic=='nitrofurantoin') %>%
+  select(bene_id, year, n=n_claims) %>%
+  right_join(filter(bene, in_cohort), by=c('bene_id', 'year')) %>%
+  replace_na(list(n=0)) %>%
+  group_by(bene_id) %>%
+  summarize(is_female=any(is_female),
+            uti=any(uti),
+            n=sum(n))
+
+nitro_results = bind_rows(model_f(nitro_dat, n ~ is_female) %>% mutate(model='base'),
+                          model_f(nitro_dat, n ~ is_female + uti) %>% mutate(model='expl'))
+
+nitro_results %>% kable
+```
+
+The nitrofurantoin model doesn't explain as much of the difference (`r pct_show(frre(nitro_results))`), which is maybe not surprising or troubling because nitrofurantoin is mainly prescribed to women.
+We're not confused about why this consumption is higher.
 
@@
 ```{r women_azithro_vs_uai, eval=FALSE}
-any_in = function(lst, elts) match(elts, lst) %>% { .[!is.na(.)] } %>% { length(.) > 0 }
+azithro_dat = pde %>%
+  filter(in_cohort, antibiotic=='azithromycin') %>%
+  select(bene_id, year, n=n_claims) %>%
+  right_join(filter(bene, in_cohort), by=c('bene_id', 'year')) %>%
+  replace_na(list(n=0)) %>%
+  group_by(bene_id) %>%
+  summarize(is_female=any(is_female),
+            rc=any(rc),
+            copd=any(COPD),
+            n=sum(n))
 
-usage %>%
-  filter(age >= 65, sex=='female') %>%
-  group_by(bene_id) %>%
-  mutate(used_azithro='azithromycin' %in% antibiotic,
-         used_uai=any_in(antibiotic, c('nitrofurantoin', 'trimethoprim/sulfamethoxazole', 'ciprofloxacin')),
-         used_both=used_azithro && used_uai,
-         is_white=race=='white') %>%
-  ungroup() %>%
-  mutate(user_type=case_when(.$used_both ~ 'both',
-                             .$used_azithro ~ 'azithro',
-                             .$used_uai ~ 'UAI',
-                             TRUE ~ 'none')) %>%
-  group_by(bene_id) %>%
-  summarize(age=age[1],
-            is_white=is_white[1],
-            chronic=chronic[1],
-            user_type=user_type[1],
-            n_claims=sum(n_claims)) %>%
-  group_by(user_type) %>%
-  summarize(n_beneficiaries=n(),
-            mean_age=mean(age),
-            percent_white=mean(is_white)*100,
-            mean_n_chronic_cond=mean(chronic),
-            mean_n_claims=mean(n_claims)) %>%
-  arrange(desc(n_beneficiaries)) %>%
-  kable
+azithro_results = bind_rows(model_f(azithro_dat, n ~ is_female) %>% mutate(model='base'),
+                            model_f(azithro_dat, n ~ is_female + rc + copd) %>% mutate(model='expl'))
+
+azithro_results %>% kable
 ```
+
+But we do a bad job: `r pct_show(frre(azithro_results))` gets explained by COPD and respiratory complaints.
 
@@
 - women have more encounters with doctors and therefore are prescribed more?
-
-### Chronic conditions
-
-Cf. the chronic conditions report.
 

### 2017-07-03 ms2/data/collect-data.R

@@ -8,21 +8,22 @@ fraction_nonzero = function(x) length(nonzero(x)) / length(x)
 pad0 = function(x, to) c(x, rep(0, to - length(x)))
 
 hoover = function(x) 0.5 * sum(abs(x - mean(x))) / sum(x)
+hoover2 = function(x) 0.5 * sum(abs(x[x > 0] - mean(x))) / sum(x)
 
 inequalities = function(x) {
   nb_par = fitdistrplus::mledist(x, 'nbinom')$estimate
   nb_size = nb_par['size']
   nb_prob = nb_size / (nb_size + nb_par['mu'])
-  data_frame(total=sum(x),
+  data_frame(n=n(),
+             total=sum(x),
              mean=mean(x),
              fnz=fraction_nonzero(x),
              gini=ineq(x, type='Gini'),
              nzgini=ineq(nonzero(x), type='Gini'),
              nb_size=nb_size,
              nb_prob=nb_prob,
-             pietra=ineq(x, type='RS'),
-             nzpieta=ineq(nonzero(x), type='RS'),
              hoover=hoover(x),
+             hoover2=hoover2(x),
              nzhoover=hoover(nonzero(x))
              )
 }

### 2017-07-05 sex-difference/sex-difference.Rmd
clean up sex

@@
 
-```{r load_data}
+```{r load_data, cache=FALSE}
 sum_over = function(target_columns) {
@@
 
-Does that mean that:
-
-- women are older and therefore use more azithromycin,
-- women have more chronic conditions (esp. COPD) and therefore use more, or
-- women have more encounters with doctors and therefore are prescribed more?
-
-### Encounters by sex
-
-Do women have more encounters than men? Does that explain their increased usage?
-
-```{r load_carrier_data, eval=FALSE}
-carrier = read_tsv('../../data/carrier_count_2011.tsv') %>%
-  rename(bene_id=BENE_ID, n_carrier_dates=n_from_dates)
-
-inpatient = read_tsv('../../data/inpatient_count_2011.tsv') %>%
-  rename(bene_id=BENE_ID, n_inpatient_dates=n_from_dates)
-
-outpatient = read_tsv('../../data/outpatient_count_2011.tsv') %>%
-  rename(bene_id=BENE_ID, n_outpatient_dates=n_from_dates)
-
-denom = usage %>% select(bene_id, sex, age) %>% distinct
-```
-
-```{r azithromycin_usage, eval=FALSE}
-na_to_zero = function(x) if_else(is.na(x), 0L, x)
-
-azi_usage = usage %>%
-  filter(antibiotic=='azithromycin') %>%
-  group_by(bene_id) %>%
-  summarize(n_claims=sum(n_claims)) %>%
-  left_join(denom, by='bene_id') %>%
-  filter(age >= 65, sex %in% c('male', 'female')) %>%
-  left_join(carrier, by='bene_id') %>%
-  left_join(inpatient, by='bene_id') %>%
-  left_join(outpatient, by='bene_id') %>%
-  mutate(age_group=bounds_group(age, c(65, 75, 85, 95))) %>%
-  mutate_at(vars(n_claims, n_carrier_dates, n_inpatient_dates, n_outpatient_dates), na_to_zero)
-```
-
-```{r azithromycin_usage_table, eval=FALSE}
-azi_usage %>%
-  group_by(age_group, sex) %>%
-  summarize(n_beneficiaries=n(),
-            mean_n_azithromycin_claims=mean(n_claims),
-            med_n_carrier_dates=median(n_carrier_dates),
-            med_n_outpatient_dates=median(n_outpatient_dates),
-            med_n_inpatient_dates=median(n_inpatient_dates)) %>%
-  kable
-```
-
-And I'll do a Poisson regression of number of azithromycin claims against age, being female, and number of carrier dates:
-
-```{r azithromycin_usage_model, eval=FALSE}
-library(broom)
-azi_usage %>%
-  mutate(is_female=sex=='female') %>%
-  glm(n_claims ~ age + is_female + n_carrier_dates + n_inpatient_dates + n_outpatient_dates, data=., family='poisson') %>%
-  tidy %>%
-  mutate(pct.increase=round((exp(estimate)-1.0)*100, 2)) %>%
-  kable
-```
-
-The coefficient for being female (0.05) accounts for almost the entire difference in consumption between men and women (56 per thousand). Note that age has an approximately zero correlation (i.e.,
-
 # Supplement
@@
 
-```{r consumption_by_age_and_sex, eval=FALSE}
+```{r consumption_by_age_and_sex}
 bene %>%
-  left_join(total_pde, by='bene_id') %>%
-  replace_na(list(n_claims=0)) %>%
-  group_by(sex, age) %>%
-  summarize(mean_n_claims=mean(n_claims)) %>%
-  ungroup() %>%
-  ggplot(aes(x=age, y=mean_n_claims, color=sex)) +
-  geom_point()
+  filter(in_cohort) %>%
+  group_by(bene_id) %>%
+  summarize(is_female=any(is_female),
+            n=sum(n_claims),
+            age=min(age)) %>%
+  ggplot(aes(x=age, y=n, group=is_female)) +
+  geom_boxplot()
 ```
@@
 
-The older people are more female: age 65 is about 60% female.
-This stays constant around 60% (or dips slightly) until age 75, when it steadily increases up to 85-90% female at age 100.
-So we'll need to account for this age difference.
+The older people are more female.
 
-```{r sex_by_age, eval=FALSE}
-plot_dat = usage %>%
-  select(age, sex) %>%
-  filter(sex %in% c('male', 'female')) %>%
-  group_by(age) %>% count(sex) %>% ungroup() %>%
-  spread(sex, n, fill=0L) %>%
-  mutate(total=female+male, grand_total=sum(total),
-         fraction_year=total/grand_total, fraction_female=female/total)
-
-ggplot(mapping=aes(x=age, y=fraction)) +
-  geom_point(data=rename(plot_dat, fraction=fraction_female)) +
-  geom_line(data=mutate(plot_dat, fraction=fraction_year*10), col='red') +
-  ylab('fraction female') + ylim(0, 1) +
-  theme_minimal()
+```{r sex_by_age}
+bene %>% filter(in_cohort) %>%
+  ggplot(aes(x=age, fill=sex)) +
+  geom_bar(position='dodge')
 ```
@@
 Then we could explain the increased consumption by a small number of people taking more drug.
-(Yonatan points to Lady Windmere syndrome.) In fact, consumption is more even among women.
-As always, this is driven by the extensive fraction: more women took at least one dose:
+(Yonatan points to Lady Windmere syndrome.) In fact, consumption is more even among women; this is driven by the extensive fraction: a greater fraction of women (than men) took at least one dose.
 
-```{r, eval=FALSE}
+```{r azithro_even}
 pde %>%
-  filter(drug=='azithromycin') %>%
-  count(bene_id) %>% rename(n=n_azi) %>%
-  right_join(bene, by='bene_id') %>% replace_na(list(n_azi=0)) %>%
-  count(sex, n_azi) %>%
+  filter(antibiotic=='azithromycin') %>%
+  group_by(bene_id) %>% summarize(azi=sum(n_claims)) %>%
+  right_join(bene %>% filter(in_cohort) %>% select(bene_id, sex) %>% distinct,
+    by='bene_id') %>% replace_na(list(azi=0)) %>%
+  count(sex, azi) %>%
   group_by(sex) %>%
   mutate(f_ppl=n/sum(n), cum_f_ppl=cumsum(f_ppl),
-         cum_cons=cumsum(n_azi*n), cum_f_cons=cum_cons/max(cum_cons)) %>%
+         cum_cons=cumsum(azi*n), cum_f_cons=cum_cons/max(cum_cons)) %>%
   ggplot(aes(x=cum_f_ppl, y=cum_f_cons, color=sex)) +

### 2017-07-05 sex-difference/sex-medicare-display.Rmd
initial

--- 2017-07-05 sex-difference/sex-difference.Rmd
@@
 
-```{r load_data, cache=FALSE}
-sum_over = function(target_columns) {
-  dummy_vars = sapply(1:length(target_columns), function(i) sprintf('x%i', i))
-  dummy_vars_string = paste0(dummy_vars, collapse=',')
-  dummy_vars_list = lapply(target_columns, as.name) %>% setNames(dummy_vars)
-  f = as.formula(paste0(c('~sum(', dummy_vars_string, ')'), collapse=''))
-  lazyeval::interp(f, .values=dummy_vars_list)
-}
-
-regions = read_tsv('../../db/census-regions/census-regions.tsv') %>%
-  select(state, region)
-
-condition_names = read_tsv('../chronic-conditions/names.tsv') %>% filter(condition_code != 'ALZH')
-names_1yr_ccs = condition_names %>% filter(condition_ref_years==1) %$% condition_code
-
-dx_codes = read_tsv('../sex-difference/sex_codes.tsv')
-
-load_data = function(year) {
-  # pde data
-  pde = read_tsv(sprintf('../pde_%i.tsv', year)) %>%
-    select(bene_id, antibiotic, days_supply) %>%
-    group_by(bene_id, antibiotic) %>%
-    summarize(n_claims=n(), days_supply=sum(days_supply)) %>%
-    ungroup() %>%
-    mutate(year=year)
-
-  # diagnosis claims
-  dx_car_fn = sprintf('../../data/dx_car_claims_%i.tsv', year)
-  dx_op_fn = sprintf('../../data/dx_op_claims_%i.tsv', year)
-
-  dx = bind_rows(read_tsv(dx_car_fn),
-                 read_tsv(dx_op_fn)) %>%
-    mutate(date=dmy(from_date)) %>%
-    rename(bene_id=BENE_ID, code=diagnosis) %>%
-    left_join(dx_codes, by='code')
-
-  dx_presence = dx %>%
-    group_by(bene_id) %>%
-    summarize(rc='acute_rc' %in% diagnosis_type,
-              uti='uti' %in% diagnosis_type)
-
-  count_car_fn = sprintf('../../data/count_car_claims_%i.tsv', year)
-  count_op_fn = sprintf('../../data/count_op_claims_%i.tsv', year)
-  n_op_days = bind_rows(read_tsv(count_car_fn), read_tsv(count_op_fn)) %>%
-    rename(bene_id=BENE_ID) %>%
-    group_by(bene_id) %>%
-    summarize(n_op_days=sum(n_from_date))
-
-  # total numbers of PDEs
-  n_claims = pde %>% group_by(bene_id) %>% summarize(n_claims=sum(n_claims))
-
-  # count chronic conditions. 'n_cc' means one-year cc's
-  cc = read_feather(sprintf('../cc_%i.feather', year)) %>%
-    rowwise() %>%
-    mutate_(n_cc=sum_over(names_1yr_ccs)) %>%
-    ungroup()
-
-  # join the cc, summary PDE, and dx data into bene
-  bene = read_tsv(sprintf('../bene_%i.tsv', year)) %>%
-    mutate(is_female=sex=='female', is_white=race=='white', is_dual=buyin_months>0) %>%
-    left_join(regions, by='state') %>%
-    mutate(year=year) %>%
-    left_join(cc, by='bene_id') %>%
-    left_join(n_op_days, by='bene_id') %>% replace_na(list(n_op_days=0)) %>%
-    left_join(n_claims, by='bene_id') %>% replace_na(list(n_claims=0)) %>%
-    left_join(dx_presence, by='bene_id') %>% replace_na(list(rc=FALSE, uti=FALSE))
-
-  list(bene=bene, pde=pde, dx=dx)
-}
-
+```{r load_data}
 years = 2011:2014
 n_years = length(years)
-dat = lapply(years, load_data)
 
-bene = lapply(dat, function(df) df$bene) %>% bind_rows %>%
-  group_by(bene_id) %>%
-  mutate(n_years=n(),
-         start_age=min(age),
-         in_cohort=n_years==4) %>%
-  ungroup()
-
-pde = lapply(dat, function(df) df$pde) %>% bind_rows %>%
-  left_join(distinct(select(bene, bene_id, in_cohort)), by='bene_id')
-
-dx = lapply(dat, function(df) df$dx) %>% bind_rows() %>%
-  left_join(distinct(select(bene, bene_id, in_cohort)), by='bene_id')
+dx_codes = read_tsv('sex_codes.tsv')
 ```
@@
 ```{r dx_counts}
-dx %>%
-  filter(in_cohort) %>%
-  count(code) %>%
+read_tsv('data/tbl_dx_counts.tsv') %>%
   left_join(dx_codes, by='code') %>%
@@
 ```{r top_drugs}
-top_abx = pde %>%
-  count(antibiotic) %>%
+top_abx = read_tsv('data/tbl_top_abx.tsv') %>%
   arrange(desc(n)) %>%
@@
 
-```{r drugs_by_sex}
-usage_by_sex = function(drug) {
-  bene %>%
-    filter(in_cohort) %>%
-    left_join(pde %>% filter(antibiotic==drug) %>% count(bene_id), by='bene_id') %>%
-    replace_na(list(n=0)) %>%
-    group_by(sex) %>%
-    summarize(n_bene=n(),
-              n_using=sum(n>0),
-              n_claims=sum(n),
-              mean_n_claims=mean(n),
-              sd_n_claims=sd(n)) %>%
-    mutate(se_n_claims=sd_n_claims/sqrt(n_bene),
-           drug=drug)
-}
+```{r usage_by_sex}
+# drugs by sex
+ubs = read_tsv('data/tbl_usage_by_sex.tsv')
 
-ubs = lapply(top_abx, usage_by_sex) %>% bind_rows
-```
-
-```{r healthcare_usage_by_sex}
-hc_by_sex = dx %>%
-  count(bene_id, diagnosis_type) %>%
-  ungroup() %>%
-  right_join(bene, by='bene_id') %>%
-  replace_na(list(n=0)) %>%
-  group_by(sex) %>%
-  summarize(total_n_days=sum(n),
-            mean_n_days=mean(n),
-            n_using=sum(n>0),
-            n_bene=n())
-
+# healthcare usage by sex
+hc_by_sex = read_tsv('tbl_hc_by_sex.tsv')
 hc_use_ratio = hc_by_sex %>% mutate(ratio=mean_n_days/min(mean_n_days)) %$% max(ratio)
@@
 ```{r cipro}
-cipro_dat = pde %>%
-  filter(in_cohort, antibiotic=='ciprofloxacin') %>%
-  select(bene_id, year, cipro=n_claims) %>%
-  right_join(filter(bene, in_cohort), by=c('bene_id', 'year')) %>%
-  replace_na(list(cipro=0)) %>%
-  group_by(bene_id) %>%
-  summarize(is_female=any(is_female),
-            uti=any(uti),
-            cipro=sum(cipro))
+cipro_results = read_tsv('data/tbl_model_cipro.tsv')
 
-model_f = function(df, f) glm(f, data=df, family='poisson') %>% tidy
-
-cipro_results = bind_rows(model_f(cipro_dat, cipro ~ is_female) %>% mutate(model='base'),
-                          model_f(cipro_dat, cipro ~ is_female + uti) %>% mutate(model='expl'))
-```
-
-```{r cipro_results}
 # fraction of risk ratio explained
@@
 ```{r bactrim}
-bactrim_dat = pde %>%
-  filter(in_cohort, antibiotic=='trimethoprim/sulfamethoxazole') %>%
-  select(bene_id, year, bactrim=n_claims) %>%
-  right_join(filter(bene, in_cohort), by=c('bene_id', 'year')) %>%
-  replace_na(list(bactrim=0)) %>%
-  group_by(bene_id) %>%
-  summarize(is_female=any(is_female),
-            uti=any(uti),
-            bactrim=sum(bactrim))
-
-bactrim_results = bind_rows(model_f(bactrim_dat, bactrim ~ is_female) %>% mutate(model='base'),
-                            model_f(bactrim_dat, bactrim ~ is_female + uti) %>% mutate(model='expl'))
-
+bactrim_results = read_tsv('data/tbl_model_tmpsmx.tsv')
 bactrim_results %>% kable
@@
 ```{r nitro}
-nitro_dat = pde %>%
-  filter(in_cohort, antibiotic=='nitrofurantoin') %>%
-  select(bene_id, year, n=n_claims) %>%
-  right_join(filter(bene, in_cohort), by=c('bene_id', 'year')) %>%
-  replace_na(list(n=0)) %>%
-  group_by(bene_id) %>%
-  summarize(is_female=any(is_female),
-            uti=any(uti),
-            n=sum(n))
-
-nitro_results = bind_rows(model_f(nitro_dat, n ~ is_female) %>% mutate(model='base'),
-                          model_f(nitro_dat, n ~ is_female + uti) %>% mutate(model='expl'))
-
+nitro_results = read_tsv('data/tbl_model_nitro.tsv')
 nitro_results %>% kable
@@
 
-```{r women_azithro_vs_uai, eval=FALSE}
-azithro_dat = pde %>%
-  filter(in_cohort, antibiotic=='azithromycin') %>%
-  select(bene_id, year, n=n_claims) %>%
-  right_join(filter(bene, in_cohort), by=c('bene_id', 'year')) %>%
-  replace_na(list(n=0)) %>%
-  group_by(bene_id) %>%
-  summarize(is_female=any(is_female),
-            rc=any(rc),
-            copd=any(COPD),
-            n=sum(n))
-
-azithro_results = bind_rows(model_f(azithro_dat, n ~ is_female) %>% mutate(model='base'),
-                            model_f(azithro_dat, n ~ is_female + rc + copd) %>% mutate(model='expl'))
-
+```{r women_azithro_vs_uai}
+azithro_results = read_tsv('data/tbl_model_azithro.tsv')
 azithro_results %>% kable
@@
 ```{r consumption_by_age_and_sex}
-bene %>%
-  filter(in_cohort) %>%
-  group_by(bene_id) %>%
-  summarize(is_female=any(is_female),
-            n=sum(n_claims),
-            age=min(age)) %>%
+read_tsv('data/tbl_consumption_age_sex.tsv') %>%
   ggplot(aes(x=age, y=n, group=is_female)) +
@@
 ```{r sex_by_age}
-bene %>% filter(in_cohort) %>%
-  ggplot(aes(x=age, fill=sex)) +
-  geom_bar(position='dodge')
+read_tsv('data/tbl_sex_by_age.tsv') %>%
+  ggplot(aes(x=age, y=n, fill=sex)) +
+  geom_bar(stat='identity', position='dodge')
 ```
@@
 ```{r azithro_even}
-pde %>%
-  filter(antibiotic=='azithromycin') %>%
-  group_by(bene_id) %>% summarize(azi=sum(n_claims)) %>%
-  right_join(bene %>% filter(in_cohort) %>% select(bene_id, sex) %>% distinct,
-    by='bene_id') %>% replace_na(list(azi=0)) %>%
-  count(sex, azi) %>%
-  group_by(sex) %>%
-  mutate(f_ppl=n/sum(n), cum_f_ppl=cumsum(f_ppl),
-         cum_cons=cumsum(azi*n), cum_f_cons=cum_cons/max(cum_cons)) %>%
+read_tsv('data/tbl_azithro_evenness.tsv') %>%
   ggplot(aes(x=cum_f_ppl, y=cum_f_cons, color=sex)) +

### 2017-07-05 sex-difference/sex-namcs.Rmd
initial

--- 2017-06-28 sex-difference/namcs.Rmd
@@
   summarize(uti='uti' %in% diagnosis_type,
-            acne='acne' %in% diagnosis_type)
+            acne='acne' %in% diagnosis_type,
+            arc='acute_rc' %in% diagnosis_type)
 
@@
   left_join(dx_visit, by='visit_id') %>%
-  replace_na(list(uti=FALSE, acne=FALSE)) %>%
+  replace_na(list(uti=FALSE, acne=FALSE, arc=FALSE)) %>%
   left_join(n_abx, by='visit_id') %>% replace_na(list(n_abx=0)) %>%
@@
 
-I kept the drugs from azithromycin through clarithromycin, accounting for 80% of all prescriptions.
+I kept the drugs from azithromycin through cefdinir, but excluding clindamycin, accounting for 80% of all prescriptions.
+I exclude clindamycin because NAMCS doesn't have any information about route of administration.
+In some cases, the old IDs can be used to distinguish different form, but other times there is no hope.
+For example, in the 2011 data, there are 118 visits that include clindamycin.
+In 84 cases, it is just "clindamycin"; in 18, it is "Cleocin" (which can be oral, injected, or vaginal); in 9 cases, it is "clindamycin phosphate", which is normally topical, etc.
 
@@
 
-Overall, more of the visits were made by women.
-Women make `r round(female_visit_odds[2], 3)*100`% more visits than men.
-I'll therefore evaluate consumption by two standards: first, do women consume more antibiotics overall; second, do women consume more antibiotics *per visit*.
+Overall, more of the visits were made by women. Women make `r round(female_visit_odds[2]-1, 3)*100`% more visits than men.
+
+What fraction of these visits are due to UTIs? We can see what fraction of visits are made for UTIs:
+
+```{r}
+svytable(~sex+uti, adult_design) %>%
+  as_data_frame() %>%
+  mutate(n=mpy(n)) %>%
+  kable
+```
+
+So of the ~150 million extra visits per year made by women, a difference of ~5 million visits (or 3%) can be attributed to UTIs.
+
+Of the top 3 diagnoses (hypertension, gynecological exam, and observation of pregnancy), two are woman-specific.
+Those account for 25 million visits per year, or about one-sixth of the total.
+So the difference is diffuse across diagnoses.
+Very few antibiotics are prescribed during those visits, less than 1 million per year.
+
+Acute respiration complaints explain another ~5 million:
+
+```{r}
+svytable(~sex+arc, adult_design) %>%
+  as_data_frame() %>%
+  mutate(n=mpy(n)) %>%
+  kable
+```
 
 ## Consumption by drug
+
+I'll therefore evaluate consumption by two standards: first, do women consume more antibiotics overall; second, do women consume more antibiotics *per visit*.
 
@@
 
-## Clindamycin
-
-```{r clinda}
-m1 = svyglm(clinda~sex, design=adult_design, family='poisson')
-m2 = svyglm(clinda~sex+acne, design=adult_design, family='poisson')
-clinda_f_expl = 1 - exp(-m2$coefficients['sexmale']+m1$coefficients['sexmale'])
-```
-
-The only clindamycin-associated diagnosis that meets (maybe) the data validity criterion is acne (`7061-`).
-However, acne explains only a small fraction (`r 100*round(clinda_f_expl, digits=3)`%) of the difference between the sexes.
-
 ## Just among the elderly

### 2017-07-05 sex-difference/sex-medicare-work.R

@@ -1,32 +1,3 @@
----
-title: "Drivers of sex differences in consumption"
-author: "Scott Olesen"
-date: "2/24/2017"
-output: html_document
----
-
-```{r global_options, include=FALSE}
-knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/', dev=c('png', 'pdf'),
-                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE, autodep=TRUE)
-```
-
-# Background
-
-Why do females account for 25% more antibiotic claims per capita than males? I have two easy explanations:
-
-- More females are alive at later ages, and older people consume more antibiotics.
-- Females consume more of the same antibiotics.
-- Females consume different antibiotics.
-
-# Methods
-
-*Data*: Medicare 2011-2014 in the cohort.
-
-*Outcome*: Antibiotic claims overall and for the top 10 or so individual drugs.
-
-*Exposure*: A diagnosis with a UTI, COPD, or an acute RTI.
-
-```{r load_data, cache=FALSE}
 sum_over = function(target_columns) {
   dummy_vars = sapply(1:length(target_columns), function(i) sprintf('x%i', i))
   dummy_vars_string = paste0(dummy_vars, collapse=',')
@@ -51,22 +22,22 @@ load_data = function(year) {
     summarize(n_claims=n(), days_supply=sum(days_supply)) %>%
     ungroup() %>%
     mutate(year=year)
-  
+
   # diagnosis claims
   dx_car_fn = sprintf('../../data/dx_car_claims_%i.tsv', year)
   dx_op_fn = sprintf('../../data/dx_op_claims_%i.tsv', year)
-  
+
   dx = bind_rows(read_tsv(dx_car_fn),
                  read_tsv(dx_op_fn)) %>%
     mutate(date=dmy(from_date)) %>%
     rename(bene_id=BENE_ID, code=diagnosis) %>%
     left_join(dx_codes, by='code')
-  
+
   dx_presence = dx %>%
     group_by(bene_id) %>%
     summarize(rc='acute_rc' %in% diagnosis_type,
               uti='uti' %in% diagnosis_type)
-  
+
   count_car_fn = sprintf('../../data/count_car_claims_%i.tsv', year)
   count_op_fn = sprintf('../../data/count_op_claims_%i.tsv', year)
[…]

### 2017-07-05 sex-difference/sex-medicare-display.Rmd
sex medicare split

@@
 knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/', dev=c('png', 'pdf'),
-                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE, autodep=TRUE)
+                      echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE)
 ```
@@
 # healthcare usage by sex
-hc_by_sex = read_tsv('tbl_hc_by_sex.tsv')
+hc_by_sex = read_tsv('data/tbl_hc_by_sex.tsv')
 hc_use_ratio = hc_by_sex %>% mutate(ratio=mean_n_days/min(mean_n_days)) %$% max(ratio)
@@
 
-# fraction of risk ratio explained
-frre = function(res, tm='is_femaleTRUE') {
-  rr0 = res %>% filter(model=='base', term==tm) %$% exp(estimate)
-  rr1 = res %>% filter(model=='expl', term==tm) %$% exp(estimate)
-  (rr0 - rr1) / (rr0 - 1)
+# fraction of coefficient explained
+fce = function(res, tm='is_femaleTRUE') {
+  den = res %>% filter(model=='base', term==tm) %$% estimate
+  num = res %>% filter(model=='expl', term==tm) %$% estimate
+  1 - num/den
 }
@@
 
-Adding UTI to the model explained `r pct_show(frre(cipro_results))` of the increased mean ciprofloxacin consumption associated with being female.
+Adding UTI to the model explained `r pct_show(fce(cipro_results))` of the increased mean ciprofloxacin consumption associated with being female.
 
@@
 
-The nitrofurantoin model doesn't explain as much of the difference (`r pct_show(frre(nitro_results))`), which is maybe not surprising or troubling because nitrofurantoin is mainly prescribed to women.
+The nitrofurantoin model doesn't explain as much of the difference (`r pct_show(fce(nitro_results))`), which is maybe not surprising or troubling because nitrofurantoin is mainly prescribed to women.
 We're not confused about why this consumption is higher.
@@
 
-But we do a bad job: `r pct_show(frre(azithro_results))` gets explained by COPD and respiratory complaints.
+But we do a bad job: `r pct_show(fce(azithro_results))` gets explained by COPD and respiratory complaints.
 
@@
 read_tsv('data/tbl_consumption_age_sex.tsv') %>%
-  ggplot(aes(x=age, y=n, group=is_female)) +
+  ggplot(aes(x=factor(age), y=n, color=is_female)) +
   geom_boxplot()

### 2017-07-05 sex-difference/sex-medicare-work.R

@@ -94,9 +94,11 @@ dx %>%
   write_tsv('data/tbl_dx_counts.tsv')
 
 # save top antibiotics
-pde %>%
-  count(antibiotic) %>%
+all_abx = pde %>%
+  count(antibiotic) %T>%
   write_tsv('data/tbl_top_abx.tsv')
+
+top_abx = all_abx %>% arrange(desc(n)) %$% head(antibiotic, 10)
 
 # consumption of drugs by sex
 usage_by_sex = function(drug) {
@@ -114,11 +116,11 @@ usage_by_sex = function(drug) {
            drug=drug)
 }
 
-ubs = lapply(top_abx, usage_by_sex) %>% bind_rows
-write_tsv('data/tbl_usage_by_sex.tsv')
+lapply(top_abx, usage_by_sex) %>% bind_rows %>%
+  write_tsv('data/tbl_usage_by_sex.tsv')
 
 # healthcare_usage_by_sex
-hc_by_sex = dx %>%
+dx %>%
   count(bene_id, diagnosis_type) %>%
   ungroup() %>%
   right_join(bene, by='bene_id') %>%
@@ -127,8 +129,8 @@ hc_by_sex = dx %>%
   summarize(total_n_days=sum(n),
             mean_n_days=mean(n),
             n_using=sum(n>0),
-            n_bene=n())
-write_tsv('data/tbl_hc_by_sex.tsv')
+            n_bene=n()) %>%
+  write_tsv('data/tbl_hc_by_sex.tsv')
 
 cipro_dat = pde %>%
   filter(in_cohort, antibiotic=='ciprofloxacin') %>%

### 2017-07-05 sex-difference/sex-medicare-display.Rmd
sex medicare split complete

@@
 read_tsv('data/tbl_consumption_age_sex.tsv') %>%
-  ggplot(aes(x=factor(age), y=n, color=is_female)) +
-  geom_boxplot()
+  mutate(y=total_n_claims, ci=1.96*sem_n_claims, ymin=y-ci, ymax=y+ci) %>%
+  ggplot(aes(x=age, y=total_n_claims, ymin=ymin, ymax=ymax, color=sex)) +
+  geom_point() + geom_linerange()
 ```

### 2017-07-05 sex-difference/sex-medicare-work.R

@@ -196,9 +196,15 @@ write_tsv(azithro_results, 'data/tbl_model_azithro.tsv')
 bene %>%
   filter(in_cohort) %>%
   group_by(bene_id) %>%
-  summarize(is_female=any(is_female),
+  summarize(sex=if_else(any(is_female), 'female', 'male'),
             n=sum(n_claims),
             age=min(age)) %>%
+  group_by(age, sex) %>%
+  summarize(n_bene=n(),
+            total_n_claims=sum(n),
+            mean_n_claims=mean(n),
+            sd_n_claims=sd(n),
+            sem_n_claims=sd_n_claims/sqrt(n_bene)) %>%
   write_tsv('data/tbl_consumption_age_sex.tsv')
 
 # Sex by age

### 2017-07-07 ms1/namcs_trends.Rmd
namcs trends per capita

@@
 }
+
+census = read_tsv('../../../../db/census/census_totals.tsv')
 
@@
 
+## Trends in numbers of visits and number of Americans
+
+The number of visits went down during this period:
+
+```{r visits_census}
+design %>%
+  update(x=1) %>%
+  svyby(~x, ~year, design=., svytotal) %>%
+  as_data_frame() %>%
+  rename(n_visits=x) %>%
+  kable
+```
+
+But the number of Americans went up (but the number of children went down):
+
+```{r}
+census %>% kable
+```
+
 ## Summary of trends
 
+### Overall trends
+
+It looks by eye like the total number of antibiotic prescriptions per capita is going down:
+
+```{r overall}
+overall_trend = svyby(~n_abx, ~year, design=design, svytotal) %>%
+  as_data_frame() %>%
+  left_join(census, by='year') %>%
+  mutate(abx_per_capita=n_abx/n_total, se_per_capita=se/n_total)
+
+overall_trend %>%
+  mutate(ci=1.96*se_per_capita, ymin=abx_per_capita-ci, ymax=abx_per_capita+ci) %>%
+  ggplot(aes(x=year, y=abx_per_capita, ymin=ymin, ymax=ymax)) +
+  geom_point() + geom_line() + geom_linerange() +
+  theme_minimal()
+```
+
+There is no support for a trend per visit:
+
+```{r overall_model1}
+svyglm(n_abx ~ year, design=design, family=quasipoisson(link='identity')) %>%
+  tidy %>% kable
+```
+
+Nor for a trend per capita:
+
+```{r overall_model2}
+overall_trend %>%
+  mutate(year=year-min(year)) %>%
+  lm(abx_per_capita ~ year, weights=1.0/se_per_capita^2, data=.) %>%
+  tidy %>% kable
+```
+
+### Trends by drug
+
 ```{r drugs_time}
-f = function(design, y, d) {
-  subset(design, year==y) %>%
-    svytotal(as.formula(paste0('~', d)), design=.) %>%
-    data_frame(year=y, abx=d, total=., se=as.numeric(SE(.)))
+# apc = antibiotic per capita
+f = function(design, d) {
+  svyby(as.formula(paste0('~', d)), ~year, design=design, svytotal) %>%
+    as_data_frame() %>%
+    rename_(.dots=setNames(d, 'n_abx')) %>%
+    mutate(abx=d)
 }
@@
 g = function(design) {
-  crossing(year=2011:2014, desc=top_abx$abx) %>%
-    rowwise() %>% do(f(design, .$year, .$desc)) %>% ungroup()
+  lapply(top_abx$abx, function(x) f(design, x)) %>% bind_rows() %>%
+    left_join(census, by='year') %>%
+    mutate(apc=n_abx/n_total, se=se/n_total)
 }
@@
 abx_time %>%
-  mutate(ci=1.96*se, ymin=total-ci, ymax=total+ci) %>%
-  ggplot(aes(x=year, y=total, ymin=ymin, ymax=ymax)) +
+  mutate(ci=1.96*se, ymin=apc-ci, ymax=apc+ci) %>%
+  ggplot(aes(x=year, y=apc, ymin=ymin, ymax=ymax)) +
   facet_wrap(~abx, ncol=3, scales='free_y') +
@@
 
-It doesn't look by eye that any trends are particularly strong. I used Poisson regressions to see if anything was changing with time. Only levo:
+And the same plot again, but with fixed scales:
+
+```{r drugs_time_plot_fixed}
+abx_time %>%
+  mutate(ci=1.96*se, ymin=apc-ci, ymax=apc+ci) %>%
+  ggplot(aes(x=year, y=apc, ymin=ymin, ymax=ymax)) +
+  facet_wrap(~abx, ncol=3) +
+  geom_point() + geom_line() + geom_linerange() +
+  theme_minimal()
+```
+
+It doesn't look by eye that any trends are particularly strong.
+
+I used weighted linear regression to look for trends in totals:
+
+```{r drugs_time_totals_model}
+abx_total_models = abx_time %>%
+  group_by(abx) %>%
+  do(tidy(lm(apc ~ year, weight=1.0/se^2, data=.))) %>%
+  ungroup()
+
+abx_total_models %>%
+  filter(term=='year') %>% select(-term) %>%
+  arrange(estimate) %>%
+  kable
+
+abx_total_models %>%
+  filter(term=='year') %>% select(-term) %>%
+  mutate(ci=1.96*std.error, ymin=estimate-ci, ymax=estimate+ci) %>%
+  mutate(abx=forcats::fct_reorder(abx, estimate)) %>%
+  ggplot(aes(x=abx, y=estimate, ymin=ymin, ymax=ymax)) +
+  geom_point() + geom_linerange() +
+  coord_flip() +
+  theme_minimal()
+```
+
+I used Poisson regressions to see if anything was changing *per visit* with time. Only levo:
 
 ```{r drugs_time_model}
-qpglm = function(design, frmla) svyglm(frmla, design=design, family=quasipoisson(link='log'))
+qpglm = function(design, frmla) {
+  svyglm(frmla, design=design, family=quasipoisson(link='identity'))
+}
 
@@
 
-```{r levo_plot}
-american_population = data_frame(year=2011:2014, n_ppl=c(312.92, 315.22, 317.47, 319.85) * 1e6)
+### Levo trend
 
+```{r levo_ratio}
 levo_time = abx_time %>%
   filter(abx=='levo') %>%
-  left_join(american_population, by='year') %>%
-  mutate(total=total/n_ppl*1000, se=se/n_ppl*1000)
+  arrange(year)
 
-levo_time %>%
-  mutate(ci=1.96*se, ymin=total-ci, ymax=total+ci) %>%
-  ggplot(aes(x=year, y=total, ymin=ymin, ymax=ymax)) +
-  geom_point() + geom_line() + geom_linerange() +
-  ylim(5, 21) +
-  ylab('Levo claims per 1,000 Americans') +
-  theme_minimal()
-```
-
-```{r levo_ratio}
-t1 = levo_time %>% filter(year==2011) %$% total
+t1 = levo_time %>% filter(year==2011) %$% apc
 v1 = levo_time %>% filter(year==2011) %$% se
-t2 = levo_time %>% filter(year==2014) %$% total
+t2 = levo_time %>% filter(year==2014) %$% apc
 v2 = levo_time %>% filter(year==2014) %$% se
@@
 
-show_tot = function(tot, v) sprintf('%g (±%g)', round(tot, 2), round(v, 2))
+show_tot = function(tot, v) sprintf('%g (±%g)', round(tot, 4), round(v, 4))
 show_pct = function(tot, v) sprintf('%g%% (±%g%%)', round((tot - 1)*100, digits=1), round(sqrt(v)*100, digits=1))
@@
 
-Levo consumption increased from `r show_tot(t1, v1)` prescriptions per 1,000 Americans per year to `r show_tot(t2, v2)`. That is an increase of `r show_pct(tr, vr)`, i.e., doubled.
-
-## Presence of this trend in different age groups
-
-I can confirm that levo is the only drug with a strong trend among adults (rather than everyone):
-
-```{r drugs_time_model_adult_table}
-abx_time_model_adult %>%
-  select(abx, estimate, p.value) %>%
-  mutate(adj.p.value=p.adjust(p.value, method='BH')) %>%
-  arrange(p.value) %>%
-  kable
-```
-
-And, in fact, this trend exists just among the elderly, although it's only marginally significant if you tested for everything.
-
-```{r levo_elderly}
-design %>% subset(age_group=='elderly') %>%
-  qpglm(levo~year) %>% tidy
-```
-
-Levo may be on the decline for children, however:
-
-```{r levo_children}
-design %>% subset(age_group=='child') %>%
-  qpglm(levo~year) %>% tidy
-```
+Levo increased from `r show_tot(t1, v1)` to `r show_tot(t2, v2)` per capita, an increase of `r show_pct(tr, vr)`.
 

### 2017-07-07 ms1/namcs_trends.Rmd
cleanup namcs trend

@@
 
-g = function(design) {
+g = function(design, denom) {
   lapply(top_abx$abx, function(x) f(design, x)) %>% bind_rows() %>%
     left_join(census, by='year') %>%
-    mutate(apc=n_abx/n_total, se=se/n_total)
+    rename_(.dots=setNames(denom, 'denom')) %>%
+    mutate(apc=n_abx/denom, se=se/denom)
 }
 
-abx_time = g(design)
-abx_time_adult = g(adult_design)
+abx_time = g(design, 'n_total')
+abx_time_adult = g(adult_design, 'n_adult')
 ```
@@
 
-### Levo trend
+## Trends among the adults
+
+I do the same analysis, but just among adults. Weirdly, the results look a lot
+different. The confidence intervals get a lot bigger.
+
+```{r adult_trends}
+adult_models = abx_time_adult %>%
+  group_by(abx) %>%
+  do(tidy(lm(apc ~ year, weight=1.0/se^2, data=.))) %>%
+  ungroup()
+
+adult_models %>%
+  filter(term=='year') %>% select(-term) %>%
+  mutate(ci=1.96*std.error, ymin=estimate-ci, ymax=estimate+ci) %>%
+  mutate(abx=forcats::fct_reorder(abx, estimate)) %>%
+  ggplot(aes(x=abx, y=estimate, ymin=ymin, ymax=ymax)) +
+  geom_point() + geom_linerange() +
+  coord_flip() +
+  theme_minimal()
+```
+
+### Just levo
 
 ```{r levo_ratio}
-levo_time = abx_time %>%
+levo_time = abx_time_adult %>%
   filter(abx=='levo') %>%
@@
 
+```{r}
+levo_time %>%
+  select(year, abx_per_capita=apc, se) %>%
+  kable
+```
+
 Levo increased from `r show_tot(t1, v1)` to `r show_tot(t2, v2)` per capita, an increase of `r show_pct(tr, vr)`.

### 2017-07-20 sex-difference/sex-namcs.Rmd
sex diff totals
updates

@@
 
-sex_codes = read_tsv('namcs_sex_codes.tsv') %>%
-  rename(dx=code)
+dx_codes = read_tsv('../../data/fd_codes.tsv') %>%
+  rename(dx=code, desc=dx_desc)
+
+diagnosis_types = unique(dx_codes$diagnosis_type)
 
 dx = read_tsv_years('../../db/namcs/data/namcs_dx_%i.tsv') %>%
-  left_join(sex_codes, by='dx')
-
-dx_visit = dx %>%
-  group_by(visit_id) %>%
-  summarize(uti='uti' %in% diagnosis_type,
-            acne='acne' %in% diagnosis_type,
-            arc='acute_rc' %in% diagnosis_type)
+  left_join(dx_codes, by='dx')
 
@@
                              .$age <= 92 ~ 'elderly')) %>%
-  left_join(dx_visit, by='visit_id') %>%
-  replace_na(list(uti=FALSE, acne=FALSE, arc=FALSE)) %>%
   left_join(n_abx, by='visit_id') %>% replace_na(list(n_abx=0)) %>%
@@
 # Results
+
+## Diagnoses across the sexes
+
+What's the difference in primary diagnoses across the sexes? I look in not-children.
+
+```{r diagnoses}
+unique_primary_dx = dx %>%
+  filter(dx_number==1) %>%
+  left_join(visits, by=c('visit_id', 'year')) %>%
+  filter(age_group != 'child') %>%
+  count(dx, sort=TRUE) %>%
+  head(20) %$%
+  dx
+
+unique_primary_dx
+```
+
+```{r}
+f = function(code) {
+  filter(dx, dx_number==1, dx==code) %>%
+    mutate(n_dx=1) %>%
+    left_join(visits, by=c('visit_id', 'year')) %>%
+    replace_na(list(n_dx=0)) %>%
+    design_f() %>%
+    subset(age_group != 'child') %>%
+    svyby(~n_dx, ~sex, design=., svytotal) %>%
+    as_data_frame() %>%
+    mutate(code=code)
+}
+
+dx_by_sex = lapply(unique_primary_dx, f) %>% bind_rows()
+dx_by_sex
+```
+
+```{r}
+dx_by_sex %>%
+  rename(n=n_dx) %>%
+  mutate(ci=1.96*se, ci_low=n-ci, ci_high=n+ci) %>%
+  select(code, sex, n, ci_low, ci_high) %>%
+  gather('key', 'value', n:ci_high) %>%
+  mutate(key=case_when(.$sex=='male' & .$key=='n' ~ 'x',
+                       .$sex=='male' & .$key=='ci_low' ~ 'xmin',
+                       .$sex=='male' & .$key=='ci_high' ~ 'xmax',
+                       .$sex=='female' & .$key=='n' ~ 'y',
+                       .$sex=='female' & .$key=='ci_low' ~ 'ymin',
+                       .$sex=='female' & .$key=='ci_high' ~ 'ymax')) %>%
+  select(-sex) %>%
+  spread(key, value) %>%
+  ggplot(aes(x=x, xmin=xmin, xmax=xmax, y=y, ymin=ymin, ymax=ymax, label=code)) +
+  geom_abline(col='gray') +
+  geom_point() + geom_errorbar() + geom_errorbarh() + geom_text() +
+  coord_fixed() +
+  theme_minimal()
+```
+
+```{r}
+g = function(diagnosis_type) {
+  filter(dx, dx_number==1, diagnosis_type==diagnosis_type) %>%
+    mutate(n_dx=1) %>%
+    left_join(visits, by=c('visit_id', 'year')) %>%
+    replace_na(list(n_dx=0)) %>%
+    design_f() %>%
+    subset(age_group != 'child') %>%
+    svyby(~n_dx, ~sex, design=., svytotal) %>%
+    as_data_frame() %>%
+    mutate(diagnosis_type=diagnosis_type)
+}
+
+#dt_by_sex = lapply(diagnosis_types, f) %>% bind_rows()
+#dt_by_sex
+g('pneumonia')
+```
 
@@
 
-```{r}
+```{r, eval=FALSE}
 svytable(~sex+uti, adult_design) %>%
@@
 
-```{r}
+```{r, eval=FALSE}
 svytable(~sex+arc, adult_design) %>%
@@
   kable
+```
+
+## Overall consumption
+
+How many antibiotic prescriptions total were made by each sex?
+
+```{r overall_abx}
+svyby(~n_abx, ~sex, svytotal, design=adult_design)
 ```

### 2017-07-20 polish-bene-pde.R

@@ -1,6 +1,7 @@
 #!/usr/bin/env Rscript
 
 abx_class = read_tsv('abx_class.tsv')
+dx_class = read_tsv('../data/fd_codes.tsv') %>% select(diagnosis=code, diagnosis_type)
 
 polish_bene = function(df) {
   rename(df, bene_id=BENE_ID, age=AGE, buyin_months=BUYIN_MO)
@@ -11,6 +12,15 @@ polish_pde = function(df) {
     mutate(service_date=dmy(service_date)) %>%
     left_join(abx_class, by='generic_name') %>%
     select(-generic_name)
+}
+
+polish_dx = function(df) {
+  rename(df, bene_id=BENE_ID) %>%
+    mutate(from_date=dmy(from_date)) %>%
+    left_join(dx_class, by='diagnosis')
+    group_by(bene_id, diagnosis) %>%
+    summarize(n_dx_date=length(unique(from_date))) %>%
+    ungroup()
 }
 
 polish_year = function(y) {
@@ -27,6 +37,14 @@ polish_year = function(y) {
   read_tsv(bene_in_fn, col_types=cols(zipcode='c')) %>%
     polish_bene() %>%
     write_tsv(bene_out_fn)
+
+  car_in_fn = sprintf('../data/fd_dx_car_claims_%i.tsv', y)
+  op_in_fn = sprintf('../data/fd_dx_op_claims_%i.tsv', y)
+  dx_out_fn = sprintf('dx_%i.tsv', y)
+
+  bind_rows(read_tsv(car_in_fn), read_tsv(op_in_fn)) %>%
+    polish_dx() %>%
+    write_tsv(dx_out_fn)
 }
 
 for (year in 2011:2014) polish_year(year)

### 2017-07-21 polish-bene-pde.R
new dx polish

@@ -17,10 +17,10 @@ polish_pde = function(df) {
 polish_dx = function(df) {
   rename(df, bene_id=BENE_ID) %>%
     mutate(from_date=dmy(from_date)) %>%
+    group_by(bene_id, diagnosis, from_date) %>%
+    summarize(n_encounter_claim=n()) %>%
+    ungroup() %>%
     left_join(dx_class, by='diagnosis')
-    group_by(bene_id, diagnosis) %>%
-    summarize(n_dx_date=length(unique(from_date))) %>%
-    ungroup()
 }
 
 polish_year = function(y) {

### 2017-07-21 ms1/ms1_display.Rmd
streamlining ms1

@@
   model_ci(df) %>%
-    mutate_at(vars(ci_range, ci_low, ci_high, estimate), function(x) x * 1000) %>%
+    mutate_at(vars(ci_range, ci_low, ci_high, estimate), function(x) x*1000) %>%
     rename(cpkp=estimate)
@@
 
+# Diagnoses
+
+```{r}
+read_tsv('tables/model_abx.tsv') %>%
+  filter(term=='year') %>%
+  mutate(model=case_when(.$model=='trimethoprim/sulfamethoxazole' ~ 'TMP/SMX',
+                         .$model=='amoxicillin/clavulanate' ~ 'amox/clav',
+                         TRUE ~ .$model)) %>%
+  mutate(model=forcats::fct_reorder(model, estimate),
+         cpkp=1000*estimate) %>%
+  ggplot(aes(x=model, y=cpkp)) +
+  geom_point() +
+  geom_linerange(aes(x=model, ymin=0, ymax=cpkp)) +
+  coord_flip() +
+  xlab('') + ylab('Adjusted yearly change in claims per 1k bene. per yr') +
+  scale_y_continuous(limits=c(-16, 17), breaks=c(-15, -10, -5, 0, 5, 10, 15)) +
+  theme_minimal()
+```
+
 # Supplemental results

### 2017-07-21 ms1/ms1_work.R

@@ -1,7 +1,5 @@
 # Do all the heavy-lifting analyses so that the Rmd can just show the results
 
-import::from(broom, tidy)
-
 sum_over = function(target_columns) {
   dummy_vars = sapply(1:length(target_columns), function(i) sprintf('x%i', i))
   dummy_vars_string = paste0(dummy_vars, collapse=',')
@@ -16,6 +14,8 @@ sum_over = function(target_columns) {
 regions = read_tsv('../../db/census-regions/census-regions.tsv') %>%
   select(state, region)
 
+# Keep chronic conditions that have a single reference year. Exclude Alzheimer's,
+# which is a subset of Alz. & dementia
 condition_names = read_tsv('../chronic-conditions/names.tsv') %>% filter(condition_code != 'ALZH')
 names_1yr_ccs = condition_names %>% filter(condition_ref_years==1) %$% condition_code
 
@@ -24,29 +24,31 @@ dx_codes = read_tsv('../../data/fd_codes.tsv') %>%
 
 dx_types = unique(dx_codes$diagnosis_type)
 
+# window of days that you can look ahead from an rx to a dx
+rxdx_window = 3
+
 load_data = function(year) {
   # pde data
   pde = read_tsv(sprintf('../pde_%i.tsv', year)) %>%
-    select(bene_id, antibiotic, days_supply) %>%
-    group_by(bene_id, antibiotic) %>%
-    summarize(n_claims=n(), days_supply=sum(days_supply)) %>%
-    ungroup() %>%
-    mutate(year=year)
+    select(bene_id, service_date, antibiotic) %>%
+    distinct() %>%
+    mutate(year=year, pde_id=1:n())
 
   # diagnosis claims
-  dx_fn = sprintf('../dx_%i.tsv', year)
+  dx = read_tsv(sprintf('../dx_%i.tsv', year)) %>%
+    mutate(year=year)
   
-  dx = read_tsv(dx_fn) %>%
-    rename(code=diagnosis) %>%
-    filter(n_dx_date > 0) %>%
-    left_join(dx_codes, by='code') %>%
-    select(bene_id, diagnosis_type) %>%
-    distinct() %>%
-    mutate(dummy=TRUE) %>%
-    spread(diagnosis_type, dummy)
+  # associated each PDE with a diagnosis, if possible
+  pde_dx = pde %>%
[…]

### 2017-07-21 ms1/sankey/make_sankey_dat.py

@@ -1,37 +1,0 @@
-#!/usr/bin/env python3
-
-years = [str(x) for x in range(2011, 2014+1)]
-
-dat = {}
-with open('tbl_cohorts.tsv') as f:
-    header = next(f)
-    for line in f:
-        year_list, n = line.rstrip().split('\t')
-        dat[year_list] = int(n)
-
-n_bene = sum([dat[x] for x in dat])
-
-with open('node.tsv', 'w') as f:
-    def myprint(*x):
-        print(*x, sep='\t', file=f)
-
-    myprint('ID', 'x', 'y')
-    for i, year in enumerate(years):
-        myprint(year + 'in', i, 0)
-        myprint(year + 'out', i, 1)
-
-with open('edge.tsv', 'w') as f:
-    def myprint(*x):
-        print(*x, sep='\t', file=f)
-
-    myprint('ID', 'N1', 'N2', 'Value')
-    for year, next_year in zip(years, years[1:]):
-        stay_in = sum([dat[x] for x in dat if year in x and next_year in x])
-        leave = sum([dat[x] for x in dat if year in x and next_year not in x])
-        enter = sum([dat[x] for x in dat if year not in x and next_year in x])
-        stay_out = n_bene - (stay_in + leave + enter)
-
-        myprint(year + 'stayin', year + 'in', next_year + 'in', stay_in)
-        myprint(year + 'leave', year + 'in', next_year + 'out', leave)
-        myprint(year + 'enter', year + 'out', next_year + 'in', enter)
-        myprint(year + 'stayout', year + 'out', next_year + 'out', stay_out)

### 2017-07-21 ms1/sankey/sankey_plot.R

@@ -1,6 +1,0 @@
-edges = read_tsv('edge.tsv') %>% as.data.frame
-nodes = read_tsv('node.tsv') %>% as.data.frame
-
-pdf(file='sankey.pdf')
-riverplot::makeRiver(nodes=nodes, edges=edges) %>% plot
-dev.off()

### 2017-07-24 ms1/ms1_display.Rmd
streamlining

@@
 }
-
-trend_p = function(x) t.test(diff(x))$p.value
 ```
@@
 
-```{r}
+```{r, eval=FALSE}
 model_ci = function(df) {
@@
 ```{r}
-totals_dat = read_tsv('tables/tbl_cohort_totals.tsv') %>%
+totals_dat = read_tsv('tables/totals.tsv') %>%
   pctize('n_bene') %>%
@@
 ```{r}
-linear_trend_p = function(x) tidy(t.test(diff(x)))$p.value
+trend_p = function(x) tidy(t.test(diff(x)))$p.value
 
@@
   group_by(var) %>%
-  summarize(p=linear_trend_p(val)) %>%
+  summarize(p=trend_p(val)) %>%
   mutate(adj.p=p.adjust(p, method='BH'),
@@
 ```{r}
-cohort_dat = read_tsv('tables/tbl_cohort_claims_by_abx.tsv')
+claims_by_abx = read_tsv('tables/claims_by_abx.tsv')
 
-top_abx = cohort_dat %>%
+top_abx = claims_by_abx %>%
   group_by(antibiotic) %>%
@@
 ```{r}
-cohort_dat %>%
+claims_by_abx %>%
   top_abx_f() %>%
@@
   theme_minimal() +
-  ggtitle('Unadjusted consumption, in cohort')
+  ggtitle('Unadjusted consumption')
 
-cohort_dat %>%
+claims_by_abx %>%
+  filter(antibiotic %in% top_abx) %>%
+  mutate(antibiotic=factor(antibiotic, levels=top_abx)) %>%
+  ggplot(aes(x=year, y=cpkp)) +
+  facet_wrap(~ antibiotic, ncol=3) +
+  geom_point() + geom_line() +
+  theme_minimal() +
+  ggtitle('Unadjusted consumption')
+
+claims_by_abx %>%
   top_abx_f() %>%
@@
 
-P-values for trends in the data. (These are $t$-tests on the differences between subsequent years, comparing their mean to zero.)
-
-```{r}
-cohort_dat %>%
-  top_abx_f() %>%
-  arrange(antibiotic, year) %>%
-  group_by(antibiotic) %>%
-  summarize(p.value=trend_p(cpkp)) %>%
-  ungroup() %>%
-  mutate(adj.p.value=p.adjust(p.value, method='BH')) %>%
-  select(antibiotic, adj.p.value) %>%
-  mutate(sig=if_else(adj.p.value < 0.05, '*', '')) %>%
-  kable
-```
-
 ## Claims by age group
@@
 ```{r claims_by_age}
-read_tsv('tables/tbl_cohort_claims_by_age.tsv') %>%
+read_tsv('tables/claims_by_age_group.tsv') %>%
   mutate(cpkp=round(cpkp)) %>%
@@
 ```{r}
-read_tsv('tables/tbl_cohort_claims_by_sex.tsv') %>%
+read_tsv('tables/claims_by_sex.tsv') %>%
   select(sex, year, cpkp) %>%
@@
 ```{r}
-read_tsv('tables/tbl_cohort_claims_by_region.tsv') %>%
+read_tsv('tables/claims_by_region.tsv') %>%
   select(region, year, cpkp) %>%
@@
 ```{r}
-read_tsv('tables/tbl_cohort_claims_by_race.tsv') %>%
+read_tsv('tables/claims_by_race.tsv') %>%
   select(race, year, cpkp) %>%
@@
 
-"Adjusted" means the `year` coefficient in `y ~ year + age + n_cc + is_female + is_dual + is_white + region`.
+"Adjusted" means the `year` coefficient in `y ~ year + age + n_cc + is_dual + region`. Gender and race, though important covariates, are not confounded with time.
 
-```{r adjusted_changes, cache.extra=md5sum('tables/tbl_cohort_model_abx.tsv')}
-x = read_tsv('tables/tbl_cohort_model_abx.tsv') %>%
-  filter(term=='year', model %in% top_abx) %>%
+```{r adjusted_changes, cache.extra=md5sum('tables/model_abx.tsv')}
+x = read_tsv('tables/model_abx.tsv') %>%
+  filter(model=='poisson', term=='year', antibiotic %in% top_abx) %>%
   arrange(estimate) %>%
-  mutate(cpkp_per_year=estimate*1000,
-         antibiotic=as_factor(model)) %>%
-  select(antibiotic, cpkp_per_year)
+  mutate(ci=1.96*std.error,
+         pct_change=exp(estimate)-1,
+         ci_low=exp(estimate-ci)-1,
+         ci_high=exp(estimate+ci)-1) %>%
+  select(antibiotic, pct_change, ci_low, ci_high) %>%
+  arrange(pct_change) %>% mutate(antibiotic=forcats::fct_inorder(antibiotic)) %>%
+  mutate_if(is.numeric, function(x) x*100)
 
 p = x %>%
-  ggplot(aes(x=antibiotic, y=cpkp_per_year)) +
+  ggplot(aes(x=antibiotic, y=pct_change, ymin=ci_low, ymax=ci_high)) +
   geom_point() +
-  geom_linerange(aes(x=antibiotic, ymin=0, ymax=cpkp_per_year)) +
+  #geom_linerange(aes(x=antibiotic, ymin=0, ymax=cpkp_per_year)) +
+  geom_linerange() +
   coord_flip() +
@@
   xlab('') +
-  ylab('adjusted yearly change in claims per 1,000 beneficiaries per year') +
-  scale_y_continuous(limits=c(-16, 17), breaks=c(-15, -10, -5, 0, 5, 10, 15)) +
+  ylab('adjusted yearly change in claims per beneficiary per year') +
+  #scale_y_continuous(limits=c(-16, 17), breaks=c(-15, -10, -5, 0, 5, 10, 15)) +
   theme_minimal()
@@
 show(p)
-ggsave('fig/fig_adjusted_changes.pdf', plot=p)
+#ggsave('fig/fig_adjusted_changes.pdf', plot=p)
 

### 2017-07-24 ms1/namcs_trends.Rmd
streamlining

--- 2017-07-07 ms1/namcs_trends.Rmd
@@
 
-top_abx = read_tsv('../../db/namcs/analysis/top_abx.tsv')
+top_abx = read_tsv('namcs_top_abx.tsv')
 abx_cat = read_tsv('../../db/namcs/data/parse/drug_codes/abx_codes.tsv')
@@
 
-abx_visit = left_join(top_abx, drugs, by='desc') %>%
+abx_visit = top_abx %>%
+  select(desc, abx) %>%
+  left_join(drugs, by='desc') %>%
   count(visit_id, abx) %>% ungroup() %>%
@@
 
-I consider the same drugs as last time:
+I consider the same drugs as in Medicare, except for clindamycin, which I drop,
+because NAMCS doesn't have enough information to distinguish between oral and
+topical formulations.
 
-```{r}
-kable(top_abx)
+```{r top_abx_table}
+top_abx %>%
+  select(medicare_abx, abx) %>%
+  kable
 ```
@@
 
-But the number of Americans went up (but the number of children went down):
+However, the number of Americans went up (but the number of children went down):
 
@@
 
-```{r overall_model1}
-svyglm(n_abx ~ year, design=design, family=quasipoisson(link='identity')) %>%
-  tidy %>% kable
+```{r lm_functions}
+# survey poisson model
+svy_pm = function(design, frmla) svyglm(frmla, design=design, family='poisson') %>% tidy
+
+# log glm
+log_glm = function(df, frmla, weights=NULL) glm(formula=as.formula(frmla), data=df, family=gaussian(link='log'), weights=weights) %>% tidy
+```
+
+```{r overall_per_visit}
+svy_pm(design, n_abx ~ year) %>% kable
 ```
@@
 
-```{r overall_model2}
+```{r overall_per_capita}
 overall_trend %>%
   mutate(year=year-min(year)) %>%
-  lm(abx_per_capita ~ year, weights=1.0/se_per_capita^2, data=.) %>%
-  tidy %>% kable
+  log_glm('abx_per_capita ~ year', weights=1.0/.$se_per_capita^2) %>%
+  kable
 ```
@@
 
-```{r drugs_time}
-# apc = antibiotic per capita
-f = function(design, d) {
-  svyby(as.formula(paste0('~', d)), ~year, design=design, svytotal) %>%
+#### Per capita
+
+```{r drugs_per_capita_model}
+n_abx_f = function(design, a) {
+  svyby(as.formula(paste0('~', a)), ~year, design=design, svytotal) %>%
     as_data_frame() %>%
-    rename_(.dots=setNames(d, 'n_abx')) %>%
-    mutate(abx=d)
+    rename_(.dots=setNames(a, 'n_abx')) %>%
+    mutate(abx=a)
 }
 
-g = function(design) {
-  lapply(top_abx$abx, function(x) f(design, x)) %>% bind_rows() %>%
+# apc = antibiotic per capita
+apc_f = function(design, denom) {
+  lapply(top_abx$abx, function(x) n_abx_f(design, x)) %>% bind_rows() %>%
     left_join(census, by='year') %>%
-    mutate(apc=n_abx/n_total, se=se/n_total)
+    rename_(.dots=setNames(denom, 'denom')) %>%
+    mutate(apc=n_abx/denom, se=se/denom)
 }
 
-abx_time = g(design)
-abx_time_adult = g(adult_design)
+abx_time = apc_f(design, 'n_total')
+abx_time_adult = apc_f(adult_design, 'n_adult')
 ```
 
-```{r drugs_time_plot}
+```{r drugs_per_capita_time}
 abx_time %>%
@@
 
-It doesn't look by eye that any trends are particularly strong.
+It doesn't look by eye that any trends are particularly strong. I used weighted
+log regression to look for trends in totals:
 
-I used weighted linear regression to look for trends in totals:
-
-```{r drugs_time_totals_model}
+```{r drugs_per_capita_time_trends}
 abx_total_models = abx_time %>%
   group_by(abx) %>%
-  do(tidy(lm(apc ~ year, weight=1.0/se^2, data=.))) %>%
+  do(log_glm(., 'apc ~ year', weights=1.0/.$se^2)) %>%
   ungroup()
@@
   kable
+```
 
+```{r drugs_per_capita_figure}
+bind_rows(
+  read_tsv('tables/model_abx.tsv') %>%
+    filter(term=='year', model=='poisson') %>%
+    left_join(top_abx, by=c('antibiotic'='medicare_abx')) %>%
+    mutate(abx=if_else(is.na(abx), antibiotic, abx)) %>%
+    select(abx, estimate, std.error) %>%
+    mutate(data_source='medicare'),
   abx_total_models %>%
-  filter(term=='year') %>% select(-term) %>%
-  mutate(ci=1.96*std.error, ymin=estimate-ci, ymax=estimate+ci) %>%
-  mutate(abx=forcats::fct_reorder(abx, estimate)) %>%
-  ggplot(aes(x=abx, y=estimate, ymin=ymin, ymax=ymax)) +
-  geom_point() + geom_linerange() +
+    filter(term=='year') %>%
+    select(abx, estimate, std.error) %>%
+    mutate(data_source='namcs')
+) %>%
+  #mutate(abx=factor(abx, levels=top_abx$abx)) %>%
+#abx_total_models %>%
+#  filter(term=='year') %>%
+  mutate(ci=1.96*std.error, y=exp(estimate)-1, ymin=exp(estimate-ci)-1, ymax=exp(estimate+ci)-1) %>%
+  #select(abx, y, ymin, ymax) %>%
+  mutate_if(is.numeric, function(x) x*100) %>%
+  #arrange(y) %>% mutate(abx=forcats::fct_inorder(abx)) %>%
+  #ggplot(aes(x=abx, y=y, ymin=ymin, ymax=ymax)) +
+  ggplot(aes(x=abx, y=y, ymin=ymin, ymax=ymax, color=data_source)) +
+  geom_linerange() +
+  geom_point() +
   coord_flip() +
+  ylab('pct change in consumption per capita') +
   theme_minimal()
@@
 
-I used Poisson regressions to see if anything was changing *per visit* with time. Only levo:
+#### Per visit
 
-```{r drugs_time_model}
-qpglm = function(design, frmla) {
-  svyglm(frmla, design=design, family=quasipoisson(link='identity'))
+I used Poisson regressions to see if anything was changing *per visit* with
+time: levo and amox/clav. (These results are only with adults.)
+
+```{r drugs_per_visit}
+f = function(design) {
+  lapply(top_abx$abx, function(a) {
+    svy_pm(design, as.formula(paste0(a, '~year'))) %>%
+      mutate(abx=a)
+  }) %>% bind_rows()
 }
 
-f = function(design, d) {
-  qpglm(design, as.formula(paste0(d, '~ year'))) %>%
-    tidy %>%
-    mutate(abx=d) %>%
-    filter(term != '(Intercept)')
-}
-
-g = function(design) lapply(top_abx$abx, function(x) f(design, x)) %>% bind_rows()
-
-abx_time_model = g(design)
-abx_time_model_adult = g(adult_design)
+abx_time_model = f(design)
+abx_time_model_adult = f(adult_design)
 ```
 
-```{r drugs_time_model_table}
-abx_time_model %>%
+```{r drugs_per_visit_table}
+abx_time_model_adult %>%
+  filter(term=='year') %>%
   select(abx, estimate, p.value) %>%
   mutate(adj.p.value=p.adjust(p.value, method='BH')) %>%
-  arrange(p.value) %>%
+  mutate(pct=100*(exp(estimate)-1)) %>%
+  arrange(desc(pct)) %>%
   kable
@@
 
-### Levo trend
-
-```{r levo_ratio}
-levo_time = abx_time %>%
-  filter(abx=='levo') %>%
-  arrange(year)
-
-t1 = levo_time %>% filter(year==2011) %$% apc
-v1 = levo_time %>% filter(year==2011) %$% se
-t2 = levo_time %>% filter(year==2014) %$% apc
-v2 = levo_time %>% filter(year==2014) %$% se
-
-tr = t2 / t1
-vr = v1 / t2^2 + t1^1 / t2^4 * v2
-sdr = sqrt(vr)
-
-show_tot = function(tot, v) sprintf('%g (±%g)', round(tot, 4), round(v, 4))
-show_pct = function(tot, v) sprintf('%g%% (±%g%%)', round((tot - 1)*100, digits=1), round(sqrt(v)*100, digits=1))
+```{r drugs_per_visit_figure}
+abx_time_model_adult %>%
+  filter(term=='year') %>%
+  mutate(ci=1.96*std.error, y=exp(estimate)-1, ymin=exp(estimate-ci)-1, ymax=exp(estimate+ci)-1) %>%
+  select(abx, y, ymin, ymax) %>%
+  mutate_if(is.numeric, function(x) x*100) %>%
+  arrange(y) %>% mutate(abx=forcats::fct_inorder(abx)) %>%
+  ggplot(aes(x=abx, y=y, ymin=ymin, ymax=ymax)) +
+  geom_point() + geom_linerange() +
+  coord_flip() +
+  ylab('pct change in consumption per visit') +
+  theme_minimal()
 ```
 
-Levo increased from `r show_tot(t1, v1)` to `r show_tot(t2, v2)` per capita, an increase of `r show_pct(tr, vr)`.
+```{r special_times}
+levo_t1 = abx_time_adult %>% filter(abx=='levo', year==2011)
+levo_t2 = abx_time_adult %>% filter(abx=='levo', year==2014)
+amoxclav_t1 = abx_time_adult %>% filter(abx=='amoxclav', year==2011)
+amoxclav_t2 = abx_time_adult %>% filter(abx=='amoxclav', year==2014)
 
-# Conclusions
+show_var = function(df) sprintf('%g (±%g)', round(df$apc*1000, 1), round(df$se*1000, 1))
+```
 
-Levo is definitely going up.
+Levofloxacin consumption increased from `r show_var(levo_t1)` to `r show_var(levo_t2)` claims per 1,000 people per year.
+Amox/clav consumption increased from `r show_var(amoxclav_t1)` to `r show_var(amoxclav_t2)` claims per 1,000 people per year.
+
+#### Per visit, Among everyone
+
+I do the same analysis, but among everyone.
+
+```{r drugs_per_visit_everyone}
+abx_time_model %>%
+  filter(term=='year') %>%
+  select(-term) %>%
+  kable
+```

### 2017-07-24 ms1/ms1_work.R

@@ -150,9 +150,17 @@ bene %>%
 # run a model, tidy, and save the output
 subtract_min = function(x) x - min(x)
 model_f = function(df, frmla) {
-  df %>%
-    mutate_at(vars(year, age), subtract_min) %>%
-    lm(formula=frmla, data=.) %>% tidy
+  dat = df %>% mutate_at(vars(year, age), subtract_min)
+  f = function(...) glm(formula=frmla, data=dat, ...)
+  model_lm = f()
+  model_poisson = f(family='poisson')
+  model_log = f(family=gaussian(link='log'), start=model_poisson$coefficients)
+  
+  bind_rows(
+    mutate(tidy(model_lm), model='lm'),
+    mutate(tidy(model_poisson), model='poisson'),
+    mutate(tidy(model_log), model='log')
+  )
 }
 
 # models predicting consumption
@@ -168,13 +176,13 @@ single_abx_model = function(abx, frmla) {
     right_join(bene, by=c('year', 'bene_id')) %>%
     replace_na(list(y=0)) %>%
     model_f(frmla) %>%
-    mutate(model=abx)
+    mutate(antibiotic=abx)
 }
 
 model_abx_overall = bene %>%
   rename(y=n_claims) %>%
   model_f(y ~ year + age + n_cc + is_dual + region) %>%
-  mutate(model='overall')
+  mutate(antibiotic='overall')
 
 model_abx = lapply(top_abx, function(a) single_abx_model(a, y ~ year + age + n_cc + is_dual + region)) %>%
   bind_rows() %>%
@@ -206,7 +214,9 @@ n_dxage = dx %>%
   time_f() %>%
   count(time, age, diagnosis_type) %>% ungroup()
 
-n_rxdxage = pde %>% time_f() %>%
+n_rxdxage = pde %>%
+  left_join(select(bene, year, bene_id, age), by=c('year', 'bene_id')) %>%
+  time_f() %>%
   count(time, age, antibiotic, diagnosis_type) %>% ungroup()
 
 f_rxdx = n_rxdxage %>%

### 2017-07-24 ms1/namcs_trends.Rmd
re-streamlined

@@
 
-```{r load_data}
+```{r load_namcs_data}
 # Caveat: I removed a stratum from the 2011 data that had only one PSU in it.
@@
 n_years = length(years)
-mpy = function(x) x/(1e6 * n_years) # millions per year
 
@@
 }
-
-census = read_tsv('../../../../db/census/census_totals.tsv')
 
@@
   mutate_at(top_abx$abx, function(x) if_else(is.na(x), 0L, x))
+```
+
+```{r load_other_data}
+census = read_tsv('../../../../db/census/census_totals.tsv') %>%
+  mutate(n_nonchild=n_total - n_children)
+
+mpy = function(x) x/(1e6 * n_years) # millions per year
 
@@
 
-I used the 2011-2014 NAMCS data. All numbers of prescriptions will be presented in millions per year. I'll mostly consider only the adults.
+I used the 2011-2014 NAMCS data. All numbers of prescriptions will be presented
+in millions per year. I'll mostly consider only the adults (i.e., those over 18 years).
 
@@
 
-## Summary of trends
+This means that visits per American went down. Thus, consumption of some drug
+could be increasing per visit but not be increasing per capita, for example.
 
-### Overall trends
+## Trends per visit
 
-It looks by eye like the total number of antibiotic prescriptions per capita is going down:
+Methodologically, trends per visit is easier, because we just perform
+regressions on the records in NAMCS. The con is that the results are more
+difficult to interpret in light of the fact that there are trends in population
+size as well as in healthcare usage.
 
-```{r overall}
-overall_trend = svyby(~n_abx, ~year, design=design, svytotal) %>%
-  as_data_frame() %>%
-  left_join(census, by='year') %>%
-  mutate(abx_per_capita=n_abx/n_total, se_per_capita=se/n_total)
+In all the results that follow, I'm looking at consumption among the adult
+population only.
 
-overall_trend %>%
-  mutate(ci=1.96*se_per_capita, ymin=abx_per_capita-ci, ymax=abx_per_capita+ci) %>%
-  ggplot(aes(x=year, y=abx_per_capita, ymin=ymin, ymax=ymax)) +
-  geom_point() + geom_line() + geom_linerange() +
-  theme_minimal()
+```{r lm_functions}
+add_pct_change = function(df) {
+  mutate(df, ci=1.96*std.error,
+             pct_change=exp(estimate)-1,
+             ci_low=exp(estimate-ci)-1,
+             ci_high=exp(estimate+ci)-1) %>%
+    mutate_at(vars(pct_change, ci_low, ci_high), function(x) x*100)
+}
 ```
 
-There is no support for a trend per visit:
-
-```{r lm_functions}
+```{r per_visit_model}
 # survey poisson model
@@
 
-# log glm
-log_glm = function(df, frmla, weights=NULL) glm(formula=as.formula(frmla), data=df, family=gaussian(link='log'), weights=weights) %>% tidy
+overall_pv = adult_design %>% svy_pm(n_abx ~ year) %>% mutate(abx='overall')
+
+drug_pv = top_abx$abx %>%
+  lapply(function(a) svy_pm(adult_design, as.formula(paste0(a, '~year'))) %>% mutate(abx=a)) %>%
+  bind_rows()
 ```
 
-```{r overall_per_visit}
-svy_pm(design, n_abx ~ year) %>% kable
-```
+```{r per_visit_display}
+x = bind_rows(overall_pv, drug_pv) %>%
+  filter(term=='year') %>%
+  mutate(adj.p.value=p.adjust(p.value, method='BH')) %>%
+  add_pct_change() %>%
+  arrange(pct_change) %>%
+  mutate(abx=forcats::fct_inorder(abx))
 
-Nor for a trend per capita:
+ggplot(x, aes(x=abx, y=pct_change, ymin=ci_low, ymax=ci_high)) +
+  geom_linerange() +
+  geom_point() +
+  coord_flip() +
+  xlab('') +
+  ylab('pct change in consumption per visit') +
+  ggtitle('Per visit') +
+  theme_minimal()
 
-```{r overall_per_capita}
-overall_trend %>%
-  mutate(year=year-min(year)) %>%
-  log_glm('abx_per_capita ~ year', weights=1.0/.$se_per_capita^2) %>%
+x %>%
+  select(abx, ci_low, pct_change, ci_high, p.value, adj.p.value) %>%
+  mutate(sig=if_else(adj.p.value < 0.05, '*', '')) %>%
   kable
@@
 
-### Trends by drug
+## Trends per capita
 
-#### Per capita
+```{r per_capita_counts}
+# apc = antibiotics per capita
 
-```{r drugs_per_capita_model}
+overall_pc_f = function(design, denom) {
+  svyby(~n_abx, ~year, design=design, svytotal) %>%
+    as_data_frame() %>%
+    left_join(census, by='year') %>%
+    rename_(.dots=setNames(denom, 'denom')) %>%
+    mutate(apc=n_abx/denom, se=se/denom) %>%
+    mutate(abx='overall')
+}
+
 n_abx_f = function(design, a) {
@@
 
-# apc = antibiotic per capita
 apc_f = function(design, denom) {
-  lapply(top_abx$abx, function(x) n_abx_f(design, x)) %>% bind_rows() %>%
+  lapply(top_abx$abx, function(x) n_abx_f(design, x)) %>%
+    bind_rows() %>%
     left_join(census, by='year') %>%
     rename_(.dots=setNames(denom, 'denom')) %>%
-    mutate(apc=n_abx/denom, se=se/denom)
+    mutate(apc=n_abx/denom, se=se/denom) %>%
+    bind_rows(overall_pc_f(design, denom))
 }
 
-abx_time = apc_f(design, 'n_total')
-abx_time_adult = apc_f(adult_design, 'n_adult')
+per_capita_counts = apc_f(adult_design, 'n_nonchild')
+per_capita_counts_all = apc_f(design, 'n_total')
 ```
 
-```{r drugs_per_capita_time}
-abx_time %>%
+```{r per_capita_counts_display}
+# free scales, with overall
+per_capita_counts %>%
+  mutate(abx=factor(abx, levels=c('overall', top_abx$abx))) %>%
   mutate(ci=1.96*se, ymin=apc-ci, ymax=apc+ci) %>%
@@
   theme_minimal()
-```
 
-And the same plot again, but with fixed scales:
-
-```{r drugs_time_plot_fixed}
-abx_time %>%
+# fixed scales, without overall
+per_capita_counts %>%
+  filter(abx != 'overall') %>%
+  mutate(abx=factor(abx, levels=top_abx$abx)) %>%
   mutate(ci=1.96*se, ymin=apc-ci, ymax=apc+ci) %>%
   ggplot(aes(x=year, y=apc, ymin=ymin, ymax=ymax)) +
-  facet_wrap(~abx, ncol=3) +
+  facet_wrap(~abx, ncol=3, scales='fixed') +
   geom_point() + geom_line() + geom_linerange() +
@@
 
-It doesn't look by eye that any trends are particularly strong. I used weighted
-log regression to look for trends in totals:
-
-```{r drugs_per_capita_time_trends}
-abx_total_models = abx_time %>%
+```{r per_capita_models}
+per_capita = per_capita_counts %>%
   group_by(abx) %>%
-  do(log_glm(., 'apc ~ year', weights=1.0/.$se^2)) %>%
-  ungroup()
-
-abx_total_models %>%
-  filter(term=='year') %>% select(-term) %>%
-  arrange(estimate) %>%
-  kable
+  do(tidy(glm(apc ~ year, data=., family=gaussian(link='log'), weights=1.0/se^2)))
 ```
 
-```{r drugs_per_capita_figure}
-bind_rows(
+```{r per_capita_models_display}
+x = per_capita %>%
+  filter(term=='year') %>%
+  mutate(adj.p.value=p.adjust(p.value, method='BH')) %>%
+  add_pct_change()
+
 read_tsv('tables/model_abx.tsv') %>%
@@
   mutate(abx=if_else(is.na(abx), antibiotic, abx)) %>%
-    select(abx, estimate, std.error) %>%
-    mutate(data_source='medicare'),
-  abx_total_models %>%
-    filter(term=='year') %>%
-    select(abx, estimate, std.error) %>%
-    mutate(data_source='namcs')
-) %>%
-  #mutate(abx=factor(abx, levels=top_abx$abx)) %>%
-#abx_total_models %>%
-#  filter(term=='year') %>%
-  mutate(ci=1.96*std.error, y=exp(estimate)-1, ymin=exp(estimate-ci)-1, ymax=exp(estimate+ci)-1) %>%
-  #select(abx, y, ymin, ymax) %>%
-  mutate_if(is.numeric, function(x) x*100) %>%
-  #arrange(y) %>% mutate(abx=forcats::fct_inorder(abx)) %>%
-  #ggplot(aes(x=abx, y=y, ymin=ymin, ymax=ymax)) +
-  ggplot(aes(x=abx, y=y, ymin=ymin, ymax=ymax, color=data_source)) +
+  mutate(data_source='medicare') %>%
+  bind_rows(per_capita %>% filter(term=='year') %>% mutate(data_source='namcs')) %>%
+  add_pct_change() %>%
+  mutate(abx=forcats::fct_reorder(abx, pct_change)) %>%
+  ggplot(aes(x=abx, y=pct_change, ymin=ci_low, ymax=ci_high, color=data_source)) +
   geom_linerange() +
@@
   coord_flip() +
+  xlab('') +
   ylab('pct change in consumption per capita') +
+  ggtitle('Per capita') +
   theme_minimal()
-```
 
-#### Per visit
-
-I used Poisson regressions to see if anything was changing *per visit* with
-time: levo and amox/clav. (These results are only with adults.)
-
-```{r drugs_per_visit}
-f = function(design) {
-  lapply(top_abx$abx, function(a) {
-    svy_pm(design, as.formula(paste0(a, '~year'))) %>%
-      mutate(abx=a)
-  }) %>% bind_rows()
-}
-
-abx_time_model = f(design)
-abx_time_model_adult = f(adult_design)
-```
-
-```{r drugs_per_visit_table}
-abx_time_model_adult %>%
-  filter(term=='year') %>%
-  select(abx, estimate, p.value) %>%
-  mutate(adj.p.value=p.adjust(p.value, method='BH')) %>%
-  mutate(pct=100*(exp(estimate)-1)) %>%
-  arrange(desc(pct)) %>%
+x %>%
+  select(abx, ci_low, pct_change, ci_high, p.value, adj.p.value) %>%
+  arrange(desc(pct_change)) %>%
   kable
@@
 
-```{r drugs_per_visit_figure}
-abx_time_model_adult %>%
-  filter(term=='year') %>%
-  mutate(ci=1.96*std.error, y=exp(estimate)-1, ymin=exp(estimate-ci)-1, ymax=exp(estimate+ci)-1) %>%
-  select(abx, y, ymin, ymax) %>%
-  mutate_if(is.numeric, function(x) x*100) %>%
-  arrange(y) %>% mutate(abx=forcats::fct_inorder(abx)) %>%
-  ggplot(aes(x=abx, y=y, ymin=ymin, ymax=ymax)) +
-  geom_point() + geom_linerange() +
-  coord_flip() +
-  ylab('pct change in consumption per visit') +
-  theme_minimal()
-```
+## Special drugs
 
 ```{r special_times}
-levo_t1 = abx_time_adult %>% filter(abx=='levo', year==2011)
-levo_t2 = abx_time_adult %>% filter(abx=='levo', year==2014)
-amoxclav_t1 = abx_time_adult %>% filter(abx=='amoxclav', year==2011)
-amoxclav_t2 = abx_time_adult %>% filter(abx=='amoxclav', year==2014)
+levo_t1 = per_capita_counts %>% filter(abx=='levo', year==2011)
+levo_t2 = per_capita_counts %>% filter(abx=='levo', year==2014)
+amoxclav_t1 = per_capita_counts %>% filter(abx=='amoxclav', year==2011)
+amoxclav_t2 = per_capita_counts %>% filter(abx=='amoxclav', year==2014)
 
@@
 Amox/clav consumption increased from `r show_var(amoxclav_t1)` to `r show_var(amoxclav_t2)` claims per 1,000 people per year.
-
-#### Per visit, Among everyone
-
-I do the same analysis, but among everyone.
-
-```{r drugs_per_visit_everyone}
-abx_time_model %>%
-  filter(term=='year') %>%
-  select(-term) %>%
-  kable
-```

### 2017-07-24 ms1/namcs_trends.Rmd
validity checks

@@
 
+## Data validity
+
+The NAMCS methodology recommends that a value be considered sensible only if
+(*a*) there are 30 records to support it and (*b*) and standard error is no
+more than 30% of the estimate. There are at least 30 records per antibiotic per
+year:
+
+```{r data_validity}
+visits %>%
+  select(year, amox:tmpsmx) %>%
+  group_by(year) %>%
+  summarize_all(function(x) sum(x > 0)) %>%
+  kable
+```
+
 # Results
@@
 
+Levofloxacin has the largest apparent increase per visit (i.e., the number of
+visits in which levofloxacin was prescribed increases the most).
+
 ```{r per_visit_model}
@@
 ```
+
+Levofloxacin and amox/clav have statistically significant results, as reported
+by the Poisson regressions. Only the levofloxacin result meets the standard
+error criterion for data validity.
 
@@
 x %>%
-  select(abx, ci_low, pct_change, ci_high, p.value, adj.p.value) %>%
-  mutate(sig=if_else(adj.p.value < 0.05, '*', '')) %>%
+  mutate(valid=abs(std.error/estimate)<=0.3, sig=if_else(adj.p.value < 0.05, '*', '')) %>%
+  select(abx, ci_low, pct_change, ci_high, p.value, adj.p.value, valid, sig) %>%
   kable
@@
 
+The trends in consumption per capita are all consistent with those observed in
+Medicare, but none of them meet the standard error criterion for validity or
+are statistically significant.
+
 ```{r per_capita_models_display}
@@
   filter(term=='year') %>%
-  mutate(adj.p.value=p.adjust(p.value, method='BH')) %>%
+  mutate(valid=abs(std.error/estimate)<=0.3, adj.p.value=p.adjust(p.value, method='BH')) %>%
   add_pct_change()
@@
   ggplot(aes(x=abx, y=pct_change, ymin=ci_low, ymax=ci_high, color=data_source)) +
-  geom_linerange() +
+  #geom_linerange() +
+  geom_errorbar() +
   geom_point() +
@@
 x %>%
-  select(abx, ci_low, pct_change, ci_high, p.value, adj.p.value) %>%
+  select(abx, ci_low, pct_change, ci_high, p.value, valid, adj.p.value) %>%
   arrange(desc(pct_change)) %>%

### 2017-07-26 ms1/ms1_display.Rmd
more streamlining

@@
 import::from(knitr, kable)
-import::from(lazyeval, interp)
-import::from(broom, tidy)
 import::from(stringr, str_replace, str_detect)
-import::from(forcats, as_factor)
+library(forcats)
 import::from(tools, md5sum)
+```
 
+# Abstract
+
+```{r abstract}
 pctize = function(df, denom_name) {
@@
 }
+
+totals = read_tsv('tables/totals.tsv') %>%
+  pctize('n_bene') %>%
+  select(year, n_bene, mean_age, sd_age, mean_n_cc, sd_n_cc,
+         pct_female, pct_white, pct_dual,
+         pct_region_south, pct_region_midwest, pct_region_northeast, pct_region_west)
+
+claims_by_abx = read_tsv('tables/claims_by_abx.tsv')
+
+n_million_bene = totals %$% unique(n_bene) %>% { round(. / 1e6, 1) }
+n_million_pde = claims_by_abx %$% sum(n) %>% { round(. / 1e6, 1) }
+
+est2pct = function(x) (exp(x) - 1)*100
+model_ci = function(df) {
+  df %>%
+    filter(model=='poisson', term=='year') %>%
+    mutate(pct_change=est2pct(estimate),
+           ci=1.96*std.error,
+           ci_low=est2pct(estimate - ci),
+           ci_high=est2pct(estimate + ci),
+           change_str=sprintf('%.2f%% (%.2f--%.2f%%)', pct_change, ci_low, ci_high))
+}
+
+model_abx = read_tsv('tables/model_abx.tsv') %>%
+  model_ci()
+
+total_adjusted_change_str = model_abx %>%
+  filter(antibiotic=='overall') %$%
+  change_str
+
+azithro_adjusted_change_str = model_abx %>%
+  filter(antibiotic=='azithromycin') %$%
+  change_str
+
+levo_adjusted_change_str = model_abx %>%
+  filter(antibiotic=='levofloxacin') %$%
+  change_str
 ```
 
-# Abstract
+There are `r n_million_bene` million beneficiaries and `r n_million_pde` million PDEs.
+Adjusted total consumption changed by `r total_adjusted_change_str`.
+Azithromycin changed by `r azithro_adjusted_change_str`.
+Levofloxacin changed by `r levo_adjusted_change_str`.
 
-```{r, eval=FALSE}
-model_ci = function(df) {
-  mutate(df, ci_range=1.96*std.error,
-         ci_low=estimate-ci_range,
-         ci_high=estimate+ci_range)
-}
-
-model_ci_cpkp = function(df) {
-  model_ci(df) %>%
-    mutate_at(vars(ci_range, ci_low, ci_high, estimate), function(x) x*1000) %>%
-    rename(cpkp=estimate)
-}
-
-read_tsv('tables/tbl_cohort_model2.tsv') %>%
-  filter(term=='year') %>%
-  model_ci_cpkp() %>%
-  kable
-
-read_tsv('tables/tbl_cohort_model_abx.tsv') %>%
-  filter(model %in% c('azithromycin', 'levofloxacin'), term=='year') %>%
-  select(term, estimate, std.error, model) %>%
-  model_ci_cpkp() %>%
-  kable
-```
-
-# Summary of data
+# Beneficiary characteristics
 
@@
 
-```{r}
-totals_dat = read_tsv('tables/totals.tsv') %>%
+A star indicates that the trend (*t*-test for mean difference between years equal to zero) is significant after multiple hypothesis correction.
+
+```{r bene_char_table}
+trend_p = function(x) tidy(t.test(diff(x)))$p.value
+totals_trends = totals %>%
+  gather('var', 'val', -year) %>%
+  arrange(var, year, val) %>%
+  group_by(var) %>%
+  summarize(p=trend_p(val)) %>%
+  mutate(adj.p=p.adjust(p, method='BH')) %>%
+  mutate(sig=case_when(.$adj.p < 0.05 ~ '*',
+                       .$p < 0.05 ~ '.',
+                       TRUE ~ ''))
+
+totals %>%
   pctize('n_bene') %>%
@@
          pct_female, pct_white, pct_dual,
-         pct_region_south, pct_region_midwest, pct_region_northeast, pct_region_west)
-
-totals_dat %>%
+         pct_region_south, pct_region_midwest, pct_region_northeast, pct_region_west) %>%
   mutate_at(vars(starts_with('pct_')), function(x) round(x, digits=1)) %>%
@@
   gather('var', 'val', -year) %>% spread(year, val) %>% # transpose
-  kable
-```
-
-```{r}
-trend_p = function(x) tidy(t.test(diff(x)))$p.value
-
-totals_dat %>%
-  gather('var', 'val', -year) %>%
-  arrange(var, year, val) %>%
-  group_by(var) %>%
-  summarize(p=trend_p(val)) %>%
-  mutate(adj.p=p.adjust(p, method='BH'),
-         sig=if_else(adj.p<0.05, '*', '')) %>%
+  left_join(totals_trends %>% select(var, sig), by='var') %>%
   kable
@@
 
-```{r}
-claims_by_abx = read_tsv('tables/claims_by_abx.tsv')
-
+```{r claims_by_abx}
 top_abx = claims_by_abx %>%
@@
 
-top_abx_f = function(df) {
-  totals = group_by(df, year) %>%
-    summarize(cpkp=sum(cpkp)) %>%
-    mutate(antibiotic='all')
+# this should be the same list as for the models
+stopifnot(all(top_abx %in% model_abx$antibiotic))
 
-  df %>%
-    select(antibiotic, year, cpkp) %>%
-    filter(antibiotic %in% top_abx) %>%
-    bind_rows(totals) %>%
-    mutate(antibiotic=factor(antibiotic, levels=c('all', top_abx)))
-}
-```
+# but "overall" should not be in the claims_by_abx list
+if('overall' %in% claims_by_abx$antibiotic) stop()
 
-```{r}
-claims_by_abx %>%
-  top_abx_f() %>%
+claims_overall = claims_by_abx %>%
+  select(-n_bene) %>%
+  group_by(year) %>%
+  summarize_at(vars(n, cpkp), sum) %>%
+  mutate(antibiotic='overall')
+
+claims_by_top_abx = claims_by_abx %>%
+  bind_rows(claims_overall) %>%
+  mutate(antibiotic=fct_relevel(fct_other(antibiotic, keep=c('overall', top_abx)),
+                                c('overall', top_abx, 'Other'))) %>%
+  group_by(year, antibiotic) %>%
+  summarize_at(vars(n, cpkp), sum) %>% ungroup()
+
+claims_by_top_abx %>%
   ggplot(aes(x=year, y=cpkp)) +
-  facet_wrap(~ antibiotic, ncol=3, scales='free_y') +
+  facet_wrap(~antibiotic, ncol=3, scales='free_y') +
   geom_point() + geom_line() +
@@
 
-claims_by_abx %>%
-  filter(antibiotic %in% top_abx) %>%
-  mutate(antibiotic=factor(antibiotic, levels=top_abx)) %>%
+claims_by_top_abx %>%
+  filter(antibiotic != 'overall') %>%
   ggplot(aes(x=year, y=cpkp)) +
-  facet_wrap(~ antibiotic, ncol=3) +
+  facet_wrap(~antibiotic, ncol=3) +
   geom_point() + geom_line() +
@@
 
-claims_by_abx %>%
-  top_abx_f() %>%
+claims_by_top_abx %>%
+  select(year, antibiotic, cpkp) %>%
   filter(year %in% c(2011, 2014)) %>%
@@
 
-```{r}
+```{r claims_by_sex}
 read_tsv('tables/claims_by_sex.tsv') %>%
@@
 
-```{r}
+```{r claims_by_region}
 read_tsv('tables/claims_by_region.tsv') %>%
@@
 
-```{r}
+```{r claims_by_race}
 read_tsv('tables/claims_by_race.tsv') %>%
@@
 ```{r adjusted_changes, cache.extra=md5sum('tables/model_abx.tsv')}
-x = read_tsv('tables/model_abx.tsv') %>%
-  filter(model=='poisson', term=='year', antibiotic %in% top_abx) %>%
-  arrange(estimate) %>%
-  mutate(ci=1.96*std.error,
-         pct_change=exp(estimate)-1,
-         ci_low=exp(estimate-ci)-1,
-         ci_high=exp(estimate+ci)-1) %>%
+x = model_abx %>%
+  filter(term=='year') %>%
   select(antibiotic, pct_change, ci_low, ci_high) %>%
-  arrange(pct_change) %>% mutate(antibiotic=forcats::fct_inorder(antibiotic)) %>%
-  mutate_if(is.numeric, function(x) x*100)
+  mutate(is_overall=antibiotic=='overall') %>%
+  arrange(is_overall, pct_change) %>%
+  mutate(antibiotic=fct_inorder(antibiotic)) %>%
+  select(-is_overall)
 
-p = x %>%
+x %>%
   ggplot(aes(x=antibiotic, y=pct_change, ymin=ci_low, ymax=ci_high)) +
   geom_point() +
-  #geom_linerange(aes(x=antibiotic, ymin=0, ymax=cpkp_per_year)) +
   geom_linerange() +
@@
   theme_minimal()
+```
 
-show(p)
-#ggsave('fig/fig_adjusted_changes.pdf', plot=p)
-
-x %>% kable
+```{r adjusted_changes_table}
+x %>% arrange(-row_number()) %>% kable
 ```

### 2017-07-26 ms1/ms1_work.R

@@ -147,7 +147,7 @@ bene %>%
     between(.$age, 87, 96) ~ '87-96')) %>%
   summarize_by('age_group')
 
-# run a model, tidy, and save the output
+# run a models and tidy
 subtract_min = function(x) x - min(x)
 model_f = function(df, frmla) {
   dat = df %>% mutate_at(vars(year, age), subtract_min)
@@ -163,10 +163,11 @@ model_f = function(df, frmla) {
   )
 }
 
-# models predicting consumption
-# models with all beneficiaries
-model_f(bene, n_claims ~ year) %>% output_table('model_total_unadjusted')
-model_f(bene, n_claims ~ year + age + n_cc + is_dual + region) %>% output_table('model_total_adjusted')
+# model for all drugs combined
+model_abx_overall = bene %>%
+  rename(y=n_claims) %>%
+  model_f(y ~ year + age + n_cc + is_dual + region) %>%
+  mutate(antibiotic='overall')
 
 # models for each individual drug
 single_abx_model = function(abx, frmla) {
@@ -179,12 +180,8 @@ single_abx_model = function(abx, frmla) {
     mutate(antibiotic=abx)
 }
 
-model_abx_overall = bene %>%
-  rename(y=n_claims) %>%
-  model_f(y ~ year + age + n_cc + is_dual + region) %>%
-  mutate(antibiotic='overall')
-
-model_abx = lapply(top_abx, function(a) single_abx_model(a, y ~ year + age + n_cc + is_dual + region)) %>%
+model_abx = top_abx %>%
+  lapply(function(a) single_abx_model(a, y ~ year + age + n_cc + is_dual + region)) %>%
   bind_rows() %>%
   bind_rows(model_abx_overall) %T>%
   output_table('model_abx')

### 2017-07-26 ms1/ms1_display.Rmd
more streamlining

@@
 ```{r global_options, include=FALSE}
-knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='fig/', dev=c('pdf', 'png'),
+knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='fig/', dev=c('png', 'pdf'),
                       echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE,
@@
 
-```{r abstract}
+```{r abstract, cache.extra=md5sum('tables/model_abx.tsv')}
 pctize = function(df, denom_name) {
@@
   df %>%
-    filter(model=='poisson', term=='year') %>%
+    filter(model=='poisson') %>%
     mutate(pct_change=est2pct(estimate),
@@
            ci_high=est2pct(estimate + ci),
-           change_str=sprintf('%.2f%% (%.2f--%.2f%%)', pct_change, ci_low, ci_high))
+           change_str=sprintf('%.2f%% (%.2f -- %.2f%%)', pct_change, ci_low, ci_high))
 }
@@
 total_adjusted_change_str = model_abx %>%
-  filter(antibiotic=='overall') %$%
+  filter(term=='year', antibiotic=='overall') %$%
   change_str
@@
 azithro_adjusted_change_str = model_abx %>%
-  filter(antibiotic=='azithromycin') %$%
+  filter(term=='year', antibiotic=='azithromycin') %$%
   change_str
@@
 levo_adjusted_change_str = model_abx %>%
-  filter(antibiotic=='levofloxacin') %$%
+  filter(term=='year', antibiotic=='levofloxacin') %$%
   change_str
@@
 
-```{r adjusted_changes, cache.extra=md5sum('tables/model_abx.tsv')}
+```{r adjusted_changes}
 x = model_abx %>%
@@
 
-# Azithro and levo
+# Prescribing practice
 
-I used the encounter data to anotate which beneficiaries had COPD or an ARTI in that year. I used these primary diagnoses:
+## Methodology
 
-```{r}
-read_tsv('../sex-difference/sex_codes.tsv') %>%
-  filter(diagnosis_type %in% c('acute_rc', 'copd')) %>%
+Let $R_0$ be the event where the drug of interest is prescribed on some specific day in the first study period; let $R_1$ represent the same thing in the second period.
+Let $D_0$ be the event where the diagnosis of interest was made within some window before the day of interest; similarly $D_1$.
+Thus we have $P(R_0 | D_0)$, which represents, say, the probability that levo will be prescribed following a pneumonia diagnosis in 2011.
+The fold change in the
+probability that a drug will be prescribed for a diagnosis is
+$$
+\frac{P(R_1 | D_1)}{P(R_0 | D_0)} = \frac{P(R_1 \cap D_1) / P(R_0 \cap D_0)}{P(D_1) / P(D_0)}.
+$$
+
+The ratio of the joint probabilities is the ratio in the number of times a qualifying prescription-diagnosis pair appears in each year.
+The ratio of the diagnosis-in-window probabilities is the ratio of the number of diagnoses of that type made in each year.
+
+(There's a way to do the age adjustment here too: just replace every appearance
+of $D_y$ with $D_y \cap A_{ya}$.)
+
+```{r rxdx}
+rxdx = read_tsv('tables/rxdx.tsv')
+
+rxdx_f = function(a) {
+  rxdx %>%
+    filter(abx==a) %>%
+    select(-abx) %>%
+    arrange(desc(f_rxdx))
+}
+```
+
+## Levo
+
+Levofloxacin prescribing increased across diagnoses. Of the respiratory diagnoses, pneumonia was the one associated with the smallest fractional increase in probability of prescription.
+
+```{r rxdx_levo}
+rxdx_f('levofloxacin') %>% filter(f_rxdx > 0.03) %>% kable
+```
+
+## Azithro
+
+Azithromycin use decreased across diagnoses. Pneumonia experienced a greater decline.
+
+```{r rxdx_azithro}
+rxdx_f('azithromycin') %>% filter(f_rxdx > 0.01) %>% kable
+```
+
+## Azithromycin & heart disease
+
+Azithromycin prescribing for people with heart disease decreased slightly more than the people without heart disease.
+
+```{r model_azithro}
+model_azithro = read_tsv('tables/model_azithro.tsv') %>%
+  filter(model=='poisson')
+
+nohd_est = model_azithro %>% filter(term=='year') %$% estimate
+nohd_se = model_azithro %>% filter(term=='year') %$% std.error
+offset_est = model_azithro %>% filter(term=='year:heart_diseaseTRUE') %$% estimate
+offset_se = model_azithro %>% filter(term=='year:heart_diseaseTRUE') %$% std.error
+
+hd_est = nohd_est + offset_est
+hd_se = sqrt(nohd_se^2 + offset_se^2)
+
+data_frame(heart_disease=c(FALSE, TRUE),
+           estimate=c(nohd_est, hd_est),
+           std.error=c(nohd_se, hd_se)) %>%
+  mutate(model='poisson') %>%
+  model_ci() %>%
   kable
@@
 
-Then I ran models predicting azithromycin and levofloxacin consumption using these some standard demographic variables and a narrowly-defined health status (COPD and ARTI).
+Compare these changes to the overall azithromycin change: `r azithro_adjusted_change_str`.
 
-## Azithromycin
+A large fraction of beneficiaries have heart disease in each year:
 
-`y ~ year + start_age + is_female + is_dual + is_white + region + copd + copd:year + copd:start_age + acute_rc + acute_rc:year + acute_rc:start_age`
-
-That is, add COPD, ARTI, and the interactions of those terms with beneficiary age (in 2011) and calendar year.
-
-```{r, cache.extra=md5sum('tables/tbl_cohort_model_azithro.tsv')}
-read_tsv('tables/tbl_cohort_model_azithro.tsv') %>%
-  model_ci_cpkp() %>%
-  filter(term != '(Intercept)') %>%
-  select(term, cpkp, ci_low, ci_high, p.value) %>%
-  mutate_at(vars(cpkp, ci_low, ci_high), function(x) round(x, 1)) %>%
+```{r heart_disease_table}
+read_tsv('tables/counts_heartdisease.tsv') %>%
+  group_by(year) %>%
+  mutate(f=n/sum(n)) %>%
+  ungroup %>%
+  filter(heart_disease) %>%
   kable
@@
 
-Azithromycin prescribing decreases with calendar year and with age, even for those who have COPD or an ARTI.
+## Amox/clav
 
-## Levofloxacin
+Amoxicillin increased for sinusitis, its major associated diagnosis, and also for other respiratory diagnoses, except pneumonia (for which it is associated with a very small fraction of prescritions).
+It seems like most of the increase is attributable to sinusitis.
 
-`y ~ year + start_age + is_female + is_dual + is_white + region + acute_rc + acute_rc:year + acute_rc:start_age`
-
-That is, add ARTI and the interactions with beneficiary age and calendar year.
-
-```{r, cache.extra=md5sum('tables/tbl_cohort_model_levo.tsv')}
-read_tsv('tables/tbl_cohort_model_levo.tsv') %>%
-  model_ci_cpkp() %>%
-  filter(term != '(Intercept)') %>%
-  select(term, cpkp, ci_low, ci_high, p.value) %>%
-  mutate_at(vars(cpkp, ci_low, ci_high), function(x) round(x, 1)) %>%
-  kable
+```{r rxdx_amoxclav}
+rxdx_f('amoxicillin/clavulanate') %>% kable
 ```
 
-By contrast, levofloxacin prescribing increases with calendar year and with age, including for those who have and ARTI.
-The overall increase with time is about half the size of the increase with time for those who have an ARTI, meaning that levofloxacin, aside from being prescribed more overall, is being prescribed more especially for those with ARTIs.
+## Keflex
 
-Interestingly, the effect of dual-enrollment is much higher for levofloxacin than azithromycin (about triple). Why would being poorer make you more likely to take levofloxacin?
+Increased prescribing of cephalexin seems t obe mostly for UTIs; prescribing probabilities for skin and soft tissue infections remained nearly steady.
 
-# Diagnoses
-
-```{r}
-read_tsv('tables/model_abx.tsv') %>%
-  filter(term=='year') %>%
-  mutate(model=case_when(.$model=='trimethoprim/sulfamethoxazole' ~ 'TMP/SMX',
-                         .$model=='amoxicillin/clavulanate' ~ 'amox/clav',
-                         TRUE ~ .$model)) %>%
-  mutate(model=forcats::fct_reorder(model, estimate),
-         cpkp=1000*estimate) %>%
-  ggplot(aes(x=model, y=cpkp)) +
-  geom_point() +
-  geom_linerange(aes(x=model, ymin=0, ymax=cpkp)) +
-  coord_flip() +
-  xlab('') + ylab('Adjusted yearly change in claims per 1k bene. per yr') +
-  scale_y_continuous(limits=c(-16, 17), breaks=c(-15, -10, -5, 0, 5, 10, 15)) +
-  theme_minimal()
+```{r rxdx_cephalexin}
+rxdx_f('cephalexin') %>% filter(f_rxdx > 0.017) %>% kable
 ```
-
-# Supplemental results
-
-## Summary of data for the entire group
-
-```{r}
-read_tsv('tables/tbl_totals.tsv') %>%
-  pctize('n_bene') %>%
-  select(year, n_bene, mean_age, sem_age, mean_n_cc, sem_n_cc,
-         pct_female, pct_white, pct_dual,
-         pct_region_south, pct_region_midwest, pct_region_northeast, pct_region_west) %>%
-  mutate_at(vars(starts_with('pct_')), function(x) round(x, digits=1)) %>%
-  mutate(n_bene=prettyNum(n_bene, big.mark=','),
-         mean_age=round(mean_age, digits=1),
-         mean_n_cc=round(mean_n_cc, digits=2)) %>%
-  gather('var', 'val', -year) %>% spread(year, val) %>% # transpose
-  kable
-```
-
-## Consumption trends in both groups
-
-```{r}
-dat = read_tsv('tables/tbl_claims_by_abx.tsv')
-
-all_dat = bind_rows(
-  top_abx_f(dat) %>% mutate(group='all'),
-  top_abx_f(cohort_dat) %>% mutate(group='cohort')
-)
-
-all_dat %>%
-  ggplot(aes(x=year, y=cpkp, color=group)) +
-  facet_wrap(~ antibiotic, ncol=3, scales='free_y') +
-  geom_point() + geom_line() +
-  theme_minimal() +
-  ggtitle('Unadjusted consumption, overall or in cohort')
-```

### 2017-07-26 ms1/ms1_work.R

@@ -186,6 +186,21 @@ model_abx = top_abx %>%
   bind_rows(model_abx_overall) %T>%
   output_table('model_abx')
 
+# special azithromycin model
+# did azithro consumption decrease faster in benes with heart disease?
+model_azithro = pde %>%
+  filter(antibiotic=='azithromycin') %>%
+  count(year, bene_id) %>% ungroup() %>% rename(y=n) %>%
+  right_join(bene, by=c('year', 'bene_id')) %>% replace_na(list(y=0)) %>%
+  mutate(heart_disease=AMI | ATRIALFB | CHF | ISCHMCHT) %>%
+  model_f(y ~ year*heart_disease + age + n_cc + is_dual + region) %T>%
+  output_table('model_azithro')
+
+bene %>%
+  mutate(heart_disease=AMI | ATRIALFB | CHF | ISCHMCHT) %>%
+  count(year, heart_disease) %>%
+  output_table('counts_heartdisease')
+
 # diagnoses analysis
 times = data_frame(year=c(2011, 2014), time=c(0, 1))
 time_f = function(df) left_join(times, df, by='year')

### 2017-07-27 ms1/namcs_trends.Rmd
FD codes for namcs

@@
 
-sex_codes = read_tsv('../sex-difference/namcs_sex_codes.tsv') %>%
-  rename(dx=code)
+dx_categories = read_tsv('../../data/fd_codes.tsv') %>%
+  mutate(dx=str_pad(code, 5, 'right', '-')) %>%
+  select(dx, diagnosis_type)
+unique_dx_categories = unique(dx_categories$diagnosis_type)
 
 dx = read_tsv_years('../../db/namcs/data/namcs_dx_%i.tsv') %>%
-  left_join(sex_codes, by='dx')
+  left_join(dx_categories, by='dx')
 
 dx_visit = dx %>%
-  group_by(visit_id) %>%
-  summarize(acute_rc='acute_rc' %in% diagnosis_type)
+  filter(!is.na(diagnosis_type)) %>%
+  mutate(dummy=TRUE) %>%
+  select(visit_id, diagnosis_type, dummy) %>%
+  distinct() %>%
+  spread(diagnosis_type, dummy) %>%
+  mutate_if(is.logical, function(x) if_else(is.na(x), FALSE, x))
 
@@
   count(visit_id, abx) %>% ungroup() %>%
-  spread(abx, n)
+  #mutate(present=n>0) %>% select(-n) %>%
+  #spread(abx, present) %>%
+  #mutate_if(is.logical, function(x) if_else(is.na(x), FALSE, x))
+  spread(abx, n) %>%
+  mutate_at(vars(-visit_id), function(x) if_else(is.na(x), 0L, x))
 
@@
   left_join(dx_visit, by='visit_id') %>%
-  replace_na(list(acute_rc=FALSE)) %>%
   left_join(n_abx, by='visit_id') %>% replace_na(list(n_abx=0)) %>%
-  left_join(abx_visit, by='visit_id') %>%
-  mutate_at(top_abx$abx, function(x) if_else(is.na(x), 0L, x))
+  left_join(abx_visit, by='visit_id') #%>%
+  #mutate_at(top_abx$abx, function(x) if_else(is.na(x), 0L, x))
 ```
@@
 Amox/clav consumption increased from `r show_var(amoxclav_t1)` to `r show_var(amoxclav_t2)` claims per 1,000 people per year.
+
+## Diagnoses
+
+### Levofloxacin
+
+The NAMCS data can actually confirm that, not only is levo going up per visit, it's going up faster in those with URIs (and not for pneumonia). (Here I'm looking at the entire population.)
+
+```{r levo_dx}
+svyglm(levo ~ year*uri + year*pneumonia, design=design, family='poisson') %>%
+  tidy %>%
+  kable
+```
+
+### Cephalexin
+
+We can also confirm that keflex is going up because of UTIs and now because of SSTIs (even though the use of keflex for UTIs is so much smaller than for SSTIs):
+
+```{r keflex_dx}
+svyglm(keflex ~ year*ssti + year*uti, design=design, family='poisson') %>%
+  tidy %>%
+  kable
+```

### 2017-08-01 ms1/namcs_trends.Rmd
update analysis

@@
   distinct() %>%
-  spread(diagnosis_type, dummy) %>%
-  mutate_if(is.logical, function(x) if_else(is.na(x), FALSE, x))
+  spread(diagnosis_type, dummy)
 
@@
   #mutate_if(is.logical, function(x) if_else(is.na(x), FALSE, x))
-  spread(abx, n) %>%
-  mutate_at(vars(-visit_id), function(x) if_else(is.na(x), 0L, x))
+  spread(abx, n)
 
@@
   left_join(n_abx, by='visit_id') %>% replace_na(list(n_abx=0)) %>%
-  left_join(abx_visit, by='visit_id') #%>%
-  #mutate_at(top_abx$abx, function(x) if_else(is.na(x), 0L, x))
+  left_join(abx_visit, by='visit_id') %>%
+  mutate_at(top_abx$abx, function(x) {
+    case_when(is.na(x) ~ 0L,
+              x == 0 ~ 0L,
+              x > 0 ~ 1L)
+  }) %>%
+  mutate_at(unique_dx_categories, function(x) if_else(is.na(x), FALSE, x))
 ```
@@
 ```{r per_capita_models_display}
-x = per_capita %>%
-  filter(term=='year') %>%
-  mutate(valid=abs(std.error/estimate)<=0.3, adj.p.value=p.adjust(p.value, method='BH')) %>%
-  add_pct_change()
-
-read_tsv('tables/model_abx.tsv') %>%
+x = read_tsv('tables/model_abx.tsv') %>%
   filter(term=='year', model=='poisson') %>%
@@
   mutate(abx=if_else(is.na(abx), antibiotic, abx)) %>%
-  mutate(data_source='medicare') %>%
-  bind_rows(per_capita %>% filter(term=='year') %>% mutate(data_source='namcs')) %>%
+  mutate(data_source='medicare')
+
+x %>% head %>% kable
+
+y = per_capita %>%
+  filter(term=='year') %>%
+  mutate(data_source='namcs')
+
+abx_levels = x %>%
+  arrange(estimate) %$%
+  abx
+
+bind_rows(x, y) %>%
   add_pct_change() %>%
-  mutate(abx=forcats::fct_reorder(abx, pct_change)) %>%
+  mutate(abx=factor(abx, levels=abx_levels),
+         abx=forcats::fct_relevel(abx, 'overall', after=Inf)) %>%
   ggplot(aes(x=abx, y=pct_change, ymin=ci_low, ymax=ci_high, color=data_source)) +
@@
 
-x %>%
+per_capita %>%
+  filter(term=='year') %>%
+  mutate(valid=abs(std.error/estimate)<=0.3,
+         adj.p.value=p.adjust(p.value, method='BH')) %>%
+  add_pct_change() %>%
   select(abx, ci_low, pct_change, ci_high, p.value, valid, adj.p.value) %>%

### 2017-08-01 ms1/ms1_work.R

@@ -36,16 +36,26 @@ load_data = function(year) {
 
   # diagnosis claims
   dx = read_tsv(sprintf('../dx_%i.tsv', year)) %>%
-    mutate(year=year)
+    mutate(year=year) %>%
+    mutate(year=year, dx_id=1:n())
+  
+  # associate a PDE with each diagnosis, if possible
+  dx_pde = dx %>%
+    left_join(pde, by=c('year', 'bene_id')) %>%
+    filter(between(service_date - from_date, 0, rxdx_window)) %>%
+    group_by(dx_id) %>% filter(service_date==min(service_date)) %>% ungroup() %>%
+    select(dx_id, antibiotic) %>%
+    right_join(dx, by='dx_id') %>%
+    replace_na(list(antibiotic='none'))
   
   # associated each PDE with a diagnosis, if possible
-  pde_dx = pde %>%
-    left_join(dx, by=c('year', 'bene_id')) %>%
-    filter(between(service_date - from_date, 0, rxdx_window)) %>%
-    group_by(pde_id) %>% filter(from_date==max(from_date)) %>% ungroup() %>%
-    select(pde_id, diagnosis_type) %>%
-    right_join(pde, by='pde_id') %>%
-    replace_na(list(diagnosis_type='none'))
+  #pde_dx = pde %>%
+  #  left_join(dx, by=c('year', 'bene_id')) %>%
+  #  filter(between(service_date - from_date, 0, rxdx_window)) %>%
+  #  group_by(pde_id) %>% filter(from_date==max(from_date)) %>% ungroup() %>%
+  #  select(pde_id, diagnosis_type) %>%
+  #  right_join(pde, by='pde_id') %>%
+  #  replace_na(list(diagnosis_type='none'))
 
   # total numbers of PDEs
   n_claims = count(pde, bene_id) %>% rename(n_claims=n)
@@ -65,7 +75,7 @@ load_data = function(year) {
     left_join(n_claims, by='bene_id') %>% replace_na(list(n_claims=0))
 
   # take the bene, pde (with diagnoses) and diagnoses (separately)
-  list(bene=bene, pde=pde_dx, dx=dx)
+  list(bene=bene, pde=pde, dx=dx_pde)
 }
 
 dat = lapply(2011:2014, load_data)
@@ -83,7 +93,8 @@ pde = lapply(dat, function(df) df$pde) %>% bind_rows %>%
[…]

### 2017-08-08 ms1/ms1_display.Rmd
new dxrx

@@
 ```{r claims_by_age}
-read_tsv('tables/claims_by_age_group.tsv') %>%
-  mutate(cpkp=round(cpkp)) %>%
+read_tsv('tables/claims_by_age.tsv') %>%
+  mutate(age_group=case_when(between(.$age, 65, 75) ~ '65-75',
+                             between(.$age, 76, 85) ~ '76-85',
+                             between(.$age, 86, 95) ~ '86-95')) %>%
+  group_by(year, age_group) %>%
+  summarize_at(vars(n_bene, n_claims), sum) %>%
+  ungroup() %>%
+  mutate(cpkp=round(1000*n_claims/n_bene)) %>%
   select(year, age_group, cpkp) %>%

### 2017-08-08 ms1/ms1_work.R

@@ -1,12 +1,5 @@
 # Do all the heavy-lifting analyses so that the Rmd can just show the results
-
-sum_over = function(target_columns) {
-  dummy_vars = sapply(1:length(target_columns), function(i) sprintf('x%i', i))
-  dummy_vars_string = paste0(dummy_vars, collapse=',')
-  dummy_vars_list = lapply(target_columns, as.name) %>% setNames(dummy_vars)
-  f = as.formula(paste0(c('~sum(', dummy_vars_string, ')'), collapse=''))
-  lazyeval::interp(f, .values=dummy_vars_list)
-}
+library(forcats)
 
 # Load the census regions. Code them as factors so that Northeast is taken as
 # the baseline in the linear models.
@@ -18,6 +11,7 @@ regions = read_tsv('../../db/census-regions/census-regions.tsv') %>%
 # which is a subset of Alz. & dementia
 condition_names = read_tsv('../chronic-conditions/names.tsv') %>% filter(condition_code != 'ALZH')
 names_1yr_ccs = condition_names %>% filter(condition_ref_years==1) %$% condition_code
+sum_1yr_cc = paste0(names_1yr_ccs, collapse=' + ')
 
 dx_codes = read_tsv('../../data/fd_codes.tsv') %>%
   select(code, diagnosis_type)
@@ -25,7 +19,7 @@ dx_codes = read_tsv('../../data/fd_codes.tsv') %>%
 dx_types = unique(dx_codes$diagnosis_type)
 
 # window of days that you can look ahead from an rx to a dx
-rxdx_window = 3
+rxdx_window = 7
 
 load_data = function(year) {
   # pde data
@@ -34,45 +28,35 @@ load_data = function(year) {
     distinct() %>%
     mutate(year=year, pde_id=1:n())
 
-  # diagnosis claims
-  dx = read_tsv(sprintf('../dx_%i.tsv', year)) %>%
-    mutate(year=year) %>%
-    mutate(year=year, dx_id=1:n())
-  
-  # associate a PDE with each diagnosis, if possible
-  dx_pde = dx %>%
-    left_join(pde, by=c('year', 'bene_id')) %>%
-    filter(between(service_date - from_date, 0, rxdx_window)) %>%
-    group_by(dx_id) %>% filter(service_date==min(service_date)) %>% ungroup() %>%
-    select(dx_id, antibiotic) %>%
[…]

### 2017-08-09 ms1/ms1_display.Rmd
improved ms1 display

@@
 import::from(knitr, kable)
-import::from(stringr, str_replace, str_detect)
+import::from(tools, md5sum)
 library(forcats)
-import::from(tools, md5sum)
 ```
@@
   df %>%
-    filter(model=='poisson') %>%
     mutate(pct_change=est2pct(estimate),
@@
 model_abx = read_tsv('tables/model_abx.tsv') %>%
+  rename(antibiotic=name) %>%
   model_ci()
@@
 total_adjusted_change_str = model_abx %>%
-  filter(term=='year', antibiotic=='overall') %$%
+  filter(term=='year0', antibiotic=='overall') %$%
   change_str
@@
 azithro_adjusted_change_str = model_abx %>%
-  filter(term=='year', antibiotic=='azithromycin') %$%
+  filter(term=='year0', antibiotic=='azithromycin') %$%
   change_str
@@
 levo_adjusted_change_str = model_abx %>%
-  filter(term=='year', antibiotic=='levofloxacin') %$%
+  filter(term=='year0', antibiotic=='levofloxacin') %$%
   change_str
@@
 There are `r n_million_bene` million beneficiaries and `r n_million_pde` million PDEs.
-Adjusted total consumption changed by `r total_adjusted_change_str`.
-Azithromycin changed by `r azithro_adjusted_change_str`.
-Levofloxacin changed by `r levo_adjusted_change_str`.
+Adjusted total consumption changed by `r total_adjusted_change_str` per year.
+Azithromycin changed by `r azithro_adjusted_change_str` per year.
+Levofloxacin changed by `r levo_adjusted_change_str` per year.
 
@@
 A star indicates that the trend (*t*-test for mean difference between years equal to zero) is significant after multiple hypothesis correction.
+A dot indicates significance ($p < 0.05$) before multiple hypothesis correction.
 
@@
   summarize_at(vars(n, cpkp), sum) %>% ungroup()
+```
+
+```{r claims_by_top_abx_figs}
+claims_by_top_abx %>%
+  filter(antibiotic=='overall') %>%
+  ggplot(aes(x=year, y=cpkp)) +
+  geom_point() + geom_line() +
+  theme_minimal() +
+  ggtitle('Unadjusted consumption, all abx')
 
 claims_by_top_abx %>%
+  filter(antibiotic %in% c('azithromycin', 'ciprofloxacin', 'amoxicillin', 'cephalexin', 'trimethoprim/sulfamethoxazole')) %>%
   ggplot(aes(x=year, y=cpkp)) +
-  facet_wrap(~antibiotic, ncol=3, scales='free_y') +
+  facet_wrap(~antibiotic, ncol=5) +
   geom_point() + geom_line() +
   theme_minimal() +
-  ggtitle('Unadjusted consumption')
+  ggtitle('Unadjusted consumption, top 5')
 
 claims_by_top_abx %>%
-  filter(antibiotic != 'overall') %>%
+  filter(antibiotic %in% c('levofloxacin', 'amoxicillin/clavulanate', 'doxycycline', 'nitrofurantoin', 'clindamycin')) %>%
   ggplot(aes(x=year, y=cpkp)) +
-  facet_wrap(~antibiotic, ncol=3) +
+  facet_wrap(~antibiotic, ncol=5) +
   geom_point() + geom_line() +
   theme_minimal() +
-  ggtitle('Unadjusted consumption')
-
-claims_by_top_abx %>%
-  select(year, antibiotic, cpkp) %>%
-  filter(year %in% c(2011, 2014)) %>%
-  spread(year, cpkp) %>%
-  mutate(pct_change=(`2014`-`2011`)/`2011`*100) %>%
-  kable
+  ggtitle('Unadjusted consumption, next 5')
 ```
@@
 
-"Adjusted" means the `year` coefficient in `y ~ year + age + n_cc + is_dual + region`. Gender and race, though important covariates, are not confounded with time.
+Adjusted changes were computed using a Poisson regression predicting claims per capita.
+The unit of analysis were individual beneficiaries in each year.
+The covariates in the regression were age, number of chronic conditions, dual eligibility, sex, race (white or not), and region.
 
@@
 x = model_abx %>%
-  filter(term=='year') %>%
+  filter(term=='year0') %>%
   select(antibiotic, pct_change, ci_low, ci_high) %>%
-  mutate(is_overall=antibiotic=='overall') %>%
-  arrange(is_overall, pct_change) %>%
-  mutate(antibiotic=fct_inorder(antibiotic)) %>%
-  select(-is_overall)
+  arrange(pct_change) %>%
+  mutate(antibiotic=fct_inorder(antibiotic),
+         antibiotic=fct_relevel(antibiotic, 'overall', after=Inf))
 
@@
   ggplot(aes(x=antibiotic, y=pct_change, ymin=ci_low, ymax=ci_high)) +
-  geom_point() +
-  geom_linerange() +
+  geom_bar(stat='identity', fill=NA, color='black') +
+  geom_errorbar(width=0.5) +
   coord_flip() +
@@
   ylab('adjusted yearly change in claims per beneficiary per year') +
-  #scale_y_continuous(limits=c(-16, 17), breaks=c(-15, -10, -5, 0, 5, 10, 15)) +
+  scale_y_continuous(limits=c(-10, 10), breaks=c(-10, -5, 0, 5, 10)) +
   theme_minimal()
@@
 ```{r adjusted_changes_table}
-x %>% arrange(-row_number()) %>% kable
+x %>% arrange(desc(antibiotic)) %>% kable
 ```
+
+## Fluoroquinolones
+
+**swo**: Overall, FQs haven't been changing
+
+## Heart disease
+
+**swo**: Trends for azithro & levo
 
@@
 
-Let $R_0$ be the event where the drug of interest is prescribed on some specific day in the first study period; let $R_1$ represent the same thing in the second period.
-Let $D_0$ be the event where the diagnosis of interest was made within some window before the day of interest; similarly $D_1$.
-Thus we have $P(R_0 | D_0)$, which represents, say, the probability that levo will be prescribed following a pneumonia diagnosis in 2011.
-The fold change in the
-probability that a drug will be prescribed for a diagnosis is
-$$
-\frac{P(R_1 | D_1)}{P(R_0 | D_0)} = \frac{P(R_1 \cap D_1) / P(R_0 \cap D_0)}{P(D_1) / P(D_0)}.
-$$
+The unit of analysis are encounters recorded in the Carrier Line File whose HCPCS code indicated an office visit.
+We only examined encounters whose primary diagnosis fell into set of groups: skin and soft tissue infections (SSTIs), UTIs, probably viral URIs, pneumonia, bronchitis, pharyngitis, sinusitis.
 
-The ratio of the joint probabilities is the ratio in the number of times a qualifying prescription-diagnosis pair appears in each year.
-The ratio of the diagnosis-in-window probabilities is the ratio of the number of diagnoses of that type made in each year.
+## Trends in diagnoses
 
-(There's a way to do the age adjustment here too: just replace every appearance
-of $D_y$ with $D_y \cap A_{ya}$.)
+Trends per 1,000 beneficiaries in each year. The $p$-value indicates a linear trend.
 
-```{r rxdx}
-rxdx = read_tsv('tables/rxdx.tsv')
+```{r dx_counts}
+dx_counts = read_tsv('tables/dx_counts.tsv') %>%
+  left_join(totals, by='year') %>%
+  mutate(dx_pkb=round(n/n_bene*1000, 1)) %>%
+  select(diagnosis_type, year, dx_pkb)
 
-rxdx_f = function(a) {
-  rxdx %>%
-    filter(abx==a) %>%
-    select(-abx) %>%
-    arrange(desc(f_rxdx))
-}
-```
+dx_counts_trends = dx_counts %>%
+  group_by(diagnosis_type) %>%
+  arrange(year) %>%
+  summarize(p=round(trend_p(dx_pkb), 2))
 
-## Levo
-
-Levofloxacin prescribing increased across diagnoses. Of the respiratory diagnoses, pneumonia was the one associated with the smallest fractional increase in probability of prescription.
-
-```{r rxdx_levo}
-rxdx_f('levofloxacin') %>% filter(f_rxdx > 0.03) %>% kable
-```
-
-## Azithro
-
-Azithromycin use decreased across diagnoses. Pneumonia experienced a greater decline.
-
-```{r rxdx_azithro}
-rxdx_f('azithromycin') %>% filter(f_rxdx > 0.01) %>% kable
-```
-
-## Azithromycin & heart disease
-
-Azithromycin prescribing for people with heart disease decreased slightly more than the people without heart disease.
-
-```{r model_azithro}
-model_azithro = read_tsv('tables/model_azithro.tsv') %>%
-  filter(model=='poisson')
-
-nohd_est = model_azithro %>% filter(term=='year') %$% estimate
-nohd_se = model_azithro %>% filter(term=='year') %$% std.error
-offset_est = model_azithro %>% filter(term=='year:heart_diseaseTRUE') %$% estimate
-offset_se = model_azithro %>% filter(term=='year:heart_diseaseTRUE') %$% std.error
-
-hd_est = nohd_est + offset_est
-hd_se = sqrt(nohd_se^2 + offset_se^2)
-
-data_frame(heart_disease=c(FALSE, TRUE),
-           estimate=c(nohd_est, hd_est),
-           std.error=c(nohd_se, hd_se)) %>%
-  mutate(model='poisson') %>%
-  model_ci() %>%
+dx_counts %>%
+  spread(year, dx_pkb) %>%
+  left_join(dx_counts_trends, by='diagnosis_type') %>%
+  arrange(desc(`2011`)) %>%
   kable
@@
 
-Compare these changes to the overall azithromycin change: `r azithro_adjusted_change_str`.
+## Delays between diagnosis and prescription
 
-A large fraction of beneficiaries have heart disease in each year:
+The table shows the number of diagnoses that were followed by that number of days' delay before the following antibiotic prescription. Most prescriptions were filled on the same day as the diagnosis.
 
-```{r heart_disease_table}
-read_tsv('tables/counts_heartdisease.tsv') %>%
-  group_by(year) %>%
+```{r dx_delay}
+dx_delay = read_tsv('tables/dx_delay.tsv')
+
+dx_delay %>%
+  filter(!is.na(delay)) %>%
   mutate(f=n/sum(n)) %>%
-  ungroup %>%
-  filter(heart_disease) %>%
+  kable
+
+f_dx_not_followed = dx_delay %>%
+  group_by(a=!is.na(delay)) %>%
+  summarize(n=sum(n)) %>%
+  spread(a, n) %>%
+  rename(yes=`TRUE`, no=`FALSE`) %$%
+  { yes / (yes + no) }
+```
+
+However, only `r f_dx_not_followed*100`% of these diagnoses were followed by an antibiotic prescription fill.
+
+## Unadjusted changes in prescribing practice
+
+For each diagnosis type, I computed what fraction of diagnoses were followed by treatment with each antibiotic.
+In the table, I only include dx/rx combinations where the antibiotic was filled after at least 5% of diagnoses.
+
+```{r dxrx_fraction}
+dxrx_frac = read_tsv('tables/dxrx_fractions.tsv') %>%
+  group_by(year, diagnosis_type) %>%
+  mutate(total=sum(n)) %>%
+  ungroup() %>%
+  mutate(f=round(100*n/total, 1)) %>%
+  filter(antibiotic %in% top_abx) %>%
+  select(diagnosis_type, antibiotic, year, f) %>%
+  spread(year, f) %>%
+  filter(`2011` >= 5) %>%
+  arrange(diagnosis_type, desc(`2011`))
+
+dxrx_frac %>%
   kable
@@
 
-## Amox/clav
+## Adjusted changes in prescribing practice
 
-Amoxicillin increased for sinusitis, its major associated diagnosis, and also for other respiratory diagnoses, except pneumonia (for which it is associated with a very small fraction of prescritions).
-It seems like most of the increase is attributable to sinusitis.
+To correct for the changing population, I ran a logistic regression for each diagnosis/antibiotic combination.
+The units of analysis were encounters with a primary diagnosis in that diagnosis group, and the outcome was whether that antibiotic was filled after the encounter.
+The main predictor variable was the year, and adjusted covariates included the same as before (age, # chronic conditions, sex, white race, region, dual eligibility).
 
-```{r rxdx_amoxclav}
-rxdx_f('amoxicillin/clavulanate') %>% kable
+```{r dxrx_models}
+read_tsv('tables/model_dxrx.tsv') %>%
+  model_ci() %>%
+  filter(term=='year0') %>%
+  # swo: pick the right models to show
+  right_join(dxrx_frac, by=c('diagnosis_type', 'antibiotic')) %>%
+  filter(!is.na(pct_change)) %>%
+  select(antibiotic, diagnosis_type, pct_change, ci_low, ci_high, p.value) %>%
+  arrange(antibiotic, pct_change) %>%
+  kable
 ```
-
-## Keflex
-
-Increased prescribing of cephalexin seems t obe mostly for UTIs; prescribing probabilities for skin and soft tissue infections remained nearly steady.
-
-```{r rxdx_cephalexin}
-rxdx_f('cephalexin') %>% filter(f_rxdx > 0.017) %>% kable
-```

### 2017-08-09 ms1/ms1_display.Rmd
updated rmd

@@
 
-n_million_bene = totals %$% unique(n_bene) %>% { round(. / 1e6, 1) }
-n_million_pde = claims_by_abx %$% sum(n) %>% { round(. / 1e6, 1) }
+n_unique_bene = scan('tables/n_unique_bene.tsv')
+n_million_unique_bene = round(scan('tables/n_unique_bene.tsv') / 1e6, 2)
+n_million_pde = claims_by_abx %$% sum(n) %>% { round(. / 1e6, 2) }
 
@@
 
-There are `r n_million_bene` million beneficiaries and `r n_million_pde` million PDEs.
+There are `r n_million_unique_bene` million unique beneficiaries and `r n_million_pde` million PDEs across all study years.
 Adjusted total consumption changed by `r total_adjusted_change_str` per year.
@@
 
-In the cohort, all the values remain fixed, except for:
-
-- age
-- chronic conditions
-- to a lesser degree, dual enrollment
-- to a very small degree, geographical location
-
 A star indicates that the trend (*t*-test for mean difference between years equal to zero) is significant after multiple hypothesis correction.
 A dot indicates significance ($p < 0.05$) before multiple hypothesis correction.
+Mean age decreased during the study.
 
@@
   geom_point() + geom_line() +
+  #ylim(0, 1520) +
+  scale_y_continuous(limits=c(0, 1525), breaks=c(0, 250, 500, 1000, 1500)) +
   theme_minimal() +
@@
 claims_by_top_abx %>%
-  filter(antibiotic %in% c('azithromycin', 'ciprofloxacin', 'amoxicillin', 'cephalexin', 'trimethoprim/sulfamethoxazole')) %>%
-  ggplot(aes(x=year, y=cpkp)) +
-  facet_wrap(~antibiotic, ncol=5) +
+  filter(antibiotic %in% c('azithromycin', 'ciprofloxacin', 'amoxicillin', 'cephalexin', 'trimethoprim/sulfamethoxazole', 'levofloxacin')) %>%
+  ggplot(aes(x=year, y=cpkp, color=antibiotic)) +
+  #ggplot(aes(x=year, y=cpkp)) +
+  #facet_wrap(~antibiotic, ncol=5) +
   geom_point() + geom_line() +
+  scale_y_continuous(limits=c(0, 250), breaks=c(0, 50, 100, 150, 200, 250)) +
   theme_minimal() +
@@
 claims_by_top_abx %>%
-  filter(antibiotic %in% c('levofloxacin', 'amoxicillin/clavulanate', 'doxycycline', 'nitrofurantoin', 'clindamycin')) %>%
-  ggplot(aes(x=year, y=cpkp)) +
-  facet_wrap(~antibiotic, ncol=5) +
+  filter(antibiotic %in% c('amoxicillin/clavulanate', 'doxycycline', 'nitrofurantoin', 'clindamycin')) %>%
+  ggplot(aes(x=year, y=cpkp, color=antibiotic)) +
+  #ggplot(aes(x=year, y=cpkp)) +
+  #facet_wrap(~antibiotic, ncol=5) +
   geom_point() + geom_line() +
+  scale_y_continuous(limits=c(0, 100), breaks=c(0, 50, 100)) +
+  #ylim(0, 140) +
   theme_minimal() +
@@
 
-**swo**: Overall, FQs haven't been changing
+Consumption of ciprofloxacin is slightly decreasing, and consumption of levofloxacin is increasing. How does fluoroquinolone consumption overall change?
+
+```{r fq}
+read_tsv('tables/model_fq.tsv') %>%
+  model_ci() %>%
+  select(term, pct_change, ci_low, ci_high, p.value) %>%
+  kable
+```
+
+Overall, fluoroquinolones changed by about 0% per year during this period.
 
@@
 
-**swo**: Trends for azithro & levo
+We suspected that changes in azithromycin and levofloxacin consumption could be due to altered prescribing guidelines, including one that recommends that azithromycin not be used to treat people at risk for heart fibrillations.
+This includes all elderly, but we epxect the risk would be especially high for those people with heart disease (i.e., AMI | ATRIALFB | CHF | ISCHMCHT).
+I ran separate models for azithromycin and levofloxacin where I include an interaction between a heart disease flag and the year.
+
+```{r heart_model}
+bind_rows(
+  read_tsv('tables/model_azithro.tsv') %>% mutate(model='azithro'),
+  read_tsv('tables/model_levo.tsv') %>% mutate(model='levo')
+) %>%
+  model_ci() %>%
+  select(term, model, pct_change) %>%
+  spread(model, pct_change) %>%
+  kable
+```
+
+For azithromycin, there is a moderate interaction: azithromycin consumption decrease 6.6% per year amongst those without heart disease and 10.2% per year ($1.06588 \times 1.03344$) among those with heart disease.
+For levofloxacin, there is a small interaction that is the opposite of what we might expect: levo consumption increased 10.1% per year among those without heart disease and at a lower rate, 9.4% ($1.10149*(1-0.006782)$), among those with heart disease.
+
+The weakness to this analysis is that I use the CCW flag for heart disease during each year, which means that someone might be getting a lot of azithromycin before they have a heart attack, and afterward they get none.
+I'm also using the one-year chronic conditions, so if someone has a heart attack last year, I don't know about that for the next eyar.
+Nevertheless, as a rough slice, it seems like heart disease doesn't explain what's going on.
 
@@
   select(antibiotic, diagnosis_type, pct_change, ci_low, ci_high, p.value) %>%
-  arrange(antibiotic, pct_change) %>%
+  arrange(antibiotic, desc(pct_change)) %>%
   kable

### 2017-08-09 ms1/ms1_2process.R

@@ -53,6 +53,10 @@ totals = bene %>%
   summarize_totals() %T>%
   output_table('totals')
 
+n_unique_bene = bene$bene_id %>%
+  unique %>% length %T>%
+  write('tables/n_unique_bene.tsv')
+
 # the top few inidividual antibiotics
 claims_by_abx = pde %>%
   count(year, antibiotic) %>% ungroup() %>%

### 2017-08-10 ms1/ms1_display.Rmd
ms1 display update

@@
 n_million_unique_bene = round(scan('tables/n_unique_bene.tsv') / 1e6, 2)
-n_million_pde = claims_by_abx %$% sum(n) %>% { round(. / 1e6, 2) }
+n_pde = sum(claims_by_abx$n)
+n_million_pde = round(n_pde / 1e6, 2)
 
@@
 
-total_adjusted_change_str = model_abx %>%
-  filter(term=='year0', antibiotic=='overall') %$%
-  change_str
+claims_overall = claims_by_abx %>%
+  select(-n_bene) %>%
+  group_by(year) %>%
+  summarize_at(vars(n, cpkp), sum) %>%
+  mutate(antibiotic='overall')
 
-azithro_adjusted_change_str = model_abx %>%
-  filter(term=='year0', antibiotic=='azithromycin') %$%
-  change_str
-
-levo_adjusted_change_str = model_abx %>%
-  filter(term=='year0', antibiotic=='levofloxacin') %$%
-  change_str
+bind_rows(claims_by_abx, claims_overall) %>%
+  filter(antibiotic %in% c('overall', 'azithromycin', 'levofloxacin'), year %in% c(2011, 2014)) %>%
+  select(antibiotic, year, cpkp) %>%
+  spread(year, cpkp) %>%
+  mutate(pct_change=round(100*(`2014`-`2011`)/`2011`, 2)) %>%
+  kable
 ```
@@
 There are `r n_million_unique_bene` million unique beneficiaries and `r n_million_pde` million PDEs across all study years.
-Adjusted total consumption changed by `r total_adjusted_change_str` per year.
-Azithromycin changed by `r azithro_adjusted_change_str` per year.
-Levofloxacin changed by `r levo_adjusted_change_str` per year.
 
 # Beneficiary characteristics
+
+There were `r prettyNum(n_unique_bene, big.mark=',')` unique beneficiaries and `r prettyNum(n_pde, big.mark=',')` PDEs.
 
@@
 stopifnot(all(top_abx %in% model_abx$antibiotic))
-
-# but "overall" should not be in the claims_by_abx list
-if('overall' %in% claims_by_abx$antibiotic) stop()
-
-claims_overall = claims_by_abx %>%
-  select(-n_bene) %>%
-  group_by(year) %>%
-  summarize_at(vars(n, cpkp), sum) %>%
-  mutate(antibiotic='overall')
 
@@
   coord_flip() +
-  ggtitle('Adjusted changes in consumption') +
   xlab('') +
-  ylab('adjusted yearly change in claims per beneficiary per year') +
+  ylab('adjusted yearly change in consumption (%)') +
   scale_y_continuous(limits=c(-10, 10), breaks=c(-10, -5, 0, 5, 10)) +
-  theme_minimal()
+  theme_minimal() +
+  theme(axis.title.x = element_text(size=16),
+        axis.text = element_text(size=16))
 ```
@@
 x %>% arrange(desc(antibiotic)) %>% kable
+
+adjusted_az_decrease = x %>%
+  filter(antibiotic=='azithromycin') %$%
+  { (1 + pct_change/100)^3 } %>%
+  { (1 - .) * 100 } %>%
+  round(2)
+
+adjusted_levo_increase = x %>%
+  filter(antibiotic=='levofloxacin') %$%
+  { (1 + pct_change/100)^3 } %>%
+  { (. - 1) * 100 } %>%
+  round(2)
 ```
+
+Consumption of azithromycin fell by `r adjusted_az_decrease`% over the entire study period. Levofloxacin consumption increased by `r adjusted_levo_increase`% over the entire study period.
 
@@
 ```{r fq}
+fq_names = claims_by_abx %>%
+  filter(str_detect(antibiotic, 'floxacin')) %>%
+  group_by(antibiotic) %>%
+  summarize(n=sum(n)) %>%
+  arrange(desc(n)) %$%
+  antibiotic
+
 read_tsv('tables/model_fq.tsv') %>%
@@
 
-Overall, fluoroquinolones changed by about 0% per year during this period.
+Overall, fluoroquinolones changed by about 0% per year during this period. This included `r fq_names`.
 
@@
   left_join(totals, by='year') %>%
-  mutate(dx_pkb=round(n/n_bene*1000, 1)) %>%
+  mutate(dx_pkb=n/n_bene*1000) %>%
   select(diagnosis_type, year, dx_pkb)
@@
   arrange(year) %>%
-  summarize(p=round(trend_p(dx_pkb), 2))
+  summarize(p=trend_p(dx_pkb))
 
@@
   spread(year, dx_pkb) %>%
+  mutate(pct_change_per_year=(`2014`-`2011`)/`2011`*100/3) %>%
   left_join(dx_counts_trends, by='diagnosis_type') %>%
   arrange(desc(`2011`)) %>%
+  mutate_at(vars(`2011`:p), function(x) round(x, 2)) %>%
   kable
@@
   spread(year, f) %>%
-  filter(`2011` >= 5) %>%
+  filter(`2011` >= 3) %>%
   arrange(diagnosis_type, desc(`2011`))

### 2017-08-10 ms2/ms2.Rmd
moved to summarize simply

@@
 
-```{r fnz_mean}
+```{r fnz_mean_old}
 all_ineq %>%

### 2017-08-10 ms2/data/collect-data.R

@@ -1,108 +1,0 @@
-#!/usr/bin/env Rscript
-
-# inequality function
-import::from(ineq, ineq)
-nonzero = function(x) x[x > 0]
-fraction_nonzero = function(x) length(nonzero(x)) / length(x)
-
-pad0 = function(x, to) c(x, rep(0, to - length(x)))
-
-hoover = function(x) 0.5 * sum(abs(x - mean(x))) / sum(x)
-hoover2 = function(x) 0.5 * sum(abs(x[x > 0] - mean(x))) / sum(x)
-
-inequalities = function(x) {
-  nb_par = fitdistrplus::mledist(x, 'nbinom')$estimate
-  nb_size = nb_par['size']
-  nb_prob = nb_size / (nb_size + nb_par['mu'])
-  data_frame(n=n(),
-             total=sum(x),
-             mean=mean(x),
-             fnz=fraction_nonzero(x),
-             gini=ineq(x, type='Gini'),
-             nzgini=ineq(nonzero(x), type='Gini'),
-             nb_size=nb_size,
-             nb_prob=nb_prob,
-             hoover=hoover(x),
-             hoover2=hoover2(x),
-             nzhoover=hoover(nonzero(x))
-             )
-}
-
-consumption_groups = read_tsv('../consumption_groups.tsv')
-
-regions = read_tsv('../../../db/census-regions/census-regions.tsv') %>%
-  select(state, state_abbreviation, region)
-
-hrr_all = read_tsv('../../../db/hrr/hrr.tsv')
-
-summarize_inequality = function(year) {
-  hrr = hrr_all %>%
-    filter(year==year) %>%
-    select(zipcode, hrr)
-
-  bene_fn = sprintf('../../bene_%i.tsv', year)
-  pde_fn = sprintf('../../pde_%i.tsv', year)
-  state_all_fn = sprintf('state_ineq_all_%i.tsv', year)
-  state_fn = sprintf('state_ineq_%i.tsv', year)
-  hrr_all_fn = sprintf('hrr_all_%i.tsv', year)
-  hrr_fn = sprintf('hrr_ineq_%i.tsv', year)
-
-  bene = read_tsv(bene_fn) %>%
-    left_join(regions, by='state') %>%
-    left_join(hrr, by='zipcode')
-
-  pde = read_tsv(pde_fn) %>%
-    rename(drug=antibiotic) %>%
-    left_join(consumption_groups, by='drug') %>%
[…]

### 2017-08-10 ms2/ms2_1summarize.R

@@ -1,0 +1,93 @@
+#!/usr/bin/env Rscript
+
+# utility functions function
+import::from(ineq, ineq)
+nonzero = function(x) x[x > 0]
+fraction_nonzero = function(x) length(nonzero(x)) / length(x)
+pad0 = function(x, to) c(x, rep(0, to - length(x)))
+filter_eq_ = function(df, col, val) filter_(df, str_interp("${col}=='${val}'"))
+
+inequalities = function(pde, drug_group, unit_type, unit, n_bene) {
+  if (drug_group=='overall') {
+    filter_drug_f = function(df) df
+  } else {
+    filter_drug_f = function(df) filter_eq_(df, 'drug_group', drug_group)
+  }
+
+  x = pde %>%
+    filter_eq_(unit_type, unit) %>%
+    filter_drug_f() %$%
+    pad0(n_claims, n_bene)
+
+  nb_par = fitdistrplus::mledist(x, 'nbinom')$estimate
+  nb_size = nb_par['size']
+  nb_prob = nb_size / (nb_size + nb_par['mu'])
+
+  data_frame(drug_group=drug_group,
+             unit_type=unit_type,
+             unit=unit,
+             n_bene=n_bene,
+             n=length(x),
+             total=sum(x),
+             mean=mean(x),
+             fnz=fraction_nonzero(x),
+             gini=ineq(x, type='Gini'),
+             nzgini=ineq(nonzero(x), type='Gini'),
+             nb_size=nb_size,
+             nb_prob=nb_prob)
+}
+
+# auxiliary data
+consumption_groups = read_tsv('consumption_groups.tsv')
+
+regions = read_tsv('../../db/census-regions/census-regions.tsv') %>%
+  select(state, state_abbreviation, region) %>%
+  filter(state != 'District of Columbia')
+stopifnot('DC' %in% regions$state)
+
+# read in HRR designations for all years
+hrr_all = read_tsv('../../db/hrr/hrr.tsv') %>%
+  mutate(hrr=as.character(hrr))
+
+summarize_inequality = function(y) {
+  # get HRR designations just for that year
+  hrr = hrr_all %>%
+    filter(year==y) %>%
+    select(zipcode, hrr)
+
+  # set up input/output filenames
+  bene_fn = sprintf('../bene_%i.tsv', y)
+  pde_fn = sprintf('../pde_%i.tsv', y)
+
[…]

### 2017-08-10 ms2/ms2_2analyze.R

@@ -1,26 +1,29 @@
 # Consumption data
+ineq = read_tsv('ineq.tsv')
 
-regions = read_tsv('../../db/census-regions/census-regions.tsv') %>%
-  select(state, state_abbreviation, region)
-
-zipcodes = read_tsv('../../db/zipcode/zipcode.tsv') %>%
-  select(zipcode=zip, state)
-
-read_years = function(base_fn, years) {
-  lapply(years, function(y) sprintf(base_fn, y) %>% read_tsv %>% mutate(year=y)) %>%
-    bind_rows
-}
-
-hrr = read_tsv('../../db/hrr/hrr.tsv') %>%
-  select(zipcode, hrr) %>%
-  left_join(zipcodes, by='zipcode') %>%
-  left_join(regions, by='state')
-
-years = 2011:2014
-all_ineq = read_years('data/state_ineq_all_%i.tsv', years) %>%
-  left_join(regions, by='state')
-state_ineq = read_years('data/state_ineq_%i.tsv', years)
-hrr_ineq = read_years('data/hrr_ineq_%i.tsv', years)
+# regions = read_tsv('../../db/census-regions/census-regions.tsv') %>%
+#   select(state, state_abbreviation, region)
+# 
+# zipcodes = read_tsv('../../db/zipcode/zipcode.tsv') %>%
+#   select(zipcode=zip, state)
+# 
+# read_years = function(base_fn, years) {
+#   lapply(years, function(y) {
+#       read_tsv(sprintf(base_fn, y)) %>% mutate(year=y)
+#     }) %>%
+#     bind_rows
+# }
+# 
+# hrr = read_tsv('../../db/hrr/hrr.tsv') %>%
+#   select(zipcode, hrr) %>%
+#   left_join(zipcodes, by='zipcode') %>%
+#   left_join(regions, by='state')
+# 
+# years = 2011:2014
+# all_ineq = read_years('data/state_ineq_all_%i.tsv', years) %>%
+#   left_join(regions, by='state')
+# state_ineq = read_years('data/state_ineq_%i.tsv', years)
+# hrr_ineq = read_years('data/hrr_ineq_%i.tsv', years)
 
 # Aggregate usage correlates with fraction of users
 p = ggplot(all_ineq, aes(x=mean, y=fnz, color=factor(year))) +

### 2017-08-10 ms2/ms2.Rmd
ms2 total rework

--- 2017-06-22 ms2/ms2.Rmd
@@
                       echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE, autodep=TRUE)
-import::from(broom, tidy)
 import::from(tools, md5sum)
+library(forcats)
 ```
@@
 ```{r load_data}
+ineq = read_tsv('ineq.tsv') %>%
+  group_by(drug_group, unit_type, unit) %>%
+  summarize_at(vars(mean, fnz, nzgini, gini), mean) %>%
+  ungroup()
+
+results = read_tsv('results.tsv')
+
+consumption_groups = read_tsv('consumption_groups.tsv')
+resistance_groups = read_tsv('resistance_groups.tsv')
+possible_bug_drug = read_tsv('bug_drug.tsv')
+
 regions = read_tsv('../../db/census-regions/census-regions.tsv') %>%
@@
   left_join(regions, by='state')
-
-years = 2011:2014
-read_years = function(template_fn) {
-  lapply(years, function(y) read_tsv(sprintf(template_fn, y)) %>% mutate(year=y)) %>%
-    bind_rows()
-}
-
-all_ineq = read_years('data/state_ineq_all_%i.tsv') %>%
-  group_by(state) %>%
-  summarize(mean_cpb=mean(mean),
-            sem_cpb=sd(mean)/sqrt(n()),
-            mean_fnz=mean(fnz),
-            sem_fnz=sd(fnz)/sqrt(n()),
-            mean_nzgini=mean(nzgini),
-            sem_nzgini=sd(nzgini)/sqrt(n())) %>%
-  left_join(regions, by='state')
-
-state_ineq = read_years('data/state_ineq_%i.tsv') %>%
-  rename(cpb=mean, nzg=nzgini, nzr=mean_fnz_residual) %>%
-  group_by(state, drug_group) %>%
-  summarize_at(vars(-state, -drug_group, -total), mean) %>%
-  ungroup()
-
-hrr_ineq = read_years('data/hrr_ineq_%i.tsv') %>%
-  rename(cpb=mean, nzg=nzgini, nzr=mean_fnz_residual) %>%
-  group_by(hrr, drug_group) %>%
-  summarize_at(vars(-hrr, -drug_group, -total), mean) %>%
-  ungroup()
 ```
 
-```{r load_abg, cache.extra=md5sum('bug_drug.tsv')}
-resistance_groups = read_tsv('resistance_groups.tsv')
-bug_names = read_tsv('bug_names.tsv')
-bug_groups = read_tsv('bug_groups.tsv')
-possible_bug_drug = read_tsv('bug_drug.tsv')
+# Methods
 
-raw_abg = read_tsv('../../../antibiogram/data/abg.tsv', col_types=cols(zipcode='c')) %>%
-  mutate(drug=tolower(drug), percent_nonsusceptible=100-percent_susceptible) %>%
-  left_join(hrr, by=c('zipcode', 'state')) %>%
-  right_join(resistance_groups, by='drug') %>%
-  select(bug, raw_drug=drug, drug_group, state, hrr, percent_nonsusceptible, n_isolates)
+## Consumption data
 
-abg = raw_abg %>%
-  # shorten the bug names
-  left_join(bug_names, by='bug') %>%
-  select(-bug) %>% rename(bug=bug_short) %>%
-  # group the different S. aureus into meta-groups depending on the drug
-  left_join(bug_groups, by=c('bug', 'drug_group')) %>%
-  mutate(bug_group=if_else(is.na(bug_group), bug, bug_group)) %>%
-  select(-bug) %>% rename(bug=bug_group) %>%
-  # keep only particular bug/drug combos
-  right_join(possible_bug_drug, by=c('bug', 'drug_group'))
+I used the same inclusion criteria as in ms. #1. I grouped the beneficiaries by state or by HRR. (When zipcode-to-HRR links differed between years, I used the one from 2014.)
 
-# get the median number of isolates for each bug/drug
-median_n_isolates = abg %>%
-  filter(!is.na(n_isolates)) %>%
-  group_by(bug, drug_group) %>%
-  summarize(median_n_isolates=as.integer(median(n_isolates))) %>%
-  ungroup()
+Then I grouped the drugs into consumption groups:
 
-# replace the NA number of isolates with the median for that bug/drug
-abg %<>%
-  left_join(median_n_isolates) %>%
-  mutate(n_isolates=if_else(is.na(n_isolates), median_n_isolates, n_isolates)) %>%
-  select(-median_n_isolates)
+`r kable(consumption_groups)`
 
-# check that all the short names made it in
-for (x in c('bug', 'n_isolates', 'percent_nonsusceptible')) {
-  if (any(is.na(abg[[x]]))) stop(sprintf('NAs in field "%s"', x))
-}
+For each year, in each geographical unit, for each drug group, I computed the mean consumption per capita, the fraction of beneficiaries who had at least one claim ("fraction nonzero"), and the Gini coefficient among those nonzero users.
 
-summarize_abg = function(df, place_name) {
-    group_by_(df, 'bug', 'drug_group', place_name) %>%
-    summarize(mean_percent_nonsusceptible=weighted.mean(percent_nonsusceptible, sqrt(n_isolates)),
-              n_place_antibiograms=n()) %>%
-    ungroup() %>%
-    mutate(place_weight=sqrt(n_place_antibiograms))
-}
+I then took the mean of these statistics across years.
 
-state_abg = summarize_abg(abg, 'state')
-hrr_abg = summarize_abg(abg, 'hrr')
-```
+## Resistance data
 
-```{r do_models}
-linear_interval = function(model, level) {
-  confint(model, level=level) %>%
-    tidy %>%
-    setNames(c('term', 'ci_low', 'ci_high'))
-}
-
-linear_model = function(df, dependent_var, explanatory_vars) {
-  f = paste0(dependent_var, ' ~ ', paste0(explanatory_vars, collapse = ' + ')) %>% formula
-  m = lm(formula=f, weights=place_weight, data=df)
-  tidy(m) %>%
-    select(term, estimate, p.value) %>%
-    mutate(n=nrow(df)) %>%
-    left_join(linear_interval(m, 0.95), by='term')
-}
-
-spearman_interval = function(df, rho_name, n_name, level) {
-  mada::CIrho(df[[rho_name]], df[[n_name]], level=level) %>%
-    tidy %>%
-    setNames(c('rho', 'ci_low', 'ci_high')) %>%
-    select(-rho)
-}
-
-spearman_model = function(df, dependent_var, explanatory_var) {
-  cor.test(df[[dependent_var]], df[[explanatory_var]], method='spearman') %>%
-    tidy %>%
-    mutate(term=explanatory_var, method='spearman') %>%
-    mutate(n=nrow(df)) %>%
-    group_by(term, estimate, p.value, n) %>%
-    do(spearman_interval(., 'estimate', 'n', 0.95)) %>%
-    ungroup()
-}
-
-uni_term = 'cpb' # univariable explanatory term
-multi_term = 'nzg' #'nb_prob' # multivariable explanatory term
-multi_term_adjuster = 'cpb' # other multivarible explanatory term
-
-models = function(x) {
-  dep = 'mean_percent_nonsusceptible'
-  bind_rows(
-    linear_model(x, dep, uni_term) %>% mutate(model='uni'),
-    linear_model(x, dep, c(multi_term_adjuster, multi_term)) %>% mutate(model='multi'),
-    spearman_model(x, dep, uni_term) %>% mutate(model='spearman')
-  ) %>%
-    mutate(n_data=nrow(x))
-}
-
-results = left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
-  group_by(bug, drug_group) %>%
-  do(models(.)) %>%
-  ungroup
-
-hrr_results = left_join(hrr_abg, hrr_ineq, by=c('hrr', 'drug_group')) %>% 
-  group_by(bug, drug_group) %>%
-  do(models(.)) %>%
-  ungroup
-```
-
-# Landscape of bug/drug correlations
-
-## Selecting correlations to test
-
-I join the resistance drugs into the classes shown below. (I also joined the consumption drugs into similar classes.)
+I join the resistance drugs into the classes shown below:
 
@@
 
+I included both inpatient and outpatient data, because although inpatient resistances were mostly higher than outpatient resistances, the overall trends by state appeared to be the same for both sets of data, and the difference between in- vs.
+outpatient were smaller than the differences between extreme ends of the states.
+
+## Bug/drug combinations
+
 I started by considering these `r nrow(possible_bug_drug)` bug/drug combinations on the basis of their clinical relevance and plausibility:
@@
 
-## Preparing the consumption data
+For each combination, I computed three models:
 
-Unlike previous analyses, this time I treated the consumption and resistance data similarly: I averaged both across all data years.
-Thus, to get a state's Gini coefficient, I computed that number each year and then took the average.
+- a Spearman correlation of mean claims per beneficiary against mean percent nonsusceptible
+- a linear model predicting mean resistance from mean consumption
+- a linear model predicting mean resistance from fraction nonzero and nonzero Gini
+
+# Results
+
+## Landscape of correlations
 
@@
 
-For each bug/drug combination, I computed `mean percent nonsusceptible ~ claims per beneficiary`.
-A star after the bug/drug combination implies that the adjusted $p < 0.05$, and a dot means the unadjusted $p < 0.05$.
-Intervals show 95% confidence intervals.
+Two stars after the bug/drug combination implies that the adjusted $p < 0.05$, and one star means the unadjusted $p < 0.05$. Intervals show 95% confidence intervals.
 
-```{r hrr_spearman_interval}
+```{r hrr_spearman}
 interval_plot = function(df) {
@@
   mutate(p.adj=p.adjust(p.value, method='BH')) %>%
-  mutate(sig=case_when(.$p.adj < 0.05 ~ '*',
-                       .$p.value < 0.05 ~ '.',
+  mutate(sig=case_when(.$p.adj < 0.05 ~ '**',
+                       .$p.value < 0.05 ~ '*',
                        TRUE ~ ''),
-         bug_drug=interaction(bug, drug_group, sig),
-         bug_drug=factor(bug_drug, levels=bug_drug, ordered=TRUE)) %>% 
+         bug_drug=fct_inorder(str_c(bug, drug_group, sig, sep=' '))) %>% 
     ggplot(aes(x=bug_drug, y=estimate, ymin=ci_low, ymax=ci_high, color=drug_group)) +
     geom_point() +
-    geom_linerange() +
+    geom_errorbar(width=0.5) +
     coord_flip()
@@
 
-hrr_results %>%
-  filter(model=='spearman') %>%
+results %>%
+  filter(model=='spearman', unit_type=='hrr') %>%
   interval_plot() +
-  ylab("Spearman's rho")
+  ylab("Spearman's rho") +
+  ggtitle("Spearman, HRR")
 ```
 
-All the demonstrably-different-from-zero relationships are positive.
-This jives with our general understanding that more consumption means more resistance.
-Note also the bunching by drug type, rather than organism.
+```{r state_spearman}
+results %>%
+  filter(model=='spearman', unit_type=='state') %>%
+  interval_plot() +
+  ylab("Spearman's rho") +
+  ggtitle("Spearman, state")
+```
+
+All the demonstrably-different-from-zero relationships are positive. This jibes with our general understanding that more consumption means more resistance.
 
@@
 
-```{r hrr_spearman_table}
-hrr_results %>%
-  filter(model=='spearman') %>%
-  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
-  filter(p.value < 0.05) %>%
-  select(bug, drug_group, p.adj, p.value, rho=estimate, ci_low, ci_high) %>%
-  arrange(p.value) %>%
-  kable
+## Linear models
+
+```{r hrr_linear}
+results %>%
+  filter(model=='univariate', unit_type=='hrr', term=='mean') %>%
+  interval_plot() +
+  ylab("Linear coeff., mean cons.") +
+  ggtitle("Linear univariate, HRR")
 ```
 
-Univariable linear regressions are in the Supplement. A crazy visualization of all the bug/drug combinations and resistances is also in the Supplement.
+```{r state_linear}
+results %>%
+  filter(model=='univariate', unit_type=='state', term=='mean') %>%
+  interval_plot() +
+  ylab("Linear coeff., mean cons.") +
+  ggtitle("Linear univariate, state")
+```
 
@@
 cbbPalette <- c("#000000", "#D55E00", "#CC79A7", "#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2")
-state_ineq %>%
-  ggplot(aes(x=cpb, y=nzg, color=drug_group)) +
+
+ineq %>%
+  filter(unit_type=='state', drug_group != 'overall') %>%
+  ggplot(aes(x=mean, y=nzgini, color=drug_group)) +
   geom_point() +
   ylab('Gini coefficient among users') + xlab('claims per beneficiary') +
+  ggtitle('Unevenness and mean consumption, states') +
+  theme_minimal() +
+  scale_color_manual(values=cbbPalette)
+```
+
+## Unevenness is mostly related to the "extensive margin"
+
+```{r nzgini_fnz}
+ineq %>%
+  filter(unit_type=='state', drug_group != 'overall') %>%
+  ggplot(aes(x=fnz, y=mean, color=drug_group)) +
+  geom_point() +
+  xlab('Fraction nonzero') + ylab('claims per beneficiary') +
+  ggtitle('Fraction consuming and mean consumption, states') +
   theme_minimal() +
@@
 
-Here we use the same bug/drug combinations, but instead we test the multivariable model mean percent nonsusceptible ~ fraction nonzero + nonzero Gini.
-
-There is only one significant and one marginally significant result:
-
-```{r multi_models}
+```{r hrr_multi}
 results %>%
-  filter(model=='multi', term==multi_term) %>%
-  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
-  filter(p.value < 0.05) %>%
-  arrange(p.value) %>%
-  select(bug, drug_group, estimate, p.value, p.adj) %>%
-  kable
+  filter(model=='multivariate', unit_type=='hrr', term=='nzgini') %>%
+  interval_plot() +
+  ylab("Linear coeff., nonzero Gini") +
+  ggtitle("Linear multivariate, HRR")
 ```
 
-```{r}
+```{r state_multi}
 results %>%
-  filter(model=='multi', term==multi_term) %>%
+  filter(model=='multivariate', unit_type=='state', term=='nzgini') %>%
   interval_plot() +
-  ylab('Gini coefficient')
-```
-
-Using HRRs, one of the answers remains the same, but a different marginally-significant result pops up.
-
-```{r}
-hrr_results %>%
-  filter(model=='multi', term==multi_term) %>%
-  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
-  filter(p.value < 0.05) %>%
-  arrange(p.value) %>%
-  select(bug, drug_group, estimate, p.value, p.adj) %>%
-  kable
-```
-
-```{r}
-hrr_results %>%
-  filter(model=='multi', term==multi_term) %>%
-  interval_plot()
+  ylab("Linear coeff., nonzero Gini") +
+  ggtitle("Linear multivariate, state")
 ```
@@
 
-Quinolone consumption predicts *E. coli* quinolone resistance:
+### E. coli & quinolones
 
-```{r ecoli_quin}
-left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
+```{r ecoli_quin_state}
+state_abg = read_tsv('abg_state.tsv')
+
+state_cons_res_plot = function(b, d) {
+  p = inner_join(
+    ineq %>%
+      filter(unit_type=='state', drug_group==d) %>%
+      select(state=unit, mean),
+    state_abg %>%
+      filter(bug==b, drug_group==d) %>%
+      select(state, mean_percent_nonsusceptible),
+    by='state') %>%
     left_join(regions, by='state') %>%
-  filter(bug=='E. coli', drug_group=='quinolone') %>%
-  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, label=state)) +
-  geom_smooth(method='lm', col='gray', se=FALSE) +
+    ggplot(aes(x=mean, y=mean_percent_nonsusceptible, label=state, color=region)) +
+    geom_smooth(method='lm', se=F, col='gray') +
     geom_point() +
     geom_text() +
-  xlim(0.2, 0.45) + ylim(0, 50) +
-  xlab('quinolone claims per beneficiary') + ylab('mean percent nonsusceptible E. coli isolates') +
-  theme_minimal()
-```
+    xlab(str_interp('mean ${d} claims per beneficiary')) +
+    ylab(str_interp('mean percent nonsusceptible ${b} isolates')) +
+    ggtitle(str_interp('${b} & ${d}, state'))
 
-## E.coli and bactrim
+  p
+}
 
-```{r ecoli_bactrim}
-ecoli_bactrim_dat = left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
-  left_join(regions, by='state') %>%
-  filter(bug=='E. coli', drug_group=='tmp_smx') #%>%
-  #filter(!is.na(mean_percent_nonsusceptible))
-
-ecoli_bactrim_dat %>%
-  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, label=state)) +
-  geom_smooth(method='lm', col='gray', se=FALSE) +
-  geom_point() +
-  geom_text() +
-  scale_x_continuous(limits=c(0.08, 0.2), breaks=c(0.08, 0.1, 0.15, 0.2)) +
-  ylim(10, 45) +
-  xlab('TMP/SMX claims per beneficiary') + ylab('mean percent nonsusceptible E. coli isolates') +
-  theme_minimal()
-```
-
-```{r}
-ecoli_bactrim_dat %>%
-  ggplot(aes(x=nzg, y=mean_percent_nonsusceptible, label=state)) +
-  geom_smooth(method='lm', col='gray', se=FALSE) +
-  geom_point() +
-  geom_text() +
-  xlab('Gini coefficient)') +
-  ylab('E. coli resistance') +
-  theme_minimal()
+show(state_cons_res_plot('E. coli', 'quinolone'))
 ```
@@
 
-```{r}
-left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
-  left_join(regions, by='state') %>%
-  filter(bug=='S. pneumoniae', drug_group=='macrolide') %>%
-  filter(!is.na(mean_percent_nonsusceptible)) %>% 
-  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, label=state)) +
-  geom_smooth(method='lm', col='gray', se=FALSE) +
-  geom_point() +
-  geom_text() +
-  theme_minimal()
+```{r pneumo_macro_state}
+show(state_cons_res_plot('S. pneumoniae', 'macrolide'))
 ```
 
-## E. cloacae and quinolones
-
-```{r}
-left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
-  left_join(regions, by='state') %>%
-  filter(bug=='Eb. cloacae', drug_group=='quinolone') %>%
-  filter(!is.na(mean_percent_nonsusceptible)) %>%
-  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, label=state)) +
-  geom_smooth(method='lm', col='gray', se=FALSE) +
-  geom_point() +
-  geom_text() +
-  theme_minimal()
-```
-
-## S. aureus and clindamycin
-
-```{r staph_clinda}
-left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
-  left_join(regions, by='state') %>%
-  filter(bug=='S. aureus', drug_group=='clindamycin') %>%
-  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, label=state)) +
-  geom_smooth(method='lm', col='gray', se=FALSE) +
-  geom_point() +
-  geom_text() +
-  xlab('clindamycin claims per beneficiary') + ylab('mean percent nonsusceptible S. aureus isolates') +
-  theme_minimal()
-```
+- *E. cloacae* and quinolone
+- *S. aureus* and clindamycin
 
@@
 
-## Amount of data vs. strength of correlation
+## Correlations between the metrics
 
-There is a correlation between the amount of data (i.e., the number of HRRs in which there is an antibiogram for a bug/drug combination) and the correlation coefficient for that relationship.
-The bug/drug combinations with more data tend to be the ones with more positive $\rho$ values.
-This makes it advantageous to the study's power to filter the poorly-represented combinations.
+```{r metrics}
+f = function(df, x, y) {
+  cor.test(df[[x]], df[[y]], method='spearman') %>%
+    tidy %>%
+    mutate(comparison=str_interp("${x}/${y}"))
+}
 
-```{r}
-abg %>% 
-  select(bug, drug_group, hrr) %>%
-  distinct() %>%
-  count(bug, drug_group) %>% ungroup() %>% rename(n_places=n) %>%
-  right_join(filter(hrr_results, model=='spearman')) %>%
-  ggplot(aes(x=n_places, y=estimate)) +
-  geom_point()
+x = ineq %>%
+  group_by(unit_type, drug_group) %>%
+  do(bind_rows(f(., 'mean', 'fnz'),
+               f(., 'fnz', 'gini'),
+               f(., 'mean', 'gini'),
+               f(., 'mean', 'nzgini'),
+               f(., 'fnz', 'nzgini'))) %>%
+  ungroup() %>%
+  select(comparison, unit_type, drug_group, estimate, p.value)
+
+comparison_plot_f = function(comp) {
+  p = x %>%
+    filter(comparison==comp) %>%
+    ggplot(aes(x=drug_group, y=estimate, color=unit_type)) +
+    geom_point() +
+    ylab('Spearman correlation between metrics') +
+    ylim(-1, 1) +
+    coord_flip() +
+    ggtitle(comp)
+  p
+}
+
+comparison_plot_f('mean/fnz') %>% show
+comparison_plot_f('fnz/gini') %>% show
+comparison_plot_f('mean/gini') %>% show
+comparison_plot_f('mean/nzgini') %>% show
+comparison_plot_f('fnz/nzgini') %>% show
 ```
 
-## Metrics of unevenness
+## Comparing state and HRR Spearman models
 
-### Aggregate usage correlates with fraction of users
-
-Error bars show the s.e.m. across the four data years.
-
-```{r fnz_mean}
-all_ineq %>%
-  mutate_at(vars(matches('fnz')), function(x) x*100) %>%
-  ggplot(aes(x=mean_cpb, xmin=mean_cpb-sem_cpb, xmax=mean_cpb+sem_cpb,
-             y=mean_fnz, ymin=mean_fnz-sem_fnz, ymax=mean_fnz+sem_fnz,
-             color=region, label=state_abbreviation)) +
-  geom_point() +
-  geom_linerange() +
-  geom_errorbarh(height=0) +
-  geom_text(nudge_x=0.04) +
-  xlab('mean claims per beneficiary') +
-  ylab('mean % beneficiaries with at least one claim') +
-  theme_minimal()
-```
-
-```{r}
-all_ineq %$%
-  cor.test(mean_cpb, mean_fnz, method='spearman')
-```
-
-### The two inequality metrics not correlate well
-
-Error bars show the s.e.m. across the four data years.
-
-```{r nzgini_mean}
-all_ineq %>%
-  ggplot(aes(x=mean_fnz, xmin=mean_fnz-sem_fnz, xmax=mean_fnz+sem_fnz,
-             y=mean_nzgini, ymin=mean_nzgini-sem_nzgini, ymax=mean_nzgini+sem_nzgini,
-             color=region, label=state_abbreviation)) +
-  geom_point() +
-  geom_linerange() +
-  geom_errorbarh(height=0) +
-  geom_text(nudge_x=0.00) +
-  xlab('mean fraction with at least one claim') +
-  ylab('mean Gini coefficient') +
-  theme_minimal()
-```
-
-The correlation exists but it's much less strong.
-
-```{r}
-all_ineq %$%
-  cor.test(mean_fnz, mean_nzgini, method='spearman')
-```
-
-There is an intermediate correlation between claims per beneficiary and the Gini coefficient (see below), so in multivariable regressions I will use those two (rather than CPB and Gini).
-
-```{r}
-all_ineq %$% cor.test(mean_cpb, mean_nzgini, method='spearman')
-```
-
-## State-level Spearman models
-
-```{r state_spearman_table}
+```{r spearman_state_hrr_compare}
 results %>%
   filter(model=='spearman') %>%
-  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
-  filter(p.value < 0.05) %>%
-  arrange(p.adj) %>%
-  select(bug, drug_group, estimate, p.value, p.adj, n) %>%
-  kable
+  mutate(group=interaction(bug, drug_group)) %>%
+  ggplot(aes(x=estimate, y=-log10(p.value), color=unit_type, group=group)) +
+  geom_point() +
+  geom_line() +
+  geom_hline(yintercept=-log10(0.05))
 ```
-
-```{r state_spearman_interval}
-results %>%
-  filter(model=='spearman') %>%
-  interval_plot() +
-  ylab("Spearman's rho")
-```
-
-## Comparing Spearman models state and HRR
-
-HRR-level Spearman correlations between consumption and resistance (almost) always underestimate the correlation reported at the state level:
-
-```{r}
-bind_rows(mutate(results, group='state'),
-          mutate(hrr_results, group='hrr')) %>%
-  filter(model=='spearman') %>%
-  arrange(estimate) %>%
-  mutate(bug_drug=interaction(bug, drug_group)) %>%
-  select(bug_drug, group, estimate) %>%
-  spread(group, estimate) %>%
-  ggplot(aes(x=state, y=hrr, label=bug_drug)) +
-  geom_point() +
-  geom_text() +
-  geom_abline()
-```
-
-## Univariable regressions
-
-Like the Spearman models, the model here is `mean percent nonsusceptible ~ claims per beneficiary`, only this time it's a normal linear regression. At the HRR level:
-
-```{r hrr_uni_table}
-hrr_results %>%
-  filter(model=='uni', term==uni_term) %>%
-  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
-  filter(p.value < 0.05) %>%
-  arrange(p.value) %>%
-  select(bug, drug_group, estimate, p.value, p.adj) %>%
-  kable
-```
-
-```{r hrr_uni_interval}
-hrr_results %>%
-  filter(model=='uni', term==uni_term) %>%
-  interval_plot()
-```
-
-And at the state level:
-
-```{r state_uni_interval}
-results %>%
-  filter(model=='uni', term==uni_term) %>%
-  interval_plot()
-```
-
-```{r state_uni_table}
-results %>%
-  filter(model=='uni', term==uni_term) %>%
-  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
-  filter(p.value < 0.05) %>%
-  arrange(p.value) %>%
-  select(bug, drug_group, estimate, p.value, p.adj, n) %>%
-  kable
-```
-
-## Consumption/resistance patterns
-
-Different bug/drug combinations have different spreads along the consumption and resistance axes, so we would expect to see different strengths of association for each one.
-
-```{r consumption_resistance_patterns}
-left_join(state_abg, state_ineq, by=c('state', 'drug_group')) %>%
-  filter(bug %in% c('E. coli', 'Eb. cloacae', 'CoNS', 'MSSA', 'MRSA', 'S. pneumoniae')) %>%
-  ggplot(aes(x=cpb, y=mean_percent_nonsusceptible, color=drug_group, shape=bug)) +
-  geom_point() +
-  xlab('claims per beneficiary')
-```

### 2017-08-10 ms2/ms2_2analyze.R

@@ -1,22 +1,8 @@
 # Consumption data
 ineq = read_tsv('ineq.tsv') %>%
   group_by(drug_group, unit_type, unit) %>%
-  summarize_at(vars(mean, fnz, nzgini), mean) %>%
+  summarize_at(vars(mean, fnz, nzgini, gini), mean) %>%
   ungroup()
-
-# # Aggregate usage correlates with fraction of users
-# p = ggplot(all_ineq, aes(x=mean, y=fnz, color=factor(year))) +
-#   geom_point() +
-#   xlab('claims per beneficiary') +
-#   ylab('fraction of beneficiaries w/ >= 1 claim') +
-#   theme_minimal()
-
-# # Inequality among users does not correlate well
-# p = ggplot(all_ineq, aes(x=mean, y=nzgini, color=factor(year))) +
-#   geom_point() +
-#   xlab('claims per beneficiary') +
-#   ylab('Gini coefficient among benes. w/ >= 1 claim') +
-#   theme_minimal()
 
 # Resistance data
 resistance_groups = read_tsv('resistance_groups.tsv')
@@ -48,9 +34,9 @@ good_combos = raw_abg %>%
   semi_join(possible_bug_drug, by=c('bug', 'drug_group')) %>%
   group_by(bug, drug_group) %>%
   summarize_at(vars(state, hrr), function(x) length(unique(x))) %>%
-  ungroup() %>%
-  filter(state >= 35) %>%
-  select(bug, drug_group)
+  ungroup() #%>%
+  #filter(state >= 35) %>%
+  #select(bug, drug_group)
 
 abg = raw_abg %>%
   semi_join(good_combos, by=c('bug', 'drug_group')) %>%
@@ -75,8 +61,8 @@ summarize_abg = function(df, place_name) {
     ungroup()
 }
 
-state_abg = summarize_abg(abg, 'state')
-hrr_abg = summarize_abg(abg, 'hrr')
+state_abg = summarize_abg(abg, 'state') %T>% write_tsv('abg_state.tsv')
+hrr_abg = summarize_abg(abg, 'hrr') %T>% write_tsv('abg_hrr.tsv')
 
 linear_model = function(df, y, xs) {
   frmla = as.formula(str_interp("${y} ~ ${str_c(xs, collapse=' + ')}"))
@@ -121,7 +107,7 @@ hrr_results = ineq %>%
   mutate(hrr=as.integer(unit)) %>%
[…]

### 2017-08-11 ms2/ms2_display.Rmd
updated ms2 display

--- 2017-08-10 ms2/ms2.Rmd
@@
 
+I use a few metrics: the mean consumption $\mu$, the fraction nonzero $f_+$ (i.e., the fraction who are consumers), the Gini coefficient $G$, and the nonzero Gini $G_+$ (i.e., the Gini coefficient ignoring the nonconsumers).
+Here I explore the relationship between these.
+
+Some math shows that
+$$
+G = (1 - f_+) + f_+ G_+,
+$$
+i.e., the Gini coefficient is a mixture of a "subcoefficient" of 1 for the nonconsumers and $G_+$ for the consumers.
+
+In the plots below, I compare the correlations between these metrics across states and HRRs. Blessedly, the correlations are very similar across both sets of geographic units.
+
 ```{r metrics}
@@
 }
+```
 
+The mean and the fraction nonzero are highly correlated ($\rho \sim 1$) for all drug groups:
+
+```{r mean_fnz}
 comparison_plot_f('mean/fnz') %>% show
+```
+
+In contrast, the fraction nonzero is almost perfectly *anticorrelated* with the Gini coefficient, i.e., the more people consuming, the more even the consumption:
+
+```{r fnz_gini}
 comparison_plot_f('fnz/gini') %>% show
+```
+
+$\mu$ is very well correlated with $f_+$, and $f_+$ is very well anticorrelated with $G$, so, not surprisingly, the anticorrelation between $\mu$ and $G$ is also strong:
+
+```{r mean_gini}
 comparison_plot_f('mean/gini') %>% show
+```
+
+However, mean consumption $\mu$ and the nonzero Gini coefficient $G_+$ are not as well correlated with one another.
+For most drug groups, increased mean consumption is somewhat correlated with increased unevenness among consumers.
+For two drug groups, the relationship is near zero:
+
+```{r mean_nzgini}
 comparison_plot_f('mean/nzgini') %>% show
+```
+
+When replacing $\mu$ with $f_+$ (recall that the two are closely correlated), all correlations get more negative, i.e., the positive ones get less positive and the near-zero ones get negative:
+
+```{r fnz_nzgini}
 comparison_plot_f('fnz/nzgini') %>% show
@@
 
+Each point shows the $\rho$ and $p$-value from a Spearman correlation test for a bug/drug combination in a geographic unit (state or HRR).
+Gray lines connect the two points representing the test for the same bug/drug combination; one point is for states, the other for HRRs.
+
+It seems like the states generally have more positive estimates.
+I wonder if that means that states are the "correct" geographical level at which to look at these effects, but there just aren't enough of them to be able to make a sufficiently small $p$-value.
+
+The dumbbells that cross the black, $p = 0.05$ line are the ones that are significant for one geographical unit but not the other.
+
 ```{r spearman_state_hrr_compare}
@@
   filter(model=='spearman') %>%
-  mutate(group=interaction(bug, drug_group)) %>%
-  ggplot(aes(x=estimate, y=-log10(p.value), color=unit_type, group=group)) +
+  mutate(group=interaction(bug, drug_group),
+         adj.p.value=p.adjust(p.value, method='BH')) %>%
+  ggplot(aes(x=estimate, y=-log10(adj.p.value), color=unit_type, group=group)) +
+  geom_line(color='gray') +
+  geom_hline(yintercept=-log10(0.05)) +
   geom_point() +
-  geom_line() +
-  geom_hline(yintercept=-log10(0.05))
+  xlab("Spearman's rho")
 ```

### 2017-08-14 ms1/ms1_display.Rmd
adjusted OR dx

@@
 
+I also performed a similar analysis using logistic regression.
+I predicted whether a beneficiary would have a diagnosis at any point in the year.
+The main predictor variable was year, and the covariates were as elsewhere.
+
+```{r adjusted_dx_or}
+read_tsv('tables/model_dx.tsv') %>%
+  filter(term=='year0') %>%
+  model_ci() %>%
+  select(diagnosis_type, change_str) %>%
+  kable
+```
+
 ## Delays between diagnosis and prescription

### 2017-08-14 ms1/ms1_2process.R

@@ -164,6 +164,30 @@ dxrx = dx_raw %>%
   mutate(dummy=TRUE) %>% spread(antibiotic, dummy, fill=FALSE) %>%
   left_join(bene, by=c('year', 'bene_id'))
 
+dx_types = unique(dxrx$diagnosis_type)
+dx_replace = rep(FALSE, length(dx_types)) %>% as.list %>% setNames(dx_types)
+
+dxrx_binary = dxrx %>%
+  select(year, bene_id, diagnosis_type) %>%
+  distinct() %>%
+  mutate(dummy=TRUE) %>%
+  spread(diagnosis_type, dummy) %>%
+  right_join(bene, by=c('year', 'bene_id')) %>%
+  replace_na(dx_replace)
+
+# adjusted OR for diagnoses
+# i.e., logistic regression bene_had_dx_in_year? ~ year + covariates
+dx_adjust_f = function(dxrx, dx) {
+  frmla = as.formula(str_interp("${dx} ~ year0 + age0 + n_cc + is_white + is_dual + is_female + region"))
+  glm(frmla, data=dxrx, family='binomial') %>%
+    tidy %>%
+    mutate(diagnosis_type=dx)
+}
+
+dx_res = lapply(dx_types, function(x) dx_adjust_f(dxrx_binary, x)) %>%
+  bind_rows() %T>%
+  output_table('model_dx')
+
 # for a diagnosis type dx and an antibiotic, what is the trend in the probability
 # that that drug will be used for that diagnosis?
 # i.e., logistic regression drug_used? ~ dx_present?*year + covariates
@@ -175,7 +199,7 @@ dxrx_f = function(dxrx, dx, abx) {
     filter_f = function(df) filter(df, diagnosis_type==dx)
   }
   
-  frmla = as.formula(str_interp("`${abx}` ~ year0 + age0 + is_white + is_dual + is_female + region"))
+  frmla = as.formula(str_interp("`${abx}` ~ year0 + age0 + n_cc + is_white + is_dual + is_female + region"))
   
   dxrx %>%
     filter_f() %>%
@@ -184,7 +208,7 @@ dxrx_f = function(dxrx, dx, abx) {
     mutate(diagnosis_type=dx, antibiotic=abx)
 }
 
-dxrx_res = crossing(diagnosis_type=unique(dxrx$diagnosis_type),
+dxrx_res = crossing(diagnosis_type=dx_types,
[…]

### 2017-08-15 ms2/ms2_1summarize.R
ms2 fixed bug names

@@ -1,10 +1,15 @@
 #!/usr/bin/env Rscript
 
 # utility functions function
-import::from(ineq, ineq)
+gini = function(x) ineq::ineq(x, type='Gini')
 nonzero = function(x) x[x > 0]
 fraction_nonzero = function(x) length(nonzero(x)) / length(x)
 pad0 = function(x, to) c(x, rep(0, to - length(x)))
+
+without_min = function(x) x[-which.min(x)]
+without_max = function(x) x[-which.max(x)]
+without_1 = function(x) x[-match(1, x)]
+
 filter_eq_ = function(df, col, val) filter_(df, str_interp("${col}=='${val}'"))
 
 inequalities = function(pde, drug_group, unit_type, unit, n_bene) {
@@ -23,6 +28,7 @@ inequalities = function(pde, drug_group, unit_type, unit, n_bene) {
   }
 
   x = pad0(filtered_pde$n_claims, n_bene)
+  nzx = nonzero(x)
 
   nb_par = fitdistrplus::mledist(x, 'nbinom')$estimate
   nb_size = nb_par['size']
@@ -36,9 +42,13 @@ inequalities = function(pde, drug_group, unit_type, unit, n_bene) {
              n_bene=n_bene,
              total=sum(x),
              mean=mean(x),
+             mean_without_1=mean(without_1(x)),
+             mean_without_max=mean(without_max(x)),
              fnz=fraction_nonzero(x),
-             gini=ineq(x, type='Gini'),
-             nzgini=ineq(nonzero(x), type='Gini'),
+             gini=gini(x),
+             nzgini=gini(nzx),
+             nzgini_wo_min=gini(without_min(nzx)),
+             nzgini_wo_max=gini(without_max(nzx)),
              nb_size=nb_size,
              nb_prob=nb_prob)
 }
@@ -94,6 +104,14 @@ summarize_inequality = function(y) {
     mutate(year=y)
 }
 
-lapply(2011:2014, summarize_inequality) %>%
+# prepare for parallel execution
+library(parallel)
+n_cores = detectCores() - 1
+cluster = makeCluster(n_cores, type='FORK')
+
+#lapply(2011:2014, summarize_inequality) %>%
+parLapply(cluster, 2011:2014, summarize_inequality) %>%
   bind_rows() %>%
   write_tsv('ineq.tsv')
+
+stopCluster(cluster)

### 2017-08-15 ms2/ms2_2analyze.R

@@ -115,7 +115,7 @@ state_results = ineq %>%
   rename(state=unit) %>%
   right_join(state_abg, by=c('state', 'drug_group')) %>%
   group_by(bug, drug_group) %>%
-  do(models(., 'mean_percent_nonsusceptible', 'mean', c('fnz', 'nzgini'))) %>%
+  do(models(., 'mean_percent_nonsusceptible', 'mean', c('mean', 'nzgini'))) %>%
   ungroup()
 
 bind_rows(

### 2017-08-16 ms2/ms2_display.Rmd
ms2 fixed bug names
ms2 display

--- 2017-08-10 ms2/ms2.Rmd
@@
   group_by(drug_group, unit_type, unit) %>%
-  summarize_at(vars(mean, fnz, nzgini, gini), mean) %>%
+  summarize_at(vars(mean, fnz), mean) %>%
   ungroup()
@@
   left_join(regions, by='state')
+
+hrr_out = hrr %>%
+  count(hrr, region) %>%
+  filter(n==max(n)) %>%
+  ungroup() %>%
+  select(hrr, region)
+
+state_abg = read_tsv('abg_state.tsv')
+hrr_abg = read_tsv('abg_hrr.tsv')
 ```
@@
 
-(In previous iterations, I filtered by the number of states/HRRs in which each bug/drug combination appeared.
-Weirdly, there is a relationship in the data, that better-represented bug/drug combinations are the ones with the stronger relationships [i.e., the higher $\rho$, not necessarily the lower $p$].
-When using the HRR data, that kind of fancy footwork wasn't necessary, so I just didn't do it.)
-
 For each combination, I computed three models:
@@
 
-Two stars after the bug/drug combination implies that the adjusted $p < 0.05$, and one star means the unadjusted $p < 0.05$. Intervals show 95% confidence intervals.
+The plot shows the Spearman correlation coefficient $\rho$ between resistance $R$ and fraction $f_>$ of beneficiaries who are consumers.
+Two stars after the bug/drug combination implies that the adjusted $p < 0.05$, and one star means the unadjusted $p < 0.05$.
+Intervals show 95% confidence intervals.
 
@@
 
+I run models $R \sim f_>$, that is $R^{(u)} = \beta_0 + \beta_{f_>} f_>^{(u)} + \varepsilon^{(u)}$, where the superscript $u$ indicates the geographical unit (HRR or state).
+The plots show $\beta_{f_>}$.
+This model is equivalent to one in which the population resistance $R^{(u)}$ in unit $u$ is the mean of the resistance $r_i^{(u)}$ in the individual people $i$ in unit $u$, where $r_i^{(u)}$ is $\beta_0$ if person $i$ took no drug and $\beta_0 + \beta_{f_>}$ if that person took at least one dose.
+
+The gray line shows $\beta_{f_>} = 1$.
+
 ```{r hrr_linear}
 results %>%
-  filter(model=='univariate', unit_type=='hrr', term=='mean') %>%
+  filter(model=='univariate', unit_type=='hrr', term=='fnz') %>%
   interval_plot() +
-  ylab("Linear coeff., mean cons.") +
+  geom_hline(yintercept=1, col='gray') +
+  ylab(expression(beta[f])) +
   ggtitle("Linear univariate, HRR")
@@
 results %>%
-  filter(model=='univariate', unit_type=='state', term=='mean') %>%
+  filter(model=='univariate', unit_type=='state', term=='fnz') %>%
   interval_plot() +
-  ylab("Linear coeff., mean cons.") +
+  ylab(expression(beta[f])) +
   ggtitle("Linear univariate, state")
@@
 
-## Drugs are consumed differently in different places
+## Unevenness is mostly related to the "extensive margin"
 
-```{r nzgini_cpb_drug}
+```{r fnz_mean_drug}
 cbbPalette <- c("#000000", "#D55E00", "#CC79A7", "#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2")
 
-ineq %>%
-  filter(unit_type=='state', drug_group != 'overall') %>%
-  ggplot(aes(x=mean, y=nzgini, color=drug_group)) +
-  geom_point() +
-  ylab('Gini coefficient among users') + xlab('claims per beneficiary') +
-  ggtitle('Unevenness and mean consumption, states') +
-  theme_minimal() +
-  scale_color_manual(values=cbbPalette)
-```
-
-## Unevenness is mostly related to the "extensive margin"
-
-```{r nzgini_fnz}
 ineq %>%
@@
 
+## The South has more consumers per capita
+
+However, among those people who do consume, consumption is smaller per capita (i.e., the slope is similar but the intercept is smaller).
+In this sense, consumption in the South is less concentrated than in the Midwest.
+
+```{r fnz_mean_region}
+ineq %>%
+  filter(unit_type=='state', drug_group == 'overall') %>%
+  rename(state=unit) %>%
+  filter(state != 'DC') %>%
+  left_join(regions, by='state') %>%
+  ggplot(aes(x=fnz, y=mean, color=region)) +
+  geom_smooth(aes(color=NULL), method='lm', se=F, color='gray') +
+  geom_point() +
+  xlab('Fraction nonzero') + ylab('claims per beneficiary') +
+  ggtitle('Fraction consuming and mean consumption, states') +
+  theme_minimal()
+```
+
 ## Effect of unevenness on antibiotic resistance
+
+We ask if mean consumption $\mu$ has any information beyond fraction nonzero $f_>$ for predicting resistance $R$.
+We might expect that the simple two-part model represented by the regressions $R \sim f_>$ doesn't capture the fact that taking two doses should have a greater effect than taking one, etc.
+
+These plots show the coefficient $\beta_\mu$ from the regression $R^{(u)} = \beta_0 + \beta_{f_>} f_>^{(u)} + \beta_{\mu} \mu^{(u)} + \varepsilon^{(u)}$.
 
@@
 results %>%
-  filter(model=='multivariate', unit_type=='hrr', term=='nzgini') %>%
+  filter(model=='multivariate', unit_type=='hrr', term=='mean') %>%
   interval_plot() +
-  ylab("Linear coeff., nonzero Gini") +
+  ylab(expression(beta[mu])) +
   ggtitle("Linear multivariate, HRR")
@@
 results %>%
-  filter(model=='multivariate', unit_type=='state', term=='nzgini') %>%
+  filter(model=='multivariate', unit_type=='state', term=='mean') %>%
   interval_plot() +
-  ylab("Linear coeff., nonzero Gini") +
+  ylab(expression(beta[mu])) +
   ggtitle("Linear multivariate, state")
+```
+
+There are a few significant results:
+
+- *E. faecium* and $\beta$-lactams have a postive relationship (implying higher mean consumption, given the same fraction of people consuming, goes with more resistance)
+- *E. coli* and Bactrim have a *negative* relationship (higher consumption goes with *less* resistance)
+- *S. pneumoniae* and tetracyclines have a negative relationship
+
+And some marginal results:
+
+- *E. coli* and $\beta$-lactams, negative relationship, HRR only
+- Non-MRSA *S. aureus* and first-generation cephalosporins, negative, HRR only
+- *E. faecium* and nitro, negative, both HRR and state
+- *S. aureus* and clinda, positive, state only
+- *S. aureus* and tetracyclines, positive, state only
+- *E. cloacae* and quinolones, positive, state only
+
+The positive relationships are not troubling: it means that it might be that increasing consumption among the consumers can increase resistance.
+The negative relationships are troubling: how could decreasing consumption among consumers increase resistance?
+
+### E. faecium and beta-lactams
+
+This one doesn't worry me.
+First, the univariate regression $R \sim f_>$ isn't significant.
+Second, the ANOVA gives different results depending on the order of the variables: if $\mu$ goes first, it is non-significant but $f_>$, going second, is; conversely, if $f_>$ goes first, then $\mu$ is significant.
+So I think this is a weird case about collinearity.
+
+```{r efaecium_betalactams}
+ineq %>%
+  filter(unit_type=='hrr') %>%
+  mutate(hrr=as.integer(unit)) %>% left_join(hrr_out, by='hrr') %>%
+  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
+  filter(bug=='Ec. faecium', drug_group=='beta_lactam') %>%
+  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region)) +
+  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
+  geom_smooth(method='lm', se=F) +
+  geom_point() +
+  xlab('fraction who consumed beta-lactams') +
+  ylab('Beta-lactam resistance E. faecium') +
+  ggtitle('E. faecium & beta-lactams across HRRs')
+
+ineq %>%
+  filter(unit_type=='state') %>%
+  rename(state=unit) %>%
+  left_join(regions, by='state') %>%
+  right_join(state_abg, by=c('state', 'drug_group')) %>%
+  filter(bug=='Ec. faecium', drug_group=='beta_lactam') %>%
+  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region)) +
+  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
+  geom_smooth(method='lm', se=F) +
+  geom_point() +
+  xlab('fraction who consumed beta-lactams') +
+  ylab('Beta-lactam resistance E. faecium') +
+  ggtitle('E. faecium & beta-lactams across states')
+```
+
+### E. coli and Bactrim
+
+From the Spearman and univariate models, Bactrim resistance in *E.
+coli* is correlated with increased resistance (black line).
+When breaking that apart by region, however, there are different relationship (colored lines).
+
+```{r ecoli_bactrim}
+ineq %>%
+  filter(unit_type=='hrr') %>%
+  mutate(hrr=as.integer(unit)) %>% left_join(hrr_out, by='hrr') %>%
+  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
+  filter(bug=='E. coli', drug_group=='tmp_smx') %>%
+  filter(mean_percent_nonsusceptible < 50) %>% # filter one crazy Nevada point
+  ggplot(aes(x=fnz, y=mean_percent_nonsusceptible, color=region)) +
+  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
+  geom_smooth(method='lm', se=F) +
+  geom_point() +
+  xlab('fraction who consumed Bactrim') +
+  ylab('Bactrim-resistance E. coli') +
+  ggtitle('E. coli & Bactrim across HRRs')
+
+ineq %>%
+  filter(unit_type=='state') %>%
+  rename(state=unit) %>%
+  left_join(regions, by='state') %>%
+  right_join(state_abg, by=c('state', 'drug_group')) %>%
+  filter(bug=='E. coli', drug_group=='tmp_smx') %>%
+  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region)) +
+  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
+  geom_smooth(method='lm', se=F) +
+  geom_point() +
+  xlab('fraction who consumed Bactrim') +
+  ylab('Bactrim resistance E. coli') +
+  ggtitle('E. coli & Bactrim across states')
+```
+
+### S. pneumonia and tetracyclines
+
+This one doesn't bother me either.
+Tetracycline resistance in pneumococcus not significantly related ot tetracycline consumption in the univariate models; the coefficient is almost exactly zero.
+If you run a regression $R \sim \mu$, you get a non-significant result.
+So you only get this relationship *after* you correct for an approximately flat relationship.
+Not promising.
+
+```{r pneumo_tetra}
+ineq %>%
+  filter(unit_type=='hrr') %>%
+  mutate(hrr=as.integer(unit)) %>% left_join(hrr_out, by='hrr') %>%
+  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
+  filter(bug=='S. pneumoniae', drug_group=='tetracycline') %>%
+  filter(mean_percent_nonsusceptible < 50) %>% # filter one crazy Nevada point
+  ggplot(aes(x=fnz, y=mean_percent_nonsusceptible, color=region)) +
+  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
+  geom_smooth(method='lm', se=F) +
+  geom_point() +
+  xlab('fraction who consumed Bactrim') +
+  ylab('Bactrim-resistance E. coli') +
+  ggtitle('S. pneumonia & Bactrim across HRRs')
 ```
@@
 ```{r ecoli_quin_state}
-state_abg = read_tsv('abg_state.tsv')
-
 state_cons_res_plot = function(b, d) {
@@
     ggplot(aes(x=mean, y=mean_percent_nonsusceptible, label=state, color=region)) +
+    geom_smooth(method='lm', se=F) +
     geom_smooth(method='lm', se=F, col='gray') +
@@
 
-## Pneumococcus and macrolides
+### Pneumococcus and macrolides
 
@@
 
+The mean and the fraction nonzero are highly correlated ($\rho \sim 1$) for all drug groups:
+
 ```{r metrics}
 f = function(df, x, y) {
-  cor.test(df[[x]], df[[y]], method='spearman') %>%
-    tidy %>%
-    mutate(comparison=str_interp("${x}/${y}"))
+  cor.test(df[[x]], df[[y]], method='spearman') %>% tidy
 }
 
-x = ineq %>%
+ineq %>%
   group_by(unit_type, drug_group) %>%
-  do(bind_rows(f(., 'mean', 'fnz'),
-               f(., 'fnz', 'gini'),
-               f(., 'mean', 'gini'),
-               f(., 'mean', 'nzgini'),
-               f(., 'fnz', 'nzgini'))) %>%
+  do(f(., 'fnz', 'mean')) %>%
   ungroup() %>%
-  select(comparison, unit_type, drug_group, estimate, p.value)
-
-comparison_plot_f = function(comp) {
-  p = x %>%
-    filter(comparison==comp) %>%
+  select(unit_type, drug_group, estimate, p.value) %>%
   ggplot(aes(x=drug_group, y=estimate, color=unit_type)) +
@@
   ylab('Spearman correlation between metrics') +
-    ylim(-1, 1) +
-    coord_flip() +
-    ggtitle(comp)
-  p
-}
-
-comparison_plot_f('mean/fnz') %>% show
-comparison_plot_f('fnz/gini') %>% show
-comparison_plot_f('mean/gini') %>% show
-comparison_plot_f('mean/nzgini') %>% show
-comparison_plot_f('fnz/nzgini') %>% show
+  coord_flip()
 ```

### 2017-08-16 ms2/ms2_display.Rmd
ms2 fewer plots

@@
 
-```{r load_data}
+```{r load_data, cache.extra=c(md5sum('ineq.tsv'), md5sum('results.tsv'))}
 ineq = read_tsv('ineq.tsv') %>%
@@
 results %>%
-  filter(model=='univariate', unit_type=='hrr', term=='fnz') %>%
+  filter(model=='univariate_fnz', unit_type=='hrr', term=='fnz') %>%
+  interval_plot() +
+  geom_hline(yintercept=1, col='gray') +
+  ylab(expression(beta[f])) +
+  ggtitle("Linear univariate, HRR")
+```
+
+```{r}
+results %>%
+  filter(model=='univariate_mean', unit_type=='hrr', term=='mean') %>%
   interval_plot() +
@@
 results %>%
-  filter(model=='univariate', unit_type=='state', term=='fnz') %>%
+  filter(model=='univariate_fnz', unit_type=='state', term=='fnz') %>%
   interval_plot() +
@@
 
+- *E. coli* and Bactrim have a *negative* relationship (higher consumption goes with *less* resistance)
 - *E. faecium* and $\beta$-lactams have a postive relationship (implying higher mean consumption, given the same fraction of people consuming, goes with more resistance)
-- *E. coli* and Bactrim have a *negative* relationship (higher consumption goes with *less* resistance)
 - *S. pneumoniae* and tetracyclines have a negative relationship
+
+The second two, *E.
+faecium* with $\beta$-lactams and pneumococcus with tetracyclines, are not problematic: in both cases, neither $R \sim f_>$ nor $R \sim \mu$ is significant.
+We're only interested in cases where $f_>$ gives information and $\mu$ gives information beyond that.
+There are cases where $f_>$ doesn't appear to give any information.
 
@@
 
-- *E. coli* and $\beta$-lactams, negative relationship, HRR only
-- Non-MRSA *S. aureus* and first-generation cephalosporins, negative, HRR only
-- *E. faecium* and nitro, negative, both HRR and state
-- *S. aureus* and clinda, positive, state only
-- *S. aureus* and tetracyclines, positive, state only
-- *E. cloacae* and quinolones, positive, state only
-
-The positive relationships are not troubling: it means that it might be that increasing consumption among the consumers can increase resistance.
-The negative relationships are troubling: how could decreasing consumption among consumers increase resistance?
-
-### E. faecium and beta-lactams
-
-This one doesn't worry me.
-First, the univariate regression $R \sim f_>$ isn't significant.
-Second, the ANOVA gives different results depending on the order of the variables: if $\mu$ goes first, it is non-significant but $f_>$, going second, is; conversely, if $f_>$ goes first, then $\mu$ is significant.
-So I think this is a weird case about collinearity.
-
-```{r efaecium_betalactams}
-ineq %>%
-  filter(unit_type=='hrr') %>%
-  mutate(hrr=as.integer(unit)) %>% left_join(hrr_out, by='hrr') %>%
-  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
-  filter(bug=='Ec. faecium', drug_group=='beta_lactam') %>%
-  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region)) +
-  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
-  geom_smooth(method='lm', se=F) +
-  geom_point() +
-  xlab('fraction who consumed beta-lactams') +
-  ylab('Beta-lactam resistance E. faecium') +
-  ggtitle('E. faecium & beta-lactams across HRRs')
-
-ineq %>%
-  filter(unit_type=='state') %>%
-  rename(state=unit) %>%
-  left_join(regions, by='state') %>%
-  right_join(state_abg, by=c('state', 'drug_group')) %>%
-  filter(bug=='Ec. faecium', drug_group=='beta_lactam') %>%
-  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region)) +
-  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
-  geom_smooth(method='lm', se=F) +
-  geom_point() +
-  xlab('fraction who consumed beta-lactams') +
-  ylab('Beta-lactam resistance E. faecium') +
-  ggtitle('E. faecium & beta-lactams across states')
-```
+- *E. coli* and $\beta$-lactams, negative relationship, HRR only; significant in univariate regressions $R \sim f_>$
+- *E. cloacae* and quinolones, positive, state only; marginally significant in univariate
+- Non-MRSA *S. aureus* and first-generation cephalosporins, negative, HRR only; but not significant in univariate
+- *E. faecium* and nitro, negative, both HRR and state; not significant in univariate
+- *S. aureus* and clinda, positive, state only; not significant in univariate
+- *S. aureus* and tetracyclines, positive, state only; not significant in univariate
 
@@
 
-From the Spearman and univariate models, Bactrim resistance in *E.
+This negative relationship is the most tricky.
+
+First, the regression results are equivocal.
+Regressing $r ~ \mu$, where $r$ are the residuals from the regression $R ~ f_>$, gives a non-significant result.
+An ANOVA analysis shows that sequential application of $f_>$ and $\mu$ as predictors of $R$ both explain non-trivial variance; but if the order is reversed, and $\mu$ is used first, then it *doesn't* explain any variance.
+This makes me think that there is just some weird interaction between the two terms in the multiple regression.
+
+Second, from the Spearman and univariate models, Bactrim resistance in *E.
 coli* is correlated with increased resistance (black line).
 When breaking that apart by region, however, there are different relationship (colored lines).
+This makes it seem like there is something strange going on.
 
@@
 
-### S. pneumonia and tetracyclines
+### E. coli and beta-lactams
 
-This one doesn't bother me either.
-Tetracycline resistance in pneumococcus not significantly related ot tetracycline consumption in the univariate models; the coefficient is almost exactly zero.
-If you run a regression $R \sim \mu$, you get a non-significant result.
-So you only get this relationship *after* you correct for an approximately flat relationship.
-Not promising.
+This negative relationship was significant only for HRRs, not states.
+Not surprisingly, there is something weird going on with the HRRs: the Southern HRRs are all above the $R \sim f_>$ line, suggesting that the relationship with the mean is just a predictor about region.
 
-```{r pneumo_tetra}
+```{r ecoli_betalactam}
 ineq %>%
@@
   right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
-  filter(bug=='S. pneumoniae', drug_group=='tetracycline') %>%
+  filter(bug=='E. coli', drug_group=='beta_lactam') %>%
   filter(mean_percent_nonsusceptible < 50) %>% # filter one crazy Nevada point
@@
   geom_point() +
-  xlab('fraction who consumed Bactrim') +
-  ylab('Bactrim-resistance E. coli') +
-  ggtitle('S. pneumonia & Bactrim across HRRs')
+  xlab('fraction who consumed beta-lactams') +
+  ylab('Beta-lactam resistance in E. coli') +
+  ggtitle('E. coli & beta-lactams across HRRs')
+
+ineq %>%
+  filter(unit_type=='state') %>%
+  rename(state=unit) %>%
+  left_join(regions, by='state') %>%
+  right_join(state_abg, by=c('state', 'drug_group')) %>%
+  filter(bug=='E. coli', drug_group=='beta_lactam') %>%
+  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region)) +
+  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
+  geom_smooth(method='lm', se=F) +
+  geom_point() +
+  xlab('fraction who consumed beta-lactams') +
+  ylab('Beta-lactam resistance E. coli') +
+  ggtitle('E. coli & beta-lactams across states')
 ```

### 2017-08-16 ms2/ms2_2analyze.R

@@ -56,7 +56,7 @@ if (any(is.na(abg$bug))) stop('missing bug names')
 
 summarize_abg = function(df, place_name) {
     group_by_(df, 'bug', 'drug_group', place_name) %>%
-    summarize(mean_percent_nonsusceptible=weighted.mean(percent_nonsusceptible, n_isolates),
+    summarize(mean_percent_nonsusceptible=weighted.mean(percent_nonsusceptible, sqrt(n_isolates)),
               n_place_antibiograms=n()) %>%
     ungroup()
 }
@@ -101,19 +101,21 @@ spearman_model = function(df, y_name, x_name) {
              p.value=m$p.value)
 }
 
-models = function(df, y, uni_x, multi_xs) {
+models = function(df) {
   bind_rows(
-    linear_model(df, y, uni_x) %>% mutate(model='univariate'),
-    spearman_model(df, y, uni_x) %>% mutate(model='spearman'),
-    linear_model(df, y, multi_xs) %>% mutate(model='multivariate')
+    linear_model(df, 'y', 'mean') %>% mutate(model='univariate_mean'),
+    linear_model(df, 'y', 'fnz') %>% mutate(model='univariate_fnz'),
+    spearman_model(df, 'y', 'fnz') %>% mutate(model='spearman'),
+    linear_model(df, 'y', c('fnz', 'mean')) %>% mutate(model='multivariate')
   ) %>%
     mutate(n_data=nrow(df))
 }
 
 hrr_out = hrr %>%
-  select(hrr, state, region) %>%
-  group_by(hrr) %>%
-  do(head(., 1))
+  count(hrr, region) %>%
+  filter(n==max(n)) %>%
+  ungroup() %>%
+  select(hrr, region)
 
 hrr_results = ineq %>%
   filter(unit_type=='hrr') %>%
@@ -122,7 +124,7 @@ hrr_results = ineq %>%
   left_join(hrr_out, by='hrr') %>%
   mutate(y=mean_percent_nonsusceptible/100) %>%
   group_by(bug, drug_group) %>%
-  do(models(., 'y', 'fnz', c('fnz', 'mean'))) %>%
+  do(models(.)) %>%
   ungroup()
 
 state_results = ineq %>%
@@ -132,7 +134,7 @@ state_results = ineq %>%
   right_join(state_abg, by=c('state', 'drug_group')) %>%
[…]

### 2017-08-16 ms2/ms2_display.Rmd
ms2 coefficient correlations

@@
 
+## Comparing the predictive power of mean consumption and fraction nonzero
+
+How do the regressions $R \sim f_>$ compare to $R \sim \mu$?
+Here I look at the correlation between the coefficients $\beta_\mu$ and $\beta_{f_>}$ across bug/drug combinations.
+The correlation is really high for HRRs and just regular high for states.
+
+```{r univariate_result_correlation}
+results %>%
+  filter(str_detect(model, '^uni'), term != '(Intercept)') %>%
+  select(unit_type, bug, drug_group, term, estimate) %>%
+  spread(term, estimate) %>%
+  group_by(unit_type) %>%
+  summarize(cor_between_betas=cor(fnz, mean)) %>%
+  kable
+```
+
+In other words, if you do the regression with $f_>$, you can predict with very high accuracy the coefficient from the univariate regression on $\mu$.
+
 ## Specific plots

### 2017-08-16 ms2/ms2_2analyze.R

@@ -3,7 +3,8 @@ ineq = read_tsv('ineq.tsv') %>%
   mutate(f0=n_0/n_bene, f1=n_1/n_bene, f2=n_2p/n_bene) %>%
   group_by(drug_group, unit_type, unit) %>%
   summarize_at(vars(mean, fnz, f0, f1, f2, mean), mean) %>%
-  ungroup()
+  ungroup() %>%
+  mutate(mup=mean/fnz)
 
 # Resistance data
 resistance_groups = read_tsv('resistance_groups.tsv')
@@ -110,12 +111,6 @@ models = function(df) {
   ) %>%
     mutate(n_data=nrow(df))
 }
-
-hrr_out = hrr %>%
-  count(hrr, region) %>%
-  filter(n==max(n)) %>%
-  ungroup() %>%
-  select(hrr, region)
 
 hrr_results = ineq %>%
   filter(unit_type=='hrr') %>%

### 2017-08-16 ms2/ms2_display.Rmd
new figures; mup

--- 2017-08-16 ms2/ms2_display.Rmd
@@
   summarize_at(vars(mean, fnz), mean) %>%
-  ungroup()
+  ungroup() %>%
+  mutate(mup=mean/fnz)
 
@@
 
-```{r}
-results %>%
-  filter(model=='univariate_mean', unit_type=='hrr', term=='mean') %>%
-  interval_plot() +
-  geom_hline(yintercept=1, col='gray') +
-  ylab(expression(beta[f])) +
-  ggtitle("Linear univariate, HRR")
-```
-
 ```{r state_linear}
@@
 
+Rather than mean consumption $\mu$, I'll use the mean consumption among consumers $\mu_> = \mu / f_>$.
+
+```{r fnz_mup}
+ineq %>%
+  filter(unit_type=='state', drug_group != 'overall') %>%
+  ggplot(aes(x=fnz, y=mup, color=drug_group)) +
+  geom_smooth(method='lm', se=F) +
+  geom_point() +
+  xlab(expression(f['>'])) +
+  ylab(expression(mu['>'])) +
+  ggtitle('Mean nonzero consumption vs. fraction consuming, states') +
+  theme_minimal() +
+  scale_color_manual(values=cbbPalette)
+```
+
+The tetracyclines and Bactrim are the interesting exceptions: in these drugs, states where consumption is more widespread are the states where consumption per consumer is lower.
+
 ## The South has more consumers per capita
@@
 
-We ask if mean consumption $\mu$ has any information beyond fraction nonzero $f_>$ for predicting resistance $R$.
+We ask if mean nonzero consumption $\mu_>$ has any information beyond fraction nonzero $f_>$ for predicting resistance $R$.
 We might expect that the simple two-part model represented by the regressions $R \sim f_>$ doesn't capture the fact that taking two doses should have a greater effect than taking one, etc.
 
-These plots show the coefficient $\beta_\mu$ from the regression $R^{(u)} = \beta_0 + \beta_{f_>} f_>^{(u)} + \beta_{\mu} \mu^{(u)} + \varepsilon^{(u)}$.
+These plots show the coefficient $\beta_{\mu_>}$ from the regression $R^{(u)} = \beta_0 + \beta_{f_>} f_>^{(u)} + \beta_{\mu_>} \mu_>^{(u)} + \varepsilon^{(u)}$.
 
@@
 results %>%
-  filter(model=='multivariate', unit_type=='hrr', term=='mean') %>%
+  filter(model=='multivariate', unit_type=='hrr', term=='mup') %>%
   interval_plot() +
-  ylab(expression(beta[mu])) +
+  ylab(expression(beta[mu['>']])) +
   ggtitle("Linear multivariate, HRR")
@@
 
+At the state level, none of the $\beta_{\mu_>}$ coefficients are significantly different from zero (after multiple-hypothesis correction).
+
 ```{r state_multi}
 results %>%
-  filter(model=='multivariate', unit_type=='state', term=='mean') %>%
+  filter(model=='multivariate', unit_type=='state', term=='mup') %>%
   interval_plot() +
-  ylab(expression(beta[mu])) +
+  ylab(expression(beta[mu['>']])) +
   ggtitle("Linear multivariate, state")
@@
 
-The second two, *E.
-faecium* with $\beta$-lactams and pneumococcus with tetracyclines, are not problematic: in both cases, neither $R \sim f_>$ nor $R \sim \mu$ is significant.
-We're only interested in cases where $f_>$ gives information and $\mu$ gives information beyond that.
-There are cases where $f_>$ doesn't appear to give any information.
+For *E.
+faecium* with $\beta$-lactams, the univariate regression with $f_>$ is not significant, and the one with $\mu_>$ is marginally so.
+Similarly, for *S.
+pneumoniae* and tetracyclines, the regression with $f_>$ is not significant, and the regression wiht $\mu_>$ is.
+Thus, it looks like $\mu_>$ isn't adding information in addition to $f_>$; instead, it just has all the information.
 
@@
 
-First, the regression results are equivocal.
+**swo double-check** First, the regression results are equivocal.
 Regressing $r ~ \mu$, where $r$ are the residuals from the regression $R ~ f_>$, gives a non-significant result.
@@
   filter(bug=='E. coli', drug_group=='tmp_smx') %>%
-  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region)) +
+  ggplot(aes(x=fnz, y=mean_percent_nonsusceptible, color=region)) +
   geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
@@
   ggtitle('E. coli & Bactrim across states')
+```
+
+### E. faecium and beta-lactams
+
+```{r efaecium_betalactam}
+ineq %>%
+  filter(unit_type=='hrr') %>%
+  mutate(hrr=as.integer(unit)) %>% left_join(hrr_out, by='hrr') %>%
+  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
+  filter(bug=='Ec. faecium', drug_group=='beta_lactam') %>%
+  ggplot(aes(x=fnz, y=mean_percent_nonsusceptible, color=region)) +
+  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
+  geom_smooth(method='lm', se=F) +
+  geom_point() +
+  xlab('fraction who consumed beta-lactams') +
+  ylab('Beta-lactam resistance in E. faecium') +
+  ggtitle('E. faecium & beta-lactams across HRRs')
+
+ineq %>%
+  filter(unit_type=='state') %>%
+  rename(state=unit) %>%
+  left_join(regions, by='state') %>%
+  right_join(state_abg, by=c('state', 'drug_group')) %>%
+  filter(bug=='Ec. faecium', drug_group=='beta_lactam') %>%
+  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region)) +
+  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
+  geom_smooth(method='lm', se=F) +
+  geom_point() +
+  xlab('fraction who consumed beta-lactams') +
+  ylab('Beta-lactam resistance E. coli') +
+  ggtitle('E. faecium & beta-lactams across states')
+```
+
+### S. pneumonia and tetracyclines
+
+```{r pneumo_tetra}
+ineq %>%
+  filter(unit_type=='hrr') %>%
+  mutate(hrr=as.integer(unit)) %>% left_join(hrr_out, by='hrr') %>%
+  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
+  filter(bug=='S. pneumoniae', drug_group=='tetracycline') %>%
+  ggplot(aes(x=fnz, y=mean_percent_nonsusceptible, color=region)) +
+  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
+  geom_smooth(method='lm', se=F) +
+  geom_point() +
+  xlab('fraction who consumed tetra') +
+  ylab('tetra resistance pneumo') +
+  ggtitle('pneumo & tetra across HRRs')
+
+ineq %>%
+  filter(unit_type=='state') %>%
+  rename(state=unit) %>%
+  left_join(regions, by='state') %>%
+  right_join(state_abg, by=c('state', 'drug_group')) %>%
+  filter(bug=='S. pneumoniae', drug_group=='tetracycline') %>%
+  ggplot(aes(x=fnz, y=mean_percent_nonsusceptible, color=region)) +
+  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
+  geom_smooth(method='lm', se=F) +
+  geom_point() +
+  xlab('fraction who consumed tetra') +
+  ylab('tetra resistance pneumo') +
+  ggtitle('pneumo & tetra across states')
 ```
@@
   filter(bug=='E. coli', drug_group=='beta_lactam') %>%
+  filter(mean_percent_nonsusceptible < 50) %>% # filter one crazy point
   ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region)) +
@@
 ```
+
+## Comparing the predictive power of mean consumption and fraction nonzero
+
+Not surprisingly, linear regressions $R \sim \mu_>$ give slightly different results from $R \sim f_>$. The negative correlations are all in those drugs where $\mu_>$ was anticorrelated with $f_>$.
+
+```{r hrr_linear_mup}
+results %>%
+  filter(model=='univariate_mup', unit_type=='hrr', term=='mup') %>%
+  interval_plot() +
+  geom_hline(yintercept=1, col='gray') +
+  ylab(expression(beta[mu['+']])) +
+  ggtitle("Linear univariate, HRR")
+```
+
+How do the regressions $R \sim f_>$ compare to $R \sim \mu_>$?
+Here I look at the correlation between the coefficients $\beta_{\mu_>}$ and $\beta_{f_>}$ across bug/drug combinations.
+Not surprisingly, $\beta_\mu$ and $\beta_{f_>}$ are highly correlated: because $\mu$ and $f_>$ are correlated, you can guess the result of the one regression from the other.
+However, $\beta_{\mu_>}$ is not as well correlated with $\beta_{f_>}$, indicating that there might be some extra information there.
+
+```{r univariate_result_correlation}
+results %>%
+  filter(str_detect(model, '^uni'), term != '(Intercept)') %>%
+  select(unit_type, bug, drug_group, term, estimate) %>%
+  spread(term, estimate) %>%
+  group_by(unit_type, drug_group) %>%
+  filter(n() > 2) %>%
+  do(tidy(cor(select(., -unit_type, -bug, -drug_group)))) %>%
+  ungroup() %>%
+  filter(!is.na(fnz)) %>%
+  rename(x=.rownames) %>%
+  gather('y', 'value', fnz:mup) %>%
+  filter(x > y) %>%
+  mutate(key=str_c(x, '_', y)) %>%
+  select(-x, -y) %>%
+  spread(key, value) %>%
+  kable
+```
+
+In other words, if you do the regression with $f_>$, you can predict with very high accuracy the coefficient from the univariate regression on $\mu$.
 

### 2017-08-16 ms2/ms2_2analyze.R

@@ -106,8 +106,9 @@ models = function(df) {
   bind_rows(
     linear_model(df, 'y', 'mean') %>% mutate(model='univariate_mean'),
     linear_model(df, 'y', 'fnz') %>% mutate(model='univariate_fnz'),
+    linear_model(df, 'y', 'mup') %>% mutate(model='univariate_mup'),
     spearman_model(df, 'y', 'fnz') %>% mutate(model='spearman'),
-    linear_model(df, 'y', c('fnz', 'mean')) %>% mutate(model='multivariate')
+    linear_model(df, 'y', c('fnz', 'mup')) %>% mutate(model='multivariate')
   ) %>%
     mutate(n_data=nrow(df))
 }

### 2017-08-17 ms2/ms2_display.Rmd
added NB

@@
   group_by(drug_group, unit_type, unit) %>%
-  summarize_at(vars(mean, fnz), mean) %>%
+  summarize_if(is.numeric, mean) %>%
   ungroup() %>%
@@
 The tetracyclines and Bactrim are the interesting exceptions: in these drugs, states where consumption is more widespread are the states where consumption per consumer is lower.
+(The trend in clindamycin is not significant.) For tetracyclines, there appear to be downward trends within regions (except the Northeast, where the trend is up).
+For Bactrim, there's a Simpson's paradox: the trend is down overall, and down for the South, but it's up for all the other regions.
+
+In both cases, it seems like the South is the mass with the most leverage, and it leverages downward.
+
+For $\beta$-lactams, cephalosporins, clindamycin, macrolides, and quinolones, the slopes seem to mostly agree with one another.
+This suggests that tetracycline and Bactrim prescribing are somehow quite variable between regions.
+
+```{r fnz_mup_tetra_bactrim}
+ineq %>%
+  #filter(unit_type=='hrr', drug_group %in% c('tetracycline', 'tmp_smx')) %>%
+  filter(unit_type=='hrr') %>%
+  filter(mup < 2.2) %>% # get rid of one outlier
+  mutate(hrr=as.integer(unit)) %>% left_join(hrr_out, by='hrr') %>%
+  ggplot(aes(x=fnz, y=mup, color=region)) +
+  facet_wrap(~drug_group, scales='free') +
+  geom_smooth(method='lm', se=F, aes(color=NULL), color='gray') +
+  geom_smooth(method='lm', se=F) +
+  geom_point(shape=1) +
+  xlab(expression(f['>'])) +
+  ylab(expression(mu['>'])) +
+  ggtitle('Mean nonzero consumption vs. fraction consuming, states') +
+  theme_minimal()
+```
+
+Tetracycline and Bactrim are interesting: the South has this long tail where $f_>$ is exceptionally high and $\mu_>$ is low.
+I imagined a model where the South was like the Midwest, but just with more people taking one does of, say, Bactrim.
+So it goes from $d$ doses into $c$ consumers to $d + \Delta c$ doses into $c + \Delta c$ consumers.
+This predicts that $\mu'_> = (\mu_> + \Delta f_>)/(1 + \Delta f_>)$.
+This prediction is too high: it's not *just* that more people in some parts of the South are getting these drugs; it's also that those getting the drug are getting less of it.
 

### 2017-08-17 ms2/ms2_1summarize.R

@@ -30,16 +30,16 @@ inequalities = function(pde, drug_group, unit_type, unit, n_bene) {
   x = pad0(filtered_pde$n_claims, n_bene)
   nzx = nonzero(x)
 
-  nb_par = fitdistrplus::mledist(x, 'nbinom')$estimate
+  nb_par = fitdistrplus::fitdist(x, 'nbinom')$estimate
   nb_size = nb_par['size']
-  nb_prob = nb_size / (nb_size + nb_par['mu'])
+  nb_mu = nb_par['mu']
 
   stopifnot(length(x) == n_bene)
 
-  data_frame(drug_group=drug_group,
-             unit_type=unit_type,
-             unit=unit,
-             n_bene=n_bene,
+  data_frame(drug_group,
+             unit_type,
+             unit,
+             n_bene,
              n_0=sum(x==0),
              n_1=sum(x==1),
              n_2p=sum(x>1),
@@ -47,7 +47,9 @@ inequalities = function(pde, drug_group, unit_type, unit, n_bene) {
              mean=mean(x),
              mean_without_1=mean(without_1(x)),
              mean_without_max=mean(without_max(x)),
-             fnz=fraction_nonzero(x))
+             fnz=fraction_nonzero(x),
+             nb_size,
+             nb_mu)
 }
 
 # auxiliary data

### 2017-08-17 ms2/ms2_2analyze.R

@@ -1,8 +1,7 @@
 # Consumption data
 ineq = read_tsv('ineq.tsv') %>%
-  mutate(f0=n_0/n_bene, f1=n_1/n_bene, f2=n_2p/n_bene) %>%
   group_by(drug_group, unit_type, unit) %>%
-  summarize_at(vars(mean, fnz, f0, f1, f2, mean), mean) %>%
+  summarize_if(is.numeric, mean) %>%
   ungroup() %>%
   mutate(mup=mean/fnz)
 
@@ -107,8 +106,10 @@ models = function(df) {
     linear_model(df, 'y', 'mean') %>% mutate(model='univariate_mean'),
     linear_model(df, 'y', 'fnz') %>% mutate(model='univariate_fnz'),
     linear_model(df, 'y', 'mup') %>% mutate(model='univariate_mup'),
+    linear_model(df, 'y', 'nb_mu') %>% mutate(model='univariate_nb_mu'),
     spearman_model(df, 'y', 'fnz') %>% mutate(model='spearman'),
-    linear_model(df, 'y', c('fnz', 'mup')) %>% mutate(model='multivariate')
+    linear_model(df, 'y', c('fnz', 'mup')) %>% mutate(model='multivariate'),
+    linear_model(df, 'y', c('nb_mu', 'nb_size')) %>% mutate(model='nb')
   ) %>%
     mutate(n_data=nrow(df))
 }

### 2017-08-18 ms1/ms1_display.Rmd
ms1 raw trends

@@
 ```{r adjusted_changes}
-x = model_abx %>%
+adjusted_changes = model_abx %>%
   filter(term=='year0') %>%
@@
 
-x %>%
+unadjusted_changes = bind_rows(claims_by_abx, claims_overall) %>%
+  filter(antibiotic %in% c('overall', top_abx), year %in% c(2011, 2014)) %>%
+  select(antibiotic, year, cpkp) %>%
+  spread(year, cpkp) %>%
+  mutate(change=(`2014`-`2011`)/3,
+         fold_change=change/`2011`,
+         pct_change=100*fold_change) %>%
+  select(antibiotic, pct_change)
+
+adjusted_changes %>%
   ggplot(aes(x=antibiotic, y=pct_change, ymin=ci_low, ymax=ci_high)) +
@@
   geom_errorbar(width=0.5) +
+  geom_point(dat=unadjusted_changes, aes(x=antibiotic, y=pct_change, ymin=NULL, ymax=NULL), color='red') +
   coord_flip() +

### 2017-08-18 ms2/ms2_display.Rmd
ms2 more models

@@
 
+I summarized the resistance data for each geographical unit with a weighted mean, where the weight was the square root of the number of isolates.
+(In the rare cases where the number of isolates was not reported, I put in the median number of isolates for that organism.)
+
 ## Bug/drug combinations
@@
 
-- a Spearman correlation of mean claims per beneficiary against mean percent nonsusceptible
-- a linear model predicting mean resistance from mean consumption
-- a linear model predicting mean resistance from fraction nonzero and nonzero Gini
+- a Spearman correlation of mean claims per beneficiary $\mu$ against mean percent nonsusceptible $R$
+- a linear model predicting mean resistance from the fraction $f_>$ of people who are consumers
+- a linear model predicting mean resistance from $f_>$ and mean consumption among consumers $\mu_> = \mu/f_>$.
+
+When running regressions, I weight the data points by the number of antibiograms present in that unit.
+(The weights in a linear regression should be proportional to $1/\sigma^2$, and we generally expect that the standard error of estimators scales like $1/\sqrt{n}$.)
 
@@
 
-The plot shows the Spearman correlation coefficient $\rho$ between resistance $R$ and fraction $f_>$ of beneficiaries who are consumers.
+The plot shows the Spearman correlation coefficient $\rho$ between resistance $R$ and mean consumption $\mu$.
 Two stars after the bug/drug combination implies that the adjusted $p < 0.05$, and one star means the unadjusted $p < 0.05$.
@@
   ylab("Spearman's rho") +
-  ggtitle("Spearman, HRR")
+  ggtitle("Spearman, res. vs. mean consumption, HRR")
 ```
@@
   ylab("Spearman's rho") +
-  ggtitle("Spearman, state")
+  ggtitle("Spearman, res. vs. mean consumption, state")
 ```
@@
 ## Linear models
+
+We're interested in how it is that mean consumption drives resistance. Is it at the intensive or extensive margin? We separate that by writing $\mu = f_> \mu_>$.
 
@@
   geom_hline(yintercept=1, col='gray') +
-  ylab(expression(beta[f])) +
+  ylab(expression(beta[f['>']])) +
   ggtitle("Linear univariate, HRR")
@@
   interval_plot() +
-  ylab(expression(beta[f])) +
+  ylab(expression(beta[f['>']])) +
   ggtitle("Linear univariate, state")
 ```
+
+Why should the linear models do so much better than the Spearman?
+I think it's because, in the linear model, I can weight the geographical units that have more antibiograms.
+The Spearman model treats them all equally.
 
@@
 
-```{r fnz_mup_tetra_bactrim}
+```{r fnz_mup_by_drug}
 ineq %>%
-  #filter(unit_type=='hrr', drug_group %in% c('tetracycline', 'tmp_smx')) %>%
   filter(unit_type=='hrr') %>%
@@
   xlab('Fraction nonzero') + ylab('claims per beneficiary') +
-  ggtitle('Fraction consuming and mean consumption, states') +
+  ggtitle('All antibiotics, f> and mean consumption, states') +
   theme_minimal()
@@
 ```
+
+Only the West has a significantly non-zero slope.
+
+### E. coli & Bactrim, Part II
+
+Weirdly, within each region, Bactrim resistance goes does as $\mu_>$ increases.
+
+```{r}
+hrr_ir = ineq %>%
+  filter(unit_type=='hrr') %>%
+  mutate(hrr=as.integer(unit)) %>% left_join(hrr_out, by='hrr') %>%
+  right_join(hrr_abg, by=c('hrr', 'drug_group'))
+
+hrr_ir %>%
+  filter(bug=='E. coli', drug_group=='tmp_smx') %>%
+  filter(mean_percent_nonsusceptible < 50) %>% # filter one crazy Nevada point
+  ggplot(aes(x=mup, y=mean_percent_nonsusceptible, color=region)) +
+  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
+  geom_smooth(method='lm', se=F) +
+  geom_point() +
+  xlab('mean Bactrim cons. among Bactrim-eaters') +
+  ylab('Bactrim-resistance E. coli') +
+  ggtitle('E. coli & Bactrim, mu>, HRRs')
+```
+
+Here, the South and West have significantly non-zero slopes.
+
+At least this explains, in a sense, why this relationship looks negative: the South is perched up high, so overall the black line goes down.
 

### 2017-08-18 ms2/ms2_2analyze.R

@@ -40,7 +40,9 @@ good_combos = raw_abg %>%
   #select(bug, drug_group)
 
 abg = raw_abg %>%
+  # filter for only the "good" combinations
   semi_join(good_combos, by=c('bug', 'drug_group')) %>%
+  # compute the median number of isolates for each bug/drug combo
   (function(df) {
     filter(df, !is.na(n_isolates)) %>%
       group_by(bug, drug_group) %>%
@@ -103,12 +105,13 @@ spearman_model = function(df, y_name, x_name) {
 
 models = function(df) {
   bind_rows(
+    spearman_model(df, 'y', 'mean') %>% mutate(model='spearman'),
     linear_model(df, 'y', 'mean') %>% mutate(model='univariate_mean'),
     linear_model(df, 'y', 'fnz') %>% mutate(model='univariate_fnz'),
     linear_model(df, 'y', 'mup') %>% mutate(model='univariate_mup'),
     linear_model(df, 'y', 'nb_mu') %>% mutate(model='univariate_nb_mu'),
-    spearman_model(df, 'y', 'fnz') %>% mutate(model='spearman'),
     linear_model(df, 'y', c('fnz', 'mup')) %>% mutate(model='multivariate'),
+    linear_model(df, 'y', c('mup', 'fnz')) %>% mutate(model='multivariate'),
     linear_model(df, 'y', c('nb_mu', 'nb_size')) %>% mutate(model='nb')
   ) %>%
     mutate(n_data=nrow(df))

### 2017-08-18 ms1/ms1_display.Rmd
overall by populations

@@
 ```{r claims_by_age}
-read_tsv('tables/claims_by_age.tsv') %>%
-  mutate(age_group=case_when(between(.$age, 65, 75) ~ '65-75',
-                             between(.$age, 76, 85) ~ '76-85',
-                             between(.$age, 86, 95) ~ '86-95')) %>%
-  group_by(year, age_group) %>%
+change_demog = read_tsv('tables/model_overall_demography.tsv') %>%
+  mutate(estimate=3*estimate,
+         std.error=sqrt(3)*std.error) %>%
+  model_ci() %>%
+  filter(term=='year0') %>%
+  select(name, adj_change=change_str)
+
+claims_by = function(x) {
+  read_tsv(str_interp("tables/claims_by_${x}.tsv")) %>%
+    rename_(.dots=list(name=x)) %>%
+    select(name, year, cpkp) %>%
+    filter(year %in% c(2011, 2014)) %>%
+    spread(year, cpkp) %>%
+    mutate(pct_change=round((`2014`-`2011`)/`2011`*100, 2)) %>%
+    mutate_at(vars(`2011`, `2014`), round) %>%
+    mutate(group=x)
+}
+
+# do claims by age group separate
+# this could be baked into the analysis step
+claims_by_age = read_tsv('tables/claims_by_age.tsv') %>%
+  filter(year %in% c(2011, 2014)) %>%
+  mutate(name=case_when(between(.$age, 65, 75) ~ 'age65_75',
+                        between(.$age, 76, 85) ~ 'age76_85',
+                        between(.$age, 86, 95) ~ 'age86_95')) %>%
+  group_by(year, name) %>%
   summarize_at(vars(n_bene, n_claims), sum) %>%
   ungroup() %>%
-  mutate(cpkp=round(1000*n_claims/n_bene)) %>%
-  select(year, age_group, cpkp) %>%
+  mutate(cpkp=1000*n_claims/n_bene) %>%
+  select(year, name, cpkp) %>%
   spread(year, cpkp) %>%
+  mutate(pct_change=round((`2014`-`2011`)/`2011`*100, 2)) %>%
+  mutate_at(vars(`2011`, `2014`), round) %>%
+  mutate(group='age')
+
+# use a common interface for all the others
+bind_rows(claims_by_age,
+          claims_by('sex'),
+          claims_by('race'),
+          claims_by('region')) %>%
+  left_join(change_demog, by='name') %>%
+  select(group, name, `2011`, `2014`, pct_change, adj_change) %>%
   kable
@@
 
-## Consumption by sex
-
-```{r claims_by_sex}
-read_tsv('tables/claims_by_sex.tsv') %>%
-  select(sex, year, cpkp) %>%
-  spread(year, cpkp) %>%
-  arrange(desc(`2011`)) %>% 
-  mutate_at(vars(-sex), round) %>%
-  kable
-```
-
-## Consumption by region
-
-```{r claims_by_region}
-read_tsv('tables/claims_by_region.tsv') %>%
-  select(region, year, cpkp) %>%
-  spread(year, cpkp) %>%
-  arrange(desc(`2011`)) %>% 
-  mutate_at(vars(-region), round) %>%
-  kable
-```
-
-## Consumption by race
-
-```{r claims_by_race}
-read_tsv('tables/claims_by_race.tsv') %>%
-  select(race, year, cpkp) %>%
-  spread(year, cpkp) %>%
-  arrange(desc(`2011`)) %>% 
-  mutate_at(vars(-race), round) %>%
-  kable
-```
-
-# Adjusted consumption changes
+# Adjusted consumption changes by drug
 
@@
 The covariates in the regression were age, number of chronic conditions, dual eligibility, sex, race (white or not), and region.
+
+## By drug
 

### 2017-08-18 ms1/ms1_2process.R

@@ -100,6 +100,34 @@ overall_model = bene %>%
   rename(y=n_claims) %>%
   model_f(abx_frmla, 'overall')
 
+# model overall consumption, but look at individual populations
+covariates = c('age0', 'is_female', 'race', 'is_dual', 'region')
+reduced_model = function(df, missing_term, name) {
+  remaining_covariates = covariates[-which(covariates == missing_term)] %>% str_c(collapse=' + ')
+  frmla = str_interp("y ~ year0 + n_cc + ${remaining_covariates}") %>% as.formula
+  
+  df %>%
+    rename(y=n_claims) %>%
+    model_f(frmla, name)
+}
+
+bind_rows(
+  bene %>% filter(between(age, 65, 75)) %>% reduced_model('age0', 'age65_75'),
+  bene %>% filter(between(age, 76, 85)) %>% reduced_model('age0', 'age76_85'),
+  bene %>% filter(between(age, 86, 95)) %>% reduced_model('age0', 'age86_95'),
+  bene %>% filter(is_female) %>% reduced_model('is_female', 'female'),
+  bene %>% filter(!is_female) %>% reduced_model('is_female', 'male'),
+  bene %>% filter(race=='white') %>% reduced_model('race', 'white'),
+  bene %>% filter(race=='black') %>% reduced_model('race', 'black'),
+  bene %>% filter(race=='Hispanic') %>% reduced_model('race', 'Hispanic'),
+  bene %>% filter(race=='other') %>% reduced_model('race', 'other'),
+  bene %>% filter(region=='South') %>% reduced_model('region', 'South'),
+  bene %>% filter(region=='Midwest') %>% reduced_model('region', 'Midwest'),
+  bene %>% filter(region=='West') %>% reduced_model('region', 'West'),
+  bene %>% filter(region=='Northeast') %>% reduced_model('region', 'Northeast')
+) %>%
+  output_table('model_overall_demography')
+
 # get models for individual abx
 single_f = function(abx, frmla) {
   pde %>%
@@ -175,12 +203,22 @@ dxrx_binary = dxrx %>%
   right_join(bene, by=c('year', 'bene_id')) %>%
   replace_na(dx_replace)
 
+# risk ratio
+rr = function(or, p) or / (1 - p + (p * or))
+
 # adjusted OR for diagnoses
 # i.e., logistic regression bene_had_dx_in_year? ~ year + covariates
[…]

### 2017-08-22 ms1/ms1_display.Rmd
ms2; new display

@@
 ```{r adjusted_changes_table}
-x %>% arrange(desc(antibiotic)) %>% kable
+adjusted_changes %>% arrange(desc(antibiotic)) %>% kable
 
-adjusted_az_decrease = x %>%
+adjusted_az_decrease = adjusted_changes %>%
   filter(antibiotic=='azithromycin') %$%
@@
 
-adjusted_levo_increase = x %>%
+adjusted_levo_increase = adjusted_changes %>%
   filter(antibiotic=='levofloxacin') %$%
@@
 
-## Heart disease
-
-We suspected that changes in azithromycin and levofloxacin consumption could be due to altered prescribing guidelines, including one that recommends that azithromycin not be used to treat people at risk for heart fibrillations.
-This includes all elderly, but we epxect the risk would be especially high for those people with heart disease (i.e., AMI | ATRIALFB | CHF | ISCHMCHT).
-I ran separate models for azithromycin and levofloxacin where I include an interaction between a heart disease flag and the year.
-
-```{r heart_model}
-bind_rows(
-  read_tsv('tables/model_azithro.tsv') %>% mutate(model='azithro'),
-  read_tsv('tables/model_levo.tsv') %>% mutate(model='levo')
-) %>%
-  model_ci() %>%
-  select(term, model, pct_change) %>%
-  spread(model, pct_change) %>%
-  kable
-```
-
-For azithromycin, there is a moderate interaction: azithromycin consumption decrease 6.6% per year amongst those without heart disease and 10.2% per year ($1.06588 \times 1.03344$) among those with heart disease.
-For levofloxacin, there is a small interaction that is the opposite of what we might expect: levo consumption increased 10.1% per year among those without heart disease and at a lower rate, 9.4% ($1.10149*(1-0.006782)$), among those with heart disease.
-
-The weakness to this analysis is that I use the CCW flag for heart disease during each year, which means that someone might be getting a lot of azithromycin before they have a heart attack, and afterward they get none.
-I'm also using the one-year chronic conditions, so if someone has a heart attack last year, I don't know about that for the next eyar.
-Nevertheless, as a rough slice, it seems like heart disease doesn't explain what's going on.
-
 # Prescribing practice
@@
 
-The unit of analysis are encounters recorded in the Carrier Line File whose HCPCS code indicated an office visit.
-We only examined encounters whose primary diagnosis fell into set of groups: skin and soft tissue infections (SSTIs), UTIs, probably viral URIs, pneumonia, bronchitis, pharyngitis, sinusitis.
+I examined encounters recorded in the Carrier Line File whose HCPCS code indicated an office visit. I grouped diagnoses into categories using the Fleming-Dutra eTable 2.
 
-## Trends in diagnoses
+To link diagnoses and PDEs, I looked for all diagnosis-PDE pairs where the PDE came 0-3 days after the diagnosis.
 
-Trends per 1,000 beneficiaries in each year. The $p$-value indicates a linear trend.
-
-```{r dx_counts}
-dx_counts = read_tsv('tables/dx_counts.tsv') %>%
-  left_join(totals, by='year') %>%
-  mutate(dx_pkb=n/n_bene*1000) %>%
-  select(diagnosis_type, year, dx_pkb)
-
-dx_counts_trends = dx_counts %>%
-  group_by(diagnosis_type) %>%
-  arrange(year) %>%
-  summarize(p=trend_p(dx_pkb))
-
-dx_counts %>%
-  spread(year, dx_pkb) %>%
-  mutate(pct_change_per_year=(`2014`-`2011`)/`2011`*100/3) %>%
-  left_join(dx_counts_trends, by='diagnosis_type') %>%
-  arrange(desc(`2011`)) %>%
-  mutate_at(vars(`2011`:p), function(x) round(x, 2)) %>%
-  kable
-```
-
-I also performed a similar analysis using logistic regression.
-I predicted whether a beneficiary would have a diagnosis at any point in the year.
-The main predictor variable was year, and the covariates were as elsewhere.
-
-```{r adjusted_dx_or}
-read_tsv('tables/model_dx.tsv') %>%
-  filter(term=='year0') %>%
-  model_ci() %>%
-  select(diagnosis_type, change_str) %>%
-  kable
-```
-
-## Delays between diagnosis and prescription
+### Delays between diagnosis and prescription
 
@@
   kable
-
-f_dx_not_followed = dx_delay %>%
-  group_by(a=!is.na(delay)) %>%
-  summarize(n=sum(n)) %>%
-  spread(a, n) %>%
-  rename(yes=`TRUE`, no=`FALSE`) %$%
-  { yes / (yes + no) }
 ```
 
-However, only `r f_dx_not_followed*100`% of these diagnoses were followed by an antibiotic prescription fill.
+## Trends in diagnoses
 
-## Unadjusted changes in prescribing practice
+Are some diagnoses becoming more or less common?
 
-For each diagnosis type, I computed what fraction of diagnoses were followed by treatment with each antibiotic.
-In the table, I only include dx/rx combinations where the antibiotic was filled after at least 5% of diagnoses.
+### Unadjusted trends in diagnoses
 
-```{r dxrx_fraction}
-dxrx_frac = read_tsv('tables/dxrx_fractions.tsv') %>%
-  group_by(year, diagnosis_type) %>%
-  mutate(total=sum(n)) %>%
-  ungroup() %>%
-  mutate(f=round(100*n/total, 1)) %>%
-  filter(antibiotic %in% top_abx) %>%
-  select(diagnosis_type, antibiotic, year, f) %>%
-  spread(year, f) %>%
-  filter(`2011` >= 3) %>%
-  arrange(diagnosis_type, desc(`2011`))
+First, unadjusted trends in the number of diagnoses per 1,000 beneficiaries in each year:
 
-dxrx_frac %>%
+```{r dx_counts}
+read_tsv('tables/dx_counts.tsv') %>%
+  select(diagnosis_category, year, dxpkp) %>%
+  filter(year %in% c(2011, 2014)) %>%
+  spread(year, dxpkp) %>%
+  mutate(pct_change=(`2014`/`2011`-1)*100) %>%
+  arrange(desc(`2011`)) %>%
+  mutate_if(is.numeric, function(x) round(x, 2)) %>%
   kable
@@
 
-## Adjusted changes in prescribing practice
+### Adjusted trends in diagnoses
 
-To correct for the changing population, I ran a logistic regression for each diagnosis/antibiotic combination.
-The units of analysis were encounters with a primary diagnosis in that diagnosis group, and the outcome was whether that antibiotic was filled after the encounter.
-The main predictor variable was the year, and adjusted covariates included the same as before (age, # chronic conditions, sex, white race, region, dual eligibility).
+Second, adjusted trends, using Poisson regressions.
+The unit of analysis is the beneficiary in a year.
+The outcome variable is number of diagnoses in that category.
+The main predictor is year; the other covariates are as above.
+The percent change is between 2011 and 2014.
 
-```{r dxrx_models}
-read_tsv('tables/model_dxrx.tsv') %>%
-  model_ci() %>%
+```{r dx_trends_adjusted}
+read_tsv('tables/dx_models.tsv') %>%
   filter(term=='year0') %>%
-  # swo: pick the right models to show
-  right_join(dxrx_frac, by=c('diagnosis_type', 'antibiotic')) %>%
-  filter(!is.na(pct_change)) %>%
-  select(antibiotic, diagnosis_type, pct_change, ci_low, ci_high, p.value) %>%
-  arrange(antibiotic, desc(pct_change)) %>%
+  mutate(pct_change=100*(exp(estimate*3)-1),
+         p.value=p.adjust(p.value, method='BH')) %>%
+  mutate_if(is.numeric, function(x) round(x, 2)) %>%
+  select(diagnosis_category, pct_change, p.value) %>%
   kable
 ```
+
+## Common diagnoses by drug
+
+For each drug, I asked what fraction of PDEs are associated with each diagnosis.
+This gives you a sense that, say, the diagnoses that give rise to azithromycin are bronchitis, URI, etc.
+The values are percents.
+I might display just the 2011 data in the paper.
+
+```{r pde_dx}
+read_tsv('tables/pde_dx_counts.tsv') %>%
+  filter(antibiotic %in% top_abx, year %in% c(2011, 2014)) %>%
+  rename(pct=f_pde_with_dx) %>%
+  select(antibiotic, year, diagnosis_category, pct) %>%
+  spread(year, pct) %>%
+  mutate(pct_change=`2014`/`2011`-1) %>%
+  gather('key', 'value', c('2011', '2014', 'pct_change')) %>%
+  mutate(value=round(value*100, 1)) %>%
+  spread(diagnosis_category, value) %>%
+  kable
+```
+
+I also did logistic regressions for these trends, but I don't think it's very important.
+We're not particulary interested in whether the diagnoses that contribute to a drug are changing.
+We're more interested in whether the drugs that *follow* diagnoses are changing, which is next.
+
+## Trends in prescribing practice
+
+### Unadjusted trends
+
+For each diagnosis and each drug, I ask what fraction of diagnoses are followed by a PDE for that drug.
+
+```{r dx_pde}
+read_tsv('tables/dx_pde_counts.tsv') %>%
+  filter(antibiotic %in% top_abx, year %in% c(2011, 2014)) %>%
+  rename(pct=f_dx_with_pde) %>%
+  select(antibiotic, year, diagnosis_category, pct) %>%
+  spread(year, pct) %>%
+  mutate(pct_change=`2014`/`2011`-1) %>%
+  gather('key', 'value', c('2011', '2014', 'pct_change')) %>%
+  mutate(value=round(value*100, 1)) %>%
+  spread(diagnosis_category, value) %>%
+  kable
+```
+
+### Adjusted trends
+
+These are logistic regressions.
+The unit of analysis is the diagnosis.
+The outcome is whether a PDE for that drug follows a diagnosis in that category.
+The main predictor is year; covariates are as above.
+The odds ratio is between 2011 and 2014.
+I'll convert that into a risk ratio using the 2011 probabilities from the previous table.
+
+```{r dx_pde_adjusted}
+read_tsv('tables/dx_pde_model.tsv') %>%
+  filter(term=='year0') %>%
+  mutate(p.value=p.adjust(p.value, method='BH'),
+         OR=round(exp(3*estimate), 2)) %>%
+  filter(p.value < 0.05) %>%
+  select(antibiotic, diagnosis_category, OR) %>%
+  spread(diagnosis_category, OR) %>%
+  head %>%
+  kable
+```

### 2017-08-22 ms1/ms1_2process.R

@@ -238,7 +238,7 @@ dx_pde_adjust_f = function(dx, abx) {
     mutate(y=antibiotic==abx) %>%
     glm(abx_frmla, data=., family='binomial') %>%
     tidy %>%
-    mutate(antibiotic=abx, diagnosis_type=dx)
+    mutate(antibiotic=abx, diagnosis_category=dx)
 }
 
 top_dx = dx_counts %>%

### 2017-08-23 ms1/ms1_display.Rmd
ms1 cleanup

@@
 
-First, unadjusted trends in the number of diagnoses per 1,000 beneficiaries in each year:
-
-```{r dx_counts}
-read_tsv('tables/dx_counts.tsv') %>%
-  select(diagnosis_category, year, dxpkp) %>%
-  filter(year %in% c(2011, 2014)) %>%
-  spread(year, dxpkp) %>%
-  mutate(pct_change=(`2014`/`2011`-1)*100) %>%
-  arrange(desc(`2011`)) %>%
-  mutate_if(is.numeric, function(x) round(x, 2)) %>%
-  kable
-```
-
-### Adjusted trends in diagnoses
+First, unadjusted trends in the number of diagnoses per 1,000 beneficiaries in each year. The percent change is between 2011 and 2014.
 
@@
 
-```{r dx_trends_adjusted}
-read_tsv('tables/dx_models.tsv') %>%
+```{r dx_trends}
+dx_unadj = read_tsv('tables/dx_counts.tsv') %>%
+  select(diagnosis_category, year, dxpkp) %>%
+  filter(year %in% c(2011, 2014)) %>%
+  spread(year, dxpkp) %>%
+  mutate(unadj_pct_change=(`2014`/`2011`-1)*100) %>%
+  arrange(desc(`2011`)) %>%
+  mutate_if(is.numeric, function(x) round(x, 2))
+
+dx_adj = read_tsv('tables/dx_models.tsv') %>%
   filter(term=='year0') %>%
-  mutate(pct_change=100*(exp(estimate*3)-1),
+  mutate(adj_pct_change=100*(exp(estimate*3)-1),
          p.value=p.adjust(p.value, method='BH')) %>%
   mutate_if(is.numeric, function(x) round(x, 2)) %>%
-  select(diagnosis_category, pct_change, p.value) %>%
+  select(diagnosis_category, adj_pct_change, p.value)
+
+full_join(dx_unadj, dx_adj, by='diagnosis_category') %>%
   kable
@@
 This gives you a sense that, say, the diagnoses that give rise to azithromycin are bronchitis, URI, etc.
-The values are percents.
-I might display just the 2011 data in the paper.
+The values are percents in 2011.
+
+The first table includes a "no diagnosis" category.
+This means that the majority of PDEs were not downstream of a diagnosis.
+In the second table, I remove the "none" diagnosis and rescale the values, which helps make it easier to read.
 
@@
 read_tsv('tables/pde_dx_counts.tsv') %>%
-  filter(antibiotic %in% top_abx, year %in% c(2011, 2014)) %>%
-  rename(pct=f_pde_with_dx) %>%
-  select(antibiotic, year, diagnosis_category, pct) %>%
-  spread(year, pct) %>%
-  mutate(pct_change=`2014`/`2011`-1) %>%
-  gather('key', 'value', c('2011', '2014', 'pct_change')) %>%
-  mutate(value=round(value*100, 1)) %>%
-  spread(diagnosis_category, value) %>%
+  filter(antibiotic %in% top_abx, year==2011) %>%
+  mutate(pct=round(f_pde_with_dx*100, 1)) %>%
+  select(antibiotic, diagnosis_category, pct) %>%
+  spread(diagnosis_category, pct) %>%
+  kable
+
+read_tsv('tables/pde_dx_counts.tsv') %>%
+  filter(antibiotic %in% top_abx, year==2011, diagnosis_category != 'none') %>%
+  group_by(antibiotic) %>%
+  mutate(f=f_pde_with_dx/sum(f_pde_with_dx),
+         pct=round(f*100, 1)) %>%
+  ungroup() %>%
+  select(antibiotic, diagnosis_category, pct) %>%
+  spread(diagnosis_category, pct) %>%
   kable
@@
 
-For each diagnosis and each drug, I ask what fraction of diagnoses are followed by a PDE for that drug.
+For each diagnosis and each drug, I ask what fraction of diagnoses are followed by a PDE for that drug in 2011 and 2014. I compute the percent change in those fractions.
 
-```{r dx_pde}
-read_tsv('tables/dx_pde_counts.tsv') %>%
-  filter(antibiotic %in% top_abx, year %in% c(2011, 2014)) %>%
-  rename(pct=f_dx_with_pde) %>%
-  select(antibiotic, year, diagnosis_category, pct) %>%
-  spread(year, pct) %>%
-  mutate(pct_change=`2014`/`2011`-1) %>%
-  gather('key', 'value', c('2011', '2014', 'pct_change')) %>%
-  mutate(value=round(value*100, 1)) %>%
-  spread(diagnosis_category, value) %>%
-  kable
-```
-
-### Adjusted trends
-
-These are logistic regressions.
+Second, I do logistic regressions.
 The unit of analysis is the diagnosis.
@@
 The odds ratio is between 2011 and 2014.
-I'll convert that into a risk ratio using the 2011 probabilities from the previous table.
+(I'll convert that into a risk ratio using the 2011 probabilities.)
 
-```{r dx_pde_adjusted}
-read_tsv('tables/dx_pde_model.tsv') %>%
+```{r dx_pde_trends}
+dx_pde_unadj = read_tsv('tables/dx_pde_counts.tsv') %>%
+  filter(antibiotic %in% top_abx, year %in% c(2011, 2014)) %>%
+  mutate(key=str_c('y', year)) %>%
+  select(diagnosis_category, antibiotic, key, value=f_dx_with_pde)
+
+dx_pde_adj = read_tsv('tables/dx_pde_model.tsv') %>%
   filter(term=='year0') %>%
-  mutate(p.value=p.adjust(p.value, method='BH'),
+  mutate(key='OR',
+         p.value=p.adjust(p.value, method='BH'),
          OR=round(exp(3*estimate), 2)) %>%
   filter(p.value < 0.05) %>%
-  select(antibiotic, diagnosis_category, OR) %>%
-  spread(diagnosis_category, OR) %>%
-  head %>%
+  select(diagnosis_category, antibiotic, key, value=OR)
+
+# risk ratio
+rr = function(odds_ratio, p0) odds_ratio / (1 - p0 + (p0 * odds_ratio))
+
+ bind_rows(dx_pde_unadj, dx_pde_adj) %>%
+   spread(key, value) %>%
+   mutate(pct_change=(y2014/y2011-1)*100,
+          adj_pct_change=(rr(OR, y2011)-1)*100) %>%
+   select(-OR) %>%
+   #mutate_if(is.numeric, function(x) round(x, 2)) %>%
+   gather('key', 'value', y2011:adj_pct_change) %>%
+   mutate(value=round(value, 2)) %>%
+   spread(diagnosis_category, value) %>%
+   mutate(key=factor(key, levels=c('y2011', 'y2014', 'pct_change', 'adj_pct_change'))) %>%
+   arrange(antibiotic, key) %>%
    kable

### 2017-08-23 ms1/ms1_display.Rmd
RR, broke up table

@@
 
-### Unadjusted trends
-
 For each diagnosis and each drug, I ask what fraction of diagnoses are followed by a PDE for that drug in 2011 and 2014. I compute the percent change in those fractions.
@@
 The main predictor is year; covariates are as above.
-The odds ratio is between 2011 and 2014.
-(I'll convert that into a risk ratio using the 2011 probabilities.)
+The adjusted percent change is between 2011 and 2014.
+(I computed this adjusted change, which is a relative risk, using the odds ratio from the regression and the mean probability of PDE for that diagnosis in 2011.
+This strictly incorrect: the mean relative risk is the mean of the risks computed using the predicted probabilities for each 2011 diagnosis, *not* the relative risk computed using the mean probability in 2011.
+But it seems like it's decently close, and I think insisting on the thing that is technically correct will only create confusion.)
+
+I broke this into three tables by antibiotic, just to make it easier to navigate.
 
@@
 
- bind_rows(dx_pde_unadj, dx_pde_adj) %>%
+dx_pde_both= bind_rows(dx_pde_unadj, dx_pde_adj) %>%
    spread(key, value) %>%
@@
    mutate(key=factor(key, levels=c('y2011', 'y2014', 'pct_change', 'adj_pct_change'))) %>%
-   arrange(antibiotic, key) %>%
+   arrange(antibiotic, key)
+
+dx_pde_both %>%
+  filter(str_detect(antibiotic, '^a')) %>%
+   kable
+
+dx_pde_both %>%
+  filter(str_detect(antibiotic, '^c')) %>%
+   kable
+
+dx_pde_both %>%
+  filter(str_detect(antibiotic, '^[^ac]')) %>%
    kable

### 2017-08-23 ms2/ms2_display.Rmd
ms2; cleaner models; collinearity

@@
   count(hrr, region) %>%
+  group_by(hrr) %>%
   filter(n==max(n)) %>%
@@
 The gray line shows $\beta_{f_>} = 1$.
+Most confidence intervals include that line, i.e., we can explain our data with a transmission-model where individuals who take drug have 100% resistance and others have some smaller amount.
 
@@
 I think it's because, in the linear model, I can weight the geographical units that have more antibiograms.
+That is, we give more weight to the greater amounts of data.
 The Spearman model treats them all equally.
@@
 ## Unevenness is mostly related to the "extensive margin"
+
+The line has slope 1.5.
 
@@
   ggplot(aes(x=fnz, y=mean, color=drug_group)) +
+  geom_abline(slope=1.5, color='gray') +
   geom_point() +
@@
 
-Rather than mean consumption $\mu$, I'll use the mean consumption among consumers $\mu_> = \mu / f_>$.
+Because $f_>$ is so collinear with $\mu$, multiple regressions using both as predictors will cause problems (cf.
+the Supplement).
+Rather than mean consumption $\mu$, I'll use the mean consumption among consumers $\mu_> = \mu / f_>$ as my second predictor.
 
@@
 
-We ask if mean nonzero consumption $\mu_>$ has any information beyond fraction nonzero $f_>$ for predicting resistance $R$.
+We ask if mean consumption $\mu$ has any information beyond fraction nonzero $f_>$ for predicting resistance $R$.
 We might expect that the simple two-part model represented by the regressions $R \sim f_>$ doesn't capture the fact that taking two doses should have a greater effect than taking one, etc.
 
-These plots show the coefficient $\beta_{\mu_>}$ from the regression $R^{(u)} = \beta_0 + \beta_{f_>} f_>^{(u)} + \beta_{\mu_>} \mu_>^{(u)} + \varepsilon^{(u)}$.
+These plots show the coefficient $\beta_\mu$ from the regression $R^{(u)} = \beta_0 + \beta_{f_>} f_>^{(u)} + \beta_\mu \mu^{(u)} + \varepsilon^{(u)}$.
 
@@
 results %>%
-  filter(model=='multivariate', unit_type=='hrr', term=='mup') %>%
+  filter(model=='multivariate_fnz_mup', unit_type=='hrr', term=='mup') %>%
   interval_plot() +
   ylab(expression(beta[mu['>']])) +
-  ggtitle("Linear multivariate, HRR")
+  ggtitle("Linear multivariate (f> and mu>), HRR")
 ```
 
-At the state level, none of the $\beta_{\mu_>}$ coefficients are significantly different from zero (after multiple-hypothesis correction).
+At the state level, none of the $\beta_\mu$ coefficients are significantly different from zero (after multiple-hypothesis correction).
 
@@
 results %>%
-  filter(model=='multivariate', unit_type=='state', term=='mup') %>%
+  filter(model=='multivariate_fnz_mean', unit_type=='state', term=='mean') %>%
   interval_plot() +
-  ylab(expression(beta[mu['>']])) +
-  ggtitle("Linear multivariate, state")
+  ylab(expression(beta[mu])) +
+  ggtitle("Linear multivariate (f> and mean), state")
 ```
@@
 - *E. coli* and Bactrim have a *negative* relationship (higher consumption goes with *less* resistance)
-- *E. faecium* and $\beta$-lactams have a postive relationship (implying higher mean consumption, given the same fraction of people consuming, goes with more resistance)
 - *S. pneumoniae* and tetracyclines have a negative relationship
+- *E. faecium* and $\beta$-lactams have a positive relationship (implying higher mean consumption, given the same fraction of people consuming, goes with more resistance)
+
+Interestingly, these are the same three values that are significant when performing the (ill-advised) regression using $f_>$ and $\mu$:
+
+```{r fnz_mu_results}
+results %>%
+  filter(unit_type=='hrr', model=='multivariate_fnz_mup', term=='mup') %>%
+  filter(p.adjust(p.value, method='BH') < 0.05) %>%
+  select(bug, drug_group, estimate, p.value) %>%
+  arrange(desc(estimate)) %>%
+  kable
+```
 
 For *E.
-faecium* with $\beta$-lactams, the univariate regression with $f_>$ is not significant, and the one with $\mu_>$ is marginally so.
-Similarly, for *S.
-pneumoniae* and tetracyclines, the regression with $f_>$ is not significant, and the regression wiht $\mu_>$ is.
-Thus, it looks like $\mu_>$ isn't adding information in addition to $f_>$; instead, it just has all the information.
+faecium* with $\beta$-lactams, the univariate regression with $f_>$ gives a slightly negative slope but is not significant.
+The univariate regression with $\mu$ gives a slightly positive slope and is not significant.
+In the multiple regression, the signs stay the same but the coefficients are very different.
+Thus, we can't trust that $\mu$ provides data in addition to $f_>$ when it's not clear that either provides the information on its own.
 
-And some marginal results:
+For *S.
+pneumoniae* and tetracyclines, the regression with $f_>$ and the regression with $\mu$ give negative slopes, but neither are significant.
+In the multiple regression, the $f_>$ coefficient becomes positive.
+Again, I'm not sure I can trust this.
 
-- *E. coli* and $\beta$-lactams, negative relationship, HRR only; significant in univariate regressions $R \sim f_>$
-- *E. cloacae* and quinolones, positive, state only; marginally significant in univariate
-- Non-MRSA *S. aureus* and first-generation cephalosporins, negative, HRR only; but not significant in univariate
-- *E. faecium* and nitro, negative, both HRR and state; not significant in univariate
-- *S. aureus* and clinda, positive, state only; not significant in univariate
-- *S. aureus* and tetracyclines, positive, state only; not significant in univariate
+For *E.
+coli* and Bactrim, the univariate regression with $f_>$ is positive and signficant.
+The regression with $\mu$ is positive and not significant.
+Then, in the multiple regression, $\mu$ flips signs and becomes highly significant (different from zero), and it's significant in an ANOVA where $f_>$ comes first, but it's not significant in an ANOVA where it ($\mu$) comes first.
 
@@
 This negative relationship is the most tricky.
-
-**swo double-check** First, the regression results are equivocal.
-Regressing $r ~ \mu$, where $r$ are the residuals from the regression $R ~ f_>$, gives a non-significant result.
-An ANOVA analysis shows that sequential application of $f_>$ and $\mu$ as predictors of $R$ both explain non-trivial variance; but if the order is reversed, and $\mu$ is used first, then it *doesn't* explain any variance.
-This makes me think that there is just some weird interaction between the two terms in the multiple regression.
 
@@
 ```{r}
-hrr_ir = ineq %>%
+ineq %>%
   filter(unit_type=='hrr') %>%
   mutate(hrr=as.integer(unit)) %>% left_join(hrr_out, by='hrr') %>%
-  right_join(hrr_abg, by=c('hrr', 'drug_group'))
-
-hrr_ir %>%
+  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
   filter(bug=='E. coli', drug_group=='tmp_smx') %>%
   filter(mean_percent_nonsusceptible < 50) %>% # filter one crazy Nevada point
-  ggplot(aes(x=mup, y=mean_percent_nonsusceptible, color=region)) +
+  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region)) +
   geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
@@
 
-Not surprisingly, linear regressions $R \sim \mu_>$ give slightly different results from $R \sim f_>$. The negative correlations are all in those drugs where $\mu_>$ was anticorrelated with $f_>$.
-
-```{r hrr_linear_mup}
-results %>%
-  filter(model=='univariate_mup', unit_type=='hrr', term=='mup') %>%
-  interval_plot() +
-  geom_hline(yintercept=1, col='gray') +
-  ylab(expression(beta[mu['+']])) +
-  ggtitle("Linear univariate, HRR")
-```
-
-How do the regressions $R \sim f_>$ compare to $R \sim \mu_>$?
-Here I look at the correlation between the coefficients $\beta_{\mu_>}$ and $\beta_{f_>}$ across bug/drug combinations.
+How do the regressions $R \sim f_>$ compare to $R \sim \mu$?
+Here I look at the correlation between the coefficients $\beta_\mu$ and $\beta_{f_>}$ across bug/drug combinations.
 Not surprisingly, $\beta_\mu$ and $\beta_{f_>}$ are highly correlated: because $\mu$ and $f_>$ are correlated, you can guess the result of the one regression from the other.
 However, $\beta_{\mu_>}$ is not as well correlated with $\beta_{f_>}$, indicating that there might be some extra information there.
+
+The column titles are `metric1_metric2`, and the value shows the correlation between the univariate models' coefficients for predicting resistance to that drug across bugs.
 
@@
   ungroup() %>%
-  filter(!is.na(fnz)) %>%
   rename(x=.rownames) %>%
-  gather('y', 'value', fnz:mup) %>%
+  gather('y', 'value', fnz:nb_mu) %>%
   filter(x > y) %>%
@@
 
+## Collinearity
+
+This shows why it's a bad idea to do a multiple regression with $f_>$ and $\mu$. It confirms that we're mostly getting bad results due to multicollinearity.
+
+The gray lines link the coefficients from a univariate model and a multivariate model (i.e., with $f_>$ and $\mu$) for a certain bug/drug combination.
+
+From a previous plot, we know that $\mu \approx \alpha f_>$, where $\alpha \approx 3/2$. Then if $R \approx \beta_0 + \beta_{f_>}$, then we should guess that $\beta_\mu = \beta_{f_>} / \alpha$.
+
+Furthermore, in doing a multivariate regression, using the same math, we find that we expect
+$$
+\beta_\mu^{(2)} = \frac{1}{\alpha} \left( \beta_{f_>}^{(1)} - \beta_{f_>}^{(2)} \right),
+$$
+where the superscripts indicate that the coefficients are from the univariate ("1") and multivariate ("2") regressions.
+Thus, for $\beta_{f_>}^{(1)} \approx 0$, we expect the multivariate regression coefficients be related by a slope of $-1/\alpha$.
+
+On this plot, the pink lines show $\pm 1/\alpha$, which lines up very nicely with the data (especially considering that I picked $\alpha=3/2$ just by eyeballing that other plot).
+
+```{r uni_multi}
+results %>%
+  filter(unit_type=='hrr', term != '(Intercept)',
+         model %in% c('univariate_fnz', 'univariate_mean', 'multivariate_fnz_mean')) %>%
+  mutate(model=str_extract(model, '(uni|multi)'),
+         bug_drug=interaction(bug, drug_group)) %>%
+  select(bug_drug, model, term, estimate) %>%
+  spread(term, estimate) %>%
+  ggplot(aes(x=fnz, y=mean, color=model, group=bug_drug)) +
+  geom_abline(slope=2/3, color='pink') +
+  geom_abline(slope=-2/3, intercept=0, color='pink') +
+  geom_line(color='gray') +
+  geom_point() +
+  xlab(expression(beta[f['>']])) +
+  ylab(expression(beta[mu])) +
+  ggtitle('Relationship between model coefficients (f>, mean), HRR')
+```
+
+But the result is not so pronounced for $f_>$ and $\mu_>$:
+
+```{r uni_multi_mup}
+results %>%
+  filter(unit_type=='hrr', term != '(Intercept)',
+         model %in% c('univariate_fnz', 'univariate_mup', 'multivariate_fnz_mup')) %>%
+  mutate(model=str_extract(model, '(uni|multi)'),
+        bug_drug=str_c(bug, drug_group)) %>%
+  select(bug_drug, model, term, estimate) %>%
+  spread(term, estimate) %>%
+  ggplot(aes(x=fnz, y=mup, color=model, group=bug_drug, label=bug_drug)) +
+  #geom_text(size=2) +
+  geom_line(color='gray') +
+  geom_point() +
+  xlab(expression(beta[f['>']])) +
+  ylab(expression(beta[mu['>']])) +
+  ggtitle('Relationship between model coefficients (f>, mu>) , HRR')
+```
+
 ## Correlations between the metrics
 
-The mean and the fraction nonzero are highly correlated ($\rho \sim 1$) for all drug groups:
+The mean and the fraction nonzero are highly correlated ($\rho \sim 1$) for all drug groups. (Note the range on the $x$-axis!)
 

### 2017-08-23 ms2/ms2_2analyze.R

@@ -110,8 +110,10 @@ models = function(df) {
     linear_model(df, 'y', 'fnz') %>% mutate(model='univariate_fnz'),
     linear_model(df, 'y', 'mup') %>% mutate(model='univariate_mup'),
     linear_model(df, 'y', 'nb_mu') %>% mutate(model='univariate_nb_mu'),
-    linear_model(df, 'y', c('fnz', 'mup')) %>% mutate(model='multivariate'),
-    linear_model(df, 'y', c('mup', 'fnz')) %>% mutate(model='multivariate'),
+    linear_model(df, 'y', c('fnz', 'mup')) %>% mutate(model='multivariate_fnz_mup'),
+    linear_model(df, 'y', c('mup', 'fnz')) %>% mutate(model='multivariate_mup_fnz'),
+    linear_model(df, 'y', c('fnz', 'mean')) %>% mutate(model='multivariate_fnz_mean'),
+    linear_model(df, 'y', c('mean', 'fnz')) %>% mutate(model='multivariate_mean_fnz'),
     linear_model(df, 'y', c('nb_mu', 'nb_size')) %>% mutate(model='nb')
   ) %>%
     mutate(n_data=nrow(df))
@@ -121,7 +123,7 @@ hrr_results = ineq %>%
   filter(unit_type=='hrr') %>%
   mutate(hrr=as.integer(unit)) %>%
   right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
-  left_join(hrr_out, by='hrr') %>%
+  #left_join(hrr_out, by='hrr') %>%
   mutate(y=mean_percent_nonsusceptible/100) %>%
   group_by(bug, drug_group) %>%
   do(models(.)) %>%

### 2017-08-23 ms1/ms1_display.Rmd
notes

@@
 ```
+
+# Summary of results for each drug
+
+For azithromycin, the most common diagnoses are bronchitis, URI, and sinusitis.
+Bronchitis diagnoses are on a steep decline (15% over the study period), URI
+diagnoses rose slightly (~5%), and sinusitis diagnoses remained nearly constant
+(+0.13%). Azithromycin prescribing per diagnosis also fell during the study
+period for each of these diagnoses (10-15%). So the diagnoses for which
+azithromycin is used were overall less common, and azithromycin was also less
+used for all those diagnoses.
+
+For levofloxacin, the most common diagnoses are bronchitis, "other
+respiratory", and URI. Bronchitis fell steeply, other respiratory fell somewhat
+(5%) and URIs rose slightly (2%). However, levofloxacin prescribing rose
+sharply for bronchitis (75%), other respiratory (60%), and URI (75%). So even
+though bronchitis diagnosing fell, levo prescribing for its diagnoses rose so
+much.
+
+The most common diagnoses for clindamycin, which rose, were SSTIs (tier 2 and
+tier 3). Tier 3 diagnoses rose somewhat (2%) and Tier 2 diagnoses fell somewhat
+(4%). Clindamycin prescribing for Tier 2 and Tier 3 diagnoses rose (27%, 22%).
+So clinda simply became a more popular treatment for SSTIs.
+
+The most common diagnoses for amox/clav, which rose, were sinusitis, SSTIs,
+bronchitis, and URIs. Amox/clav prescribing rose sharply for sinusitis (42%),
+rose for Tier 3 SSTIs (10%) but not Tier 2 SSTIs (n.s.), and rose for
+bronchitis (29%) and URIs (36%). So amox/clav became more popular for most of
+its diagnoses, like levo.
+
+The most common diagnoses for amox, which rose, were similar to those for
+amox/clav: sinusitis, SSTIs, bronchitis, and URIs. Amox prescribing rose
+slightly for sinusitis (3%), rose for Tier 2 and Tier 3 SSTIs (7%, 10%), and
+rose for bronchitis (12%) and URI (9%). The story is similar to amox/clav, just
+not as pronounced.
+
+The most common diagnoses for cephalexin, which rose slightly, were SSTIs, and
+somewhat UTIs. Prescribing for SSTIs rose (6%) and rose more for UTIs (32% for
+Tier 1 UTIs, 20% for Tier 3). So there was some growth in SSTIs, which make up
+more of cephalexin's portfolio, and big growth in UTIs, which make up a small
+part.
+
+The most common diagnoses for Bactrim, which fell slightly, are UTIs and
+SSTIs.  UTIs are nearly constant and SSTIs increased a little. However, Bactrim
+prescribing *increased* for SSTIs and UTIs. This means that the decrease
+occurred among the PDEs that were not associated with diagnoses. My first guess
+is that there is less long-term treatment (since, in long-term treatment, most
+of the PDEs will nt be associated with a visit).
+
+The most common diagnoses for doxy, which fell, are SSTIs and bronchitis.
+Doxy prescribing increased for Tier 2 SSTIs (5%), remained constant for Tier 3
+SSTIs (n.s.), and decreased slightly for bronchitis (2%). So it seems there was
+not big change in prescribing practice, and maybe this is more a reflection of
+what gets diagnosed as bronchitis.
+
+The most common diagnoses for cipro, which fell, are UTIs, followed by GI
+infections. GI infections fell (5%). Cipro prescribing for UTIs and GI
+infections remained constant (n.s. for all diagnoses). My first guess is that
+cipro fell because GI infection diagnoses fell and because of a decrease in
+long-term treatment for UTIs, like in Bactrim.
+
+The most common diagnoses for nitrofurantoin, which fell, were UTIs. UTIs
+remained nearly constant, and nitro prescribing for UTIs only fell slightly (3%
+for Tier 1 UTIs, n.s. for Tier 3). Again, my first guess is that there was a
+decrease in the number of PDEs not associated with diagnoses, i.e., those in
+long-term treatment.
+
+## Questions for the paper
+
+### Breaking up the tables
+
+I don't want to show this enormous table in the main text. I might break this into three tables:
+
+- Skin and soft tissue infections (Tiers 2 and 3): clindamycin, cephalexin, Bactrim, doxy, amox, amox/clav
+- Respiratory conditions (bronchitis, URI, sinusitis, "other respiratory"): azithro, levo, amox, amox/clav, doxy
+- UTIs (Tier 1 infections and Tier 3 conditions): cipro, Bactrim, nitro, cephalexin
+- GI (optional): but this is just cipro
+
+Alternatively, I could keep one table that has these diagnoses as columns and all these drugs going down as rows.
+
+### Lumping together diagnoses.
+
+I could lump the multiple tiers for diagnoses together. This
+makes the table easier to interpret, but it doesn't make the point that
+diagnosing patterns for multiple tiers for the same diagnosis genus have
+similar dynamics, i.e., appropriate and inappropriate prescribing are changing
+the same ways.

### 2017-08-23 ms2/ms2_display.Rmd
notes

@@
 The gray line shows $\beta_{f_>} = 1$.
-Most confidence intervals include that line, i.e., we can explain our data with a transmission-model where individuals who take drug have 100% resistance and others have some smaller amount.
+Most confidence intervals include that line, i.e., we can explain our data with a transmission-free model where individuals who take drug have 100% resistance and others have some smaller amount.
+In only one case do we have a place where the data are inconsistent with this simple model.
+(Note that the $\beta_0$ can mean the resistance among non-consumers or the fraction of non-consumers who are cryptic consumers.)
 
@@
 
-Second, from the Spearman and univariate models, Bactrim resistance in *E.
+From the Spearman and univariate models, Bactrim resistance in *E.
 coli* is correlated with increased resistance (black line).
@@
 
-```{r ecoli_bactrim}
+It's the L-shape!
+
+```{r ecoli_bactrim_fnz_hrr}
 ineq %>%
@@
   ggplot(aes(x=fnz, y=mean_percent_nonsusceptible, color=region)) +
+  geom_abline(intercept=16.86, slope=114.8, color='pink') +
   geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
@@
   ggtitle('E. coli & Bactrim across HRRs')
+```
 
+These internal trends are a lot weaker when looking at the state level:
+
+```{r ecoli_bactrim_fnz_state}
 ineq %>%
@@
 
-Only the West has a significantly non-zero slope.
+But the same patterns are visible when looking at $\mu_>$ across HRRs:
 
-### E. coli & Bactrim, Part II
-
-Weirdly, within each region, Bactrim resistance goes does as $\mu_>$ increases.
-
-```{r}
+```{r ecoli_bactrim_mup}
 ineq %>%
@@
 ```
-
-Here, the South and West have significantly non-zero slopes.
-
-At least this explains, in a sense, why this relationship looks negative: the South is perched up high, so overall the black line goes down.
 

### 2017-08-23 ms2/ms2_display.Rmd
ms2 model notes

@@
 
-Interestingly, these are the same three values that are significant when performing the (ill-advised) regression using $f_>$ and $\mu$:
+Interestingly, these are the same three values that are significant when performing the (ill-advised) regression using $f_>$ and $\mu$.
+The table shows that bug/drug combinations in which the multiple regressions have significantly non-zero $\beta_{\mu_>}$.
+The values in the columns show the estimates of $\beta$ in each model.
+Note the big differences in $\beta_{f_>}$ between the two models.
+Note that although the regression $R \sim f_> + \mu$ distorts $\beta_{f_>}$ more than the regression $R \sim f_> + \mu_>$, the coefficient still gets changed quite a bit nonetheless.
+We were hoping for $\mu_>$ to give information in addition to $f_>$, not to just change it completely.
 
@@
 results %>%
-  filter(unit_type=='hrr', model=='multivariate_fnz_mup', term=='mup') %>%
-  filter(p.adjust(p.value, method='BH') < 0.05) %>%
-  select(bug, drug_group, estimate, p.value) %>%
-  arrange(desc(estimate)) %>%
+  filter(unit_type=='hrr', model %in% c('univariate_fnz', 'multivariate_fnz_mean', 'multivariate_fnz_mup'), term != '(Intercept)') %>%
+  (function(df) {
+    filter(df, term != 'fnz', p.adjust(p.value, method='BH') < 0.05) %>%
+      select(bug, drug_group) %>%
+      distinct() %>%
+      inner_join(df, by=c('bug', 'drug_group')) }) %>%
+  mutate(model=str_replace(model, '^multivariate_', '')) %>%
+  select(bug, drug_group, model, term, estimate) %>%
+  spread(term, estimate) %>%
   kable

### 2017-08-24 ms2/ms2_display.Rmd
ms2 new plot

@@
       inner_join(df, by=c('bug', 'drug_group')) }) %>%
+  mutate(model=str_replace(model, '^multivariate_', '')) %>%
+  select(bug, drug_group, model, term, estimate) %>%
+  spread(term, estimate) %>%
+  kable
+```
+
+I contrast this situation against *E. coli* and quinolones. In that case, adding terms to the regression doesn't drastically alter $\beta_{f_>}$.
+
+```{r fnz_mu_ecoliquin}
+results %>%
+  filter(unit_type=='hrr', model %in% c('univariate_fnz', 'multivariate_fnz_mean', 'multivariate_fnz_mup'), term != '(Intercept)') %>%
+  filter(bug=='E. coli', drug_group=='quinolone') %>%
   mutate(model=str_replace(model, '^multivariate_', '')) %>%

### 2017-08-28 ms2/ms2_display.Rmd
cleanup

@@
   ggplot(aes(x=fnz, y=mean, color=drug_group)) +
-  geom_abline(slope=1.5, color='gray') +
+  #geom_abline(slope=1.5, color='gray') +
+  stat_smooth(method='lm', se=F, fullrange=TRUE) +
   geom_point() +
   xlab('Fraction nonzero') + ylab('claims per beneficiary') +
+  geom_hline(yintercept=0) + geom_vline(xintercept=0) +
   ggtitle('Fraction consuming and mean consumption, states') +
@@
 
-```{r univariate_result_correlation}
+```{r univariate_result_correlation, eval=FALSE}
 results %>%

### 2017-08-28 ms1/ms1_1clean.R

@@ -13,14 +13,11 @@ regions = read_tsv('../../db/census-regions/census-regions.tsv') %>%
 # window of days that you can look ahead from an rx to a dx
 rxdx_window = 3
 
-# diagnosis types
-fd = read_tsv('../../db/fd_icd/fd_categories.tsv') %>%
-  mutate(dc_id=1:n())
-
 load_data = function(year) {
   # pde data
   pde = read_tsv(sprintf('../pde_%i.tsv', year)) %>%
-    select(bene_id, service_date, antibiotic) %>%
+    select(bene_id, pde_date, antibiotic) %>%
+    distinct() %>%
     mutate(year=year, pde_id=1:n())
 
   # total numbers of PDEs
@@ -41,16 +38,14 @@ load_data = function(year) {
     replace_na(list(n_claims=0))
 
   # diagnosis claims
-  dx = read_tsv(str_interp('../../data/dx_${year}.tsv')) %>%
-    select(bene_id=BENE_ID, code=dx_code, diagnosis_category, from_date=dt1) %>%
-    # filter for bene's we have
-    semi_join(bene, by='bene_id') %>%
+  dx = read_tsv(str_interp('../dx_${year}.tsv')) %>%
+    select(bene_id, dx_date, dx_category) %>%
     distinct() %>%
-    mutate(from_date=dmy(from_date), year=year, dx_id=1:n())
+    mutate(year=year, dx_id=1:n())
 
   # find all possible dx-PDE combinations
   dx_pde = inner_join(dx, pde, by=c('year', 'bene_id')) %>%
-    mutate(delay=service_date - from_date) %>%
+    mutate(delay=pde_date - dx_date) %>%
     filter(between(delay, 0, rxdx_window)) %>%
     mutate(year=year)
 
@@ -66,8 +61,7 @@ save_dat = function(dat, name) {
 
 # prepare for parallel execution
 library(parallel)
-n_cores = detectCores() - 1
-cluster = makeCluster(n_cores, type='FORK')
+cluster = makeCluster(3, type='FORK')
 
 dat = parLapply(cluster, 2011:2014, load_data)
 

### 2017-08-28 ms2/ms2_2analyze.R

@@ -3,7 +3,8 @@ ineq = read_tsv('ineq.tsv') %>%
   group_by(drug_group, unit_type, unit) %>%
   summarize_if(is.numeric, mean) %>%
   ungroup() %>%
-  mutate(mup=mean/fnz)
+  mutate(mup=mean/fnz,
+         emup=mup - 1)
 
 # Resistance data
 resistance_groups = read_tsv('resistance_groups.tsv')
@@ -22,6 +23,11 @@ hrr = read_tsv('../../db/hrr/hrr.tsv') %>%
   select(zipcode, hrr) %>%
   left_join(zipcodes, by='zipcode') %>%
   left_join(regions, by='state')
+
+# Associate each HRR with a single region by just taking the first one
+hrr_region = hrr %>%
+  group_by(hrr) %>%
+  summarize(region=first(region))
 
 raw_abg = read_tsv('../../../antibiogram/data/abg.tsv', col_types=cols(zipcode='c')) %>%
   mutate(drug=tolower(drug), percent_nonsusceptible=100-percent_susceptible) %>%
@@ -108,9 +114,14 @@ models = function(df) {
     spearman_model(df, 'y', 'mean') %>% mutate(model='spearman'),
     linear_model(df, 'y', 'mean') %>% mutate(model='univariate_mean'),
     linear_model(df, 'y', 'fnz') %>% mutate(model='univariate_fnz'),
+    linear_model(df, 'y', c('fnz', '0')) %>% mutate(model='univariate_fnz_0'),
     linear_model(df, 'y', 'mup') %>% mutate(model='univariate_mup'),
     linear_model(df, 'y', 'nb_mu') %>% mutate(model='univariate_nb_mu'),
     linear_model(df, 'y', c('fnz', 'mup')) %>% mutate(model='multivariate_fnz_mup'),
+    linear_model(df, 'y', c('fnz', 'emup')) %>% mutate(model='multivariate_fnz_emup'),
+    linear_model(df, 'y', c('fnz', 'mup', '0')) %>% mutate(model='multivariate_fnz_mup_0'),
+    linear_model(df, 'y', c('fnz', 'I(1/fnz)')) %>% mutate(model='multivariate_fnz_1fnz'),
+    linear_model(df, 'y', c('fnz', 'I(fnz^2)')) %>% mutate(model='multivariate_fnz_fnz2'),
[…]

### 2017-08-28 ms2/ms2_display.Rmd
simplified ineq

@@
 # Unevenness of consumption and antibiotic resistance
+
+## Unevenness of consumption
+
+For each drug, what fraction of people took that drug? Of people who took it, how many had one claim? How many claims are accounted for by singleton users? This is data from 2011.
+
+```{r ineq_table}
+pct = function(x) round(x*100, 2)
+
+ineq %>%
+  # use only 2011 data
+  filter(year==2011, unit_type=='state') %>%
+  # sum across unit types
+  group_by(drug_group, n_claims) %>%
+  summarize(n_bene=sum(n_bene)) %>%
+  ungroup() %>%
+  mutate(n_class_claims = n_claims * n_bene) %>%
+  group_by(drug_group) %>%
+  mutate(pct_bene=pct(n_bene/sum(n_bene)),
+         pct_claims=pct(n_class_claims/sum(n_class_claims))) %>%
+  select(drug_group, n_claims, pct_bene, pct_claims) %>%
+  mutate(n_claims=fct_other(factor(n_claims), keep=0:2, other_level='3+')) %>%
+  group_by(drug_group, n_claims) %>%
+  summarize_at(vars(pct_bene, pct_claims), sum) %>%
+  ungroup() %>%
+  gather('key', 'val', pct_bene:pct_claims) %>%
+  mutate(key2=str_c(n_claims, ':', key)) %>%
+  select(drug_group, key2, val) %>%
+  spread(key2, val) %>%
+  kable
+```
 

### 2017-08-28 ms2/ms2_1summarize.R

@@ -1,68 +1,12 @@
 #!/usr/bin/env Rscript
 
-# utility functions function
-gini = function(x) ineq::ineq(x, type='Gini')
-nonzero = function(x) x[x > 0]
-fraction_nonzero = function(x) length(nonzero(x)) / length(x)
-pad0 = function(x, to) c(x, rep(0, to - length(x)))
-
-without_min = function(x) x[-which.min(x)]
-without_max = function(x) x[-which.max(x)]
-without_1 = function(x) x[-match(1, x)]
-
-filter_eq_ = function(df, col, val) filter_(df, str_interp("${col}=='${val}'"))
-
-inequalities = function(pde, drug_group, unit_type, unit, n_bene) {
-  if (drug_group=='overall') {
-    filter_drug_f = function(df) group_by(df, bene_id) %>% summarize(n_claims=sum(n_claims))
-  } else {
-    filter_drug_f = function(df) filter_eq_(df, 'drug_group', drug_group)
-  }
-
-  filtered_pde = pde %>%
-    filter_eq_(unit_type, unit) %>%
-    filter_drug_f()
-
-  if(n_bene < nrow(filtered_pde)) {
-    stop(str_interp("n_bene=${n_bene}, but got ${nrow(filtered_pde)} rows when doing ${drug_group}, ${unit_type}, ${unit}"))
-  }
-
-  x = pad0(filtered_pde$n_claims, n_bene)
-  nzx = nonzero(x)
-
-  nb_par = fitdistrplus::fitdist(x, 'nbinom')$estimate
-  nb_size = nb_par['size']
-  nb_mu = nb_par['mu']
-
-  stopifnot(length(x) == n_bene)
-
-  data_frame(drug_group,
-             unit_type,
-             unit,
-             n_bene,
-             n_0=sum(x==0),
-             n_1=sum(x==1),
-             n_2p=sum(x>1),
-             total=sum(x),
-             mean=mean(x),
-             mean_without_1=mean(without_1(x)),
-             mean_without_max=mean(without_max(x)),
-             fnz=fraction_nonzero(x),
-             nb_size,
-             nb_mu)
-}
-
-# auxiliary data
-consumption_groups = read_tsv('consumption_groups.tsv')
-
-regions = read_tsv('../../db/census-regions/census-regions.tsv') %>%
-  select(state, state_abbreviation, region) %>%
-  filter(state != 'District of Columbia')
-stopifnot('DC' %in% regions$state)
+library(rlang)
 
 # read in HRR designations for all years
[…]

### 2017-08-30 ms2/ms2_display.Rmd
ms2: new ineq

--- 2017-08-28 ms2/ms2_display.Rmd
@@
 ```{r load_data, cache.extra=c(md5sum('ineq.tsv'), md5sum('results.tsv'))}
-ineq = read_tsv('ineq.tsv') %>%
-  group_by(drug_group, unit_type, unit) %>%
-  summarize_if(is.numeric, mean) %>%
-  ungroup() %>%
-  mutate(mup=mean/fnz)
+ineq = read_tsv('ineq.tsv')
+
+ineq_summary = ineq %>%
+  group_by(year, unit_type, unit, drug_group) %>%
+  summarize(mean=weighted.mean(n_claims, w=n_bene),
+            fnz=sum(n_bene[n_claims > 0]) / sum(n_bene),
+            mup=mean/fnz) %>%
+  group_by(unit_type, unit, drug_group) %>%
+  summarize_at(vars(mean, fnz,mup), mean) %>%
+  ungroup()
 
@@
 
-hrr_out = hrr %>%
+hrr_regions = hrr %>%
   count(hrr, region) %>%
@@
   interval_plot() +
+  geom_hline(yintercept=1, color='gray') +
   ylab(expression(beta[f['>']])) +
@@
 
+## Unevenness of consumption
+
+For each drug, what fraction of people took that drug? Of people who took it, how many had one claim? How many claims are accounted for by singleton users? This is data from 2011.
+
+```{r ineq_table}
+pct = function(x) round(x*100, 2)
+
+ineq %>%
+  filter(unit_type=='state') %>%
+  # sum across units
+  group_by(year, drug_group, n_claims) %>%
+  summarize(n_bene=sum(n_bene)) %>%
+  ungroup() %>%
+  # compute the number of claims made by people making 1 claim, etc.
+  mutate(n_class_claims = n_claims * n_bene) %>%
+  # compute percentages: fr. of benes w/ 1 claim, fr. of claims made by 1-claimers, etc.
+  group_by(year, drug_group) %>%
+  mutate(pct_bene=pct(n_bene/sum(n_bene)),
+         pct_claims=pct(n_class_claims/sum(n_class_claims))) %>%
+  select(year, drug_group, n_claims, pct_bene, pct_claims) %>%
+  # group 3+ claims into one group, then sum up fractions in that new group
+  mutate(n_claims=fct_other(factor(n_claims), keep=0:2, other_level='3+')) %>%
+  group_by(year, drug_group, n_claims) %>%
+  summarize_at(vars(pct_bene, pct_claims), sum) %>%
+  ungroup() %>%
+  # merge the claims group and metric into one key, and spread
+  gather('key', 'val', pct_bene:pct_claims) %>%
+  mutate(key2=str_c(n_claims, ':', key)) %>%
+  select(year, drug_group, key2, val) %>%
+  spread(key2, val) %>%
+  # keep only 2011 data
+  filter(year==2011) %>%
+  select(-year) %>%
+  kable
+```
+
 ## Unevenness is mostly related to the "extensive margin"
@@
 
-ineq %>%
+ineq_summary %>%
   filter(unit_type=='state', drug_group != 'overall') %>%
   ggplot(aes(x=fnz, y=mean, color=drug_group)) +
-  #geom_abline(slope=1.5, color='gray') +
-  stat_smooth(method='lm', se=F, fullrange=TRUE) +
-  geom_point() +
+  stat_smooth(aes(color=NA), color='gray', method='lm', se=F, fullrange=TRUE) +
+  geom_point(shape=1) +
   xlab('Fraction nonzero') + ylab('claims per beneficiary') +
@@
 ```{r fnz_mup}
-ineq %>%
+ineq_summary %>%
   filter(unit_type=='state', drug_group != 'overall') %>%
@@
   geom_smooth(method='lm', se=F) +
-  geom_point() +
+  geom_point(shape=1) +
   xlab(expression(f['>'])) +
@@
 ```{r fnz_mup_by_drug}
-ineq %>%
+ineq_summary %>%
   filter(unit_type=='hrr') %>%
-  filter(mup < 2.2) %>% # get rid of one outlier
-  mutate(hrr=as.integer(unit)) %>% left_join(hrr_out, by='hrr') %>%
+  #rename(state=unit) %>% left_join(regions, by='state') %>%
+  mutate(hrr=as.integer(unit)) %>% left_join(hrr_regions) %>%
   ggplot(aes(x=fnz, y=mup, color=region)) +
@@
 ```{r fnz_mean_region}
-ineq %>%
-  filter(unit_type=='state', drug_group == 'overall') %>%
+ineq_summary %>%
+  filter(unit_type=='state', drug_group=='overall') %>%
   rename(state=unit) %>%
-  filter(state != 'DC') %>%
+  filter(state != 'District of Columbia') %>%
   left_join(regions, by='state') %>%

### 2017-08-30 ms2/ms2_1summarize.R

@@ -28,21 +28,28 @@ summarize_inequality = function(y) {
     rename(drug=antibiotic) %>%
     inner_join(consumption_groups, by='drug') %>%
     count(bene_id, drug_group) %>% rename(n_claims=n) %>%
+    # add the "overall" category
+    (function(df) {
+      group_by(df, bene_id) %>%
+        summarize(n_claims=sum(n_claims)) %>%
+        mutate(drug_group='overall') %>%
+        bind_rows(df)
+    }) %>%
     left_join(bene, by='bene_id')
-  
+
   # function to compute the denominators, i.e., the number of benes in each unit
   denom_f = function(bene, unit_type_) {
     bene %>%
       rename(unit=!!sym(unit_type_)) %>%
       count(unit) %>% rename(n_bene=n)
   }
-  
+
   # function to compute inequality, i.e., the number of beneficiaries taking
   # each number of drug
   ineq_f = function(pde, bene, unit_type_) {
     # get denominator, which we'll use to fill in the non-consumers
     denom = denom_f(bene, unit_type_)
-    
+
     pde %>%
       # rename, e.g., the "state" column to "unit"
       rename(unit=!!sym(unit_type_)) %>%
@@ -66,14 +73,14 @@ summarize_inequality = function(y) {
       arrange(unit, drug_group) %>%
       mutate(unit_type=unit_type_)
   }
-  
+
   # get inequalities for each drug/unit at the state and HRR levels
   ineq = bind_rows(
     ineq_f(pde, bene, 'state'),
     ineq_f(pde, bene, 'hrr')
   ) %>%
     mutate(year=y)
-  
+
   ineq
 }
 

### 2017-08-30 ms2/ms2_2analyze.R

@@ -1,10 +1,14 @@
+#!/usr/bin/env Rscript
+
 # Consumption data
 ineq = read_tsv('ineq.tsv') %>%
-  group_by(drug_group, unit_type, unit) %>%
-  summarize_if(is.numeric, mean) %>%
-  ungroup() %>%
-  mutate(mup=mean/fnz,
-         emup=mup - 1)
+  group_by(year, unit_type, unit, drug_group) %>%
+  summarize(mean=weighted.mean(n_claims, w=n_bene),
+            fnz=sum(n_bene[n_claims > 0]) / sum(n_bene),
+            mup=mean/fnz) %>%
+  group_by(unit_type, unit, drug_group) %>%
+  summarize_at(vars(mean, fnz,mup), mean) %>%
+  ungroup()
 
 # Resistance data
 resistance_groups = read_tsv('resistance_groups.tsv')
@@ -114,18 +118,13 @@ models = function(df) {
     spearman_model(df, 'y', 'mean') %>% mutate(model='spearman'),
     linear_model(df, 'y', 'mean') %>% mutate(model='univariate_mean'),
     linear_model(df, 'y', 'fnz') %>% mutate(model='univariate_fnz'),
-    linear_model(df, 'y', c('fnz', '0')) %>% mutate(model='univariate_fnz_0'),
     linear_model(df, 'y', 'mup') %>% mutate(model='univariate_mup'),
-    linear_model(df, 'y', 'nb_mu') %>% mutate(model='univariate_nb_mu'),
     linear_model(df, 'y', c('fnz', 'mup')) %>% mutate(model='multivariate_fnz_mup'),
-    linear_model(df, 'y', c('fnz', 'emup')) %>% mutate(model='multivariate_fnz_emup'),
-    linear_model(df, 'y', c('fnz', 'mup', '0')) %>% mutate(model='multivariate_fnz_mup_0'),
     linear_model(df, 'y', c('fnz', 'I(1/fnz)')) %>% mutate(model='multivariate_fnz_1fnz'),
     linear_model(df, 'y', c('fnz', 'I(fnz^2)')) %>% mutate(model='multivariate_fnz_fnz2'),
     linear_model(df, 'y', c('mup', 'fnz')) %>% mutate(model='multivariate_mup_fnz'),
     linear_model(df, 'y', c('fnz', 'mean')) %>% mutate(model='multivariate_fnz_mean'),
[…]

### 2017-08-30 ms2/ms2_display.Rmd
ms2 analyze, display

@@
 ```{r load_data, cache.extra=c(md5sum('ineq.tsv'), md5sum('results.tsv'))}
-ineq = read_tsv('ineq.tsv')
+ineq_table = read_tsv('ineq.tsv')
 
-ineq_summary = ineq %>%
+ineq = ineq_table %>%
   group_by(year, unit_type, unit, drug_group) %>%
@@
 
-ineq %>%
+ineq_table %>%
   filter(unit_type=='state') %>%
@@
 
-ineq_summary %>%
+ineq %>%
   filter(unit_type=='state', drug_group != 'overall') %>%
@@
 ```{r fnz_mup}
-ineq_summary %>%
+ineq %>%
   filter(unit_type=='state', drug_group != 'overall') %>%
@@
 ```{r fnz_mup_by_drug}
-ineq_summary %>%
+ineq %>%
   filter(unit_type=='hrr') %>%
@@
 ```{r fnz_mean_region}
-ineq_summary %>%
+ineq %>%
   filter(unit_type=='state', drug_group=='overall') %>%
@@
   filter(unit_type=='hrr') %>%
-  mutate(hrr=as.integer(unit)) %>% left_join(hrr_out, by='hrr') %>%
+  mutate(hrr=as.integer(unit)) %>% left_join(hrr_regions, by='hrr') %>%
   right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
@@
 
+## Using logistic models
+
+We don't expect that, over a large range, the consumption/resistance relationship will be linear.
+We expect it would be something more like a logistic function.
+To evaluate the difference between these two approaches, I performed quasibinomial regressions $R \sim f_>$.
+If $\bar{f}$ is the midpoint of the $f_>$ values (i.e., the weighted mean) and $\hat{y}$ is the predicted resistance from the model, then the slope at the midpoint is $\beta_f \hat{y}(\bar{x}) \left[ 1- \hat{y}(\bar{x}) \right]$.
+I compare the $p$-values and slopes in the linear and quasibinomial regressions:
+
+```{r linear_quasibin}
+results %>%
+  filter(unit_type=='hrr', model %in% c('univariate_fnz', 'univariate_fnz_log'), term=='fnz') %>%
+  select(bug, drug_group, model, estimate, p.value) %>%
+  group_by(model) %>%
+  mutate(p.value = p.adjust(p.value, method='BH')) %>%
+  ungroup() %>%
+  ggplot(aes(x=estimate, y=-log10(p.value), color=model, group=interaction(bug, drug_group))) +
+  geom_hline(yintercept = -log10(0.05)) +
+  geom_point() +
+  geom_line(color='gray')
+```
+
+Points above the black line are significant.
+Note that the $p$-values in the two cases are really similar (i.e., most dumbbells are on the same horizontal line) but they can differ somewhat in their slope.
+The right-most pink dot is pneumococcus/macrolides, and the next is *P.
+mirabilis*/Bactrim.
+
+I also tried doing beta regressions, which get similar-sized changes in the slope but really significant $p$-values. Not sure what to make of those.
+
 ## Correlations between the metrics

### 2017-08-30 ms2/ms2_2analyze.R

@@ -113,6 +113,39 @@ spearman_model = function(df, y_name, x_name) {
              p.value=m$p.value)
 }
 
+beta_model = function(df_raw, y, x) {
+  # replace 0 and 1 with slightly different values
+  eps = 1e-6
+  df = df_raw %>%
+    mutate(y=case_when(.$y == 0 ~ eps,
+                       .$y == 1 ~ 1 - eps,
+                       TRUE ~ .$y))
+  
+  frmla = as.formula(str_interp("${y} ~ ${x}"))
+  m = betareg::betareg(formula=frmla, weights=n_place_antibiograms, data=df, link='logit')
+  
+  xbar = weighted.mean(df[[y]], w=df$n_place_antibiograms)
+  newdata = setNames(list(xbar), x) %>% as_data_frame
+
+  yhat = predict(m, newdata=newdata, type='response')
+  slope = m$coefficients$mean[[x]] * yhat * (1 - yhat)
+  pvalue = tidy(m) %>% filter(term==x) %>% pull(p.value)
+  data_frame(term=x, estimate=slope, p.value=pvalue)
+}
+
+logistic_model = function(df, y, xs) {
+  frmla = as.formula(str_interp("${y} ~ ${str_c(xs, collapse='+')}"))
+  m = glm(formula=frmla, weights=n_place_antibiograms, data=df, family=quasibinomial)
+  
+  xbar = weighted.mean(df[[y]], w=df$n_place_antibiograms)
+  newdata = setNames(list(xbar), x) %>% as_data_frame
+  
+  yhat = predict(m, newdata=newdata, type='response')
+  slope = m$coefficients[[x]] * yhat * (1 - yhat)
+  pvalue = tidy(m) %>% filter(term==x) %>% pull(p.value)
+  data_frame(term=x, estimate=slope, p.value=pvalue)
+}
+
 models = function(df) {
   bind_rows(
     spearman_model(df, 'y', 'mean') %>% mutate(model='spearman'),
@@ -124,7 +157,9 @@ models = function(df) {
     linear_model(df, 'y', c('fnz', 'I(fnz^2)')) %>% mutate(model='multivariate_fnz_fnz2'),
     linear_model(df, 'y', c('mup', 'fnz')) %>% mutate(model='multivariate_mup_fnz'),
     linear_model(df, 'y', c('fnz', 'mean')) %>% mutate(model='multivariate_fnz_mean'),
[…]

### 2017-08-30 ms2/ms2_display.Rmd
ms2 display

@@
   kable
+```
+
+```{r ineq_hist}
+cbbPalette <- c("#000000", "#D55E00", "#CC79A7", "#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2")
+
+ineq_table %>%
+  filter(unit_type=='state') %>%
+  # sum across units
+  group_by(year, drug_group, n_claims) %>%
+  summarize(n_bene=sum(n_bene)) %>%
+  ungroup() %>%
+  # compute the number of claims made by people making 1 claim, etc.
+  mutate(n_class_claims = n_claims * n_bene) %>%
+  # compute percentages: fr. of benes w/ 1 claim, fr. of claims made by 1-claimers, etc.
+  group_by(year, drug_group) %>%
+  mutate(pct_bene=pct(n_bene/sum(n_bene))) %>%
+  select(year, drug_group, n_claims, pct_bene) %>%
+  # group 3+ claims into one group, then sum up fractions in that new group
+  mutate(n_claims=fct_other(factor(n_claims), keep=0:2, other_level='3+')) %>%
+  group_by(year, drug_group, n_claims) %>%
+  summarize_at(vars(pct_bene), sum) %>%
+  ungroup() %>%
+  # show only 2011 data
+  filter(year==2011) %>%
+  mutate(drug_group=fct_reorder(drug_group, pct_bene, fun=max)) %>%
+  ggplot(aes(x=n_claims, y=pct_bene, fill=drug_group)) +
+  geom_bar(stat='identity', position='dodge') +
+  scale_fill_manual(values=cbbPalette) +
+  theme_minimal()
 ```
@@
 
+```{r linear_beta}
+results %>%
+  filter(unit_type=='hrr', model %in% c('univariate_fnz', 'univariate_fnz_beta'), term=='fnz') %>%
+  select(bug, drug_group, model, estimate, p.value) %>%
+  group_by(model) %>%
+  mutate(p.value = p.adjust(p.value, method='BH')) %>%
+  ungroup() %>%
+  ggplot(aes(x=estimate, y=-log10(p.value), color=model, group=interaction(bug, drug_group))) +
+  geom_hline(yintercept = -log10(0.05)) +
+  geom_point() +
+  geom_line(color='gray')
+```
+
 ## Correlations between the metrics

### 2017-09-01 ms1/ms1_display.Rmd
adapting

@@
 est2pct = function(x) (exp(x) - 1)*100
+mult_est = function(df, m) {
+  mutate(df, estimate=m*estimate, std.error=sqrt(m)*estimate)
+}
 model_ci = function(df) {
@@
 }
-
-model_abx = read_tsv('tables/model_abx.tsv') %>%
-  rename(antibiotic=name) %>%
-  model_ci()
 
@@
 
-# this should be the same list as for the models
-stopifnot(all(top_abx %in% model_abx$antibiotic))
-
 claims_by_top_abx = claims_by_abx %>%
@@
   model_ci() %>%
-  filter(term=='year0') %>%
+  filter(term=='year') %>%
   select(name, adj_change=change_str)
@@
 ```{r adjusted_changes}
-adjusted_changes = model_abx %>%
-  filter(term=='year0') %>%
+adjusted_changes = read_tsv('tables/model_abx.tsv') %>%
+  rename(antibiotic=name) %>%
+  filter(term=='year') %>%
+  mutate(estimate=3*estimate,
+         std.error=sqrt(3)*std.error) %>%
+  model_ci() %>%
   select(antibiotic, pct_change, ci_low, ci_high) %>%
@@
   spread(year, cpkp) %>%
-  mutate(change=(`2014`-`2011`)/3,
+  mutate(change=`2014`-`2011`,
          fold_change=change/`2011`,
@@
   xlab('') +
-  ylab('adjusted yearly change in consumption (%)') +
-  scale_y_continuous(limits=c(-10, 10), breaks=c(-10, -5, 0, 5, 10)) +
+  ylab('adjusted change in consumption 2011 to 2014 (%)') +
+  scale_y_continuous(limits=c(-22, 32), breaks=c(-20, -10, 0, 10, 20, 30)) +
   theme_minimal() +
@@
 adjusted_az_decrease = adjusted_changes %>%
-  filter(antibiotic=='azithromycin') %$%
-  { (1 + pct_change/100)^3 } %>%
-  { (1 - .) * 100 } %>%
+  filter(antibiotic=='azithromycin') %>%
+  pull(pct_change) %>%
   round(2)
@@
 adjusted_levo_increase = adjusted_changes %>%
-  filter(antibiotic=='levofloxacin') %$%
-  { (1 + pct_change/100)^3 } %>%
-  { (. - 1) * 100 } %>%
+  filter(antibiotic=='levofloxacin') %>%
+  pull(pct_change) %>%
   round(2)
@@
 
-read_tsv('tables/model_fq.tsv') %>%
+fq_trends = read_tsv('tables/model_fq.tsv') %>%
+  mult_est(3) %>%
   model_ci() %>%
-  select(term, pct_change, ci_low, ci_high, p.value) %>%
+  select(term, pct_change, ci_low, ci_high, p.value) %T>%
   kable
+
+fq_change_pct = fq_trends %>%
+  filter(term=='year') %>%
+  pull(pct_change) %>%
+  round(2)
 ```
 
-Overall, fluoroquinolones changed by about 0% per year during this period. This included `r fq_names`.
+Overall, fluoroquinolones changed by a tiny amount, `r fq_change_pct`%, over this period. This included `r fq_names`.
 
@@
 
-I examined encounters recorded in the Carrier Line File whose HCPCS code indicated an office visit. I grouped diagnoses into categories using the Fleming-Dutra eTable 2.
+First, I looked at all encounters recorded in the Carrier Line File and the Outpatient file. A large fraction, I think 78%, of PDEs had an encounter 0-7 days before.
 
-To link diagnoses and PDEs, I looked for all diagnosis-PDE pairs where the PDE came 0-3 days after the diagnosis.
+Next, I limited by analysis to encounters who HCPCS code indicated an E&M visit, and I grouped diagnostic codes into categories using Fleming-Dutra eTable 2.
+I determined the trends in the prevalence of these diagnosis categories by counting the diagnoses per beneficiary per year and by using Poisson regressions that account for the unusual covariates.
 
-### Delays between diagnosis and prescription
+I analyzed trends in prescribing practice by computing the fraction of encounters in each diagnosis category that were followed by PDEs for each drug.
+I computed adjusted trends using logistic regression.
 
-The table shows the number of diagnoses that were followed by that number of days' delay before the following antibiotic prescription. Most prescriptions were filled on the same day as the diagnosis.
-
-```{r dx_delay}
-dx_delay = read_tsv('tables/dx_delay.tsv')
-
-dx_delay %>%
-  filter(!is.na(delay)) %>%
-  mutate(f=n/sum(n)) %>%
-  kable
-```
+Finally, I considered only those PDEs that are downstream of those E&M diagnoses. For each antibiotic, I computed the fraction of PDEs that have each diagnosis category upstream.
 
@@
 Are some diagnoses becoming more or less common?
-
-### Unadjusted trends in diagnoses
 
@@
 ```{r dx_trends}
-dx_unadj = read_tsv('tables/dx_counts.tsv') %>%
-  select(diagnosis_category, year, dxpkp) %>%
+dx_unadj = read_tsv('tables/dx_trends_table.tsv') %>%
   filter(year %in% c(2011, 2014)) %>%
+  left_join(select(totals, year, n_bene), by='year') %>%
+  mutate(dxpkp=1000*n_dx/n_bene) %>%
+  select(dx_cat, year, dxpkp) %>%
   spread(year, dxpkp) %>%
@@
 
-dx_adj = read_tsv('tables/dx_models.tsv') %>%
-  filter(term=='year0') %>%
-  mutate(adj_pct_change=100*(exp(estimate*3)-1),
-         p.value=p.adjust(p.value, method='BH')) %>%
-  mutate_if(is.numeric, function(x) round(x, 2)) %>%
-  select(diagnosis_category, adj_pct_change, p.value)
+dx_adj = read_tsv('tables/dx_trends.tsv') %>%
+  rename(dx_cat=model) %>%
+  filter(term=='year') %>%
+  mult_est(3) %>%
+  model_ci() %>%
+  select(dx_cat, change_str, p.value)
 
-full_join(dx_unadj, dx_adj, by='diagnosis_category') %>%
+full_join(dx_unadj, dx_adj, by='dx_cat') %>%
   kable
@@
 
-The first table includes a "no diagnosis" category.
-This means that the majority of PDEs were not downstream of a diagnosis.
-In the second table, I remove the "none" diagnosis and rescale the values, which helps make it easier to read.
-
 ```{r pde_dx}
-read_tsv('tables/pde_dx_counts.tsv') %>%
-  filter(antibiotic %in% top_abx, year==2011) %>%
-  mutate(pct=round(f_pde_with_dx*100, 1)) %>%
-  select(antibiotic, diagnosis_category, pct) %>%
-  spread(diagnosis_category, pct) %>%
-  kable
-
-read_tsv('tables/pde_dx_counts.tsv') %>%
-  filter(antibiotic %in% top_abx, year==2011, diagnosis_category != 'none') %>%
-  group_by(antibiotic) %>%
-  mutate(f=f_pde_with_dx/sum(f_pde_with_dx),
-         pct=round(f*100, 1)) %>%
-  ungroup() %>%
-  select(antibiotic, diagnosis_category, pct) %>%
-  spread(diagnosis_category, pct) %>%
-  kable
+read_tsv('tables/dx_from_pde.tsv') %>% 
+  spread(present, n) %>%
+  mutate(pct=`TRUE`/(`TRUE` + `FALSE`)*100 %>% round(2)) %>%
+  select(antibiotic, dx_cat, pct) %>%
+  spread(dx_cat, pct) %>%
+  head %>% kable
 ```
-
-I also did logistic regressions for these trends, but I don't think it's very important.
-We're not particulary interested in whether the diagnoses that contribute to a drug are changing.
-We're more interested in whether the drugs that *follow* diagnoses are changing, which is next.
 
@@
 For each diagnosis and each drug, I ask what fraction of diagnoses are followed by a PDE for that drug in 2011 and 2014. I compute the percent change in those fractions.
+
+```{r dx_to_pde_table}
+dx_to_pde_unadj = read_tsv('tables/dx_rx_table.tsv') %>%
+  filter(year %in% c(2011, 2014)) %>%
+  select(dx_cat, antibiotic, year, f) %>%
+  spread(year, f)
+
+dx_to_pde_adj = read_tsv('tables/dx_rx_trends.tsv') %>%
+  filter(term=='year') %>%
+  mult_est(3) %>%
+  mutate(ci=1.96*std.error,
+         ci_low=estimate-ci,
+         ci_high=estimate+ci) %>%
+  select(dx_cat, antibiotic, estimate, ci_low, ci_high)
+
+risk_ratio_f = function(or, p) or / (1 + p * (or - 1))
+
+inner_join(dx_to_pde_unadj, dx_to_pde_adj, by=c('dx_cat', 'antibiotic')) %>%
+  mutate(estimate=risk_ratio_f(estimate, `2011`),
+         ci_low=risk_ratio_f(ci_low, `2011`),
+         ci_high=risk_ratio_f(ci_high, `2011`),
+         pct_change=(`2014`-`2011`)/`2011`) %>%
+  mutate_if(is.numeric, function(x) round(x*100, 2)) %>%
+  mutate(adj_change=sprintf('%.2f%% (%.2f -- %.2f%%)', estimate, ci_low, ci_high)) %>%
+  select(dx_cat, antibiotic, `2011`, `2014`, pct_change, adj_change) %>%
+  kable
+```
 
@@
 
-```{r dx_pde_trends}
-dx_pde_unadj = read_tsv('tables/dx_pde_counts.tsv') %>%
-  filter(antibiotic %in% top_abx, year %in% c(2011, 2014)) %>%
-  mutate(key=str_c('y', year)) %>%
-  select(diagnosis_category, antibiotic, key, value=f_dx_with_pde)
+```{r dx_to_pde_trends}
 
-dx_pde_adj = read_tsv('tables/dx_pde_model.tsv') %>%
-  filter(term=='year0') %>%
-  mutate(key='OR',
-         p.value=p.adjust(p.value, method='BH'),
-         OR=round(exp(3*estimate), 2)) %>%
-  filter(p.value < 0.05) %>%
-  select(diagnosis_category, antibiotic, key, value=OR)
-
-# risk ratio
-rr = function(odds_ratio, p0) odds_ratio / (1 - p0 + (p0 * odds_ratio))
-
-dx_pde_both= bind_rows(dx_pde_unadj, dx_pde_adj) %>%
-   spread(key, value) %>%
-   mutate(pct_change=(y2014/y2011-1)*100,
-          adj_pct_change=(rr(OR, y2011)-1)*100) %>%
-   select(-OR) %>%
-   #mutate_if(is.numeric, function(x) round(x, 2)) %>%
-   gather('key', 'value', y2011:adj_pct_change) %>%
-   mutate(value=round(value, 2)) %>%
-   spread(diagnosis_category, value) %>%
-   mutate(key=factor(key, levels=c('y2011', 'y2014', 'pct_change', 'adj_pct_change'))) %>%
-   arrange(antibiotic, key)
-
-dx_pde_both %>%
-  filter(str_detect(antibiotic, '^a')) %>%
-   kable
-
-dx_pde_both %>%
-  filter(str_detect(antibiotic, '^c')) %>%
-   kable
-
-dx_pde_both %>%
-  filter(str_detect(antibiotic, '^[^ac]')) %>%
-   kable
 ```

### 2017-09-01 ms1/ms1_3dx.R

@@ -12,7 +12,6 @@ bene = read_tsv('data/bene.tsv') %>%
          race=factor(race, levels=c('white', 'black', 'Hispanic', 'other'))) %>%
   mutate(age=age-1)
 
-# trends in diagnoses by beneficiary
 read_years = function(years, template) {
   lapply(years, function(y) {
     read_tsv(sprintf(template, y)) %>%
@@ -21,14 +20,23 @@ read_years = function(years, template) {
     bind_rows()
 }
 
+# trends in diagnoses by beneficiary
 dx_by_bene = read_years(2011:2014, '../../data/dx_by_bene_%i.tsv') %>%
   rename(bene_id=BENE_ID, dx_cat=diagnosis_category)
 
 dxs = unique(dx_by_bene$dx_cat)
 
+# first, a table
+dx_trends_table = dx_by_bene %>%
+  group_by(year, dx_cat) %>%
+  summarize(n_dx=sum(n_dx)) %>%
+  ungroup() %T>%
+  output_table('dx_trends_table')
+
 glm_f = function(df, frmla, ...) eval(substitute(function(df2) glm(frmla, data=df2, ...)))(df)
 frmla = y ~ year + age + n_cc + sex + race + dual + region
 
+# then models
 dx_trends_models = lapply(dxs, function(dx) {
   dx_by_bene %>%
     filter(dx_cat==dx) %>%
@@ -63,7 +71,7 @@ dx_from_pde_table = crossing(antibiotic=top_abx, dx_cat=dxs) %>%
     # how what fraction of PDEs for this abx had this dx upstream?
     pde_denom %>%
       filter(antibiotic==a) %>%
-      left_join(filter(dx_from_pde, dx_cat==dx), by='pde_id') %>%
+      left_join(filter(dx_from_pde, antibiotic==a, dx_cat==dx), by='pde_id') %>%
       mutate(present=!is.na(dx_cat)) %>%
       count(present)
   })(.$antibiotic, .$dx_cat)) %T>%
@@ -73,26 +81,28 @@ dx_from_pde_table = crossing(antibiotic=top_abx, dx_cat=dxs) %>%
 # swo: test this part
 dx_to_pde = read_years(2011:2014, '../../data/dx_to_pde_%i.tsv') %>%
   rename(bene_id=BENE_ID, dx_cat=diagnosis_category) %>%
[…]

### 2017-09-01 ms1/ms1_display.Rmd
fixed or

@@
 mult_est = function(df, m) {
-  mutate(df, estimate=m*estimate, std.error=sqrt(m)*estimate)
+  mutate(df, estimate=m*estimate, std.error=sqrt(m)*std.error)
 }
@@
   spread(dx_cat, pct) %>%
-  head %>% kable
+  kable
 ```
@@
 
-```{r dx_to_pde_table}
+Second, I do logistic regressions.
+The unit of analysis is the diagnosis.
+The outcome is whether a PDE for that drug follows a diagnosis in that category.
+The main predictor is year; covariates are as above.
+The adjusted percent change is between 2011 and 2014.
+(I computed this adjusted change, which is a relative risk, using the odds ratio from the regression and the mean probability of PDE for that diagnosis in 2011.
+This strictly incorrect: the mean relative risk is the mean of the risks computed using the predicted probabilities for each 2011 diagnosis, *not* the relative risk computed using the mean probability in 2011.)
+
+```{r dx_to_pde}
 dx_to_pde_unadj = read_tsv('tables/dx_rx_table.tsv') %>%
@@
          ci_high=estimate+ci) %>%
-  select(dx_cat, antibiotic, estimate, ci_low, ci_high)
+  select(dx_cat, antibiotic, estimate, ci_low, ci_high) %>%
+  arrange(dx_cat, antibiotic)
 
-risk_ratio_f = function(or, p) or / (1 + p * (or - 1))
+risk_ratio_f = function(odr, p) odr / (1 + p * (odr - 1))
 
 inner_join(dx_to_pde_unadj, dx_to_pde_adj, by=c('dx_cat', 'antibiotic')) %>%
-  mutate(estimate=risk_ratio_f(estimate, `2011`),
-         ci_low=risk_ratio_f(ci_low, `2011`),
-         ci_high=risk_ratio_f(ci_high, `2011`),
+  # convert model estimates to odds ratios, then to risk ratios,
+  # then to adjusted fractional changes, then to percentages
+  mutate(or=exp(estimate),
+         or_low=exp(ci_low),
+         or_high=exp(ci_high),
+         rr=risk_ratio_f(or, `2011`),
+         rr_low=risk_ratio_f(or_low, `2011`),
+         rr_high=risk_ratio_f(or_high, `2011`),
+         adj_change=rr-1,
+         adj_change_low=rr_low-1,
+         adj_change_high=rr_high-1,
          pct_change=(`2014`-`2011`)/`2011`) %>%
   mutate_if(is.numeric, function(x) round(x*100, 2)) %>%
-  mutate(adj_change=sprintf('%.2f%% (%.2f -- %.2f%%)', estimate, ci_low, ci_high)) %>%
+  mutate(adj_change=sprintf('%.2f%% (%.2f -- %.2f%%)', adj_change, adj_change_low, adj_change_high)) %>%
   select(dx_cat, antibiotic, `2011`, `2014`, pct_change, adj_change) %>%
   kable
-```
-
-Second, I do logistic regressions.
-The unit of analysis is the diagnosis.
-The outcome is whether a PDE for that drug follows a diagnosis in that category.
-The main predictor is year; covariates are as above.
-The adjusted percent change is between 2011 and 2014.
-(I computed this adjusted change, which is a relative risk, using the odds ratio from the regression and the mean probability of PDE for that diagnosis in 2011.
-This strictly incorrect: the mean relative risk is the mean of the risks computed using the predicted probabilities for each 2011 diagnosis, *not* the relative risk computed using the mean probability in 2011.
-But it seems like it's decently close, and I think insisting on the thing that is technically correct will only create confusion.)
-
-I broke this into three tables by antibiotic, just to make it easier to navigate.
-
-```{r dx_to_pde_trends}
-
 ```

### 2017-09-01 ms1/ms1_display.Rmd
ms1 fixes

@@
   spread(year, dxpkp) %>%
-  mutate(unadj_pct_change=(`2014`/`2011`-1)*100) %>%
+  mutate(pct_change=(`2014`/`2011`-1)*100) %>%
   arrange(desc(`2011`)) %>%
@@
   model_ci() %>%
-  select(dx_cat, change_str, p.value)
+  select(dx_cat, adj_change=change_str, p.value)
 
-full_join(dx_unadj, dx_adj, by='dx_cat') %>%
+dx_trends = full_join(dx_unadj, dx_adj, by='dx_cat') %T>%
   kable
@@
 
-```{r pde_dx}
-read_tsv('tables/dx_from_pde.tsv') %>% 
+```{r dx_from_pde}
+dx_from_pde = read_tsv('tables/dx_from_pde.tsv') %>% 
   spread(present, n) %>%
@@
   select(antibiotic, dx_cat, pct) %>%
-  spread(dx_cat, pct) %>%
+  spread(dx_cat, pct) %T>%
   kable
@@
 
-inner_join(dx_to_pde_unadj, dx_to_pde_adj, by=c('dx_cat', 'antibiotic')) %>%
+dx_to_pde = inner_join(dx_to_pde_unadj,
+                       dx_to_pde_adj,
+                       by=c('dx_cat', 'antibiotic')) %>%
   # convert model estimates to odds ratios, then to risk ratios,
@@
   mutate(adj_change=sprintf('%.2f%% (%.2f -- %.2f%%)', adj_change, adj_change_low, adj_change_high)) %>%
-  select(dx_cat, antibiotic, `2011`, `2014`, pct_change, adj_change) %>%
+  select(dx_cat, antibiotic, `2011`, `2014`, pct_change, adj_change) %T>%
+  kable
+```
+
+## Combined tables
+
+```{r}
+x = dx_trends %>%
+  select(-p.value) %>%
+  gather('key', 'value', -dx_cat) %>%
+  spread(dx_cat, value) %>%
+  mutate(antibiotic='dx_trends')
+
+y = dx_from_pde %>%
+  mutate(key='dx_from_pde') %>%
+  mutate_if(is.numeric, function(x) as.character(round(x, 2)))
+
+z = dx_to_pde %>%
+  gather('key', 'value', -dx_cat, -antibiotic) %>%
+  spread(dx_cat, value)
+
+combined_table = bind_rows(x, y, z) %>%
+  mutate(antibiotic=factor(antibiotic, levels=c('dx_trends', top_abx)),
+         key=factor(key, levels=c('dx_from_pde', '2011', '2014', 'pct_change', 'adj_change'))) %>%
+  arrange(antibiotic, key)
+```
+
+### Respiratory stuff
+
+```{r respiratory}
+combined_table %>%
+  filter(antibiotic %in% c('dx_trends', 'azithromycin', 'levofloxacin', 'amoxicillin', 'amoxicillin/clavulanate')) %>%
+  select(antibiotic, key, not_infectious, asthma, bronchitis, other_resp, pharyngitis, pneumonia, sinusitis, uri) %>%
+  kable
+```
+
+### UTIs
+
+```{r uti}
+combined_table %>%
+  filter(antibiotic %in% c('dx_trends', 'ciprofloxacin', 'trimethoprim/sulfamethoxazole', 'nitrofurantoin')) %>%
+  select(antibiotic, key, not_infectious, uti, uti_t3) %>%
+  kable
+```
+
+### SSTI
+
+```{r ssti}
+combined_table %>%
+  filter(antibiotic %in% c('dx_trends', 'clindamycin', 'cephalexin', 'trimethoprim/sulfamethoxazole')) %>%
+  select(antibiotic, key, not_infectious, ssti, ssti_t3) %>%
+  kable
+```
+
+### GI
+
+```{r gi}
+combined_table %>%
+  filter(antibiotic %in% c('dx_trends', 'ciprofloxacin')) %>%
+  select(antibiotic, key, not_infectious, gi, gi_t3) %>%
   kable

### 2017-09-01 ms2/ms2_display.Rmd
ms1 fixes

@@
          pct_claims=pct(n_class_claims/sum(n_class_claims))) %>%
+  ungroup() %>%
   select(year, drug_group, n_claims, pct_bene, pct_claims) %>%

### 2017-09-01 ms1/ms1_3dx.R

@@ -50,6 +50,10 @@ dx_trends_models = lapply(dxs, function(dx) {
   bind_rows() %T>%
   output_table('dx_trends')
 
+# swo:
+# NB! SAS truncated the abx names in some files
+# esp., "trimethoprim/sulfamethoxazole" got chopped to "..metho"
+
 # which dx's contribute to each abx?
 # in 2011
 dx_from_pde = read_tsv(sprintf('../../data/dx_from_pde_%i.tsv', 2011)) %>%
@@ -80,13 +84,18 @@ dx_from_pde_table = crossing(antibiotic=top_abx, dx_cat=dxs) %>%
 # what are trends in prescribing practice?
 # swo: test this part
 dx_to_pde = read_years(2011:2014, '../../data/dx_to_pde_%i.tsv') %>%
+  # hack to fix the weird TMP/SMX values
+  mutate(antibiotic=if_else(antibiotic=='trimethoprim/sulfametho',
+                            'trimethoprim/sulfamethoxazole',
+                            antibiotic)) %>%
   rename(bene_id=BENE_ID, dx_cat=diagnosis_category) %>%
   replace_na(list(antibiotic='no_abx')) %>%
   # drop no infection and no abx, since we won't look at those
   filter(!(dx_cat=='not_infectious' & antibiotic=='no_abx')) %>%
   # lump abx outside of top 10 into "other"
   mutate(antibiotic=fct_other(factor(antibiotic), keep=c(top_abx, 'no_abx'))) %>%
-  count(year, bene_id, dx_cat, antibiotic)
+  count(year, bene_id, dx_cat, antibiotic) %>%
+  left_join(bene, by=c('year', 'bene_id'))
 
 dx_rx_table = dx_to_pde %>%
   group_by(year, dx_cat, antibiotic) %>%
@@ -104,5 +113,6 @@ dx_rx_trends = crossing(antibiotic=top_abx, dx_cat=dxs) %>%
       mutate(y=antibiotic==a) %>%
       glm_f(frmla, family='binomial', weights=n) %>%
       tidy
-  })(.$antibiotic, .$dx_cat)) %T>%
+  })(.$antibiotic, .$dx_cat)) %>%
+  ungroup() %T>%
   output_table('dx_rx_trends')

### 2017-09-01 ms2/ms2_2analyze.R

@@ -120,7 +120,7 @@ beta_model = function(df_raw, y, x) {
     mutate(y=case_when(.$y == 0 ~ eps,
                        .$y == 1 ~ 1 - eps,
                        TRUE ~ .$y))
-  
+
   frmla = as.formula(str_interp("${y} ~ ${x}"))
   m = betareg::betareg(formula=frmla, weights=n_place_antibiograms, data=df, link='logit')
   

### 2017-09-06 ms1/ms1_display.Rmd
simpler beta reg

@@
   geom_point() + geom_line() +
-  #ylim(0, 1520) +
-  scale_y_continuous(limits=c(0, 1525), breaks=c(0, 250, 500, 1000, 1500)) +
+  scale_y_continuous(limits=c(975, 1525), breaks=c(1000, 1125, 1250, 1375, 1500)) +
   theme_minimal() +
@@
   ggplot(aes(x=year, y=cpkp, color=antibiotic)) +
-  #ggplot(aes(x=year, y=cpkp)) +
-  #facet_wrap(~antibiotic, ncol=5) +
   geom_point() + geom_line() +
@@
   ggplot(aes(x=year, y=cpkp, color=antibiotic)) +
-  #ggplot(aes(x=year, y=cpkp)) +
-  #facet_wrap(~antibiotic, ncol=5) +
   geom_point() + geom_line() +
   scale_y_continuous(limits=c(0, 100), breaks=c(0, 50, 100)) +
+  #ylim(0, 140) +
+  theme_minimal() +
+  ggtitle('Unadjusted consumption, next 5')
+```
+
+```{r claims_by_top_abx_facet, eval=FALSE}
+claims_by_top_abx %>%
+  filter(antibiotic %in% c('azithromycin', 'ciprofloxacin', 'amoxicillin', 'cephalexin', 'trimethoprim/sulfamethoxazole')) %>%
+  ggplot(aes(x=year, y=cpkp, color=antibiotic)) +
+  #ggplot(aes(x=year, y=cpkp)) +
+  facet_wrap(~antibiotic, ncol=5) +
+  geom_point() + geom_line() +
+  scale_y_continuous(limits=c(0, 250), breaks=c(0, 50, 100, 150, 200, 250)) +
+  theme_minimal() +
+  ggtitle('Unadjusted consumption, top 5')
+
+claims_by_top_abx %>%
+  filter(antibiotic %in% c('levofloxacin', 'amoxicillin/clavulanate', 'doxycycline', 'nitrofurantoin', 'clindamycin')) %>%
+  ggplot(aes(x=year, y=cpkp, color=antibiotic)) +
+  #ggplot(aes(x=year, y=cpkp)) +
+  facet_wrap(~antibiotic, ncol=5) +
+  geom_point() + geom_line() +
+  scale_y_continuous(limits=c(0, 150), breaks=c(0, 50, 100)) +
   #ylim(0, 140) +
@@
   geom_errorbar(width=0.5) +
-  geom_point(dat=unadjusted_changes, aes(x=antibiotic, y=pct_change, ymin=NULL, ymax=NULL), color='red') +
+  geom_point(dat=unadjusted_changes, aes(x=antibiotic, y=pct_change, ymin=NULL, ymax=NULL), shape=1) +
   coord_flip() +
@@
 
-```{r fq}
+```{r fq_table}
 fq_names = claims_by_abx %>%
@@
 
+fq_table = claims_by_abx %>%
+  filter(antibiotic %in% fq_names) %>%
+  group_by(year) %>%
+  summarize(cpkp=sum(cpkp))
+
+fq_change_pct_unadj = fq_table %>%
+  spread(year, cpkp) %$%
+  { (`2014`-`2011`)/`2011` } %>%
+  { round(100*., 2) }
+```
+
+```{r fq_trends}
 fq_trends = read_tsv('tables/model_fq.tsv') %>%
@@
   model_ci() %>%
-  select(term, pct_change, ci_low, ci_high, p.value) %T>%
-  kable
+  select(term, pct_change, ci_low, ci_high, p.value)
 
-fq_change_pct = fq_trends %>%
+fq_change_adj = fq_trends %>%
   filter(term=='year') %>% 
-  pull(pct_change) %>%
-  round(2)
+  mutate_at(vars(pct_change, ci_low, ci_high), function(x) round(x, 2)) %$%
+  str_interp("${pct_change}% (${ci_low}% -- ${ci_high}%)")
 ```
 
-Overall, fluoroquinolones changed by a tiny amount, `r fq_change_pct`%, over this period. This included `r fq_names`.
+Overall, fluoroquinolones changed by a tiny amount, `r fq_change_pct_unadj`% (unadjusted). The adjusted change is `r fq_change_adj` over this period. This included `r fq_names`.
 
@@
 
-dx_trends = full_join(dx_unadj, dx_adj, by='dx_cat') %T>%
-  kable
+dx_trends = full_join(dx_unadj, dx_adj, by='dx_cat')
+
+dx_trends %>% kable
 ```
@@
 
+All respiratory diagnoses fell, bronchitis the most sharply. Azithromycin prescribing fell for all diagnoses, while levofloxacin, amox, and amox/clav prescribing rose for all diagnoses.
+
 ```{r respiratory}
@@
 ### UTIs
+
+UTIs declined somewhat in prevalence. Ciprofloxacin prescribing remained nearly constant, nitrofurantion fell 3%, and Bactrim rose 4-6%.
 
@@
 
+Although skin and soft tissue infections declined in prevalence, prescribing of clindamycin, cephalexin, Bactrim, amox, and amox/clav increased, with one exception: amox/clav use for Tier 2 SSTIs stay nearly constant.
+Clindamycin was the biggest winner.
+
 ```{r ssti}
+# swo: I removed amox and amox/clav because they mostly
+# showed trends in Tier 3, which is confusing
 combined_table %>%
@@
 
+GI indications became less prevalent, and cipro prescribing fell 1-2%.
+Only a small amount of cipro is used for GI indications: compare the 5-6% of cipro PDEs preceded by GI indications with the 14-20% preceded by UTIs.
+
 ```{r gi}
@@
 
-# Summary of results for each drug
+# Questions for the paper
 
-For azithromycin, the most common diagnoses are bronchitis, URI, and sinusitis.
-Bronchitis diagnoses are on a steep decline (15% over the study period), URI
-diagnoses rose slightly (~5%), and sinusitis diagnoses remained nearly constant
-(+0.13%). Azithromycin prescribing per diagnosis also fell during the study
-period for each of these diagnoses (10-15%). So the diagnoses for which
-azithromycin is used were overall less common, and azithromycin was also less
-used for all those diagnoses.
-
-For levofloxacin, the most common diagnoses are bronchitis, "other
-respiratory", and URI. Bronchitis fell steeply, other respiratory fell somewhat
-(5%) and URIs rose slightly (2%). However, levofloxacin prescribing rose
-sharply for bronchitis (75%), other respiratory (60%), and URI (75%). So even
-though bronchitis diagnosing fell, levo prescribing for its diagnoses rose so
-much.
-
-The most common diagnoses for clindamycin, which rose, were SSTIs (tier 2 and
-tier 3). Tier 3 diagnoses rose somewhat (2%) and Tier 2 diagnoses fell somewhat
-(4%). Clindamycin prescribing for Tier 2 and Tier 3 diagnoses rose (27%, 22%).
-So clinda simply became a more popular treatment for SSTIs.
-
-The most common diagnoses for amox/clav, which rose, were sinusitis, SSTIs,
-bronchitis, and URIs. Amox/clav prescribing rose sharply for sinusitis (42%),
-rose for Tier 3 SSTIs (10%) but not Tier 2 SSTIs (n.s.), and rose for
-bronchitis (29%) and URIs (36%). So amox/clav became more popular for most of
-its diagnoses, like levo.
-
-The most common diagnoses for amox, which rose, were similar to those for
-amox/clav: sinusitis, SSTIs, bronchitis, and URIs. Amox prescribing rose
-slightly for sinusitis (3%), rose for Tier 2 and Tier 3 SSTIs (7%, 10%), and
-rose for bronchitis (12%) and URI (9%). The story is similar to amox/clav, just
-not as pronounced.
-
-The most common diagnoses for cephalexin, which rose slightly, were SSTIs, and
-somewhat UTIs. Prescribing for SSTIs rose (6%) and rose more for UTIs (32% for
-Tier 1 UTIs, 20% for Tier 3). So there was some growth in SSTIs, which make up
-more of cephalexin's portfolio, and big growth in UTIs, which make up a small
-part.
-
-The most common diagnoses for Bactrim, which fell slightly, are UTIs and
-SSTIs.  UTIs are nearly constant and SSTIs increased a little. However, Bactrim
-prescribing *increased* for SSTIs and UTIs. This means that the decrease
-occurred among the PDEs that were not associated with diagnoses. My first guess
-is that there is less long-term treatment (since, in long-term treatment, most
-of the PDEs will nt be associated with a visit).
-
-The most common diagnoses for doxy, which fell, are SSTIs and bronchitis.
-Doxy prescribing increased for Tier 2 SSTIs (5%), remained constant for Tier 3
-SSTIs (n.s.), and decreased slightly for bronchitis (2%). So it seems there was
-not big change in prescribing practice, and maybe this is more a reflection of
-what gets diagnosed as bronchitis.
-
-The most common diagnoses for cipro, which fell, are UTIs, followed by GI
-infections. GI infections fell (5%). Cipro prescribing for UTIs and GI
-infections remained constant (n.s. for all diagnoses). My first guess is that
-cipro fell because GI infection diagnoses fell and because of a decrease in
-long-term treatment for UTIs, like in Bactrim.
-
-The most common diagnoses for nitrofurantoin, which fell, were UTIs. UTIs
-remained nearly constant, and nitro prescribing for UTIs only fell slightly (3%
-for Tier 1 UTIs, n.s. for Tier 3). Again, my first guess is that there was a
-decrease in the number of PDEs not associated with diagnoses, i.e., those in
-long-term treatment.
-
-## Questions for the paper
-
-### Breaking up the tables
-
-I don't want to show this enormous table in the main text. I might break this into three tables:
-
-- Skin and soft tissue infections (Tiers 2 and 3): clindamycin, cephalexin, Bactrim, doxy, amox, amox/clav
-- Respiratory conditions (bronchitis, URI, sinusitis, "other respiratory"): azithro, levo, amox, amox/clav, doxy
-- UTIs (Tier 1 infections and Tier 3 conditions): cipro, Bactrim, nitro, cephalexin
-- GI (optional): but this is just cipro
-
-Alternatively, I could keep one table that has these diagnoses as columns and all these drugs going down as rows.
-
-### Lumping together diagnoses.
+## Lumping together diagnoses
 

### 2017-09-06 ms2/ms2_2analyze.R

@@ -84,22 +84,22 @@ linear_model = function(df, y, xs) {
     tidy %>%
     filter(term != 'Residuals') %>%
     select(term, anova.p.value=p.value)
- 
+
   coef_res = tidy(m) %>%
     mutate(ci=1.96*std.error,
            ci_low=estimate-ci,
            ci_high=estimate+ci) %>%
     select(term, estimate, ci_low, ci_high, p.value)
- 
+
   left_join(coef_res, anova_res, by='term')
 }
 
 spearman_model = function(df, y_name, x_name) {
   y = df[[y_name]]
   x = df[[x_name]]
-  
+
   m = cor.test(y, x, method='spearman')
-  
+
   if (length(y) > 3) {
     cis = mada::CIrho(m$estimate, length(y), level=0.95)
   } else {
@@ -113,37 +113,34 @@ spearman_model = function(df, y_name, x_name) {
              p.value=m$p.value)
 }
 
-beta_model = function(df_raw, y, x) {
+beta_model = function(df_raw, y, xs, link='logit', eps=1e-6) {
   # replace 0 and 1 with slightly different values
-  eps = 1e-6
   df = df_raw %>%
-    mutate(y=case_when(.$y == 0 ~ eps,
-                       .$y == 1 ~ 1 - eps,
+    mutate(y=case_when(.$y==0 ~ eps,
+                       .$y==1 ~ 1 - eps,
                        TRUE ~ .$y))
 
-  frmla = as.formula(str_interp("${y} ~ ${x}"))
-  m = betareg::betareg(formula=frmla, weights=n_place_antibiograms, data=df, link='logit')
-  
-  xbar = weighted.mean(df[[y]], w=df$n_place_antibiograms)
-  newdata = setNames(list(xbar), x) %>% as_data_frame
+  frmla = as.formula(str_interp("${y} ~ ${str_c(xs, collapse=' + ')}"))
+  m = betareg::betareg(formula=frmla, weights=n_place_antibiograms, data=df, link=link)
 
-  yhat = predict(m, newdata=newdata, type='response')
-  slope = m$coefficients$mean[[x]] * yhat * (1 - yhat)
-  pvalue = tidy(m) %>% filter(term==x) %>% pull(p.value)
-  data_frame(term=x, estimate=slope, p.value=pvalue)
+  coef_res = tidy(m) %>%
+    mutate(ci=1.96*std.error,
[…]

### 2017-09-11 ms1/ms1_display.Rmd
updated tables

@@
            ci_high=est2pct(estimate + ci),
-           change_str=sprintf('%.2f%% (%.2f -- %.2f%%)', pct_change, ci_low, ci_high))
+           change_str=sprintf('%.2f (%.2f -- %.2f)', pct_change, ci_low, ci_high))
 }
@@
 
-Overall, fluoroquinolones changed by a tiny amount, `r fq_change_pct_unadj`% (unadjusted). The adjusted change is `r fq_change_adj` over this period. This included `r fq_names`.
+Overall, fluoroquinolones changed by a tiny amount, `r fq_change_pct_unadj`% (unadjusted). The adjusted change is `r fq_change_adj` (%) over this period. This included `r fq_names`.
 
@@
 dx_unadj = read_tsv('tables/dx_trends_table.tsv') %>%
-  filter(year %in% c(2011, 2014)) %>%
+  #filter(year %in% c(2011, 2014)) %>%
   left_join(select(totals, year, n_bene), by='year') %>%
@@
   mutate_if(is.numeric, function(x) round(x*100, 2)) %>%
-  mutate(adj_change=sprintf('%.2f%% (%.2f -- %.2f%%)', adj_change, adj_change_low, adj_change_high)) %>%
+  mutate(adj_change=sprintf('%.2f (%.2f -- %.2f)', adj_change, adj_change_low, adj_change_high)) %>%
   select(dx_cat, antibiotic, `2011`, `2014`, pct_change, adj_change) %T>%
@@
 x = dx_trends %>%
-  select(-p.value) %>%
-  gather('key', 'value', -dx_cat) %>%
+  select(dx_cat, `2011`:`2014`) %>%
+  gather('year', 'value', `2011`:`2014`) %>%
+  group_by(dx_cat) %>%
+  summarize(value=mean(value)) %>%
+  # convert to character to fit with other parts
+  mutate(value=as.character(value)) %>%
   spread(dx_cat, value) %>%
-  mutate(antibiotic='dx_trends')
+  mutate(antibiotic='dx', key='mean_per_1kyr')
+  # select(-p.value) %>%
+  # gather('key', 'value', -dx_cat) %>%
+  # spread(dx_cat, value) %>%
+  # mutate(antibiotic='dx_trends')
 
@@
 combined_table = bind_rows(x, y, z) %>%
-  mutate(antibiotic=factor(antibiotic, levels=c('dx_trends', top_abx)),
-         key=factor(key, levels=c('dx_from_pde', '2011', '2014', 'pct_change', 'adj_change'))) %>%
+  mutate(antibiotic=factor(antibiotic, levels=c('dx', top_abx)),
+         key=factor(key, levels=c('mean_per_1kyr', 'dx_from_pde', '2011', '2014', 'pct_change', 'adj_change'))) %>%
   arrange(antibiotic, key)
@@
 combined_table %>%
-  filter(antibiotic %in% c('dx_trends', 'azithromycin', 'levofloxacin', 'amoxicillin', 'amoxicillin/clavulanate')) %>%
-  select(antibiotic, key, not_infectious, asthma, bronchitis, other_resp, pharyngitis, pneumonia, sinusitis, uri) %>%
+  filter(antibiotic %in% c('dx', 'azithromycin', 'levofloxacin', 'amoxicillin', 'amoxicillin/clavulanate')) %>%
+  select(antibiotic, key, asthma, bronchitis, other_resp, pharyngitis, pneumonia, sinusitis, uri) %>%
   kable
@@
 combined_table %>%
-  filter(antibiotic %in% c('dx_trends', 'ciprofloxacin', 'trimethoprim/sulfamethoxazole', 'nitrofurantoin')) %>%
-  select(antibiotic, key, not_infectious, uti, uti_t3) %>%
+  filter(antibiotic %in% c('dx', 'ciprofloxacin', 'trimethoprim/sulfamethoxazole', 'nitrofurantoin')) %>%
+  select(antibiotic, key, uti, uti_t3) %>%
   kable
@@
 combined_table %>%
-  filter(antibiotic %in% c('dx_trends', 'clindamycin', 'cephalexin', 'trimethoprim/sulfamethoxazole')) %>%
-  select(antibiotic, key, not_infectious, ssti, ssti_t3) %>%
+  filter(antibiotic %in% c('dx', 'clindamycin', 'cephalexin', 'trimethoprim/sulfamethoxazole')) %>%
+  select(antibiotic, key, ssti, ssti_t3) %>%
   kable
@@
 combined_table %>%
-  filter(antibiotic %in% c('dx_trends', 'ciprofloxacin')) %>%
-  select(antibiotic, key, not_infectious, gi, gi_t3) %>%
+  filter(antibiotic %in% c('dx', 'ciprofloxacin')) %>%
+  select(antibiotic, key, gi, gi_t3) %>%
   kable

### 2017-09-15 ms1/ms1_display.Rmd
added appropriateness

@@
 
+To compute appropriateness, I classified each PDE with the maximum appropriateness of all its linked diagnoses.
+I put PDEs with no associated infectious disease (i.e., FD) diagnoses into their own, lower tier.
+
+## Trends in appropriateness
+
+Just by eye, it looks like the changes in appropriateness are small.
+
+```{r approp_graph}
+pde_approp_table = read_tsv('tables/pde_approp_table.tsv') %>%
+  filter(year %in% c(2011, 2014)) %>%
+  (function(df) bind_rows(
+    df,
+    df %>% group_by(year, tier) %>%
+      summarize(n=sum(n)) %>%
+      mutate(antibiotic='overall')
+  )) %>%
+  filter(antibiotic %in% c(top_abx, 'overall')) %>%
+  group_by(year, antibiotic) %>%
+  mutate(f=n/sum(n)) %>%
+  ungroup()
+
+pde_approp_table %>%
+  ggplot(aes(x=factor(year), y=f, fill=factor(tier))) +
+  geom_bar(stat='identity') +
+  facet_grid(~antibiotic) +
+  theme_classic()
+```
+
+I'll focus on tier 1 and 2 because the SSTI drugs don't have common tier 1 uses.
+I'll call this "high tier" prescribign.
+If prescribing had changed in an optimal way, only low-tier prescribing would have declined, or only high-tier prescribing would have been added.
+
+If a drug declined by $X$, then the optimal change in the high tier fraction would be $1/(1-X) \approx 1+X$.
+For example, overall prescribing fell by about 5%, and high-tier prescribing increased by 2.5%.
+Azithromycin prescribing fell by about 20%, and high-tier azithro prescribing rose by only 6%.
+
+I'm not sure what to make of these results, because they depend strongly on how you define the tiers.
+If you use only tier 1 to describe "appropriate", then levofloxacin has the biggest decline in appropriateness, but if you use tiers 1 and 2, then it's change isn't significant.
+
+```{r pre_approp_table}
+pde_approp_trends = read_tsv('tables/pde_approp_trends.tsv') %>%
+  filter(term=='year') %>%
+  mutate(or=exp(3*estimate), adj.p=p.adjust(p.value, method='BH')) %>%
+  select(antibiotic, or, adj.p)
+
+risk_ratio_f = function(odr, p) odr / (1 + p * (odr - 1))
+
+pde_approp_table %>%
+  filter(tier <= 2) %>%
+  select(antibiotic, year, f) %>%
+  group_by(antibiotic, year) %>%
+  summarize(f=sum(f)) %>%
+  ungroup() %>%
+  spread(year, f) %>%
+  left_join(pde_approp_trends, by='antibiotic') %>%
+  mutate(raw_pct_change=(`2014`-`2011`)/`2011`,
+         adj_pct_change=risk_ratio_f(or, `2011`) - 1) %>%
+  mutate_at(vars(`2011`, `2014`, raw_pct_change, adj_pct_change),
+            function(x) round(x*100, 2)) %>%
+  select(antibiotic, `2011`, `2014`, raw_pct_change, adj_pct_change, adj.p) %>%
+  arrange(desc(`2011`)) %>%
+  kable
+```
+
+And, to be more careful, I show the answer to a similar question but using a regression to correct for covariates:
+
+```{r pde_approp_trends}
+
+```
+
 ## Trends in diagnoses

### 2017-09-15 ms1/ms1_3dx.R

@@ -21,65 +21,106 @@ read_years = function(years, template) {
 }
 
 # trends in diagnoses by beneficiary
-dx_by_bene = read_years(2011:2014, '../../data/dx_by_bene_%i.tsv') %>%
-  rename(bene_id=BENE_ID, dx_cat=diagnosis_category)
-
-dxs = unique(dx_by_bene$dx_cat)
-
-# first, a table
-dx_trends_table = dx_by_bene %>%
-  group_by(year, dx_cat) %>%
-  summarize(n_dx=sum(n_dx)) %>%
-  ungroup() %T>%
-  output_table('dx_trends_table')
+# dx_by_bene = read_years(2011:2014, '../../data/dx_by_bene_%i.tsv') %>%
+#   rename(bene_id=BENE_ID, dx_cat=diagnosis_category)
+# 
+# dxs = unique(dx_by_bene$dx_cat)
+# 
+# # first, a table
+# dx_trends_table = dx_by_bene %>%
+#   group_by(year, dx_cat) %>%
+#   summarize(n_dx=sum(n_dx)) %>%
+#   ungroup() %T>%
+#   output_table('dx_trends_table')
 
 glm_f = function(df, frmla, ...) eval(substitute(function(df2) glm(frmla, data=df2, ...)))(df)
 frmla = y ~ year + age + n_cc + sex + race + dual + region
 
 # then models
-dx_trends_models = lapply(dxs, function(dx) {
-  dx_by_bene %>%
-    filter(dx_cat==dx) %>%
-    right_join(bene, by=c('year', 'bene_id')) %>%
-    replace_na(list(n_dx=0)) %>%
-    rename(y=n_dx) %>%
-    glm_f(frmla, family='poisson') %>%
-    tidy %>%
-    mutate(model=dx)
-}) %>%
-  bind_rows() %T>%
-  output_table('dx_trends')
+# dx_trends_models = lapply(dxs, function(dx) {
+#   dx_by_bene %>%
+#     filter(dx_cat==dx) %>%
+#     right_join(bene, by=c('year', 'bene_id')) %>%
+#     replace_na(list(n_dx=0)) %>%
+#     rename(y=n_dx) %>%
+#     glm_f(frmla, family='poisson') %>%
+#     tidy %>%
+#     mutate(model=dx)
+# }) %>%
+#   bind_rows() %T>%
+#   output_table('dx_trends')
 
 # swo:
 # NB! SAS truncated the abx names in some files
[…]

### 2017-09-25 ms1/ms1_display.Rmd
ms1 dx

@@
   select(antibiotic, `2011`, `2014`, raw_pct_change, adj_pct_change, adj.p) %>%
-  arrange(desc(`2011`)) %>%
+  arrange(raw_pct_change) %>%
   kable

### 2017-09-25 ms1/ms1_3dx.R

@@ -107,20 +107,32 @@ pde_approp_table = pde_approp %>%
 # logistic regression
 # two ways to do this: either keep the "not infectious" diagnoses in the
 # denominator, or leave them out (i.e., filter tier != 4)
+# swo: i think this formula is repeated
 frmla = y ~ year + age + n_cc + sex + race + dual + region
+approp_trends_f = function(df) {
+  df %>%
+    mutate(y=tier <= 2) %>%
+    (function(df) {
+      bind_rows(
+        glm_f(df, frmla, family='binomial') %>% tidy %>% mutate(antibiotic='overall'),
+        df %>%
+          filter(antibiotic %in% top_abx) %>%
+          group_by(antibiotic) %>%
+          do(tidy(glm_f(., frmla, family='binomial'))) %>%
+          ungroup()
+      )
+    })
+}
+
 pde_approp_trends = pde_approp %>%
-  mutate(y=tier <= 2) %>%
-  (function(df) {
-    bind_rows(
-      glm_f(df, frmla, family='binomial') %>% tidy %>% mutate(antibiotic='overall'),
-      df %>%
-        filter(antibiotic %in% top_abx) %>%
-        group_by(antibiotic) %>%
-        do(tidy(glm_f(., frmla, family='binomial'))) %>%
-        ungroup()
-    )
-  }) %T>%
+  filter(fill_num==0) %>% # leave out the refills
+  approp_trends_f() %T>%
   output_table('pde_approp_trends')
+
+pde_approp_trends_wrefills = pde_approp %>%
+  # don't filter out the refills
+  approp_trends_f() %T>%
+  output_table('pde_approp_trends_wrefills')
 
 # what are trends in prescribing practice?
 # swo: test this part
@@ -157,3 +169,19 @@ dx_rx_trends = crossing(antibiotic=top_abx, dx_cat=dxs) %>%
   })(.$antibiotic, .$dx_cat)) %>%
   ungroup() %T>%
   output_table('dx_to_pde_trends')
+
+# pde appropriateness at the margins
+pde_approp_margin = pde_approp %>%
+  # filter out refills if desired!
+  mutate(antibiotic=if_else(str_detect(antibiotic, '^trimethoprim/sulfa'), 'tmp_smx', antibiotic)) %>%
+  filter(antibiotic %in% c('tmp_smx', top_abx)) %>%
[…]

### 2017-09-25 ms2/ms2_display.Rmd
new graphs
cleaner ms2 display

@@
 
+### Effect along extensive margin $\mu_>$
+
+```{r hrr_linear_mup}
+results %>%
+  filter(model=='univariate_mup', unit_type=='hrr', term=='mup') %>%
+  interval_plot() +
+  ylab(expression(beta[nu])) +
+  ggtitle("Linear univariate, HRR")
+```
+
+### Effect beyond $f_>$
+
 We ask if mean consumption $\mu$ has any information beyond fraction nonzero $f_>$ for predicting resistance $R$.
-We might expect that the simple two-part model represented by the regressions $R \sim f_>$ doesn't capture the fact that taking two doses should have a greater effect than taking one, etc.
+We might expect that the simple two-part model (i.e., consuming or not) represented by the regressions $R \sim f_>$ doesn't capture the fact that taking two doses should have a greater effect than taking one, etc.
 
-These plots show the coefficient $\beta_\mu$ from the regression $R^{(u)} = \beta_0 + \beta_{f_>} f_>^{(u)} + \beta_\mu \mu^{(u)} + \varepsilon^{(u)}$.
+These plots show the coefficient $\beta_\nu$ from the regression $R = \beta_0 + \beta_{f_>} f_> + \beta_\nu \nu + \varepsilon$.
 
@@
   interval_plot() +
-  ylab(expression(beta[mu['>']])) +
-  ggtitle("Linear multivariate (f> and mu>), HRR")
+  ylab(expression(beta[nu])) +
+  ggtitle("Linear multivariate (f> and nu), HRR")
 ```
 
-At the state level, none of the $\beta_\mu$ coefficients are significantly different from zero (after multiple-hypothesis correction).
+At the state level, none of the $\beta_\nu$ coefficients are significantly different from zero (after multiple-hypothesis correction).
 
@@
 results %>%
-  filter(model=='multivariate_fnz_mean', unit_type=='state', term=='mean') %>%
+  filter(model=='multivariate_fnz_mup', unit_type=='state', term=='mup') %>%
   interval_plot() +
-  ylab(expression(beta[mu])) +
-  ggtitle("Linear multivariate (f> and mean), state")
+  ylab(expression(beta[nu])) +
+  ggtitle("Linear multivariate (f> and nu), state")
 ```

### 2017-09-25 ms2/ms2_display.Rmd
improved lm summaries

@@
 # Use the zipcode-HRR mapping from 2011
-hrr = read_tsv('../../db/hrr/hrr.tsv') %>%
+hrr = read_tsv('../../db/hrr/hrr.tsv', col_types=list(hrr='c')) %>%
   filter(year==2011) %>%
@@
 state_abg = read_tsv('abg_state.tsv')
-hrr_abg = read_tsv('abg_hrr.tsv')
+hrr_abg = read_tsv('abg_hrr.tsv', col_types=list(hrr='c'))
 ```
@@
   #rename(state=unit) %>% left_join(regions, by='state') %>%
-  mutate(hrr=as.integer(unit)) %>% left_join(hrr_regions) %>%
+  rename(hrr=unit) %>% left_join(hrr_regions, by='hrr') %>%
   ggplot(aes(x=fnz, y=mup, color=region)) +
@@
   filter(unit_type=='hrr') %>%
-  mutate(hrr=as.integer(unit)) %>% left_join(hrr_regions, by='hrr') %>%
+  rename(hrr=unit) %>% left_join(hrr_regions, by='hrr') %>%
   right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
   filter(bug=='E. coli', drug_group=='tmp_smx') %>%
-  filter(mean_percent_nonsusceptible < 50) %>% # filter one crazy Nevada point
-  ggplot(aes(x=fnz, y=mean_percent_nonsusceptible, color=region)) +
+  filter(percent_nonsusceptible < 50) %>% # filter one crazy Nevada point
+  ggplot(aes(x=fnz, y=percent_nonsusceptible, color=region)) +
   geom_abline(intercept=16.86, slope=114.8, color='pink') +
@@
   filter(bug=='E. coli', drug_group=='tmp_smx') %>%
-  ggplot(aes(x=fnz, y=mean_percent_nonsusceptible, color=region)) +
+  ggplot(aes(x=fnz, y=percent_nonsusceptible, color=region)) +
   geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
@@
   filter(unit_type=='hrr') %>%
-  mutate(hrr=as.integer(unit)) %>% left_join(hrr_regions, by='hrr') %>%
+  rename(hrr=unit) %>% left_join(hrr_regions, by='hrr') %>%
   right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
   filter(bug=='E. coli', drug_group=='tmp_smx') %>%
-  filter(mean_percent_nonsusceptible < 50) %>% # filter one crazy Nevada point
-  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region)) +
+  filter(percent_nonsusceptible < 50) %>% # filter one crazy Nevada point
+  ggplot(aes(x=mean, y=percent_nonsusceptible, color=region)) +
   geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
@@
   filter(unit_type=='hrr') %>%
-  mutate(hrr=as.integer(unit)) %>% left_join(hrr_regions, by='hrr') %>%
+  rename(hrr=unit) %>% left_join(hrr_regions, by='hrr') %>%
   right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
   filter(bug=='Ec. faecium', drug_group=='beta_lactam') %>%
-  ggplot(aes(x=fnz, y=mean_percent_nonsusceptible, color=region)) +
+  ggplot(aes(x=fnz, y=percent_nonsusceptible, color=region)) +
   geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
@@
   filter(bug=='Ec. faecium', drug_group=='beta_lactam') %>%
-  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region)) +
+  ggplot(aes(x=mean, y=percent_nonsusceptible, color=region)) +
   geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
@@
   filter(unit_type=='hrr') %>%
-  mutate(hrr=as.integer(unit)) %>% left_join(hrr_regions, by='hrr') %>%
+  rename(hrr=unit) %>% left_join(hrr_regions, by='hrr') %>%
   right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
   filter(bug=='S. pneumoniae', drug_group=='tetracycline') %>%
-  ggplot(aes(x=fnz, y=mean_percent_nonsusceptible, color=region)) +
+  ggplot(aes(x=fnz, y=percent_nonsusceptible, color=region)) +
   geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
@@
   filter(bug=='S. pneumoniae', drug_group=='tetracycline') %>%
-  ggplot(aes(x=fnz, y=mean_percent_nonsusceptible, color=region)) +
+  ggplot(aes(x=fnz, y=percent_nonsusceptible, color=region)) +
   geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
@@
   filter(unit_type=='hrr') %>%
-  mutate(hrr=as.integer(unit)) %>% left_join(hrr_regions, by='hrr') %>%
+  rename(hrr=unit) %>% left_join(hrr_regions, by='hrr') %>%
   right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
   filter(bug=='E. coli', drug_group=='beta_lactam') %>%
-  filter(mean_percent_nonsusceptible < 50) %>% # filter one crazy Nevada point
-  ggplot(aes(x=fnz, y=mean_percent_nonsusceptible, color=region)) +
+  filter(percent_nonsusceptible < 50) %>% # filter one crazy Nevada point
+  ggplot(aes(x=fnz, y=percent_nonsusceptible, color=region)) +
   geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
@@
   filter(bug=='E. coli', drug_group=='beta_lactam') %>%
-  filter(mean_percent_nonsusceptible < 50) %>% # filter one crazy point
-  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region)) +
+  filter(percent_nonsusceptible < 50) %>% # filter one crazy point
+  ggplot(aes(x=mean, y=percent_nonsusceptible, color=region)) +
   geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
@@
       filter(bug==b, drug_group==d) %>%
-      select(state, mean_percent_nonsusceptible),
+      select(state, percent_nonsusceptible),
     by='state') %>%
     left_join(regions, by='state') %>%
-    ggplot(aes(x=mean, y=mean_percent_nonsusceptible, label=state, color=region)) +
+    ggplot(aes(x=mean, y=percent_nonsusceptible, label=state, color=region)) +
     geom_smooth(method='lm', se=F) +

### 2017-09-25 ms2/ms2_2analyze.R

@@ -22,7 +22,7 @@ zipcodes = read_tsv('../../db/zipcode/zipcode.tsv') %>%
   select(zipcode=zip, state)
 
 # Just use 2014 HRRs
-hrr = read_tsv('../../db/hrr/hrr.tsv') %>%
+hrr = read_tsv('../../db/hrr/hrr.tsv', col_types=list(hrr='c')) %>%
   filter(year==2014) %>%
   select(zipcode, hrr) %>%
   left_join(zipcodes, by='zipcode') %>%
@@ -55,17 +55,39 @@ abg = read_tsv('../../../antibiogram/data/abg.tsv', col_types=cols(zipcode='c'))
 # check that all the short names made it in
 if (any(is.na(abg$bug))) stop('missing bug names')
 
+# summarize antibiograms by just taking the sum of the isolates
+# regression weight is the number of antibiograms
+summarize_abg = function(df, place) {
+  place_var = enquo(place)
+  place_str = quo_name(place_var)
+
+  df %>%
+    group_by(bug, drug_group, !!place_var) %>%
+    summarize(n_place_antibiograms=n(),
+              percent_nonsusceptible=weighted.mean(percent_nonsusceptible, w=n_isolates)) %>%
+    ungroup() %>%
+    mutate(regression_weight=n_place_antibiograms)
+}
+
+# summarize antibiograms by correcting for in- vs. outpatient
+# regression weight goes like the inverse square of the standard error
+# (i.e., it's easier to estimate outpatient for some states)
 summarize_abg = function(df, place_name) {
   place_name_var = enquo(place_name)
   place_name_str = quo_name(place_name_var)
   
   frmla = str_interp("percent_nonsusceptible ~ ${place_name_str} + has_inpatient + 0") %>% as.formula
+  place_name_regex = str_c('^', place_name_str)
   
   models = df %>%
     group_by(bug, drug_group) %>%
-    do(tidy(lm(frmla, data=., weights=median_n_isolates))) %>%
-    mutate(!!place_name_str := str_replace(term, str_c('^', place_name_str), '')) %>%
-    select(-term)
+    do(tidy(lm(frmla, data=., weights=n_isolates))) %>%
+    ungroup() %>%
+    filter(str_detect(term, place_name_regex)) %>%
[…]

### 2017-09-26 ms1/ms1_display.Rmd
small cleanups

@@
 pde_approp_table = read_tsv('tables/pde_approp_table.tsv') %>%
-  filter(year %in% c(2011, 2014)) %>%
+  #filter(year %in% c(2011, 2014)) %>%
   (function(df) bind_rows(
     df,
-    df %>% group_by(year, tier) %>%
+    df %>% group_by(year, tier, refill) %>%
       summarize(n=sum(n)) %>%
@@
 pde_approp_table %>%
+  filter(year %in% c(2011, 2014)) %>%
   ggplot(aes(x=factor(year), y=f, fill=factor(tier))) +
@@
 
-If a drug declined by $X$, then the optimal change in the high tier fraction would be $1/(1-X) \approx 1+X$.
+If a drug declined by $X$, then the optimal change in the high tier fraction would be $1/(1-X) = 1+X + \mathcal{O}(X^2)$.
 For example, overall prescribing fell by about 5%, and high-tier prescribing increased by 2.5%.

### 2017-09-26 ms2/ms2_display.Rmd
small cleanups

@@
 results %>%
-  filter(unit_type=='hrr', model %in% c('univariate_fnz', 'multivariate_fnz_mean', 'multivariate_fnz_mup'), term != '(Intercept)') %>%
+  filter(unit_type=='hrr', model %in% c('univariate_fnz', 'univariate_mup', 'multivariate_fnz_mean', 'multivariate_fnz_mup'), term != '(Intercept)') %>%
   (function(df) {

### 2017-09-27 ms2/ms2_display.Rmd
ms2: great simplification

--- 2017-09-25 ms2/ms2_display.Rmd
@@
 - a Spearman correlation of mean claims per beneficiary $\mu$ against mean percent nonsusceptible $R$
-- a linear model predicting mean resistance from the fraction $f_>$ of people who are consumers
-- a linear model predicting mean resistance from $f_>$ and mean consumption among consumers $\mu_> = \mu/f_>$.
+- a linear model predicting mean resistance from the fraction $f$ of people who are consumers
+- a linear model predicting resistance from mean consumption among consumers $\nu = \mu/f$
+- if both univariate linear models are significant, a linear model predicting resistance from both $f$ and $\nu$
 
-When running regressions, I weight the data points by the number of antibiograms present in that unit.
-(The weights in a linear regression should be proportional to $1/\sigma^2$, and we generally expect that the standard error of estimators scales like $1/\sqrt{n}$.)
+When running regressions, I weight the states/HRRs by the inverse square of the standard error of the unit-level estimates.
 
@@
 The plot shows the Spearman correlation coefficient $\rho$ between resistance $R$ and mean consumption $\mu$.
-Two stars after the bug/drug combination implies that the adjusted $p < 0.05$, and one star means the unadjusted $p < 0.05$.
+Two stars after the bug/drug combination indicates that the adjusted $p < 0.05$, and one star means the unadjusted $p < 0.05$.
 Intervals show 95% confidence intervals.
@@
 
-We're interested in how it is that mean consumption drives resistance. Is it at the intensive or extensive margin? We separate that by writing $\mu = f_> \mu_>$.
+We're interested in how it is that mean consumption drives resistance. Is it at the intensive or extensive margin? We separate that by writing $\mu = f \nu$.
 
-I run models $R \sim f_>$, that is $R^{(u)} = \beta_0 + \beta_{f_>} f_>^{(u)} + \varepsilon^{(u)}$, where the superscript $u$ indicates the geographical unit (HRR or state).
-The plots show $\beta_{f_>}$.
-This model is equivalent to one in which the population resistance $R^{(u)}$ in unit $u$ is the mean of the resistance $r_i^{(u)}$ in the individual people $i$ in unit $u$, where $r_i^{(u)}$ is $\beta_0$ if person $i$ took no drug and $\beta_0 + \beta_{f_>}$ if that person took at least one dose.
+I run models $R \sim f$, that is $R^{(u)} = \beta_0 + \beta_f f^{(u)} + \varepsilon^{(u)}$, where the superscript $u$ indicates the geographical unit (HRR or state).
+The plots show $\beta_f$.
+This model is equivalent to one in which the population resistance $R^{(u)}$ in unit $u$ is the mean of the resistance $r_i^{(u)}$ in the individual people $i$ in unit $u$, where $r_i^{(u)}$ is $\beta_0$ if person $i$ took no drug and $\beta_0 + \beta_f$ if that person took at least one dose.
 
-The gray line shows $\beta_{f_>} = 1$.
+The gray line shows $\beta_f = 1$.
 Most confidence intervals include that line, i.e., we can explain our data with a transmission-free model where individuals who take drug have 100% resistance and others have some smaller amount.
@@
   interval_plot() +
+  geom_hline(yintercept=0, col='gray') +
   geom_hline(yintercept=1, col='gray') +
-  ylab(expression(beta[f['>']])) +
+  ylab(expression(beta[f])) +
   ggtitle("Linear univariate, HRR")
@@
   interval_plot() +
+  geom_hline(yintercept=0, color='gray') +
   geom_hline(yintercept=1, color='gray') +
-  ylab(expression(beta[f['>']])) +
+  ylab(expression(beta[f])) +
   ggtitle("Linear univariate, state")
 ```
-
-Why should the linear models do so much better than the Spearman?
-I think it's because, in the linear model, I can weight the geographical units that have more antibiograms.
-That is, we give more weight to the greater amounts of data.
-The Spearman model treats them all equally.
 
@@
   geom_point(shape=1) +
-  xlab('Fraction nonzero') + ylab('claims per beneficiary') +
+  xlab('fraction nonzero (f)') + ylab('mean consumption (mu)') +
   geom_hline(yintercept=0) + geom_vline(xintercept=0) +
@@
 
-Because $f_>$ is so collinear with $\mu$, multiple regressions using both as predictors will cause problems (cf.
+Because $f$ is so collinear with $\mu$, multiple regressions using both as predictors will cause problems (cf.
 the Supplement).
-Rather than mean consumption $\mu$, I'll use the mean consumption among consumers $\mu_> = \mu / f_>$ as my second predictor.
+Rather than mean consumption $\mu$, I'll use the mean consumption among consumers $\nu = \mu / f$ as my second predictor.
 
@@
   geom_point(shape=1) +
-  xlab(expression(f['>'])) +
-  ylab(expression(mu['>'])) +
-  ggtitle('Mean nonzero consumption vs. fraction consuming, states') +
+  xlab(expression(f)) +
+  ylab(expression(nu)) +
+  ggtitle('Consumption among consumers (nu) vs. fraction consuming (f), states') +
   theme_minimal() +
@@
   filter(unit_type=='hrr') %>%
-  #rename(state=unit) %>% left_join(regions, by='state') %>%
   rename(hrr=unit) %>% left_join(hrr_regions, by='hrr') %>%
@@
   geom_point(shape=1) +
-  xlab(expression(f['>'])) +
-  ylab(expression(mu['>'])) +
-  ggtitle('Mean nonzero consumption vs. fraction consuming, states') +
+  xlab(expression(f)) +
+  ylab(expression(nu)) +
+  ggtitle('Mean nonzero consumption vs. fraction consuming, HRR') +
   theme_minimal()
@@
 
-Tetracycline and Bactrim are interesting: the South has this long tail where $f_>$ is exceptionally high and $\mu_>$ is low.
+Tetracycline and Bactrim are interesting: the South has this long tail where $f$ is exceptionally high and $\nu$ is low.
 I imagined a model where the South was like the Midwest, but just with more people taking one does of, say, Bactrim.
 So it goes from $d$ doses into $c$ consumers to $d + \Delta c$ doses into $c + \Delta c$ consumers.
-This predicts that $\mu'_> = (\mu_> + \Delta f_>)/(1 + \Delta f_>)$.
+This predicts that $\nu' = (\nu + \Delta f)/(1 + \Delta f)$.
 This prediction is too high: it's not *just* that more people in some parts of the South are getting these drugs; it's also that those getting the drug are getting less of it.
@@
   geom_point() +
-  xlab('Fraction nonzero') + ylab('claims per beneficiary') +
-  ggtitle('All antibiotics, f> and mean consumption, states') +
+  xlab('Fraction nonzero (f)') + ylab('claims per beneficiary (mu)') +
+  ggtitle('All antibiotics, f and mean consumption, states') +
   theme_minimal()
@@
 
-### Effect along extensive margin $\mu_>$
+### Effect along intensive margin $\nu$
 
@@
 
-### Effect beyond $f_>$
+### Effect of $f$ beyond $\nu$, or vice versa
 
-We ask if mean consumption $\mu$ has any information beyond fraction nonzero $f_>$ for predicting resistance $R$.
-We might expect that the simple two-part model (i.e., consuming or not) represented by the regressions $R \sim f_>$ doesn't capture the fact that taking two doses should have a greater effect than taking one, etc.
+We ask if consumption among consumers $\nu$ has any information beyond fraction nonzero $f$ for predicting resistance $R$ (and vice versa).
+We do regressions $R = \beta_0 + \beta_f f + \beta_\nu \nu + \varepsilon$ and perform $F$-tests on the nested regression models.
 
-These plots show the coefficient $\beta_\nu$ from the regression $R = \beta_0 + \beta_{f_>} f_> + \beta_\nu \nu + \varepsilon$.
+```{r double_sig_hrr}
+double_sig = results %>%
+  filter(model %in% c('univariate_fnz', 'univariate_mup'),
+         term != '(Intercept)',
+         unit_type == 'hrr') %>%
+  group_by(model) %>%
+  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
+  ungroup() %>%
+  filter(p.value < 0.05) %>%
+  group_by(bug, drug_group) %>%
+  summarize(n=n()) %>%
+  filter(n == 2) %>%
+  select(-n)
 
-```{r hrr_multi}
 results %>%
-  filter(model=='multivariate_fnz_mup', unit_type=='hrr', term=='mup') %>%
-  interval_plot() +
-  ylab(expression(beta[nu])) +
-  ggtitle("Linear multivariate (f> and nu), HRR")
-```
-
-At the state level, none of the $\beta_\nu$ coefficients are significantly different from zero (after multiple-hypothesis correction).
-
-```{r state_multi}
-results %>%
-  filter(model=='multivariate_fnz_mup', unit_type=='state', term=='mup') %>%
-  interval_plot() +
-  ylab(expression(beta[nu])) +
-  ggtitle("Linear multivariate (f> and nu), state")
-```
-
-There are a few significant results:
-
-- *E. coli* and Bactrim have a *negative* relationship (higher consumption goes with *less* resistance)
-- *S. pneumoniae* and tetracyclines have a negative relationship
-- *E. faecium* and $\beta$-lactams have a positive relationship (implying higher mean consumption, given the same fraction of people consuming, goes with more resistance)
-
-Interestingly, these are the same three values that are significant when performing the (ill-advised) regression using $f_>$ and $\mu$.
-The table shows that bug/drug combinations in which the multiple regressions have significantly non-zero $\beta_{\mu_>}$.
-The values in the columns show the estimates of $\beta$ in each model.
-Note the big differences in $\beta_{f_>}$ between the two models.
-Note that although the regression $R \sim f_> + \mu$ distorts $\beta_{f_>}$ more than the regression $R \sim f_> + \mu_>$, the coefficient still gets changed quite a bit nonetheless.
-We were hoping for $\mu_>$ to give information in addition to $f_>$, not to just change it completely.
-
-```{r fnz_mu_results}
-results %>%
-  filter(unit_type=='hrr', model %in% c('univariate_fnz', 'multivariate_fnz_mean', 'multivariate_fnz_mup'), term != '(Intercept)') %>%
-  (function(df) {
-    filter(df, term != 'fnz', p.adjust(p.value, method='BH') < 0.05) %>%
-      select(bug, drug_group) %>%
-      distinct() %>%
-      inner_join(df, by=c('bug', 'drug_group')) }) %>%
-  mutate(model=str_replace(model, '^multivariate_', '')) %>%
-  select(bug, drug_group, model, term, estimate) %>%
-  spread(term, estimate) %>%
+  semi_join(double_sig, by=c('bug', 'drug_group')) %>%
+  filter(unit_type == 'hrr',
+         model %in% c('multivariate_fnz_mup', 'multivariate_mup_fnz'),
+         term != '(Intercept)') %>%
+  mutate(second_anova_term=str_extract(model, '(mup|fnz)$')) %>%
+  filter(term == second_anova_term) %>%
+  select(bug, drug_group, term, anova.p.value) %>%
+  spread(term, anova.p.value) %>%
   kable
@@
 
-I contrast this situation against *E. coli* and quinolones. In that case, adding terms to the regression doesn't drastically alter $\beta_{f_>}$.
+At the HRR level, we see that:
 
-```{r fnz_mu_ecoliquin}
+- For CoNS/FQ and *E. cloacae*/FQ, neither $f$ nor $\nu$ gives information beyond the other
+- For *E. coli*/FQ and *S. pneumonia*/macrolides, $f$ gives informatio beyond $\nu$
+- For *P. mirabilis*/quinolones, $\nu$ gives information beyond $f$
+- For *E. coli*/Bactrim, $f$ and $\nu$ each give information beyond the other
+
+And at the state level:
+
+```{r double_sig_state}
+double_sig = results %>%
+  filter(model %in% c('univariate_fnz', 'univariate_mup'),
+         term != '(Intercept)',
+         unit_type == 'state') %>%
+  group_by(model) %>%
+  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
+  ungroup() %>%
+  filter(p.value < 0.05) %>%
+  group_by(bug, drug_group) %>%
+  summarize(n=n()) %>%
+  filter(n == 2) %>%
+  select(-n)
+
 results %>%
-  filter(unit_type=='hrr', model %in% c('univariate_fnz', 'multivariate_fnz_mean', 'multivariate_fnz_mup'), term != '(Intercept)') %>%
-  filter(bug=='E. coli', drug_group=='quinolone') %>%
-  mutate(model=str_replace(model, '^multivariate_', '')) %>%
-  select(bug, drug_group, model, term, estimate) %>%
-  spread(term, estimate) %>%
+  semi_join(double_sig, by=c('bug', 'drug_group')) %>%
+  filter(unit_type == 'state',
+         model %in% c('multivariate_fnz_mup', 'multivariate_mup_fnz'),
+         term != '(Intercept)') %>%
+  mutate(second_anova_term=str_extract(model, '(mup|fnz)$')) %>%
+  filter(term == second_anova_term) %>%
+  select(bug, drug_group, term, anova.p.value) %>%
+  spread(term, anova.p.value) %>%
   kable
@@
 
-For *E.
-faecium* with $\beta$-lactams, the univariate regression with $f_>$ gives a slightly negative slope but is not significant.
-The univariate regression with $\mu$ gives a slightly positive slope and is not significant.
-In the multiple regression, the signs stay the same but the coefficients are very different.
-Thus, we can't trust that $\mu$ provides data in addition to $f_>$ when it's not clear that either provides the information on its own.
-
-For *S.
-pneumoniae* and tetracyclines, the regression with $f_>$ and the regression with $\mu$ give negative slopes, but neither are significant.
-In the multiple regression, the $f_>$ coefficient becomes positive.
-Again, I'm not sure I can trust this.
-
-For *E.
-coli* and Bactrim, the univariate regression with $f_>$ is positive and signficant.
-The regression with $\mu$ is positive and not significant.
-Then, in the multiple regression, $\mu$ flips signs and becomes highly significant (different from zero), and it's significant in an ANOVA where $f_>$ comes first, but it's not significant in an ANOVA where it ($\mu$) comes first.
+There are only two bug/drug combinations that are double-significant, and in those cases, neither metric gives information beyond the other.
 
@@
 
-From the Spearman and univariate models, Bactrim resistance in *E.
-coli* is correlated with increased resistance (black line).
-When breaking that apart by region, however, there are different relationship (colored lines).
-This makes it seem like there is something strange going on.
-
-It's the L-shape!
+Increased fraction consuming $f$ is correlated with increased resistance:
 
@@
   ggplot(aes(x=fnz, y=percent_nonsusceptible, color=region)) +
-  geom_abline(intercept=16.86, slope=114.8, color='pink') +
+  #geom_abline(intercept=16.86, slope=114.8, color='pink') +
   geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
@@
   ylab('Bactrim-resistance E. coli') +
-  ggtitle('E. coli & Bactrim across HRRs')
+  ggtitle('E. coli & Bactrim, f, HRRs')
 ```
 
-These internal trends are a lot weaker when looking at the state level:
-
-```{r ecoli_bactrim_fnz_state}
-ineq %>%
-  filter(unit_type=='state') %>%
-  rename(state=unit) %>%
-  left_join(regions, by='state') %>%
-  right_join(state_abg, by=c('state', 'drug_group')) %>%
-  filter(bug=='E. coli', drug_group=='tmp_smx') %>%
-  ggplot(aes(x=fnz, y=percent_nonsusceptible, color=region)) +
-  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
-  geom_smooth(method='lm', se=F) +
-  geom_point() +
-  xlab('fraction who consumed Bactrim') +
-  ylab('Bactrim resistance E. coli') +
-  ggtitle('E. coli & Bactrim across states')
-```
-
-But the same patterns are visible when looking at $\mu_>$ across HRRs:
+But increased consumption among consumers $\nu$ is anticorrelated resistance, both within and across regions:
 
@@
   filter(percent_nonsusceptible < 50) %>% # filter one crazy Nevada point
-  ggplot(aes(x=mean, y=percent_nonsusceptible, color=region)) +
+  ggplot(aes(x=mup, y=percent_nonsusceptible, color=region)) +
   geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
@@
   ylab('Bactrim-resistance E. coli') +
-  ggtitle('E. coli & Bactrim, mu>, HRRs')
+  ggtitle('E. coli & Bactrim, nu, HRRs')
 ```
@@
   geom_point() +
-  xlab('fraction who consumed tetra') +
+  xlab('f') +
   ylab('tetra resistance pneumo') +
-  ggtitle('pneumo & tetra across HRRs')
+  ggtitle('pneumo & tetra, f, HRRs')
 
 ineq %>%
-  filter(unit_type=='state') %>%
-  rename(state=unit) %>%
-  left_join(regions, by='state') %>%
-  right_join(state_abg, by=c('state', 'drug_group')) %>%
+  filter(unit_type=='hrr') %>%
+  rename(hrr=unit) %>% left_join(hrr_regions, by='hrr') %>%
+  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
   filter(bug=='S. pneumoniae', drug_group=='tetracycline') %>%
-  ggplot(aes(x=fnz, y=percent_nonsusceptible, color=region)) +
+  ggplot(aes(x=mup, y=percent_nonsusceptible, color=region)) +
   geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
@@
   geom_point() +
-  xlab('fraction who consumed tetra') +
-  ylab('tetra resistance pneumo') +
-  ggtitle('pneumo & tetra across states')
+  xlab('nu') +
+  ylab('R') +
+  ggtitle('pneumo & tetra, nu, HRRs')
 ```
@@
 ### E. coli and beta-lactams
-
-This negative relationship was significant only for HRRs, not states.
-Not surprisingly, there is something weird going on with the HRRs: the Southern HRRs are all above the $R \sim f_>$ line, suggesting that the relationship with the mean is just a predictor about region.
 
@@
   geom_point() +
-  xlab('fraction who consumed beta-lactams') +
+  xlab('f') +
   ylab('Beta-lactam resistance in E. coli') +
-  ggtitle('E. coli & beta-lactams across HRRs')
+  ggtitle('E. coli & beta-lactams, f, HRRs')
 
 ineq %>%
-  filter(unit_type=='state') %>%
-  rename(state=unit) %>%
-  left_join(regions, by='state') %>%
-  right_join(state_abg, by=c('state', 'drug_group')) %>%
+  filter(unit_type=='hrr') %>%
+  rename(hrr=unit) %>% left_join(hrr_regions, by='hrr') %>%
+  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
   filter(bug=='E. coli', drug_group=='beta_lactam') %>%
-  filter(percent_nonsusceptible < 50) %>% # filter one crazy point
-  ggplot(aes(x=mean, y=percent_nonsusceptible, color=region)) +
+  filter(percent_nonsusceptible < 50) %>% # filter one crazy Nevada point
+  ggplot(aes(x=mup, y=percent_nonsusceptible, color=region)) +
   geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
@@
   geom_point() +
-  xlab('fraction who consumed beta-lactams') +
-  ylab('Beta-lactam resistance E. coli') +
-  ggtitle('E. coli & beta-lactams across states')
+  xlab('nu') +
+  ylab('Beta-lactam resistance in E. coli') +
+  ggtitle('E. coli & beta-lactams, nu, HRRs')
 ```
-
-## Comparing the predictive power of mean consumption and fraction nonzero
-
-How do the regressions $R \sim f_>$ compare to $R \sim \mu$?
-Here I look at the correlation between the coefficients $\beta_\mu$ and $\beta_{f_>}$ across bug/drug combinations.
-Not surprisingly, $\beta_\mu$ and $\beta_{f_>}$ are highly correlated: because $\mu$ and $f_>$ are correlated, you can guess the result of the one regression from the other.
-However, $\beta_{\mu_>}$ is not as well correlated with $\beta_{f_>}$, indicating that there might be some extra information there.
-
-The column titles are `metric1_metric2`, and the value shows the correlation between the univariate models' coefficients for predicting resistance to that drug across bugs.
-
-```{r univariate_result_correlation, eval=FALSE}
-results %>%
-  filter(str_detect(model, '^uni'), term != '(Intercept)') %>%
-  select(unit_type, bug, drug_group, term, estimate) %>%
-  spread(term, estimate) %>%
-  group_by(unit_type, drug_group) %>%
-  filter(n() > 2) %>%
-  do(tidy(cor(select(., -unit_type, -bug, -drug_group)))) %>%
-  ungroup() %>%
-  rename(x=.rownames) %>%
-  gather('y', 'value', fnz:nb_mu) %>%
-  filter(x > y) %>%
-  mutate(key=str_c(x, '_', y)) %>%
-  select(-x, -y) %>%
-  spread(key, value) %>%
-  kable
-```
-
-In other words, if you do the regression with $f_>$, you can predict with very high accuracy the coefficient from the univariate regression on $\mu$.
 
@@
 
-- *E. cloacae* and quinolone
-- *S. aureus* and clindamycin
-
 # Supplementary results
-
-## Collinearity
-
-This shows why it's a bad idea to do a multiple regression with $f_>$ and $\mu$. It confirms that we're mostly getting bad results due to multicollinearity.
-
-The gray lines link the coefficients from a univariate model and a multivariate model (i.e., with $f_>$ and $\mu$) for a certain bug/drug combination.
-
-From a previous plot, we know that $\mu \approx \alpha f_>$, where $\alpha \approx 3/2$. Then if $R \approx \beta_0 + \beta_{f_>}$, then we should guess that $\beta_\mu = \beta_{f_>} / \alpha$.
-
-Furthermore, in doing a multivariate regression, using the same math, we find that we expect
-$$
-\beta_\mu^{(2)} = \frac{1}{\alpha} \left( \beta_{f_>}^{(1)} - \beta_{f_>}^{(2)} \right),
-$$
-where the superscripts indicate that the coefficients are from the univariate ("1") and multivariate ("2") regressions.
-Thus, for $\beta_{f_>}^{(1)} \approx 0$, we expect the multivariate regression coefficients be related by a slope of $-1/\alpha$.
-
-On this plot, the pink lines show $\pm 1/\alpha$, which lines up very nicely with the data (especially considering that I picked $\alpha=3/2$ just by eyeballing that other plot).
-
-```{r uni_multi}
-results %>%
-  filter(unit_type=='hrr', term != '(Intercept)',
-         model %in% c('univariate_fnz', 'univariate_mean', 'multivariate_fnz_mean')) %>%
-  mutate(model=str_extract(model, '(uni|multi)'),
-         bug_drug=interaction(bug, drug_group)) %>%
-  select(bug_drug, model, term, estimate) %>%
-  spread(term, estimate) %>%
-  ggplot(aes(x=fnz, y=mean, color=model, group=bug_drug)) +
-  geom_abline(slope=2/3, color='pink') +
-  geom_abline(slope=-2/3, intercept=0, color='pink') +
-  geom_line(color='gray') +
-  geom_point() +
-  xlab(expression(beta[f['>']])) +
-  ylab(expression(beta[mu])) +
-  ggtitle('Relationship between model coefficients (f>, mean), HRR')
-```
-
-But the result is not so pronounced for $f_>$ and $\mu_>$:
-
-```{r uni_multi_mup}
-results %>%
-  filter(unit_type=='hrr', term != '(Intercept)',
-         model %in% c('univariate_fnz', 'univariate_mup', 'multivariate_fnz_mup')) %>%
-  mutate(model=str_extract(model, '(uni|multi)'),
-        bug_drug=str_c(bug, drug_group)) %>%
-  select(bug_drug, model, term, estimate) %>%
-  spread(term, estimate) %>%
-  ggplot(aes(x=fnz, y=mup, color=model, group=bug_drug, label=bug_drug)) +
-  #geom_text(size=2) +
-  geom_line(color='gray') +
-  geom_point() +
-  xlab(expression(beta[f['>']])) +
-  ylab(expression(beta[mu['>']])) +
-  ggtitle('Relationship between model coefficients (f>, mu>) , HRR')
-```
-
-## Using logistic models
-
-We don't expect that, over a large range, the consumption/resistance relationship will be linear.
-We expect it would be something more like a logistic function.
-To evaluate the difference between these two approaches, I performed quasibinomial regressions $R \sim f_>$.
-If $\bar{f}$ is the midpoint of the $f_>$ values (i.e., the weighted mean) and $\hat{y}$ is the predicted resistance from the model, then the slope at the midpoint is $\beta_f \hat{y}(\bar{x}) \left[ 1- \hat{y}(\bar{x}) \right]$.
-I compare the $p$-values and slopes in the linear and quasibinomial regressions:
-
-```{r linear_quasibin}
-results %>%
-  filter(unit_type=='hrr', model %in% c('univariate_fnz', 'univariate_fnz_log'), term=='fnz') %>%
-  select(bug, drug_group, model, estimate, p.value) %>%
-  group_by(model) %>%
-  mutate(p.value = p.adjust(p.value, method='BH')) %>%
-  ungroup() %>%
-  ggplot(aes(x=estimate, y=-log10(p.value), color=model, group=interaction(bug, drug_group))) +
-  geom_hline(yintercept = -log10(0.05)) +
-  geom_point() +
-  geom_line(color='gray')
-```
-
-Points above the black line are significant.
-Note that the $p$-values in the two cases are really similar (i.e., most dumbbells are on the same horizontal line) but they can differ somewhat in their slope.
-The right-most pink dot is pneumococcus/macrolides, and the next is *P.
-mirabilis*/Bactrim.
-
-I also tried doing beta regressions, which get similar-sized changes in the slope but really significant $p$-values. Not sure what to make of those.
-
-```{r linear_beta}
-results %>%
-  filter(unit_type=='hrr', model %in% c('univariate_fnz', 'univariate_fnz_beta'), term=='fnz') %>%
-  select(bug, drug_group, model, estimate, p.value) %>%
-  group_by(model) %>%
-  mutate(p.value = p.adjust(p.value, method='BH')) %>%
-  ungroup() %>%
-  ggplot(aes(x=estimate, y=-log10(p.value), color=model, group=interaction(bug, drug_group))) +
-  geom_hline(yintercept = -log10(0.05)) +
-  geom_point() +
-  geom_line(color='gray')
-```
 

### 2017-09-28 ms2/ms2_display.Rmd
ms2: added approp margin

@@
 
+# Appropriateness at the margins
+
+Stewardship should also be motivated by whether consumption is more appropriate along the extensive or intensive margin.
+
+For each drug, take all claims made by beneficiaries who had $N$ claims for that drug.
+Of those $N$ claims, what fraction were associated with appropriate diagnoses?
+I exclude re-fills, since we don't expect them to be associated with diagnoses.
+
+I also do logistic regressions predicting inappropriate prescribing from the number of claims.
+(In other words, we predict whether a claim is appropriate using the number of claims for that drug its beneficiary had.) I used all the data, rather than truncating at 9 claims like for the figure.
+Where inappropriate prescribing increases (significantly, after multiple hypothesis correction), I put "+" after the name; where it decreases, "-".
+
+```{r approp_margin}
+top_abx = c('azithromycin', 'ciprofloxacin', 'amoxicillin',
+            'cephalexin', 'trimethoprim/sulfamethoxazole',
+            'levofloxacin', 'amoxicillin/clavulanate',
+            'doxycycline', 'nitrofurantoin', 'clindamycin')
+
+approp = read_tsv('../ms1/tables/pde_approp_margin.tsv') %>%
+  filter(year==2014, antibiotic %in% top_abx) %>%
+  select(-year, -f_app) %>%
+  complete(antibiotic, n_abx_claims, app, fill=list(n_abx_app_claims=0)) %>%
+  spread(app, n_abx_app_claims) %>%
+  rename(n_app=`TRUE`, n_inapp=`FALSE`) %>%
+  mutate(n_total=n_app+n_inapp)
+
+approp_trends = approp %>%
+  mutate(y=n_inapp / n_total) %>%
+  group_by(antibiotic) %>%
+  do(tidy(glm(y ~ n_abx_claims, data=., weights=n_total, family='quasibinomial'))) %>%
+  filter(term=='n_abx_claims') %>%
+  mutate(p.adj=p.adjust(p.value, method='BH')) %>%
+  filter(p.adj < 0.05) %>%
+  select(antibiotic, estimate, std.error, p.adj)
+
+top_abx_short = c('azithromycin', 'ciprofloxacin', 'amoxicillin',
+                  'cephalexin', 'TMP/SMX', 'levofloxacin',
+                  'amox/clav', 'doxycycline', 'nitrofurantoin',
+                  'clindamycin')
+
+approp_abx_labels = data_frame(antibiotic=top_abx,
+                               short_antibiotic=top_abx_short) %>%
+  left_join(approp_trends, by='antibiotic') %>%
+  mutate(sig=case_when(is.na(.$estimate) ~ '',
+                       estimate > 0 ~ ' (+)',
+                       estimate < 0 ~ ' (-)')) %>%
+  mutate(short_antibiotic=str_c(short_antibiotic, sig)) %>%
+  pull(short_antibiotic)
+
+approp %>%
+  filter(n_total > 0) %>%
+  group_by(antibiotic, n_abx_claims) %>%
+  do((function (x, n) {
+    m = binom.test(x, n)
+    data_frame(estimate=m$estimate,
+               ci_low=m$conf.int[1],
+               ci_high=m$conf.int[2])
+  })(.$n_inapp, .$n_total)) %>%
+  ungroup() %>%
+  # throw out the messy end margins
+  filter(n_abx_claims <= 9) %>%
+  # order the antibiotics
+  mutate(antibiotic=factor(antibiotic,
+                           levels=top_abx,
+                           labels=approp_abx_labels)) %>%
+  ggplot(aes(x=n_abx_claims, y=estimate, ymin=ci_low, ymax=ci_high)) +
+  geom_point() +
+  geom_line() +
+  geom_errorbar() +
+  facet_wrap(~antibiotic) +
+  ylab('fraction inappropriate') +
+  scale_x_continuous(breaks=c(1, 3, 5, 7, 9))
+```
+
+This table has the data from the tests: the estimate is the odds ratio for each additional claim.
+Positive estimates mean that inappropriate increases with more intensive usage; negative means that inappropriate prescribing decreases.
+I only show the drugs that has significant results after multiple hypothesis correction.
+
+```{r approp_trends}
+approp_trends %>% kable
+```
+
+Thus, amoxicillin is the only drug that is more inappropriate among high intensity users.
+All the other drugs are either the same across intensiveness or actually more appropriate among intensive users.
+This is another argument for, in general, focusing on the extensive margin.
+
 # Supplementary results
@@
 ```
+
+## Comparing state and HRR univariate $f$ models
+
+For univariate models on $f$, HRRs have better $p$-values but similar estimates.
+
+```{r compare_uni_f}
+results %>%
+  filter(model=='univariate_fnz', term=='fnz') %>%
+  mutate(group=interaction(bug, drug_group),
+         adj.p=p.adjust(p.value, method='BH')) %>%
+  ggplot(aes(x=estimate, y=-log10(adj.p), color=unit_type, group=group)) +
+  geom_line(color='gray') +
+  geom_hline(yintercept=-log10(0.05)) +
+  geom_point() +
+  xlab(expression(beta[f]))
+```
+
+## Comparing state and HRR univariate $\nu$ models
+
+For univariate models on $\nu$, the two negative-estimate results follow the same pattern: HRRs have better $p$-values but similar estimates.
+*Proteus* is way up in the corner.
+For the other positive-estimate results, it's a little more muddy.
+
+```{r compare_uni_nu}
+results %>%
+  filter(model=='univariate_mup', term=='mup') %>%
+  mutate(group=interaction(bug, drug_group),
+         adj.p=p.adjust(p.value, method='BH')) %>%
+  ggplot(aes(x=estimate, y=-log10(adj.p), color=unit_type, group=group)) +
+  geom_line(color='gray') +
+  geom_hline(yintercept=-log10(0.05)) +
+  geom_point() +
+  xlab(expression(beta[nu]))
+```

### 2017-09-28 ms1/ms1_display.Rmd
ms1: rearrange approp

@@
 
+```{r approp_changes}
+# read in the raw and trend data
+pde_approp_table = read_tsv('tables/pde_approp_table.tsv') %>%
+  (function(df) bind_rows(
+    df,
+    df %>% group_by(year, tier) %>%
+      summarize(n=sum(n)) %>%
+      mutate(antibiotic='overall')
+  )) %>%
+  filter(antibiotic %in% c(top_abx, 'overall')) %>%
+  group_by(year, antibiotic) %>%
+  mutate(f=n/sum(n)) %>%
+  ungroup()
+
+pde_approp_risk_2011 = pde_approp_table %>%
+  # add up tiers 1 and 2
+  filter(year==2011, tier <= 2) %>%
+  group_by(antibiotic) %>%
+  summarize(risk_2011=sum(f))
+
+risk_ratio_f = function(odr, p) odr / (1 + p * (odr - 1))
+
+# compute risk ratios
+pde_approp_trends = read_tsv('tables/pde_approp_trends.tsv') %>%
+  filter(term=='year') %>%
+  mutate(estimate=exp(3*estimate), # odds ratio over the whole time period
+         std.error=sqrt(3)*std.error,
+         adj.p=p.adjust(p.value, method='BH'),
+         ci=1.96*std.error,
+         ci_low=estimate-ci,
+         ci_high=estimate+ci) %>%
+  left_join(pde_approp_risk_2011, by='antibiotic') %>%
+  # convert OR to RR (i.e., adjusted change)
+  mutate_at(vars(estimate, ci_low, ci_high), function(x) risk_ratio_f(x, .$risk_2011 - 1)) %>%
+  # convert to percents, and reverse direction (to make it *inapprop*)
+  mutate_at(vars(estimate, ci_low, ci_high), function(x) -(x - 1)*100) %>%
+  select(antibiotic, pct_change=estimate, ci_low, ci_high, adj.p) # now estimate is adjusted change
+```
+
 ```{r adjusted_changes}
@@
 
+ggplot(adjusted_changes, aes(x=antibiotic, y=pct_change, ymin=ci_low, ymax=ci_high)) +
+  geom_hline(yintercept=0) +
+  geom_bar(dat=adjusted_changes, stat='identity', fill=NA, color='black') +
+  geom_errorbar(dat=adjusted_changes, width=0.5) +
+  geom_point(dat=unadjusted_changes, aes(ymin=NULL, ymax=NULL), shape=1) +
+  geom_point(dat=pde_approp_trends, color='blue', shape=1) +
+  geom_errorbar(dat=pde_approp_trends, width=0.25, color='blue') +
+  coord_flip() +
+  xlab('') +
+  ylab('adjusted change from 2011 to 2014 (%)') +
+  scale_y_continuous(limits=c(-22, 32), breaks=c(-20, -10, 0, 10, 20, 30)) +
+  theme_classic() +
+  theme(axis.title.x = element_text(size=16),
+        axis.text = element_text(size=16))
+
 adjusted_changes %>%
@@
 ```
+
+The bars show the adjusted change in consumption. Black circles show unadjusted change in consumption. Blue shows the adjusted change in inappropriate prescribing.
+
+### Adjusted changes in usage
 
@@
 Consumption of azithromycin fell by `r adjusted_az_decrease`% over the entire study period. Levofloxacin consumption increased by `r adjusted_levo_increase`% over the entire study period.
+
+### Adjusted changes in inappropriate prescribing
+
+```{r pde_approp_trends_table}
+pde_approp_trends %>% kable
+```
 
@@
 ```{r approp_graph}
-pde_approp_table = read_tsv('tables/pde_approp_table.tsv') %>%
+pde_approp_table %>%
   #filter(year %in% c(2011, 2014)) %>%
-  (function(df) bind_rows(
-    df,
-    df %>% group_by(year, tier, refill) %>%
-      summarize(n=sum(n)) %>%
-      mutate(antibiotic='overall')
-  )) %>%
-  filter(antibiotic %in% c(top_abx, 'overall')) %>%
-  group_by(year, antibiotic) %>%
-  mutate(f=n/sum(n)) %>%
-  ungroup()
-
-pde_approp_table %>%
-  filter(year %in% c(2011, 2014)) %>%
   ggplot(aes(x=factor(year), y=f, fill=factor(tier))) +
@@
 ```{r pre_approp_table}
-pde_approp_trends = read_tsv('tables/pde_approp_trends.tsv') %>%
-  filter(term=='year') %>%
-  mutate(or=exp(3*estimate), adj.p=p.adjust(p.value, method='BH')) %>%
-  select(antibiotic, or, adj.p)
 
-risk_ratio_f = function(odr, p) odr / (1 + p * (odr - 1))
-
-pde_approp_table %>%
-  filter(tier <= 2) %>%
-  select(antibiotic, year, f) %>%
-  group_by(antibiotic, year) %>%
-  summarize(f=sum(f)) %>%
-  ungroup() %>%
-  spread(year, f) %>%
-  left_join(pde_approp_trends, by='antibiotic') %>%
-  mutate(raw_pct_change=(`2014`-`2011`)/`2011`,
-         adj_pct_change=risk_ratio_f(or, `2011`) - 1) %>%
-  mutate_at(vars(`2011`, `2014`, raw_pct_change, adj_pct_change),
-            function(x) round(x*100, 2)) %>%
-  select(antibiotic, `2011`, `2014`, raw_pct_change, adj_pct_change, adj.p) %>%
-  arrange(raw_pct_change) %>%
-  kable
 ```

### 2017-09-28 ms1/ms1_display.Rmd
ms1: first fills

@@
 
+### All fills
+
+Bars show adjusted trends in usage, including refills. The dots shows unadjusted trends.
+
+```{r adjusted_changes}
+adjusted_changes = read_tsv('tables/model_abx.tsv') %>%
+  rename(antibiotic=name) %>%
+  filter(term=='year') %>%
+  mutate(estimate=3*estimate,
+         std.error=sqrt(3)*std.error) %>%
+  model_ci() %>%
+  select(antibiotic, pct_change, ci_low, ci_high) %>%
+  arrange(pct_change) %>%
+  mutate(antibiotic=fct_inorder(antibiotic),
+         antibiotic=fct_relevel(antibiotic, 'overall', after=Inf))
+
+unadjusted_changes = bind_rows(claims_by_abx, claims_overall) %>%
+  filter(antibiotic %in% c('overall', top_abx), year %in% c(2011, 2014)) %>%
+  select(antibiotic, year, cpkp) %>%
+  spread(year, cpkp) %>%
+  mutate(change=`2014`-`2011`,
+         fold_change=change/`2011`,
+         pct_change=100*fold_change) %>%
+  select(antibiotic, pct_change)
+
+adjusted_changes %>%
+  ggplot(aes(x=antibiotic, y=pct_change, ymin=ci_low, ymax=ci_high)) +
+  geom_bar(stat='identity', fill=NA, color='black') +
+  geom_errorbar(width=0.5) +
+  geom_point(dat=unadjusted_changes, aes(x=antibiotic, y=pct_change, ymin=NULL, ymax=NULL), shape=1) +
+  coord_flip() +
+  xlab('') +
+  ylab('adjusted change in consumption 2011 to 2014 (%)') +
+  scale_y_continuous(limits=c(-22, 32), breaks=c(-20, -10, 0, 10, 20, 30)) +
+  theme_minimal() +
+  theme(axis.title.x = element_text(size=16),
+        axis.text = element_text(size=16))
+```
+
+The bars show the adjusted change in consumption. Black circles show unadjusted change in consumption. Blue shows the adjusted change in inappropriate prescribing.
+
+### First fills
+
+Bars show the adjusted change in prescribing, considering only first fills. The blue dots show the change in inappropriate prescribing (which only includes first fills).
+
 ```{r approp_changes}
@@
   mutate_at(vars(estimate, ci_low, ci_high), function(x) -(x - 1)*100) %>%
-  select(antibiotic, pct_change=estimate, ci_low, ci_high, adj.p) # now estimate is adjusted change
-```
+  select(antibiotic, pct_change=estimate, ci_low, ci_high, adj.p)
+  # now estimate is adjusted change in INappropriate prescribing
 
-```{r adjusted_changes}
-adjusted_changes = read_tsv('tables/model_abx.tsv') %>%
+adjusted_changes_firstfill = read_tsv('tables/model_abx_firstfill.tsv') %>%
   rename(antibiotic=name) %>%
@@
 
-unadjusted_changes = bind_rows(claims_by_abx, claims_overall) %>%
-  filter(antibiotic %in% c('overall', top_abx), year %in% c(2011, 2014)) %>%
-  select(antibiotic, year, cpkp) %>%
-  spread(year, cpkp) %>%
-  mutate(change=`2014`-`2011`,
-         fold_change=change/`2011`,
-         pct_change=100*fold_change) %>%
-  select(antibiotic, pct_change)
-
-ggplot(adjusted_changes, aes(x=antibiotic, y=pct_change, ymin=ci_low, ymax=ci_high)) +
+ggplot(adjusted_changes_firstfill, aes(x=antibiotic, y=pct_change, ymin=ci_low, ymax=ci_high)) +
   geom_hline(yintercept=0) +
-  geom_bar(dat=adjusted_changes, stat='identity', fill=NA, color='black') +
-  geom_errorbar(dat=adjusted_changes, width=0.5) +
-  geom_point(dat=unadjusted_changes, aes(ymin=NULL, ymax=NULL), shape=1) +
+  geom_bar(stat='identity', fill=NA, color='black') +
+  geom_errorbar(width=0.5) +
   geom_point(dat=pde_approp_trends, color='blue', shape=1) +
@@
   ylab('adjusted change from 2011 to 2014 (%)') +
-  scale_y_continuous(limits=c(-22, 32), breaks=c(-20, -10, 0, 10, 20, 30)) +
+  scale_y_continuous(limits=c(-22, 35), breaks=c(-20, -10, 0, 10, 20, 30)) +
   theme_classic() +
-  theme(axis.title.x = element_text(size=16),
-        axis.text = element_text(size=16))
-
-adjusted_changes %>%
-  ggplot(aes(x=antibiotic, y=pct_change, ymin=ci_low, ymax=ci_high)) +
-  geom_bar(stat='identity', fill=NA, color='black') +
-  geom_errorbar(width=0.5) +
-  geom_point(dat=unadjusted_changes, aes(x=antibiotic, y=pct_change, ymin=NULL, ymax=NULL), shape=1) +
-  coord_flip() +
-  xlab('') +
-  ylab('adjusted change in consumption 2011 to 2014 (%)') +
-  scale_y_continuous(limits=c(-22, 32), breaks=c(-20, -10, 0, 10, 20, 30)) +
-  theme_minimal() +
   theme(axis.title.x = element_text(size=16),
@@
 
-The bars show the adjusted change in consumption. Black circles show unadjusted change in consumption. Blue shows the adjusted change in inappropriate prescribing.
-
-### Adjusted changes in usage
+### Adjusted changes in usage table
 

### 2017-09-28 ms1/ms1_1clean.R

@@ -10,18 +10,22 @@ sum_1yr_cc = paste0(names_1yr_ccs, collapse=' + ')
 regions = read_tsv('../../db/census-regions/census-regions.tsv') %>%
   select(state, region)
 
-# window of days that you can look ahead from an rx to a dx
-rxdx_window = 3
-
 load_data = function(year) {
   # pde data
   pde = read_tsv(sprintf('../pde_%i.tsv', year)) %>%
-    select(bene_id, pde_date, antibiotic) %>%
-    distinct() %>%
+    select(bene_id, pde_date, antibiotic, fill_num) %>%
+    # for each bene, date, and drug, keep only one record. give preference to first fills.
+    group_by(bene_id, pde_date, antibiotic) %>%
+    summarize(fill_num=min(fill_num)) %>%
+    ungroup() %>%
     mutate(year=year, pde_id=1:n())
 
   # total numbers of PDEs
   n_claims = count(pde, bene_id) %>% rename(n_claims=n)
+  n_claims_firstfill = pde %>%
+    filter(fill_num==0) %>%
+    count(bene_id) %>%
+    rename(n_claims_firstfill=n)
 
   # count chronic conditions. 'n_cc' means one-year cc's
   cc = read_feather(sprintf('../cc_%i.feather', year)) %>%
@@ -35,7 +39,8 @@ load_data = function(year) {
     left_join(regions, by='state') %>%
     left_join(cc, by='bene_id') %>%
     left_join(n_claims, by='bene_id') %>%
-    replace_na(list(n_claims=0))
+    left_join(n_claims_firstfill, by='bene_id') %>%
+    replace_na(list(n_claims=0, n_claims_firstfill=0))
 
   # take the bene, pde (with diagnoses) and diagnoses (separately)
   list(bene=bene, pde=pde)

### 2017-09-28 ms1/ms1_2process.R

@@ -44,6 +44,8 @@ n_unique_bene = bene$bene_id %>%
 pde = read_tsv('data/pde.tsv') %>%
   left_join(bene, by=c('year', 'bene_id'))
 
+pde_firstfill = pde %>% filter(fill_num==0)
+
 # the top few inidividual antibiotics
 claims_by_abx = pde %>%
   count(year, antibiotic) %>% ungroup() %>%
@@ -91,6 +93,10 @@ overall_model = bene %>%
   rename(y=n_claims) %>%
   model_f(abx_frmla, 'overall')
 
+overall_model_firstfill = bene %>%
+  rename(y=n_claims_firstfill) %>%
+  model_f(abx_frmla, 'overall')
+
 # model overall consumption, but look at individual populations
 covariates = c('age', 'sex', 'race', 'dual', 'region')
 reduced_model = function(df, missing_term, name) {
@@ -120,7 +126,7 @@ bind_rows(
   output_table('model_overall_demography')
 
 # get models for individual abx
-single_f = function(abx, frmla) {
+single_f = function(pde, abx, frmla) {
   pde %>%
     filter(antibiotic==abx) %>%
     count(year, bene_id) %>% ungroup() %>% rename(y=n) %>%
@@ -129,11 +135,15 @@ single_f = function(abx, frmla) {
     model_f(frmla, abx)
 }
 
-single_models = lapply(top_abx, function(a) single_f(a, abx_frmla))
+single_models = lapply(top_abx, function(a) single_f(pde, a, abx_frmla))
+single_models_firstfill = lapply(top_abx, function(a) single_f(pde_firstfill, a, abx_frmla))
 
 # report the model parameters
 bind_rows(overall_model, single_models) %>%
   output_table('model_abx')
+
+bind_rows(overall_model_firstfill, single_models_firstfill) %>%
+  output_table('model_abx_firstfill')
 
 # trends in all fluoroquinolones
 model_fq = pde %>%

### 2017-10-06 ms1/ms1_display.Rmd
lots of updates

@@
 ## Claims by abx
+
+The top 10 antibiotics account for most of the claims:
+
+```{r top_10_fraction}
+claims_by_abx %>%
+  group_by(antibiotic) %>%
+  summarize(n=sum(n)) %>%
+  arrange(desc(n)) %>%
+  mutate(f=n/sum(n), cumf=cumsum(f)) %>%
+  head(10) %>%
+  kable
+```
 
@@
 
+## Numbers of diagnoses
+
+The fraction of diagnoses that have E&M HCPCS codes:
+
+```{r hcpcs_count}
+read_tsv('tables/hcpcs_count.tsv') %>%
+  group_by(hcpcs_flag) %>%
+  summarize(n_dx=sum(n_dx)) %>%
+  mutate(f=n_dx/sum(n_dx)) %>%
+  kable
+```
+
+The fraction of first-fill PDEs that have upstream:
+
+```{r pde_with_dx_upstream}
+read_tsv('tables/pde_with_dx_upstream.tsv') %>%
+  group_by(has_dx, has_em_dx) %>%
+  summarize(n=sum(n)) %>%
+  ungroup() %>%
+  mutate(f=n/sum(n), cumf=cumsum(f)) %>%
+  kable
+```
+
 ## Trends in appropriateness
@@
 
-```{r pre_approp_table}
-
-```
-
 And, to be more careful, I show the answer to a similar question but using a regression to correct for covariates:
@@
 ```{r pde_approp_trends}
-
-```
-
-## Trends in diagnoses
-
-Are some diagnoses becoming more or less common?
-
-First, unadjusted trends in the number of diagnoses per 1,000 beneficiaries in each year. The percent change is between 2011 and 2014.
-
-Second, adjusted trends, using Poisson regressions.
-The unit of analysis is the beneficiary in a year.
-The outcome variable is number of diagnoses in that category.
-The main predictor is year; the other covariates are as above.
-The percent change is between 2011 and 2014.
-
-```{r dx_trends}
-dx_unadj = read_tsv('tables/dx_trends_table.tsv') %>%
-  #filter(year %in% c(2011, 2014)) %>%
-  left_join(select(totals, year, n_bene), by='year') %>%
-  mutate(dxpkp=1000*n_dx/n_bene) %>%
-  select(dx_cat, year, dxpkp) %>%
-  spread(year, dxpkp) %>%
-  mutate(pct_change=(`2014`/`2011`-1)*100) %>%
-  arrange(desc(`2011`)) %>%
-  mutate_if(is.numeric, function(x) round(x, 2))
-
-dx_adj = read_tsv('tables/dx_trends.tsv') %>%
-  rename(dx_cat=model) %>%
+read_tsv('tables/pde_approp_trends.tsv') %>%
   filter(term=='year') %>%
-  mult_est(3) %>%
-  model_ci() %>%
-  select(dx_cat, adj_change=change_str, p.value)
-
-dx_trends = full_join(dx_unadj, dx_adj, by='dx_cat')
-
-dx_trends %>% kable
+  select(antibiotic, estimate, p.value) %>%
+  mutate(antibiotic=factor(antibiotic, levels=c('overall', top_abx))) %>%
+  arrange(antibiotic) %>%
+  kable
 ```
@@
 ```{r dx_to_pde}
-dx_to_pde_unadj = read_tsv('tables/dx_rx_table.tsv') %>%
+dx_to_pde_unadj = read_tsv('tables/dx_to_pde_table.tsv') %>%
   filter(year %in% c(2011, 2014)) %>%
@@
 
-dx_to_pde_adj = read_tsv('tables/dx_rx_trends.tsv') %>%
+dx_to_pde_adj = read_tsv('tables/dx_to_pde_trends.tsv') %>%
   filter(term=='year') %>%
@@
 ```{r}
-x = dx_trends %>%
-  select(dx_cat, `2011`:`2014`) %>%
-  gather('year', 'value', `2011`:`2014`) %>%
+x = read_tsv('tables/dx_cat_counts.tsv') %>%
+  mutate(value=n_dx/n_bene*1000) %>%
   group_by(dx_cat) %>%
@@
   # convert to character to fit with other parts
-  mutate(value=as.character(value)) %>%
+  mutate(value=as.character(round(value))) %>%
   spread(dx_cat, value) %>%
   mutate(antibiotic='dx', key='mean_per_1kyr')
-  # select(-p.value) %>%
-  # gather('key', 'value', -dx_cat) %>%
-  # spread(dx_cat, value) %>%
-  # mutate(antibiotic='dx_trends')
 
@@
 ```
-
-# Questions for the paper
-
-## Lumping together diagnoses
-
-I could lump the multiple tiers for diagnoses together. This
-makes the table easier to interpret, but it doesn't make the point that
-diagnosing patterns for multiple tiers for the same diagnosis genus have
-similar dynamics, i.e., appropriate and inappropriate prescribing are changing
-the same ways.

### 2017-10-06 ms2/ms2_display.Rmd
lots of updates

@@
 ```{r global_options, include=FALSE}
-knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/', dev=c('png', 'pdf'),
-                      pdf.options(useDingbats=FALSE),
-                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE, autodep=TRUE)
+knitr::opts_chunk$set    
 import::from(tools, md5sum)
 library(forcats)
+BH = function(x) p.adjust(x, method='BH')
 ```
@@
   group_by(unit_type, unit, drug_group) %>%
-  summarize_at(vars(mean, fnz,mup), mean) %>% 
+  summarize_at(vars(mean, fnz, mup), mean) %>% 
   ungroup()
@@
   ggtitle("Linear univariate, HRR")
+```
+
+```{r hrr_linear_fnz_table}
+results %>%
+  filter(model=='univariate_fnz', unit_type=='hrr', term=='fnz') %>%
+  select(bug, drug_group, estimate, anova.p.value) %>%
+  filter(anova.p.value < 0.05) %>%
+  arrange(desc(estimate)) %>%
+  kable
 ```
@@
   ggtitle("Linear univariate, HRR")
+```
+
+```{r hrr_linear_mup_table}
+results %>%
+  filter(model=='univariate_mup', unit_type=='hrr', term=='mup') %>%
+  select(bug, drug_group, estimate, anova.p.value) %>%
+  filter(anova.p.value < 0.05) %>%
+  arrange(desc(estimate)) %>%
+  kable
 ```
@@
     geom_point() +
-    geom_text() +
+    #geom_text() +
+    ggrepel::geom_text_repel() +
     xlab(str_interp('mean ${d} claims per beneficiary')) +
@@
 show(state_cons_res_plot('S. pneumoniae', 'macrolide'))
+```
+
+```{r pneumo_macro_minimal}
+inner_join(
+  ineq %>%
+    filter(unit_type=='state', drug_group=='macrolide') %>%
+    select(state=unit, fnz),
+  state_abg %>%
+    filter(bug=='S. pneumoniae', drug_group=='macrolide') %>%
+    select(state, percent_nonsusceptible),
+by='state') %>%
+  left_join(regions, by='state') %>%
+  ggplot(aes(x=fnz, y=percent_nonsusceptible, label=state, color=region)) +
+  geom_smooth(method='lm', se=F, col='gray') +
+  geom_point() +
+  #geom_text() +
+  ggrepel::geom_text_repel() +
+  xlab(str_interp('fraction f of beneficiaries with a macrolide claim')) +
+  ylab(str_interp('percent nonsusceptible pneumococcus isolates')) +
+  theme_classic()
 ```

### 2017-10-06 ms1/ms1_3dx.R

@@ -23,6 +23,22 @@ read_years = function(years, template) {
 glm_f = function(df, frmla, ...) eval(substitute(function(df2) glm(frmla, data=df2, ...)))(df)
 frmla = y ~ year + age + n_cc + sex + race + dual + region
 
+# count total dx's and hcpcs counts
+read_years(2011:2014, '../../data/dx_hcpcs_count_%i.tsv') %>%
+  output_table('hcpcs_count')
+
+# count dx by categories
+read_years(2011:2014, '../../data/dx_cat_count_%i.tsv') %>%
+  rename(dx_cat=diagnosis_category) %>%
+  left_join(bene %>% count(year) %>% rename(n_bene=n), by='year') %>%
+  output_table('dx_cat_counts')
+
+# count many PDEs have any dx or E&M dx upstream
+read_years(2011:2014, '../../data/dx_any_pde_%i.tsv') %>%
+  mutate_at(vars(has_dx, has_em_dx), as.logical) %>%
+  count(year, has_dx, has_em_dx) %>%
+  output_table('pde_with_dx_upstream')
+
 # which dx's contribute to each abx?
 dx_from_pde = read_years(2011:2014, '../../data/dx_from_pde_%i.tsv') %>%
   rename(bene_id=BENE_ID, pde_id=PDE_ID, dx_cat=diagnosis_category)
@@ -72,8 +88,6 @@ pde_approp_table = pde_approp %>%
 # logistic regression
 # two ways to do this: either keep the "not infectious" diagnoses in the
 # denominator, or leave them out (i.e., filter tier != 4)
-# swo: i think this formula is repeated
-frmla = y ~ year + age + n_cc + sex + race + dual + region
 approp_trends_f = function(df) {
   df %>%
     mutate(y=tier <= 2) %>%

### 2017-10-07 ms1/ms1_display.Rmd
cleanups

@@
 est2pct = function(x) (exp(x) - 1)*100
+
 mult_est = function(df, m) {
@@
 }
+
 model_ci = function(df) {
@@
 
-pde_approp_risk_2011 = pde_approp_table %>%
-  # add up tiers 1 and 2
-  filter(year==2011, tier <= 2) %>%
+pde_inapprop_risk_2011 = pde_approp_table %>%
+  # add up tiers 3 and 4
+  filter(year==2011, tier > 2) %>%
   group_by(antibiotic) %>%
@@
 # compute risk ratios
-pde_approp_trends = read_tsv('tables/pde_approp_trends.tsv') %>%
+pde_inapprop_trends = read_tsv('tables/pde_inapprop_trends.tsv') %>%
   filter(term=='year') %>%
@@
          ci_high=estimate+ci) %>%
-  left_join(pde_approp_risk_2011, by='antibiotic') %>%
+  left_join(pde_inapprop_risk_2011, by='antibiotic') %>%
   # convert OR to RR (i.e., adjusted change)
@@
 ```{r hcpcs_count}
-read_tsv('tables/hcpcs_count.tsv') %>%
+hcpcs_count = read_tsv('tables/hcpcs_count.tsv') %>%
   group_by(hcpcs_flag) %>%
   summarize(n_dx=sum(n_dx)) %>%
-  mutate(f=n_dx/sum(n_dx)) %>%
-  kable
+  mutate(f=n_dx/sum(n_dx))
+
+hcpcs_count %>% kable
 ```
+
+This equals a total of `r prettyNum(sum(hcpcs_count$n_dx), big.mark=',')` diagnoses.
 
@@
   ungroup() %>%
-  mutate(f=n/sum(n), cumf=cumsum(f)) %>%
+  mutate(f=n/sum(n), cumf=cumsum(f), cumf1=1-cumf) %>%
   kable
@@
   theme_classic()
+```
+
+```{r pde_approp_table}
+pde_approp_table %>%
+  mutate(approp=tier <= 2) %>%
+  group_by(approp) %>%
+  summarize(n=sum(n)) %>%
+  mutate(f=n/sum(n)) %>%
+  kable
 ```
@@
   filter(term=='year') %>%
-  select(antibiotic, estimate, p.value) %>%
+  #select(antibiotic, estimate, p.value) %>%
+  mult_est(3) %>%
+  model_ci() %>%
   mutate(antibiotic=factor(antibiotic, levels=c('overall', top_abx))) %>%
   arrange(antibiotic) %>%
+  select(antibiotic, pct_change, ci_low, ci_high, p.value) %>%
   kable

### 2017-10-07 ms1/ms1_3dx.R

@@ -88,9 +88,9 @@ pde_approp_table = pde_approp %>%
 # logistic regression
 # two ways to do this: either keep the "not infectious" diagnoses in the
 # denominator, or leave them out (i.e., filter tier != 4)
-approp_trends_f = function(df) {
+inapprop_trends_f = function(df) {
   df %>%
-    mutate(y=tier <= 2) %>%
+    mutate(y=tier > 2) %>%
     (function(df) {
       bind_rows(
         glm_f(df, frmla, family='binomial') %>% tidy %>% mutate(antibiotic='overall'),
@@ -103,9 +103,9 @@ approp_trends_f = function(df) {
     })
 }
 
-pde_approp_trends = pde_approp %>%
-  approp_trends_f() %T>%
-  output_table('pde_approp_trends')
+pde_inapprop_trends = pde_approp %>%
+  inapprop_trends_f() %T>%
+  output_table('pde_inapprop_trends')
 
 # what are trends in prescribing practice?
 # swo: test this part

### 2017-10-10 ms2/ms2_display.Rmd
ms2: try combining

@@
   ungroup()
+
+ineq_hrr_mean = ineq %>%
+  filter(unit_type=='hrr') %>%
+  group_by(drug_group) %>%
+  summarize_at(vars(mean, fnz, mup), mean)
 
@@
 
-The plot shows the Spearman correlation coefficient $\rho$ between resistance $R$ and mean consumption $\mu$.
-Two stars after the bug/drug combination indicates that the adjusted $p < 0.05$, and one star means the unadjusted $p < 0.05$.
-Intervals show 95% confidence intervals.
+The plot shows the Spearman correlation coefficient $\rho$ between resistance
+$R$ and mean consumption $\mu$. Two stars after the bug/drug combination
+indicates that the adjusted $p < 0.05$, and one star means the unadjusted $p <
+0.05$. Intervals show 95% confidence intervals.
 
@@
 ```{r hrr_linear_fnz_table}
+results %>%
+  filter(model=='univariate_fnz', unit_type=='hrr', term=='fnz') %>%
+  select(bug, drug_group, estimate, anova.p.value) %>%
+  filter(anova.p.value < 0.05) %>%
+  arrange(desc(estimate)) %>%
+  kable
+
 results %>%
@@
   arrange(desc(estimate)) %>%
+  # swo: fanciful now
+  left_join(ineq_hrr_mean, by='drug_group') %>%
+  mutate(dR_nu_ext=-(mup-1)/fnz*estimate,
+         dR_nu_int=estimate/fnz) %>%
+  select(bug, drug_group, estimate, anova.p.value, dR_nu_ext, dR_nu_int) %>%
   kable
@@
 There are only two bug/drug combinations that are double-significant, and in those cases, neither metric gives information beyond the other.
+
+### Effect of both together
+
+```{r both_effects}
+results %>%
+  filter(unit_type=='hrr', model=='multivariate_fnz_mup', term != '(Intercept)') %>%
+  select(bug, drug_group, term, estimate) %>%
+  spread(term, estimate) %>%
+  rename(beta_fnz=fnz, beta_mup=mup) %>%
+  left_join(ineq_hrr_mean, by='drug_group') %>%
+  mutate(dR_ext=beta_fnz - (mup - 1)/fnz * beta_mup,
+         dR_int=beta_mup / fnz) %>%
+  kable
+```
 

### 2017-10-13 ms1/ms1_display.Rmd
ms1: better dx

@@
 
+adjusted_changes_firstfill = read_tsv('tables/model_abx_firstfill.tsv') %>%
+  rename(antibiotic=name) %>%
+  filter(term=='year') %>%
+  mutate(estimate=3*estimate,
+         std.error=sqrt(3)*std.error) %>%
+  model_ci() %>%
+  select(antibiotic, pct_change, ci_low, ci_high) %>%
+  arrange(pct_change) %>%
+  mutate(antibiotic=fct_relevel(fct_inorder(factor(antibiotic)), 'overall', after=Inf))
+
 # compute risk ratios
 pde_inapprop_trends = read_tsv('tables/pde_inapprop_trends.tsv') %>%
-  filter(term=='year') %>%
+  filter(term=='year', antibiotic != 'Other') %>%
   mutate(estimate=exp(3*estimate), # odds ratio over the whole time period
@@
   mutate_at(vars(estimate, ci_low, ci_high), function(x) risk_ratio_f(x, .$risk_2011 - 1)) %>%
-  # convert to percents, and reverse direction (to make it *inapprop*)
-  mutate_at(vars(estimate, ci_low, ci_high), function(x) -(x - 1)*100) %>%
-  select(antibiotic, pct_change=estimate, ci_low, ci_high, adj.p)
+  # convert to percents
+  mutate_at(vars(estimate, ci_low, ci_high), function(x) (x - 1)*100) %>%
+  select(antibiotic, model, pct_change=estimate, ci_low, ci_high, adj.p) %>%
   # now estimate is adjusted change in INappropriate prescribing
+  mutate(antibiotic=factor(antibiotic, levels=levels(adjusted_changes_firstfill$antibiotic), ordered=TRUE))
 
-adjusted_changes_firstfill = read_tsv('tables/model_abx_firstfill.tsv') %>%
-  rename(antibiotic=name) %>%
-  filter(term=='year') %>%
-  mutate(estimate=3*estimate,
-         std.error=sqrt(3)*std.error) %>%
-  model_ci() %>%
-  select(antibiotic, pct_change, ci_low, ci_high) %>%
-  arrange(pct_change) %>%
-  mutate(antibiotic=fct_inorder(antibiotic),
-         antibiotic=fct_relevel(antibiotic, 'overall', after=Inf))
-
-ggplot(adjusted_changes_firstfill, aes(x=antibiotic, y=pct_change, ymin=ci_low, ymax=ci_high)) +
+ggplot(adjusted_changes_firstfill, aes(x=antibiotic, y=pct_change, ymin=ci_low, ymax=ci_high, color=model)) +
   geom_hline(yintercept=0) + 
-  geom_bar(stat='identity', fill=NA, color='black') +
-  geom_errorbar(width=0.5) +
-  geom_point(dat=pde_approp_trends, color='blue', shape=1) +
-  geom_errorbar(dat=pde_approp_trends, width=0.25, color='blue') +
+  geom_bar(aes(color=NULL), stat='identity', fill=NA, color='black') +
+  geom_errorbar(aes(color=NULL), width=0.5) +
+  geom_point(dat=pde_inapprop_trends, shape=1) +
+  geom_errorbar(dat=pde_inapprop_trends, width=0.25) +
   coord_flip() +
@@
 ```{r pde_approp_trends_table}
-pde_approp_trends %>% kable
+pde_inapprop_trends %>% kable
 ```
@@
 If you use only tier 1 to describe "appropriate", then levofloxacin has the biggest decline in appropriateness, but if you use tiers 1 and 2, then it's change isn't significant.
-
-And, to be more careful, I show the answer to a similar question but using a regression to correct for covariates:
-
-```{r pde_approp_trends}
-read_tsv('tables/pde_approp_trends.tsv') %>%
-  filter(term=='year') %>%
-  #select(antibiotic, estimate, p.value) %>%
-  mult_est(3) %>%
-  model_ci() %>%
-  mutate(antibiotic=factor(antibiotic, levels=c('overall', top_abx))) %>%
-  arrange(antibiotic) %>%
-  select(antibiotic, pct_change, ci_low, ci_high, p.value) %>%
-  kable
-```
 

### 2017-10-13 ms1/ms1_3dx.R

@@ -71,13 +71,17 @@ dx_from_pde_table = crossing(antibiotic=top_abx, dx_cat=dxs) %>%
 
 # appropriateness
 fd = read_tsv('../../db/fd_icd/fd_categories.tsv') %>%
-  select(dx_cat=diagnosis_category, tier)
+  select(dx_cat=diagnosis_category, tier, t3_acute_respiratory) %>%
+  mutate(t3_acute_respiratory=t3_acute_respiratory=='Y')
 
 pde_approp = dx_from_pde %>%
   left_join(fd, by='dx_cat') %>%
-  replace_na(list(tier=4)) %>%
+  # "Remaining codes not listed elsewhere" are tier 3
+  replace_na(list(tier=3)) %>%
   group_by(year, bene_id, pde_id) %>%
-  summarize(antibiotic=unique(antibiotic), tier=min(tier)) %>%
+  summarize(antibiotic=unique(antibiotic),
+            tier=min(tier),
+            t3_acute_respiratory=all(t3_acute_respiratory)) %>%
   ungroup() %>%
   left_join(bene, by=c('year', 'bene_id'))
 
@@ -85,17 +89,18 @@ pde_approp_table = pde_approp %>%
   count(year, antibiotic, tier) %T>%
   output_table('pde_approp_table')
 
+pde_t3_acute_resp_table = pde_approp %>%
+  count(year, antibiotic, t3_acute_respiratory) %T>%
+  output_table('pde_t3_acute_resp_table')
+
 # logistic regression
-# two ways to do this: either keep the "not infectious" diagnoses in the
-# denominator, or leave them out (i.e., filter tier != 4)
-inapprop_trends_f = function(df) {
+inapprop_trends_model_f = function(df) {
   df %>%
-    mutate(y=tier > 2) %>%
     (function(df) {
       bind_rows(
         glm_f(df, frmla, family='binomial') %>% tidy %>% mutate(antibiotic='overall'),
         df %>%
-          filter(antibiotic %in% top_abx) %>%
+          mutate(antibiotic=fct_other(factor(antibiotic), keep=top_abx)) %>%
           group_by(antibiotic) %>%
           do(tidy(glm_f(., frmla, family='binomial'))) %>%
           ungroup()
@@ -103,19 +108,22 @@ inapprop_trends_f = function(df) {
     })
 }
 
+inapprop_trends_f = function(df) {
[…]

### 2017-10-16 ms2/ms2_2analyze.R
changing weights

@@ -64,9 +64,9 @@ summarize_abg = function(df, place) {
   df %>%
     group_by(bug, drug_group, !!place_var) %>%
     summarize(n_place_antibiograms=n(),
-              percent_nonsusceptible=weighted.mean(percent_nonsusceptible, w=n_isolates)) %>%
-    ungroup() %>%
-    mutate(regression_weight=n_place_antibiograms)
+              f_nonsusceptible=weighted.mean(f_nonsusceptible, w=n_isolates),
+              n_isolates=sum(n_isolates)) %>%
+    ungroup()
 }
 
 # summarize antibiograms by correcting for IP/OP and lab standard.
@@ -96,7 +96,7 @@ summarize_abg = function(df, place_name) {
         mutate(f_nonsusceptible=pred$fit,
                std.error=pred$se.fit,
                regression_weight=std.error ** (-2)) %>%
-        select(-has_inpatient, -standard)
+        select(-has_inpatient)
     })(.))
   
   abg_counts = df %>%
@@ -111,8 +111,8 @@ hrr_abg = summarize_abg(abg, hrr) %T>% write_tsv('abg_hrr.tsv')
 
 linear_model = function(df, y, xs) {
   frmla = as.formula(str_interp("${y} ~ ${str_c(xs, collapse=' + ')}"))
-  m = lm(frmla, weights=n_place_antibiograms, data=df)
-  #m = lm(frmla, weights=std.error ** (-2), data=df)
+  #m = lm(frmla, weights=n_place_antibiograms, data=df)
+  m = lm(frmla, weights=std.error ** (-2), data=df)
 
   anova_res = anova(m) %>%
     tidy %>%
@@ -137,7 +137,7 @@ spearman_model = function(df, y_name, x_name) {
   if (length(y) > 3) {
     cis = mada::CIrho(m$estimate, length(y), level=0.95)
   } else {
-    cis = c(0, 0, 1)
+    cis = c(NA, 0, 1)
   }
 
   data_frame(term=x_name,
@@ -182,13 +182,12 @@ ims = read_tsv('../../db/ims/ims.tsv') %>%
   select(state, drug_group, rate)
 
 ims_results = ims %>%
-  right_join(state_abg, by=c('state', 'drug_group')) %>%
-  mutate(y=percent_nonsusceptible/100) %>%
+  inner_join(state_abg, by=c('state', 'drug_group')) %>%
   group_by(bug, drug_group) %>%
   do((function (df) {
[…]

### 2017-10-16 ms1/ms1_display.Rmd
ms1 simplify

@@
   select(antibiotic, dx_cat, pct) %>%
-  spread(dx_cat, pct) %T>%
-  kable
+  spread(dx_cat, pct)
+
+dx_from_pde %>% kable
 ```

### 2017-10-16 ms1/ms1_3dx.R

@@ -1,5 +1,3 @@
-library(forcats)
-
 # load data
 # NB: I put DC into the South
 

### 2017-10-16 ms1/ms1_display.Rmd
ms1: kludgy new graphs

@@
 
-```{r adjusted_changes}
-adjusted_changes = read_tsv('tables/model_abx.tsv') %>%
-  rename(antibiotic=name) %>%
-  filter(term=='year') %>%
-  mutate(estimate=3*estimate,
-         std.error=sqrt(3)*std.error) %>%
-  model_ci() %>%
-  select(antibiotic, pct_change, ci_low, ci_high) %>%
-  arrange(pct_change) %>%
-  mutate(antibiotic=fct_inorder(antibiotic),
-         antibiotic=fct_relevel(antibiotic, 'overall', after=Inf))
-
-unadjusted_changes = bind_rows(claims_by_abx, claims_overall) %>%
-  filter(antibiotic %in% c('overall', top_abx), year %in% c(2011, 2014)) %>%
-  select(antibiotic, year, cpkp) %>%
-  spread(year, cpkp) %>%
-  mutate(change=`2014`-`2011`,
-         fold_change=change/`2011`,
-         pct_change=100*fold_change) %>%
-  select(antibiotic, pct_change)
-
-adjusted_changes %>%
-  ggplot(aes(x=antibiotic, y=pct_change, ymin=ci_low, ymax=ci_high)) +
-  geom_bar(stat='identity', fill=NA, color='black') +
-  geom_errorbar(width=0.5) +
-  geom_point(dat=unadjusted_changes, aes(x=antibiotic, y=pct_change, ymin=NULL, ymax=NULL), shape=1) +
-  coord_flip() +
-  xlab('') +
-  ylab('adjusted change in consumption 2011 to 2014 (%)') +
-  scale_y_continuous(limits=c(-22, 32), breaks=c(-20, -10, 0, 10, 20, 30)) +
-  theme_minimal() +
-  theme(axis.title.x = element_text(size=16),
-        axis.text = element_text(size=16))
-```
-
-The bars show the adjusted change in consumption. Black circles show unadjusted change in consumption. Blue shows the adjusted change in inappropriate prescribing.
-
-### First fills
-
-Bars show the adjusted change in prescribing, considering only first fills. The blue dots show the change in inappropriate prescribing (which only includes first fills).
-
 ```{r approp_changes}
@@
   mutate(antibiotic=factor(antibiotic, levels=levels(adjusted_changes_firstfill$antibiotic), ordered=TRUE))
+```
 
+```{r adjusted_changes}
+adjusted_changes = read_tsv('tables/model_abx.tsv') %>%
+  rename(antibiotic=name) %>%
+  filter(term=='year') %>%
+  mutate(estimate=3*estimate,
+         std.error=sqrt(3)*std.error) %>%
+  model_ci() %>%
+  select(antibiotic, pct_change, ci_low, ci_high) %>%
+  arrange(pct_change) %>%
+  mutate(antibiotic=fct_inorder(antibiotic),
+         antibiotic=fct_relevel(antibiotic, 'overall', after=Inf))
+
+adjusted_change_levels = adjusted_changes %>%
+  filter(antibiotic != 'overall') %>%
+  pull(antibiotic) %>%
+  as.character
+
+unadjusted_changes = bind_rows(claims_by_abx, claims_overall) %>%
+  filter(antibiotic %in% c('overall', top_abx), year %in% c(2011, 2014)) %>%
+  select(antibiotic, year, cpkp) %>%
+  spread(year, cpkp) %>%
+  mutate(change=`2014`-`2011`,
+         fold_change=change/`2011`,
+         pct_change=100*fold_change) %>%
+  select(antibiotic, pct_change)
+
+adjusted_changes %>%
+  mutate(antibiotic=as.character(antibiotic)) %>%
+  bind_rows(pde_inapprop_trends %>%
+              filter(antibiotic=='overall', model=='12/3') %>%
+              mutate(antibiotic='overall_inapp')) %>%
+  mutate(antibiotic=factor(antibiotic, levels=c(adjusted_change_levels, 'overall_inapp', 'overall'))) %>%
+  ggplot(aes(x=antibiotic, y=pct_change, ymin=ci_low, ymax=ci_high)) +
+  geom_bar(stat='identity', fill=NA, color='black') +
+  geom_errorbar(width=0.5) +
+  coord_flip() +
+  xlab('') +
+  ylab('adjusted change in consumption 2011 to 2014 (%)') +
+  scale_y_continuous(limits=c(-22, 32), breaks=c(-20, -10, 0, 10, 20, 30)) +
+  theme_minimal() +
+  theme(axis.title.x = element_text(size=16),
+        axis.text = element_text(size=16))
+```
+
+The bars show the adjusted change in consumption. Black circles show unadjusted change in consumption. Blue shows the adjusted change in inappropriate prescribing.
+
+### First fills
+
+Bars show the adjusted change in prescribing, considering only first fills. The blue dots show the change in inappropriate prescribing (which only includes first fills).
+
+```{r}
 ggplot(adjusted_changes_firstfill, aes(x=antibiotic, y=pct_change, ymin=ci_low, ymax=ci_high, color=model)) +
@@
         axis.text = element_text(size=16))
+```
+
+```{r}
+bind_rows(
+  unadjusted_changes %>% mutate(src='unadjusted, all fills'),
+  adjusted_changes %>% mutate(src='adjusted, all fills'),
+  adjusted_changes_firstfill %>% mutate(src='adjusted, first fills')
+) %>%
+  mutate(antibiotic=factor(antibiotic, levels=c(adjusted_change_levels, 'overall'))) %>%
+  ggplot(aes(x=antibiotic, y=pct_change, ymin=ci_low, ymax=ci_high, color=src)) +
+  geom_hline(yintercept=0, color='black') +
+  geom_point(shape=1) +
+  geom_errorbar(width=0.5) +
+  coord_flip() +
+  xlab('') + ylab('change from 2011 to 2014 (%)') +
+  theme_minimal()
 ```

### 2017-10-23 ms1/ms1_display.Rmd
fixing dx-to-pde

@@
   scale_y_continuous(limits=c(0, 100), breaks=c(0, 50, 100)) +
-  #ylim(0, 140) +
-  theme_minimal() +
-  ggtitle('Unadjusted consumption, next 5')
-```
-
-```{r claims_by_top_abx_facet, eval=FALSE}
-claims_by_top_abx %>%
-  filter(antibiotic %in% c('azithromycin', 'ciprofloxacin', 'amoxicillin', 'cephalexin', 'trimethoprim/sulfamethoxazole')) %>%
-  ggplot(aes(x=year, y=cpkp, color=antibiotic)) +
-  #ggplot(aes(x=year, y=cpkp)) +
-  facet_wrap(~antibiotic, ncol=5) +
-  geom_point() + geom_line() +
-  scale_y_continuous(limits=c(0, 250), breaks=c(0, 50, 100, 150, 200, 250)) +
-  theme_minimal() +
-  ggtitle('Unadjusted consumption, top 5')
-
-claims_by_top_abx %>%
-  filter(antibiotic %in% c('levofloxacin', 'amoxicillin/clavulanate', 'doxycycline', 'nitrofurantoin', 'clindamycin')) %>%
-  ggplot(aes(x=year, y=cpkp, color=antibiotic)) +
-  #ggplot(aes(x=year, y=cpkp)) +
-  facet_wrap(~antibiotic, ncol=5) +
-  geom_point() + geom_line() +
-  scale_y_continuous(limits=c(0, 150), breaks=c(0, 50, 100)) +
   #ylim(0, 140) +

### 2017-10-23 ms1/ms1_3dx.R

@@ -102,28 +102,38 @@ dx_to_pde = read_years(2011:2014, '../../data/dx_to_pde_%i.tsv') %>%
   rename(bene_id=BENE_ID, dx_cat=diagnosis_category) %>%
   # lump abx outside of top 10 into "other"
   mutate(antibiotic=fct_other(factor(antibiotic), keep=top_abx)) %>%
-  count(year, bene_id, dx_cat, antibiotic) %>%
   left_join(bene, by=c('year', 'bene_id'))
 
+# count how many of each encounter type there are in a year
+dx_denom = dx_to_pde %>%
+  select(year, encounter_id, dx_cat) %>%
+  distinct() %>%
+  count(year, dx_cat) %>% ungroup() %>%
+  rename(n_dx=n)
+
 dx_to_pde_table = dx_to_pde %>%
-  group_by(year, dx_cat, antibiotic) %>%
-  summarize(n=sum(n)) %>%
-  mutate(f=n/sum(n)) %>%
-  ungroup() %T>%
+  count(year, dx_cat, antibiotic) %>% ungroup() %>%
+  rename(n_dx_with_abx=n) %>%
+  left_join(dx_denom, by=c('year', 'dx_cat')) %>%
+  mutate(f_dx_with_abx=n_dx_with_abx/n_dx) %T>%
   output_table('dx_to_pde_table')
+
+encounter = dx_to_pde %>%
+  select(year, bene_id, encounter_id, dx_cat) %>%
+  distinct()
 
 dx_to_pde_trend_f = function (a, dx) {
   dx_to_pde %>%
     filter(dx_cat==dx) %>%
-    group_by(year, bene_id, encounter_id) %>%
-    summarize(age=unique(age), sex=unique(sex), race=unique(race), dual=unique(dual), n_cc=unique(n_cc), region=unique(region),
+    group_by(year, encounter_id) %>%
+    summarize(bene_id=unique(bene_id), age=unique(age), sex=unique(sex), race=unique(race), dual=unique(dual), n_cc=unique(n_cc), region=unique(region),
               y=a %in% antibiotic) %>%
     ungroup() %>%
     glm_f(frmla, family='binomial') %>%
     sandwich_tidy
 }
 
-dx_rx_trends = crossing(antibiotic=top_abx, dx_cat=dxs) %>%
+dx_to_pde_trends = crossing(antibiotic=top_abx, dx_cat=dxs) %>%
[…]

### 2017-10-26 ms1/ms1_display.Rmd
updating display

@@
 
-n_unique_bene = scan('tables/n_unique_bene.tsv')
-n_million_unique_bene = round(scan('tables/n_unique_bene.tsv') / 1e6, 2)
-n_pde = sum(claims_by_abx$n)
-n_million_pde = round(n_pde / 1e6, 2)
-
 est2pct = function(x) (exp(x) - 1)*100
@@
 
-There are `r n_million_unique_bene` million unique beneficiaries and `r n_million_pde` million PDEs across all study years.
-
 # Beneficiary characteristics
 
-There were `r prettyNum(n_unique_bene, big.mark=',')` unique beneficiaries and `r prettyNum(n_pde, big.mark=',')` PDEs.
+```{r unique_bene_pde}
+round_millions = function(x) round(x / 1e6, 2)
+pn = function(x) prettyNum(x, big.mark=',')
+
+n_unique_bene = scan('tables/n_unique_bene.tsv')
+
+claims_by_fill = read_tsv('tables/claims_by_fill.tsv')
+n_pde = sum(claims_by_fill$n)
+n_refills = claims_by_fill %>% filter(!first_fill) %$% n
+```
+
+There were `r round_millions(n_unique_bene)` million unique beneficiaries. They had `r round_millions(n_pde)` million PDEs, of which `r round_millions(n_refills)` million were refills.
 
@@
 
-## Claims by age group
+## Claims by demography
 
-```{r claims_by_age}
+```{r claims_by_demography}
 change_demog = read_tsv('tables/model_overall_demography.tsv') %>%
@@
                         between(.$age, 76, 85) ~ 'age76_85',
-                        between(.$age, 86, 95) ~ 'age86_95')) %>%
+                        between(.$age, 86, 95) ~ 'age86_95',
+                        .$age >= 96 ~ 'age96_')) %>%
   group_by(year, name) %>%
@@
 
+That last table could also use an "overall" row:
+
+```{r}
+claims_overall %>%
+  filter(year %in% c(2011, 2014)) %>%
+  mutate(cpkp=round(cpkp)) %>%
+  kable
+```
+
 # Adjusted consumption changes by drug
@@
 
+### Trends in appropriateness
+
+Just by eye, it looks like the changes in appropriateness are small.
+
+```{r tier_graph}
+pde_tier_table = read_tsv('tables/pde_approp_table.tsv') %>%
+  group_by(year, antibiotic) %>%
+  mutate(f=n/sum(n)) %>%
+  ungroup()
+
+pde_tier_table %>%
+  filter(antibiotic %in% top_abx) %>%
+  ggplot(aes(x=factor(year), y=f, fill=factor(tier))) +
+  geom_bar(stat='identity') +
+  facet_grid(~antibiotic) +
+  theme_classic()
+```
+
+```{r approp_trends}
+raw_adjusted_changes = read_tsv('tables/model_abx.tsv') %>%
+  rename(antibiotic=name) %>%
+  filter(term=='year') %>%
+  mutate(estimate=3*estimate,
+         std.error=sqrt(3)*std.error)
+
+# fraction inappropriate in 2011
+f_i = pde_tier_table %>%
+  filter(year==2011) %>%
+  mutate(inapp=tier==3) %>%
+  group_by(inapp) %>%
+  summarize(n=sum(n)) %>%
+  mutate(f=n/sum(n)) %>%
+  filter(inapp) %$%
+  f
+
+# fold change in consumption
+pde_trend = raw_adjusted_changes %>%
+  filter(antibiotic=='overall') %>%
+  # multiplications by 3 done above
+  mutate(ci=1.96*std.error,
+         rr=exp(estimate),
+         rr_lo=exp(estimate-ci),
+         rr_hi=exp(estimate+ci))
+
+d_c = pde_trend$rr
+d_c_lo = pde_trend$rr_lo
+d_c_hi = pde_trend$rr_hi
+
+risk_ratio_f = function(odr, p) odr / (1 + p * (odr - 1))
+
+pde_inapprop_trend = read_tsv('tables/pde_inapprop_trend.tsv') %>%
+  filter(term=='year') %>%
+  mutate(estimate=3*estimate,
+         std.error=sqrt(3)*std.error,
+         ci=1.96*std.error,
+         or=exp(estimate),
+         or_lo=exp(estimate-ci),
+         or_hi=exp(estimate+ci))
+
+d_f = risk_ratio_f(pde_inapprop_trend$or, f_i)
+d_f_lo = risk_ratio_f(pde_inapprop_trend$or_lo, f_i)
+d_f_hi = risk_ratio_f(pde_inapprop_trend$or_hi, f_i)
+
+rr2pct = function(x) round((x - 1) * 100, 2)
+
+inapprop_changes = data_frame(
+  antibiotic=c('overall_inapp', 'overall_app'),
+  pct_change=c(d_c * d_f, (1.0 - d_f * f_i) / (1.0 - f_i) * d_c),
+  ci_low=c(d_c_lo * d_f_lo, (1.0 - d_f_hi * f_i) / (1.0 - f_i) * d_c_lo),
+  ci_high=c(d_c_hi * d_f_hi, (1.0 - d_f_lo * f_i) / (1.0 - f_i) * d_c_hi)
+) %>%
+  mutate_if(is.numeric, rr2pct)
+```
+
+Inappropriate prescribing changed by `r rr2pct(d_c * d_f)`% (95% CI `r rr2pct(d_c_lo * d_f_lo)` -- `r rr2pct(d_c_hi * d_f_hi)`).
+Appropriate prescribing changed by `r rr2pct((1.0 - d_f * f_i) / (1.0 - f_i) * d_c)`% (95% CI `r rr2pct((1.0 - d_f_hi * f_i) / (1.0 - f_i) * d_c_lo)` -- `r rr2pct((1.0 - d_f_lo * f_i) / (1.0 - f_i) * d_c_hi)`).
+
 ### All fills
@@
 
-```{r approp_changes}
-# read in the raw and trend data
-pde_approp_table = read_tsv('tables/pde_approp_table.tsv') %>%
-  (function(df) bind_rows(
-    df,
-    df %>% group_by(year, tier) %>%
-      summarize(n=sum(n)) %>%
-      mutate(antibiotic='overall')
-  )) %>%
-  filter(antibiotic %in% c(top_abx, 'overall')) %>%
-  group_by(year, antibiotic) %>%
-  mutate(f=n/sum(n)) %>%
-  ungroup()
-
-pde_inapprop_risk_2011 = pde_approp_table %>%
-  # add up tiers 3 and 4
-  filter(year==2011, tier > 2) %>%
-  group_by(antibiotic) %>%
-  summarize(risk_2011=sum(f))
-
-risk_ratio_f = function(odr, p) odr / (1 + p * (odr - 1))
-
-adjusted_changes_firstfill = read_tsv('tables/model_abx_firstfill.tsv') %>%
-  rename(antibiotic=name) %>%
-  filter(term=='year') %>%
-  mutate(estimate=3*estimate,
-         std.error=sqrt(3)*std.error) %>%
+```{r adjusted_changes}
+adjusted_changes = raw_adjusted_changes %>%
   model_ci() %>%
-  select(antibiotic, pct_change, ci_low, ci_high) %>%
-  arrange(pct_change) %>%
-  mutate(antibiotic=fct_relevel(fct_inorder(factor(antibiotic)), 'overall', after=Inf))
-
-# compute risk ratios
-pde_inapprop_trends = read_tsv('tables/pde_inapprop_trends.tsv') %>%
-  filter(term=='year', antibiotic != 'Other') %>%
-  mutate(estimate=exp(3*estimate), # odds ratio over the whole time period
-         std.error=sqrt(3)*std.error,
-         adj.p=p.adjust(p.value, method='BH'),
-         ci=1.96*std.error,
-         ci_low=estimate-ci,
-         ci_high=estimate+ci) %>%
-  left_join(pde_inapprop_risk_2011, by='antibiotic') %>%
-  # convert OR to RR (i.e., adjusted change)
-  mutate_at(vars(estimate, ci_low, ci_high), function(x) risk_ratio_f(x, .$risk_2011 - 1)) %>%
-  # convert to percents
-  mutate_at(vars(estimate, ci_low, ci_high), function(x) (x - 1)*100) %>%
-  select(antibiotic, model, pct_change=estimate, ci_low, ci_high, adj.p) %>%
-  # now estimate is adjusted change in INappropriate prescribing
-  mutate(antibiotic=factor(antibiotic, levels=levels(adjusted_changes_firstfill$antibiotic), ordered=TRUE))
-```
-
-```{r adjusted_changes}
-adjusted_changes = read_tsv('tables/model_abx.tsv') %>%
-  rename(antibiotic=name) %>%
-  filter(term=='year') %>%
-  mutate(estimate=3*estimate,
-         std.error=sqrt(3)*std.error) %>%
-  model_ci() %>%
-  select(antibiotic, pct_change, ci_low, ci_high) %>%
-  arrange(pct_change) %>%
-  mutate(antibiotic=fct_inorder(antibiotic),
-         antibiotic=fct_relevel(antibiotic, 'overall', after=Inf))
+  arrange(desc(pct_change)) %>%
+  select(antibiotic, pct_change, ci_low, ci_high)
 
 adjusted_change_levels = adjusted_changes %>%
-  filter(antibiotic != 'overall') %>%
-  pull(antibiotic) %>%
-  as.character
+  filter(antibiotic != 'overall') %$%
+  { c('overall', 'overall_app', 'overall_inapp', antibiotic) } %>%
+  rev
 
@@
 adjusted_changes %>%
+  bind_rows(inapprop_changes) %>%
   mutate(antibiotic=as.character(antibiotic)) %>%
-  bind_rows(pde_inapprop_trends %>%
-              filter(antibiotic=='overall', model=='12/3') %>%
-              mutate(antibiotic='overall_inapp')) %>%
-  mutate(antibiotic=factor(antibiotic, levels=c(adjusted_change_levels, 'overall_inapp', 'overall'))) %>%
+  mutate(antibiotic=factor(antibiotic, levels=adjusted_change_levels)) %>%
   ggplot(aes(x=antibiotic, y=pct_change, ymin=ci_low, ymax=ci_high)) +
@@
 The bars show the adjusted change in consumption. Black circles show unadjusted change in consumption. Blue shows the adjusted change in inappropriate prescribing.
-
-### First fills
-
-Bars show the adjusted change in prescribing, considering only first fills. The blue dots show the change in inappropriate prescribing (which only includes first fills).
-
-```{r}
-ggplot(adjusted_changes_firstfill, aes(x=antibiotic, y=pct_change, ymin=ci_low, ymax=ci_high, color=model)) +
-  geom_hline(yintercept=0) + 
-  geom_bar(aes(color=NULL), stat='identity', fill=NA, color='black') +
-  geom_errorbar(aes(color=NULL), width=0.5) +
-  geom_point(dat=pde_inapprop_trends, shape=1) +
-  geom_errorbar(dat=pde_inapprop_trends, width=0.25) +
-  coord_flip() +
-  xlab('') +
-  ylab('adjusted change from 2011 to 2014 (%)') +
-  scale_y_continuous(limits=c(-22, 35), breaks=c(-20, -10, 0, 10, 20, 30)) +
-  theme_classic() +
-  theme(axis.title.x = element_text(size=16),
-        axis.text = element_text(size=16))
-```
-
-```{r}
-bind_rows(
-  unadjusted_changes %>% mutate(src='unadjusted, all fills'),
-  adjusted_changes %>% mutate(src='adjusted, all fills'),
-  adjusted_changes_firstfill %>% mutate(src='adjusted, first fills')
-) %>%
-  mutate(antibiotic=factor(antibiotic, levels=c(adjusted_change_levels, 'overall'))) %>%
-  ggplot(aes(x=antibiotic, y=pct_change, ymin=ci_low, ymax=ci_high, color=src)) +
-  geom_hline(yintercept=0, color='black') +
-  geom_point(shape=1) +
-  geom_errorbar(width=0.5) +
-  coord_flip() +
-  xlab('') + ylab('change from 2011 to 2014 (%)') +
-  theme_minimal()
-```
 
@@
 Consumption of azithromycin fell by `r adjusted_az_decrease`% over the entire study period. Levofloxacin consumption increased by `r adjusted_levo_increase`% over the entire study period.
-
-### Adjusted changes in inappropriate prescribing
-
-```{r pde_approp_trends_table}
-pde_inapprop_trends %>% kable
-```
-
-## Fluoroquinolones
-
-Consumption of ciprofloxacin is slightly decreasing, and consumption of levofloxacin is increasing. How does fluoroquinolone consumption overall change?
-
-```{r fq_table}
-fq_names = claims_by_abx %>%
-  filter(str_detect(antibiotic, 'floxacin')) %>%
-  group_by(antibiotic) %>%
-  summarize(n=sum(n)) %>%
-  arrange(desc(n)) %$%
-  antibiotic
-
-fq_table = claims_by_abx %>%
-  filter(antibiotic %in% fq_names) %>%
-  group_by(year) %>%
-  summarize(cpkp=sum(cpkp))
-
-fq_change_pct_unadj = fq_table %>%
-  spread(year, cpkp) %$%
-  { (`2014`-`2011`)/`2011` } %>%
-  { round(100*., 2) }
-```
-
-```{r fq_trends}
-fq_trends = read_tsv('tables/model_fq.tsv') %>%
-  mult_est(3) %>%
-  model_ci() %>%
-  select(term, pct_change, ci_low, ci_high, p.value)
-
-fq_change_adj = fq_trends %>%
-  filter(term=='year') %>% 
-  mutate_at(vars(pct_change, ci_low, ci_high), function(x) round(x, 2)) %$%
-  str_interp("${pct_change}% (${ci_low}% -- ${ci_high}%)")
-```
-
-Overall, fluoroquinolones changed by a tiny amount, `r fq_change_pct_unadj`% (unadjusted). The adjusted change is `r fq_change_adj` (%) over this period. This included `r fq_names`.
 
@@
 
-## Trends in appropriateness
-
-Just by eye, it looks like the changes in appropriateness are small.
-
-```{r approp_graph}
-pde_approp_table %>%
-  #filter(year %in% c(2011, 2014)) %>%
-  ggplot(aes(x=factor(year), y=f, fill=factor(tier))) +
-  geom_bar(stat='identity') +
-  facet_grid(~antibiotic) +
-  theme_classic()
-```
-
-```{r pde_approp_table}
-pde_approp_table %>%
-  mutate(approp=tier <= 2) %>%
-  group_by(approp) %>%
-  summarize(n=sum(n)) %>%
-  mutate(f=n/sum(n)) %>%
-  kable
-```
-
-I'll focus on tier 1 and 2 because the SSTI drugs don't have common tier 1 uses.
-I'll call this "high tier" prescribign.
-If prescribing had changed in an optimal way, only low-tier prescribing would have declined, or only high-tier prescribing would have been added.
-
-If a drug declined by $X$, then the optimal change in the high tier fraction would be $1/(1-X) = 1+X + \mathcal{O}(X^2)$.
-For example, overall prescribing fell by about 5%, and high-tier prescribing increased by 2.5%.
-Azithromycin prescribing fell by about 20%, and high-tier azithro prescribing rose by only 6%.
-
-I'm not sure what to make of these results, because they depend strongly on how you define the tiers.
-If you use only tier 1 to describe "appropriate", then levofloxacin has the biggest decline in appropriateness, but if you use tiers 1 and 2, then it's change isn't significant.
-
 ## Common diagnoses by drug
@@
 The values are percents in 2011.
+A row can add up to more than 100% because one PDE can be associated with multiple diagnoses.
 
@@
   filter(year %in% c(2011, 2014)) %>%
-  select(dx_cat, antibiotic, year, f) %>%
+  select(dx_cat, antibiotic, year, f=f_encounters_with_abx) %>%
   spread(year, f)
@@
   arrange(dx_cat, antibiotic)
-
-risk_ratio_f = function(odr, p) odr / (1 + p * (odr - 1))
 

### 2017-10-30 ms1/ms1_display.Rmd
display: small fixes

@@
 
-Inappropriate prescribing changed by `r rr2pct(d_c * d_f)`% (95% CI `r rr2pct(d_c_lo * d_f_lo)` -- `r rr2pct(d_c_hi * d_f_hi)`).
-Appropriate prescribing changed by `r rr2pct((1.0 - d_f * f_i) / (1.0 - f_i) * d_c)`% (95% CI `r rr2pct((1.0 - d_f_hi * f_i) / (1.0 - f_i) * d_c_lo)` -- `r rr2pct((1.0 - d_f_lo * f_i) / (1.0 - f_i) * d_c_hi)`).
-
 ### All fills
@@
   filter(antibiotic != 'overall') %$%
-  { c('overall', 'overall_app', 'overall_inapp', antibiotic) } %>%
-  rev
+  { c('overall', 'overall_app', 'overall_inapp', antibiotic) }
 
@@
   mutate(antibiotic=as.character(antibiotic)) %>%
-  mutate(antibiotic=factor(antibiotic, levels=adjusted_change_levels)) %>%
+  mutate(antibiotic=factor(antibiotic, levels=rev(adjusted_change_levels))) %>%
   ggplot(aes(x=antibiotic, y=pct_change, ymin=ci_low, ymax=ci_high)) +
-  geom_bar(stat='identity', fill=NA, color='black') +
+  geom_bar(stat='identity', fill='white', color='black') +
   geom_errorbar(width=0.5) +
@@
 
-The bars show the adjusted change in consumption. Black circles show unadjusted change in consumption. Blue shows the adjusted change in inappropriate prescribing.
+The bars show the adjusted change in consumption.
 
@@
 ```{r adjusted_changes_table}
-adjusted_changes %>% arrange(desc(antibiotic)) %>% kable
+adjusted_changes %>%
+  bind_rows(inapprop_changes) %>%
+  mutate(antibiotic=factor(antibiotic, levels=adjusted_change_levels)) %>%
+  arrange(antibiotic) %>%
+  kable
 

### 2017-11-01 ms1/ms1_display.Rmd
collapse

@@
 ```{r tier_graph}
+pde_tier_table = dx_pde_counts %>%
+  filter(has_dx_em) %>%
+
+pde_tier_table = read_tsv('raw_data/dx_pde_counts.tsv') %>%
+  group_by(year,
+
 pde_tier_table = read_tsv('tables/pde_approp_table.tsv') %>%
@@
 ```{r hcpcs_count}
-hcpcs_count = read_tsv('tables/hcpcs_count.tsv') %>%
-  group_by(hcpcs_flag) %>%
-  summarize(n_dx=sum(n_dx)) %>%
-  mutate(f=n_dx/sum(n_dx))
+encounter_counts = read_tsv('raw_data/dx_cat_count.tsv') %>%
+  summarize(n_encounters=sum(n_encounters),
+            n_em_encounters=sum(n_encounters[hcpcs_em]),
+            f_em_encounters=weighted.mean(hcpcs_em, weights=n_encounters))
 
@@
 
-This equals a total of `r prettyNum(sum(hcpcs_count$n_dx), big.mark=',')` diagnoses.
+This equals a total of `r prettyNum(sum(encounter_counts$n_encounters), big.mark=',')` encounters.
 

### 2017-11-01 ms1/ms1_0organize_dx2pde.py

@@ -1,91 +1,0 @@
-#!/usr/bin/env python3
-
-import fileinput
-
-class DxReader:
-    top_abx = [None, 'azithromycin', 'ciprofloxacin', 'amoxicillin', 'cephalexin',
-        'trimethoprim/sulfamethoxazole', 'levofloxacin', 'amoxicillin/clavulanate',
-        'doxycycline', 'nitrofurantoin', 'clindamycin']
-
-    out_abx = ['.n', 'azithromycin', 'ciprofloxacin', 'amoxicillin', 'cephalexin',
-        'trimethoprim/sulfamethoxazole', 'levofloxacin', 'amoxicillin/clavulanate',
-        'doxycycline', 'nitrofurantoin', 'clindamycin']
-
-    def __init__(self):
-        self.bene = None
-        self.dat = {}
-        self.first_line = True
-
-        print('bene_id', 'dx_cat', 'n_encounters', *self.out_abx[1:], sep='\t')
-
-    def parse_line(self, line):
-        fields = line.rstrip().split('\t')
-        bene, enc, dx = fields[0:3]
-
-        if len(fields) == 4:
-            abx = fields[3]
-        else:
-            abx = None
-
-        return (bene, enc, dx, abx)
-
-    def process_line(self, line):
-        bene, enc, dx, abx = self.parse_line(line)
-
-        if self.first_line:
-            self.bene = bene
-            self.dat = {}
-            self.add_to_dat(enc, dx, abx)
-            self.first_line = False
-        else:
-            if bene != self.bene:
-                self.report()
-                self.dat = {}
-                self.bene = bene
-
-            self.add_to_dat(enc, dx, abx)
-
-    def add_to_dat(self, enc, dx, abx):
-        if enc in self.dat:
-            self.dat[enc]['.dx'].append(dx)
-
-            if abx in self.dat[enc]:
-                self.dat[enc][abx] += 1
-            else:
-                self.dat[enc][abx] = 1
-        else:
-            self.dat[enc] = {'.dx': [dx], abx: 1}
-
-    def report(self):
-        dxs = set()
-        for enc in self.dat:
-            dxs.update(self.dat[enc]['.dx'])
-
-        out = {dx: {'.n': 0} for dx in dxs}
-
-        for enc in self.dat:
-            for dx in self.dat[enc]['.dx']:
-                out[dx]['.n'] += 1
-                for abx in self.dat[enc].keys():
-                    if abx == '.dx':
-                        continue
-
-                    if abx not in self.top_abx:
[…]

### 2017-11-01 ms1/ms1_1clean.R

@@ -1,55 +1,28 @@
 #!/usr/bin/env Rscript
 
-# Process the bene, pde, and dx data so it's ready for analysis
-
-# lower the names of a tibble
-lower_names = function(df) {
-  ns = names(df)
-  rename(df, !!!setNames(as.list(ns), sapply(ns, tolower)))
-}
-
-regions = read_tsv('db/census-regions/census-regions.tsv') %>%
+# Process the bene and pde
+region = read_tsv('db/census-regions.tsv') %>%
   select(state, region)
 
-load_data = function(year) {
-  # pde data
-  pde = read_tsv(sprintf('raw_data/pde_%i.tsv', year)) %>%
-    lower_names() %>%
-    mutate(pde_date=dmy(pde_date)) %>%
-    select(bene_id, pde_id, pde_date, antibiotic, fill_num) %>%
-    # for each bene, date, and drug, keep only one record, giving preference to first fills.
-    group_by(bene_id, pde_date, antibiotic) %>%
-    summarize(pde_id=min(pde_id), fill_num=min(fill_num)) %>%
-    ungroup() %>%
-    mutate(year=year) %>%
-    select(year, bene_id, pde_id, antibiotic, fill_num)
+pde = read_tsv('raw_data/pde.tsv') %>%
+  mutate(first_fill=fill_num==0) %>%
+  select(year, bene_id, antibiotic, first_fill)
 
-  # total numbers of PDEs
-  n_claims = count(pde, bene_id) %>% rename(n_claims=n)
-  n_claims_firstfill = pde %>%
-    filter(fill_num==0) %>%
-    count(bene_id) %>%
-    rename(n_claims_firstfill=n)
+pde_denom = pde %>%
+  group_by(year, bene_id) %>%
+  summarize(n_claims=n(),
+            n_claims_firstfill=sum(first_fill)) %>%
+  ungroup()
 
-  # join the summary PDE
-  bene = read_tsv(sprintf('raw_data/bene_%i.tsv', year)) %>%
-    lower_names() %>%
-    left_join(regions, by='state') %>%
-    left_join(n_claims, by='bene_id') %>%
-    left_join(n_claims_firstfill, by='bene_id') %>%
-    replace_na(list(n_claims=0, n_claims_firstfill=0)) %>%
-    mutate(year=year, dual=buyin_mo>0) %>%
[…]

### 2017-11-01 ms1/ms1_2process.R

@@ -3,12 +3,8 @@
 # load data
 # NB: I put DC into the South
 
-# bene
-bene = read_tsv('data/bene.tsv') %>%
-  mutate(region=factor(region, levels=c('Northeast', 'West', 'Midwest', 'South')),
-         sex=factor(sex, levels=c('male', 'female')),
-         race=factor(race, levels=c('white', 'black', 'Hispanic', 'other'))) %>%
-  mutate(age=age-1)
+bene = readRDS('data/bene.rds')
+pde = readRDS('data/pde.rds')
 
 output_table = function(x, base) write_tsv(x, sprintf('tables/%s.tsv', base))
 sem = function(x) sd(x) / sqrt(length(x))
@@ -39,13 +35,6 @@ totals = bene %>%
 n_unique_bene = bene$bene_id %>%
   unique %>% length %T>%
   write('tables/n_unique_bene.tsv')
-
-# pde
-pde = read_tsv('data/pde.tsv') %>%
-  mutate(first_fill=fill_num==0) %>%
-  left_join(bene, by=c('year', 'bene_id'))
-
-pde_firstfill = pde %>% filter(first_fill)
 
 # count first fills vs. refills
 pde %>%

### 2017-11-01 ms1/ms1_3dx.R

@@ -14,11 +14,7 @@ top_abx = c('azithromycin', 'ciprofloxacin', 'amoxicillin', 'cephalexin',
 output_table = function(x, base) write_tsv(x, sprintf('tables/%s.tsv', base))
 
 # bene
-bene = read_tsv('data/bene.tsv') %>%
-  mutate(region=factor(region, levels=c('Northeast', 'West', 'Midwest', 'South')),
-         sex=factor(sex, levels=c('male', 'female')),
-         race=factor(race, levels=c('white', 'black', 'Hispanic', 'other'))) %>%
-  mutate(age=age-1)
+bene = readRSD('data/bene.rsd')
 
 read_years = function(years, template) {
   lapply(years, function(y) {
@@ -30,41 +26,24 @@ read_years = function(years, template) {
 
 glm_f = function(df, frmla, ...) eval(substitute(function(df2) glm(frmla, data=df2, ...)))(df)
 sandwich_tidy = function(m) tidy(lmtest::coeftest(m, vcov=sandwich::vcovHC(m, type='HC3')))
-  
+
 frmla = y ~ year + age + n_cc + sex + race + dual + region
 
-# count total dx's and hcpcs counts
-read_years(2011:2014, 'raw_data/dx_hcpcs_count_%i.tsv') %>%
-  output_table('hcpcs_count')
+# swo:
+# need to be able to count E&M here
+dx_from_pde = read_tsv('raw_data/dx_from_pde.tsv')
 
-# count many PDEs have any dx or E&M dx upstream
-read_years(2011:2014, 'raw_data/dx_any_pde_%i.tsv') %>%
-  mutate_at(vars(has_dx, has_em_dx), as.logical) %>%
-  count(year, has_dx, has_em_dx) %>%
+dx_from_pde %>%
+  group_by(year, has_encounter, has_em_encounter) %>%
+  summarize(n_pde=sum(n_pde)) %>% ungroup() %>%
   output_table('pde_with_dx_upstream')
 
 # which dx's contribute to each abx?
-dx_from_pde = read_years(2011:2014, 'raw_data/dx_from_pde_%i.tsv') %>%
-  rename(bene_id=BENE_ID, pde_id=PDE_ID, dx_cat=diagnosis_category)
-
-# get a denominator for total number of PDEs
-pde_denom_2011 = dx_from_pde %>%
-  filter(year==2011) %>%
-  filter(antibiotic %in% top_abx) %>%
[…]

### 2017-11-01 ms1/ms1_display.Rmd
refactoring display

@@
 import::from(knitr, kable)
-import::from(tools, md5sum)
-library(forcats)
 ```
 
-# Abstract
+# Population characteristics
 
-```{r abstract, cache.extra=md5sum('tables/model_abx.tsv')}
-pctize = function(df, denom_name) {
-  for (n_name in Filter(function(x) str_detect(x, '^n_'), names(df))) {
-    pct_name = str_replace(n_name, '^n_', 'pct_')
-    df[[pct_name]] = df[[n_name]] / df[[denom_name]] * 100
-  }
-  df
-}
+```{r pop_chars}
+pop_chars = read_tsv('tables/pop_chars.tsv') %>%
+  select(-sem_age, -sem_n_cc) %>%
+  mutate_at(vars(n_female:n_region_midwest), function(x) x / .$n_bene * 100)
 
-totals = read_tsv('tables/totals.tsv') %>%
-  pctize('n_bene') %>%
-  select(year, n_bene, mean_age, sd_age, mean_n_cc, sd_n_cc,
-         pct_female, pct_white, pct_dual,
-         pct_region_south, pct_region_midwest, pct_region_northeast, pct_region_west)
+trend_p = function(x) tidy(t.test(diff(x)))$p.value
 
-claims_by_abx = read_tsv('tables/claims_by_abx.tsv')
+pop_char_trends = pop_chars %>%
+  select(year, n_bene, mean_age, mean_n_cc, n_female:n_region_midwest) %>%
+  gather('key', 'value', -year) %>%
+  arrange(key, year) %>%
+  group_by(key) %>%
+  summarize(p=trend_p(value)) %>%
+  mutate(adj.p=p.adjust(p, method='BH'))
 
-est2pct = function(x) (exp(x) - 1)*100
+pop_chars %>%
+  mutate(n_bene=prettyNum(n_bene, big.mark=',')) %>%
+  mutate_at(vars(mean_age, sd_age, n_female:n_region_midwest), function(x) format(x, digits=1, nsmall=1)) %>%
+  mutate_at(vars(mean_n_cc, sd_n_cc), function(x) format(x, digits=2, nsmall=2)) %>%
+  gather('key', 'value', -year) %>%
+  spread(year, value) %>%
+  mutate(key=factor(key, levels=names(pop_chars))) %>%
+  arrange(key) %>%
+  kable(caption='Population characteristics')
 
-mult_est = function(df, m) {
-  mutate(df, estimate=m*estimate, std.error=sqrt(m)*std.error)
-}
-
-model_ci = function(df) {
-  df %>%
-    mutate(pct_change=est2pct(estimate),
-           ci=1.96*std.error,
-           ci_low=est2pct(estimate - ci),
-           ci_high=est2pct(estimate + ci),
-           change_str=sprintf('%.2f (%.2f -- %.2f)', pct_change, ci_low, ci_high))
-}
-
-claims_overall = claims_by_abx %>%
-  select(-n_bene) %>%
-  group_by(year) %>%
-  summarize_at(vars(n, cpkp), sum) %>%
-  mutate(antibiotic='overall')
-
-bind_rows(claims_by_abx, claims_overall) %>%
-  filter(antibiotic %in% c('overall', 'azithromycin', 'levofloxacin'), year %in% c(2011, 2014)) %>%
-  select(antibiotic, year, cpkp) %>%
-  spread(year, cpkp) %>%
-  mutate(pct_change=round(100*(`2014`-`2011`)/`2011`, 2)) %>%
-  kable
+pop_char_trends %>%
+  kable(caption='Trends in population characteristics')
 ```
 
-# Beneficiary characteristics
+There are `r scan('tables/n_unique_bene.txt') %>% prettyNum(big.mark=',')` unique beneficiaries.
 
-```{r unique_bene_pde}
-round_millions = function(x) round(x / 1e6, 2)
-pn = function(x) prettyNum(x, big.mark=',')
+**swo**: Double-check that number.
 
-n_unique_bene = scan('tables/n_unique_bene.tsv')
+There are **swo: PDES** and first fills.
 
-claims_by_fill = read_tsv('tables/claims_by_fill.tsv')
-n_pde = sum(claims_by_fill$n)
-n_refills = claims_by_fill %>% filter(!first_fill) %$% n
-```
+# Trends in drug usage
 
-There were `r round_millions(n_unique_bene)` million unique beneficiaries. They had `r round_millions(n_pde)` million PDEs, of which `r round_millions(n_refills)` million were refills.
+## Raw trends
 
-A star indicates that the trend (*t*-test for mean difference between years equal to zero) is significant after multiple hypothesis correction.
-A dot indicates significance ($p < 0.05$) before multiple hypothesis correction.
-Mean age decreased during the study.
+```{r raw_drug_trends}
+claims_by_abx = read_tsv('tables/claims_by_abx.tsv') %>%
+  mutate(abx=fct_inorder(factor(abx)))
 
-```{r bene_char_table}
-trend_p = function(x) tidy(t.test(diff(x)))$p.value
-totals_trends = totals %>%
-  gather('var', 'val', -year) %>%
-  arrange(var, year, val) %>%
-  group_by(var) %>%
-  summarize(p=trend_p(val)) %>%
-  mutate(adj.p=p.adjust(p, method='BH')) %>%
-  mutate(sig=case_when(.$adj.p < 0.05 ~ '*',
-                       .$p < 0.05 ~ '.',
-                       TRUE ~ ''))
+abx_levels = levels(claims_by_abx$abx)
 
-totals %>%
-  pctize('n_bene') %>%
-  select(year, n_bene, mean_age, sd_age, mean_n_cc, sd_n_cc,
-         pct_female, pct_white, pct_dual,
-         pct_region_south, pct_region_midwest, pct_region_northeast, pct_region_west) %>%
-  mutate_at(vars(starts_with('pct_')), function(x) round(x, digits=1)) %>%
-  mutate(n_bene=prettyNum(n_bene, big.mark=','),
-         mean_age=round(mean_age, digits=1)) %>%
-  mutate_at(vars(mean_n_cc, sd_age, sd_n_cc),
-            function(x) round(x, digits=2)) %>% 
-  gather('var', 'val', -year) %>% spread(year, val) %>% # transpose
-  left_join(totals_trends %>% select(var, sig), by='var') %>%
-  kable
-```
-
-# Patterns of consumption
-
-All consumption rates are reported as claims per thousand people per year (CPKP).
-
-## Claims by abx
-
-The top 10 antibiotics account for most of the claims:
-
-```{r top_10_fraction}
 claims_by_abx %>%
-  group_by(antibiotic) %>%
-  summarize(n=sum(n)) %>%
-  arrange(desc(n)) %>%
-  mutate(f=n/sum(n), cumf=cumsum(f)) %>%
-  head(10) %>%
-  kable
-```
-
-```{r claims_by_abx}
-top_abx = claims_by_abx %>%
-  group_by(antibiotic) %>%
-  summarize(cpkp=mean(cpkp)) %>%
-  arrange(desc(cpkp)) %$%
-  head(antibiotic, 10)
-
-claims_by_top_abx = claims_by_abx %>%
-  bind_rows(claims_overall) %>%
-  mutate(antibiotic=fct_relevel(fct_other(antibiotic, keep=c('overall', top_abx)),
-                                c('overall', top_abx, 'Other'))) %>%
-  group_by(year, antibiotic) %>%
-  summarize_at(vars(n, cpkp), sum) %>% ungroup()
-```
-
-```{r claims_by_top_abx_figs}
-claims_by_top_abx %>%
-  filter(antibiotic=='overall') %>%
+  filter(abx=='overall') %>%
   ggplot(aes(x=year, y=cpkp)) +
@@
 
-claims_by_top_abx %>%
-  filter(antibiotic %in% c('azithromycin', 'ciprofloxacin', 'amoxicillin', 'cephalexin', 'trimethoprim/sulfamethoxazole', 'levofloxacin')) %>%
-  ggplot(aes(x=year, y=cpkp, color=antibiotic)) +
+claims_by_abx %>%
+  filter(abx %in% c('azithromycin', 'ciprofloxacin', 'amoxicillin', 'cephalexin', 'tmpsmx', 'levofloxacin')) %>%
+  ggplot(aes(x=year, y=cpkp, color=abx)) +
   geom_point() + geom_line() +
@@
 
-claims_by_top_abx %>%
-  filter(antibiotic %in% c('amoxicillin/clavulanate', 'doxycycline', 'nitrofurantoin', 'clindamycin')) %>%
-  ggplot(aes(x=year, y=cpkp, color=antibiotic)) +
+claims_by_abx %>%
+  filter(abx %in% c('amoxclav', 'doxycycline', 'nitrofurantoin', 'clindamycin')) %>%
+  ggplot(aes(x=year, y=cpkp, color=abx)) +
   geom_point() + geom_line() +
   scale_y_continuous(limits=c(0, 100), breaks=c(0, 50, 100)) +
-  #ylim(0, 140) +
   theme_minimal() +
@@
 
-## Claims by demography
-
-```{r claims_by_demography}
-change_demog = read_tsv('tables/model_overall_demography.tsv') %>%
-  mutate(estimate=3*estimate,
-         std.error=sqrt(3)*std.error) %>%
-  model_ci() %>%
-  filter(term=='year') %>%
-  select(name, adj_change=change_str)
-
-claims_by = function(x) {
-  read_tsv(str_interp("tables/claims_by_${x}.tsv")) %>%
-    rename_(.dots=list(name=x)) %>%
-    select(name, year, cpkp) %>%
-    filter(year %in% c(2011, 2014)) %>%
-    spread(year, cpkp) %>%
-    mutate(pct_change=round((`2014`-`2011`)/`2011`*100, 2)) %>%
-    mutate_at(vars(`2011`, `2014`), round) %>%
-    mutate(group=x)
-}
-
-# do claims by age group separate
-# this could be baked into the analysis step
-claims_by_age = read_tsv('tables/claims_by_age.tsv') %>%
-  filter(year %in% c(2011, 2014)) %>%
-  mutate(name=case_when(between(.$age, 65, 75) ~ 'age65_75',
-                        between(.$age, 76, 85) ~ 'age76_85',
-                        between(.$age, 86, 95) ~ 'age86_95',
-                        .$age >= 96 ~ 'age96_')) %>%
-  group_by(year, name) %>%
-  summarize_at(vars(n_bene, n_claims), sum) %>%
-  ungroup() %>%
-  mutate(cpkp=1000*n_claims/n_bene) %>%
-  select(year, name, cpkp) %>%
-  spread(year, cpkp) %>%
-  mutate(pct_change=round((`2014`-`2011`)/`2011`*100, 2)) %>%
-  mutate_at(vars(`2011`, `2014`), round) %>%
-  mutate(group='age')
-
-# use a common interface for all the others
-bind_rows(claims_by_age,
-          claims_by('sex'),
-          claims_by('race'),
-          claims_by('region')) %>%
-  left_join(change_demog, by='name') %>%
-  select(group, name, `2011`, `2014`, pct_change, adj_change) %>%
-  kable
-```
-
-That last table could also use an "overall" row:
-
-```{r}
-claims_overall %>%
-  filter(year %in% c(2011, 2014)) %>%
-  mutate(cpkp=round(cpkp)) %>%
-  kable
-```
-
-# Adjusted consumption changes by drug
-
-Adjusted changes were computed using a Poisson regression predicting claims per capita.
-The unit of analysis were individual beneficiaries in each year.
-The covariates in the regression were age, number of chronic conditions, dual eligibility, sex, race (white or not), and region.
-
-## By drug
-
-### Trends in appropriateness
-
-Just by eye, it looks like the changes in appropriateness are small.
-
-```{r tier_graph}
-pde_tier_table = dx_pde_counts %>%
-  filter(has_dx_em) %>%
-
-pde_tier_table = read_tsv('raw_data/dx_pde_counts.tsv') %>%
-  group_by(year,
-
-pde_tier_table = read_tsv('tables/pde_approp_table.tsv') %>%
-  group_by(year, antibiotic) %>%
-  mutate(f=n/sum(n)) %>%
-  ungroup()
-
-pde_tier_table %>%
-  filter(antibiotic %in% top_abx) %>%
-  ggplot(aes(x=factor(year), y=f, fill=factor(tier))) +
-  geom_bar(stat='identity') +
-  facet_grid(~antibiotic) +
-  theme_classic()
-```
+## Trends in appropriateness
 
 ```{r approp_trends}
-raw_adjusted_changes = read_tsv('tables/model_abx.tsv') %>%
-  rename(antibiotic=name) %>%
-  filter(term=='year') %>%
-  mutate(estimate=3*estimate,
-         std.error=sqrt(3)*std.error)
-
-# fraction inappropriate in 2011
-f_i = pde_tier_table %>%
-  filter(year==2011) %>%
-  mutate(inapp=tier==3) %>%
-  group_by(inapp) %>%
-  summarize(n=sum(n)) %>%
-  mutate(f=n/sum(n)) %>%
-  filter(inapp) %$%
-  f
-
 # fold change in consumption
-pde_trend = raw_adjusted_changes %>%
-  filter(antibiotic=='overall') %>%
-  # multiplications by 3 done above
-  mutate(ci=1.96*std.error,
-         rr=exp(estimate),
-         rr_lo=exp(estimate-ci),
-         rr_hi=exp(estimate+ci))
+pde_trend = drug_trends %>%
+  filter(abx=='overall') %>%
+  mutate(rr=exp(estimate),
+         rr_lo=exp(ci_lo),
+         rr_hi=exp(ci_hi))
 
@@
 
-risk_ratio_f = function(odr, p) odr / (1 + p * (odr - 1))
+# fraction inappropriate in 2011
+risk_ratio_f = function(or, p) or / (1 + p * (or - 1))
+f_i = scan('tables/inapprop_risk_2011.txt')
 
-pde_inapprop_trend = read_tsv('tables/pde_inapprop_trend.tsv') %>%
+inapprop_trend = read_tsv('tables/inapprop_trend.tsv') %>%
   filter(term=='year') %>%
   mutate(estimate=3*estimate,
-         std.error=sqrt(3)*std.error,
-         ci=1.96*std.error,
-         or=exp(estimate),
-         or_lo=exp(estimate-ci),
-         or_hi=exp(estimate+ci))
+         std.error=sqrt(3)*std.error) %>%
+  mutate(ci=1.96*std.error,
+         ci_lo=estimate-ci,
+         ci_hi=estimate+ci) %>%
+  mutate(or   =exp(estimate),
+         or_lo=exp(ci_lo),
+         or_hi=exp(ci_hi)) %>%
+  mutate(rr   =risk_ratio_f(or, f_i),
+         rr_lo=risk_ratio_f(or_lo, f_i),
+         rr_hi=risk_ratio_f(or_hi, f_i))
 
-d_f = risk_ratio_f(pde_inapprop_trend$or, f_i)
-d_f_lo = risk_ratio_f(pde_inapprop_trend$or_lo, f_i)
-d_f_hi = risk_ratio_f(pde_inapprop_trend$or_hi, f_i)
+d_f = inapprop_trend$rr
+d_f_lo = inapprop_trend$rr_lo
+d_f_hi = inapprop_trend$rr_hi
 
@@
 
-inapprop_changes = data_frame(
-  antibiotic=c('overall_inapp', 'overall_app'),
-  pct_change=c(d_c * d_f, (1.0 - d_f * f_i) / (1.0 - f_i) * d_c),
-  ci_low=c(d_c_lo * d_f_lo, (1.0 - d_f_hi * f_i) / (1.0 - f_i) * d_c_lo),
-  ci_high=c(d_c_hi * d_f_hi, (1.0 - d_f_lo * f_i) / (1.0 - f_i) * d_c_hi)
+inapprop_trends = data_frame(
+  abx=c('overall_inapp', 'overall_app'),
+  y   =c(d_c    * d_f,    (1.0 - d_f    * f_i) / (1.0 - f_i) * d_c),
+  ymin=c(d_c_lo * d_f_lo, (1.0 - d_f_hi * f_i) / (1.0 - f_i) * d_c_lo),
+  ymax=c(d_c_hi * d_f_hi, (1.0 - d_f_lo * f_i) / (1.0 - f_i) * d_c_hi)
 ) %>%
   mutate_if(is.numeric, rr2pct)
+
+kable(inapprop_changes)
 ```
 
-### All fills
+## Adjusted trends
 
-Bars show adjusted trends in usage, including refills. The dots shows unadjusted trends.
+```{r drug_trends}
+est2pct = function(x) (exp(x) - 1)*100
 
-```{r adjusted_changes}
-adjusted_changes = raw_adjusted_changes %>%
-  model_ci() %>%
-  arrange(desc(pct_change)) %>%
-  select(antibiotic, pct_change, ci_low, ci_high)
+drug_trends = read_tsv('tables/trend_models.tsv') %>%
+  filter(term=='year') %>%
+  arrange(estimate) %>%
+  mutate(estimate=3*estimate,
+         std.error=sqrt(3)*std.error) %>%
+  mutate(ci=1.96*std.error,
+         ci_lo=estimate-ci,
+         ci_hi=estimate+ci) %>%
+  mutate(y=est2pct(estimate),
+         ymin=est2pct(ci_lo),
+         ymax=est2pct(ci_hi)) %>%
+  mutate(change=sprintf('%.2f (%.2f -- %.2f)', y, ymin, ymax))
 
-adjusted_change_levels = adjusted_changes %>%
-  filter(antibiotic != 'overall') %$%
-  { c('overall', 'overall_app', 'overall_inapp', antibiotic) }
+drug_trends %>%
+  select(abx, y, ymin, ymax) %>%
+  mutate_if(is.numeric, function(x) format(x, digits=2, nsmall=2)) %>%
+  kable(caption='Drug trends: Percent change from 2011 to 2014')
 
-unadjusted_changes = bind_rows(claims_by_abx, claims_overall) %>%
-  filter(antibiotic %in% c('overall', top_abx), year %in% c(2011, 2014)) %>%
-  select(antibiotic, year, cpkp) %>%
-  spread(year, cpkp) %>%
-  mutate(change=`2014`-`2011`,
-         fold_change=change/`2011`,
-         pct_change=100*fold_change) %>%
-  select(antibiotic, pct_change)
+abx_trends_levels = drug_trends %>%
+  filter(abx != 'overall') %$%
+  { c('overall', 'overall_app', 'overall_inapp', rev(abx)) }
 
-adjusted_changes %>%
-  bind_rows(inapprop_changes) %>%
-  mutate(antibiotic=as.character(antibiotic)) %>%
-  mutate(antibiotic=factor(antibiotic, levels=rev(adjusted_change_levels))) %>%
-  ggplot(aes(x=antibiotic, y=pct_change, ymin=ci_low, ymax=ci_high)) +
+drug_and_inapprop_trends = drug_trends %>%
+  select(abx, y, ymin, ymax) %>%
+  bind_rows(inapprop_trends) %>%
+  mutate(abx=factor(abx, levels=rev(abx_trends_levels)))
+
+drug_and_inapprop_trends %>%
+  ggplot(aes(x=abx, y=y, ymin=ymin, ymax=ymax)) +
   geom_bar(stat='identity', fill='white', color='black') +
@@
         axis.text = element_text(size=16))
+
+drug_and_inapprop_trends %>%
+  arrange(desc(abx)) %>%
+  kable(caption='Percent changes; same data as in figure')
 ```
 
-The bars show the adjusted change in consumption.
+# Supplement
 
-### Adjusted changes in usage table
+## Claims by demography
 
-```{r adjusted_changes_table}
-adjusted_changes %>%
-  bind_rows(inapprop_changes) %>%
-  mutate(antibiotic=factor(antibiotic, levels=adjusted_change_levels)) %>%
-  arrange(antibiotic) %>%
+```{r claims_by_demography}
+claims_by = function(by) {
+  read_tsv(str_interp("tables/claims_by_${by}.tsv")) %>%
+    mutate(group_type=by) %>%
+    rename(group=!!rlang::sym(by)) %>%
+    filter(year %in% c(2011, 2014)) %>%
+    select(group_type, group, year, cpkp) %>%
+    spread(year, cpkp) %>%
+    mutate_at(vars(`2011`, `2014`), round)
+}
+
+trends_by_demography = read_tsv('tables/demography_models.tsv') %>%
+  filter(term=='year') %>%
+  mutate(estimate=3*estimate,
+         std.error=sqrt(3)*std.error) %>%
+  mutate(ci=1.96*std.error,
+         ci_lo=estimate-ci,
+         ci_hi=estimate+ci) %>%
+  mutate(y=est2pct(estimate),
+         ymin=est2pct(ci_lo),
+         ymax=est2pct(ci_hi)) %>%
+  mutate(change=sprintf('%.2f (%.2f -- %.2f)', y, ymin, ymax)) %>%
+  select(group_type, group, change)
+
+# overall trend row
+overall_change = drug_trends %>% filter(abx=='overall') %>% pull(change)
+overall_row = claims_by_abx %>%
+  filter(abx=='overall', year %in% c(2011, 2014)) %>%
+  mutate(cpkp=round(cpkp)) %>%
+  spread(year, cpkp) %>%
+  mutate(group_type='overall', group='overall', change=overall_change) %>%
+  select(group_type, group, `2011`, `2014`, change)
+
+bind_rows(
+  claims_by('age_group'),
+  claims_by('sex'),
+  claims_by('race'),
+  claims_by('region')
+) %>%
+  left_join(trends_by_demography, by=c('group', 'group_type')) %>%
+  bind_rows(overall_row, .) %>%
   kable
-
-adjusted_az_decrease = adjusted_changes %>%
-  filter(antibiotic=='azithromycin') %>%
-  pull(pct_change) %>%
-  round(2)
-
-adjusted_levo_increase = adjusted_changes %>%
-  filter(antibiotic=='levofloxacin') %>%
-  pull(pct_change) %>%
-  round(2)
 ```
 
-Consumption of azithromycin fell by `r adjusted_az_decrease`% over the entire study period. Levofloxacin consumption increased by `r adjusted_levo_increase`% over the entire study period.
+# ***Earlier content
+
+### All fills
 

### 2017-11-01 ms1/ms1_display.Rmd
refactor display

@@
 
-**swo**: Double-check that number.
-
-There are **swo: PDES** and first fills.
+```{r pde_counts}
+read_tsv('raw_data/firstfill.tsv') %>%
+  group_by(firstfill) %>%
+  summarize(n_pde=sum(n_pde)) %>%
+  #arrange(desc(firstfill)) %>%
+  mutate(cum_n=prettyNum(cumsum(n_pde), big.mark=',')) %>%
+  kable(caption='PDEs by firstfill (or refill)')
+```
 
@@
 ## Trends in appropriateness
+
+```{r drug_trends}
+est2pct = function(x) (exp(x) - 1)*100
+
+drug_trends = read_tsv('tables/trend_models.tsv') %>%
+  filter(term=='year') %>%
+  arrange(estimate) %>%
+  mutate(estimate=3*estimate,
+         std.error=sqrt(3)*std.error) %>%
+  mutate(ci=1.96*std.error,
+         ci_lo=estimate-ci,
+         ci_hi=estimate+ci) %>%
+  mutate(y=est2pct(estimate),
+         ymin=est2pct(ci_lo),
+         ymax=est2pct(ci_hi)) %>%
+  mutate(change=sprintf('%.2f (%.2f -- %.2f)', y, ymin, ymax))
+```
 
@@
 
-kable(inapprop_changes)
+kable(inapprop_trends, caption='Percent changes in in/appropriate prescribing')
 ```
@@
 
-```{r drug_trends}
-est2pct = function(x) (exp(x) - 1)*100
-
-drug_trends = read_tsv('tables/trend_models.tsv') %>%
-  filter(term=='year') %>%
-  arrange(estimate) %>%
-  mutate(estimate=3*estimate,
-         std.error=sqrt(3)*std.error) %>%
-  mutate(ci=1.96*std.error,
-         ci_lo=estimate-ci,
-         ci_hi=estimate+ci) %>%
-  mutate(y=est2pct(estimate),
-         ymin=est2pct(ci_lo),
-         ymax=est2pct(ci_hi)) %>%
-  mutate(change=sprintf('%.2f (%.2f -- %.2f)', y, ymin, ymax))
-
+```{r drug_trends_display}
 drug_trends %>%
@@
   kable(caption='Percent changes; same data as in figure')
+```
+
+# Prescribing practice
+
+```{r prescribing_practice}
+dx_counts_2011 = read_tsv('raw_data/dx_counts.tsv') %>%
+  left_join(select(pop_chars, year, n_bene), by='year') %>%
+  mutate(dkpy=round(n_encounter/n_bene*1000)) %>%
+  filter(year==2011) %>%
+  select(dx_cat, dkpy)
+
+dx_from_pde_2011 = read_tsv('raw_data/dx_from_pde.tsv') %>%
+  filter(year==2011) %>%
+  mutate(f_with_dx=n_pde_with_encounter/n_pde_with_any_encounter) %>%
+  select(antibiotic, dx_cat, year, f_with_dx)
+
+dx_to_pde_table = read_tsv('tables/dx_to_pde_table.tsv') %>%
+  select(-n_encounters) %>%
+  filter(year %in% c(2011, 2014)) %>%
+  gather('abx', 'f_with_abx', azithromycin:clindamycin)
+
+# swo: convert ORs to RRs!
+rr2pct = function(x) (x - 1) * 100
+
+dx_to_pde_trends = read_tsv('tables/dx_to_pde_trends.tsv') %>%
+  filter(term=='year') %>%
+  mutate(estimate=3*estimate,
+         std.error=sqrt(3)*std.error) %>%
+  mutate(ci=1.96*std.error,
+         ci_lo=estimate-ci,
+         ci_hi=estimate+ci) %>%
+  mutate(or=   exp(estimate),
+         or_lo=exp(ci_lo),
+         or_hi=exp(ci_hi)) %>%
+  left_join(filter(dx_to_pde_table, year==2011), by=c('dx_cat', 'abx')) %>%
+  mutate(rr=   risk_ratio_f(or,    f_with_abx),
+         rr_lo=risk_ratio_f(or_lo, f_with_abx),
+         rr_hi=risk_ratio_f(or_hi, f_with_abx)) %>%
+  mutate(y=rr2pct(rr),
+         ymin=rr2pct(rr_lo),
+         ymax=rr2pct(rr_hi)) %>%
+  mutate(change=sprintf('%.2f (%.2f -- %.2f)', y, ymin, ymax)) %>%
+  select(dx_cat, abx, change)
+```
+
+```{r combined_table}
+format2 = function(x) format(round(x, 2), digits=2, nsmall=2)
+
+part1 = dx_counts_2011 %>%
+  mutate(dkpy=as.character(dkpy)) %>%
+  spread(dx_cat, dkpy) %>%
+  mutate(abx='dx', key1='dkpy', key2='')
+
+part2 = dx_from_pde_2011 %>%
+  # throw out some rare drugs
+  filter(!is.na(dx_cat)) %>%
+  mutate(f_with_dx=format2(f_with_dx*100)) %>%
+  spread(dx_cat, f_with_dx) %>%
+  mutate(key1='pde2dx', key2=as.character(year)) %>% rename(abx=antibiotic)
+
+part3 = dx_to_pde_table %>%
+  mutate(f_with_abx=format2(f_with_abx*100)) %>%
+  spread(dx_cat, f_with_abx) %>%
+  mutate(key1='dx2pde', key2=as.character(year))
+
+part4 = dx_to_pde_trends %>%
+  spread(dx_cat, change) %>%
+  mutate(key1='dx2pde_trend', key2='')
+
+combined_table = bind_rows(part1, part2, part3, part4) %>%
+  mutate(abx=factor(abx, levels=c('dx', abx_levels)),
+         key1=factor(key1, levels=c('dkpy', 'pde2dx', 'dx2pde', 'dx2pde_trend'))) %>%
+  arrange(abx, key1, key2)
+```
+
+## Respiratory
+
+```{r respiratory}
+combined_table %>%
+  select(abx, key1, key2, pneumonia, sinusitis, uri, bronchitis) %>%
+  filter(abx %in% c('dx', 'azithromycin', 'levofloxacin', 'amoxclav')) %>%
+  kable
 ```
@@
 
-# ***Earlier content
+## Other respiratory
 
-### All fills
-
-# Prescribing practice
-
-## Methodology
-
-First, I looked at all encounters recorded in the Carrier Line File and the Outpatient file. A large fraction, I think 78%, of PDEs had an encounter 0-7 days before.
-
-Next, I limited by analysis to encounters who HCPCS code indicated an E&M visit, and I grouped diagnostic codes into categories using Fleming-Dutra eTable 2.
-I determined the trends in the prevalence of these diagnosis categories by counting the diagnoses per beneficiary per year and by using Poisson regressions that account for the unusual covariates.
-
-I analyzed trends in prescribing practice by computing the fraction of encounters in each diagnosis category that were followed by PDEs for each drug.
-I computed adjusted trends using logistic regression.
-
-Finally, I considered only those PDEs that are downstream of those E&M diagnoses. For each antibiotic, I computed the fraction of PDEs that have each diagnosis category upstream.
-
-To compute appropriateness, I classified each PDE with the maximum appropriateness of all its linked diagnoses.
-I put PDEs with no associated infectious disease (i.e., FD) diagnoses into their own, lower tier.
-
-## Numbers of diagnoses
-
-The fraction of diagnoses that have E&M HCPCS codes:
-
-```{r hcpcs_count}
-encounter_counts = read_tsv('raw_data/dx_cat_count.tsv') %>%
-  summarize(n_encounters=sum(n_encounters),
-            n_em_encounters=sum(n_encounters[hcpcs_em]),
-            f_em_encounters=weighted.mean(hcpcs_em, weights=n_encounters))
-
-hcpcs_count %>% kable
-```
-
-This equals a total of `r prettyNum(sum(encounter_counts$n_encounters), big.mark=',')` encounters.
-
-The fraction of first-fill PDEs that have upstream:
-
-```{r pde_with_dx_upstream}
-read_tsv('tables/pde_with_dx_upstream.tsv') %>%
-  group_by(has_dx, has_em_dx) %>%
-  summarize(n=sum(n)) %>%
-  ungroup() %>%
-  mutate(f=n/sum(n), cumf=cumsum(f), cumf1=1-cumf) %>%
+```{r other_respiratory}
+combined_table %>%
+  select(abx, key1, key2, pharyngitis, asthma, other_resp) %>%
+  filter(abx %in% c('dx', 'azithromycin', 'levofloxacin', 'amoxclav')) %>%
   kable
@@
 
-## Common diagnoses by drug
+## UTI
 
-For each drug, I asked what fraction of PDEs are associated with each diagnosis.
-This gives you a sense that, say, the diagnoses that give rise to azithromycin are bronchitis, URI, etc.
-The values are percents in 2011.
-A row can add up to more than 100% because one PDE can be associated with multiple diagnoses.
-
-```{r dx_from_pde}
-dx_from_pde = read_tsv('tables/dx_from_pde.tsv') %>% 
-  spread(present, n) %>%
-  mutate(pct=`TRUE`/(`TRUE` + `FALSE`)*100 %>% round(2)) %>%
-  select(antibiotic, dx_cat, pct) %>%
-  spread(dx_cat, pct)
-
-dx_from_pde %>% kable
-```
-
-## Trends in prescribing practice
-
-For each diagnosis and each drug, I ask what fraction of diagnoses are followed by a PDE for that drug in 2011 and 2014. I compute the percent change in those fractions.
-
-Second, I do logistic regressions.
-The unit of analysis is the diagnosis.
-The outcome is whether a PDE for that drug follows a diagnosis in that category.
-The main predictor is year; covariates are as above.
-The adjusted percent change is between 2011 and 2014.
-(I computed this adjusted change, which is a relative risk, using the odds ratio from the regression and the mean probability of PDE for that diagnosis in 2011.
-This strictly incorrect: the mean relative risk is the mean of the risks computed using the predicted probabilities for each 2011 diagnosis, *not* the relative risk computed using the mean probability in 2011.)
-
-```{r dx_to_pde}
-dx_to_pde_unadj = read_tsv('tables/dx_to_pde_table.tsv') %>%
-  filter(year %in% c(2011, 2014)) %>%
-  select(dx_cat, antibiotic, year, f=f_encounters_with_abx) %>%
-  spread(year, f)
-
-dx_to_pde_adj = read_tsv('tables/dx_to_pde_trends.tsv') %>%
-  filter(term=='year') %>%
-  mult_est(3) %>%
-  mutate(ci=1.96*std.error,
-         ci_low=estimate-ci,
-         ci_high=estimate+ci) %>%
-  select(dx_cat, antibiotic, estimate, ci_low, ci_high) %>%
-  arrange(dx_cat, antibiotic)
-
-dx_to_pde = inner_join(dx_to_pde_unadj,
-                       dx_to_pde_adj,
-                       by=c('dx_cat', 'antibiotic')) %>%
-  # convert model estimates to odds ratios, then to risk ratios,
-  # then to adjusted fractional changes, then to percentages
-  mutate(or=exp(estimate),
-         or_low=exp(ci_low),
-         or_high=exp(ci_high),
-         rr=risk_ratio_f(or, `2011`),
-         rr_low=risk_ratio_f(or_low, `2011`),
-         rr_high=risk_ratio_f(or_high, `2011`),
-         adj_change=rr-1,
-         adj_change_low=rr_low-1,
-         adj_change_high=rr_high-1,
-         pct_change=(`2014`-`2011`)/`2011`) %>%
-  mutate_if(is.numeric, function(x) round(x*100, 2)) %>%
-  mutate(adj_change=sprintf('%.2f (%.2f -- %.2f)', adj_change, adj_change_low, adj_change_high)) %>%
-  select(dx_cat, antibiotic, `2011`, `2014`, pct_change, adj_change) %T>%
+```{r uti}
+combined_table %>%
+  select(abx, key1, key2, uti, uti_t3) %>%
+  filter(abx %in% c('dx', 'ciprofloxacin', 'tmpsmx', 'nitrofurantoin')) %>%
   kable
@@
 
-## Combined tables
+## SSTI
 
-```{r}
-x = read_tsv('tables/dx_cat_counts.tsv') %>%
-  mutate(value=n_dx/n_bene*1000) %>%
-  group_by(dx_cat) %>%
-  summarize(value=mean(value)) %>%
-  # convert to character to fit with other parts
-  mutate(value=as.character(round(value))) %>%
-  spread(dx_cat, value) %>%
-  mutate(antibiotic='dx', key='mean_per_1kyr')
-
-y = dx_from_pde %>%
-  mutate(key='dx_from_pde') %>%
-  mutate_if(is.numeric, function(x) as.character(round(x, 2)))
-
-z = dx_to_pde %>%
-  gather('key', 'value', -dx_cat, -antibiotic) %>%
-  spread(dx_cat, value)
-
-combined_table = bind_rows(x, y, z) %>%
-  mutate(antibiotic=factor(antibiotic, levels=c('dx', top_abx)),
-         key=factor(key, levels=c('mean_per_1kyr', 'dx_from_pde', '2011', '2014', 'pct_change', 'adj_change'))) %>%
-  arrange(antibiotic, key)
-```
-
-### Respiratory stuff
-
-All respiratory diagnoses fell, bronchitis the most sharply. Azithromycin prescribing fell for all diagnoses, while levofloxacin, amox, and amox/clav prescribing rose for all diagnoses.
-
-```{r respiratory}
+```{r ssti}
 combined_table %>%
-  filter(antibiotic %in% c('dx', 'azithromycin', 'levofloxacin', 'amoxicillin', 'amoxicillin/clavulanate')) %>%
-  select(antibiotic, key, asthma, bronchitis, other_resp, pharyngitis, pneumonia, sinusitis, uri) %>%
+  select(abx, key1, key2, ssti, ssti_t3) %>%
+  filter(abx %in% c('dx', 'cephalexin', 'tmpsmx', 'clindamycin')) %>%
   kable
 ```
-
-### UTIs
-
-UTIs declined somewhat in prevalence. Ciprofloxacin prescribing remained nearly constant, nitrofurantion fell 3%, and Bactrim rose 4-6%.
-
-```{r uti}
-combined_table %>%
-  filter(antibiotic %in% c('dx', 'ciprofloxacin', 'trimethoprim/sulfamethoxazole', 'nitrofurantoin')) %>%
-  select(antibiotic, key, uti, uti_t3) %>%
-  kable
-```
-
-### SSTI
-
-Although skin and soft tissue infections declined in prevalence, prescribing of clindamycin, cephalexin, Bactrim, amox, and amox/clav increased, with one exception: amox/clav use for Tier 2 SSTIs stay nearly constant.
-Clindamycin was the biggest winner.
-
-```{r ssti}
-# swo: I removed amox and amox/clav because they mostly
-# showed trends in Tier 3, which is confusing
-combined_table %>%
-  filter(antibiotic %in% c('dx', 'clindamycin', 'cephalexin', 'trimethoprim/sulfamethoxazole')) %>%
-  select(antibiotic, key, ssti, ssti_t3) %>%
-  kable
-```
-
-### GI
-
-GI indications became less prevalent, and cipro prescribing fell 1-2%.
-Only a small amount of cipro is used for GI indications: compare the 5-6% of cipro PDEs preceded by GI indications with the 14-20% preceded by UTIs.
-
-```{r gi}
-combined_table %>%
-  filter(antibiotic %in% c('dx', 'ciprofloxacin')) %>%
-  select(antibiotic, key, gi, gi_t3) %>%
-  kable
-```

### 2017-11-01 ms1/ms1_analyze.R

@@ -10,6 +10,9 @@ bene = read_tsv('raw_data/bene.tsv') %>%
   rename(bene_id=BENE_ID) %>%
   left_join(regions, by='state')
 
+n_unique_bene = length(unique(bene$bene_id)) %T>%
+  write('tables/n_unique_bene.txt')
+
 pde = read_tsv('raw_data/bene_pde.tsv') %>%
   rename(bene_id=BENE_ID) %>%
   rename(overall=n_pde) %>%
@@ -37,9 +40,6 @@ pop_chars = bene %>%
             n_region_northeast=sum(region=='Northeast'),
             n_region_midwest=sum(region=='Midwest')) %T>%
   output_table('pop_chars')
-
-n_unique_bene = length(unique(bene$bene_id)) %T>%
-  write('tables/n_unique_bene.txt')
 
 claims_by_abx = pde %>%
   select(year, overall:clindamycin) %>%

### 2017-11-08 ms1/ms1_display.Rmd
display: new no refills

@@
 pop_chars = read_tsv('tables/pop_chars.tsv') %>%
-  select(-sem_age, -sem_n_cc) %>%
   mutate_at(vars(n_female:n_region_midwest), function(x) x / .$n_bene * 100)
@@
 
-```{r pde_counts}
-read_tsv('raw_data/firstfill.tsv') %>%
-  group_by(firstfill) %>%
-  summarize(n_pde=sum(n_pde)) %>%
-  #arrange(desc(firstfill)) %>%
-  mutate(cum_n=prettyNum(cumsum(n_pde), big.mark=',')) %>%
-  kable(caption='PDEs by firstfill (or refill)')
+```{r claims_by_fill}
+claims_by_fill = read_tsv('tables/claims_by_fill.tsv') %>%
+  mutate(f_refill=refills/overall_wrefill)
+
+claims_by_fill %>%
+  kable(caption='PDEs by refill')
+
+#  mutate(cum_n=prettyNum(cumsum(n_pde), big.mark=',')) %>%
 ```
@@
 claims_by_abx %>%
-  filter(abx=='overall') %>%
-  ggplot(aes(x=year, y=cpkp)) +
+  filter(abx %in% c('overall', 'overall_approp', 'overall_inapprop')) %>%
+  ggplot(aes(x=year, y=cpkp, color=abx)) +
   geom_point() + geom_line() +
-  scale_y_continuous(limits=c(975, 1525), breaks=c(1000, 1125, 1250, 1375, 1500)) +
+  #scale_y_continuous(limits=c(0, 1525), breaks=c(1000, 1125, 1250, 1375, 1500)) +
+  scale_y_continuous(limits=c(0, 1525), breaks=c(0, 250, 500, 750, 1000, 1250, 1500)) +
   theme_minimal() +
@@
 
-## Trends in appropriateness
+## Adjusted trends
 
@@
 
-```{r approp_trends}
-# fold change in consumption
-pde_trend = drug_trends %>%
-  filter(abx=='overall') %>%
-  mutate(rr=exp(estimate),
-         rr_lo=exp(ci_lo),
-         rr_hi=exp(ci_hi))
+```{r drug_trends_display}
+format_p = function(p) {
+  if_else(p < 0.001, '< 0.001', as.character(round(p, 3)))
+}
 
-d_c = pde_trend$rr
-d_c_lo = pde_trend$rr_lo
-d_c_hi = pde_trend$rr_hi
-
-# fraction inappropriate in 2011
-risk_ratio_f = function(or, p) or / (1 + p * (or - 1))
-f_i = scan('tables/inapprop_risk_2011.txt')
-
-inapprop_trend = read_tsv('tables/inapprop_trend.tsv') %>%
-  filter(term=='year') %>%
-  mutate(estimate=3*estimate,
-         std.error=sqrt(3)*std.error) %>%
-  mutate(ci=1.96*std.error,
-         ci_lo=estimate-ci,
-         ci_hi=estimate+ci) %>%
-  mutate(or   =exp(estimate),
-         or_lo=exp(ci_lo),
-         or_hi=exp(ci_hi)) %>%
-  mutate(rr   =risk_ratio_f(or, f_i),
-         rr_lo=risk_ratio_f(or_lo, f_i),
-         rr_hi=risk_ratio_f(or_hi, f_i))
-
-d_f = inapprop_trend$rr
-d_f_lo = inapprop_trend$rr_lo
-d_f_hi = inapprop_trend$rr_hi
-
-rr2pct = function(x) round((x - 1) * 100, 2)
-
-inapprop_trends = data_frame(
-  abx=c('overall_inapp', 'overall_app'),
-  y   =c(d_c    * d_f,    (1.0 - d_f    * f_i) / (1.0 - f_i) * d_c),
-  ymin=c(d_c_lo * d_f_lo, (1.0 - d_f_hi * f_i) / (1.0 - f_i) * d_c_lo),
-  ymax=c(d_c_hi * d_f_hi, (1.0 - d_f_lo * f_i) / (1.0 - f_i) * d_c_hi)
-) %>%
-  mutate_if(is.numeric, rr2pct)
-
-kable(inapprop_trends, caption='Percent changes in in/appropriate prescribing')
-```
-
-## Adjusted trends
-
-```{r drug_trends_display}
 drug_trends %>%
-  select(abx, y, ymin, ymax) %>%
-  mutate_if(is.numeric, function(x) format(x, digits=2, nsmall=2)) %>%
+  filter(abx != 'overall_wrefill') %>%
+  mutate(adj.p=p.adjust(p.value, method='BH')) %>%
+  select(abx, y, ymin, ymax, p.value) %>%
+  mutate_at(vars(y, ymin, ymax), function(x) format(x, digits=2, nsmall=2)) %>%
+  mutate(p.value=format_p(p.value)) %>%
   kable(caption='Drug trends: Percent change from 2011 to 2014')
 
-abx_trends_levels = drug_trends %>%
-  filter(abx != 'overall') %$%
-  { c('overall', 'overall_app', 'overall_inapp', rev(abx)) }
-
-drug_and_inapprop_trends = drug_trends %>%
-  select(abx, y, ymin, ymax) %>%
-  bind_rows(inapprop_trends) %>%
-  mutate(abx=factor(abx, levels=rev(abx_trends_levels)))
-
-drug_and_inapprop_trends %>%
+drug_trends %>%
+  filter(abx != 'overall_wrefill') %>%
+  # put the abx in order
+  arrange(desc(y)) %>%
+  mutate(abx=fct_inorder(factor(abx))) %>%
+  mutate(abx=fct_relevel(abx, c('overall', 'overall_inapprop', 'overall_approp'))) %>%
+  mutate(abx=fct_rev(abx)) %>%
+  # plot
   ggplot(aes(x=abx, y=y, ymin=ymin, ymax=ymax)) +
@@
   ylab('adjusted change in consumption 2011 to 2014 (%)') +
-  scale_y_continuous(limits=c(-22, 32), breaks=c(-20, -10, 0, 10, 20, 30)) +
+  scale_y_continuous(limits=c(-22, 32.5), breaks=c(-20, -10, 0, 10, 20, 30)) +
   theme_minimal() +
@@
         axis.text = element_text(size=16))
-
-drug_and_inapprop_trends %>%
-  arrange(desc(abx)) %>%
-  kable(caption='Percent changes; same data as in figure')
 ```
@@
 ```{r prescribing_practice}
-dx_counts_2011 = read_tsv('raw_data/dx_counts.tsv') %>%
-  left_join(select(pop_chars, year, n_bene), by='year') %>%
-  mutate(dkpy=round(n_encounter/n_bene*1000)) %>%
-  filter(year==2011) %>%
-  select(dx_cat, dkpy)
-
-dx_from_pde_2011 = read_tsv('raw_data/dx_from_pde.tsv') %>%
-  filter(year==2011) %>%
-  mutate(f_with_dx=n_pde_with_encounter/n_pde_with_any_encounter) %>%
-  select(antibiotic, dx_cat, year, f_with_dx)
-
 dx_to_pde_table = read_tsv('tables/dx_to_pde_table.tsv') %>%
@@
 # swo: convert ORs to RRs!
+risk_ratio_f = function(or, p) or / (1 + p * (or - 1))
 rr2pct = function(x) (x - 1) * 100
@@
 ```{r combined_table}
-format2 = function(x) format(round(x, 2), digits=2, nsmall=2)
+format_pct = function(x) format(round(x*100, 2), digits=2, nsmall=2)
 
-part1 = dx_counts_2011 %>%
-  mutate(dkpy=as.character(dkpy)) %>%
-  spread(dx_cat, dkpy) %>%
-  mutate(abx='dx', key1='dkpy', key2='')
-
-part2 = dx_from_pde_2011 %>%
-  # throw out some rare drugs
-  filter(!is.na(dx_cat)) %>%
-  mutate(f_with_dx=format2(f_with_dx*100)) %>%
-  spread(dx_cat, f_with_dx) %>%
-  mutate(key1='pde2dx', key2=as.character(year)) %>% rename(abx=antibiotic)
-
-part3 = dx_to_pde_table %>%
-  mutate(f_with_abx=format2(f_with_abx*100)) %>%
-  spread(dx_cat, f_with_abx) %>%
-  mutate(key1='dx2pde', key2=as.character(year))
-
-part4 = dx_to_pde_trends %>%
-  spread(dx_cat, change) %>%
-  mutate(key1='dx2pde_trend', key2='')
-
-combined_table = bind_rows(part1, part2, part3, part4) %>%
-  mutate(abx=factor(abx, levels=c('dx', abx_levels)),
-         key1=factor(key1, levels=c('dkpy', 'pde2dx', 'dx2pde', 'dx2pde_trend'))) %>%
-  arrange(abx, key1, key2)
+combined_table = dx_to_pde_table %>%
+  mutate(f_with_abx=format_pct(f_with_abx)) %>%
+  spread(year, f_with_abx) %>%
+  left_join(dx_to_pde_trends, by=c('dx_cat', 'abx')) %>%
+  arrange(abx, dx_cat)
 ```
@@
 combined_table %>%
-  select(abx, key1, key2, pneumonia, sinusitis, uri, bronchitis) %>%
-  filter(abx %in% c('dx', 'azithromycin', 'levofloxacin', 'amoxclav')) %>%
+  filter(dx_cat %in% c('pneumonia', 'sinusitis', 'uri', 'bronchitis'),
+         abx %in% c('dx', 'azithromycin', 'levofloxacin', 'amoxclav')) %>%
   kable
@@
 combined_table %>%
-  select(abx, key1, key2, pharyngitis, asthma, other_resp) %>%
-  filter(abx %in% c('dx', 'azithromycin', 'levofloxacin', 'amoxclav')) %>%
+  filter(dx_cat %in% c('pharyngitis', 'asthma', 'other_resp'),
+         abx %in% c('azithromycin', 'levofloxacin', 'amoxclav')) %>%
   kable
@@
 combined_table %>%
-  select(abx, key1, key2, uti, uti_t3) %>%
-  filter(abx %in% c('dx', 'ciprofloxacin', 'tmpsmx', 'nitrofurantoin')) %>%
+  filter(dx_cat %in% c('uti', 'uti_t3'),
+         abx %in% c('dx', 'ciprofloxacin', 'tmpsmx', 'nitrofurantoin')) %>%
   kable
@@
 combined_table %>%
-  select(abx, key1, key2, ssti, ssti_t3) %>%
-  filter(abx %in% c('dx', 'cephalexin', 'tmpsmx', 'clindamycin')) %>%
+  filter(dx_cat %in% c('ssti', 'ssti_t3'),
+         abx %in% c('dx', 'cephalexin', 'tmpsmx', 'clindamycin')) %>%
   kable
 ```
+
+## Projected appropriateness trend
+
+```{r approp_trends}
+# fold change in consumption
+pde_trend = drug_trends %>%
+  filter(abx=='overall') %>%
+  mutate(rr=exp(estimate),
+         rr_lo=exp(ci_lo),
+         rr_hi=exp(ci_hi))
+
+d_c = pde_trend$rr
+d_c_lo = pde_trend$rr_lo
+d_c_hi = pde_trend$rr_hi
+
+# fraction inappropriate in 2011
+f_i = scan('tables/inapprop_risk_2011.txt')
+
+inapprop_trend = read_tsv('tables/inapprop_trend.tsv') %>%
+  filter(term=='year') %>%
+  mutate(estimate=3*estimate,
+         std.error=sqrt(3)*std.error) %>%
+  mutate(ci=1.96*std.error,
+         ci_lo=estimate-ci,
+         ci_hi=estimate+ci) %>%
+  mutate(or   =exp(estimate),
+         or_lo=exp(ci_lo),
+         or_hi=exp(ci_hi)) %>%
+  mutate(rr   =risk_ratio_f(or, f_i),
+         rr_lo=risk_ratio_f(or_lo, f_i),
+         rr_hi=risk_ratio_f(or_hi, f_i))
+
+d_f = inapprop_trend$rr
+d_f_lo = inapprop_trend$rr_lo
+d_f_hi = inapprop_trend$rr_hi
+
+rr2pct = function(x) round((x - 1) * 100, 2)
+
+inapprop_trends = data_frame(
+  abx=c('overall_inapp', 'overall_app'),
+  y   =c(d_c    * d_f,    (1.0 - d_f    * f_i) / (1.0 - f_i) * d_c),
+  ymin=c(d_c_lo * d_f_lo, (1.0 - d_f_hi * f_i) / (1.0 - f_i) * d_c_lo),
+  ymax=c(d_c_hi * d_f_hi, (1.0 - d_f_lo * f_i) / (1.0 - f_i) * d_c_hi)
+) %>%
+  mutate_if(is.numeric, rr2pct)
+
+kable(inapprop_trends, caption='Percent changes in in/appropriate prescribing')
+```

### 2017-11-08 ms1/ms1_analyze.R

@@ -1,26 +1,19 @@
 #!/usr/bin/env Rscript
 
 output_table = function(x, base) write_tsv(x, sprintf('tables/%s.tsv', base))
-sem = function(x) sd(x) / sqrt(length(x))
-
-regions = read_tsv('db/census-regions.tsv') %>%
-  select(state, region)
 
 bene = read_tsv('raw_data/bene.tsv') %>%
-  rename(bene_id=BENE_ID) %>%
-  left_join(regions, by='state')
-
-n_unique_bene = length(unique(bene$bene_id)) %T>%
-  write('tables/n_unique_bene.txt')
-
-pde = read_tsv('raw_data/bene_pde.tsv') %>%
-  rename(bene_id=BENE_ID) %>%
-  rename(overall=n_pde) %>%
-  left_join(bene, by=c('year', 'bene_id')) %>%
+  rename(overall=n_pde,
+         overall_inapprop=n_pde_inapprop,
+         overall_wrefill=n_pde_wrefill) %>%
+  mutate(overall_approp=overall - overall_inapprop) %>%
   mutate(age_group=case_when(between(.$age, 65, 75) ~ 'age65_75',
                              between(.$age, 76, 85) ~ 'age76_85',
                              between(.$age, 86, 95) ~ 'age86_95',
                              .$age > 95 ~ 'age96_'))
+
+n_unique_bene = length(unique(bene$BENE_ID)) %T>%
+  write('tables/n_unique_bene.txt')
 
 # summary characteristics
 pop_chars = bene %>%
@@ -28,10 +21,8 @@ pop_chars = bene %>%
   summarize(n_bene=n(),
             mean_age=mean(age),
             sd_age=sd(age),
-            sem_age=sem(age),
             mean_n_cc=mean(n_cc),
             sd_n_cc=sd(n_cc),
-            sem_n_cc=sem(n_cc),
             n_female=sum(sex=='female'),
             n_white=sum(race=='white'),
             n_dual=sum(dual),
@@ -41,12 +32,18 @@ pop_chars = bene %>%
             n_region_midwest=sum(region=='Midwest')) %T>%
   output_table('pop_chars')
 
-claims_by_abx = pde %>%
-  select(year, overall:clindamycin) %>%
+claims_by_fill = bene %>%
+  select(overall, overall_wrefill) %>%
[…]

### 2017-11-08 ms1/ms1_display.Rmd
removed projected

@@
 ```
-
-## Projected appropriateness trend
-
-```{r approp_trends}
-# fold change in consumption
-pde_trend = drug_trends %>%
-  filter(abx=='overall') %>%
-  mutate(rr=exp(estimate),
-         rr_lo=exp(ci_lo),
-         rr_hi=exp(ci_hi))
-
-d_c = pde_trend$rr
-d_c_lo = pde_trend$rr_lo
-d_c_hi = pde_trend$rr_hi
-
-# fraction inappropriate in 2011
-f_i = scan('tables/inapprop_risk_2011.txt')
-
-inapprop_trend = read_tsv('tables/inapprop_trend.tsv') %>%
-  filter(term=='year') %>%
-  mutate(estimate=3*estimate,
-         std.error=sqrt(3)*std.error) %>%
-  mutate(ci=1.96*std.error,
-         ci_lo=estimate-ci,
-         ci_hi=estimate+ci) %>%
-  mutate(or   =exp(estimate),
-         or_lo=exp(ci_lo),
-         or_hi=exp(ci_hi)) %>%
-  mutate(rr   =risk_ratio_f(or, f_i),
-         rr_lo=risk_ratio_f(or_lo, f_i),
-         rr_hi=risk_ratio_f(or_hi, f_i))
-
-d_f = inapprop_trend$rr
-d_f_lo = inapprop_trend$rr_lo
-d_f_hi = inapprop_trend$rr_hi
-
-rr2pct = function(x) round((x - 1) * 100, 2)
-
-inapprop_trends = data_frame(
-  abx=c('overall_inapp', 'overall_app'),
-  y   =c(d_c    * d_f,    (1.0 - d_f    * f_i) / (1.0 - f_i) * d_c),
-  ymin=c(d_c_lo * d_f_lo, (1.0 - d_f_hi * f_i) / (1.0 - f_i) * d_c_lo),
-  ymax=c(d_c_hi * d_f_hi, (1.0 - d_f_lo * f_i) / (1.0 - f_i) * d_c_hi)
-) %>%
-  mutate_if(is.numeric, rr2pct)
-
-kable(inapprop_trends, caption='Percent changes in in/appropriate prescribing')
-```

### 2017-11-08 ms1/ms1_analyze.R

@@ -152,13 +152,3 @@ dx_to_pde_trends = dx_to_pde_trend_pairs %>%
   do(dx_to_pde_trend_f(.$abx, .$dx_cat)) %>%
   ungroup() %T>%
   output_table('dx_to_pde_trends')
-
-# projected appropriateness trend
-bene %>%
-  glm_f(cbind(overall_inapprop, overall_approp) ~ year + age + n_cc + sex + race + dual + region, family='binomial') %>%
-  sandwich_tidy() %T>%
-  output_table('projected_inapprop_trend')
-
-bene %>%
-  filter(year==2011) %$%
-  write(sum(overall_inapprop) / sum(overall), 'tables/inapprop_risk_2011.txt')

### 2017-11-10 ms1/ms1_display.Rmd
small display

@@
 claims_by_fill %>%
+  mutate_at(vars(-f_refill), function(x) prettyNum(x, big.mark=',')) %>%
   kable(caption='PDEs by refill')
+```
 
-#  mutate(cum_n=prettyNum(cumsum(n_pde), big.mark=',')) %>%
-```
+The top 10 drugs take up `r round(scan('tables/top10_cumf.txt') * 100, 2)`% of all claims.
 
@@
 
+claims_by_abx %>%
+  mutate(cpkp=round(cpkp)) %>%
+  spread(year, cpkp) %>%
+  kable(caption='Raw drug CPKP')
+```
+
+```{r raw_drug_trends_figure}
 abx_levels = levels(claims_by_abx$abx)
@@
   mutate(change=sprintf('%.2f (%.2f -- %.2f)', y, ymin, ymax)) %>%
-  select(dx_cat, abx, change)
+  select(dx_cat, abx, change, p.value)
 ```
@@
   left_join(dx_to_pde_trends, by=c('dx_cat', 'abx')) %>%
+  mutate(p.value=format_p(p.value)) %>%
   arrange(abx, dx_cat)

### 2017-11-10 ms1/ms1_analyze.R

@@ -48,6 +48,16 @@ claims_by_abx = bene %>%
   gather('abx', 'cpkp', -year) %T>%
   output_table('claims_by_abx')
 
+# summary of claims taken up by top 10
+bene %>%
+  select(overall, azithromycin:clindamycin) %>%
+  summarize_all(sum) %>%
+  gather('abx', 'n', azithromycin:clindamycin) %>%
+  mutate(cumf=cumsum(n)/overall) %>%
+  tail(1) %>%
+  pull(cumf) %>%
+  write('tables/top10_cumf.txt')
+
 # from here on, the denominators are taken from the beneficiary data grouped
 # in the same way as the consumption data, so we can use the summarize_by function
 summarize_by = function(df, by) {

### 2017-11-22 ms1/ms1_display.Rmd
minor

@@
   ggplot(aes(x=year, y=cpkp, color=abx)) +
-  geom_point() + geom_line() +
-  #scale_y_continuous(limits=c(0, 1525), breaks=c(1000, 1125, 1250, 1375, 1500)) +
-  scale_y_continuous(limits=c(0, 1525), breaks=c(0, 250, 500, 750, 1000, 1250, 1500)) +
+  geom_line() +
+  scale_y_continuous(limits=c(0, 1400), breaks=c(0, 250, 500, 750, 1000, 1250, 1400)) +
+  coord_fixed(ratio=4/1400*1.25) +
   theme_minimal() +
+  theme(panel.grid.minor.x=element_blank()) +
   ggtitle('Unadjusted consumption, all abx')
@@
   ggplot(aes(x=year, y=cpkp, color=abx)) +
-  geom_point() + geom_line() +
+  geom_line() +
   scale_y_continuous(limits=c(0, 250), breaks=c(0, 50, 100, 150, 200, 250)) +
+  coord_fixed(ratio=4/250*1.25) +
   theme_minimal() +
+  theme(panel.grid.minor.x=element_blank()) +
   ggtitle('Unadjusted consumption, top 5')
@@
   ggplot(aes(x=year, y=cpkp, color=abx)) +
-  geom_point() + geom_line() +
+  geom_line() +
   scale_y_continuous(limits=c(0, 100), breaks=c(0, 50, 100)) +
+  coord_fixed(ratio=4/100*1.25) +
   theme_minimal() +
+  theme(panel.grid.minor.x=element_blank()) +
   ggtitle('Unadjusted consumption, next 5')
@@
 
+## Raw inappropriate fractions
+
+```{r inappropriate_fraction}
+claims_by_abx %>%
+  spread(abx, cpkp) %>%
+  mutate(pct_inapprop = overall_inapprop / overall * 100) %>%
+  select(year, pct_inapprop) %>%
+  kable(caption='Raw fraction of inappropriate prescribing')
+```
+
 # Prescribing practice
@@
   mutate(p.value=format_p(p.value)) %>%
-  arrange(abx, dx_cat)
+  arrange(dx_cat, abx)
 ```
@@
   mutate(change=sprintf('%.2f (%.2f -- %.2f)', y, ymin, ymax)) %>%
-  select(group_type, group, change)
+  select(group_type, group, change, p.value)
 

### 2017-11-27 ms1/ms1_display.Rmd
minor

@@
 
-drug_trends = read_tsv('tables/trend_models.tsv') %>%
+drug_trends = read_tsv('raw_data/poisson_models.tsv') %>%
+  select(abx=model, term=Parameter, estimate=Estimate, std.error=StdErr, p.value=ProbChiSq) %>%
   filter(term=='year') %>%
@@
          ymax=est2pct(ci_hi)) %>%
-  mutate(change=sprintf('%.2f (%.2f -- %.2f)', y, ymin, ymax))
+  mutate(change=sprintf('%.1f (%.1f -- %.1f)', y, ymin, ymax))
 ```
@@
 ```{r drug_trends_display}
-format_p = function(p) {
-  if_else(p < 0.001, '< 0.001', as.character(round(p, 3)))
-}
-
 drug_trends %>%
   filter(abx != 'overall_wrefill') %>%
-  mutate(adj.p=p.adjust(p.value, method='BH')) %>%
   select(abx, y, ymin, ymax, p.value) %>%
-  mutate_at(vars(y, ymin, ymax), function(x) format(x, digits=2, nsmall=2)) %>%
-  mutate(p.value=format_p(p.value)) %>%
+  mutate_at(vars(y, ymin, ymax), function(x) format(x, digits=1, nsmall=1)) %>%
   kable(caption='Drug trends: Percent change from 2011 to 2014')
@@
 drug_trends %>%
-  filter(abx != 'overall_wrefill') %>%
   # put the abx in order
@@
   kable(caption='Raw fraction of inappropriate prescribing')
+```
+
+## Odds and stuff
+
+```{r odds}
+inapp_odds_trend = read_tsv('raw_data/logistic_models.tsv') %>%
+  select(model, term=Variable, estimate=Estimate, std.error=StdErr, p.value=ProbChiSq) %>%
+  filter(term=='year') %>%
+  mutate(estimate=3*estimate,
+         ci=1.96*sqrt(3)*std.error,
+         y=exp(estimate),
+         ymin=exp(estimate-ci),
+         ymax=exp(estimate+ci)) %>%
+  select(model, y, ymin, ymax) %>%
+  mutate_if(is.numeric, function(x) round((x-1)*100, 1))
+
+inapp_odds_trend %>% kable
 ```
@@
          ymax=rr2pct(rr_hi)) %>%
-  mutate(change=sprintf('%.2f (%.2f -- %.2f)', y, ymin, ymax)) %>%
+  mutate(change=sprintf('%.1f (%.1f -- %.1f)', y, ymin, ymax)) %>%
   select(dx_cat, abx, change, p.value)
@@
 ```{r combined_table}
-format_pct = function(x) format(round(x*100, 2), digits=2, nsmall=2)
+format_pct = function(x) format(round(x*100, 1), digits=1, nsmall=1)
+format_p = function(p) if_else(p<0.0001, '<.0001', format(round(p, 4), digits=4, nsmall=4))
 
@@
          ymax=est2pct(ci_hi)) %>%
-  mutate(change=sprintf('%.2f (%.2f -- %.2f)', y, ymin, ymax)) %>%
+  mutate(change=sprintf('%.1f (%.1f -- %.1f)', y, ymin, ymax)) %>%
   select(group_type, group, change, p.value)

### 2018-01-09 ms1/ms1_display.Rmd
major cleanups

--- 2017-11-22 ms1/ms1_display.Rmd
@@
 ```{r pop_chars}
-pop_chars = read_tsv('tables/pop_chars.tsv') %>%
-  mutate_at(vars(n_female:n_region_midwest), function(x) x / .$n_bene * 100)
+pop_chars = read_tsv('tables/new/study_pop_chars.tsv')
+
+pop_chars = read_tsv('tables/new/study_pop_chars.tsv') %>%
+  spread(key, value) %>%
+  (function (df) {
+    ns1 = names(df)
+    ns2 = str_replace(ns1, '^n_(?!bene|cc)', 'pct_')
+    rename(df, !!!setNames(as.list(ns1), ns2))
+  }) %>%
+  mutate_at(vars(matches('^pct_')), function(x) x / .$n_bene * 100) %>%
+  gather('key', 'value', -year)
+
+pop_chars %>%
+  mutate(key=factor(key, levels=c('n_bene', 'mean_age', 'sd_age', 'n_cc', 'sd_cc', 'pct_female', 'pct_white', 'pct_dual', 'pct_south', 'pct_midwest', 'pct_west', 'pct_northeast'))) %>%
+  mutate(value=case_when(
+    .$key == 'n_bene' ~ prettyNum(.$value, big.mark=','),
+    .$key %in% c('n_cc', 'sd_cc') ~ format(.$value, digits=2, nsmall=2, scientific=FALSE),
+    TRUE ~ format(.$value, digits=2, nsmall=1, scientific=FALSE)
+  )) %>%
+  spread(year, value) %>%
+  kable(caption='Population characteristics')
 
@@
 
-pop_char_trends = pop_chars %>%
-  select(year, n_bene, mean_age, mean_n_cc, n_female:n_region_midwest) %>%
-  gather('key', 'value', -year) %>%
+pop_chars %>%
   arrange(key, year) %>%
   group_by(key) %>%
-  summarize(p=trend_p(value)) %>%
-  mutate(adj.p=p.adjust(p, method='BH'))
-
-pop_chars %>%
-  mutate(n_bene=prettyNum(n_bene, big.mark=',')) %>%
-  mutate_at(vars(mean_age, sd_age, n_female:n_region_midwest), function(x) format(x, digits=1, nsmall=1)) %>%
-  mutate_at(vars(mean_n_cc, sd_n_cc), function(x) format(x, digits=2, nsmall=2)) %>%
-  gather('key', 'value', -year) %>%
-  spread(year, value) %>%
-  mutate(key=factor(key, levels=names(pop_chars))) %>%
-  arrange(key) %>%
-  kable(caption='Population characteristics')
-
-pop_char_trends %>%
+  summarize(p_value=trend_p(value)) %>%
   kable(caption='Trends in population characteristics')
@@
 There are `r scan('tables/n_unique_bene.txt') %>% prettyNum(big.mark=',')` unique beneficiaries.
-
-```{r claims_by_fill}
-claims_by_fill = read_tsv('tables/claims_by_fill.tsv') %>%
-  mutate(f_refill=refills/overall_wrefill)
-
-claims_by_fill %>%
-  mutate_at(vars(-f_refill), function(x) prettyNum(x, big.mark=',')) %>%
-  kable(caption='PDEs by refill')
-```
 
@@
 
-```{r raw_drug_trends}
-claims_by_abx = read_tsv('tables/claims_by_abx.tsv') %>%
-  mutate(abx=fct_inorder(factor(abx)))
+```{r raw_drug_trends_table}
+claims_by_abx = read_tsv('tables/new/claims_by_abx.tsv') %>%
+  gather('abx', 'n', c(-year, -bene)) %>%
+  mutate(cpkp=n/bene*1000)
 
@@
   mutate(cpkp=round(cpkp)) %>%
+  select(year, abx, cpkp) %>%
   spread(year, cpkp) %>%
+  arrange(desc(`2011`)) %>%
   kable(caption='Raw drug CPKP')
@@
 
-```{r raw_drug_trends_figure}
+```{r raw_drug_trends}
 abx_levels = levels(claims_by_abx$abx)
@@
 claims_by_abx %>%
-  filter(abx %in% c('overall', 'overall_approp', 'overall_inapprop')) %>%
+  filter(abx %in% c('pde', 'pde_approp', 'pde_inapprop', 'pde_nodx')) %>%
   ggplot(aes(x=year, y=cpkp, color=abx)) +
@@
 ```{r drug_trends}
-est2pct = function(x) (exp(x) - 1)*100
+est2pct = function(x) (x - 1)*100
 
-drug_trends = read_tsv('tables/trend_models.tsv') %>%
-  filter(term=='year') %>%
-  arrange(estimate) %>%
-  mutate(estimate=3*estimate,
-         std.error=sqrt(3)*std.error) %>%
-  mutate(ci=1.96*std.error,
-         ci_lo=estimate-ci,
-         ci_hi=estimate+ci) %>%
-  mutate(y=est2pct(estimate),
-         ymin=est2pct(ci_lo),
-         ymax=est2pct(ci_hi)) %>%
-  mutate(change=sprintf('%.2f (%.2f -- %.2f)', y, ymin, ymax))
-```
-
-```{r drug_trends_display}
-format_p = function(p) {
-  if_else(p < 0.001, '< 0.001', as.character(round(p, 3)))
-}
+drug_trends = read_tsv('tables/new/abx_models.tsv') %>%
+  mutate_at(vars(estimate, lower_ci, upper_ci), est2pct) %>%
+  mutate(change=sprintf('%.1f (%.1f -- %.1f)', estimate, lower_ci, upper_ci))
 
 drug_trends %>%
-  filter(abx != 'overall_wrefill') %>%
-  mutate(adj.p=p.adjust(p.value, method='BH')) %>%
-  select(abx, y, ymin, ymax, p.value) %>%
-  mutate_at(vars(y, ymin, ymax), function(x) format(x, digits=2, nsmall=2)) %>%
-  mutate(p.value=format_p(p.value)) %>%
+  mutate_at(vars(estimate, lower_ci, upper_ci), function(x) format(x, digits=1, nsmall=1, scientific=FALSE)) %>%
   kable(caption='Drug trends: Percent change from 2011 to 2014')
@@
 drug_trends %>%
-  filter(abx != 'overall_wrefill') %>%
   # put the abx in order
-  arrange(desc(y)) %>%
-  mutate(abx=fct_inorder(factor(abx))) %>%
-  mutate(abx=fct_relevel(abx, c('overall', 'overall_inapprop', 'overall_approp'))) %>%
-  mutate(abx=fct_rev(abx)) %>%
+  arrange(desc(estimate)) %>%
+  mutate(antibiotic=fct_inorder(factor(antibiotic))) %>%
+  mutate(antibiotic=fct_relevel(antibiotic, c('n_pde', 'n_pde_approp', 'n_pde_inapprop', 'n_pde_nodx'))) %>%
+  mutate(antibiotic=fct_rev(antibiotic)) %>%
   # plot
-  ggplot(aes(x=abx, y=y, ymin=ymin, ymax=ymax)) +
+  ggplot(aes(x=antibiotic, y=estimate, ymin=lower_ci, ymax=upper_ci)) +
   geom_bar(stat='identity', fill='white', color='black') +
@@
 claims_by_abx %>%
-  spread(abx, cpkp) %>%
-  mutate(pct_inapprop = overall_inapprop / overall * 100) %>%
-  select(year, pct_inapprop) %>%
-  kable(caption='Raw fraction of inappropriate prescribing')
+  filter(str_detect(abx, '^pde')) %>%
+  select(year, abx, n) %>%
+  spread(abx, n) %>%
+  mutate_at(vars(matches('^pde_')), function(x) round(x / .$pde * 100, 1)) %>%
+  select(-pde) %>%
+  kable(caption='Proportion of app., inapp., and no-dx prescribing')
 ```
@@
 
-```{r prescribing_practice}
-dx_to_pde_table = read_tsv('tables/dx_to_pde_table.tsv') %>%
-  select(-n_encounters) %>%
-  filter(year %in% c(2011, 2014)) %>%
-  gather('abx', 'f_with_abx', azithromycin:clindamycin)
+```{r top_dx}
+denom = claims_by_abx %>%
+  select(year, bene) %>%
+  distinct()
 
-# swo: convert ORs to RRs!
-risk_ratio_f = function(or, p) or / (1 + p * (or - 1))
-rr2pct = function(x) (x - 1) * 100
+rxdx_table = read_tsv('tables/new/rxdx_counts.tsv') %>%
+  left_join(denom, by='year') %>%
+  mutate(cpkp=n_pdedx / bene * 1000)
 
-dx_to_pde_trends = read_tsv('tables/dx_to_pde_trends.tsv') %>%
-  filter(term=='year') %>%
-  mutate(estimate=3*estimate,
-         std.error=sqrt(3)*std.error) %>%
-  mutate(ci=1.96*std.error,
-         ci_lo=estimate-ci,
-         ci_hi=estimate+ci) %>%
-  mutate(or=   exp(estimate),
-         or_lo=exp(ci_lo),
-         or_hi=exp(ci_hi)) %>%
-  left_join(filter(dx_to_pde_table, year==2011), by=c('dx_cat', 'abx')) %>%
-  mutate(rr=   risk_ratio_f(or,    f_with_abx),
-         rr_lo=risk_ratio_f(or_lo, f_with_abx),
-         rr_hi=risk_ratio_f(or_hi, f_with_abx)) %>%
-  mutate(y=rr2pct(rr),
-         ymin=rr2pct(rr_lo),
-         ymax=rr2pct(rr_hi)) %>%
-  mutate(change=sprintf('%.2f (%.2f -- %.2f)', y, ymin, ymax)) %>%
-  select(dx_cat, abx, change, p.value)
+top_dx = rxdx_table %>%
+  filter(year==2011) %>%
+  group_by(dx_cat) %>%
+  summarize(max_cpkp=max(cpkp)) %>%
+  mutate(over_10=max_cpkp >= 10) %>%
+  arrange(desc(max_cpkp))
+
+top_dx %>%
+  kable(caption='Diagnoses by maximum drug usage in 2011')
 ```
 
-```{r combined_table}
-format_pct = function(x) format(round(x*100, 2), digits=2, nsmall=2)
+```{r top_rxdx}
+infection_sites = read_tsv('infection_sites.tsv')
 
-combined_table = dx_to_pde_table %>%
-  mutate(f_with_abx=format_pct(f_with_abx)) %>%
-  spread(year, f_with_abx) %>%
-  left_join(dx_to_pde_trends, by=c('dx_cat', 'abx')) %>%
-  mutate(p.value=format_p(p.value)) %>%
-  arrange(dx_cat, abx)
+top_dx %>%
+  filter(over_10) %>%
+  slice(-1) %>% # drop "remaining codes"
+  select(dx_cat) %>%
+  inner_join(rxdx_table, by='dx_cat') %>%
+  left_join(infection_sites, by='dx_cat') %>%
+  filter(year==2011) %>%
+  group_by(site, antibiotic) %>%
+  summarize(cpkp=sum(cpkp)) %>%
+  arrange(site, desc(cpkp)) %>%
+  group_by(site) %>%
+  mutate(f=cpkp/sum(cpkp), cumf=cumsum(f)/sum(f)) %>%
+  do(head(., 5)) %>%
+  kable(caption='Top drugs by diagnosis in 2011')
+```
+
+```{r prescribing_practice}
+rxdx_models = read_tsv('tables/new/rxdx_models.tsv')
+
+format_p = function(p) if_else(p<0.0001, '<.0001', format(round(p, 4), digits=4, nsmall=4))
+
+combined_table = rxdx_table %>%
+  mutate(cpkp=round(n_pdedx / bene * 1000, 1)) %>%
+  filter(year %in% c(2011, 2014)) %>%
+  select(antibiotic, dx_cat, year, cpkp) %>%
+  spread(year, cpkp) %>%
+  right_join(rxdx_models, by=c('antibiotic', 'dx_cat')) %>%
+  mutate_at(vars(estimate, lower_ci, upper_ci), est2pct) %>%
+  mutate(change=sprintf('%.1f (%.1f -- %.1f)', estimate, lower_ci, upper_ci)) %>%
+  mutate(d=pmax(estimate-lower_ci, upper_ci-estimate)) %>%
+  mutate(change2=sprintf('%.1f ± %.1f', estimate, d)) %>%
+  mutate(p.value=format_p(p_value)) %>%
+  select(dx_cat, antibiotic, `2011`, `2014`, change2, p.value) %>%
+  arrange(dx_cat, antibiotic)
 ```
@@
 ```{r respiratory}
+rxdx_kable = function(abxs, dxs) {
   combined_table %>%
-  filter(dx_cat %in% c('pneumonia', 'sinusitis', 'uri', 'bronchitis'),
-         abx %in% c('dx', 'azithromycin', 'levofloxacin', 'amoxclav')) %>%
-  kable
+    filter(dx_cat %in% dxs, antibiotic %in% abxs) %>%
+    mutate(dx_cat=factor(dx_cat, levels=dxs),
+           antibiotic=factor(antibiotic, levels=abxs)) %>%
+    arrange(dx_cat, antibiotic) %>%
+    kable()
+}
+
+rxdx_kable(c('azithromycin', 'levofloxacin', 'amoxclav'),
+           c('pneumonia', 'sinusitis', 'uri', 'bronchitis'))
 ```
@@
 ```{r claims_by_demography}
-claims_by = function(by) {
-  read_tsv(str_interp("tables/claims_by_${by}.tsv")) %>%
-    mutate(group_type=by) %>%
-    rename(group=!!rlang::sym(by)) %>%
+subpop_table = read_tsv('tables/new/claims_by_subpop.tsv') %>%
+  mutate(cpkp=round(n_pde/n_bene*1000)) %>%
   filter(year %in% c(2011, 2014)) %>%
-    select(group_type, group, year, cpkp) %>%
-    spread(year, cpkp) %>%
-    mutate_at(vars(`2011`, `2014`), round)
-}
+  select(subpop, year, cpkp) %>%
+  spread(year, cpkp)
 
-trends_by_demography = read_tsv('tables/demography_models.tsv') %>%
-  filter(term=='year') %>%
-  mutate(estimate=3*estimate,
-         std.error=sqrt(3)*std.error) %>%
-  mutate(ci=1.96*std.error,
-         ci_lo=estimate-ci,
-         ci_hi=estimate+ci) %>%
-  mutate(y=est2pct(estimate),
-         ymin=est2pct(ci_lo),
-         ymax=est2pct(ci_hi)) %>%
-  mutate(change=sprintf('%.2f (%.2f -- %.2f)', y, ymin, ymax)) %>%
-  select(group_type, group, change, p.value)
+subpop_models = read_tsv('tables/new/subpop_models.tsv') %>%
+  mutate_at(vars(estimate, lower_ci, upper_ci), est2pct) %>%
+  mutate(pm=pmax(estimate-lower_ci, upper_ci-estimate)) %>%
+  mutate(change=sprintf('%.1f (%.1f -- %.1f)', estimate, lower_ci, upper_ci)) %>%
+  mutate(change2=sprintf('%.1f ± %.1f', estimate, pm)) %>%
+  mutate(p.value=format_p(p_value)) %>%
+  select(subpop, change2, p.value)
 
-# overall trend row
-overall_change = drug_trends %>% filter(abx=='overall') %>% pull(change)
-overall_row = claims_by_abx %>%
-  filter(abx=='overall', year %in% c(2011, 2014)) %>%
-  mutate(cpkp=round(cpkp)) %>%
-  spread(year, cpkp) %>%
-  mutate(group_type='overall', group='overall', change=overall_change) %>%
-  select(group_type, group, `2011`, `2014`, change)
-
-bind_rows(
-  claims_by('age_group'),
-  claims_by('sex'),
-  claims_by('race'),
-  claims_by('region')
-) %>%
-  left_join(trends_by_demography, by=c('group', 'group_type')) %>%
-  bind_rows(overall_row, .) %>%
-  kable
+left_join(subpop_table, subpop_models, by='subpop') %>%
+  kable(caption='Claims and trends by subpopulation')
 ```
@@
 ```{r other_respiratory}
-combined_table %>%
-  filter(dx_cat %in% c('pharyngitis', 'asthma', 'other_resp'),
-         abx %in% c('azithromycin', 'levofloxacin', 'amoxclav')) %>%
-  kable
+rxdx_kable(c('azithromycin', 'levofloxacin', 'amoxclav'),
+           c('pharyngitis', 'asthma', 'other_resp'))
+```
+
+## GI
+
+```{r gi}
+rxdx_kable(c('ciprofloxacin', 'metronidazole', 'levofloxacin'),
+           c('gi', 'gi_t3'))
 ```
@@
 ```{r uti}
-combined_table %>%
-  filter(dx_cat %in% c('uti', 'uti_t3'),
-         abx %in% c('dx', 'ciprofloxacin', 'tmpsmx', 'nitrofurantoin')) %>%
-  kable
+rxdx_kable(c('ciprofloxacin', 'tmpsmx', 'nitrofurantoin'),
+           c('uti', 'uti_t3'))
 ```
@@
 ```{r ssti}
-combined_table %>%
-  filter(dx_cat %in% c('ssti', 'ssti_t3'),
-         abx %in% c('dx', 'cephalexin', 'tmpsmx', 'clindamycin')) %>%
-  kable
+rxdx_kable(c('cephalexin', 'tmpsmx'),
+           c('ssti', 'ssti_t3'))
 ```

### 2018-01-09 ms1/ms1_display.Rmd
combined resp tables

@@
 
-```{r drug_trends}
+```{r abx_trends}
 est2pct = function(x) (x - 1)*100
+format_p = function(p) if_else(p<0.0001, '<.0001', format(round(p, 4), digits=4, nsmall=4))
 
-drug_trends = read_tsv('tables/new/abx_models.tsv') %>%
+ceiling_at = function(x, n) {
+  mult = 10 ** n
+  x %>% { . * mult } %>% ceiling %>% { . / mult }
+}
+
+format_model = function(df) {
+  df %>%
     mutate_at(vars(estimate, lower_ci, upper_ci), est2pct) %>%
-  mutate(change=sprintf('%.1f (%.1f -- %.1f)', estimate, lower_ci, upper_ci))
+    mutate(pm=pmax(estimate-lower_ci, upper_ci-estimate)) %>%
+    mutate(change=sprintf('%.1f ± %.1f', estimate, ceiling_at(pm, 1))) %>%
+    mutate(p.value=format_p(p_value))
+}
 
-drug_trends %>%
+abx_trends = read_tsv('tables/new/abx_models.tsv') %>%
+  format_model()
+
+abx_trends %>%
   mutate_at(vars(estimate, lower_ci, upper_ci), function(x) format(x, digits=1, nsmall=1, scientific=FALSE)) %>%
@@
 
-drug_trends %>%
+abx_trends %>%
   # put the abx in order
@@
 
-format_p = function(p) if_else(p<0.0001, '<.0001', format(round(p, 4), digits=4, nsmall=4))
-
 combined_table = rxdx_table %>%
@@
   right_join(rxdx_models, by=c('antibiotic', 'dx_cat')) %>%
-  mutate_at(vars(estimate, lower_ci, upper_ci), est2pct) %>%
-  mutate(change=sprintf('%.1f (%.1f -- %.1f)', estimate, lower_ci, upper_ci)) %>%
-  mutate(d=pmax(estimate-lower_ci, upper_ci-estimate)) %>%
-  mutate(change2=sprintf('%.1f ± %.1f', estimate, d)) %>%
-  mutate(p.value=format_p(p_value)) %>%
-  select(dx_cat, antibiotic, `2011`, `2014`, change2, p.value) %>%
+  format_model() %>%
+  select(dx_cat, antibiotic, `2011`, `2014`, change, p.value) %>%
   arrange(dx_cat, antibiotic)
@@
 rxdx_kable(c('azithromycin', 'levofloxacin', 'amoxclav'),
-           c('pneumonia', 'sinusitis', 'uri', 'bronchitis'))
+           c('pneumonia', 'sinusitis', 'uri', 'bronchitis', 'other_resp', 'asthma'))
 ```
@@
 
-## Other respiratory
-
-```{r other_respiratory}
-rxdx_kable(c('azithromycin', 'levofloxacin', 'amoxclav'),
-           c('pharyngitis', 'asthma', 'other_resp'))
-```
-
 ## GI

### 2018-01-09 ms1/ms1_display.Rmd
panels for figure 1

@@
 
-claims_by_abx %>%
-  filter(abx %in% c('pde', 'pde_approp', 'pde_inapprop', 'pde_nodx')) %>%
-  ggplot(aes(x=year, y=cpkp, color=abx)) +
-  geom_line() +
-  scale_y_continuous(limits=c(0, 1400), breaks=c(0, 250, 500, 750, 1000, 1250, 1400)) +
-  coord_fixed(ratio=4/1400*1.25) +
-  theme_minimal() +
-  theme(panel.grid.minor.x=element_blank()) +
-  ggtitle('Unadjusted consumption, all abx')
+dummy = data_frame(
+  year = rep(c(2011, 2014), 3),
+  cpkp = c(0, 1500, 0, 250, 0, 100),
+  panel = factor(c(1, 1, 2, 2, 3, 3)),
+  abx = 'pde'
+)
+
+panel_labels = c('1'='a) Overall antibiotic consumption',
+                 '2'='b) Top 6 most-used antibiotics',
+                 '3'='c) Next 4 most-used antibiotics')
 
 claims_by_abx %>%
-  filter(abx %in% c('azithromycin', 'ciprofloxacin', 'amoxicillin', 'cephalexin', 'tmpsmx', 'levofloxacin')) %>%
-  ggplot(aes(x=year, y=cpkp, color=abx)) +
+  mutate(panel = factor(case_when(
+    .$abx %in% c('pde', 'pde_approp', 'pde_inapprop', 'pde_nodx') ~ 1,
+    .$abx %in% c('azithromycin', 'ciprofloxacin', 'amoxicillin', 'cephalexin', 'tmpsmx', 'levofloxacin') ~ 2,
+    .$abx %in% c('amoxclav', 'doxycycline', 'nitrofurantoin', 'clindamycin') ~ 3
+  ))) %>%
+  mutate(abx=case_when(
+    .$abx == 'pde' ~ 'Overall use',
+    .$abx == 'pde_inapprop' ~ 'Inappropriate use',
+    .$abx == 'pde_approp' ~ 'Appropriate use',
+    .$abx == 'pde_nodx' ~ 'Indeterminate use',
+    .$abx == 'amoxclav' ~ 'amox/clav',
+    .$abx == 'tmpsmx' ~ 'TMP/SMX',
+    TRUE ~ .$abx)
+  ) %>%
+  arrange(year, desc(cpkp)) %>%
+  {
+  ggplot(., aes(x=year, y=cpkp, group=abx)) +
+  geom_blank(data=dummy) +
   geom_line() +
-  scale_y_continuous(limits=c(0, 250), breaks=c(0, 50, 100, 150, 200, 250)) +
-  coord_fixed(ratio=4/250*1.25) +
-  theme_minimal() +
-  theme(panel.grid.minor.x=element_blank()) +
-  ggtitle('Unadjusted consumption, top 5')
+  geom_text_repel(
+    data=filter(., year==2013),
+    aes(label=abx)
+  ) +
+  facet_wrap(~panel, scales='free_y', labeller=labeller(panel=panel_labels)) +
+  theme_classic() +
+  theme(panel.grid.minor.x=element_blank(),
+        strip.background = element_blank()) +
+  scale_x_continuous(expand=c(0, 0), name='') +
+  scale_y_continuous(expand=c(0, 0), name='claims per 1,000 beneficiaries per year')
+  }
 
-claims_by_abx %>%
-  filter(abx %in% c('amoxclav', 'doxycycline', 'nitrofurantoin', 'clindamycin')) %>%
-  ggplot(aes(x=year, y=cpkp, color=abx)) +
-  geom_line() +
-  scale_y_continuous(limits=c(0, 100), breaks=c(0, 50, 100)) +
-  coord_fixed(ratio=4/100*1.25) +
-  theme_minimal() +
-  theme(panel.grid.minor.x=element_blank()) +
-  ggtitle('Unadjusted consumption, next 5')
+ggsave('fig/raw_drug.pdf', height=5.0, width=9.0, unit='in')
 ```
@@
 abx_trends %>%
+  mutate(antibiotic=case_when(
+    .$antibiotic == 'n_pde' ~ 'Overall use',
+    .$antibiotic == 'n_pde_inapprop' ~ 'Inappropriate use',
+    .$antibiotic == 'n_pde_approp' ~ 'Appropriate use',
+    .$antibiotic == 'n_pde_nodx' ~ 'Indeterminate use',
+    .$antibiotic == 'amoxclav' ~ 'amox/clav',
+    .$antibiotic == 'tmpsmx' ~ 'TMP/SMX',
+    TRUE ~ .$antibiotic)
+  ) %>%
   # put the abx in order
@@
   mutate(antibiotic=fct_inorder(factor(antibiotic))) %>%
-  mutate(antibiotic=fct_relevel(antibiotic, c('n_pde', 'n_pde_approp', 'n_pde_inapprop', 'n_pde_nodx'))) %>%
+  mutate(antibiotic=fct_relevel(antibiotic, c('Overall use', 'Inappropriate use', 'Appropriate use', 'Indeterminate use'))) %>%
   mutate(antibiotic=fct_rev(antibiotic)) %>%
@@
   xlab('') +
-  ylab('adjusted change in consumption 2011 to 2014 (%)') +
-  scale_y_continuous(limits=c(-22, 32.5), breaks=c(-20, -10, 0, 10, 20, 30)) +
+  ylab('adjusted change in claims per beneficiary per year from 2011 to 2014 (%)') +
+  scale_y_continuous(limits=c(-22, 33), breaks=c(-20, -10, 0, 10, 20, 30)) +
   theme_minimal() +
-  theme(axis.title.x = element_text(size=16),
-        axis.text = element_text(size=16))
+  theme(axis.title.x = element_text(size=12),
+        axis.text = element_text(size=12))
+
+ggsave('fig/abx_trends.pdf', width=6.4, height=4.5, units='in')
 ```

### 2018-01-19 ms1/ms1_display.Rmd
small updates

@@
   ))) %>%
+  filter(!is.na(panel)) %>% # remove metro
   mutate(abx=case_when(
@@
 abx_trends %>%
+  filter(antibiotic != 'metronidazole') %>%
   mutate(antibiotic=case_when(
@@
   spread(abx, n) %>%
-  mutate_at(vars(matches('^pde_')), function(x) round(x / .$pde * 100, 1)) %>%
-  select(-pde) %>%
+  mutate(denom=pde_approp + pde_inapprop + pde_nodx) %>%
+  mutate_at(vars(matches('^pde_')), function(x) round(x / .$denom * 100, 1)) %>%
+  select(-pde, -denom) %>%
   kable(caption='Proportion of app., inapp., and no-dx prescribing')
@@
   mutate(cpkp=round(n_pde/n_bene*1000)) %>%
-  filter(year %in% c(2011, 2014)) %>%
+  #filter(year %in% c(2011, 2014)) %>%
   select(subpop, year, cpkp) %>%

### 2018-01-22 ms1/ms1_display.Rmd
moved tables, updated tables

@@
 ```{r pop_chars}
-pop_chars = read_tsv('tables/new/study_pop_chars.tsv')
+pop_chars = read_tsv('tables/study_pop_chars.tsv')
 
-pop_chars = read_tsv('tables/new/study_pop_chars.tsv') %>%
+pop_chars = read_tsv('tables/study_pop_chars.tsv') %>%
   spread(key, value) %>%
@@
 ```{r raw_drug_trends_table}
-claims_by_abx = read_tsv('tables/new/claims_by_abx.tsv') %>%
+claims_by_abx = read_tsv('tables/claims_by_abx.tsv') %>%
   gather('abx', 'n', c(-year, -bene)) %>%
+  mutate(n=if_else(year == 2015 & abx %in% c('pde_approp', 'pde_inapprop', 'pde_nodx'),
+                   as.integer(round(n * 4/3)),
+                   n)) %>%
   mutate(cpkp=n/bene*1000)
@@
   arrange(desc(`2011`)) %>%
-  kable(caption='Raw drug CPKP')
+  kable(caption='Raw drug CPKP. 2015 values for abx-assoc. are projected.')
 ```
@@
 
-abx_trends = read_tsv('tables/new/abx_models.tsv') %>%
+abx_trends = read_tsv('tables/abx_models.tsv') %>%
   format_model()
@@
   xlab('') +
-  ylab('adjusted change in claims per beneficiary per year from 2011 to 2014 (%)') +
+  ylab('adjusted change in claims per beneficiary per year from 2011 to 2015 (%)') +
   scale_y_continuous(limits=c(-22, 33), breaks=c(-20, -10, 0, 10, 20, 30)) +
@@
 
-rxdx_table = read_tsv('tables/new/rxdx_counts.tsv') %>%
+rxdx_table = read_tsv('tables/rxdx_counts.tsv') %>%
   left_join(denom, by='year') %>%
-  mutate(cpkp=n_pdedx / bene * 1000)
+  mutate(cpkp=n_pdedx / bene * 1000) %>%
+  mutate(cpkp=if_else(year==2015,
+                      cpkp * 4/3,
+                      cpkp))
 
@@
 ```{r prescribing_practice}
-rxdx_models = read_tsv('tables/new/rxdx_models.tsv')
+rxdx_models = read_tsv('tables/rxdx_models.tsv')
 
 combined_table = rxdx_table %>%
-  mutate(cpkp=round(n_pdedx / bene * 1000, 1)) %>%
-  filter(year %in% c(2011, 2014)) %>%
+  mutate(cpkp=round(cpkp, 1)) %>%
+  filter(year %in% c(2011, 2015)) %>%
   select(antibiotic, dx_cat, year, cpkp) %>%
@@
   format_model() %>%
-  select(dx_cat, antibiotic, `2011`, `2014`, change, p.value) %>%
+  select(dx_cat, antibiotic, `2011`, `2015`, change, p.value) %>%
   arrange(dx_cat, antibiotic)
@@
 
+## Any respiratory
+
+```{r any_respiratory}
+read_tsv('tables/rxdx_resp_counts.tsv') %>%
+  left_join(denom, by='year') %>%
+  mutate(n_pdedx=if_else(year==2015, n_pdedx * 4/3, as.numeric(n_pdedx))) %>%
+  mutate(cpkp=round(n_pdedx / bene * 1000, 1)) %>%
+  filter(year %in% c(2011, 2015)) %>%
+  select(antibiotic, year, cpkp) %>%
+  spread(year, cpkp) %>%
+  mutate(change=`2015`-`2011`) %>%
+  kable(caption='Claims per 1,000 beneficiaries of each drug in association with any of the considered respiratory diagnoses')
+```
+
 # Supplement
@@
 ```{r claims_by_demography}
-subpop_table = read_tsv('tables/new/claims_by_subpop.tsv') %>%
+subpop_table = read_tsv('tables/claims_by_subpop.tsv') %>%
   mutate(cpkp=round(n_pde/n_bene*1000)) %>%
-  #filter(year %in% c(2011, 2014)) %>%
+  filter(year %in% c(2011, 2015)) %>%
   select(subpop, year, cpkp) %>%
@@
 
-subpop_models = read_tsv('tables/new/subpop_models.tsv') %>%
+subpop_models = read_tsv('tables/subpop_models.tsv') %>%
   mutate_at(vars(estimate, lower_ci, upper_ci), est2pct) %>%
   mutate(pm=pmax(estimate-lower_ci, upper_ci-estimate)) %>%
-  mutate(change=sprintf('%.1f (%.1f -- %.1f)', estimate, lower_ci, upper_ci)) %>%
-  mutate(change2=sprintf('%.1f ± %.1f', estimate, pm)) %>%
+  mutate(change=sprintf('%.1f ± %.1f', estimate, ceiling_at(pm, 1))) %>%
   mutate(p.value=format_p(p_value)) %>%
-  select(subpop, change2, p.value)
+  select(subpop, change, p.value)
 

### 2018-01-30 ms1/ms1_display.Rmd
study pop chars

@@
 import::from(knitr, kable)
+
+est2pct = function(x) (x - 1)*100
 ```
@@
 ```{r pop_chars}
-pop_chars = read_tsv('tables/study_pop_chars.tsv')
-
 pop_chars = read_tsv('tables/study_pop_chars.tsv') %>%
@@
   kable(caption='Population characteristics')
-
-trend_p = function(x) tidy(t.test(diff(x)))$p.value
-
-pop_chars %>%
-  arrange(key, year) %>%
-  group_by(key) %>%
-  summarize(p_value=trend_p(value)) %>%
-  kable(caption='Trends in population characteristics')
 ```
 
-There are `r scan('tables/n_unique_bene.txt') %>% prettyNum(big.mark=',')` unique beneficiaries.
+```{r popchar_models}
+pop_char_models = read_tsv('tables/pop_char_models.tsv')
 
-The top 10 drugs take up `r round(scan('tables/top10_cumf.txt') * 100, 2)`% of all claims.
+pop_char_models %>%
+  mutate_at(vars(estimate, lower_ci, upper_ci), function(x) prettyNum(est2pct(x), digits=3, width=0, format='fg')) %>%
+  kable(caption='Trends in study population characteristics')
+```
 
@@
   kable(caption='Raw drug CPKP. 2015 values for abx-assoc. are projected.')
+```
+
+The change in overall use is so small that I'll give it another decimal place:
+
+```{r}
+claims_by_abx %>%
+  filter(abx=='pde', year %in% c(2011, 2015)) %>%
+  select(year, cpkp) %>%
+  mutate(cpkp=round(cpkp, 1)) %>%
+  kable(caption='Overall use')
 ```
@@
   theme(panel.grid.minor.x=element_blank(),
-        strip.background = element_blank()) +
+        strip.background = element_blank(),
+        plot.margin=unit(c(1.0, 1.0, 0.5, 0.5), 'lines')) +
   scale_x_continuous(expand=c(0, 0), name='') +
@@
 ```{r abx_trends}
-est2pct = function(x) (x - 1)*100
 format_p = function(p) if_else(p<0.0001, '<.0001', format(round(p, 4), digits=4, nsmall=4))
@@
     mutate(pm=pmax(estimate-lower_ci, upper_ci-estimate)) %>%
-    mutate(change=sprintf('%.1f ± %.1f', estimate, ceiling_at(pm, 1))) %>%
+    #mutate(change=sprintf('%.1f ± %.1f', estimate, ceiling_at(pm, 1))) %>%
+    mutate(change=sprintf('%.1f (%.1f -- %.1f)', estimate, lower_ci, upper_ci)) %>%
     mutate(p.value=format_p(p_value))
@@
 
-abx_trends = read_tsv('tables/abx_models.tsv') %>%
+abx_trends_raw = read_tsv('tables/abx_models.tsv')
+
+abx_trends = abx_trends_raw %>%
   format_model()
@@
   ggplot(aes(x=antibiotic, y=estimate, ymin=lower_ci, ymax=upper_ci)) +
-  geom_bar(stat='identity', fill='white', color='black') +
+  geom_col(fill='white', color='black') +
   geom_errorbar(width=0.5) +
@@
   ylab('adjusted change in claims per beneficiary per year from 2011 to 2015 (%)') +
-  scale_y_continuous(limits=c(-22, 33), breaks=c(-20, -10, 0, 10, 20, 30)) +
+  scale_y_continuous(limits=c(-20, 30), breaks=c(-20, -10, 0, 10, 20, 30), expand=c(0.0, 0.00)) +
   theme_minimal() +
@@
 ggsave('fig/abx_trends.pdf', width=6.4, height=4.5, units='in')
+```
+
+Again, overall use declines so little that I'll give it one more decimal place:
+
+```{r}
+abx_trends_raw %>%
+  filter(antibiotic=='n_pde') %>%
+  select(-antibiotic) %>%
+  mutate_at(vars(estimate, lower_ci, upper_ci), est2pct) %>%
+  mutate(pm=pmax(estimate-lower_ci, upper_ci-estimate)) %>%
+  #mutate(change=sprintf('%.2f ± %.2f', estimate, ceiling_at(pm, 2))) %>%
+  mutate(change=sprintf('%.2f (%.2f -- %.2f)', estimate, lower_ci, upper_ci)) %>%
+  mutate(p.value=format_p(p_value)) %>%
+  kable
 ```
@@
   right_join(rxdx_models, by=c('antibiotic', 'dx_cat')) %>%
-  format_model() %>%
+  #format_model() %>%
+  mutate_at(vars(estimate, lower_ci, upper_ci), est2pct) %>%
+  mutate(pm=pmax(estimate-lower_ci, upper_ci-estimate)) %>%
+  mutate(change=sprintf('%.1f ± %.1f', estimate, ceiling_at(pm, 1))) %>%
+  mutate(p.value=format_p(p_value)) %>%
   select(dx_cat, antibiotic, `2011`, `2015`, change, p.value) %>%
@@
   mutate(pm=pmax(estimate-lower_ci, upper_ci-estimate)) %>%
-  mutate(change=sprintf('%.1f ± %.1f', estimate, ceiling_at(pm, 1))) %>%
+  #mutate(change=sprintf('%.1f ± %.1f', estimate, ceiling_at(pm, 1))) %>%
+  mutate(change=sprintf('%.2f (%.2f -- %.2f)', estimate, lower_ci, upper_ci)) %>%
   mutate(p.value=format_p(p_value)) %>%
@@
 ```{r ssti}
-rxdx_kable(c('cephalexin', 'tmpsmx'),
+rxdx_kable(c('cephalexin', 'tmpsmx', 'ciprofloxacin'),
            c('ssti', 'ssti_t3'))

### 2018-01-31 ms1/ms1_display.Rmd
a few more displays

@@
 
-The change in overall use is so small that I'll give it another decimal place:
+The changes in overall use and appropriate use are so small that I'll give it another decimal place:
 
 ```{r}
+# N.B.: 2015 claims were adjusted during loading of the data file
 claims_by_abx %>%
-  filter(abx=='pde', year %in% c(2011, 2015)) %>%
-  select(year, cpkp) %>%
+  filter(abx %in% c('pde', 'pde_approp'), year %in% c(2011, 2015)) %>%
+  select(year, abx, cpkp) %>%
   mutate(cpkp=round(cpkp, 1)) %>%
-  kable(caption='Overall use')
+  spread(year, cpkp) %>%
+  kable(caption='CPKP for small changes')
 ```

### 2018-06-01 ms1/ms1_display.Rmd
new ms1 analyses

@@
 pop_chars %>%
-  mutate(key=factor(key, levels=c('n_bene', 'mean_age', 'sd_age', 'n_cc', 'sd_cc', 'pct_female', 'pct_white', 'pct_dual', 'pct_south', 'pct_midwest', 'pct_west', 'pct_northeast'))) %>%
+  mutate(key=factor(key, levels=c(
+    'n_bene',
+    'pct_65-74', 'pct_75-84', 'pct_85-94', 'pct_95+',
+    'pct_female', 'pct_white', 'pct_dual',
+    'mean_cc', 'sd_cc',
+    'pct_south', 'pct_midwest', 'pct_west', 'pct_northeast'))) %>%
   mutate(value=case_when(
     .$key == 'n_bene' ~ prettyNum(.$value, big.mark=','),
-    .$key %in% c('n_cc', 'sd_cc') ~ format(.$value, digits=2, nsmall=2, scientific=FALSE),
+    .$key %in% c('mean_cc', 'sd_cc', 'pct_95+') ~ format(.$value, digits=2, nsmall=2, scientific=FALSE),
     TRUE ~ format(.$value, digits=2, nsmall=1, scientific=FALSE)
@@
 
+## Attrition in the sample
+
+```{r}
+attrition = read_tsv('tables/clean_bene_log.tsv')
+
+attrition %>%
+  mutate(step=fct_inorder(factor(step))) %>%
+  spread(year, n_bene) %>%
+  kable()
+
+denom = attrition %>%
+  filter(step=='final') %>%
+  select(-step)
+```
+
 # Trends in drug usage
@@
 claims_by_abx = read_tsv('tables/claims_by_abx.tsv') %>%
-  gather('abx', 'n', c(-year, -bene)) %>%
-  mutate(n=if_else(year == 2015 & abx %in% c('pde_approp', 'pde_inapprop', 'pde_nodx'),
-                   as.integer(round(n * 4/3)),
-                   n)) %>%
-  mutate(cpkp=n/bene*1000)
+  left_join(denom, by='year') %>%
+  mutate(cpkp=n_pde/n_bene*1000)
 
@@
   mutate(cpkp=round(cpkp)) %>%
-  select(year, abx, cpkp) %>%
+  select(year, antibiotic, cpkp) %>%
   spread(year, cpkp) %>%
@@
   kable(caption='Raw drug CPKP. 2015 values for abx-assoc. are projected.')
+```
+
+## Appropriateness
+
+```{r}
+approp_counts = read_tsv('tables/approp_counts.tsv') %>%
+  left_join(denom, by='year') %>%
+  mutate(antibiotic=case_when(
+    .$appropriateness == 'appropriate' ~ 'approp',
+    .$appropriateness == 'inappropriate' ~ 'inapp',
+    .$appropriateness == 'indeterminate' ~ 'indet'
+  )) %>%
+  select(-appropriateness) %>%
+  mutate(cpkp=n_pde/n_bene*1000)
+
+approp_counts %>%
+  select(year, antibiotic, cpkp) %>%
+  spread(year, cpkp)
 ```
@@
 claims_by_abx %>%
-  filter(abx %in% c('pde', 'pde_approp'), year %in% c(2011, 2015)) %>%
-  select(year, abx, cpkp) %>%
+  filter(antibiotic=='n_pde') %>%
   mutate(cpkp=round(cpkp, 1)) %>%
-  spread(year, cpkp) %>%
   kable(caption='CPKP for small changes')
@@
 claims_by_abx %>%
+  bind_rows(approp_counts) %>%
+  filter(antibiotic != 'n_pde_top10') %>%
+  rename(abx=antibiotic) %>%
   mutate(panel = factor(case_when(
-    .$abx %in% c('pde', 'pde_approp', 'pde_inapprop', 'pde_nodx') ~ 1,
-    .$abx %in% c('azithromycin', 'ciprofloxacin', 'amoxicillin', 'cephalexin', 'tmpsmx', 'levofloxacin') ~ 2,
-    .$abx %in% c('amoxclav', 'doxycycline', 'nitrofurantoin', 'clindamycin') ~ 3
+    .$abx %in% c('n_pde', 'approp', 'inapp', 'indet') ~ 1,
+    .$abx %in% c('azithromycin', 'ciprofloxacin', 'amoxicillin', 'cephalexin', 'trimethoprim/sulfamethoxazole', 'levofloxacin') ~ 2,
+    .$abx %in% c('amoxicillin/clavulanate', 'doxycycline', 'nitrofurantoin', 'clindamycin') ~ 3
   ))) %>%
@@
   mutate(abx=case_when(
-    .$abx == 'pde' ~ 'Overall use',
-    .$abx == 'pde_inapprop' ~ 'Inappropriate use',
-    .$abx == 'pde_approp' ~ 'Appropriate use',
-    .$abx == 'pde_nodx' ~ 'Indeterminate use',
-    .$abx == 'amoxclav' ~ 'amox/clav',
-    .$abx == 'tmpsmx' ~ 'TMP/SMX',
+    .$abx == 'n_pde' ~ 'Overall use',
+    .$abx == 'inapp' ~ 'Inappropriate use',
+    .$abx == 'approp' ~ 'Appropriate use',
+    .$abx == 'indet' ~ 'Indeterminate use',
+    .$abx == 'amoxicillin/clavulanate' ~ 'amox/clav',
+    .$abx == 'trimethoprim/sulfamethoxazole' ~ 'TMP/SMX',
     TRUE ~ .$abx)
@@
     mutate(pm=pmax(estimate-lower_ci, upper_ci-estimate)) %>%
-    #mutate(change=sprintf('%.1f ± %.1f', estimate, ceiling_at(pm, 1))) %>%
     mutate(change=sprintf('%.1f (%.1f -- %.1f)', estimate, lower_ci, upper_ci)) %>%
@@
 
-abx_trends_raw = read_tsv('tables/abx_models.tsv')
-
-abx_trends = abx_trends_raw %>%
+abx_trends = read_tsv('tables/abx_models.tsv') %>%
   format_model()
@@
 abx_trends %>%
-  filter(antibiotic != 'metronidazole') %>%
   mutate(antibiotic=case_when(
-    .$antibiotic == 'n_pde' ~ 'Overall use',
-    .$antibiotic == 'n_pde_inapprop' ~ 'Inappropriate use',
-    .$antibiotic == 'n_pde_approp' ~ 'Appropriate use',
-    .$antibiotic == 'n_pde_nodx' ~ 'Indeterminate use',
-    .$antibiotic == 'amoxclav' ~ 'amox/clav',
-    .$antibiotic == 'tmpsmx' ~ 'TMP/SMX',
+    .$antibiotic == 'n_abx' ~ 'Overall use',
+    .$antibiotic == 'amoxicillin/clavulanate' ~ 'amox/clav',
+    .$antibiotic == 'trimethoprim/sulfamethoxazole' ~ 'TMP/SMX',
     TRUE ~ .$antibiotic)
@@
 
+```{r}
+app_trends = read_tsv('tables/approp_models.tsv') %>%
+  mutate(antibiotic=case_when(
+    .$appropriateness == 'inappropriate' ~ 'Inappropriate use',
+    .$appropriateness == 'appropriate' ~ 'Appropriate use',
+    .$appropriateness == 'indeterminate' ~ 'Indeterminate use')
+  ) %>%
+  select(-appropriateness) %>%
+  format_model()
+
+app_trends %>%
+  select(antibiotic, change, p.value) %>%
+  kable()
+```
+
 Again, overall use declines so little that I'll give it one more decimal place:
@@
 ```{r}
-abx_trends_raw %>%
+abx_trends %>%
   filter(antibiotic=='n_pde') %>%
@@
   mutate(pm=pmax(estimate-lower_ci, upper_ci-estimate)) %>%
-  #mutate(change=sprintf('%.2f ± %.2f', estimate, ceiling_at(pm, 2))) %>%
   mutate(change=sprintf('%.2f (%.2f -- %.2f)', estimate, lower_ci, upper_ci)) %>%
@@
 ```{r inappropriate_fraction}
-claims_by_abx %>%
-  filter(str_detect(abx, '^pde')) %>%
-  select(year, abx, n) %>%
-  spread(abx, n) %>%
-  mutate(denom=pde_approp + pde_inapprop + pde_nodx) %>%
-  mutate_at(vars(matches('^pde_')), function(x) round(x / .$denom * 100, 1)) %>%
-  select(-pde, -denom) %>%
-  kable(caption='Proportion of app., inapp., and no-dx prescribing')
+read_tsv('tables/approp_counts.tsv') %>%
+  group_by(year) %>%
+  mutate(pct = round(100*(n_pde / sum(n_pde)), 1)) %>%
+  select(-n_pde) %>%
+  spread(year, pct)
 ```
@@
 ```{r top_dx}
-denom = claims_by_abx %>%
-  select(year, bene) %>%
-  distinct()
+dx_table = read_tsv('tables/dx_counts.tsv') %>%
+  filter(!is.na(dx_cat), dx_cat != 'remaining_codes')
 
-rxdx_table = read_tsv('tables/rxdx_counts.tsv') %>%
+dx_table %>%
   left_join(denom, by='year') %>%
-  mutate(cpkp=n_pdedx / bene * 1000) %>%
-  mutate(cpkp=if_else(year==2015,
-                      cpkp * 4/3,
-                      cpkp))
-
-top_dx = rxdx_table %>%
-  filter(year==2011) %>%
-  group_by(dx_cat) %>%
-  summarize(max_cpkp=max(cpkp)) %>%
-  mutate(over_10=max_cpkp >= 10) %>%
-  arrange(desc(max_cpkp))
-
-top_dx %>%
-  kable(caption='Diagnoses by maximum drug usage in 2011')
-```
-
-```{r top_rxdx}
-infection_sites = read_tsv('infection_sites.tsv')
-
-top_dx %>%
-  filter(over_10) %>%
-  slice(-1) %>% # drop "remaining codes"
-  select(dx_cat) %>%
-  inner_join(rxdx_table, by='dx_cat') %>%
-  left_join(infection_sites, by='dx_cat') %>%
-  filter(year==2011) %>%
-  group_by(site, antibiotic) %>%
-  summarize(cpkp=sum(cpkp)) %>%
-  arrange(site, desc(cpkp)) %>%
-  group_by(site) %>%
-  mutate(f=cpkp/sum(cpkp), cumf=cumsum(f)/sum(f)) %>%
-  do(head(., 5)) %>%
-  kable(caption='Top drugs by diagnosis in 2011')
+  mutate(dxkpy = round(n_dx / n_bene * 1000)) %>%
+  select(-n_dx, -n_bene) %>%
+  spread(year, dxkpy) %>%
+  arrange(desc(`2011`)) %>%
+  kable
 ```
@@
 ```{r prescribing_practice}
-rxdx_models = read_tsv('tables/rxdx_models.tsv')
+rxdx_table = read_tsv('tables/pde_for_dx_counts.tsv') %>%
+  left_join(denom, by='year') %>%
+  mutate(cpkp = n_pdedx / n_bene * 1000)
+
+rxdx_models = read_tsv('tables/pde_for_dx_models.tsv')
 
@@
   mutate(cpkp=round(cpkp, 1)) %>%
-  filter(year %in% c(2011, 2015)) %>%
+  filter(year %in% c(2011, 2014)) %>%
   select(antibiotic, dx_cat, year, cpkp) %>%
@@
   right_join(rxdx_models, by=c('antibiotic', 'dx_cat')) %>%
-  #format_model() %>%
   mutate_at(vars(estimate, lower_ci, upper_ci), est2pct) %>%
@@
   mutate(p.value=format_p(p_value)) %>%
-  select(dx_cat, antibiotic, `2011`, `2015`, change, p.value) %>%
+  select(dx_cat, antibiotic, `2011`, `2014`, change, p.value) %>%
   arrange(dx_cat, antibiotic)
@@
 
-rxdx_kable(c('azithromycin', 'levofloxacin', 'amoxclav'),
+rxdx_kable(c('azithromycin', 'levofloxacin', 'amoxicillin/clavulanate'),
            c('pneumonia', 'sinusitis', 'uri', 'bronchitis', 'other_resp', 'asthma'))
-```
-
-## Any respiratory
-
-```{r any_respiratory}
-read_tsv('tables/rxdx_resp_counts.tsv') %>%
-  left_join(denom, by='year') %>%
-  mutate(n_pdedx=if_else(year==2015, n_pdedx * 4/3, as.numeric(n_pdedx))) %>%
-  mutate(cpkp=round(n_pdedx / bene * 1000, 1)) %>%
-  filter(year %in% c(2011, 2015)) %>%
-  select(antibiotic, year, cpkp) %>%
-  spread(year, cpkp) %>%
-  mutate(change=`2015`-`2011`) %>%
-  kable(caption='Claims per 1,000 beneficiaries of each drug in association with any of the considered respiratory diagnoses')
 ```
@@
   filter(year %in% c(2011, 2015)) %>%
-  select(subpop, year, cpkp) %>%
+  select(pop_char, value, year, cpkp) %>%
   spread(year, cpkp)
@@
   mutate(pm=pmax(estimate-lower_ci, upper_ci-estimate)) %>%
-  #mutate(change=sprintf('%.1f ± %.1f', estimate, ceiling_at(pm, 1))) %>%
   mutate(change=sprintf('%.2f (%.2f -- %.2f)', estimate, lower_ci, upper_ci)) %>%
   mutate(p.value=format_p(p_value)) %>%
-  select(subpop, change, p.value)
+  select(pop_char, value, change, p.value)
 
-left_join(subpop_table, subpop_models, by='subpop') %>%
+left_join(subpop_table, subpop_models, by=c('pop_char', 'value')) %>%
   kable(caption='Claims and trends by subpopulation')
@@
 ```{r uti}
-rxdx_kable(c('ciprofloxacin', 'tmpsmx', 'nitrofurantoin'),
+rxdx_kable(c('ciprofloxacin', 'trimethropim/sulfamethoxazole', 'nitrofurantoin'),
            c('uti', 'uti_t3'))
@@
 ```{r ssti}
-rxdx_kable(c('cephalexin', 'tmpsmx', 'ciprofloxacin'),
+rxdx_kable(c('cephalexin', 'trimethoprim/sulfamethoxazole', 'ciprofloxacin'),
            c('ssti', 'ssti_t3'))
 ```
+
+## Prescribing practice -- single
+
+```{r prescribing_practice_similarity}
+rxdx_table_single = read_tsv('tables/pde_for_dx_single_counts.tsv') %>%
+  left_join(denom, by='year') %>%
+  mutate(cpkp = n_pdedx / n_bene * 1000)
+
+rxdx_models_single = read_tsv('tables/pde_for_dx_single_models.tsv')
+
+rxdx_comp = inner_join(rxdx_models, rxdx_models_single,
+           by=c('antibiotic', 'dx_cat'),
+           suffix=c('.multi', '.single'))
+
+rxdx_comp %$%
+  cor.test(estimate.multi, estimate.single)
+
+lm(estimate.multi  ~ estimate.single, data=rxdx_comp) %>%
+  summary()
+
+rxdx_comp %>%
+  ggplot(aes(estimate.multi, estimate.single,
+             xmin=lower_ci.multi, xmax=upper_ci.multi,
+             ymin=lower_ci.single, ymax=upper_ci.single)) +
+  geom_abline() +
+  geom_point() +
+  geom_errorbar() + geom_errorbarh()
+```
+
+```{r prescribing_practice_single}
+infection_sites = read_tsv('infection_sites.tsv')
+
+combined_table_single = rxdx_table_single %>%
+  mutate(cpkp=round(cpkp, 1)) %>%
+  filter(year %in% c(2011, 2014)) %>%
+  select(antibiotic, dx_cat, year, cpkp) %>%
+  spread(year, cpkp) %>%
+  right_join(rxdx_models_single, by=c('antibiotic', 'dx_cat')) %>%
+  mutate_at(vars(estimate, lower_ci, upper_ci), est2pct) %>%
+  mutate(pm=pmax(estimate-lower_ci, upper_ci-estimate)) %>%
+  mutate(change=sprintf('%.1f ± %.1f', estimate, ceiling_at(pm, 1))) %>%
+  mutate(p.value=format_p(p_value)) %>%
+  left_join(infection_sites, by='dx_cat') %>%
+  select(site, dx_cat, antibiotic, `2011`, `2014`, change, p.value) %>%
+  arrange(site, dx_cat, antibiotic)
+
+combined_table_single %>%
+  kable()
+```

### 2018-06-04 ms1/ms1_display.Rmd
new ms1 analys

@@
 
+## Top 10 fraction
+
+```{r}
+pct_top10 = claims_by_abx %>%
+  filter(antibiotic %in% c('n_pde', 'n_pde_top10')) %>%
+  select(year, antibiotic, n_pde) %>%
+  spread(antibiotic, n_pde) %$%
+  { 100 * sum(n_pde_top10) / sum(n_pde) }
+```
+
+The top 10 antibiotics accounted for `r round(pct_top10, 2)`% of claims.
+
 ## Appropriateness
@@
   select(year, antibiotic, cpkp) %>%
-  spread(year, cpkp)
+  spread(year, cpkp) %>%
+  kable()
 ```
@@
 ```{r raw_drug_trends}
-abx_levels = levels(claims_by_abx$abx)
-
 dummy = data_frame(
@@
 
-claims_by_abx %>%
+cba_dat = claims_by_abx %>%
   bind_rows(approp_counts) %>%
@@
   ) %>%
-  arrange(year, desc(cpkp)) %>%
+  arrange(year, desc(cpkp))
+
+cba_dat %>%
   {
@@
   geom_line() +
-  geom_text_repel(
+  geom_text(
     data=filter(., year==2013),
-    aes(label=abx)
+    aes(label=abx),
+    nudge_y=c(50, 50, -35, 35,
+              26, -8, 10, 8, -8, -6,
+              4, 8, -3, -5),
+    nudge_x=c(0, rep(-0.5, 3),
+              0, 0, 0, 0, 1.3, 0,
+              rep(0, 4))
   ) +
@@
         strip.background = element_blank(),
-        plot.margin=unit(c(1.0, 1.0, 0.5, 0.5), 'lines')) +
+        plot.margin=unit(c(1.0, 1.0, 0.5, 0.5), 'lines'),
+        axis.text.x=element_text(color='black'),
+        axis.text.y=element_text(color='black'),
+        text=element_text(color='black')) +
   scale_x_continuous(expand=c(0, 0), name='') +
@@
 ggsave('fig/raw_drug.pdf', height=5.0, width=9.0, unit='in')
+ggsave('fig/drug_trends.png', height=5.0, width=9.0, unit='in')
 ```
@@
   mutate(antibiotic=fct_inorder(factor(antibiotic))) %>%
-  mutate(antibiotic=fct_relevel(antibiotic, c('Overall use', 'Inappropriate use', 'Appropriate use', 'Indeterminate use'))) %>%
+  mutate(antibiotic=fct_relevel(antibiotic, 'Overall use')) %>%
   mutate(antibiotic=fct_rev(antibiotic)) %>%
@@
   theme_minimal() +
-  theme(axis.title.x = element_text(size=12),
-        axis.text = element_text(size=12))
+  theme(axis.title.x=element_text(size=12, color='black', hjust=1),
+        axis.text = element_text(size=12, color='black'),
+        plot.margin = unit(c(0.1, 0.2, 0.1, 0), 'in'))
 
 ggsave('fig/abx_trends.pdf', width=6.4, height=4.5, units='in')
+ggsave('fig/abx_trends.png', width=6.4, height=4.5, units='in')
 ```
@@
     .$appropriateness == 'appropriate' ~ 'Appropriate use',
-    .$appropriateness == 'indeterminate' ~ 'Indeterminate use')
+    .$appropriateness == 'indeterminate' ~ 'Indeterminate use',
+    .$appropriateness == 'overall3' ~ 'Overall use'
+    )
   ) %>%
@@
 }
+
+rxdx_models %>%
+  format_model() %>%
+  select(antibiotic, dx_cat, change, p.value) %>%
+  filter(antibiotic %in% c('azithromycin', 'levofloxacin'),
+         dx_cat %in% c('pneumonia', 'uri'))
 
@@
 ```
+
+### Use for any
+
+Use of a drug for any respiratory bit.
+
+```{r}
+rxdx_table_single %>%
+  filter(site=='respiratory') %>%
+  group_by(year, antibiotic) %>%
+  summarize(cpkp = round(sum(cpkp), 1)) %>%
+  spread(year, cpkp) %>%
+  kable()
+```

### 2018-06-12 ms1/ms1_display.Rmd
small tweaks

@@
   panel = factor(c(1, 1, 2, 2, 3, 3)),
-  abx = 'pde'
+  abx = 'n_pde'
 )
@@
     .$abx == 'n_pde' ~ 'Overall use',
-    .$abx == 'inapp' ~ 'Inappropriate use',
-    .$abx == 'approp' ~ 'Appropriate use',
+    .$abx == 'inapp' ~ 'Potentially inappropriate use',
+    .$abx == 'approp' ~ 'Potentially appropriate use',
     .$abx == 'indet' ~ 'Indeterminate use',
@@
     aes(label=abx),
-    nudge_y=c(50, 50, -35, 35,
+    nudge_y=c(50, 55, -40, 40,
               26, -8, 10, 8, -8, -6,
               4, 8, -3, -5),
-    nudge_x=c(0, rep(-0.5, 3),
+    nudge_x=c(0, -0.3, -0.3, -0.5,
               0, 0, 0, 0, 1.3, 0,
@@
 
-ggsave('fig/raw_drug.pdf', height=5.0, width=9.0, unit='in')
-ggsave('fig/drug_trends.png', height=5.0, width=9.0, unit='in')
+ggsave('fig/raw_trends.pdf', height=5.0, width=9.0, unit='in')
+ggsave('fig/raw_trends.png', height=5.0, width=9.0, unit='in')
 ```
@@
 
-ggsave('fig/abx_trends.pdf', width=6.4, height=4.5, units='in')
-ggsave('fig/abx_trends.png', width=6.4, height=4.5, units='in')
+ggsave('fig/adjusted_trends.pdf', width=6.4, height=4.5, units='in')
+ggsave('fig/adjusted_trends.png', width=6.4, height=4.5, units='in')
 ```
@@
 abx_trends %>%
-  filter(antibiotic=='n_pde') %>%
+  filter(antibiotic=='n_abx') %>%
   select(-antibiotic) %>%
@@
   mutate(p.value=format_p(p_value)) %>%
-  kable
+  kable(caption='Trends in overall use')
 ```
@@
 
+## Retention
+
+```{r}
+retention_f = function(x) {
+  n = length(x)
+  c(1.0, x[2:n] / x[1:(n-1)])
+}
+
+retention_dat = read_tsv('tables/clean_bene_log.tsv')
+
+retention_dat %>%
+  filter(!(step %in% c('total', 'final'))) %>%
+  group_by(year) %>%
+  mutate(retention=retention_f(n_bene)) %>%
+  ungroup() %>%
+  mutate(step=factor(step, levels=c('20pct', 'over65', 'abd', 'demo')),
+         n_bene=prettyNum(n_bene, big.mark=','),
+         value=sprintf('%s (%.1f)', n_bene, 100*retention)) %>%
+  select(year, step, value) %>%
+  spread(year, value) %>%
+  kable()
+```
+
 ## Claims by demography
@@
 subpop_table = read_tsv('tables/claims_by_subpop.tsv') %>%
-  mutate(cpkp=round(n_pde/n_bene*1000)) %>%
+  mutate(cpkp=round(n_pde/n_bene*1000, 2)) %>%
   filter(year %in% c(2011, 2015)) %>%
@@
 ```{r uti}
-rxdx_kable(c('ciprofloxacin', 'trimethropim/sulfamethoxazole', 'nitrofurantoin'),
+rxdx_kable(c('ciprofloxacin', 'trimethoprim/sulfamethoxazole', 'nitrofurantoin'),
            c('uti', 'uti_t3'))
@@
 ```{r prescribing_practice_similarity}
+infection_sites = read_tsv('db/infection_sites.tsv')
+
+respiratory_dxs = infection_sites %>%
+  filter(site=='resp') %>%
+  pull(dx_cat)
+
 rxdx_table_single = read_tsv('tables/pde_for_dx_single_counts.tsv') %>%
+  filter(dx_cat %in% respiratory_dxs) %>%
   left_join(denom, by='year') %>%
@@
 
-rxdx_models_single = read_tsv('tables/pde_for_dx_single_models.tsv')
+rxdx_models_single = read_tsv('tables/pde_for_dx_single_models.tsv') %>%
+  filter(dx_cat %in% respiratory_dxs)
+```
 
-rxdx_comp = inner_join(rxdx_models, rxdx_models_single,
+The two types of models give really similar results about trend estimates:
+
+```{r}
+inner_join(rxdx_models, rxdx_models_single,
            by=c('antibiotic', 'dx_cat'),
-           suffix=c('.multi', '.single'))
+           suffix=c('.multi', '.single')) %>%
+  mutate(f_diff=pmax(estimate.multi, estimate.single) / pmin(estimate.multi, estimate.single)) %$%
+  max(f_diff)
+```
 
-rxdx_comp %$%
-  cor.test(estimate.multi, estimate.single)
+But the number of claims should be different (line shows slope of 2):
 
-lm(estimate.multi  ~ estimate.single, data=rxdx_comp) %>%
-  summary()
-
-rxdx_comp %>%
-  ggplot(aes(estimate.multi, estimate.single,
-             xmin=lower_ci.multi, xmax=upper_ci.multi,
-             ymin=lower_ci.single, ymax=upper_ci.single)) +
-  geom_abline() +
-  geom_point() +
-  geom_errorbar() + geom_errorbarh()
+```{r}
+inner_join(rxdx_table, rxdx_table_single,
+           by=c('antibiotic', 'dx_cat'),
+           suffix=c('.multi', '.single')) %>%
+  ggplot(aes(cpkp.single, cpkp.multi)) +
+  geom_abline(slope=2) +
+  geom_point()
 ```
@@
 ```{r prescribing_practice_single}
-infection_sites = read_tsv('infection_sites.tsv')
-
 combined_table_single = rxdx_table_single %>%
@@
   mutate(p.value=format_p(p_value)) %>%
-  left_join(infection_sites, by='dx_cat') %>%
-  select(site, dx_cat, antibiotic, `2011`, `2014`, change, p.value) %>%
-  arrange(site, dx_cat, antibiotic)
+  select(dx_cat, antibiotic, `2011`, `2014`, change, p.value) %>%
+  mutate(antibiotic=factor(antibiotic, levels=c('azithromycin', 'levofloxacin', 'amoxicillin/clavulanate')),
+         dx_cat=factor(dx_cat, levels=c('pneumonia', 'sinusitis', 'uri', 'bronchitis', 'other_resp', 'asthma'))) %>%
+  arrange(dx_cat, antibiotic)
 
@@
 ```
+
+# Reviewer response
+
+## Print abstract figure
+
+```{r print_abstract}
+abstract_dummy = data_frame(
+  year = rep(c(2011, 2015), 2),
+  cpkp = c(0, 250, 0, 100),
+  panel = factor(c(2, 2, 3, 3)),
+  abx = 'pde'
+)
+
+abstract_panel_labels = c('2'='Top 6 most-used antibiotics',
+                 '3'='Next 4 most-used antibiotics')
+
+cba_dat %>%
+  filter(panel %in% c(2, 3)) %>%
+  {
+  ggplot(., aes(x=year, y=cpkp, group=abx)) +
+  geom_blank(data=abstract_dummy) +
+  geom_line() +
+  geom_text(
+    data=filter(., year==2013),
+    aes(label=abx),
+    nudge_y=c(26, -8, 10, 8, -8, -6,
+              4, 8, -3, -5),
+    nudge_x=c(0, 0, 0, 0, 1.3, 0,
+              rep(0, 4))
+  ) +
+  facet_wrap(~panel, scales='free_y', labeller=labeller(panel=abstract_panel_labels)) +
+  theme_classic() +
+  theme(panel.grid.minor.x=element_blank(),
+        strip.background = element_blank(),
+        plot.margin=unit(c(1.0, 1.0, 0.5, 0.5), 'lines'),
+        axis.text.x=element_text(color='black'),
+        axis.text.y=element_text(color='black'),
+        text=element_text(color='black')) +
+  scale_x_continuous(expand=c(0, 0), name='') +
+  scale_y_continuous(expand=c(0, 0), name='claims per 1,000 beneficiaries per year')
+  }
+
+ggsave('fig/abstract_fig.pdf', height=5.0, width=6.0, unit='in')
+ggsave('fig/abstract_fig.png', height=5.0, width=6.0, unit='in')
+```
+
+## Stacked bar charts
+
+```{r stacked_trends}
+p = cba_dat %>%
+  filter(panel != 1) %>%
+  mutate(abx=fct_reorder(factor(abx), cpkp)) %>%
+  ggplot(aes(x=year, y=cpkp, fill=abx)) +
+  geom_col(color='black') +
+  xlab('') +
+  ylab('Claims per 1,000 beneficiaries per year') +
+  theme_minimal() +
+  theme(legend.title=element_blank())
+
+show(p)
+ggsave(plot=p, 'fig/stacked_trends.pdf', units='in', width=4, height=3)
+```

### 2019-07-25 ms1/ms1_display.Rmd
small tweaks

@@
 
+rxdx_models_display = rxdx_models %>%
+  format_model %>%
+  select(antibiotic, dx_cat, change, p.value)
+
 combined_table = rxdx_table %>%
@@
   spread(year, cpkp) %>%
-  right_join(rxdx_models, by=c('antibiotic', 'dx_cat')) %>%
-  mutate_at(vars(estimate, lower_ci, upper_ci), est2pct) %>%
-  mutate(pm=pmax(estimate-lower_ci, upper_ci-estimate)) %>%
-  mutate(change=sprintf('%.1f ± %.1f', estimate, ceiling_at(pm, 1))) %>%
-  mutate(p.value=format_p(p_value)) %>%
+  right_join(rxdx_models_display, by=c('antibiotic', 'dx_cat')) %>%
   select(dx_cat, antibiotic, `2011`, `2014`, change, p.value) %>%
@@
 }
-
-rxdx_models %>%
-  format_model() %>%
-  select(antibiotic, dx_cat, change, p.value) %>%
-  filter(antibiotic %in% c('azithromycin', 'levofloxacin'),
-         dx_cat %in% c('pneumonia', 'uri'))
 
